### 2015-07-21 linvis.tex
Rough paper layout
This is a rough layout of how the paper will be set up.

1: Introduction.
Introduce the problem, etc...

2: Related work
Talk about work done by DMG to collect the information on the kernel.

3: Navigation
How our solution provides navigation mechanisms through the commit data
3.1: Search
Search mechanisms we provide.
3.2: Breadcrumbs
Breadcrumb mechanism allows users to move toward the root

4: Presentation

View files, modules, commit log, etc...

5: Date Range Queries (May be moved into navigation)

6: Tree Visualization (May be moved into navigation)
The tree provides a

% vim: set fenc=utf-8 ft=latex encoding=utf-8
% -*- mode: latex; coding: UTF-8; -*-

\newif\ifdraft
\drafttrue

\ifdraft
	\documentclass[conference, draftclsnofoot]{IEEEtran}
	\def\baselinestretch{1}
	\setlength{\marginparwidth}{2cm}
\else
	\documentclass[conferece, final]{IEEEtran}
\fi

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\newcommand{\TheTitle}{Visualizing Release Information of Linux}
\newcommand{\TheAuthors}{Evan Wilde}
\newcommand{\TheEmails}{etcwilde@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

\synctex=1

\usepackage[hyphens]{url}
\urlstyle{same}

\ifdraft
	\usepackage[unicode=true,bookmarks=false,breaklinks=false,
		pdfborder={0 0 0},backref=none,colorlinks=true]{hyperref}

\else
	\usepackage[unicode=true,bookmarks=false,breaklinks=false,
		pdfborder={0 0 0},backref=none,colorlinks=false]{hyperref}
\fi

\usepackage[nospace]{cite}

% Table Support
\usepackage{dcolumn}
\usepackage{longtable}

\usepackage{balance}
\usepackage{placeins}
\usepackage{multirow}

% Extra support
\usepackage{xspace}
\usepackage{caption}

% Fix any bad-hyphenations here
\hyphenation{}

\input{macros.tex}

\begin{document}

\title{\TheTitle}
\author{
\IEEEauthorblockA{\TheAuthors}
\IEEEauthorblockN{Department of Computer Science,
                    University of Victoria,
	    Canada.}
\IEEEauthorblockA{Email: \TheEmails}
}
\maketitle
\begin{abstract}

With an average of $x$ \evantodo{Fill in this number} merges into the Linux
kernel per release, containing thousands of commits, maintenance of older
versions of the kernel becomes nearly impossible. For security, performance,
and changing hardware, maintainers must understand the changes made to current
versions of the kernel, and how these changes fit into older versions in order
to merge them into the older versions. Current tools provide information about
the directed acyclic graph (DAG)\evan{Should I just replace this with DAG?}
which can be helpful for smaller projects, but with the scale and the number of
branches in the kernel, becomes overwhelming very quickly, if the tool doesn't
crash before generating the graph and handling the vast quantities information.

In this paper, I present a tool that uses the dataset collected by Daniel
German for \evantodo{Paper name\ldots} to generate a tool that can provide
better visualizations and better filtering.
\end{abstract}
\begin{IEEEkeywords}
\TheKeywords
\end{IEEEkeywords}

\section{Introduction}

Maintainers of older versions of the Linux kernel must sift through massive
number of commits with tools that are unable to filter and visualize the kernel
into a manageable size. Tools like gitk provide users with a full
view of the DAG (\ref{fig:DAG}), which for smaller projects can give a good
view of how the project is evolving. With larger, modular projects like the
Linux Kernel, the DAG becomes a tangled mess of lines that lose their meaning.
GitHub will also provide a DAG for most projects, but is unable to display it
for projects as large as the Linux Kernel (\ref{fig:github dag}).

\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/gitk.png}
	\caption{Portion of the Linux Kernel DAG}
	\label{fig:DAG}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/github_viewer.png}
	\caption{Github Error message}
	\label{fig:github dag}
\end{figure}

\section{Related Work}
Daniel German, et. al \evantodo{Find paper} recorded the Linux git repository
metadata from 2012 until 2015\evan{?}, storing information not available. They
also developed a heuristic to convert the DAG to a set of trees strung together
by the root as a linked list. We will call this structure a vine.

\evantodo{Make an picture of a vine with regards to the Linux kernel}

We built a web-based tool\footnote{\url{li.turingmachine.org}} to experiment
with the various ways to display and organize the release information of the
kernel.

%% The parts that make the tool usable
\section{Navigation}

Our system has multiple methods of allowing users to navigate through the data.

\subsection{Search}

Users search to filter out results they are not interested. We break the data
down by releases. From there, a user can further trim how many trees they want
to see by including a begin or end commit date, an author name, the commit id,
or keyword from a log preview to search for commits that match what they are
looking for.

\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/search_view.png}
	\caption{Search view}
	\label{fig:search_view}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/results_view.png}
	\caption{Results when searching for commits by Andrew Morton in Linux
	3.10}
	\label{fig:results_view}
\end{figure}

\subsection{Breadcrumbs}

\begin{figure}[h]
	\centering
	\includegraphics[height=1.0in]{figures/breadcrumbs.png}
	\caption{Generated Breadcrumbs}
	\label{fig:breadcrumbs}
\end{figure}

Our system provides breadcrumbs, allowing a user to move toward the root of the
tree and view the detailed view of the parent merge.

\section{Presentation}

\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/page.png}
	\caption{View of page}
	\label{fig:full page}
\end{figure}

We can provide information about what files were edited by a given commit or
merge. Due to the tree-structure of the data, we can perform a tree traversal
to determine the number of files and modules touched within the merges.

%% The special parts that make the tool special
\section{Date Range Queries}

Git does not store the date that a given commit is merged into another branch,
it stores the date that the commit was authored. This is why the data collected
by Daniel German becomes important. His recording system monitored each commit
as it moved through the kernel, storing the commit and author dates of each
commit.

A commit may be created and not be merged with the kernel until much later. It
is these merges that make performing date range queries using the conventional
tools useless as they will not be included even though they are having an
effect on the kernel in the requested date range. Our system splits the kernel
into releases, and only looks at the dates of the merges performed by Linus
directly into the master branch. From there, we are able to construct the trees
below each of these merges, which may contain commits from before the begin
date of the selected range, or after the end date for the given range. This
provides information about the commits that come from older branches that were
pulled forward, or patches that were merged backward into an older version of
the kernel as part of maintenance.

\section{Visualization}

Recognizing each merge with the master branch of the kernel as a subtree helps
remove information that is unnecessarily shown in other tools, like the DAG.
There are many ways of visualizing trees, including a standard tree
visualization(\ref{fig:moderate_tree_view}). An issue with a standard tree visualization
is that it can only display a certain amount of information at each node before
it becomes cluttered and incomprehensible, furthermore, as we can see in figure 
(\ref{fig:large_tree_view}), if there are many nodes, even limited information
becomes cluttered.

\evantodo{Make the lines and nodes bigger and brighter in the tree screenshots}
\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/tree_viewer.png}
	\caption{Moderately sized tree view tree}
	\label{fig:moderate_tree_view}
	% CID: 042dd60ca6dec9a02cefa8edd67de386e35755d6
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[height=2.5in]{figures/tree_viewer_large.png}
	\caption{Large tree view tree}
	\label{fig:large_tree_view}
	% CID : 5f56886521d6ddd3648777fae44d82382dd8c87f
\end{figure}

Perhaps a more useful tree is the list-tree, which allows us to show the
layers in the tree using indentation. This also allows us to create click-able
links to the corresponding items in the tree.

\begin{figure}[h]
	\centering
	\includegraphics[height=1.5in]{figures/list_viewer.png}
	\caption{Moderately sized list view tree}
	\label{fig:moderate_list_view}
	% CID: 042dd60ca6dec9a02cefa8edd67de386e35755d6
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[height=3.5in]{figures/list_viewer_large.png}
	\caption{Large list view tree}
	\label{fig:large_list_view}
	% CID : 5f56886521d6ddd3648777fae44d82382dd8c87f
\end{figure}

Figures \ref{fig:moderate_tree_view} and \ref{fig:moderate_list_view} display
the same subtree, however, the list-based subtree provides a cleaner, faster
way to navigate the tree, whereas the graphical tree view provides a way to
hide unnecessary branches of the tree.

We can see in figures \ref{fig:large_tree_view} and \ref{fig:large_list_view}
that neither representation of the trees can cope with the larger trees in a
clean manner. These larger trees are, however, and exception to the rule. These
trees are from one of the many memory management patches by Andrew Morton,
which get the name ``Andrews patch bombs''. Andrews patches are very wide trees
that have only leaves and no internal merging. Conversely, the networking
modules are far larger with massive breadth and some depth. Overall the Linux
kernel has a very flat structure wit little depth, the deepest node being $5$
merges away \evantodo{verify this} from the root merge into the kernel.

\evantodo{Build sunburst tree into visualizer, then add pictures}

\end{document}

### 2015-08-09 linvis.tex
Re-working the paper
Focusing more on the tree-centered structure, then talking about how it helps
the navigation and presentation of the information.

Not really sure what more I actually need to say in the paper though...

@@
 \IEEEauthorblockN{Department of Computer Science,
-                    University of Victoria,
-	    Canada.}
+                    University of Victoria, Canada.}
 \IEEEauthorblockA{Email: \TheEmails}
@@
 
-With an average of $x$ \evantodo{Fill in this number} merges into the Linux
-kernel per release, containing thousands of commits, maintenance of older
-versions of the kernel becomes nearly impossible. For security, performance,
-and changing hardware, maintainers must understand the changes made to current
-versions of the kernel, and how these changes fit into older versions in order
-to merge them into the older versions. Current tools provide information about
-the directed acyclic graph (DAG)\evan{Should I just replace this with DAG?}
-which can be helpful for smaller projects, but with the scale and the number of
-branches in the kernel, becomes overwhelming very quickly, if the tool doesn't
-crash before generating the graph and handling the vast quantities information.
+	With an average of over 900 top-level merges into the Linux kernel per
+	release, each containing thousands of commits, maintenance of older
+	versions of the kernel becomes nearly impossible. For security,
+	performance, and changing hardware, maintainers must understand the
+	changes made to current versions of the kernel, and show these changes
+	fit into older versions in order to make the necessary merges and
+	modifications to the older versions. Current tools provide information
+	about repositories through the directed acyclic graph (DAG) which is
+	helpful for smaller projects, but with the scale and number of branches
+	in the kernel, the DAG becomes overwhelming very quickly.
 
-In this paper, I present a tool that uses the dataset collected by Daniel
-German for \evantodo{Paper name\ldots} to generate a tool that can provide
-better visualizations and better filtering.
+	In this paper, we present a tool that uses the dataset collected by
+	German et al. to build a tool that uses a vine model instead of the
+	DAG. This tool is designed to cater to the needs of users looking for a
+	top-down approach, or a bottom-up approach to navigating the commit
+	information of the kernel.
 \end{abstract}
+
 \begin{IEEEkeywords}
@@
 
-Maintainers of older versions of the Linux kernel must sift through massive
-number of commits with tools that are unable to filter and visualize the kernel
-into a manageable size. Tools like gitk provide users with a full
-view of the DAG (\ref{fig:DAG}), which for smaller projects can give a good
-view of how the project is evolving. With larger, modular projects like the
-Linux Kernel, the DAG becomes a tangled mess of lines that lose their meaning.
-GitHub will also provide a DAG for most projects, but is unable to display it
-for projects as large as the Linux Kernel (\ref{fig:github dag}).
+Maintainers of older version of the Linux kernel must sift through thousands of
+commits with tools that are unable to filter and effectively visualize projects
+that are of the scale of the kernel. Tools like gitk provide users with the
+full view of the DAG, showing all merges and commits. For smaller projects,
+this information is useful to show the relation between various branches. With
+larger, modular projects like the Linux Kernel, the DAG becomes a tangled mess
+of lines that lose meaning. Github provides a DAG view for most projects, but is
+unable to display the DAG of projects as large as the Linux Kernel.
 
-\begin{figure}[h]
+German et al. recorded the metadata of every commit and merge in the Linux git
+repository from 2012 to 2015, including data that is not stored by git. We are
+able to use this information to generate sets of trees rather than a DAG. The
+root node of each tree is the top-level merge performed by Linus, which merges
+a given commit into the kernel.
+
+In this paper, we present the design decisions behind a web-based tool that
+demonstrates the advantages of using the tree model over the use of a DAG. Our
+tool provides information about the location of a given commit in the
+respective merge tree, the files edited in the commit or merge, the modules
+edited in the commit or merge, and the commit message of a given commit or
+merge. The tool allows users various filters, including the release, a keyword
+or phrase from a log, the name of the author, or the commit id. The user can
+request all the top-level merges containing a commit or merge that matches the
+query, or all commits and merges that match the query.
+
+\section{Related Work}
+
+Gitk, Gitg, and other similar git repository visualizing tools display the DAG.
+For smaller projects, this model is able to provide meaningful information, but
+in large projects like the Linux kernel, the branches become too tangled and
+deep to derive any meaning from the DAG.
+
+%% Check for papers if they exist
+%% This tool isn't specifically designed for showing Linux
+
+\evantodo{Picture of gitk DAG and github failing}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/gitk.png}
-	\caption{Portion of the Linux Kernel DAG}
-	\label{fig:DAG}
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-\begin{figure}[h]
+University of Maryland presented a treemap design for displaying all the files
+in a given release. This information doesn't provide meaining about what has
+changed. The information within the treemap does not provide clear insight on
+the meaning of the files, what changes were made, or even the size. The treemap
+structure provides information on where in the directory tree the file exists.
+
+%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
+%% Find paper for this maybe?
+
+\evantodo{Picture of visualization}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/github_viewer.png}
-	\caption{Github Error message}
-	\label{fig:github dag}
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-\section{Related Work}
-Daniel German, et. al \evantodo{Find paper} recorded the Linux git repository
-metadata from 2012 until 2015\evan{?}, storing information not available. They
-also developed a heuristic to convert the DAG to a set of trees strung together
-by the root as a linked list. We will call this structure a vine.
+Moghaddam designed a graph showing how various commits in the Kernel were
+linked together. This graph doesn't provide a clear picture of what is
+happening though, and gives no insight to the hierarchy of the data. It
+provides information about who authored the commit, and when the commit was
+authored.
+%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
+%% Find paper for this maybe?
 
-\evantodo{Make an picture of a vine with regards to the Linux kernel}
-
-We built a web-based tool\footnote{\url{li.turingmachine.org}} to experiment
-with the various ways to display and organize the release information of the
-kernel.
-
-%% The parts that make the tool usable
-\section{Navigation}
-
-Our system has multiple methods of allowing users to navigate through the data.
-
-\subsection{Search}
-
-Users search to filter out results they are not interested. We break the data
-down by releases. From there, a user can further trim how many trees they want
-to see by including a begin or end commit date, an author name, the commit id,
-or keyword from a log preview to search for commits that match what they are
-looking for.
-
-\begin{figure}[h]
+\evantodo{Picture of graph}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/search_view.png}
-	\caption{Search view}
-	\label{fig:search_view}
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-\begin{figure}[h]
+\section{Design}
+
+Our primary focus was looking for a way to remove as much information as
+possible to narrow the results that we wanted to return. Right now, the largest
+set of data that a search could return is the entire set of commits that
+pertain to a given release. While this set is massive, it pales in comparison
+to every commit and merge generated for all versions of the kernel. From there,
+the design is centered around the merge tree. These are all of the commits and
+merges that eventually are merged by a top-level merge into the kernel. The
+design centers around a vine-shape. The top-level merges form a linked list,
+each node being a logical section within the kernel. These sections are defined
+by Linus. Within each of these top-level merges is a tree of merges and
+commits. The merges form the inner-nodes, and the commits form the leaf-nodes.
+
+\evantodo{Make a vine diagram}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/results_view.png}
-	\caption{Results when searching for commits by Andrew Morton in Linux
-	3.10}
-	\label{fig:results_view}
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-\subsection{Breadcrumbs}
+Different kinds of users need to be able to use the tool differently.
+Some users may be looking for a top-down approach, trying to find the merge
+closest to the top-level merge that works with a given set of files. Other uses
+may be looking for a bottom-up approach, given a commit, what top-level merge
+includes this commit. Other users are interested in what happens in a given
+release cycle without any other requirements.
 
-\begin{figure}[h]
+To build an effective tool, we focused on two main areas, the navigation
+and presentation.
+
+\subsection{Navigation}
+
+A user is greeted by the search screen. We break the kernel down by release,
+limiting the maximum information a user can query to help minimize being
+overwhelmed. From there, users can either search by author name, commit id, or
+keyword from the log preview. A user may also request the top-level merges that
+meet the requirements. If the checkbox is unchecked, a user is able to query
+all commits and merges that happened in the selected release cycle. If the box
+is checked, the user will get all of the top-level merges in the release cycle
+that contain a commit or merge that match the given criteria. If no criteria
+are provided and the box is checked, the user will simply be given all
+top-level merges for that release cycle. This helps separate the use-cases,
+allowing users wanting the bottom-up approach to perform searches, and users
+wanting a top-down approach to perform searches.
+
+Once a user has found either the top-level merge or commit that they are
+interested in, they should be able to move freely between the commits and
+merges in the tree. We use three mechanisms to provide better navigation
+through the tree. The tree view, the list tree, and breadcrumbs.
+
+The tree view provides most of the
+navigational functionality, allowing users to jump to any commit or merge in
+the current tree. The main tree view displays all the commits and merges, the
+root is the top-level merge of all the commits and merges in the group. Users
+are able to see all commits and merges of all the merges.
+
+\evantodo{Picture of tree}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.0in]{figures/breadcrumbs.png}
-	\caption{Generated Breadcrumbs}
-	\label{fig:breadcrumbs}
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-Our system provides breadcrumbs, allowing a user to move toward the root of the
-tree and view the detailed view of the parent merge.
+The list tree shows the merges and commits of the current subtree, built as
+nested lists of links to the respective commits and merges. The information in
+this tree is a subset of the main tree view, providing a method users with a
+view of what commits and merges are below the current commit or merge. If the
+current position is a commit, the tree will be empty, as commits are leaf
+nodes in the tree representation. This is designed for letting users move
+quickly down in the tree.
 
-\section{Presentation}
-
-\begin{figure}[h]
+\evantodo{Picture of list tree}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/page.png}
-	\caption{View of page}
-	\label{fig:full page}
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-We can provide information about what files were edited by a given commit or
-merge. Due to the tree-structure of the data, we can perform a tree traversal
-to determine the number of files and modules touched within the merges.
+The breadcrumbs show the path of merges that are above the current commit or
+merge, up to the top-level merge. This allows users a fast way to move up in
+the tree.
 
-%% The special parts that make the tool special
-\section{Date Range Queries}
-
-Git does not store the date that a given commit is merged into another branch,
-it stores the date that the commit was authored. This is why the data collected
-by Daniel German becomes important. His recording system monitored each commit
-as it moved through the kernel, storing the commit and author dates of each
-commit.
-
-A commit may be created and not be merged with the kernel until much later. It
-is these merges that make performing date range queries using the conventional
-tools useless as they will not be included even though they are having an
-effect on the kernel in the requested date range. Our system splits the kernel
-into releases, and only looks at the dates of the merges performed by Linus
-directly into the master branch. From there, we are able to construct the trees
-below each of these merges, which may contain commits from before the begin
-date of the selected range, or after the end date for the given range. This
-provides information about the commits that come from older branches that were
-pulled forward, or patches that were merged backward into an older version of
-the kernel as part of maintenance.
-
-\section{Visualization}
-
-Recognizing each merge with the master branch of the kernel as a subtree helps
-remove information that is unnecessarily shown in other tools, like the DAG.
-There are many ways of visualizing trees, including a standard tree
-visualization(\ref{fig:moderate_tree_view}). An issue with a standard tree visualization
-is that it can only display a certain amount of information at each node before
-it becomes cluttered and incomprehensible, furthermore, as we can see in figure 
-(\ref{fig:large_tree_view}), if there are many nodes, even limited information
-becomes cluttered.
-
-\evantodo{Make the lines and nodes bigger and brighter in the tree screenshots}
-\begin{figure}[h]
+\evantodo{Picture of breadcrumbs}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/tree_viewer.png}
-	\caption{Moderately sized tree view tree}
-	\label{fig:moderate_tree_view}
-	% CID: 042dd60ca6dec9a02cefa8edd67de386e35755d6
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-\begin{figure}[h]
+\subsection{Information and Presentation}
+
+Navigation is only effective if the destination contains the desired
+information. We present the commit log, what files were edited, and what
+modules were edited at a given commit or merge. Normal tools are unable to
+provide information about what files are edited in a merge because they do no
+use the tree structure. From our tree structure, we are able to traverse the
+subtree, making note of what files are edited and how many lines are added and
+removed at every commit.
+
+\evantodo{Picture of files}
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=2.5in]{figures/tree_viewer_large.png}
-	\caption{Large tree view tree}
-	\label{fig:large_tree_view}
-	% CID : 5f56886521d6ddd3648777fae44d82382dd8c87f
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{}
 \end{figure}
 
-Perhaps a more useful tree is the list-tree, which allows us to show the
-layers in the tree using indentation. This also allows us to create click-able
-links to the corresponding items in the tree.
+The tree views provide navigation, but are also able to present implied
+information about the depth and number of commits in a given merge.
 
-\begin{figure}[h]
+\section{Discussion}
+
+Our search provides mechanisms for users looking for top-down and bottom-up
+approaches. In figures \ref{fig:AndrewMortonMerges} and \ref{fig:AndrewMorton}
+we see the difference in responses when looking for the merges by Linus and the
+commits authored by Andrew Morton.
+\evantodo{Picture of results for top-down and bottom-up}
+
+%% http://localhost:8080/search?version=Linux+3.14&commit_date_begin=&commit_date_end=&search-text=Morton&search-type=Author&merges=on
+
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=1.5in]{figures/list_viewer.png}
-	\caption{Moderately sized list view tree}
-	\label{fig:moderate_list_view}
-	% CID: 042dd60ca6dec9a02cefa8edd67de386e35755d6
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{fig:AndrewMortonMerges}
 \end{figure}
 
-\begin{figure}[h]
+\begin{figure}[h!]
 	\centering
-	\includegraphics[height=3.5in]{figures/list_viewer_large.png}
-	\caption{Large list view tree}
-	\label{fig:large_list_view}
-	% CID : 5f56886521d6ddd3648777fae44d82382dd8c87f
+	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+	\caption{}
+	\label{fig:AndrewMorton}
 \end{figure}
 
-Figures \ref{fig:moderate_tree_view} and \ref{fig:moderate_list_view} display
-the same subtree, however, the list-based subtree provides a cleaner, faster
-way to navigate the tree, whereas the graphical tree view provides a way to
-hide unnecessary branches of the tree.
+\section{Future Work}
+We have limited search functionality regarding actual files involved in
+commits. If a user knows which file they are working with and what has changed
+in a given file, they are unable to look specifically for commits pertaining to
+that file. Adding a search filter for searching by file would allow maintainers
+to perform this search.
 
-We can see in figures \ref{fig:large_tree_view} and \ref{fig:large_list_view}
-that neither representation of the trees can cope with the larger trees in a
-clean manner. These larger trees are, however, and exception to the rule. These
-trees are from one of the many memory management patches by Andrew Morton,
-which get the name ``Andrews patch bombs''. Andrews patches are very wide trees
-that have only leaves and no internal merging. Conversely, the networking
-modules are far larger with massive breadth and some depth. Overall the Linux
-kernel has a very flat structure wit little depth, the deepest node being $5$
-merges away \evantodo{verify this} from the root merge into the kernel.
+Currently, there are still various features that have bad performance. More
+work is required to get those features responding more quickly, or providing an
+asynchronous mechanism and loading bars to improve user experience.
 
-\evantodo{Build sunburst tree into visualizer, then add pictures}
+At this time, we have no evidence that our tool is able to improve the
+work-flow of maintainers. We believe that the tool is able to improve the
+work-flow and performance of maintainers because it is able to better assist
+users because it provides cleaner mechanisms of working with the data. It is
+able to provide more relevant information, while removing information that is
+irrelevant to a given module or set of merges.
 

### 2015-08-12 linvis.tex
Abstract, introduction, and gitk

@@
 	With an average of over 900 top-level merges into the Linux kernel per
-	release, each containing thousands of commits, maintenance of older
-	versions of the kernel becomes nearly impossible. For security,
-	performance, and changing hardware, maintainers must understand the
-	changes made to current versions of the kernel, and show these changes
-	fit into older versions in order to make the necessary merges and
-	modifications to the older versions. Current tools provide information
-	about repositories through the directed acyclic graph (DAG) which is
-	helpful for smaller projects, but with the scale and number of branches
-	in the kernel, the DAG becomes overwhelming very quickly.
+	release, some containing thousands of commits, many containing hundreds
+	of commits, maintenance of older versions of the kernel becomes nearly
+	impossible. For security, performance, and changing hardware,
+	maintainers must understand the changes made to current versions of the
+	kernel, and show these changes fit into older versions in order to make
+	the necessary merges and modifications to the older versions. Current
+	tools provide information about repositories through the directed
+	acyclic graph (DAG) which is helpful for smaller projects, but with the
+	scale and number of branches in the kernel, the DAG becomes
+	overwhelming very quickly.
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/gitk.png}
+	\caption{DAG view in Gitk}
+	\label{fig:gitk}
+\end{figure}
+
+The second issue with performing searches is that git log, gitk, and other
+tools only store the authored date of a commit. A commit may be merged at any
+time, regardless of when the commit was authored. Because of this, performing
+date-range queries does not provide meaningful results. Performing a date-range
+query using other tools over the span of a release will contain results that
+are for previous versions of the kernel, and may not contain all of the results
+for that kernel if patches were merged after the release date.
+
 German et al. recorded the metadata of every commit and merge in the Linux git
 repository from 2012 to 2015, including data that is not stored by git. We are
-able to use this information to generate sets of trees rather than a DAG. The
+able to use this information to generate sets of trees instead of a DAG. The
 root node of each tree is the top-level merge performed by Linus, which merges
@@
 
-In this paper, we present the design decisions behind a web-based tool that
-demonstrates the advantages of using the tree model over the use of a DAG. Our
-tool provides information about the location of a given commit in the
-respective merge tree, the files edited in the commit or merge, the modules
-edited in the commit or merge, and the commit message of a given commit or
-merge. The tool allows users various filters, including the release, a keyword
-or phrase from a log, the name of the author, or the commit id. The user can
-request all the top-level merges containing a commit or merge that matches the
-query, or all commits and merges that match the query.
+In this paper, we present the design decisions behind a web-based tool,
+demonstrating the advantages of our tree model over the use of a DAG. Our tool
+provides information about the location of a given commit in the respective
+merge tree, the files edited, the modules edited, and the commit message in a
+given commit or merge. The tool allows users to apply various filters, including
+the release, a keyword or phrase from the log preview, the name of the author,
+or the commit id. The user can request all the top-level merges containing a
+commit or merge that matches the query, or all commits and merges that match
+the query.
+
+Our model of the tool provides many advantages other tools. Breaking the data
+by merge tree removes the commits and merges to other components of the kernel
+from the visualization, providing a clearer picture of what is happening in a
+given merge.
+
+No top-level merges are merged to an earlier version of the kernel after the
+release date. Because of our tree construction, we are able to use these
+top-level merges to find all merges effecting a version of the kernel,
+regardless of when they were authored and merged. This provides us with
+mechanisms for performing date-range queries that are able to provide
+meaningful information.
 
@@
 
-Gitk, Gitg, and other similar git repository visualizing tools display the DAG.
-For smaller projects, this model is able to provide meaningful information, but
-in large projects like the Linux kernel, the branches become too tangled and
-deep to derive any meaning from the DAG.
+\subsection{Gitk}
+Gitk, Gitg, and other similar git repository visualizing tools use the DAG as
+the central organization. For smaller projects, this model is sufficient for
+providing meaningful information, but in larger projects like the Linux kernel,
+the branches become too tangled and too deep to derive any meaning from the
+DAG. The DAG model allows users to go to the next commit in the chain, go to
+the previous commit in the chain, or jump to the parent merge. This suggest
+that gitk and similar tools are designed for users working with a bottom-to-top
+approach. With this model, it is difficult for users to see what other commits
+are involved with a given merge. Gitk only provides file information on
+commits, as it is unable to quickly find all files edited by a merge. It is up
+to the user to remember what files were edited in each of the commits involved
+in a given merge. For some of the large merges in the Kernel, this is an
+impossible task. \evantodo{Find a large one and report how many files are
+there} Other DAG visualization systems, like the GitHub viewer, are unable to
+provide a visualization for projects that are as large as the
+kernel.\evantodo{Show picture of github failing}
 
@@
 
-\evantodo{Picture of gitk DAG and github failing}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
+\subsection{Treemap visualizer}
 University of Maryland presented a treemap design for displaying all the files
@@
 \evantodo{Picture of visualization}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
+\subsection{List viewer}
 Moghaddam designed a graph showing how various commits in the Kernel were
@@
 \evantodo{Picture of graph}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
@@
 \evantodo{Make a vine diagram}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
@@
 \evantodo{Picture of tree}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
@@
 \evantodo{Picture of list tree}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
@@
 \evantodo{Picture of breadcrumbs}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
@@
 \evantodo{Picture of files}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{}
+% \end{figure}
 
@@
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{fig:AndrewMortonMerges}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-	\caption{}
-	\label{fig:AndrewMorton}
-\end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{fig:AndrewMortonMerges}
+% \end{figure}
+%
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% 	\caption{}
+% 	\label{fig:AndrewMorton}
+% \end{figure}
 

### 2015-08-12 linvis.tex
Architecture, Pictures, and treemap
Adding architecture section, pictures for the treemap and list viewer, and
adding more information about why the treemap is not suitable for our use-case.

@@
 full view of the DAG, showing all merges and commits. For smaller projects,
-this information is useful to show the relation between various branches. With
-larger, modular projects like the Linux Kernel, the DAG becomes a tangled mess
-of lines that lose meaning. Github provides a DAG view for most projects, but is
+this information is useful to show the relation between various branches.
+With larger, modular projects like the Linux Kernel, the DAG becomes a tangled
+mess (\ref{fig:gitk}). Github provides a DAG view for most projects, but is
 unable to display the DAG of projects as large as the Linux Kernel.
@@
 
-The second issue with performing searches is that git log, gitk, and other
-tools only store the authored date of a commit. A commit may be merged at any
-time, regardless of when the commit was authored. Because of this, performing
+Another issue with performing searches is that git log, gitk, and other tools
+only store the authored date of a commit. A commit may be merged at any time,
+regardless of when the commit was authored. Because of this, performing
 date-range queries does not provide meaningful results. Performing a date-range
@@
 merge tree, the files edited, the modules edited, and the commit message in a
-given commit or merge. The tool allows users to apply various filters, including
-the release, a keyword or phrase from the log preview, the name of the author,
-or the commit id. The user can request all the top-level merges containing a
-commit or merge that matches the query, or all commits and merges that match
-the query.
+given commit or merge. The tool allows users to apply various filters,
+including the release, a keyword or phrase from the log preview, the name of
+the author, or the commit id. The user can request all the top-level merges
+containing a commit or merge that matches the query, or all commits and merges
+that match the query.
 
@@
 University of Maryland presented a treemap design for displaying all the files
-in a given release. This information doesn't provide meaining about what has
-changed. The information within the treemap does not provide clear insight on
-the meaning of the files, what changes were made, or even the size. The treemap
-structure provides information on where in the directory tree the file exists.
+in a given release. Treemaps are good for displaying the topology of shallow,
+wide trees, making it a good candidate for displaying file systems. Treemaps
+are limited in what information can be displayed. The information within the
+treemap does not provide clear insight on what changes were made to the files
+in the given release. The treemap visualization only provides information on
+where the file is located within the filesystem\ref{fig:treemap}. File systems
+and the merge tree of the kernel share many similarities, in both cases, the
+representative trees are wide and shallow, suggesting that a treemap could
+potentially work to visualize the commit and merge structure of the kernel.
 
+The treemap is unable to take into account multiple merges at the same level
+with the same name. There is an additional dimension, in that the commits and
+merges are ordered in time. The treemap design is unable to show the ordering
+of the merges and commits.
+
+The treemap design is also designed for working with data of differing types.
+In the filesystem, there are many kinds of files.  There are text files,
+scripts, images, binary files, and other types. These can be given colours
+making differences stand out. There is no obvious way to assign meaningful
+types to the commits such that all the commits within a given merge do not all
+have the same colour.
+
+Finally, the treemap does not provide any additional information about the
+contents of each cell, only that the cell exists. The user is unable to see
+their current position within the map, or if we assume that their current
+position is the larges surrounding box, they can no-longer see how their
+current position interacts with the rest of the system.
 %% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
@@
 
-\evantodo{Picture of visualization}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
+	\caption{Treemap of Linux 2.5.33 file structure}
+	\label{fig:treemap}
+\end{figure}
 
 \subsection{List viewer}
-Moghaddam designed a graph showing how various commits in the Kernel were
-linked together. This graph doesn't provide a clear picture of what is
-happening though, and gives no insight to the hierarchy of the data. It
-provides information about who authored the commit, and when the commit was
-authored.
+Moghaddam designed a web-based visualization tool displaying the chain of
+merges into the kernel. There is no information on what commits are within a
+given merge, or what files were edited. The visualization tool provides
+information about who authored the commit and when it was authored.
+\evantodo{Link to tool}
 %% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
@@
 
-\evantodo{Picture of graph}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/gitvis.png}
+	\caption{List Viewer}
+	\label{fig:listviewer}
+\end{figure}
+
+\section{Model}
+
+We use a tree.
 
@@
 
-To build an effective tool, we focused on two main areas, the navigation
-and presentation.
+\subsection{Architecture}
+
+% Database
+We split the data presented among 5 tables; commits, filesmod, logs,
+pathtomerge, and releases. The commits table stores the metadata about a given
+commit or merge. This includes the author, the authored date, the committer,
+commit date, and the file-level diff. The filesmod table stores the information
+about the files modified in a commit. We have the commit id, or cid, of the
+commit, the filename, the number of lines added and the number of lines removed
+from the file. There may be multiple files within a given commit, and there may
+be multiple commits editing the same file, the database uses an index as the
+primary key, though the primary key could be the cid, filename pair.
+
+% Tables
+commits
+filesmod
+logs
+pathtomerge
+releases
 
@@
 use the tree structure. From our tree structure, we are able to traverse the
-subtree, making note of what files are edited and how many lines are added and
-removed at every commit.
+subtree, making note of what files are edited and how many lines of code are
+added and removed at every commit.
 

### 2015-08-14 linvis.tex
Introduction: Re-working intro and related works

@@
 commits with tools that are unable to filter and effectively visualize projects
-that are of the scale of the kernel. Tools like gitk provide users with the
+that are of the scale of the kernel. Tools like Gitk provide users with the
 full view of the DAG, showing all merges and commits. For smaller projects,
-this information is useful to show the relation between various branches.
-With larger, modular projects like the Linux Kernel, the DAG becomes a tangled
-mess (\ref{fig:gitk}). Github provides a DAG view for most projects, but is
-unable to display the DAG of projects as large as the Linux Kernel.
+this information is useful for showing the relation between various branches.
+With large modular projects, like the Linux Kernel, the DAG becomes a tangled
+mess (\ref{fig:gitk}) of commits, merges, and the links between them, making it
+difficult to derive an explanation of the changes.  Github provides a DAG view
+for most projects, but is unable to display the DAG of projects as large as the
+Linux Kernel.
 
@@
 
-Another issue with performing searches is that git log, gitk, and other tools
-only store the authored date of a commit. A commit may be merged at any time,
-regardless of when the commit was authored. Because of this, performing
-date-range queries does not provide meaningful results. Performing a date-range
-query using other tools over the span of a release will contain results that
-are for previous versions of the kernel, and may not contain all of the results
-for that kernel if patches were merged after the release date.
+Git log, Gitk, and other similar tools only store the authored date of a
+commit. A commit may be merged into the project at any time, regardless of when
+the commit was authored. Because of this, performing date-range queries does
+not provide meaningful results. Performing a date-range query using other tools
+over the span of a release will contain results that are for previous versions
+of the kernel, and may not contain all of the results for that kernel if
+patches were merged after the release date.
 
@@
 able to use this information to generate sets of trees instead of a DAG. The
-root node of each tree is the top-level merge performed by Linus, which merges
-a given commit into the kernel.
+root of each tree is the top-level merge performed by Linus, which merges a
+given commit into the kernel.
 
-In this paper, we present the design decisions behind a web-based tool,
-demonstrating the advantages of our tree model over the use of a DAG. Our tool
-provides information about the location of a given commit in the respective
-merge tree, the files edited, the modules edited, and the commit message in a
-given commit or merge. The tool allows users to apply various filters,
-including the release, a keyword or phrase from the log preview, the name of
-the author, or the commit id. The user can request all the top-level merges
-containing a commit or merge that matches the query, or all commits and merges
-that match the query.
+In this paper, we present the design decisions behind a web-based tool built
+around a tree-based model instead of the DAG. We demonstrate the advantages of
+this model over the use of the DAG. Our tool provides information about the
+location of a given commit in the respective merge tree, the files edited, the
+modules edited, and the commit message in a given commit or merge. The tool
+allows users to apply various filters, including the release, a keyword or
+phrase from the log preview, the name of the author, or the commit id. The user
+can request all the top-level merges containing a commit or merge that matches
+the query, or all commits and merges that match the query.
 
-Our model of the tool provides many advantages other tools. Breaking the data
-by merge tree removes the commits and merges to other components of the kernel
+Our core model provides many advantages to other tools. Breaking the data by
+merge tree removes the commits and merges of other components of the kernel
 from the visualization, providing a clearer picture of what is happening in a
-given merge.
+given merge. This model provides additional support to edited file information.
 
-No top-level merges are merged to an earlier version of the kernel after the
-release date. Because of our tree construction, we are able to use these
-top-level merges to find all merges effecting a version of the kernel,
-regardless of when they were authored and merged. This provides us with
-mechanisms for performing date-range queries that are able to provide
-meaningful information.
+All top-level merges to a given version of the kernel are merged between the
+start and release date. Our tree construction allows us to use the top-level
+merges to find all merges and commits that have an effect on a given version of
+the kernel. The model allows us to better provide mechanisms for performing
+date-range queries, providing access to all commits that have an effect on a
+given release, even if the commit comes before the start date, or the release
+date.
 
@@
 the previous commit in the chain, or jump to the parent merge. This suggest
-that gitk and similar tools are designed for users working with a bottom-to-top
+that Gitk and similar tools are designed for users working with a bottom-to-top
 approach. With this model, it is difficult for users to see what other commits
@@
 to the user to remember what files were edited in each of the commits involved
-in a given merge. For some of the large merges in the Kernel, this is an
-impossible task. \evantodo{Find a large one and report how many files are
-there} Other DAG visualization systems, like the GitHub viewer, are unable to
-provide a visualization for projects that are as large as the
-kernel.\evantodo{Show picture of github failing}
+in a given merge. In some of the larger merges, some containing over 1500
+edited files, this task become nearly impossible.
+% cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
+Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}),
+are unable to provide a visualization for projects that are as large as the
+kernel.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
+	\caption{Github Failure Message}
+	\label{fig:gitfail}
+\end{figure}
 
@@
 
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
-
 \subsection{Treemap visualizer}
-University of Maryland presented a treemap design for displaying all the files
-in a given release. Treemaps are good for displaying the topology of shallow,
-wide trees, making it a good candidate for displaying file systems. Treemaps
-are limited in what information can be displayed. The information within the
-treemap does not provide clear insight on what changes were made to the files
-in the given release. The treemap visualization only provides information on
-where the file is located within the filesystem\ref{fig:treemap}. File systems
-and the merge tree of the kernel share many similarities, in both cases, the
-representative trees are wide and shallow, suggesting that a treemap could
-potentially work to visualize the commit and merge structure of the kernel.
+University of Maryland presented a treemap design (\ref{fig:treemap}) for
+displaying all the file structure of a given version. Treemaps are good for
+displaying the topology of shallow, wide trees, making it a good candidate for
+displaying file systems. Treemaps are limited in what information can be
+displayed. The information within the treemap does not provide clear insight on
+what changes were made to the files in the given release. The treemap
+visualization only provides information on where the file is located within the
+filesystem. File systems and the merge tree of the kernel share many
+similarities, in both cases, the representative trees are wide and shallow,
+suggesting that a treemap could potentially work to visualize the commit and
+merge structure of the kernel.
 
@@
 
-We use a tree.
-
 \section{Design}
 
-Our primary focus was looking for a way to remove as much information as
+The goal is to build a system that allows users to easily navigate and
+understand the changes within the kernel. To do this, the user must have the
+ability to perform searches
+
+The goal was to build a system that would allow users to easily navigate the
+commit information of
+
+Our primary focus was looking for a way to remove as much unnecessary
+information as possible narrow the results.
+
+information as
 possible to narrow the results that we wanted to return. Right now, the largest
@@
 % Database
-We split the data presented among 5 tables; commits, filesmod, logs,
+We split the data presented among five tables; commits, filesmod, logs,
 pathtomerge, and releases. The commits table stores the metadata about a given
@@
 % Tables
-commits
-filesmod
-logs
-pathtomerge
-releases
+% commits
+% filesmod
+% logs
+% pathtomerge
+% releases
+
+%
 

### 2015-08-15 linvis.tex
Core Model and Implementation Details: Starting section
I've started talking about the core model and the implementation details of the
system. I've commented out the design section.

@@
 
-\section{Model}
+\section{Core Model}
+Each merge may contain zero or more children of other merges or commits.
+Commits have no children, and therefore are always leaf nodes, but commits
+contain information about what files are edited, how many lines are edited, and
+other important information.
 
-\section{Design}
+Core Model:
+	Merges are inner nodes
+	Commits are leaf nodes
 
-The goal is to build a system that allows users to easily navigate and
-understand the changes within the kernel. To do this, the user must have the
-ability to perform searches
+\section{Implementation Details}
 
-The goal was to build a system that would allow users to easily navigate the
-commit information of
+We built a we-based tool to provide information about the kernel.
+We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
+as the framework for generating the dynamic content.
 
-Our primary focus was looking for a way to remove as much unnecessary
-information as possible narrow the results.
+Nginx is a high-performance, scalable HTTP and mail server. Started in 2002,
+the project is much younger than the Apache project, and designed with massive
+scalability and multi-processing in mind. Instead of using request-driven
+threading, Nginx uses an event-driven architecture, making it more scalable,
+proving itself as being the web server behind Netflix, Hulu, GitHub, Tumblr,
+and many other popular sites.
 
-information as
-possible to narrow the results that we wanted to return. Right now, the largest
-set of data that a search could return is the entire set of commits that
-pertain to a given release. While this set is massive, it pales in comparison
-to every commit and merge generated for all versions of the kernel. From there,
-the design is centered around the merge tree. These are all of the commits and
-merges that eventually are merged by a top-level merge into the kernel. The
-design centers around a vine-shape. The top-level merges form a linked list,
-each node being a logical section within the kernel. These sections are defined
-by Linus. Within each of these top-level merges is a tree of merges and
-commits. The merges form the inner-nodes, and the commits form the leaf-nodes.
+Postgresql is our dbms of choice, providing fast and reliable results. There
+are multiple reasons behind choosing postgresql over another dbms. The main
+reason being that the dataset provided from \evantodo{paper} was generated from
+a postgresql database, allowing us to import the data directly without any
+further conversions. The secondary reason is that we are more familiar with
+postgresql than the other systems available.
 
-\evantodo{Make a vine diagram}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
+We chose the python micro-framework Flask for generating the dynamic content.
+Working with flask is simple and easy.
 
-Different kinds of users need to be able to use the tool differently.
-Some users may be looking for a top-down approach, trying to find the merge
-closest to the top-level merge that works with a given set of files. Other uses
-may be looking for a bottom-up approach, given a commit, what top-level merge
-includes this commit. Other users are interested in what happens in a given
-release cycle without any other requirements.
+Tool Implementation Details:
+	Web Server: Nginx, Flask, postgresql
+	Database: 5 tables
 
-\subsection{Architecture}
-
-% Database
-We split the data presented among five tables; commits, filesmod, logs,
-pathtomerge, and releases. The commits table stores the metadata about a given
-commit or merge. This includes the author, the authored date, the committer,
-commit date, and the file-level diff. The filesmod table stores the information
-about the files modified in a commit. We have the commit id, or cid, of the
-commit, the filename, the number of lines added and the number of lines removed
-from the file. There may be multiple files within a given commit, and there may
-be multiple commits editing the same file, the database uses an index as the
-primary key, though the primary key could be the cid, filename pair.
-
-% Tables
-% commits
-% filesmod
-% logs
-% pathtomerge
-% releases
-
+% \section{Design}
 % 
-
-\subsection{Navigation}
-
-A user is greeted by the search screen. We break the kernel down by release,
-limiting the maximum information a user can query to help minimize being
-overwhelmed. From there, users can either search by author name, commit id, or
-keyword from the log preview. A user may also request the top-level merges that
-meet the requirements. If the checkbox is unchecked, a user is able to query
-all commits and merges that happened in the selected release cycle. If the box
-is checked, the user will get all of the top-level merges in the release cycle
-that contain a commit or merge that match the given criteria. If no criteria
-are provided and the box is checked, the user will simply be given all
-top-level merges for that release cycle. This helps separate the use-cases,
-allowing users wanting the bottom-up approach to perform searches, and users
-wanting a top-down approach to perform searches.
-
-Once a user has found either the top-level merge or commit that they are
-interested in, they should be able to move freely between the commits and
-merges in the tree. We use three mechanisms to provide better navigation
-through the tree. The tree view, the list tree, and breadcrumbs.
-
-The tree view provides most of the
-navigational functionality, allowing users to jump to any commit or merge in
-the current tree. The main tree view displays all the commits and merges, the
-root is the top-level merge of all the commits and merges in the group. Users
-are able to see all commits and merges of all the merges.
-
-\evantodo{Picture of tree}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
-
-The list tree shows the merges and commits of the current subtree, built as
-nested lists of links to the respective commits and merges. The information in
-this tree is a subset of the main tree view, providing a method users with a
-view of what commits and merges are below the current commit or merge. If the
-current position is a commit, the tree will be empty, as commits are leaf
-nodes in the tree representation. This is designed for letting users move
-quickly down in the tree.
-
-\evantodo{Picture of list tree}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
-
-The breadcrumbs show the path of merges that are above the current commit or
-merge, up to the top-level merge. This allows users a fast way to move up in
-the tree.
-
-\evantodo{Picture of breadcrumbs}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
-
-\subsection{Information and Presentation}
-
-Navigation is only effective if the destination contains the desired
-information. We present the commit log, what files were edited, and what
-modules were edited at a given commit or merge. Normal tools are unable to
-provide information about what files are edited in a merge because they do no
-use the tree structure. From our tree structure, we are able to traverse the
-subtree, making note of what files are edited and how many lines of code are
-added and removed at every commit.
-
-\evantodo{Picture of files}
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{}
-% \end{figure}
-
-The tree views provide navigation, but are also able to present implied
-information about the depth and number of commits in a given merge.
-
-\section{Discussion}
-
-Our search provides mechanisms for users looking for top-down and bottom-up
-approaches. In figures \ref{fig:AndrewMortonMerges} and \ref{fig:AndrewMorton}
-we see the difference in responses when looking for the merges by Linus and the
-commits authored by Andrew Morton.
-\evantodo{Picture of results for top-down and bottom-up}
-
-%% http://localhost:8080/search?version=Linux+3.14&commit_date_begin=&commit_date_end=&search-text=Morton&search-type=Author&merges=on
-
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{fig:AndrewMortonMerges}
-% \end{figure}
+% The goal is to build a system that allows users to easily navigate and
+% understand the changes within the kernel. We provide a mechanism for performing
+% searches to filter out irrelevant commits and merges, and only work with
+% information that they are interested in. At the top level, the commits are
+% grouped by release version. \evan{We currently do not provide a mechanism for
+% checking real versions and release candidates\ldots Might be useful to at least
+% note in search results section}.
 % 
-% \begin{figure}[h!]
-% 	\centering
-% 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% 	\caption{}
-% 	\label{fig:AndrewMorton}
-% \end{figure}
+% 
+% information as
+% possible to narrow the results that we wanted to return. Right now, the largest
+% set of data that a search could return is the entire set of commits that
+% pertain to a given release. While this set is massive, it pales in comparison
+% to every commit and merge generated for all versions of the kernel. From there,
+% the design is centered around the merge tree. These are all of the commits and
+% merges that eventually are merged by a top-level merge into the kernel. The
+% design centers around a vine-shape. The top-level merges form a linked list,
+% each node being a logical section within the kernel. These sections are defined
+% by Linus. Within each of these top-level merges is a tree of merges and
+% commits. The merges form the inner-nodes, and the commits form the leaf-nodes.
+% 
+% \evantodo{Make a vine diagram}
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{}
+% % \end{figure}
+% 
+% 
+% Different kinds of users need to be able to use the tool differently.
+% Some users may be looking for a top-down approach, trying to find the merge
+% closest to the top-level merge that works with a given set of files. Other uses
+% may be looking for a bottom-up approach, given a commit, what top-level merge
+% includes this commit. Other users are interested in what happens in a given
+% release cycle without any other requirements.
+% 
+% \subsection{Architecture}
+% 
+% % Database
+% We split the data presented among five tables; commits, filesmod, logs,
+% pathtomerge, and releases. The commits table stores the metadata about a given
+% commit or merge. This includes the author, the authored date, the committer,
+% commit date, and the file-level diff. The filesmod table stores the information
+% about the files modified in a commit. We have the commit id, or cid, of the
+% commit, the filename, the number of lines added and the number of lines removed
+% from the file. There may be multiple files within a given commit, and there may
+% be multiple commits editing the same file, the database uses an index as the
+% primary key, though the primary key could be the cid, filename pair.
+% 
+% % Tables
+% % commits
+% % filesmod
+% % logs
+% % pathtomerge
+% % releases
+% 
+% %
+% 
+% 
+% \subsection{Navigation}
+% 
+% A user is greeted by the search screen. We break the kernel down by release,
+% limiting the maximum information a user can query to help minimize being
+% overwhelmed. From there, users can either search by author name, commit id, or
+% keyword from the log preview. A user may also request the top-level merges that
+% meet the requirements. If the checkbox is unchecked, a user is able to query
+% all commits and merges that happened in the selected release cycle. If the box
+% is checked, the user will get all of the top-level merges in the release cycle
+% that contain a commit or merge that match the given criteria. If no criteria
+% are provided and the box is checked, the user will simply be given all
+% top-level merges for that release cycle. This helps separate the use-cases,
+% allowing users wanting the bottom-up approach to perform searches, and users
+% wanting a top-down approach to perform searches.
+% 
+% Once a user has found either the top-level merge or commit that they are
+% interested in, they should be able to move freely between the commits and
+% merges in the tree. We use three mechanisms to provide better navigation
+% through the tree. The tree view, the list tree, and breadcrumbs.
+% 
+% The tree view provides most of the
+% navigational functionality, allowing users to jump to any commit or merge in
+% the current tree. The main tree view displays all the commits and merges, the
+% root is the top-level merge of all the commits and merges in the group. Users
+% are able to see all commits and merges of all the merges.
+% 
+% \evantodo{Picture of tree}
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{}
+% % \end{figure}
+% 
+% The list tree shows the merges and commits of the current subtree, built as
+% nested lists of links to the respective commits and merges. The information in
+% this tree is a subset of the main tree view, providing a method users with a
+% view of what commits and merges are below the current commit or merge. If the
+% current position is a commit, the tree will be empty, as commits are leaf
+% nodes in the tree representation. This is designed for letting users move
+% quickly down in the tree.
+% 
+% \evantodo{Picture of list tree}
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{}
+% % \end{figure}
+% 
+% 
+% The breadcrumbs show the path of merges that are above the current commit or
+% merge, up to the top-level merge. This allows users a fast way to move up in
+% the tree.
+% 
+% \evantodo{Picture of breadcrumbs}
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{}
+% % \end{figure}
+% 
+% 
+% \subsection{Information and Presentation}
+% 
+% Navigation is only effective if the destination contains the desired
+% information. We present the commit log, what files were edited, and what
+% modules were edited at a given commit or merge. Normal tools are unable to
+% provide information about what files are edited in a merge because they do no
+% use the tree structure. From our tree structure, we are able to traverse the
+% subtree, making note of what files are edited and how many lines of code are
+% added and removed at every commit.
+% 
+% \evantodo{Picture of files}
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{}
+% % \end{figure}
+% 
+% The tree views provide navigation, but are also able to present implied
+% information about the depth and number of commits in a given merge.
+% 
+% \section{Discussion}
+% 
+% Our search provides mechanisms for users looking for top-down and bottom-up
+% approaches. In figures \ref{fig:AndrewMortonMerges} and \ref{fig:AndrewMorton}
+% we see the difference in responses when looking for the merges by Linus and the
+% commits authored by Andrew Morton.
+% \evantodo{Picture of results for top-down and bottom-up}
+% 
+% %% http://localhost:8080/search?version=Linux+3.14&commit_date_begin=&commit_date_end=&search-text=Morton&search-type=Author&merges=on
+% 
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{fig:AndrewMortonMerges}
+% % \end{figure}
+% %
+% % \begin{figure}[h!]
+% % 	\centering
+% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
+% % 	\caption{}
+% % 	\label{fig:AndrewMorton}
+% % \end{figure}
 

### 2015-08-15 linvis.tex
Design: CLI-Shell, Web-based tool
Re-writing design section, talking about the components of the design. Also
including information on the original design of the command line navigation
interface. A new picture is probably a good idea for the shell.

@@
 
-% \section{Design}
-% 
-% The goal is to build a system that allows users to easily navigate and
-% understand the changes within the kernel. We provide a mechanism for performing
-% searches to filter out irrelevant commits and merges, and only work with
-% information that they are interested in. At the top level, the commits are
-% grouped by release version. \evan{We currently do not provide a mechanism for
-% checking real versions and release candidates\ldots Might be useful to at least
-% note in search results section}.
-% 
-% 
-% information as
-% possible to narrow the results that we wanted to return. Right now, the largest
-% set of data that a search could return is the entire set of commits that
-% pertain to a given release. While this set is massive, it pales in comparison
-% to every commit and merge generated for all versions of the kernel. From there,
-% the design is centered around the merge tree. These are all of the commits and
-% merges that eventually are merged by a top-level merge into the kernel. The
-% design centers around a vine-shape. The top-level merges form a linked list,
-% each node being a logical section within the kernel. These sections are defined
-% by Linus. Within each of these top-level merges is a tree of merges and
-% commits. The merges form the inner-nodes, and the commits form the leaf-nodes.
-% 
-% \evantodo{Make a vine diagram}
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{}
-% % \end{figure}
-% 
-% 
-% Different kinds of users need to be able to use the tool differently.
-% Some users may be looking for a top-down approach, trying to find the merge
-% closest to the top-level merge that works with a given set of files. Other uses
-% may be looking for a bottom-up approach, given a commit, what top-level merge
-% includes this commit. Other users are interested in what happens in a given
-% release cycle without any other requirements.
-% 
-% \subsection{Architecture}
-% 
-% % Database
-% We split the data presented among five tables; commits, filesmod, logs,
-% pathtomerge, and releases. The commits table stores the metadata about a given
-% commit or merge. This includes the author, the authored date, the committer,
-% commit date, and the file-level diff. The filesmod table stores the information
-% about the files modified in a commit. We have the commit id, or cid, of the
-% commit, the filename, the number of lines added and the number of lines removed
-% from the file. There may be multiple files within a given commit, and there may
-% be multiple commits editing the same file, the database uses an index as the
-% primary key, though the primary key could be the cid, filename pair.
-% 
-% % Tables
-% % commits
-% % filesmod
-% % logs
-% % pathtomerge
-% % releases
-% 
-% %
-% 
-% 
-% \subsection{Navigation}
-% 
-% A user is greeted by the search screen. We break the kernel down by release,
-% limiting the maximum information a user can query to help minimize being
-% overwhelmed. From there, users can either search by author name, commit id, or
-% keyword from the log preview. A user may also request the top-level merges that
-% meet the requirements. If the checkbox is unchecked, a user is able to query
-% all commits and merges that happened in the selected release cycle. If the box
-% is checked, the user will get all of the top-level merges in the release cycle
-% that contain a commit or merge that match the given criteria. If no criteria
-% are provided and the box is checked, the user will simply be given all
-% top-level merges for that release cycle. This helps separate the use-cases,
-% allowing users wanting the bottom-up approach to perform searches, and users
-% wanting a top-down approach to perform searches.
-% 
-% Once a user has found either the top-level merge or commit that they are
-% interested in, they should be able to move freely between the commits and
-% merges in the tree. We use three mechanisms to provide better navigation
-% through the tree. The tree view, the list tree, and breadcrumbs.
-% 
-% The tree view provides most of the
-% navigational functionality, allowing users to jump to any commit or merge in
-% the current tree. The main tree view displays all the commits and merges, the
-% root is the top-level merge of all the commits and merges in the group. Users
-% are able to see all commits and merges of all the merges.
-% 
-% \evantodo{Picture of tree}
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{}
-% % \end{figure}
-% 
-% The list tree shows the merges and commits of the current subtree, built as
-% nested lists of links to the respective commits and merges. The information in
-% this tree is a subset of the main tree view, providing a method users with a
-% view of what commits and merges are below the current commit or merge. If the
-% current position is a commit, the tree will be empty, as commits are leaf
-% nodes in the tree representation. This is designed for letting users move
-% quickly down in the tree.
-% 
-% \evantodo{Picture of list tree}
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{}
-% % \end{figure}
-% 
-% 
-% The breadcrumbs show the path of merges that are above the current commit or
-% merge, up to the top-level merge. This allows users a fast way to move up in
-% the tree.
-% 
-% \evantodo{Picture of breadcrumbs}
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{}
-% % \end{figure}
-% 
-% 
-% \subsection{Information and Presentation}
-% 
-% Navigation is only effective if the destination contains the desired
-% information. We present the commit log, what files were edited, and what
-% modules were edited at a given commit or merge. Normal tools are unable to
-% provide information about what files are edited in a merge because they do no
-% use the tree structure. From our tree structure, we are able to traverse the
-% subtree, making note of what files are edited and how many lines of code are
-% added and removed at every commit.
-% 
-% \evantodo{Picture of files}
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{}
-% % \end{figure}
-% 
-% The tree views provide navigation, but are also able to present implied
-% information about the depth and number of commits in a given merge.
-% 
-% \section{Discussion}
-% 
-% Our search provides mechanisms for users looking for top-down and bottom-up
-% approaches. In figures \ref{fig:AndrewMortonMerges} and \ref{fig:AndrewMorton}
-% we see the difference in responses when looking for the merges by Linus and the
-% commits authored by Andrew Morton.
-% \evantodo{Picture of results for top-down and bottom-up}
-% 
-% %% http://localhost:8080/search?version=Linux+3.14&commit_date_begin=&commit_date_end=&search-text=Morton&search-type=Author&merges=on
-% 
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{fig:AndrewMortonMerges}
-% % \end{figure}
-% %
-% % \begin{figure}[h!]
-% % 	\centering
-% % 	\includegraphics[width=0.47\textwidth]{figures/empty.png}
-% % 	\caption{}
-% % 	\label{fig:AndrewMorton}
-% % \end{figure}
+\section{Design}
+
+The goal is to build a tool that make navigation of the kernel commit
+information simpler, while providing an explanation of the changes occurring at
+each commit and merge. Part of making the information usable is to narrow the
+results down into only the information that is relevant to what the user is
+looking for.
+
+\subsection{Command-line Shell}
+
+Our initial intuition was to work with the commits and merges as trees, like in
+a file structure. There are many similarities in the structure of a file system
+and the structure of the commit information within the kernel. The primary
+similarity is that they are both trees, and both trees are generally fairly
+shallow and very wide. File systems have been around for quite some time, and
+there are many tools for visualizing and working withe the file structure. The
+first tool was the command-line. We built a small command-line shell
+(\ref{fig:shell}) that allowed navigation through the commit and merges instead
+of files and directory. The shell design permitted a user to move through the
+merges like they would through directories. They could use the cat command to
+see the preview of the log, or the more command to see the preview and what
+files were edited in the commit and how many lines were added and removed from
+each file.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/shell.png}
+	\caption{Shell Design}
+	\label{fig:shell}
+\end{figure}
+
+\evantodo{Put this in discussion section}
+The main issue with this design is that a user will always move from the
+top-down without any additional filtering. Furthermore, it presents the user
+with the commit hashes, which provide no explanation of what the commit or
+merge may contain.
+
+\subsection{Web-based Tool}
+
+There are multiple reasons for building the system into a web-based tool.
+Putting the tool onto the web enables users to use the system without having to
+install anything, making the system platform independent. A web-based tool can
+utilize more interactive means of navigation, and provide a better explanation
+of what each commit or merge may contain.
+
+The web-based tool provides many mechanisms to enable users to better navigate
+and understand changes in the kernel. These mechanisms include searching,
+commit messages, files edited, modules edited, and two types of tree viewers.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/search.png}
+	\caption{Search View}
+	\label{fig:search}
+\end{figure}
+
+Searching allows a user to filter our commits and merges that are irrelevant.
+The search mechanism breaks down the results by release version. A user can
+further narrow down the search range using the commit date range. This commit
+date refers to the date of the creation of the top-level merge. Commits and
+merges that are returned may have a commit date beyond the given range if they
+were merged to a top-level merge within the date range. A user may then provide
+a search text, filtering on the author name, the commit id, or keyword from the
+log. Any part of the author name may show up in the results, including
+searching by email address. The commit ID must be specified from the first
+character to the last character. For example, the commit
+`c267548755a184ef97301071300c1739a564e135' is returned when the user searches
+for a commit id of `c26' in the 3.10 Linux kernel.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/search_results.png}
+	\caption{Search Results}
+	\label{fig:results}
+\end{figure}
+
+In the search results , the user is presented with the one-line log message
+preview, the author's name and email, the date the commit was authored, and the
+date the commit was committed. \evantodo{a better definition for commit date}
+The commit date and author date will usually be the same, however, if the
+commit has been rebased, the commit date will be different than the authored
+date, and will always be after the authored date. The first and third entries
+in figure \ref{fig:results} have been rebased in the past. \evan{We are unable
+to show the rebases.}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/message_view.png}
+	\caption{Commit Message}
+	\label{fig:message}
+\end{figure}
+
+The first view displays the full commit log. From this, a user is able to see
+hat they would see had they searched for the commit using git log. This doesn't
+provide additional information to the other tools, but helps to complete the
+functionality of the tool. This information provides a user with the
+information about the content of the commit and who has looked over the commit
+to ensure that it is good quality. The message for merges may also contain the
+information about the commits that are being merged into a given merge. The
+information within these messages is highly variable, and is completely
+dependent on the author's style. As the user moves toward the top-level merge,
+the quality of these messages generally improve.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/file_view.png}
+	\caption{Commit Files}
+	\label{fig:files}
+\end{figure}
+
+The second tab is the files tab. This tab provides information on what files
+have been edited, how many lines were added, and how many lines were removed in
+a given commit. This functionality is similar to the other tools available. Our
+tree-based design model allows us to extend this functionality to merges, which
+the other tools are unable to show. When a file is edited multiple times from
+different commits within a merge, we simply add the added and removed lines to
+the values already stored. \evan{At this time, we don't have a mechanism to
+process the diff information to determine which lines were changed, and in what
+order.}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/tree_view.png}
+	\caption{Commit Tree View}
+	\label{fig:files}
+\end{figure}
+
+The third tab shows the tree view. The tree view is what makes this tool unique
+to the other tools. It provides a full view of the entire merge tree. It
+originally centres at the current commit or merge, but allows a user to inspect
+the other commits and merges surrounding it. A user may then navigate to any of
+the commits by simply double-clicking on it.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
+	\caption{Tree zoomed out}
+	\label{fig:files}
+\end{figure}
+
+Some of the merge trees are very large, containing thousands of commits and
+merges. Clicking on subtrees will cause them to minimize, only showing parts of
+the tree that are of interest to the user. Furthermore, the user can zoom the
+tree by scrolling the mouse wheel to see more of the topology of the tree, or
+to see the details of a given part of the tree.
+
+The last tab contains another type of tree view. This tree is in the form of
+nested lists, and is designed to more closely model the tree hierarchy view of
+file browsers. This tree only contains the commits and merges that are within
+the subtrees of the current merge. A commit will never have any items in this
+tab, as it is a leaf node, so it cannot contain subtrees. To accompany the
+tree, we include breadcrumbs at the top. The last item in the breadcrumb list
+is the current commit, the next item is the parent of the current commit, and
+the first item is the top-level merge into the kernel.
 

### 2015-08-15 linvis.tex
Implementation Details: Database Tables
Working on describing the design of the database. Still need to talk about the
heuristics used for converting the DAG into a tree structure.

@@
 
-Tool Implementation Details:
-	Web Server: Nginx, Flask, postgresql
-	Database: 5 tables
+The database itself is broken down into 5 main tables; commits, filesmod, logs,
+pathtomerge, and releases. Commits contains most of the metadata for each
+commit, it contains the author, the authored date, and committer, the commit
+date, a boolean saying what is a merge, and the diff. Filesmod contains the
+information on what files were modified for the commits. Since a file may be
+modified by multiple commits, and a commit may be modifying multiple files, we
+can do one of two things for the primary key. We could use both the filename
+and the commit id as the primary key, but we just use an index. The filesmod
+table contains the filename, the commit id, the number of lines added, and the
+number of lines removed. Trivially, it contains the index, but that isn't
+necessary for anything beyond the primary key. Logs contains the information
+for each commit log, this is the one-liner preview message and the full
+message. Pathtomerge is the important table. Pathtomerge contains the commit
+id, the number of merges between the commit and the top-level merge with the
+kernel, the next merge in the merge path up to the top-level merge with the
+kernel, the top-level merge into the kernel, and when it get merged into the
+kernel. The information in this table must be generated periodically by the
+server because we are unable to generate the tree structure correctly straight
+from the information stored by git. Pathtomerge does not contain any top-level
+merges, as these will never have a next merge in the tree, since we are only
+storing the parents and not the children, storing the top-level merge is
+redundant and unnecessary. The releases table stores the release information.
+It contains the version name, whether the version is a candidate release or a
+real release, the previous version, the previous non-candidate release version
+the previous version commit id, the previous non-candidate release version
+commit id, and the commit id of this version. The version commit id represents
+the last top-level merge into the version of the kernel. Using this and the
+commits table, we are able to determine the release date of any version of the
+kernel. Then using the previous release date, we can find the range of dates
+where valid top-level merges are found for that version.
+
+\evantodo{Talk a bit about the heuristics used for building the tree from the
+DAG}
 

### 2015-08-16 linvis.tex
Design: Modules
Talking about the modules tab of the viewer.

@@
 	\centering
+	\includegraphics[width=0.47\textwidth]{figures/modules.png}
+	\caption{Commit Modules}
+	\label{fig:modules}
+\end{figure}
+
+The modules tab shows the modules that are contained within the commit. Modules
+are not natively recognized by git, and are not going to be present in all
+repositories. In the Linux repository, authors put the module they are working
+on in the one-line preview of the log-message.  We heuristically extract the
+module by taking all text in the log preview until the first colon. Modules are
+logical partitions of the information in the kernel. Depending on where the
+author was working, modules can be general, such as ``bluetooth'' and
+``wireless'', or can be quite specific for individual hardware, such as
+``ath9k\_hw'' and ``wl1251''. Like files, we are able to show what modules are
+edited in merges and in commits. The image in figure \ref{fig:modules} shows
+the modules for a commit. A commit will only ever have one module that it is
+associated with.
+
+\begin{figure}[h]
+	\centering
 	\includegraphics[width=0.47\textwidth]{figures/tree_view.png}
 	\caption{Commit Tree View}
-	\label{fig:files}
+	\label{fig:tree}
 \end{figure}
 
-The third tab shows the tree view. The tree view is what makes this tool unique
-to the other tools. It provides a full view of the entire merge tree. It
-originally centres at the current commit or merge, but allows a user to inspect
-the other commits and merges surrounding it. A user may then navigate to any of
-the commits by simply double-clicking on it.
+The third tab shows the tree view (\ref{fig:tree}). The tree view is what makes
+this tool unique to the other tools. It provides a full view of the entire
+merge tree. It originally centres at the current commit or merge, but allows a
+user to inspect the other commits and merges surrounding it. A user may then
+navigate to any of the commits by simply double-clicking on it.
 
@@
 	\caption{Tree zoomed out}
-	\label{fig:files}
+	\label{fig:zoomed_tree}
 \end{figure}
 
-Some of the merge trees are very large, containing thousands of commits and
-merges. Clicking on subtrees will cause them to minimize, only showing parts of
-the tree that are of interest to the user. Furthermore, the user can zoom the
-tree by scrolling the mouse wheel to see more of the topology of the tree, or
-to see the details of a given part of the tree.
+Some of the merge trees are very large (\ref{fig:zoomed_tree}), containing
+thousands of commits and merges. Clicking on subtrees will cause them to
+minimize, only showing parts of the tree that are of interest to the user.
+Furthermore, the user can zoom the tree by scrolling the mouse wheel to see
+more of the topology of the tree, or to see the details of a given part of the
+tree.
 

### 2015-08-17 linvis.tex
Discussion: Future Work:
We discuss the differences in our tool versus the other tool, providing a
side-by-side view of the visualization models used by our tool and gitk. We
show what information can be gathered quickly and effectively through the use
of our tool that cannot be gathered as effectively with gitk.

There is still a lot of work to be done in the future work section. We need to
talk about how our tool does not provide the diffs, though we have that
information and could very easily display it. Furthermore, we could use the
diff to determine what lines were changed, which lines we

@@
 \else
-	\documentclass[conferece, final]{IEEEtran}
+        \documentclass[conference]{IEEEtran}
+	\def\baselinestretch{1}
+	\setlength{\marginparwidth}{2cm}
+
+	% \documentclass[conferece, final]{IEEEtran}
 \fi
@@
 	of commits, maintenance of older versions of the kernel becomes nearly
-	impossible. For security, performance, and changing hardware,
+	impossible. For security, performance, and changing hardware needs,
 	maintainers must understand the changes made to current versions of the
-	kernel, and show these changes fit into older versions in order to make
-	the necessary merges and modifications to the older versions. Current
-	tools provide information about repositories through the directed
-	acyclic graph (DAG) which is helpful for smaller projects, but with the
-	scale and number of branches in the kernel, the DAG becomes
-	overwhelming very quickly.
+	kernel, and understand how these changes fit into older version in
+        order to make the necessary merges and modifications to the older
+        versions. Current tools provide information about repositories through
+        the directed acyclic graph (DAG) which is helpful for smaller projects,
+        but with the scale and number of branches in the kernel, the DAG
+        becomes overwhelming very quickly.
 
@@
 
+\section{Discussion}
+
+Our tool provides mechanisms and visualizations that other tools are unable to
+produce. Continuous mining of the repository provides an additional dimension
+to the dataset that would otherwise be lost, enabling us to reliably build
+trees from the information we have.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
+	\caption{Merge Dag View}
+	\label{fig:dag_view}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.47\textwidth]{figures/042dd_tree.png}
+	\caption{Merge Tree View}
+	\label{fig:tree_view}
+\end{figure}
+
+Our goal was to build a tool to enable maintainers to effectively navigate and
+pull explanation of the changes performed to the kernel over a release.
+Achieving this goal includes removing information that does not pertain to given
+users. In figure \ref{fig:dag_view} and figure \ref{fig:tree_view}, we can
+visually compare the results returned from gitk and our tool for the top-level
+merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6'' from kernel version 3.10.
+The relevant information in both of these figures is identical, but the
+presentation durastically changes our understanding of what we are seeing.
+
+This merge is relatively small, containing only four merges, three of which
+only contain a single commit, and the last containing two commits. With our
+tool, we are able to immediately see what commits and merges pertain to our
+section of the kernel. The DAG view of this provides almost no explanation,
+futhermore, users must work to determine where this merge ends and the next one
+begins.
+
+Our tool further enhances our understanding by immediately giving us the files
+that were edited and the modules. We are able to determine that three files,
+``palamas-regulator.c'' had two lines added and two lines removed,
+``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had
+5 lines added and 2 lines removed in the merge. Finally, we are able to
+determine that only the module ``regulator'' was modified in this merge, and
+there were 5 commits in the merge to this module.
+
+\subsection{Performance Evaluation}
+
+Re-structuring the kernel into a tree changes the runtime for the structures.
+Originally, we were performing all of our tree-navigation using recursive SQL
+queries. Recursive queries have terrible performance, with the average response
+time from the database being around 3 seconds for a single operation. We found
+that it was far more time-efficient to load an entire tree directly into
+memory, then iterate over the objects returned to build the tree structures,
+than it was to have the dbms perform the tree traversals, building the specific
+trees for a commit. At this time, we have re-implemented the breadcrumbs, the
+files, and the main tree to enhance the response times. We will discuss some We
+will discuss the changes made.
+
+The breadcrumbs are generated at response
+time, so they must be generated quickly. We began with a single recursive SQL
+query, but was unable get the results in the required time, causing all result
+pages to render slowly. We replaced the recursive query with a single query
+that would grab the entire merge tree, rooted at the top-level merge.
+
+The items returned from the database contained the commit id, the commit id of
+the merge with the Linux kernel, and the next merge in the path to the
+top-level merge.
+
+We placed each item in the results into a dictionary using the commit id as the
+key. Using the following method, we build the dictionary into a list of commit
+IDs for the breadcrumbs.
+
+\begin{verbatim}
+current = cid
+breadcrumbs = [cid]
+while current is not top-level merge:
+        breadcrumbs.append(current.next)
+return reverse(breacrumbs)
+\end{verbatim}
+
+This maintains the O(n) runtime for generating the breadcrumbs, while using an
+iterative method for traversing the tree.
+
+We use similar methods to generate the file information and the tree. Both are
+generated asynchronously, allowing us to use spinners to help improve the user
+experience. We are using AJAX requests to generate the asynchronous requests,
+are are having very good results.
+
+\subsection{Extendability}
+Due to how the asynchronous requests are performed, it is possible to query the
+server for the tree and file information directly, allowing other users to
+build their own front-ends for working with our tree-based system.
+
+The tree and file information is accessible through
+http://li.turingmachine.org/data/tree/JSON/<commit id> and
+http://li.turingmachine.org/data/files/JSON/<commit id>, respectively.
+The tree is generated as many nested javascript objects. The tree response does
+not contain any notion of the current commit and will only return the entire
+merge tree.
+
+The tree responses are generated as a single object. This object is the root
+node of the tree, and may contain many objects with the same fields in the
+``children'' field. The full response format for the tree looks like:
+
+\begin{itemize}
+        \item ``cid'': string, commit id
+        \item ``name'': string, One-line log preview
+        \item ``author'': string, Author's name
+        \item ``mlinus'': string, commit id of top-level merge
+        \item ``mnext'': string, commit id of the next merge on path to
+                top-level
+        \item ``children'': list, list of items with the same fields. Null if
+                no children
+\end{itemize}
+
+The file responses contain only the files that the current merge or commit
+works with. The response is a single object in the form of a tree. This object
+is the root of the tree, and represents the current merge or commit. If the
+current position in the tree is an inner node, the response will contain the
+children in the ``children'' field, otherwise ``children'' will be null. The
+full response format for the file response looks like:
+
+\begin{itemize}
+        \item ``cid'': string, commit id
+        \item ``files'': list, list of tulples in the format
+                \begin{enumerate}
+                        \item Filename: string
+                        \item Lines added: unsigned integer
+                        \item Lines removed: unsigned integer
+                \end{enumerate}
+        \item ``mnext'': string, the commit id of the next merge toward the
+                top-level merge
+        \item ``children'': list, List of all child objects with the same
+                fields. Null if no children.
+\end{itemize}
+
 \section{Future Work}
-We have limited search functionality regarding actual files involved in
-commits. If a user knows which file they are working with and what has changed
-in a given file, they are unable to look specifically for commits pertaining to
-that file. Adding a search filter for searching by file would allow maintainers
-to perform this search.
 
-Currently, there are still various features that have bad performance. More
-work is required to get those features responding more quickly, or providing an
-asynchronous mechanism and loading bars to improve user experience.
+This project was implemented over the course of 4 months, interleaved with
+other course projects and assignments. There are still many areas that could
+use improvements before the full potential of our model can be reached and
+understood with our tool. Here, we outline various areas that still need more
+attention.
 
+\subsection{Files}
+We have limited functionality surrounding files, for both search and presented
+information.
+
+\evantodo{User cases integrated into files}
+
+% There is currently no functionality surrounding searching for a given file.
+% Users who know what file they are working with
+%
+% We have limited search functionality regarding actual files involved in
+% commits. If a user knows which file they are working with and what has changed
+% in a given file, they are unable to look specifically for commits pertaining to
+% that file. Adding a search filter for searching by file would allow maintainers
+% to perform this search.
+
+\subsection{Testing}
 At this time, we have no evidence that our tool is able to improve the
@@
 
+We could perform user-testing, or have some example cases where the tool is
+used and the users provide feedback on how it improves or hurts their work
+flow.
+
+Currently, there are still various features that have bad performance. More
+work is required to get those features responding more quickly, or providing an
+asynchronous mechanism and loading bars to improve user experience.
+
 \end{document}

### 2015-08-17 linvis.tex
Design: Use-cases
Adding use-cases
Flipping design and implementation details sections.

@@
 In this paper, we present the design decisions behind a web-based tool built
-around a tree-based model instead of the DAG. We demonstrate the advantages of
-this model over the use of the DAG. Our tool provides information about the
-location of a given commit in the respective merge tree, the files edited, the
-modules edited, and the commit message in a given commit or merge. The tool
-allows users to apply various filters, including the release, a keyword or
-phrase from the log preview, the name of the author, or the commit id. The user
-can request all the top-level merges containing a commit or merge that matches
-the query, or all commits and merges that match the query.
+around a tree-based model instead of the DAG.
+
+We demonstrate the advantages of this model over the DAG two main use-cases.
+Our tool provides information about the location of a given commit in the
+respective merge tree, the files edited, the modules edited, and the commit
+message in a given commit or merge. The tool allows users to apply various
+filters, including the release, a keyword or phrase from the log preview, the
+name of the author, or the commit id. The user can request all the top-level
+merges containing a commit or merge that matches the query, or all commits and
+merges that match the query.
 
@@
 
-Core Model:
-	Merges are inner nodes
-	Commits are leaf nodes
-
-\section{Implementation Details}
-
-We built a we-based tool to provide information about the kernel.
-We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
-as the framework for generating the dynamic content.
-
-Nginx is a high-performance, scalable HTTP and mail server. Started in 2002,
-the project is much younger than the Apache project, and designed with massive
-scalability and multi-processing in mind. Instead of using request-driven
-threading, Nginx uses an event-driven architecture, making it more scalable,
-proving itself as being the web server behind Netflix, Hulu, GitHub, Tumblr,
-and many other popular sites.
-
-Postgresql is our dbms of choice, providing fast and reliable results. There
-are multiple reasons behind choosing postgresql over another dbms. The main
-reason being that the dataset provided from \evantodo{paper} was generated from
-a postgresql database, allowing us to import the data directly without any
-further conversions. The secondary reason is that we are more familiar with
-postgresql than the other systems available.
-
-We chose the python micro-framework Flask for generating the dynamic content.
-Working with flask is simple and easy.
-
-The database itself is broken down into 5 main tables; commits, filesmod, logs,
-pathtomerge, and releases. Commits contains most of the metadata for each
-commit, it contains the author, the authored date, and committer, the commit
-date, a boolean saying what is a merge, and the diff. Filesmod contains the
-information on what files were modified for the commits. Since a file may be
-modified by multiple commits, and a commit may be modifying multiple files, we
-can do one of two things for the primary key. We could use both the filename
-and the commit id as the primary key, but we just use an index. The filesmod
-table contains the filename, the commit id, the number of lines added, and the
-number of lines removed. Trivially, it contains the index, but that isn't
-necessary for anything beyond the primary key. Logs contains the information
-for each commit log, this is the one-liner preview message and the full
-message. Pathtomerge is the important table. Pathtomerge contains the commit
-id, the number of merges between the commit and the top-level merge with the
-kernel, the next merge in the merge path up to the top-level merge with the
-kernel, the top-level merge into the kernel, and when it get merged into the
-kernel. The information in this table must be generated periodically by the
-server because we are unable to generate the tree structure correctly straight
-from the information stored by git. Pathtomerge does not contain any top-level
-merges, as these will never have a next merge in the tree, since we are only
-storing the parents and not the children, storing the top-level merge is
-redundant and unnecessary. The releases table stores the release information.
-It contains the version name, whether the version is a candidate release or a
-real release, the previous version, the previous non-candidate release version
-the previous version commit id, the previous non-candidate release version
-commit id, and the commit id of this version. The version commit id represents
-the last top-level merge into the version of the kernel. Using this and the
-commits table, we are able to determine the release date of any version of the
-kernel. Then using the previous release date, we can find the range of dates
-where valid top-level merges are found for that version.
-
-\evantodo{Talk a bit about the heuristics used for building the tree from the
-DAG}
-
 \section{Design}
@@
 results down into only the information that is relevant to what the user is
-looking for.
+looking for. We designed the tool with two use-cases in mind, though a user may
+switch roles many times as they work.
+
+\textit{Use-case 1: top-to-bottom approach}\\
+These are users that are maintaining a section of the kernel and would like to
+pick a merge from the future merges and merge it directly into their current
+repository. This is useful for minimizing required implementation on the side
+of the maintainer. We must provide them with methods of navigating from a more
+general set of merges toward a specific set of merges and commits.
+
+\textit{Use-case 2: bottom-to-top approach}\\
+These are users that know what merge or commit they are working with and would
+like to see what else is happening around their current commit. This is useful
+to see how changes to a given file will break other changes made by other
+commits. This is primarily for maintainers that have to perform some
+implementation work in order to correctly merge the patches from the future.
+These users may also know what module(s) they are dealing with, and are looking
+for modules that are relevant to the area they are working on. We must provide
+them methods of navigating from a specific commit to a more general merge,
+allowing them to see a bigger picture.
 
@@
 the first item is the top-level merge into the kernel.
+
+\section{Implementation Details}
+
+We built a we-based tool to provide information about the kernel.
+We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
+as the framework for generating the dynamic content.
+
+Nginx is a high-performance, scalable HTTP and mail server. Started in 2002,
+the project is much younger than the Apache project, and designed with massive
+scalability and multi-processing in mind. Instead of using request-driven
+threading, Nginx uses an event-driven architecture, making it more scalable,
+proving itself as being the web server behind Netflix, Hulu, GitHub, Tumblr,
+and many other popular sites.
+
+Postgresql is our dbms of choice, providing fast and reliable results. There
+are multiple reasons behind choosing postgresql over another dbms. The main
+reason being that the dataset provided from \evantodo{paper} was generated from
+a postgresql database, allowing us to import the data directly without any
+further conversions. The secondary reason is that we are more familiar with
+postgresql than the other systems available.
+
+We chose the python micro-framework Flask for generating the dynamic content.
+Working with flask is simple and easy.
+
+The database itself is broken down into 5 main tables; commits, filesmod, logs,
+pathtomerge, and releases. Commits contains most of the metadata for each
+commit, it contains the author, the authored date, and committer, the commit
+date, a boolean saying what is a merge, and the diff. Filesmod contains the
+information on what files were modified for the commits. Since a file may be
+modified by multiple commits, and a commit may be modifying multiple files, we
+can do one of two things for the primary key. We could use both the filename
+and the commit id as the primary key, but we just use an index. The filesmod
+table contains the filename, the commit id, the number of lines added, and the
+number of lines removed. Trivially, it contains the index, but that isn't
+necessary for anything beyond the primary key. Logs contains the information
+for each commit log, this is the one-liner preview message and the full
+message. Pathtomerge is the important table. Pathtomerge contains the commit
+id, the number of merges between the commit and the top-level merge with the
+kernel, the next merge in the merge path up to the top-level merge with the
+kernel, the top-level merge into the kernel, and when it get merged into the
+kernel. The information in this table must be generated periodically by the
+server because we are unable to generate the tree structure correctly straight
+from the information stored by git. Pathtomerge does not contain any top-level
+merges, as these will never have a next merge in the tree, since we are only
+storing the parents and not the children, storing the top-level merge is
+redundant and unnecessary. The releases table stores the release information.
+It contains the version name, whether the version is a candidate release or a
+real release, the previous version, the previous non-candidate release version
+the previous version commit id, the previous non-candidate release version
+commit id, and the commit id of this version. The version commit id represents
+the last top-level merge into the version of the kernel. Using this and the
+commits table, we are able to determine the release date of any version of the
+kernel. Then using the previous release date, we can find the range of dates
+where valid top-level merges are found for that version.
+
+\evantodo{Talk a bit about the heuristics used for building the tree from the
+DAG}
 

### 2015-08-17 linvis.tex
Edits: citations:
Re-wording a bunch of things and adding citations to the paper.

@@
 \usepackage{caption}
+\usepackage[nospace]{cite}
 
@@
 \begin{abstract}
-
 	With an average of over 900 top-level merges into the Linux kernel per
@@
 	In this paper, we present a tool that uses the dataset collected by
-	German et al. to build a tool that uses a vine model instead of the
-	DAG. This tool is designed to cater to the needs of users looking for a
-	top-down approach, or a bottom-up approach to navigating the commit
-	information of the kernel.
+	German et al. to build a tool that uses a tree-based model instead of
+        the DAG. This tool is designed to cater to the needs of users looking
+        for a top-down approach, or a bottom-up approach to navigating the
+        commit information of the kernel.
 \end{abstract}
@@
 this information is useful for showing the relation between various branches.
-With large modular projects, like the Linux Kernel, the DAG becomes a tangled
-mess (\ref{fig:gitk}) of commits, merges, and the links between them, making it
-difficult to derive an explanation of the changes.  Github provides a DAG view
-for most projects, but is unable to display the DAG of projects as large as the
-Linux Kernel.
+The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges, and the
+links between them with large modular projects, like the Linux kernel. This
+makes it difficult to derive an explanation of the changes made. In other
+cases, the kernel is simply too large for the system to generate a
+visualization for. Github provides a DAG view for most projects, but is unable
+to display the visualization for projects as large as the Linux kernel.
 
@@
 commit. A commit may be merged into the project at any time, regardless of when
-the commit was authored. Because of this, performing date-range queries does
-not provide meaningful results. Performing a date-range query using other tools
-over the span of a release will contain results that are for previous versions
-of the kernel, and may not contain all of the results for that kernel if
-patches were merged after the release date.
+the commit was authored, making date-range queries meaningless. Performing a
+date-range query using other tools over the span of a release will contain
+results that are for previous versions of the kernel, and may not contain all
+of the results for that kernel if patches were merged after the release date.
 
 German et al. recorded the metadata of every commit and merge in the Linux git
-repository from 2012 to 2015, including data that is not stored by git. We are
-able to use this information to generate sets of trees instead of a DAG. The
-root of each tree is the top-level merge performed by Linus, which merges a
-given commit into the kernel.
+repository starting in 2011 \cite{German}, including chronological data that is
+not stored by git. We are able to use this information to generate sets of
+trees instead of a DAG. The root of each tree is the top-level merge performed
+by Linus, which merges branches or sets of commits into the kernel.
 
-In this paper, we present the design decisions behind a web-based tool built
-around a tree-based model instead of the DAG.
+In this paper, we present the design decisions behind a web-based
+tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
+model instead of the DAG.  We demonstrate the advantages of this model over the
+DAG two main use-cases.  Our tool provides information about the location of a
+given commit in the respective merge tree, the files edited, the modules
+edited, and the commit message in a given commit or merge.  The tool allows
+users to apply various filters, including the release, a keyword or phrase from
+the log preview, the name of the author, or the commit id. The user can request
+all the top-level merges containing a commit or merge that matches the query,
+or all commits and merges that match the query.
 
-We demonstrate the advantages of this model over the DAG two main use-cases.
-Our tool provides information about the location of a given commit in the
-respective merge tree, the files edited, the modules edited, and the commit
-message in a given commit or merge. The tool allows users to apply various
-filters, including the release, a keyword or phrase from the log preview, the
-name of the author, or the commit id. The user can request all the top-level
-merges containing a commit or merge that matches the query, or all commits and
-merges that match the query.
-
-Our core model provides many advantages to other tools. Breaking the data by
-merge tree removes the commits and merges of other components of the kernel
-from the visualization, providing a clearer picture of what is happening in a
-given merge. This model provides additional support to edited file information.
+Our core model provides advantages to the DAG model used by other tools.
+Breaking the commits and merges by merge tree removes the commits and merges of
+other components of the kernel from the visualization, providing a clearer
+picture of what is happening in a given merge. This model provides additional
+support to edited file information. Furthermore, we are able to understand the
+changes being made in the order they are brought into the kernel.
 
@@
 date-range queries, providing access to all commits that have an effect on a
-given release, even if the commit comes before the start date, or the release
-date.
+given release, even if the commit comes before the start date, or after the
+release date.
 
 \section{Related Work}
+
+To our knowledge, no repository visualization systems using a tree-based model
+are being used. This may be due to the fact that more information than what
+is stored by the normal tools is required to create the trees, so servers must
+sit and watch as the repository changes over time. Many of the visualization
+tools are designed for a general-purpose approach, where any repository is
+directly usable, without additional information. Polling for this additional
+information makes the application specific to a single repository, or project.
 
@@
 \subsection{Treemap visualizer}
-University of Maryland presented a treemap design (\ref{fig:treemap}) for
+University of Maryland presented a treemap visualization
+tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
+pictured in figure \ref{fig:treemap}, for
 displaying all the file structure of a given version. Treemaps are good for
@@
 
-\begin{figure}[h!]
+\begin{figure}[h]
 	\centering
@@
 \subsection{List viewer}
-Moghaddam designed a web-based visualization tool displaying the chain of
-merges into the kernel. There is no information on what commits are within a
-given merge, or what files were edited. The visualization tool provides
-information about who authored the commit and when it was authored.
-\evantodo{Link to tool}
+Moghaddam designed a web-based visualization
+tool\footnote{\url{http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html}}
+displaying the chain of merges into the kernel. There is no information on what
+commits are within a given merge, or what files were edited. The visualization
+tool provides information about who authored the commit and when it was
+authored. We can see the visualization tool in figure \ref{fig:listviewer}
+displaying a part of the kernel commit information, from commit ``26b23ac''
+according to the tool.
+
 %% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
@@
 \end{figure}
-
-\section{Core Model}
-Each merge may contain zero or more children of other merges or commits.
-Commits have no children, and therefore are always leaf nodes, but commits
-contain information about what files are edited, how many lines are edited, and
-other important information.
 
@@
 
+\subsection{Core Model}
+Each top-level merge into the kernel is the root-node to a tree. Each merge
+may contain zero or more children of other merges or commits. Commits have no
+children, and therefore are always leaf nodes, but commits contain information
+about what files are edited and how many lines are edited, along with what
+modules are involved in a merge.
+
 \subsection{Command-line Shell}
@@
 
-\evantodo{Put this in discussion section}
-The main issue with this design is that a user will always move from the
-top-down without any additional filtering. Furthermore, it presents the user
-with the commit hashes, which provide no explanation of what the commit or
+There are two issues with this design. One issue with this design is that a
+user will always move from the top-down without any additional filtering. This
+only serves one of our use-cases. The second issue being that it presents the
+user with the commit hashes, providing no explanation of what the commit or
 merge may contain.
@@
 There are multiple reasons for building the system into a web-based tool.
-Putting the tool onto the web enables users to use the system without having to
-install anything, making the system platform independent. A web-based tool can
-utilize more interactive means of navigation, and provide a better explanation
-of what each commit or merge may contain.
+Putting the tool on the web enables users to use the system without having to
+install any software, making the system more accessible and platform
+independent.A web-based tool can utilize more interactive means of navigation,
+and provide a better explanation of what each commit or merge may contain than
+the command-line variant.
 
-The web-based tool provides many mechanisms to enable users to better navigate
-and understand changes in the kernel. These mechanisms include searching,
-commit messages, files edited, modules edited, and two types of tree viewers.
+Our implementation of the web-based tool provides many mechanisms to enable
+users to better navigate and understand changes in the kernel. These mechanisms
+include searching, commit messages, files edited, modules edited, and two types
+of tree viewers.
 
@@
         breadcrumbs.append(current.next)
+	current = current.next
 return reverse(breacrumbs)
@@
 \section{Future Work}
-
 This project was implemented over the course of 4 months, interleaved with
@@
 
-\evantodo{User cases integrated into files}
+There is currently no functionality surrounding searching by filename. Use-case
+2 may know what file they are editing and try to determine how it may work with
+the other commits and merges in the module. It is possible that a third
+use-case may arise, where the user wants to determine all commits that effect a
+given file. In both cases, a bottom-to-top approach is used. With this
+functionality, implementing the top-to-bottom search feature becomes trivial.
 
-% There is currently no functionality surrounding searching for a given file.
-% Users who know what file they are working with
-%
-% We have limited search functionality regarding actual files involved in
-% commits. If a user knows which file they are working with and what has changed
-% in a given file, they are unable to look specifically for commits pertaining to
-% that file. Adding a search filter for searching by file would allow maintainers
-% to perform this search.
+There is limited functionality to the presentation of the file information. At
+a minimum, the patches for the commits can be displayed. The patches can then
+be used for many more additional features. The patches can be used to piece
+together parts of the file to generate a full patch at a given merge rather
+than displaying each patch individually. The patches can also be used for
+determining what kinds of changes were made, if the lines are being added,
+removed, or replaced.
 
@@
 
-Currently, there are still various features that have bad performance. More
-work is required to get those features responding more quickly, or providing an
-asynchronous mechanism and loading bars to improve user experience.
+\subsection{Performance}
+The modules and list tree view still use the old method of gathering the
+information for the respective view type. For single commits or very small
+merges, getting everything directly from the database and performing everything
+server-side performs better than using the asynchronous requests, as there
+isn't the overhead of multiple round trips; however, as the size of the merges
+grow, the performance becomes far worse than the asynchronous system and in
+some cases will time out. The asynchronous approach adds some overhead,
+profiled at roughly 500ms, regardless of the response size. The asynchronous
+approach has many benefits, over building the response on the server, including
+the ability to add spinners, which is an improvement to user-experience over
+looking at a blank page. Further work can be done to improve the results beyond
+the response time of 500ms using improved queries and more efficient methods of
+building the tree.
 
+\section{Conclusion}
+
+\evantodo{DO IT! JUST DO IT!}
+
+\bibliographystyle{plain}
+\bibliography{citations}
+
+% \balance
 \end{document}

### 2015-08-17 linvis.tex
Conclusion: Finishing up conclusion
Okay, I think we are done adding content.
On to editing...

@@
 \section{Conclusion}
+Our tool shows promising results, providing a clearer picture of the relation
+between the merges and commits for sections of the kernel in relation to the
+DAG-based models.  Not only are we able to provide a clearer visualization, we
+are able to provide information about the files edited by all of the commits
+contained within a merge.
 
-\evantodo{DO IT! JUST DO IT!}
+While we are able to gather a better explanation of the changes happening at
+each commit and merge, the tree-model requires more information than what is
+provided by git, requiring a central server to monitor changes to the project.
+This breaks the design goals of decentralized version control systems and makes
+the tool specific to the project.
 
@@
 
-% \balance
+\balance
+
 \end{document}

### 2015-08-17 linvis.tex
Editing: Introduction and Related Work

@@
 commits with tools that are unable to filter and effectively visualize projects
-that are of the scale of the kernel. Tools like Gitk provide users with the
-full view of the DAG, showing all merges and commits. For smaller projects,
-this information is useful for showing the relation between various branches.
-The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges, and the
-links between them with large modular projects, like the Linux kernel. This
-makes it difficult to derive an explanation of the changes made. In other
+that are the scale of the kernel. Tools like Gitk provide users with the
+full view of the DAG, showing all of the merges and commits. For smaller
+projects, this information is useful, showing the relation between various
+branches. The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges,
+and the links between them with large modular projects, like the Linux kernel,
+making it difficult to derive an explanation of the changes made. In other
 cases, the kernel is simply too large for the system to generate a
@@
 the commit was authored, making date-range queries meaningless. Performing a
-date-range query using other tools over the span of a release will contain
-results that are for previous versions of the kernel, and may not contain all
+date-range query using these tools over the span of a release will contain
+results that are for previous versions of the kernel or may not contain all
 of the results for that kernel if patches were merged after the release date.
@@
 not stored by git. We are able to use this information to generate sets of
-trees instead of a DAG. The root of each tree is the top-level merge performed
-by Linus, which merges branches or sets of commits into the kernel.
+trees instead of a DAG. Each tree is rooted by the top-level merge, performed
+by Linus, which merges the branches or a set of commits into the kernel.
 
-In this paper, we present the design decisions behind a web-based
+In this paper, we present our design decisions behind a web-based
 tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
-model instead of the DAG.  We demonstrate the advantages of this model over the
-DAG two main use-cases.  Our tool provides information about the location of a
+model instead of a DAG. Our tool provides information about the location of a
 given commit in the respective merge tree, the files edited, the modules
-edited, and the commit message in a given commit or merge.  The tool allows
-users to apply various filters, including the release, a keyword or phrase from
-the log preview, the name of the author, or the commit id. The user can request
-all the top-level merges containing a commit or merge that matches the query,
-or all commits and merges that match the query.
+edited, and the commit message for a given commit or merge. The tool allows
+users to apply various filters, including the release version, along with a
+keyword or phrase from the log preview, the name of the author, or the commit
+id. The user can request all the top-level merges containing a commit or merge
+that matches the query, or all commits and merges that match the query.
 
 Our core model provides advantages to the DAG model used by other tools.
-Breaking the commits and merges by merge tree removes the commits and merges of
+Breaking the commits and merges by merge tree removes commits and merges of
 other components of the kernel from the visualization, providing a clearer
-picture of what is happening in a given merge. This model provides additional
-support to edited file information. Furthermore, we are able to understand the
-changes being made in the order they are brought into the kernel.
+picture of what is happening in a given merge and providing additional
+support for edited file information. Furthermore, we are able to understand the
+changes being made, in the order they are brought into the kernel.
 
@@
 date-range queries, providing access to all commits that have an effect on a
-given release, even if the commit comes before the start date, or after the
+given release, even if the commit comes before the start date or after the
 release date.
@@
 \section{Related Work}
-
-To our knowledge, no repository visualization systems using a tree-based model
-are being used. This may be due to the fact that more information than what
-is stored by the normal tools is required to create the trees, so servers must
-sit and watch as the repository changes over time. Many of the visualization
-tools are designed for a general-purpose approach, where any repository is
-directly usable, without additional information. Polling for this additional
-information makes the application specific to a single repository, or project.
+To our knowledge, we know of no repository visualization systems using a
+tree-based model. This may be due to the fact that more information is required
+to generate the trees than what is stored by git. A server must sit and monitor
+the repository as it changes over time. Many of the visualization tools are
+designed for a general-purpose approach, where any repository is directly
+usable without additional information. Polling for this additional information
+makes the application specific to a single repository or project.
 
@@
 Gitk, Gitg, and other similar git repository visualizing tools use the DAG as
-the central organization. For smaller projects, this model is sufficient for
-providing meaningful information, but in larger projects like the Linux kernel,
-the branches become too tangled and too deep to derive any meaning from the
-DAG. The DAG model allows users to go to the next commit in the chain, go to
-the previous commit in the chain, or jump to the parent merge. This suggest
+the central organization model. For smaller projects, this model is sufficient
+for providing meaningful information, but in larger projects like the Linux
+kernel, the branches become too tangled and too deep to derive any meaning from
+the DAG. The DAG model allows users to go to the next commit in the chain, go
+to the previous commit in the chain, or jump to the parent merge. This suggest
 that Gitk and similar tools are designed for users working with a bottom-to-top
@@
 are involved with a given merge. Gitk only provides file information on
-commits, as it is unable to quickly find all files edited by a merge. It is up
-to the user to remember what files were edited in each of the commits involved
-in a given merge. In some of the larger merges, some containing over 1500
-edited files, this task become nearly impossible.
+commits, as it is unable to quickly find all files edited by the commits within
+a merge. It is up to the user to remember what files were edited in each of the
+commits involved in a given merge. In some of the larger merges, some
+containing over 1500 edited files, this task become nearly impossible.
 % cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
@@
 pictured in figure \ref{fig:treemap}, for
-displaying all the file structure of a given version. Treemaps are good for
-displaying the topology of shallow, wide trees, making it a good candidate for
-displaying file systems. Treemaps are limited in what information can be
+displaying the file structure of a given version. Treemaps are good for
+visualizing the topology of shallow, wide trees, making it a good candidate for
+showing file systems. Treemaps are limited in what information can be
 displayed. The information within the treemap does not provide clear insight on
-what changes were made to the files in the given release. The treemap
-visualization only provides information on where the file is located within the
-filesystem. File systems and the merge tree of the kernel share many
-similarities, in both cases, the representative trees are wide and shallow,
-suggesting that a treemap could potentially work to visualize the commit and
-merge structure of the kernel.
+what changes were made to the files in the given release, only providing
+information on where the file is located within the file system.
+File systems and the merge tree of the kernel share many similarities, in both
+cases, they represent trees that are wide and shallow, suggesting that a
+treemap could potentially work to visualize the commit and merge structure of
+the kernel.
 
-The treemap is unable to take into account multiple merges at the same level
-with the same name. There is an additional dimension, in that the commits and
-merges are ordered in time. The treemap design is unable to show the ordering
-of the merges and commits.
+The treemap is unable to take into account multiple commits behind the same
+merge with the same name. Commits and merges have an ordering in which they
+were added to the kernel, which cannot be represented by a treemap, but is
+important and can be used to assist in building the tree.
 
@@
 Finally, the treemap does not provide any additional information about the
-contents of each cell, only that the cell exists. The user is unable to see
-their current position within the map, or if we assume that their current
-position is the larges surrounding box, they can no-longer see how their
-current position interacts with the rest of the system.
+contents of each cell, only that the cell exists and where it exists in the
+tree. The user is unable to see their current position within the map, or if we
+assume that their current position is the largest surrounding box, they can
+no longer see how their current position interacts with the rest of the
+system.
 %% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
@@
 \section{Design}
-
-The goal is to build a tool that make navigation of the kernel commit
+The goal is to build a tool that makes navigation of the kernel commit
 information simpler, while providing an explanation of the changes occurring at
@@
 \subsection{Command-line Shell}
-
 Our initial intuition was to work with the commits and merges as trees, like in
@@
 \subsection{Web-based Tool}
-
 There are multiple reasons for building the system into a web-based tool.
@@
 \section{Implementation Details}
-
 We built a we-based tool to provide information about the kernel.

### 2015-08-17 linvis.tex
Editing: Design
Edited the design section

@@
 looking for. We designed the tool with two use-cases in mind, though a user may
-switch roles many times as they work.
+switch roles as they work.
 
@@
 These are users that are maintaining a section of the kernel and would like to
-pick a merge from the future merges and merge it directly into their current
-repository. This is useful for minimizing required implementation on the side
-of the maintainer. We must provide them with methods of navigating from a more
-general set of merges toward a specific set of merges and commits.
+pick a merge and merge it directly into their current repository. This is
+useful for minimizing required implementation on the side of the maintainer,
+allowing them to see what changes are made higher up in the kernel and bundle
+multiple changes together. We must provide them with methods of navigating from
+a more general set of merges toward a specific set of merges and commits.
 
@@
 These are users that know what merge or commit they are working with and would
-like to see what else is happening around their current commit. This is useful
-to see how changes to a given file will break other changes made by other
-commits. This is primarily for maintainers that have to perform some
-implementation work in order to correctly merge the patches from the future.
-These users may also know what module(s) they are dealing with, and are looking
-for modules that are relevant to the area they are working on. We must provide
-them methods of navigating from a specific commit to a more general merge,
-allowing them to see a bigger picture.
+like to see what else is happening around their commit. This is useful to see
+how changes to a given file will break other changes made in other commits.
+This is primarily for maintainers that have to perform some implementation work
+in order to correctly merge the patches from the future. These users may also
+know what module(s) they are dealing with, and are looking for modules that are
+relevant to the area they are working on. We must provide them methods of
+navigating from a specific commit to a more general merge, allowing them to see
+a bigger picture.
 
@@
 Each top-level merge into the kernel is the root-node to a tree. Each merge
-may contain zero or more children of other merges or commits. Commits have no
-children, and therefore are always leaf nodes, but commits contain information
-about what files are edited and how many lines are edited, along with what
-modules are involved in a merge.
+may contain zero or more children made of other merges or commits. Commits have
+no children, and therefore are always leaf nodes.  The commits contain the
+information about the files that are edited along with how many lines are added
+and removed from each file. Commits also store the module of the kernel that
+they are working with.
 
 \subsection{Command-line Shell}
-Our initial intuition was to work with the commits and merges as trees, like in
-a file structure. There are many similarities in the structure of a file system
-and the structure of the commit information within the kernel. The primary
-similarity is that they are both trees, and both trees are generally fairly
-shallow and very wide. File systems have been around for quite some time, and
-there are many tools for visualizing and working withe the file structure. The
-first tool was the command-line. We built a small command-line shell
-(\ref{fig:shell}) that allowed navigation through the commit and merges instead
-of files and directory. The shell design permitted a user to move through the
-merges like they would through directories. They could use the cat command to
-see the preview of the log, or the more command to see the preview and what
-files were edited in the commit and how many lines were added and removed from
-each file.
+Our initial intuition was to work with the commits and merges as trees. We use
+the merges as directories, and commits as files, like in a file structure.
+There are many similarities between a file system and the structure of the
+commit information. The primary similarity is that they are both trees, and
+both trees are generally fairly shallow and wide. File systems have been
+around for quite some time, and there are many tools for visualizing and
+working withe the file structure. The first tool was the command-line. We built
+a small command-line shell (\ref{fig:shell}) that allows navigation through
+the commit and merges instead of files and directory. The shell design
+permitted a user to move through the merges like they would through
+directories with the \textit{cd} and \textit{ls} commands. Users can use the
+\textit{cat} command to see the preview of the log, or the \textit{more}
+command to see the preview and what files were edited in the commit and how
+many lines were added and removed from each file.
 
@@
 
-There are two issues with this design. One issue with this design is that a
-user will always move from the top-down without any additional filtering. This
-only serves one of our use-cases. The second issue being that it presents the
-user with the commit hashes, providing no explanation of what the commit or
-merge may contain.
+There are two issues with this design. One issue with is that a user will
+always move from the top-down without any additional filtering. This only
+serves one of our use-cases and can be very overwhelming. The second issue
+being that it presents the user with the commit hashes, providing no
+explanation of what the commit or merge may contain.
 
@@
 Putting the tool on the web enables users to use the system without having to
-install any software, making the system more accessible and platform
-independent.A web-based tool can utilize more interactive means of navigation,
-and provide a better explanation of what each commit or merge may contain than
-the command-line variant.
+install any software, making the system more accessible, more easily
+maintained, and platform independent. A web-based tool can utilize more
+interactive means of navigation, and provide a better explanation of what each
+commit or merge may contain than the command-line shell.
 
@@
 users to better navigate and understand changes in the kernel. These mechanisms
-include searching, commit messages, files edited, modules edited, and two types
-of tree viewers.
+include searching, the commit messages, the files edited, the modules edited,
+and two types of tree viewers.
 
@@
 
-Searching allows a user to filter our commits and merges that are irrelevant.
+Searching allows a user to filter commits and merges that are irrelevant.
 The search mechanism breaks down the results by release version. A user can
-further narrow down the search range using the commit date range. This commit
+further narrow down the search range using the commit date range. The commit
 date refers to the date of the creation of the top-level merge. Commits and
 merges that are returned may have a commit date beyond the given range if they
-were merged to a top-level merge within the date range. A user may then provide
-a search text, filtering on the author name, the commit id, or keyword from the
-log. Any part of the author name may show up in the results, including
+were merged to a top-level merge that is within the date range. A user may then
+provide a search text, filtering on the author name, the commit id, or keyword
+from the log. Any part of the author name may show up in the results, including
 searching by email address. The commit ID must be specified from the first
@@
 
-In the search results , the user is presented with the one-line log message
+In the search results, the user is presented with the one-line log message
 preview, the author's name and email, the date the commit was authored, and the
-date the commit was committed. \evantodo{a better definition for commit date}
-The commit date and author date will usually be the same, however, if the
-commit has been rebased, the commit date will be different than the authored
-date, and will always be after the authored date. The first and third entries
-in figure \ref{fig:results} have been rebased in the past. \evan{We are unable
-to show the rebases.}
+date the commit was last committed.  The commit date and author date will
+usually be the same, however, if the commit has been rebased, the commit date
+will be different than the authored date, and will always be after the authored
+date. The first and third entries in figure \ref{fig:results} have been rebased
+in the past, though we are unable to highlight rebases at this time.
 
@@
 
-The first view displays the full commit log. From this, a user is able to see
-hat they would see had they searched for the commit using git log. This doesn't
-provide additional information to the other tools, but helps to complete the
-functionality of the tool. This information provides a user with the
-information about the content of the commit and who has looked over the commit
-to ensure that it is good quality. The message for merges may also contain the
-information about the commits that are being merged into a given merge. The
-information within these messages is highly variable, and is completely
-dependent on the author's style. As the user moves toward the top-level merge,
-the quality of these messages generally improve.
+The first tab displays the full commit log. From this, a user is able to see
+what they would see had they searched for the commit using git log. This
+doesn't provide additional information to the other tools, but helps to
+complete the functionality of our tool. The commit log provides a user with
+the information about the content of the commit and who has signed-off on the
+commit to ensure that it is good quality. The message for merges may also
+contain the information about the commits that are being merged into a given
+merge. The information within these messages is highly variable, and is
+completely dependent on the author's style. As the user moves toward the
+top-level merge, the quality of these messages generally improves.
 
@@
 different commits within a merge, we simply add the added and removed lines to
-the values already stored. \evan{At this time, we don't have a mechanism to
-process the diff information to determine which lines were changed, and in what
-order.}
+the values already stored.
 
@@
 on in the one-line preview of the log-message. We heuristically extract the
-module by taking all text in the log preview until the first colon. Modules are
-logical partitions of the information in the kernel. Depending on where the
-author was working, modules can be general, such as ``bluetooth'' and
+module by taking all text in the log preview of commits until the first colon.
+Modules are logical partitions of the information in the kernel. Depending on
+where the author was working, modules can be general, such as ``bluetooth'' and
 ``wireless'', or can be quite specific for individual hardware, such as
 ``ath9k\_hw'' and ``wl1251''. Like files, we are able to show what modules are
-edited in merges and in commits. The image in figure \ref{fig:modules} shows
+edited in merges and commits. The image in figure \ref{fig:modules} shows
 the modules for a commit. A commit will only ever have one module that it is
@@
 The third tab shows the tree view (\ref{fig:tree}). The tree view is what makes
-this tool unique to the other tools. It provides a full view of the entire
-merge tree. It originally centres at the current commit or merge, but allows a
+this tool unique to the other tools, providing a full view of the entire merge
+tree. It originally centres at the current commit or merge, but allows a
 user to inspect the other commits and merges surrounding it. A user may then

### 2015-08-17 linvis.tex
Editing: Discussion

@@
 Our tool provides mechanisms and visualizations that other tools are unable to
-produce. Continuous mining of the repository provides an additional dimension
-to the dataset that would otherwise be lost, enabling us to reliably build
-trees from the information we have.
+produce. Continuous mining of the repository\cite{German} provides an
+additional dimension to the dataset that would otherwise be lost, enabling us
+to reliably build trees from the information we have.
 
@@
 Our goal was to build a tool to enable maintainers to effectively navigate and
-pull explanation of the changes performed to the kernel over a release.
-Achieving this goal includes removing information that does not pertain to given
-users. In figure \ref{fig:dag_view} and figure \ref{fig:tree_view}, we can
-visually compare the results returned from gitk and our tool for the top-level
-merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6'' from kernel version 3.10.
-The relevant information in both of these figures is identical, but the
-presentation durastically changes our understanding of what we are seeing.
+pull explanation of the changes performed to the kernel over the period of a
+release.  Achieving this goal includes removing information that does not
+pertain to given users. In figure \ref{fig:dag_view} and figure
+\ref{fig:tree_view}, we can visually compare the results returned from gitk and
+our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
+from kernel version 3.10. The relevant information in both of these figures is
+identical, but the presentation drastically changes our ability to comprehend
+what we are seeing. The primary difference is the removal of irrelevant
+information. In gitk (\ref{fig:dag_view}), the visualization results include
+commits and merges from other components of the kernel, while our tool
+(\ref{fig:tree_view}) only includes the results that are specific to the
+component of the kernel that we are interested in.
 
@@
 section of the kernel. The DAG view of this provides almost no explanation,
-futhermore, users must work to determine where this merge ends and the next one
+furthermore, users must work to determine where this merge ends and the next one
 begins.
@@
 Our tool further enhances our understanding by immediately giving us the files
-that were edited and the modules. We are able to determine that three files,
+and modules that were edited. We are able to determine that three files,
 ``palamas-regulator.c'' had two lines added and two lines removed,
 ``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had
-5 lines added and 2 lines removed in the merge. Finally, we are able to
-determine that only the module ``regulator'' was modified in this merge, and
-there were 5 commits in the merge to this module.
+5 lines added and 2 lines removed. Finally, we are able to determine that only
+the ``regulator'' module was modified in this merge and was modified by 5
+commits.
 
 \subsection{Performance Evaluation}
+Originally, we were performing all of our tree-navigation, file information
+gathering, and module information gathering using recursive SQL queries.
+Recursive queries have terrible performance, with the average response time
+from the database being around 3 seconds for a single operation. We found that
+it was far more response time-efficient to load an entire merge tree directly
+into memory, then iterate over the objects returned to build the tree
+structures, than it was to have the dbms perform the tree traversals, building
+the specific trees for a commit. At this time, we have re-implemented the
+breadcrumbs, the files, and the main tree views to enhance the response times.
+We will discuss some We will discuss the changes made.
 
-Re-structuring the kernel into a tree changes the runtime for the structures.
-Originally, we were performing all of our tree-navigation using recursive SQL
-queries. Recursive queries have terrible performance, with the average response
-time from the database being around 3 seconds for a single operation. We found
-that it was far more time-efficient to load an entire tree directly into
-memory, then iterate over the objects returned to build the tree structures,
-than it was to have the dbms perform the tree traversals, building the specific
-trees for a commit. At this time, we have re-implemented the breadcrumbs, the
-files, and the main tree to enhance the response times. We will discuss some We
-will discuss the changes made.
-
-The breadcrumbs are generated at response
-time, so they must be generated quickly. We began with a single recursive SQL
-query, but was unable get the results in the required time, causing all result
-pages to render slowly. We replaced the recursive query with a single query
-that would grab the entire merge tree, rooted at the top-level merge.
+The breadcrumbs are generated on page-load, so they must be generated quickly.
+We began with a single recursive SQL query, but was unable get the results in
+the desired time, causing all results pages to render slowly. We replaced the
+recursive query with a single query that would grab the entire merge tree,
+rooted at the top-level merge.
 
@@
 This maintains the O(n) runtime for generating the breadcrumbs, while using an
-iterative method for traversing the tree.
+iterative method for traversing the tree and dropping the response time from
+around 3 seconds to at most 500ms.
 
@@
 Due to how the asynchronous requests are performed, it is possible to query the
-server for the tree and file information directly, allowing other users to
-build their own front-ends for working with our tree-based system.
+server for the tree and file information directly, allowing users to build
+their own front-ends for working with our tree-based system.
 
@@
 The tree responses are generated as a single object. This object is the root
-node of the tree, and may contain many objects with the same fields in the
-``children'' field. The full response format for the tree looks like:
+node of the tree, and may contain a list of children in the ``children'' field.
+The full response format for the tree looks like:
 
@@
 
-The file responses contain only the files that the current merge or commit
+The file responses contain only the files that the selected merge or commit
 works with. The response is a single object in the form of a tree. This object
@@
 current position in the tree is an inner node, the response will contain the
-children in the ``children'' field, otherwise ``children'' will be null. The
+child nodes in the ``children'' field, otherwise ``children'' will be null. The
 full response format for the file response looks like:

### 2015-08-17 linvis.tex
Editing: Future Work and Conclusions

@@
 \section{Future Work}
-This project was implemented over the course of 4 months, interleaved with
-other course projects and assignments. There are still many areas that could
-use improvements before the full potential of our model can be reached and
-understood with our tool. Here, we outline various areas that still need more
-attention.
+This project was implemented over four months, interleaved with other course
+projects and assignments. There are still many areas that could use
+improvements before the full potential of our model can be realized.
+Here, we outline various areas that still need more attention.
 
 \subsection{Files}
-We have limited functionality surrounding files, for both search and presented
-information.
-
 There is currently no functionality surrounding searching by filename. Use-case
@@
 There is limited functionality to the presentation of the file information. At
-a minimum, the patches for the commits can be displayed. The patches can then
-be used for many more additional features. The patches can be used to piece
-together parts of the file to generate a full patch at a given merge rather
-than displaying each patch individually. The patches can also be used for
-determining what kinds of changes were made, if the lines are being added,
-removed, or replaced.
+a minimum, the patches for the commits can be displayed. From there, the
+patches can be used to piece together parts of the file to generate a single
+patch at a given merge rather than displaying each patch individually. The
+patches can also be used for determining what kinds of changes were made, if
+the lines are being added, removed, or replaced.
 
@@
 work-flow of maintainers. We believe that the tool is able to improve the
-work-flow and performance of maintainers because it is able to better assist
-users because it provides cleaner mechanisms of working with the data. It is
-able to provide more relevant information, while removing information that is
-irrelevant to a given module or set of merges.
+work-flow and performance of maintainers because it provides cleaner mechanisms
+of visualizing and presenting the commit information. It is able to provide
+more relevant information, while removing information that is irrelevant to a
+given module or set of merges.
 
-We could perform user-testing, or have some example cases where the tool is
-used and the users provide feedback on how it improves or hurts their work
-flow.
+We can either perform user-testing to show how a users workflow changes, or we
+could have maintainers evaluate and critique the tool and use their feedback to
+determine if the tool meets our goals.
 
@@
 information for the respective view type. For single commits or very small
-merges, getting everything directly from the database and performing everything
-server-side performs better than using the asynchronous requests, as there
-isn't the overhead of multiple round trips; however, as the size of the merges
-grow, the performance becomes far worse than the asynchronous system and in
-some cases will time out. The asynchronous approach adds some overhead,
-profiled at roughly 500ms, regardless of the response size. The asynchronous
-approach has many benefits, over building the response on the server, including
-the ability to add spinners, which is an improvement to user-experience over
-looking at a blank page. Further work can be done to improve the results beyond
-the response time of 500ms using improved queries and more efficient methods of
-building the tree.
+merges, getting everything directly from the database and generating the pages
+server-side performs better than using the asynchronous requests, there
+isn't the overhead of multiple round trips. As the size of the merges grow,
+the performance becomes far worse than the asynchronous system and in some
+cases will time out. The asynchronous approach adds some overhead, profiled at
+roughly 500ms, regardless of the response size. The asynchronous approach has
+many benefits over building the response on the server, including spinners, a
+public API, and lower response times for larger merges. Further work can be
+done to improve the results beyond the response time of 500ms using improved
+queries and more efficient methods of building the tree. \newpage
 
@@
 Our tool shows promising results, providing a clearer picture of the relation
-between the merges and commits for sections of the kernel in relation to the
-DAG-based models. Not only are we able to provide a clearer visualization, we
-are able to provide information about the files edited by all of the commits
-contained within a merge.
+between the merges and commits when compared to DAG-based models. Not only are
+we provide a clearer visualization, we provide information about the files
+edited by all of the commits contained within a merge, and we provide the
+information about what modules are edited in a merge.
 
-While we are able to gather a better explanation of the changes happening at
+While we are able to generate a better explanation of the changes happening at
 each commit and merge, the tree-model requires more information than what is
@@
 
+A tree-based model appears to be sufficient for creating a clean visualization
+and explanation of the changes in the Linux kernel.
+
 \bibliographystyle{plain}
@@
 
-\balance
-
 \end{document}

### 2015-08-17 linvis.tex
Editing: Implementation details

@@
 \section{Implementation Details}
-We built a we-based tool to provide information about the kernel.
+We built a web-based tool to provide information about the kernel.
 We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
@@
 threading, Nginx uses an event-driven architecture, making it more scalable,
-proving itself as being the web server behind Netflix, Hulu, GitHub, Tumblr,
-and many other popular sites.
+proving itself as being the web server behind large sites like Netflix, Hulu,
+GitHub, Tumblr, and many other popular
+sites\footnote{\url{http://wiki.nginx.org/Main}}.
 
@@
 are multiple reasons behind choosing postgresql over another dbms. The main
-reason being that the dataset provided from \evantodo{paper} was generated from
+reason being that the dataset built by German et al. was generated from
 a postgresql database, allowing us to import the data directly without any
@@
 
-We chose the python micro-framework Flask for generating the dynamic content.
-Working with flask is simple and easy.
+We chose the python micro-framework
+Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the
+dynamic content.  Working with flask is simple and easy, allowing us to work in
+Python.
 
@@
 where valid top-level merges are found for that version.
-
-\evantodo{Talk a bit about the heuristics used for building the tree from the
-DAG}
 
@@
 done to improve the results beyond the response time of 500ms using improved
-queries and more efficient methods of building the tree. \newpage
+queries and more efficient methods of building the tree.
 

### 2015-08-17 linvis.tex
Formatting: Minor changes

@@
 \newif\ifdraft
-\drafttrue
+% \drafttrue
 
@@
 different commits within a merge, we simply add the added and removed lines to
-the values already stored.
+the values already stored.\newpage
 
@@
 experience. We are using AJAX requests to generate the asynchronous requests,
-are are having very good results.
+are are having very good results.\newpage
 
@@
 done to improve the results beyond the response time of 500ms using improved
-queries and more efficient methods of building the tree.
-
+queries and more efficient methods of building the tree.\newpage
 \section{Conclusion}

### 2015-08-28 linvis.tex
Fix: Spacing on verbatim section
Edits: Makefile, related work, introduction, and citations
Added citations.bib to the makefile

Added information to the related and introduction section.

Added citations for the related works.

@@
 	\includegraphics[width=0.47\textwidth]{figures/gitk.png}
-	\caption{DAG view in Gitk}
+	\caption{A sample DAG view of commits from Linux in Gitk}
 	\label{fig:gitk}
@@
 
-Git log, Gitk, and other similar tools only store the authored date of a
-commit. A commit may be merged into the project at any time, regardless of when
-the commit was authored, making date-range queries meaningless. Performing a
-date-range query using these tools over the span of a release will contain
-results that are for previous versions of the kernel or may not contain all
-of the results for that kernel if patches were merged after the release date.
+Git log, Gitk, and other similar tools only store the name of the author, authored
+date of a commit. A commit may be merged into the project at any time, regardless of
+when the commit was authored, making date-range queries meaningless. Performing a
+date-range query using these tools over the span of a release will contain results
+that are for previous versions of the kernel or may not contain all of the results
+for that kernel if patches were merged after the release date.
 
@@
 All top-level merges to a given version of the kernel are merged between the
-start and release date. Our tree construction allows us to use the top-level
-merges to find all merges and commits that have an effect on a given version of
-the kernel. The model allows us to better provide mechanisms for performing
-date-range queries, providing access to all commits that have an effect on a
-given release, even if the commit comes before the start date or after the
-release date.
+start and release date. The tree construction allows us to find all commits that are
+related to a given release of the kernel regardless of the authorship date, allowing
+our system to provide better mechanisms for performing date-range queries.
 
 \section{Related Work}
-To our knowledge, we know of no repository visualization systems using a
-tree-based model. This may be due to the fact that more information is required
-to generate the trees than what is stored by git. A server must sit and monitor
-the repository as it changes over time. Many of the visualization tools are
-designed for a general-purpose approach, where any repository is directly
-usable without additional information. Polling for this additional information
-makes the application specific to a single repository or project.
+To our knowledge, we know of no repository visualization systems that builds a tree
+from the merge. CVS and svn construct trees; however, the root node for these trees
+is the branch point. This may be due to the fact that more information is required
+to generate the trees than what is stored by git. A server must sit and monitor the
+repository as it changes over time. Many of the visualization tools are designed for
+a general-purpose approach, where any repository is directly usable without
+additional information. Polling for this additional information makes the
+application specific to a single repository or project.
 
@@
 to the previous commit in the chain, or jump to the parent merge. This suggest
-that Gitk and similar tools are designed for users working with a bottom-to-top
-approach. With this model, it is difficult for users to see what other commits
-are involved with a given merge. Gitk only provides file information on
-commits, as it is unable to quickly find all files edited by the commits within
-a merge. It is up to the user to remember what files were edited in each of the
-commits involved in a given merge. In some of the larger merges, some
-containing over 1500 edited files, this task become nearly impossible.
-% cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
-Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}),
-are unable to provide a visualization for projects that are as large as the
-kernel.
+that Gitk and similar tools are designed for users working from a given commit or
+leaf-node toward the root.  With this model, it is difficult for users to see what
+other commits are involved with a given merge. Gitk only provides file information
+on commits, as it is unable to quickly aggregate the files and commits that are part
+of a merge. It is up to the user to remember what files were edited in each of the
+commits involved in a given merge. In some of the larger merges, some containing
+over 1500 edited files, this task become nearly impossible.  % cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files 
+Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}), are
+unable to provide a visualization for projects that are as large as the kernel.
 
@@
 	\includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-	\caption{Github Failure Message}
+	\caption{Github Failure Message showing the DAG of the Linux Project}
 	\label{fig:gitfail}
@@
 pictured in figure \ref{fig:treemap}, for
-displaying the file structure of a given version. Treemaps are good for
-visualizing the topology of shallow, wide trees, making it a good candidate for
-showing file systems. Treemaps are limited in what information can be
-displayed. The information within the treemap does not provide clear insight on
-what changes were made to the files in the given release, only providing
-information on where the file is located within the file system.
-File systems and the merge tree of the kernel share many similarities, in both
-cases, they represent trees that are wide and shallow, suggesting that a
-treemap could potentially work to visualize the commit and merge structure of
-the kernel.
+displaying the file structure of a given version of the kernel. Treemaps are good
+for visualizing the topology of shallow, wide trees, making it a good candidate for
+showing file systems. Treemaps are limited in what information can be displayed. The
+information within the treemap does not provide clear insight on what changes were
+made to the files in the given release, only providing information on where the file
+is located within the file system.  File systems and the merge tree of the kernel
+share many similarities, in both cases, they represent trees that are wide and
+shallow, suggesting that a treemap could potentially work to visualize the commit
+and merge structure of the kernel.
 
-The treemap is unable to take into account multiple commits behind the same
-merge with the same name. Commits and merges have an ordering in which they
-were added to the kernel, which cannot be represented by a treemap, but is
-important and can be used to assist in building the tree.
+The treemap is unable to take into account multiple commits behind the same merge
+with the same log preview message. Commits and merges have an ordering in which they
+were added to the kernel, which cannot be represented by a treemap, but is important
+and can be used to assist in building the tree.
 
@@
 `c267548755a184ef97301071300c1739a564e135' is returned when the user searches
-for a commit id of `c26' in the 3.10 Linux kernel.
+for a commit id of `c26' in the 3.10 Linux kernel. We will use this commit for
+showing each of the features of the tool.
 
@@
 Our tool further enhances our understanding by immediately giving us the files
-and modules that were edited. We are able to determine that three files,
-``palamas-regulator.c'' had two lines added and two lines removed,
-``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had
-5 lines added and 2 lines removed. Finally, we are able to determine that only
-the ``regulator'' module was modified in this merge and was modified by 5
-commits.
+and modules that were edited in the branch. We are able to determine that three
+files, ``palamas-regulator.c'' had two lines added and two lines removed,
+``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
+lines added and 2 lines removed. Finally, we are able to determine that only the
+``regulator'' module was modified in this merge and was modified by 5 commits.
 
@@
 
-\subsection{Testing}
+\subsection{Evaluation}
 At this time, we have no evidence that our tool is able to improve the
@@
 
+\nocite{*}
 \bibliographystyle{plain}

### 2016-04-19 linvis.tex
Updated Abstract and new tabulation
Re-wrote and re-worked the abstract section.

Changed from hard-tabs to soft-tabs.

@@
 \begin{abstract}
-	With an average of over 900 top-level merges into the Linux kernel per
-	release, some containing thousands of commits, many containing hundreds
-	of commits, maintenance of older versions of the kernel becomes nearly
-	impossible. For security, performance, and changing hardware needs,
-	maintainers must understand the changes made to current versions of the
-	kernel, and understand how these changes fit into older version in
-        order to make the necessary merges and modifications to the older
-        versions. Current tools provide information about repositories through
-        the directed acyclic graph (DAG) which is helpful for smaller projects,
-        but with the scale and number of branches in the kernel, the DAG
-        becomes overwhelming very quickly.
 
-	In this paper, we present a tool that uses the dataset collected by
-	German et al. to build a tool that uses a tree-based model instead of
-        the DAG. This tool is designed to cater to the needs of users looking
-        for a top-down approach, or a bottom-up approach to navigating the
-        commit information of the kernel.
+        With an average of more than 900 top-level merges into the Linux kernel per
+        release, many containing hundreds of commits and some containing thousands,
+        maintenance of older versions of the kernel becomes nearly impossible.
+
+        Various commercial products, such as the Android platform, run older
+        versions of the kernel. Due to security, performance, and changing hardware
+        needs, maintainers must understand how updates made to the current version
+        of the kernel fit into older versions in order to make the necessary
+        patches.
+
+        Current tools provide information about repositories through the directed
+        acyclic graph (DAG), which is helpful for smaller projects. However with the
+        scale and number of branches in the kernel the DAG becomes overwhelming very
+        quickly.
+
+        In this paper, we present a tool that uses a tree-based model, as described
+        by German et al., to construct a clearer summary of the changes made. This
+        tool is designed to allow users to explore the topology of the repository
+        for a given release of the kernel and view an aggregated summary of the
+        changes made.
+
 \end{abstract}
@@
 Maintainers of older version of the Linux kernel must sift through thousands of
-commits with tools that are unable to filter and effectively visualize projects
-that are the scale of the kernel. Tools like Gitk provide users with the
-full view of the DAG, showing all of the merges and commits. For smaller
-projects, this information is useful, showing the relation between various
-branches. The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges,
-and the links between them with large modular projects, like the Linux kernel,
-making it difficult to derive an explanation of the changes made. In other
-cases, the kernel is simply too large for the system to generate a
-visualization for. Github provides a DAG view for most projects, but is unable
-to display the visualization for projects as large as the Linux kernel.
+commits with tools that are unable to filter and effectively visualize projects that
+are the scale of the kernel. Tools like Gitk provide users with the full view of the
+DAG, showing all of the merges and commits. This information is useful in smaller
+projects, showing the relation between various branches. The DAG becomes a tangled
+mess (\ref{fig:gitk}) of commits, merges, and the links between them with large
+modular projects, like the Linux kernel, making it difficult to derive an
+explanation of the changes made. In other cases, the kernel is simply too large for
+the system to generate a visualization for. Github provides a DAG view for most
+projects, but is unable to display the visualization for projects as large as the
+Linux kernel.
 

### 2016-04-22 linvis.tex
Bubble Tree, and editing
Making a bunch of edits.

@@
         maintenance of older versions of the kernel becomes nearly impossible.
-
         Various commercial products, such as the Android platform, run older
@@
 
-Maintainers of older version of the Linux kernel must sift through thousands of
-commits with tools that are unable to filter and effectively visualize projects that
-are the scale of the kernel. Tools like Gitk provide users with the full view of the
-DAG, showing all of the merges and commits. This information is useful in smaller
-projects, showing the relation between various branches. The DAG becomes a tangled
-mess (\ref{fig:gitk}) of commits, merges, and the links between them with large
-modular projects, like the Linux kernel, making it difficult to derive an
-explanation of the changes made. In other cases, the kernel is simply too large for
-the system to generate a visualization for. Github provides a DAG view for most
-projects, but is unable to display the visualization for projects as large as the
-Linux kernel.
+Maintainers of older versions of the Linux kernel must sift through thousands of
+commits with tools that are unable to filter and effectively visualize projects at
+the scale of the kernel. Tools like Gitk use the DAG model, showing all commits and
+merges in chronological order to when the commit was authored. This representation
+works in smaller projects, enabling users to see when changes are made, when these
+changes are merged, and how each branch is interacting. In large modular projects,
+like the Linux kernel, the DAG becomes a mess of merges and commits (figure
+\ref{fig:gitk}), losing its visual meaning. In some cases, the Linux kernel is
+simply too large for the system to generate a visualization. Github provides a DAG
+view for many projects, but is unable to display the visualization for projects as
+large as the Linux kernel (figure \ref{fig:gitfail}).
 
@@
 
-Git log, Gitk, and other similar tools only store the name of the author, authored
-date of a commit. A commit may be merged into the project at any time, regardless of
-when the commit was authored, making date-range queries meaningless. Performing a
-date-range query using these tools over the span of a release will contain results
-that are for previous versions of the kernel or may not contain all of the results
-for that kernel if patches were merged after the release date.
+Git cannot store the date that a commit was merged into a branch due to the rebase
+operation.  When a rebase occurs, several changes could be made to the commit. It
+may be squashed with other commits, combining the effects of several commits into a
+single commit. It may be dropped altogether, in which case the commit is no-longer
+merged.  It may be that the entire branch has simply been re-rooted to another
+position on the DAG. Due to this, tools are unable to determine when a commit is
+merged. A user may decide that it is important to be able to see only the commits
+contributing to a given release, that is commits that were merged with the
+Linus-branch of the kernel between two given dates. This operation is not possible
+without the collection of additional data that is not provided by git. Git only
+enables users to see when commit was authored, not when it was merged into a branch.
 
 German et al. recorded the metadata of every commit and merge in the Linux git
-repository starting in 2011 \cite{German}, including chronological data that is
-not stored by git. We are able to use this information to generate sets of
-trees instead of a DAG. Each tree is rooted by the top-level merge, performed
-by Linus, which merges the branches or a set of commits into the kernel.
+repository starting in 2011 \cite{German2015}, including chronological data that is
+not stored by git. We are able to use this information to determine when a commit
+has been merged into the kernel, regardless of rebases preceding and proceeding that
+point in time. This information is then converted by a heuristic method, described
+by German, to generate sets of trees, which we use as our core model instead of
+using the DAG. Each tree is rooted by the top-level merge performed by Linus, which
+merges the branches or a set of commits into the kernel.
 
 In this paper, we present our design decisions behind a web-based
-tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
-model instead of a DAG. Our tool provides information about the location of a
-given commit in the respective merge tree, the files edited, the modules
-edited, and the commit message for a given commit or merge. The tool allows
-users to apply various filters, including the release version, along with a
-keyword or phrase from the log preview, the name of the author, or the commit
-id. The user can request all the top-level merges containing a commit or merge
-that matches the query, or all commits and merges that match the query.
+tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based model
+instead of a DAG. Our tool provides information about the location of a given commit
+in the respective merge tree, the files edited, the modules edited, and the commit
+message for a given commit or merge. The tool allows users to apply various filters,
+including the release version, along with a keyword or phrase from the log preview,
+the name of the author, or the commit id. The user can request all the root-level
+merges containing a commit or merge that matches the query, or all commits and
+merges that match the query.
 
-Our core model provides advantages to the DAG model used by other tools.
-Breaking the commits and merges by merge tree removes commits and merges of
-other components of the kernel from the visualization, providing a clearer
-picture of what is happening in a given merge and providing additional
-support for edited file information. Furthermore, we are able to understand the
-changes being made, in the order they are brought into the kernel.
+Using the tree provides many advantages to the DAG model, but there are two
+drawbacks. The first drawback is that a server must query the state of the master
+repository since time information is not stored by git. The second drawback is that
+the tree construction is a heuristic approach, and in some cases may break.
 
-All top-level merges to a given version of the kernel are merged between the
-start and release date. The tree construction allows us to find all commits that are
-related to a given release of the kernel regardless of the authorship date, allowing
-our system to provide better mechanisms for performing date-range queries.
+Our core model provides advantages to the DAG model used by other tools. By
+separating and focusing on sub-trees from the whole repository, we are able to
+filter out commits and merges that do not pertain to a component of the kernel. This
+provides a clearer picture of what is happening in a given merge. The second benefit
+to the tree model is that it enables us to automatically aggregate metadata.
+
+In addition to this, the model provides us with a mechanism of performing date-range
+queries with meaningful results. Using the tree built by a continuous mining system
+enables us a mechanism to perform date-range queries with meaningful results.
 
 \section{Related Work}
-To our knowledge, we know of no repository visualization systems that builds a tree
-from the merge. CVS and svn construct trees; however, the root node for these trees
-is the branch point. This may be due to the fact that more information is required
-to generate the trees than what is stored by git. A server must sit and monitor the
-repository as it changes over time. Many of the visualization tools are designed for
-a general-purpose approach, where any repository is directly usable without
-additional information. Polling for this additional information makes the
-application specific to a single repository or project.
 
-\subsection{Gitk}
-Gitk, Gitg, and other similar git repository visualizing tools use the DAG as
-the central organization model. For smaller projects, this model is sufficient
-for providing meaningful information, but in larger projects like the Linux
-kernel, the branches become too tangled and too deep to derive any meaning from
-the DAG. The DAG model allows users to go to the next commit in the chain, go
-to the previous commit in the chain, or jump to the parent merge. This suggest
-that Gitk and similar tools are designed for users working from a given commit or
-leaf-node toward the root.  With this model, it is difficult for users to see what
-other commits are involved with a given merge. Gitk only provides file information
-on commits, as it is unable to quickly aggregate the files and commits that are part
-of a merge. It is up to the user to remember what files were edited in each of the
-commits involved in a given merge. In some of the larger merges, some containing
-over 1500 edited files, this task become nearly impossible.  % cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
-Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}), are
-unable to provide a visualization for projects that are as large as the kernel.
+To our knowledge, we know of no git repository visualization systems that builds a
+tree from the merge. This may be because more information is required to generate
+the trees than what is stored by git. Git uses a DAG model, which enables
+history-altering actions like rebases to occur, whereas older tools like CVS and svn
+construct trees\cite{CVS2008} but do not work with history-altering actions.
 
-\begin{figure}[h!]
+A server must sit and monitor the git repository as it changes over time to handle
+changes to history. Many of the visualization tools are designed for a
+general-purpose approach, where any repository is directly usable without additional
+information. Polling for this additional information makes the application specific
+to a repository or project.
+
+\subsection{Other Visualizers}
+
+Gitk, Gitg, and other similar git repository visualizing tools use the DAG as the
+central organization model. This model is sufficient for smaller projects, providing
+meaningful information; however, in larger projects, like the Linux kernel, the
+branches become too tangled and too deep to derive any meaning from the DAG. Users
+can interact with the DAG by going to the previous commit in the chain or jumping to
+the parent merge. This suggests that Gitk and similar tools are designed for users
+working from a given commit or leaf-node toward the root. With this model, it is
+difficult for users to see what other commits are involved with a given merge.
+
+Furthermore, the DAG model is only able to provide file-level information on a
+commit level. It is unable to quickly aggregate the file data over all commits that
+are part of a merge. If a user would like to view what files were edited at a merge,
+they must manually aggregate this information by hand. While some merges are
+relatively small, containing less than 10 commits, other merges are large,
+containing over 1500 edited files, as in the case of commit \textit{73287a4}. This
+issue is important for authorship and licensing. A merge can only show one author,
+the contributor making the merge, but does not show who wrote the commits contained
+within that merge.
+
+\begin{figure}
         \centering
@@
 \subsection{Treemap visualizer}
+
 University of Maryland presented a treemap visualization
@@
 
-\begin{figure}[h]
+\begin{figure}
         \centering
@@
 \subsection{List viewer}
+
 Moghaddam designed a web-based visualization
@@
 
-\section{Design}
+\section{Methodologies}
+\evan{Correct section title? Other than just trying stuff, what were the
+methodologies?}
+
 The goal is to build a tool that makes navigation of the kernel commit
 information simpler, while providing an explanation of the changes occurring at
-each commit and merge. Part of making the information usable is to narrow the
-results down into only the information that is relevant to what the user is
-looking for. We designed the tool with two use-cases in mind, though a user may
-switch roles as they work.
+each commit and merge. We designed the tool with two use-cases in mind, though a
+sumer may switch between the cases as they work.
 
-\textit{Use-case 1: top-to-bottom approach}\\
-These are users that are maintaining a section of the kernel and would like to
-pick a merge and merge it directly into their current repository. This is
-useful for minimizing required implementation on the side of the maintainer,
-allowing them to see what changes are made higher up in the kernel and bundle
-multiple changes together. We must provide them with methods of navigating from
-a more general set of merges toward a specific set of merges and commits.
+\textit{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
+These are users that are maintaining a section of the kernel and would like to pick
+a merge and merge it directly into their current repository, which is useful for
+reducing the amount of re-implementation work. For them, it is
+important to have the ability to aggregate files and modules that are being changed
+somewhere in the merge. It is also important for them to be able to navigate from
+the root toward the largest merge that contains only the commits that modify the
+modules that they are working with.
 
-\textit{Use-case 2: bottom-to-top approach}\\
-These are users that know what merge or commit they are working with and would
-like to see what else is happening around their commit. This is useful to see
-how changes to a given file will break other changes made in other commits.
-This is primarily for maintainers that have to perform some implementation work
-in order to correctly merge the patches from the future. These users may also
-know what module(s) they are dealing with, and are looking for modules that are
-relevant to the area they are working on. We must provide them methods of
-navigating from a specific commit to a more general merge, allowing them to see
-a bigger picture.
+\textit{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
+These are users that start with a known merge or commit and would like to see
+what other changes are being made in commits that are in the same merge. This is
+useful to see how changes in a given file will break other changes made in the
+surrounding commits in the same merge. This is primarily for maintainers that need
+to perform some implementation work in order to correctly merge patches from the
+future. We must provide these users a mechanism for navigating from a single commit
+toward the root, allowing them to see a bigger picture.
 
 \subsection{Core Model}
-Each top-level merge into the kernel is the root-node to a tree. Each merge
-may contain zero or more children made of other merges or commits. Commits have
-no children, and therefore are always leaf nodes.  The commits contain the
-information about the files that are edited along with how many lines are added
-and removed from each file. Commits also store the module of the kernel that
-they are working with.
+
+We root the tree with the merge made by Linus. Each merge may contain zero or more
+children, which are other merges and commits. If a merge is cherry-picked or rebased
+there will be no children. Commits have no children and are always leaf nodes. The
+commits contains the metadata for the changes made. This metadata includes the files
+changed, the lines added and removed from each file, the author, the date the commit
+was merged into the merge that led to being merged into the kernel, and the date the
+commit was authored, the diff, and the commit log. Merges contain less metadata,
+only storing the author of the merge, the log, the commit date, and the author date.
+The details of the model are outlined by German \cite{German2015}.
+
+\section{Design}
+
+In this section, we describe the progression of our tool design, showing the
+benefits and drawbacks of each component.
 
 \subsection{Command-line Shell}
-Our initial intuition was to work with the commits and merges as trees. We use
-the merges as directories, and commits as files, like in a file structure.
-There are many similarities between a file system and the structure of the
-commit information. The primary similarity is that they are both trees, and
-both trees are generally fairly shallow and wide. File systems have been
-around for quite some time, and there are many tools for visualizing and
-working withe the file structure. The first tool was the command-line. We built
-a small command-line shell (\ref{fig:shell}) that allows navigation through
-the commit and merges instead of files and directory. The shell design
-permitted a user to move through the merges like they would through
-directories with the \textit{cd} and \textit{ls} commands. Users can use the
-\textit{cat} command to see the preview of the log, or the \textit{more}
-command to see the preview and what files were edited in the commit and how
-many lines were added and removed from each file.
 
-\begin{figure}[h]
+Our initial intuition was to work with the commits and merges as directory trees
+where the commits are files and the merges are directories. There are many
+similarities between the tree structure of a file system and the tree structure of
+the repository. Both trees are generally shallow, but very wide. File systems have
+been studied for quite some time and there are many tools capable of visualizing and
+working with the file structure.
+
+The first tool we implemented was a small command-line shell (figure
+\ref{fig:shell}) that allows users to navigate through the commits and merges rather
+than files and directories. The shell design permitted a user to move through merges
+like they would through directories with the \verb|cd| and \verb|ls| commands. Users
+can use the \verb|cat| command to see the log preview, or the \verb|more| command to
+see the preview and how many lines were added and removed from the files in the
+commit.
+
+\begin{figure}
         \centering
@@
 
-There are two issues with this design. One issue with is that a user will
-always move from the top-down without any additional filtering. This only
-serves one of our use-cases and can be very overwhelming. The second issue
-being that it presents the user with the commit hashes, providing no
-explanation of what the commit or merge may contain.
+The shell does not meet our design goals. It only provides users with the ability to
+navigate from the root. If there is a commit or merge that they would like to
+investigate, they must find it with another tool and navigate through the tree to
+find it. Furthermore, it does not aggregate information in the merges, so users must
+manually aggregate the metadata. Another issue with this design is that it presents
+the users with the commit hashes, which are quite long and unpleasant to remember
+type into the shell, but also do not provide any quick explanation of what the
+commit or merge may contain.
 
 \subsection{Web-based Tool}
-There are multiple reasons for building the system into a web-based tool.
-Putting the tool on the web enables users to use the system without having to
-install any software, making the system more accessible, more easily
-maintained, and platform independent. A web-based tool can utilize more
-interactive means of navigation, and provide a better explanation of what each
-commit or merge may contain than the command-line shell.
 
-Our implementation of the web-based tool provides many mechanisms to enable
-users to better navigate and understand changes in the kernel. These mechanisms
-include searching, the commit messages, the files edited, the modules edited,
-and two types of tree viewers.
+We replace the command-line interface with a web-based tool. Creating a web-based
+tool enables users to use the system without having to install software or carry a
+large database, making it more accessible, more easily maintainable, and platform
+independent. A web-based tool can utilize more interactive means of navigation and
+provide better explanations of commits and merges than the command-line interface.
 
-\begin{figure}[h]
+Our tool uses the following mechanisms to reach our goals of better navigation and
+better explanation of the changes.
+
+\begin{itemize}
+        \item Filter by search
+        \item View files edited
+        \item View modules
+        \item Tree viewer
+\end{itemize}
+
+\begin{figure}
         \centering
@@
 
-Searching allows a user to filter commits and merges that are irrelevant.
-The search mechanism breaks down the results by release version. A user can
-further narrow down the search range using the commit date range. The commit
-date refers to the date of the creation of the top-level merge. Commits and
-merges that are returned may have a commit date beyond the given range if they
-were merged to a top-level merge that is within the date range. A user may then
-provide a search text, filtering on the author name, the commit id, or keyword
-from the log. Any part of the author name may show up in the results, including
-searching by email address. The commit ID must be specified from the first
-character to the last character. For example, the commit
-`c267548755a184ef97301071300c1739a564e135' is returned when the user searches
-for a commit id of `c26' in the 3.10 Linux kernel. We will use this commit for
-showing each of the features of the tool.
+Searching allows a user to filter commits and merges that are irrelevant.  The
+search mechanism breaks down the results by release version. A user can further
+narrow down the search range using the commit date range. The commit date refers to
+the date of the creation of the top-level merge. Commits and merges that are
+returned may have a commit date beyond the given range if they were merged to a
+top-level merge that is within the date range. A user may then provide a search
+text, filtering on the author name, the commit id, or keyword from the log. Any part
+of the author name may show up in the results, including searching by email address.
+The commit ID must be specified from the first character to the last character. For
+example, the commit `c267548755a184ef97301071300c1739a564e135' is returned when the
+user searches for a commit id of `c26' in the 3.10 Linux kernel. We will use this
+commit for showing each of the features of the tool.
 
-\begin{figure}[h]
+\begin{figure}
         \centering
@@
 
-In the search results, the user is presented with the one-line log message
-preview, the author's name and email, the date the commit was authored, and the
-date the commit was last committed.  The commit date and author date will
-usually be the same, however, if the commit has been rebased, the commit date
-will be different than the authored date, and will always be after the authored
-date. The first and third entries in figure \ref{fig:results} have been rebased
-in the past, though we are unable to highlight rebases at this time.
+In the search results, the user is presented with the one-line log message preview,
+the author's name and email, the date the commit was authored, and the date the
+commit was last committed. The commit date and author date will usually be the same;
+however, if the commit has been rebased or cherry-picked into a merge that
+eventually is merged into the kernel, the commit date and authored date will be
+different. We can see this in the first and third entries in figure
+\ref{fig:results}.
 
-\begin{figure}[h]
+\begin{figure}
         \centering
@@
 
-The first tab displays the full commit log. From this, a user is able to see
-what they would see had they searched for the commit using git log. This
-doesn't provide additional information to the other tools, but helps to
-complete the functionality of our tool. The commit log provides a user with
-the information about the content of the commit and who has signed-off on the
-commit to ensure that it is good quality. The message for merges may also
-contain the information about the commits that are being merged into a given
-merge. The information within these messages is highly variable, and is
-completely dependent on the author's style. As the user moves toward the
-top-level merge, the quality of these messages generally improves.
+Once a user selects a commit or merge to investigate, they are presented with a
+tabbed pane allowing them to view the full commit log, the files edited, the modules
+involved, and the tree views.
 
-\begin{figure}[h]
+The first tab displays the full commit log. From this, a user is able to see what
+they would see had they searched for the commit using git log. This doesn't provide
+additional information to the other tools, but helps to complete the functionality
+of our tool. The commit log provides a user with the information about the content
+of the commit and who has signed-off on the commit to ensure that it is good
+quality. The message for merges may also contain the information about the commits
+that are being merged into a given merge. The information within these messages is
+highly variable, and is completely dependent on the author's style. As the user
+moves toward the top-level merge, the quality of these messages generally improves.
+
+\begin{figure}
         \centering
@@
 
-The second tab is the files tab. This tab provides information on what files
-have been edited, how many lines were added, and how many lines were removed in
-a given commit. This functionality is similar to the other tools available. Our
-tree-based design model allows us to extend this functionality to merges, which
-the other tools are unable to show. When a file is edited multiple times from
-different commits within a merge, we simply add the added and removed lines to
-the values already stored.\newpage
+The second tab is the files tab. This tab provides information on what files have
+been edited, how many lines were added, and how many lines were removed in a given
+commit. This functionality is similar to the other tools available. Our tree-based
+design model allows us to extend this functionality to merges, which the other tools
+are unable to show. When a file is edited multiple times from different commits
+within a merge, we simply add the added and removed lines to the values already
+stored.
 
-\begin{figure}[h]
+\begin{figure}
         \centering
@@
 
-The modules tab shows the modules that are contained within the commit. Modules
-are not natively recognized by git, and are not going to be present in all
-repositories. In the Linux repository, authors put the module they are working
-on in the one-line preview of the log-message. We heuristically extract the
-module by taking all text in the log preview of commits until the first colon.
-Modules are logical partitions of the information in the kernel. Depending on
-where the author was working, modules can be general, such as ``bluetooth'' and
-``wireless'', or can be quite specific for individual hardware, such as
-``ath9k\_hw'' and ``wl1251''. Like files, we are able to show what modules are
-edited in merges and commits. The image in figure \ref{fig:modules} shows
-the modules for a commit. A commit will only ever have one module that it is
-associated with.
+The modules tab shows the modules that are contained within the commit. Modules are
+not natively recognized by git, and are not going to be present in all repositories.
+In the Linux repository, authors put the module they are working on in the one-line
+preview of the log-message. We heuristically extract the module by taking all text
+in the log preview of commits until the first colon.  Modules are logical partitions
+of the information in the kernel. Depending on where the author was working, modules
+can be general, such as ``bluetooth'' and ``wireless'', or can be quite specific for
+individual hardware, such as ``ath9k\_hw'' and ``wl1251''. Like files, we are able
+to show what modules are edited in merges and commits. The image in figure
+\ref{fig:modules} shows the modules for a commit. A commit will only ever have one
+module that it is associated with.
 
-\begin{figure}[h]
+Finally, we have the tree view tab. The tree view is designed for providing easy
+navigation within a tree and providing a clear topological view of the branch off
+the kernel. We have experimented with various tree designs to find a design that
+allows for easy navigation and visualization of both large and small trees. We
+discuss the various tree designs in section \ref{treeview_section}.
+
+\subsection{Tree Views} \label{treeview_section}
+
+The tree view is what makes this tool unique to other tools, providing a full view
+of the entire merge tree. We have experimented with various types of trees; list
+trees are similar to the representation by many file system explorers,
+Reingold-Tilford trees which clearly represent the tree structure, and bubble trees
+which organize the data hierarchically by having the parent node contain the child
+nodes.
+
+\subsubsection{List Tree}
+
+The list tree  viewer is in the form of nested lists, and is designed to more
+closely model the tree hierarchy view found in file browsers. This tree only
+contains the commits and merges that are within the subtrees of the current merge. A
+commit will never have any items in this tree, as it is a leaf. To accompany the
+tree, we include breadcrumbs at the top of the page to enable a user to navigate
+both from the root to the leaves and from a leaf to the root. The last item in the
+breadcrumb list is the current commit, the next item is the parent of the current
+commit, and the first item is the root merge into the kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
+        \caption{List Tree}
+        \label{fig:list_tree}
+\end{figure}
+
+\subsubsection{Reingold-Tilford Tree}
+
+The Reingold-Tilford tree produces very nice results, creating an intuitive
+representation of the tree. The representation has a clear notion of root and
+leaves, and how to navigate in either direction.
+
+\begin{figure}
         \centering
         \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
-        \caption{Commit Tree View}
+        \caption{Reingold-Tilford Tree}
         \label{fig:tree}
@@
 
-The third tab shows the tree view (\ref{fig:tree}). The tree view is what makes
-this tool unique to the other tools, providing a full view of the entire merge
-tree. It originally centres at the current commit or merge, but allows a
-user to inspect the other commits and merges surrounding it. A user may then
-navigate to any of the commits by simply double-clicking on it.
+Clicking on subtrees will cause them to minimize, only showing parts of the tree
+that are of interest to the user. Furthermore, the user can zoom the tree by
+scrolling the mouse wheel to see more of the topology of the tree, or to see the
+details of a given part of the tree.
 
-\begin{figure}[h]
+Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
+thousands of commits and merges. While the tree is capable of producing a
+visualization, it becomes much more difficult to understand.
+
+\begin{figure}
         \centering
         \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
-        \caption{Tree zoomed out}
+        \caption{Reingold-Tilford Tree zoomed out}
         \label{fig:zoomed_tree}
@@
 
-Some of the merge trees are very large (\ref{fig:zoomed_tree}), containing
-thousands of commits and merges. Clicking on subtrees will cause them to
-minimize, only showing parts of the tree that are of interest to the user.
-Furthermore, the user can zoom the tree by scrolling the mouse wheel to see
-more of the topology of the tree, or to see the details of a given part of the
-tree.
+\subsubsection{Bubble Tree}
 
-The last tab contains another type of tree view. This tree is in the form of
-nested lists, and is designed to more closely model the tree hierarchy view of
-file browsers. This tree only contains the commits and merges that are within
-the subtrees of the current merge. A commit will never have any items in this
-tab, as it is a leaf node, so it cannot contain subtrees. To accompany the
-tree, we include breadcrumbs at the top. The last item in the breadcrumb list
-is the current commit, the next item is the parent of the current commit, and
-the first item is the top-level merge into the kernel.
+The bubble tree provides a very clear picture of where a commit is located in the
+hierarchy. The currently selected commit or merge is colored red, while non-selected
+merges use a blue color and commits use white. The depth of the blue colouring is
+dependent on how deep in the hierarchy that node is. The root is the lightest blue,
+and the contained merges are progressively darker.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
+        \caption{Bubble Tree}
+        \label{fig:bubble_tree}
+\end{figure}
+
+The bubble tree doesn't have an implicit way of providing additional information,
+placing any text near the nodes makes the tree impossible to read, so we include a
+separate pane in the web page. When a user hovers over a node, the pane shows
+additional information about the author and the commit message and a link to the
+detailed page for that commit or merge. If the user clicks on a node, the tree will
+zoom to that node and the information in the info pane will persist, enabling the
+user to click the link. This tree provides an easy mechanism for users to navigate
+from the root to the leaves and vice versa.
 
 \section{Implementation Details}
-We built a web-based tool to provide information about the kernel.
-We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
-as the framework for generating the dynamic content.
 
-Nginx is a high-performance, scalable HTTP and mail server. Started in 2002,
-the project is much younger than the Apache project, and designed with massive
-scalability and multi-processing in mind. Instead of using request-driven
-threading, Nginx uses an event-driven architecture, making it more scalable,
-proving itself as being the web server behind large sites like Netflix, Hulu,
-GitHub, Tumblr, and many other popular
-sites\footnote{\url{http://wiki.nginx.org/Main}}.
+\evan{Is this section necessary?}
 
-Postgresql is our dbms of choice, providing fast and reliable results. There
-are multiple reasons behind choosing postgresql over another dbms. The main
-reason being that the dataset built by German et al. was generated from
-a postgresql database, allowing us to import the data directly without any
-further conversions. The secondary reason is that we are more familiar with
-postgresql than the other systems available.
+We built a web-based tool to provide information about the kernel.  We chose to use
+Nginx as the main web server, postgresql as the dbms, and Flask as the framework for
+generating the dynamic content.
+
+Nginx is a high-performance, scalable HTTP and mail server. Started in 2002, the
+project is much younger than the Apache project, and designed with massive
+scalability and multi-processing in mind. Instead of using request-driven threading,
+Nginx uses an event-driven architecture, making it more scalable, proving itself as
+being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
+other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.
+
+Postgresql is our dbms of choice, providing fast and reliable results. There are
+multiple reasons behind choosing postgresql over another dbms. The main reason being
+that the dataset built by German et al. was generated from a postgresql database,
+allowing us to import the data directly without any further conversions. The
+secondary reason is that we are more familiar with postgresql than the other systems
+available.
 
 We chose the python micro-framework
-Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the
-dynamic content.  Working with flask is simple and easy, allowing us to work in
-Python.
+Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the dynamic
+content. Working with flask is simple and easy, allowing us to work in Python.
 
 The database itself is broken down into 5 main tables; commits, filesmod, logs,
-pathtomerge, and releases. Commits contains most of the metadata for each
-commit, it contains the author, the authored date, and committer, the commit
-date, a boolean saying what is a merge, and the diff. Filesmod contains the
-information on what files were modified for the commits. Since a file may be
-modified by multiple commits, and a commit may be modifying multiple files, we
-can do one of two things for the primary key. We could use both the filename
-and the commit id as the primary key, but we just use an index. The filesmod
-table contains the filename, the commit id, the number of lines added, and the
-number of lines removed. Trivially, it contains the index, but that isn't
-necessary for anything beyond the primary key. Logs contains the information
-for each commit log, this is the one-liner preview message and the full
-message. Pathtomerge is the important table. Pathtomerge contains the commit
-id, the number of merges between the commit and the top-level merge with the
-kernel, the next merge in the merge path up to the top-level merge with the
-kernel, the top-level merge into the kernel, and when it get merged into the
-kernel. The information in this table must be generated periodically by the
-server because we are unable to generate the tree structure correctly straight
-from the information stored by git. Pathtomerge does not contain any top-level
-merges, as these will never have a next merge in the tree, since we are only
-storing the parents and not the children, storing the top-level merge is
-redundant and unnecessary. The releases table stores the release information.
-It contains the version name, whether the version is a candidate release or a
-real release, the previous version, the previous non-candidate release version
-the previous version commit id, the previous non-candidate release version
-commit id, and the commit id of this version. The version commit id represents
-the last top-level merge into the version of the kernel. Using this and the
-commits table, we are able to determine the release date of any version of the
-kernel. Then using the previous release date, we can find the range of dates
-where valid top-level merges are found for that version.
+pathtomerge, and releases. Commits contains most of the metadata for each commit, it
+contains the author, the authored date, and committer, the commit date, a boolean
+saying what is a merge, and the diff. Filesmod contains the information on what
+files were modified for the commits. Since a file may be modified by multiple
+commits, and a commit may be modifying multiple files, we can do one of two things
+for the primary key. We could use both the filename and the commit id as the primary
+key, but we just use an index. The filesmod table contains the filename, the commit
+id, the number of lines added, and the number of lines removed. Trivially, it
+contains the index, but that isn't necessary for anything beyond the primary key.
+Logs contains the information for each commit log, this is the one-liner preview
+message and the full message. Pathtomerge is the important table. Pathtomerge
+contains the commit id, the number of merges between the commit and the top-level
+merge with the kernel, the next merge in the merge path up to the top-level merge
+with the kernel, the top-level merge into the kernel, and when it get merged into
+the kernel. The information in this table must be generated periodically by the
+server because we are unable to generate the tree structure correctly straight from
+the information stored by git. Pathtomerge does not contain any top-level merges, as
+these will never have a next merge in the tree, since we are only storing the
+parents and not the children, storing the top-level merge is redundant and
+unnecessary. The releases table stores the release information.  It contains the
+version name, whether the version is a candidate release or a real release, the
+previous version, the previous non-candidate release version the previous version
+commit id, the previous non-candidate release version commit id, and the commit id
+of this version. The version commit id represents the last top-level merge into the
+version of the kernel. Using this and the commits table, we are able to determine
+the release date of any version of the kernel. Then using the previous release date,
+we can find the range of dates where valid top-level merges are found for that
+version.
 
@@
 Our tool provides mechanisms and visualizations that other tools are unable to
-produce. Continuous mining of the repository\cite{German} provides an
-additional dimension to the dataset that would otherwise be lost, enabling us
-to reliably build trees from the information we have.
+produce. Continuous mining of the repository\cite{German} provides more information
+about how the repository is changing. It gives us the ability to see how commits are
+being merged into the kernel, allowing us to more reliably build trees from the
+snapshots of the repository.
 
-\begin{figure}[h]
+\begin{figure}
         \centering
@@
 
-\subsection{Performance Evaluation}
-Originally, we were performing all of our tree-navigation, file information
-gathering, and module information gathering using recursive SQL queries.
-Recursive queries have terrible performance, with the average response time
-from the database being around 3 seconds for a single operation. We found that
-it was far more response time-efficient to load an entire merge tree directly
-into memory, then iterate over the objects returned to build the tree
-structures, than it was to have the dbms perform the tree traversals, building
-the specific trees for a commit. At this time, we have re-implemented the
-breadcrumbs, the files, and the main tree views to enhance the response times.
-We will discuss some We will discuss the changes made.
+\subsection{Extendability}
 
-The breadcrumbs are generated on page-load, so they must be generated quickly.
-We began with a single recursive SQL query, but was unable get the results in
-the desired time, causing all results pages to render slowly. We replaced the
-recursive query with a single query that would grab the entire merge tree,
-rooted at the top-level merge.
+The front-end uses asynchronous requests to gather the information for the trees and
+tables. This enables third-parties to implement new front-ends, trying other
+designs, though this interface may change in the future. The tree, file, and module
+information is accessible through 
+http://li.turingmachine.org/data/tree/JSON/\verb|<commit id>|,
+http://li.turingmachine.org/data/files/JSON/\verb|<commit id>|,
+http://li.turingmachine.org/data/modules/JSON/\verb|<commit id>|, respectively.
 
-The items returned from the database contained the commit id, the commit id of
-the merge with the Linux kernel, and the next merge in the path to the
-top-level merge.
+The response for the tree is a single object. This object is the root node of the
+tree, and may contain an object of children objects in the \verb|children| field,
+which are in the same format as the root object. The children object uses the commit
+id as the key.
 
-We placed each item in the results into a dictionary using the commit id as the
-key. Using the following method, we build the dictionary into a list of commit
-IDs for the breadcrumbs.
+The tree node objects have the following fields:
 
-\begin{verbatim}
-current = cid
-breadcrumbs = [cid]
-while current is not top-level merge:
-        breadcrumbs.append(current.next)
-        current = current.next
-return reverse(breacrumbs)
-\end{verbatim}
+\begin{tabular}{ccl}
+        Field name & Data type & Description\\\hline
+        cid & string & Commit id\\
+        name & string & One-line log preview\\
+        mlinus & string & Root merge commit id\\
+        author & string & Author name and email\\
+        mnext & string & Parent merge commit id\\
+        children & object & Object of tree objects\\
+\end{tabular}
 
-This maintains the O(n) runtime for generating the breadcrumbs, while using an
-iterative method for traversing the tree and dropping the response time from
-around 3 seconds to at most 500ms.
-
-We use similar methods to generate the file information and the tree. Both are
-generated asynchronously, allowing us to use spinners to help improve the user
-experience. We are using AJAX requests to generate the asynchronous requests,
-are are having very good results.\newpage
-
-\subsection{Extendability}
-Due to how the asynchronous requests are performed, it is possible to query the
-server for the tree and file information directly, allowing users to build
-their own front-ends for working with our tree-based system.
-
-The tree and file information is accessible through
-http://li.turingmachine.org/data/tree/JSON/<commit id> and
-http://li.turingmachine.org/data/files/JSON/<commit id>, respectively.
-The tree is generated as many nested javascript objects. The tree response does
-not contain any notion of the current commit and will only return the entire
-merge tree.
-
-The tree responses are generated as a single object. This object is the root
-node of the tree, and may contain a list of children in the ``children'' field.
-The full response format for the tree looks like:
+The file responses contain only the files that the selected merge or commit works
+with. The response is a single object in the form of a tree. This object is the root
+of the tree, and represents the current merge or commit. If the current position in
+the tree is an inner node, the response will contain the child nodes in the
+``children'' field, otherwise ``children'' will be an empty list. The full response
+format for the file response looks like:
 
@@
         \item ``cid'': string, commit id
-        \item ``name'': string, One-line log preview
-        \item ``author'': string, Author's name
-        \item ``mlinus'': string, commit id of top-level merge
-        \item ``mnext'': string, commit id of the next merge on path to
-                top-level
-        \item ``children'': list, list of items with the same fields. Null if
-                no children
-\end{itemize}
-
-The file responses contain only the files that the selected merge or commit
-works with. The response is a single object in the form of a tree. This object
-is the root of the tree, and represents the current merge or commit. If the
-current position in the tree is an inner node, the response will contain the
-child nodes in the ``children'' field, otherwise ``children'' will be null. The
-full response format for the file response looks like:
-
-\begin{itemize}
-        \item ``cid'': string, commit id
-        \item ``files'': list, list of tulples in the format
+        \item ``files'': list, list of tuples in the format
                 \begin{enumerate}
@@
 \section{Future Work}
-This project was implemented over four months, interleaved with other course
-projects and assignments. There are still many areas that could use
-improvements before the full potential of our model can be realized.
-Here, we outline various areas that still need more attention.
+
+There are still many areas that can be improved before the full potential of our
+model can be realized. Here, we outline various areas that still need more
+attention.
 
 \subsection{Files}
-There is currently no functionality surrounding searching by filename. Use-case
-2 may know what file they are editing and try to determine how it may work with
-the other commits and merges in the module. It is possible that a third
-use-case may arise, where the user wants to determine all commits that effect a
-given file. In both cases, a bottom-to-top approach is used. With this
-functionality, implementing the top-to-bottom search feature becomes trivial.
 
-There is limited functionality to the presentation of the file information. At
-a minimum, the patches for the commits can be displayed. From there, the
-patches can be used to piece together parts of the file to generate a single
-patch at a given merge rather than displaying each patch individually. The
-patches can also be used for determining what kinds of changes were made, if
-the lines are being added, removed, or replaced.
+There is currently no functionality surrounding searching by filename. Use-case 2
+may know what file they are editing and try to determine how it may work with the
+other commits and merges in the module. It is possible that a third use-case may
+arise, where the user wants to determine all commits that effect a given file. In
+both cases, a bottom-to-top approach is used. With this functionality, implementing
+the top-to-bottom search feature becomes trivial.
+
+There is limited functionality to the presentation of the file information. At a
+minimum, the patches for the commits can be displayed. From there, the patches can
+be used to piece together parts of the file to generate a single patch at a given
+merge rather than displaying each patch individually. The patches can also be used
+for determining what kinds of changes were made, if the lines are being added,
+removed, or replaced.
+
+\subsection{Authorship}
+
+Our model can aggregate more information than what we have implemented. The
+authorship information is important for licensing purposes, and we can show all
+authors contributing to a merge and how many commits they contributed to the merge.
+We could go further, providing information about what files they edited, how many
+lines they added and removed, and more.
+
+\subsection{Diff generation}
+
+Due to the construction of the tree, it is possible to generate a diff of a given
+file in a merge. This would remove areas that overlap, and provide users with a
+single patch at a merge rather than attempting to combine all of the diff files from
+the leaves.
 
 \subsection{Evaluation}
-At this time, we have no evidence that our tool is able to improve the
-work-flow of maintainers. We believe that the tool is able to improve the
-work-flow and performance of maintainers because it provides cleaner mechanisms
-of visualizing and presenting the commit information. It is able to provide
-more relevant information, while removing information that is irrelevant to a
-given module or set of merges.
 
-We can either perform user-testing to show how a users workflow changes, or we
-could have maintainers evaluate and critique the tool and use their feedback to
-determine if the tool meets our goals.
+At this time, we have no evidence that our tool is able to improve the work-flow of
+maintainers. We believe that the tool is able to improve the work-flow and
+performance of maintainers because it provides cleaner mechanisms of visualizing and
+presenting the commit information. It is able to provide more relevant information,
+while removing information that is irrelevant to a given module or set of merges.
 
-\subsection{Performance}
-The modules and list tree view still use the old method of gathering the
-information for the respective view type. For single commits or very small
-merges, getting everything directly from the database and generating the pages
-server-side performs better than using the asynchronous requests, there
-isn't the overhead of multiple round trips. As the size of the merges grow,
-the performance becomes far worse than the asynchronous system and in some
-cases will time out. The asynchronous approach adds some overhead, profiled at
-roughly 500ms, regardless of the response size. The asynchronous approach has
-many benefits over building the response on the server, including spinners, a
-public API, and lower response times for larger merges. Further work can be
-done to improve the results beyond the response time of 500ms using improved
-queries and more efficient methods of building the tree.\newpage
+We can either perform user-testing to show how a users workflow changes, or we could
+have maintainers evaluate and critique the tool and use their feedback to determine
+if the tool meets our goals.
+
 \section{Conclusion}
+
 Our tool shows promising results, providing a clearer picture of the relation

### 2016-04-22 linvis.tex
Another run at the introduction
Editing the introduction again.

@@
 \newif\ifdraft
-% \drafttrue
+\drafttrue
 
@@
 commits with tools that are unable to filter and effectively visualize projects at
-the scale of the kernel. Tools like Gitk use the DAG model, showing all commits and
-merges in chronological order to when the commit was authored. This representation
-works in smaller projects, enabling users to see when changes are made, when these
-changes are merged, and how each branch is interacting. In large modular projects,
-like the Linux kernel, the DAG becomes a mess of merges and commits (figure
-\ref{fig:gitk}), losing its visual meaning. In some cases, the Linux kernel is
-simply too large for the system to generate a visualization. Github provides a DAG
-view for many projects, but is unable to display the visualization for projects as
-large as the Linux kernel (figure \ref{fig:gitfail}).
+the scale of the kernel. Tools like Gitk use a directed acyclic graph (DAG) model,
+showing all commits and merges in chronological order by when the commit was
+authored. This representation works in smaller projects; it enables users to see
+when changes are made, when these changes are merged, and how each branch is
+interacting.  In large modular projects, like the Linux kernel, the DAG becomes a
+mess of merges and commits (figure \ref{fig:gitk}) losing its visual meaning. In
+some cases, the Linux kernel is simply too large for the system to generate a
+visualization. Github provides a DAG view for many projects, but is unable to
+display the visualization for projects as large as the Linux kernel (figure
+\ref{fig:gitfail}).
 
@@
 Git cannot store the date that a commit was merged into a branch due to the rebase
-operation.  When a rebase occurs, several changes could be made to the commit. It
-may be squashed with other commits, combining the effects of several commits into a
-single commit. It may be dropped altogether, in which case the commit is no-longer
-merged.  It may be that the entire branch has simply been re-rooted to another
-position on the DAG. Due to this, tools are unable to determine when a commit is
-merged. A user may decide that it is important to be able to see only the commits
-contributing to a given release, that is commits that were merged with the
-Linus-branch of the kernel between two given dates. This operation is not possible
-without the collection of additional data that is not provided by git. Git only
-enables users to see when commit was authored, not when it was merged into a branch.
+operation. A rebase can perform serveral changes to commits and merges. It may be
+squashed with other commits, combining the effects of several commits into a single
+commit. It may be dropped altogether, in which case the commit is no-longer merged.
+It may be that the entire branch has simply been re-rooted to another position on
+the DAG. Due to this, tools are unable to determine when a commit is merged. A user
+may decide that it is important to be able to see only the commits contributing to a
+given release, that is commits that were merged with the Linus-branch of the kernel
+between two given dates. This operation is not possible without the collection of
+additional data that is not provided by git. Git only enables users to see when a
+commit was authored, not when it was merged into a branch.
 
@@
 In this paper, we present our design decisions behind a web-based
-tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based model
+tool\footnote{\url{http://li.turingmachine.org}} built around the tree-based model
 instead of a DAG. Our tool provides information about the location of a given commit
@@
 
-Our core model provides advantages to the DAG model used by other tools. By
-separating and focusing on sub-trees from the whole repository, we are able to
-filter out commits and merges that do not pertain to a component of the kernel. This
-provides a clearer picture of what is happening in a given merge. The second benefit
-to the tree model is that it enables us to automatically aggregate metadata.
-
-In addition to this, the model provides us with a mechanism of performing date-range
-queries with meaningful results. Using the tree built by a continuous mining system
-enables us a mechanism to perform date-range queries with meaningful results.
+Our core model provides advantages to the DAG model used by other tools. We separate
+the commits and merges into sub-trees and focus on a single sub-tree at a time. This
+enables us to filter out commits and merges that do not pertain to the component of
+the kernel that we are interested in, providing a clearer picture of what is
+happening in a given merge. The tree model also enables us to aggregate metadata
+easily. With the addition of the continuous polling of the repository, the model
+provides us with a mechanism of performing date-range queries with meaningful
+results.
 

### 2016-04-22 linvis.tex
Editing Related Work and Methodologies

@@
 tool\footnote{\url{http://li.turingmachine.org}} built around the tree-based model
-instead of a DAG. Our tool provides information about the location of a given commit
-in the respective merge tree, the files edited, the modules edited, and the commit
-message for a given commit or merge. The tool allows users to apply various filters,
-including the release version, along with a keyword or phrase from the log preview,
-the name of the author, or the commit id. The user can request all the root-level
-merges containing a commit or merge that matches the query, or all commits and
-merges that match the query.
+instead of the DAG. Our tool provides information about the location of a given
+commit in the respective merge tree, the files edited, the modules edited, and the
+commit message for a given commit or merge. The tool allows users to apply various
+filters, including the release version, along with a keyword or phrase from the log
+preview, the name of the author, or the commit id. The user can request all the
+root-level merges containing a commit or merge that matches the query, or all
+commits and merges that match the query.
 
@@
 
-To our knowledge, we know of no git repository visualization systems that builds a
-tree from the merge. This may be because more information is required to generate
-the trees than what is stored by git. Git uses a DAG model, which enables
-history-altering actions like rebases to occur, whereas older tools like CVS and svn
-construct trees\cite{CVS2008} but do not work with history-altering actions.
+To our knowledge, we know of no git repository visualization systems that build
+trees from the DAG provided by git. This may be because more information is required
+to generate the trees than what is stored by git. Git uses a DAG model, which
+enables history-altering actions, like rebases to occur, whereas older tools like
+CVS and svn construct trees\cite{CVS2008} but do not work with history-altering
+actions.
 
 A server must sit and monitor the git repository as it changes over time to handle
-changes to history. Many of the visualization tools are designed for a
+the changes to history. Many of the visualization tools are designed for a
 general-purpose approach, where any repository is directly usable without additional
-information. Polling for this additional information makes the application specific
+information. Polling for this additional information makes the visualizer specific
 to a repository or project.
@@
 
-Gitk, Gitg, and other similar git repository visualizing tools use the DAG as the
+Gitk, Gitg, and other similar git repository visualization tools use the DAG as the
 central organization model. This model is sufficient for smaller projects, providing
@@
 
-Furthermore, the DAG model is only able to provide file-level information on a
-commit level. It is unable to quickly aggregate the file data over all commits that
+Furthermore, the DAG model is only able to provide file-level information for
+commits. It is unable to quickly aggregate the file data over all commits that
 are part of a merge. If a user would like to view what files were edited at a merge,
@@
 tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
-pictured in figure \ref{fig:treemap}, for
-displaying the file structure of a given version of the kernel. Treemaps are good
-for visualizing the topology of shallow, wide trees, making it a good candidate for
-showing file systems. Treemaps are limited in what information can be displayed. The
-information within the treemap does not provide clear insight on what changes were
-made to the files in the given release, only providing information on where the file
-is located within the file system. File systems and the merge tree of the kernel
-share many similarities, in both cases, they represent trees that are wide and
-shallow, suggesting that a treemap could potentially work to visualize the commit
-and merge structure of the kernel.
+pictured in figure \ref{fig:treemap}, for displaying the file structure of a given
+version of the kernel. Treemaps are good for visualizing the topology of shallow,
+wide trees, making it a good candidate for displaying file systems. Treemaps are
+limited in what information can be displayed. The information within the treemap
+does not provide clear insight on what changes were made to the files in the given
+release, only providing information on where the file is located within the file
+system, that is they do no provide temporal information. File systems and the merge
+tree of the kernel share many similarities, in both cases, they represent trees that
+are wide and shallow, suggesting that a treemap could potentially work to visualize
+the commit and merge structure of the kernel.
 
@@
 
-The treemap design is also designed for working with data of differing types.
-In the filesystem, there are many kinds of files. There are text files,
-scripts, images, binary files, and other types. These can be given colours
-making differences stand out. There is no obvious way to assign meaningful
-types to the commits such that all the commits within a given merge do not all
-have the same colour.
+The treemap design is also designed for working with data of differing types. In the
+filesystem, there are many kinds of files. There are text files, scripts, images,
+binary files, and other types. These can be given colours making differences stand
+out. There is no obvious way to assign meaningful types to the commits such that all
+the commits within a given merge do not all have the same colour, although a user
+could be given a controller to colour by author or number of files edited.
 
@@
 
-The goal is to build a tool that makes navigation of the kernel commit
-information simpler, while providing an explanation of the changes occurring at
-each commit and merge. We designed the tool with two use-cases in mind, though a
-sumer may switch between the cases as they work.
+The goal is to build a tool that makes navigation of the kernel commit information
+simpler, while providing an explanation of the changes occurring at each commit and
+merge. We designed the tool with two use-cases in mind, though a user may switch
+between the cases as they work.
 
@@
 a merge and merge it directly into their current repository, which is useful for
-reducing the amount of re-implementation work. For them, it is
-important to have the ability to aggregate files and modules that are being changed
-somewhere in the merge. It is also important for them to be able to navigate from
-the root toward the largest merge that contains only the commits that modify the
-modules that they are working with.
+reducing the amount of re-implementation work. For them, it is important to have the
+ability to aggregate metadata about files and modules being effected by the merge.
+It is also important for them to be able to navigate from the root toward the
+leaves.
 
 \textit{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
-These are users that start with a known merge or commit and would like to see
-what other changes are being made in commits that are in the same merge. This is
-useful to see how changes in a given file will break other changes made in the
-surrounding commits in the same merge. This is primarily for maintainers that need
-to perform some implementation work in order to correctly merge patches from the
-future. We must provide these users a mechanism for navigating from a single commit
-toward the root, allowing them to see a bigger picture.
+These are users that start with a known merge or commit and would like to see what
+other changes are being made in commits that are in the same merge. This is useful
+to see how changes in a given file will break other changes made in the surrounding
+commits in the same merge. This is primarily for maintainers that need to perform
+some implementation work in order to correctly merge patches from the future. We
+must provide these users a mechanism for navigating from a single commit toward the
+root, allowing them to see the bigger picture.
 
@@
 children, which are other merges and commits. If a merge is cherry-picked or rebased
-there will be no children. Commits have no children and are always leaf nodes. The
-commits contains the metadata for the changes made. This metadata includes the files
-changed, the lines added and removed from each file, the author, the date the commit
-was merged into the merge that led to being merged into the kernel, and the date the
-commit was authored, the diff, and the commit log. Merges contain less metadata,
-only storing the author of the merge, the log, the commit date, and the author date.
-The details of the model are outlined by German \cite{German2015}.
+there will be no children as this information is lost when the rebase occurs.
+Commits have no children and are always leaf nodes. The commits contains the
+metadata for the changes made. This metadata includes the files changed, the lines
+added and removed from each file, the author, the date the commit was merged into
+the merge that led to being merged into the kernel, and the date the commit was
+authored, the diff, and the commit log. Merges contain less metadata, only storing
+the author of the merge, the log, the commit date, and the author date. The details
+of the model are outlined by German \cite{German2015}.
 

### 2016-04-22 linvis.tex
Editing Design section

@@
 
-In this section, we describe the progression of our tool design, showing the
-benefits and drawbacks of each component.
+We came up with two designs. We recognized that the structure of the repository was
+similar to the wide and shallow tree-structure of a file system. We began with a
+command-line shell interface. Then transitioned to a web-based tool where we looked
+beyond the file system structure. In this section, we describe the progression of
+our tool design, showing the benefits and drawbacks of each component.
 
@@
 The first tool we implemented was a small command-line shell (figure
-\ref{fig:shell}) that allows users to navigate through the commits and merges rather
-than files and directories. The shell design permitted a user to move through merges
-like they would through directories with the \verb|cd| and \verb|ls| commands. Users
-can use the \verb|cat| command to see the log preview, or the \verb|more| command to
-see the preview and how many lines were added and removed from the files in the
-commit.
+\ref{fig:shell}) that allows users to navigate through the commits and merges as one
+might navigate through files and directories. A user could navigate toward the
+leaves of the tree by using the \verb|cd| command followed by the commit id of the
+next merge. The user could then navigate toward the root by issuing \verb|cd ../|,
+as they would in a file system. Users can use \verb|ls| to view what commits and
+merges exist in the current working merge, then use \verb|cat| to view the log
+preview of a commit or \verb|more| to see the preview and how many lines were added
+and removed from the files in the commit.
 
@@
 the users with the commit hashes, which are quite long and unpleasant to remember
-type into the shell, but also do not provide any quick explanation of what the
+and type into the shell, but also do not provide any quick explanation of what the
 commit or merge may contain.
@@
 We replace the command-line interface with a web-based tool. Creating a web-based
-tool enables users to use the system without having to install software or carry a
-large database, making it more accessible, more easily maintainable, and platform
-independent. A web-based tool can utilize more interactive means of navigation and
-provide better explanations of commits and merges than the command-line interface.
+tool enables users to use the system without having to install additional software
+or carry a large database, making it more accessible, more easily maintainable, and
+platform independent. A web-based tool can utilize more interactive means of
+navigation and provide better explanations of commits and merges than the
+command-line interface.
 
@@
 
-Searching allows a user to filter commits and merges that are irrelevant.  The
-search mechanism breaks down the results by release version. A user can further
-narrow down the search range using the commit date range. The commit date refers to
-the date of the creation of the top-level merge. Commits and merges that are
-returned may have a commit date beyond the given range if they were merged to a
-top-level merge that is within the date range. A user may then provide a search
-text, filtering on the author name, the commit id, or keyword from the log. Any part
-of the author name may show up in the results, including searching by email address.
-The commit ID must be specified from the first character to the last character. For
-example, the commit `c267548755a184ef97301071300c1739a564e135' is returned when the
-user searches for a commit id of `c26' in the 3.10 Linux kernel. We will use this
-commit for showing each of the features of the tool.
+Searching allows a user to filter commits and merges that are irrelevant. The search
+mechanism breaks down the results by release version. A user can further narrow down
+the search range using the commit date range. The commit date refers to the date of
+the creation of the top-level merge. The commits and merges that are returned may
+have a commit date beyond the given range if they were merged by Linus within the
+date range. A user may then provide a search text, filtering on the author name,
+the commit id, or keyword from the log. Any part of the author name may show up in
+the results, including searching by email address. The commit ID must be specified
+from the first character to the last character. For example, the commit
+`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
+commit id of `c26' in the 3.10 Linux kernel. We will use this commit for showing
+each of the features of the tool.
 
@@
 within a merge, we simply add the added and removed lines to the values already
-stored.
+stored, though this could be extended to use the diff information to determine when
+a line has been replaced rather than incrementing the lines removed and the lines
+added.
 
@@
 can be general, such as ``bluetooth'' and ``wireless'', or can be quite specific for
-individual hardware, such as ``ath9k\_hw'' and ``wl1251''. Like files, we are able
-to show what modules are edited in merges and commits. The image in figure
-\ref{fig:modules} shows the modules for a commit. A commit will only ever have one
-module that it is associated with.
+individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a few cases, the
+author of a commit does not correctly follow this format and the hueristic approach
+fails. Like files, we are able to show what modules are edited in merges and
+commits. The image in figure \ref{fig:modules} shows the modules for a commit. A
+commit will only ever have one module that it is associated with, merges with more
+than one child will have more.
 
@@
 of the entire merge tree. We have experimented with various types of trees; list
-trees are similar to the representation by many file system explorers,
-Reingold-Tilford trees which clearly represent the tree structure, and bubble trees
-which organize the data hierarchically by having the parent node contain the child
+trees are similar to the representation used in many file system explorers,
+Reingold-Tilford trees clearly represent the tree structure, and bubble trees
+organize the data hierarchically by having the parent node contain the child nodes
+similarly to tree maps, but clearly showing the parent-child relationships between
 nodes.
@@
 
-The list tree  viewer is in the form of nested lists, and is designed to more
-closely model the tree hierarchy view found in file browsers. This tree only
-contains the commits and merges that are within the subtrees of the current merge. A
-commit will never have any items in this tree, as it is a leaf. To accompany the
-tree, we include breadcrumbs at the top of the page to enable a user to navigate
-both from the root to the leaves and from a leaf to the root. The last item in the
-breadcrumb list is the current commit, the next item is the parent of the current
-commit, and the first item is the root merge into the kernel.
+The list tree viewer is in the form of nested lists, and is designed to more closely
+model the tree hierarchy view found in file browsers. This tree only contains the
+commits and merges that are within the subtrees of the current merge. A commit will
+never have any items in this tree as it is a leaf. To accompany the tree, we include
+breadcrumbs at the top of the page to enable a user to navigate both from the root
+to the leaves and from a leaf to the root. The last item in the breadcrumb list is
+the current commit, the previous item is the parent of the current commit, and the
+first item is the root merge into the kernel.
 
@@
 hierarchy. The currently selected commit or merge is colored red, while non-selected
-merges use a blue color and commits use white. The depth of the blue colouring is
+merges use a shade of blue and commits use white. The depth of the blue colouring is
 dependent on how deep in the hierarchy that node is. The root is the lightest blue,

### 2016-04-22 linvis.tex
Editing Discussion

@@
 
-Our goal was to build a tool to enable maintainers to effectively navigate and
-pull explanation of the changes performed to the kernel over the period of a
-release. Achieving this goal includes removing information that does not
-pertain to given users. In figure \ref{fig:dag_view} and figure
-\ref{fig:tree_view}, we can visually compare the results returned from gitk and
-our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. The relevant information in both of these figures is
-identical, but the presentation drastically changes our ability to comprehend
-what we are seeing. The primary difference is the removal of irrelevant
-information. In gitk (\ref{fig:dag_view}), the visualization results include
-commits and merges from other components of the kernel, while our tool
-(\ref{fig:tree_view}) only includes the results that are specific to the
-component of the kernel that we are interested in.
+Our goal was to build a tool to enable maintainers to effectively navigate and pull
+explanation of the changes performed to the kernel over the period of a release.
+Achieving this goal includes removing information that does not pertain to the area
+of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
+figure \ref{fig:tree_view}, we can visually compare the results returned from gitk
+and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
+from kernel version 3.10. The relevant information to this commit in both of these
+figures is identical, but the presentation drastically changes our ability to
+comprehend what we are seeing. The primary difference is the removal of irrelevant
+information. In gitk (\ref{fig:dag_view}), the visualization results include commits
+and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
+only includes the results that are specific to the component of the kernel that we
+are interested in.
 
@@
 information is accessible through
-http://li.turingmachine.org/data/tree/JSON/\verb|<commit id>|,
-http://li.turingmachine.org/data/files/JSON/\verb|<commit id>|,
-http://li.turingmachine.org/data/modules/JSON/\verb|<commit id>|, respectively.
+http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
+http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
+http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.
 
@@
         Field name      & Data type & Description\\\hline
-        cid & string & Commit id\\
-        name & string & One-line log preview\\
-        mlinus & string & Root merge commit id\\
-        author & string & Author name and email\\
-        mnext & string & Parent merge commit id\\
-        children & object & Object of tree objects\\
+        \verb|cid|      & string    & Commit id\\
+        \verb|name|     & string    & One-line log preview\\
+        \verb|mlinus|   & string    & Root merge commit id\\
+        \verb|author|   & string    & Author name and email\\
+        \verb|mnext|    & string    & Parent merge commit id\\
+        \verb|children| & object    & Object of tree objects\\
 \end{tabular}
@@
 
-\begin{itemize}
-        \item ``cid'': string, commit id
-        \item ``files'': list, list of tuples in the format
-                \begin{enumerate}
-                        \item Filename: string
-                        \item Lines added: unsigned integer
-                        \item Lines removed: unsigned integer
-                \end{enumerate}
-        \item ``mnext'': string, the commit id of the next merge toward the
-                top-level merge
-        \item ``children'': list, List of all child objects with the same
-                fields. Null if no children.
-\end{itemize}
+\begin{tabular}{lll}
+        Field name & Data type & Description\\\hline
+        \verb|cid| & string & Commit id\\
+        \verb|mnext| & string & Parent merge commit id\\
+        \verb|children| & list & List of tree objects\\
+        \verb|files| & list of tuples&
+        \footnotesize{
+        \begin{tabular}{lc}
+                \verb|Filename| & String\\
+                \verb|Lines added| & uint\\
+                \verb|Lines removed| & uint\\
+        \end{tabular}}\\
+\end{tabular}
 

### 2016-04-22 linvis.tex
Editing Future Work

@@
 There are still many areas that can be improved before the full potential of our
-model can be realized. Here, we outline various areas that still need more
-attention.
+model can be realized. Here, we outline various areas of the tool that still need
+more attention.
 
@@
 
-There is currently no functionality surrounding searching by filename. Use-case 2
-may know what file they are editing and try to determine how it may work with the
-other commits and merges in the module. It is possible that a third use-case may
-arise, where the user wants to determine all commits that effect a given file. In
-both cases, a bottom-to-top approach is used. With this functionality, implementing
-the top-to-bottom search feature becomes trivial.
+There is currently no functionality surrounding searching by filename. Users falling
+under use-case 2 may know what file they are editing and try to determine how it may
+work with the other commits and merges in the module. It is possible that a third
+use-case may arise, where the user wants to determine all commits that effect a
+given file. In both cases, bottom-to-top approach is applied.
 
@@
 for determining what kinds of changes were made, if the lines are being added,
-removed, or replaced.
+removed, or being replaced.
 
@@
 authorship information is important for licensing purposes, and we can show all
-authors contributing to a merge and how many commits they contributed to the merge.
-We could go further, providing information about what files they edited, how many
-lines they added and removed, and more.
-
-\subsection{Diff generation}
-
-Due to the construction of the tree, it is possible to generate a diff of a given
-file in a merge. This would remove areas that overlap, and provide users with a
-single patch at a merge rather than attempting to combine all of the diff files from
-the leaves.
+authors contributing to commits in a  merge and how many commits they contributed to
+the merge. We could go further, providing information about what files they edited,
+how many lines they added and removed, and more.
 

### 2016-04-22 linvis.tex
Editing Conclusion

@@
 
-Our tool shows promising results, providing a clearer picture of the relation
-between the merges and commits when compared to DAG-based models. Not only are
-we provide a clearer visualization, we provide information about the files
-edited by all of the commits contained within a merge, and we provide the
-information about what modules are edited in a merge.
+Our tool shows promising results in working toward our goals, building a tool that
+can more easily navigate the repository, and provide clearer explanations of the
+changes made. The filtering by searches and the various tree visualizations enable
+users to easily navigate through the kernel, finding the commits and merges that
+pertain to what they are interested in. The tree visualizations are able to both
+serve as a navigation piece, providing simpler navigation through the kernel, but
+also demonstrate which commits contribute to which merges, providing further
+explanation over the DAG. The tree model further improves the explanation by enabling
+the tool to aggregate metadata at the merges instead of requiring a user to manually
+aggregate the information. Our tool currently aggregates file information and the
+modules edited, but can be extended to include authorship information and aggregated
+patch information.
 
-While we are able to generate a better explanation of the changes happening at
-each commit and merge, the tree-model requires more information than what is
-provided by git, requiring a central server to monitor changes to the project.
-This breaks the design goals of decentralized version control systems and makes
-the tool specific to the project.
+The tree model requires more information than what is provided by git, requiring a
+central server to monitor changes to the project. This breaks the design goals of
+decentralized version control systems and makes the visualization tool specific to
+the project.
 
-A tree-based model appears to be sufficient for creating a clean visualization
-and explanation of the changes in the Linux kernel.
+While these are significant drawbacks for small projects, large projects will likely
+have a more centralized work flow with a master repository where contributions will
+eventually land, along with various servers that can watch this master repository
+for changes. Given that the importance of clear visualizations of the repository
+outweighs the cost of having a server monitor the state of the repository, the
+tree-based model appears to provide clean visualizations and explanations of the
+changes in the Linux kernel.
 

### 2016-04-24 linvis.tex
Final editing
Making final changes to text
Final positioning of images

Adding 2016 IEEEtran.cls and bst files for proper formatting of document and
references.

@@
 \newif\ifdraft
-\drafttrue
+% \drafttrue
 
 \ifdraft
-        \documentclass[conference, draftclsnofoot]{IEEEtran}
+        \documentclass[conference, draftclsnofoot, draft]{IEEEtran}
         \def\baselinestretch{1}
@@
         \documentclass[conference]{IEEEtran}
-        \def\baselinestretch{1}
-        \setlength{\marginparwidth}{2cm}
-
-        % \documentclass[conferece, final]{IEEEtran}
 \fi
@@
 % Fix any bad-hyphenations here
-\hyphenation{}
+\hyphenation{contained}
 
@@
         Current tools provide information about repositories through the directed
-        acyclic graph (DAG), which is helpful for smaller projects. However with the
-        scale and number of branches in the kernel the DAG becomes overwhelming very
-        quickly.
+        acyclic graph (DAG), which is helpful for smaller projects. However, with
+        the scale and number of branches in the kernel the DAG becomes overwhelming
+        very quickly.
 
@@
 
+\begin{figure}
+        \centering
+        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
+        \caption{Github Failure Message showing the DAG of the Linux Project}
+        \label{fig:gitfail}
+\end{figure}
+
 Git cannot store the date that a commit was merged into a branch due to the rebase
-operation. A rebase can perform serveral changes to commits and merges. It may be
+operation. A rebase can perform several changes to commits and merges. It may be
 squashed with other commits, combining the effects of several commits into a single
@@
 filters, including the release version, along with a keyword or phrase from the log
-preview, the name of the author, or the commit id. The user can request all the
+preview, the name of the author, or the commit ID. The user can request all the
 root-level merges containing a commit or merge that matches the query, or all
@@
 
-To our knowledge, we know of no git repository visualization systems that build
+To our knowledge, we know of no git repository visualization tool that builds
 trees from the DAG provided by git. This may be because more information is required
@@
 
-A server must sit and monitor the git repository as it changes over time to handle
+A server must sit and monitor the git repository as it changes over time to record
 the changes to history. Many of the visualization tools are designed for a
@@
 working from a given commit or leaf-node toward the root. With this model, it is
-difficult for users to see what other commits are involved with a given merge.
+difficult for users to see what other commits are involved within a given merge.
 
@@
 within that merge.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project}
-        \label{fig:gitfail}
-\end{figure}
 
@@
 limited in what information can be displayed. The information within the treemap
-does not provide clear insight on what changes were made to the files in the given
-release, only providing information on where the file is located within the file
-system, that is they do no provide temporal information. File systems and the merge
+does not provide clear insight into what changes were made to the files in the given
+release. It only provides information on where the file is located within the file
+system; they do not provide temporal information. File systems and the merge
 tree of the kernel share many similarities, in both cases, they represent trees that
@@
 the commit and merge structure of the kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
+        \caption{Treemap of Linux 2.5.33 file structure}
+        \label{fig:treemap}
+\end{figure}
 
@@
 
-The treemap design is also designed for working with data of differing types. In the
+The treemap is also designed for working with data of differing types. In the
 filesystem, there are many kinds of files. There are text files, scripts, images,
@@
 out. There is no obvious way to assign meaningful types to the commits such that all
-the commits within a given merge do not all have the same colour, although a user
+the commits within a given merge do not have the same colour, although a user
 could be given a controller to colour by author or number of files edited.
@@
 %% Find paper for this maybe?
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
-        \caption{Treemap of Linux 2.5.33 file structure}
-        \label{fig:treemap}
-\end{figure}
 
@@
 the merge that led to being merged into the kernel, and the date the commit was
-authored, the diff, and the commit log. Merges contain less metadata, only storing
+authored, the patch, and the commit log. Merges contain less metadata, only storing
 the author of the merge, the log, the commit date, and the author date. The details
@@
 might navigate through files and directories. A user could navigate toward the
-leaves of the tree by using the \verb|cd| command followed by the commit id of the
-next merge. The user could then navigate toward the root by issuing \verb|cd ../|,
+leaves of the tree by using the \verb|cd| command followed by the commit ID of the
+next merge. The user could then navigate toward the root by issuing \verb|cd ../|
 as they would in a file system. Users can use \verb|ls| to view what commits and
@@
 tool enables users to use the system without having to install additional software
-or carry a large database, making it more accessible, more easily maintainable, and
+or store a large database, making it more accessible, more easily maintainable, and
 platform independent. A web-based tool can utilize more interactive means of
@@
 
-Searching allows a user to filter commits and merges that are irrelevant. The search
-mechanism breaks down the results by release version. A user can further narrow down
-the search range using the commit date range. The commit date refers to the date of
-the creation of the top-level merge. The commits and merges that are returned may
-have a commit date beyond the given range if they were merged by Linus within the
-date range. A user may then provide a search text, filtering on the author name,
-the commit id, or keyword from the log. Any part of the author name may show up in
+Searching (figure \ref{fig:search}) allows a user to filter commits and merges that
+are irrelevant. The search mechanism breaks down the results by release version. A
+user can further narrow down the search range using the commit date range. The
+commit date refers to the date of the creation of the top-level merge. The commits
+and merges returned may have a commit date beyond the queried dates; the date-range
+query refers to the date that the commit is finally merged into the kernel by Linus.
+A commit may not be merged immediately, and in some cases may take years before
+being applied. A user may then provide a search text, filtering by the author name,
+the commit ID, or keyword from the log. Any part of the author name may show up in
 the results, including searching by email address. The commit ID must be specified
@@
 `c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
-commit id of `c26' in the 3.10 Linux kernel. We will use this commit for showing
+commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
 each of the features of the tool.
@@
 
-The first tab displays the full commit log. From this, a user is able to see what
-they would see had they searched for the commit using git log. This doesn't provide
-additional information to the other tools, but helps to complete the functionality
-of our tool. The commit log provides a user with the information about the content
-of the commit and who has signed-off on the commit to ensure that it is good
-quality. The message for merges may also contain the information about the commits
-that are being merged into a given merge. The information within these messages is
-highly variable, and is completely dependent on the author's style. As the user
-moves toward the top-level merge, the quality of these messages generally improves.
+The first tab displays the full commit log (figure \ref{fig:message}). From this, a
+user is able to see what they would see had they searched for the commit using git
+log. This doesn't provide additional information to the other tools, but helps to
+complete the functionality of our tool. The commit log provides a user with the
+information about the content of the commit and who has signed-off on the commit to
+ensure that it is good quality. The message for merges may contain a summary of the
+commits being merged.  The information within these messages is highly variable, and
+is completely dependent on the author's style. As the user moves toward the
+top-level merge, the quality of these messages generally improves.
 
@@
 
-The second tab is the files tab. This tab provides information on what files have
-been edited, how many lines were added, and how many lines were removed in a given
-commit. This functionality is similar to the other tools available. Our tree-based
-design model allows us to extend this functionality to merges, which the other tools
-are unable to show. When a file is edited multiple times from different commits
-within a merge, we simply add the added and removed lines to the values already
-stored, though this could be extended to use the diff information to determine when
-a line has been replaced rather than incrementing the lines removed and the lines
-added.
+The second tab is the files tab (figure \ref{fig:files}). This tab provides
+information on what files have been edited, how many lines were added, and how many
+lines were removed in a given commit. This functionality is similar to the other
+tools available. Our tree-based design model allows us to extend this functionality
+to merges, which the other tools are unable to show. To find the number of lines
+added to a file in a merge, we take the sum of the lines added to that file in each
+of the children of that merge. We do the same for calculating the number of lines
+removed. This could be extended to use the patch information to determine when a
+line has been replaced rather than incrementing the lines added and removed.
 
@@
 
-The modules tab shows the modules that are contained within the commit. Modules are
-not natively recognized by git, and are not going to be present in all repositories.
-In the Linux repository, authors put the module they are working on in the one-line
-preview of the log-message. We heuristically extract the module by taking all text
-in the log preview of commits until the first colon.  Modules are logical partitions
-of the information in the kernel. Depending on where the author was working, modules
-can be general, such as ``bluetooth'' and ``wireless'', or can be quite specific for
-individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a few cases, the
-author of a commit does not correctly follow this format and the hueristic approach
-fails. Like files, we are able to show what modules are edited in merges and
-commits. The image in figure \ref{fig:modules} shows the modules for a commit. A
-commit will only ever have one module that it is associated with, merges with more
-than one child will have more.
+The modules tab (figure \ref{fig:modules}) shows the modules that are contained
+within the commit. Modules are not natively recognized by git, and are not going to
+be present in all repositories.  In the Linux repository, authors put the module
+they are working on in the one-line preview of the log-message. We heuristically
+extract the module by taking all text in the log preview of commits until the first
+colon.  Modules are logical partitions of the information in the kernel. Depending
+on where the author was working, modules can be general, such as ``bluetooth'' and
+``wireless'', or can be quite specific for individual hardware, such as
+``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
+correctly follow this format and the hueristic approach fails. Like files, we are
+able to show what modules are edited in merges and commits. The image in figure
+\ref{fig:modules} shows the modules for a commit.  As commits have no children, they
+will only have a single module; merges with more than one child will have more.
 
 Finally, we have the tree view tab. The tree view is designed for providing easy
-navigation within a tree and providing a clear topological view of the branch off
-the kernel. We have experimented with various tree designs to find a design that
+navigation within a tree and providing a clear topological view of the branch of the
+main kernel. We have experimented with various tree designs to find a design that
 allows for easy navigation and visualization of both large and small trees. We
@@
 
-The list tree viewer is in the form of nested lists, and is designed to more closely
-model the tree hierarchy view found in file browsers. This tree only contains the
-commits and merges that are within the subtrees of the current merge. A commit will
-never have any items in this tree as it is a leaf. To accompany the tree, we include
-breadcrumbs at the top of the page to enable a user to navigate both from the root
-to the leaves and from a leaf to the root. The last item in the breadcrumb list is
-the current commit, the previous item is the parent of the current commit, and the
-first item is the root merge into the kernel.
+The list tree viewer (figure \ref{fig:list_tree}) is in the form of nested lists,
+and is designed to more closely model the tree view found in file browsers. This
+tree only contains the commits and merges that are within the subtrees of the
+current merge. A commit will never have any items in this tree as it is a leaf. To
+accompany the tree, we include breadcrumbs at the top of the page to enable a user
+to navigate both from the root to the leaves and from a leaf to the root. The last
+item in the breadcrumb list is the current commit, the previous item is the parent
+of the current commit, and the first item is the root merge into the kernel.
 
@@
 
-The Reingold-Tilford tree produces very nice results, creating an intuitive
-representation of the tree. The representation has a clear notion of root and
-leaves, and how to navigate in either direction.
+The Reingold-Tilford tree (figure \ref{fig:tree})produces very nice results,
+creating an intuitive representation of the tree. This illustrates a clear notion of
+root and leaves, and how to navigate in either direction.
 
@@
 
-The bubble tree provides a very clear picture of where a commit is located in the
-hierarchy. The currently selected commit or merge is colored red, while non-selected
-merges use a shade of blue and commits use white. The depth of the blue colouring is
-dependent on how deep in the hierarchy that node is. The root is the lightest blue,
-and the contained merges are progressively darker.
+The bubble tree (figure \ref{fig:bubble_tree}) provides a very clear picture of
+where a commit is located in the hierarchy. The currently selected commit or merge
+is colored red, while non-selected merges use a shade of blue and commits use white.
+The depth of the blue colouring is dependent on how deep in the hierarchy that node
+is. The root is the lightest blue, and the contained merges are progressively
+darker.
 
@@
 
-\evan{Is this section necessary?}
+\evan{Is this section necessary? The tables are probably useful, but the rest... not
+sure on}
 
@@
 
-Nginx is a high-performance, scalable HTTP and mail server. Started in 2002, the
-project is much younger than the Apache project, and designed with massive
+Nginx is a high-performance, open source, scalable HTTP and mail server. Started in
+2002, the project is much younger than the Apache project, and designed with massive
 scalability and multi-processing in mind. Instead of using request-driven threading,
@@
 that the dataset built by German et al. was generated from a postgresql database,
-allowing us to import the data directly without any further conversions. The
-secondary reason is that we are more familiar with postgresql than the other systems
-available.
+allowing us to import the data directly without any further conversions.
 
@@
 The database itself is broken down into 5 main tables; commits, filesmod, logs,
-pathtomerge, and releases. Commits contains most of the metadata for each commit, it
+pathtomerge, and releases. Commits contains most of the metadata for each commit; it
 contains the author, the authored date, and committer, the commit date, a boolean
-saying what is a merge, and the diff. Filesmod contains the information on what
+saying what is a merge, and the patch. Filesmod contains the information on what
 files were modified for the commits. Since a file may be modified by multiple
 commits, and a commit may be modifying multiple files, we can do one of two things
-for the primary key. We could use both the filename and the commit id as the primary
+for the primary key. We could use both the filename and the commit ID as the primary
 key, but we just use an index. The filesmod table contains the filename, the commit
-id, the number of lines added, and the number of lines removed. Trivially, it
+ID, the number of lines added, and the number of lines removed. Trivially, it
 contains the index, but that isn't necessary for anything beyond the primary key.
-Logs contains the information for each commit log, this is the one-liner preview
-message and the full message. Pathtomerge is the important table. Pathtomerge
-contains the commit id, the number of merges between the commit and the top-level
-merge with the kernel, the next merge in the merge path up to the top-level merge
-with the kernel, the top-level merge into the kernel, and when it get merged into
-the kernel. The information in this table must be generated periodically by the
-server because we are unable to generate the tree structure correctly straight from
-the information stored by git. Pathtomerge does not contain any top-level merges, as
-these will never have a next merge in the tree, since we are only storing the
-parents and not the children, storing the top-level merge is redundant and
-unnecessary. The releases table stores the release information.  It contains the
-version name, whether the version is a candidate release or a real release, the
-previous version, the previous non-candidate release version the previous version
-commit id, the previous non-candidate release version commit id, and the commit id
-of this version. The version commit id represents the last top-level merge into the
-version of the kernel. Using this and the commits table, we are able to determine
-the release date of any version of the kernel. Then using the previous release date,
-we can find the range of dates where valid top-level merges are found for that
-version.
+Logs contains the commit log preview and full commit log. Pathtomerge is the
+important table.  Pathtomerge contains the commit ID, the number of merges between
+the commit and the top-level merge with the kernel, the next merge in the merge path
+up to the top-level merge with the kernel, the top-level merge into the kernel, and
+the date it is merged into the kernel. The information in this table must be
+generated periodically by the server because we are unable to generate the tree
+structure correctly straight from the information stored by git. Pathtomerge does
+not contain any top-level merges, as these will never have a next merge in the tree;
+since we are only storing the parents and not the children, storing the top-level
+merge is redundant and unnecessary. The releases table stores the release
+information.
+
+It contains
+\begin{itemize}
+        \item Version name
+        \item Candidate releases versus real releases
+        \item Previous release
+        \item Previous real release
+        \item Previous release commit ID
+        \item Previous real release commit ID
+        \item Current version commit ID
+\end{itemize}
+The current version commit ID represents the last top-level merge into the version
+of the kernel. Using this and the commits table, we are able to determine the
+release date of any version of the kernel. Then using the previous release date, we
+can find the range of dates where valid top-level merges are found for that version.
 
@@
 Our tool provides mechanisms and visualizations that other tools are unable to
-produce. Continuous mining of the repository\cite{German} provides more information
-about how the repository is changing. It gives us the ability to see how commits are
-being merged into the kernel, allowing us to more reliably build trees from the
-snapshots of the repository.
+produce. Continuous mining of the repository provides more information about how the
+repository is changing. It gives us the ability to see how commits are being merged
+into the kernel, allowing us to more reliably build trees from the snapshots of the
+repository.
 
@@
 Our goal was to build a tool to enable maintainers to effectively navigate and pull
-explanation of the changes performed to the kernel over the period of a release.
+an explanation of the changes performed to the kernel over the period of a release.
 Achieving this goal includes removing information that does not pertain to the area
 of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
-figure \ref{fig:tree_view}, we can visually compare the results returned from gitk
+figure \ref{fig:tree_view} we can visually compare the results returned from gitk
 and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
@@
 which are in the same format as the root object. The children object uses the commit
-id as the key.
+ID as the key.
 
@@
         Field name      & Data type & Description\\\hline
-        \verb|cid|      & string    & Commit id\\
+        \verb|cid|      & string    & Commit ID\\
         \verb|name|     & string    & One-line log preview\\
-        \verb|mlinus|   & string    & Root merge commit id\\
+        \verb|mlinus|   & string    & Root merge commit ID\\
         \verb|author|   & string    & Author name and email\\
-        \verb|mnext|    & string    & Parent merge commit id\\
+        \verb|mnext|    & string    & Parent merge commit ID\\
         \verb|children| & object    & Object of tree objects\\
@@
 the tree is an inner node, the response will contain the child nodes in the
-``children'' field, otherwise ``children'' will be an empty list. The full response
-format for the file response looks like:
+``children'' field, otherwise ``children'' will be an empty list. The format of file
+response is as follows,
 
@@
         Field name & Data type & Description\\\hline
-        \verb|cid| & string & Commit id\\
-        \verb|mnext| & string & Parent merge commit id\\
+        \verb|cid| & string & Commit ID\\
+        \verb|mnext| & string & Parent merge commit ID\\
         \verb|children| & list & List of tree objects\\
@@
         \begin{tabular}{lc}
-                \verb|Filename| & String\\
+                \verb|Filename| & string\\
                 \verb|Lines added| & uint\\
@@
 
-There is currently no functionality surrounding searching by filename. Users falling
+There currently is no functionality surrounding searching by filename. Users falling
 under use-case 2 may know what file they are editing and try to determine how it may
@@
 \nocite{*}
-\bibliographystyle{plain}
+
+\bibliographystyle{IEEEtran}
 \bibliography{citations}

### 2016-04-24 linvis.tex
Adding Daniel as co-author
modified abstract

@@
 \newcommand{\TheTitle}{Visualizing Release Information of Linux}
-\newcommand{\TheAuthors}{Evan Wilde}
-\newcommand{\TheEmails}{etcwilde@uvic.ca}
+\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
+\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
 \newcommand{\TheSubject}{Digesting large amounts of commit data}
@@
         versions of the kernel. Due to security, performance, and changing hardware
-        needs, maintainers must understand how updates made to the current version
-        of the kernel fit into older versions in order to make the necessary
+        needs, maintainers must understand what changes (commits) are added the current version of the kernel 
+        since the last time they inspected it in order to make the necessary
         patches.
@@
         Current tools provide information about repositories through the directed
-        acyclic graph (DAG), which is helpful for smaller projects. However, with
+        acyclic graph (DAG) of the repostory, which is helpful for smaller projects. However, with
         the scale and number of branches in the kernel the DAG becomes overwhelming
-        very quickly.
+        very quickly. Furthermore, the DAG contains every ancestor of every commit, while maintainers are 
+        more interested in how and when a commmit arrives to Linux official repository.
 

### 2016-04-24 linvis.tex
i have to go for lunch... back in couple of hours

@@
 
-Maintainers of older versions of the Linux kernel must sift through thousands of
+Maintainers of older versions of the Linux kernel (such as those used in embedded systems and mobile phones) must keep track of changes to the current version
+of the current in order to determine what commits are relevant to their versions.
+Due to the ratio of growth of the kernel (every year, between 50 and 70 thousand commits get added to it)
+this process requires
+sifting through thousands of
 commits with tools that are unable to filter and effectively visualize projects at
@@
 showing all commits and merges in chronological order by when the commit was
-authored. This representation works in smaller projects; it enables users to see
+authored, and not by then it arrives in the official repository.
+
+This representation works in smaller projects; it enables users to see
 when changes are made, when these changes are merged, and how each branch is
-interacting.  In large modular projects, like the Linux kernel, the DAG becomes a
+interacting.  
+In large modular projects, like the Linux kernel, the DAG becomes a
 mess of merges and commits (figure \ref{fig:gitk}) losing its visual meaning.  In
 some cases, the Linux kernel is simply too large for the system to generate a
-visualization. Github provides a DAG view for many projects, but is unable to
+visualization (eg. Github provides a DAG view for many projects, but is unable to
 display the visualization for projects as large as the Linux kernel (figure
 \ref{fig:gitfail}).
+Furthermore, Linux creates between 60 and 70k new commits every year and, according to \cite{emsePaper}, 
+a commit a median of YYYY days to arrive to the official repository from the moment it is authored. 
+This means that, if a person inspected the commits in this repository today, tomorrow it might contain  hundreds
+of commits that were authored months ago. However, distinguishing these new commits is not trivial.
 
@@
 
-Git cannot store the date that a commit was merged into a branch due to the rebase
-operation. A rebase can perform several changes to commits and merges. It may be
-squashed with other commits, combining the effects of several commits into a single
-commit. It may be dropped altogether, in which case the commit is no-longer merged.
-It may be that the entire branch has simply been re-rooted to another position on
-the DAG. Due to this, tools are unable to determine when a commit is merged. A user
-may decide that it is important to be able to see only the commits contributing to a
-given release, that is commits that were merged with the Linus-branch of the kernel
-between two given dates. This operation is not possible without the collection of
-additional data that is not provided by git. Git only enables users to see when a
-commit was authored, not when it was merged into a branch.
+One of the challenges with visualzing the arrival of commits to a repository is that
+in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master)
+To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
+would prefer to know the path that a commit followed to reach the master repository. \dmg{perhaps a figure is good here}.
+In the context of Linux tracing the path that a commit followed to master would imply that for any given merge, it would be possible to know what commits were merged.
+With such a feature,  a user
+could inspect the commits that have arrived into master within a given time frame.
+ and decide that it is important.
+
+\dmg{I am going to describe the method we used here... }
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2016-04-24 linvis.tex
small commit, more needs to be done

@@
 
-\dmg{I am going to describe the method we used here... }
+% \dmg{I am going to describe the method we used here... }
 
-German et al. recorded the metadata of every commit and merge in the Linux git
-repository starting in 2011 \cite{German2015}, including chronological data that is
-not stored by git. We are able to use this information to determine when a commit
-has been merged into the kernel, regardless of rebases preceding and proceeding that
-point in time. This information is then converted by a heuristic method, described
-by German, to generate sets of trees, which we use as our core model instead of
-using the DAG. Each tree is rooted by the top-level merge performed by Linus, which
-merges the branches or a set of commits into the kernel.
+% German et al. recorded the metadata of every commit and merge in the Linux git
+% repository starting in 2011 \cite{German2015}, including chronological data that is
+% not stored by git. We are able to use this information to determine when a commit
+% has been merged into the kernel, regardless of rebases preceding and proceeding that
+% point in time. This information is then converted by a heuristic method, described
+% by German, to generate sets of trees, which we use as our core model instead of
+% using the DAG. Each tree is rooted by the top-level merge performed by Linus, which
+% merges the branches or a set of commits into the kernel.
 
-In this paper, we present our design decisions behind a web-based
-tool\footnote{\url{http://li.turingmachine.org}} built around the tree-based model
-instead of the DAG. Our tool provides information about the location of a given
+ The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the tree model of the
+ reposisory) that represents the path used by
+ any commit to reach master. Second, 
+ we present a method to inspect and visualize the history of merges in the  Linux project using the tree-model of the repository
+
+These methods and visualizations have been implemented in a web-based
+tool\footnote{It is currently available at \url{http://li.turingmachine.org}}
+instead of the DAG. Our visualizations and tool provide information about the location of any given
 commit in the respective merge tree, the files edited, the modules edited, and the
-commit message for a given commit or merge. The tool allows users to apply various
+commit message for a given commit or merge. Our tool allows users to apply various
 filters, including the release version, along with a keyword or phrase from the log
@@
 commits and merges that match the query.
+
+\evantodo{move to discussion}
 

### 2016-04-25 linvis.tex
added tree model to the paper

@@
 \ref{fig:gitfail}).
-Furthermore, Linux creates between 60 and 70k new commits every year and, according to \cite{emsePaper}, 
-a commit a median of YYYY days to arrive to the official repository from the moment it is authored. 
-This means that, if a person inspected the commits in this repository today, tomorrow it might contain  hundreds
-of commits that were authored months ago. However, distinguishing these new commits is not trivial.
+Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}, 
+a commit takes a median of 30 days (from the moment it is authored) to the time it arrives to the official repository
+This means that, if a person inspected the commits in this repository today, tomorrow it might contain  many more
+of commits that were authored long before. However, distinguishing these new commits is not trivial.
 
@@
 
-One of the challenges with visualzing the arrival of commits to a repository is that
-in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master)
+One of the challenges with visualizing the arrival of commits to a repository is that
+in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master).
 To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
 would prefer to know the path that a commit followed to reach the master repository. \dmg{perhaps a figure is good here}.
-In the context of Linux tracing the path that a commit followed to master would imply that for any given merge, it would be possible to know what commits were merged.
-With such a feature,  a user
-could inspect the commits that have arrived into master within a given time frame.
- and decide that it is important.
+Hence, tracing the path that any commit followed to master would imply that for any given merge, it would be possible to know what commits were merged. A user could
+inspect the commits that have arrived into master within a given time frame by simply checking which commits where merged during that time frame.
 
@@
 
- The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the tree model of the
- reposisory) that represents the path used by
+ The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{tree-model of the
+ repository}) that represents the path used by
  any commit to reach master. Second, 
@@
 
+\section{Data model}
+
+Git uses a Directed Acyclic Graph (DAG) as its main data model. In this model, a commit has one or more ordered parents (ancestors), except the root commits of a repository which
+do not have ancestors--Linux git repository has two such commits. Commits are divided into merge commits (they merge two or more branches) and non-merge
+commits. Non-merge commits have only once parent, while merge commits have two or more. The order of the parents matter: the first parent is the branch in which
+the merge is being done; while the rest indicate the branches being merged (git allows any number of branches to be merged at the same time). Their order in the
+parents relationship indicate the order in which the branches are being merged (second parent first, third parent second, etc.).
+
+In Git commits are immutable. Once they are created, they never change. Git allows operations to alter commits or to reorder them. However, in such case, the
+commit changes its commit id, effectively becoming a new commit. This means that commits do not record the branches or repositories they have traversed in their
+way to master. 
+For instance, assume the sequence of commits  exemplified in Figure~\ref{fig:repoEvents}. In this case commits are performed in various repos and branches. The
+DAG representation of these commits is shown in Figure~\ref{fig:repoDAG}. Notice that this DAG loses information about what repository is Master (in fact, it
+loses all information about what repositories any commit has passed through). The tree-model of this DAG is shown in Figure~
+\ref{fig:reposTree}. Note that the direction edges of the DAG have been inverted. Instead of pointing to the ancestors, it points to its successors (the path to
+Master). Also, the DAG has been simplified, and it only shows for any commit, one edge (its path to Master).
+
+\subsection{Computing the tree-model of the DAG of Linux}
+
+Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it possible to
+compute the its tree model. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
+assertion if previous research~\cite{German2015}.
+
+\dmg{I'll continue after dinner}
+
+\evanTodo{Move related work to the end, try to match the text of the paper to my intro}
+
+\begin{figure}[htbp]
+  \centering
+        \includegraphics[width=\columnwidth]{figures/events.pdf}
+  \caption{Example of a sequence of events performed in different repositories. The horizontal axis represents time. Each horizontal section represents a
+    different branch and/or repository. Each commit points to its ancestor.}
+  \label{fig:repoEvents}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+        \includegraphics[width=.8\columnwidth]{figures/dag.pdf}
+  \caption{DAG representation of the commits represented in Figure~\ref{fig:repoEvents}. It loses information about what repository the commit is performed and
+    through which repositories it has passed in its way to Master. In fact, the DAG does not even know what Master is.}
+  \label{fig:repoDAG}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+        \includegraphics[width=\columnwidth]{figures/tree.pdf}
+  \caption{Tree-model of the commits  represented in Figure~\ref{fig:repoEvents} showing the path they followed to reach master. In this model the successors of
+    each commit represents the path followed by that commit to reach Master.}
+  \label{fig:repoTree}
+\end{figure}
+
 \section{Related Work}
@@
 To our knowledge, we know of no git repository visualization tool that builds
-trees from the DAG provided by git. This may be because more information is required
-to generate the trees than what is stored by git. Git uses a DAG model, which
-enables history-altering actions, like rebases to occur, whereas older tools like
-CVS and svn construct trees\cite{CVS2008} but do not work with history-altering
-actions.
+a tree that maps the path commits follow to master
+from the DAG provided by git. This may be because more information is required
+to generate the tree-model than what is stored by git. % Git uses a DAG model, which
+% enables history-altering actions, like rebases to occur, whereas older tools like
+% CVS and svn construct trees\cite{CVS2008} but do not work with history-altering
+% actions.
 
-A server must sit and monitor the git repository as it changes over time to record
-the changes to history. Many of the visualization tools are designed for a
-general-purpose approach, where any repository is directly usable without additional
-information. Polling for this additional information makes the visualizer specific
+% A server must sit and monitor the git repository as it changes over time to record
+% the changes to history. Many of the visualization tools are designed for a
+% general-purpose approach, where any repository is directly usable without additional
+% information. Polling for this additional information makes the visualizer specific
 to a repository or project.
@@
 can interact with the DAG by going to the previous commit in the chain or jumping to
-the parent merge. This suggests that Gitk and similar tools are designed for users
-working from a given commit or leaf-node toward the root. With this model, it is
-difficult for users to see what other commits are involved within a given merge.
+the parent merge. Given the direction that each node in the DAG has (pointing to its parent)
+it can be assumed that tools that exploit the DAG (such as gitk and similar tools) are designed for users
+working from a given commit towards the past. Working in the opposite direction is hard, since merge commits provide multiple paths towards the present, with no
+way to distinguish which path is the one used by the current commit to be merged into master.
 

### 2016-04-25 linvis.tex
many changes

@@
 \newif\ifdraft
-% \drafttrue
+%\drafttrue
 
@@
 \usepackage[nospace]{cite}
+
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \ref{fig:gitfail}).
-Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}, 
+Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}pdfl, 
 a commit takes a median of 30 days (from the moment it is authored) to the time it arrives to the official repository
@@
 
-\section{Data model}
+\section{Merge-Tree model}
 
@@
 do not have ancestors--Linux git repository has two such commits. Commits are divided into merge commits (they merge two or more branches) and non-merge
-commits. Non-merge commits have only once parent, while merge commits have two or more. The order of the parents matter: the first parent is the branch in which
+commits. Non-merge commits have only one parent, while merge commits have two or more. The order of the parents matter: the first parent is the branch in which
 the merge is being done; while the rest indicate the branches being merged (git allows any number of branches to be merged at the same time). Their order in the
@@
 
-In Git commits are immutable. Once they are created, they never change. Git allows operations to alter commits or to reorder them. However, in such case, the
+In Git, commits are immutable. Once they are created, they never change. Git allows operations to alter commits or to reorder them. However, in such case, the
 commit changes its commit id, effectively becoming a new commit. This means that commits do not record the branches or repositories they have traversed in their
@@
 DAG representation of these commits is shown in Figure~\ref{fig:repoDAG}. Notice that this DAG loses information about what repository is Master (in fact, it
-loses all information about what repositories any commit has passed through). The tree-model of this DAG is shown in Figure~
-\ref{fig:reposTree}. Note that the direction edges of the DAG have been inverted. Instead of pointing to the ancestors, it points to its successors (the path to
+loses all information about what repositories any commit has passed through). The merge-tree view of this DAG is shown in Figure~
+\ref{fig:repoTree}. Note that the direction of the edges of the DAG has been inverted. Instead of pointing to the ancestors, it points to its successors (the path to
 Master). Also, the DAG has been simplified, and it only shows for any commit, one edge (its path to Master).
-
-\subsection{Computing the tree-model of the DAG of Linux}
-
-Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it possible to
-compute the its tree model. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
-assertion if previous research~\cite{German2015}.
-
-\dmg{I'll continue after dinner}
-
-\evanTodo{Move related work to the end, try to match the text of the paper to my intro}
 
@@
         \includegraphics[width=\columnwidth]{figures/tree.pdf}
-  \caption{Tree-model of the commits  represented in Figure~\ref{fig:repoEvents} showing the path they followed to reach master. In this model the successors of
+  \caption{Merge-tree view of the commits  represented in Figure~\ref{fig:repoEvents} showing the path they followed to reach master. In this model the successors of
     each commit represents the path followed by that commit to reach Master.}
@@
 
+\subsection{Computing the tree-model of the DAG of Linux}
+
+Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it feasible to
+compute the its merge-tree model. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
+assertion if previous research~\cite{German2015}. We have developed a heuristic that is presented in Algorithm~\ref{fig:alg}. In a nutshell, first identify the commits directly made
+to the master branch. Then,
+recursively, for
+every commit, using the DAG, determine the shortest path (in terms of time) from such commit to master using the inverted DAG.
+
+\begin{algorithm}
+\caption{Computing the tree-model of Linux Git's DAG}\label{fig:alg}
+\begin{algorithmic}
+\Function{ComputeTreeModel}{DAG}: tree 
+\State {\# Compute the tree from the DAG of a git repository.}
+\State {\#  Returns $Tree$, a graph containing every commit }
+\State {\# in DAG with the path it followed to master.\\}
+
+\State $head \gets \textit{Head of master of git repository}$
+\State $master \gets \textit{traverse DAG from head using }$
+\State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
+\State $nodes(Tree) \gets nodes(DAG)$
+\State \Function{distance2Master}{cid} : seconds 
+\State {\# Compute the shortest distance to master setting}
+\State {\# cid's successor (next) in its way to master.}
+\State {\# This function should be memoized. Otherwise it}
+\State {\# would run in exponential time.}
+\If {\textit{cid in master}} 
+\State \Return 0
+\EndIf
+\State    $d \gets 	\infty$
+   \State {\# Traverse the inverted DAG}
+   \For{$c \in children(cid, DAG)$} 
+      \If {$c \in master$} 
+      \State $d_1 \gets commitTime(c)-commitTime(cid)$
+      \Else
+      \State {$d_1 \gets distance2Master(c)$} 
+      \EndIf
+    \If {$d_1 < d $} 
+    \State $next \gets c$
+   \State  $d \gets d_1$
+   \EndIf
+   \EndFor
+   \State {\# $c$ is the commit that follows $cid$}
+   \State {\# in its way to master}
+   \State add edge $(cid, next)$ to $Tree$
+   \State \Return $d$ 
+\EndFunction
+
+\State {\# Compute the distance for each commit}
+\State {\# discarding result}
+\For{$c \in nodes(DAG)$}
+\State $distance2Master(c)$
+\EndFor
+\State \Return $Tree$
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Evaluation}
+
+Merges that do not have conflicts provide information to verify this heuristic. If a merge does not contain a conflict, it records a summary of the commits that
+it merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a list of the first 20 non-merge commits in the merge, including their one-line
+log description, the full logs of the merge commits that merge this subset, and the total number of non-merge commits in the merge.
+
+\begin{figure}[htbp]
+  \centering
+{\tiny
+\begin{verbatim}
+Merge: 8cbd84f fd8aa2c
+Author: Linus Torvalds <torvalds@linux-foundation.org>
+Date:   Tue Aug 10 15:38:19 2010 -0700
+
+    Merge branch 'for-linus' of git://neil.brown.name/md
+
+    * 'for-linus' of git://neil.brown.name/md: (24 commits)
+      md: clean up do_md_stop
+[... edited for the sake of space]
+      md: split out md_rdev_init
+      md: be more careful setting MD_CHANGE_CLEAN
+      md/raid5: ensure we create a unique name for kmem_cache when mddev has no gendisk
+      ...
+\end{verbatim}}
+  \caption{Example of how merges record a subset of commits being merged. The commit only shows the first 20 one line log messages for the 24 non-merge commits
+    it merged. The ending ... is part of the log and represents that other commits were merged.}
+  \label{fig:sampleMerge}
+\end{figure}
+
+We use this information to evaluate the accuracy of the merge tree-model extracted from the DAG. We therefore compare, for commits with no conflict: a) the number of
+non-merge commits, b) the summaries of the commits available in the log, and c) the summaries of the merges included.
+
+The results of this verification show that this heuristic detected no incorrect merges since 2013, two incorrect merges in 2012, and two in 2011. However, due
+to fast-forward commits performed in Linus Torvalds repository, this heuristic breaks apart on Oct 15, 2009, and no longer works for older merges.
+
 \section{Related Work}
+
+\evantodo{Move related work to the end, try to match the text of the paper to my intro}
 
@@
 
-\section{Methodologies}
-\evan{Correct section title? Other than just trying stuff, what were the
-methodologies?}
+\section{Visualizing the merge-tree of Linux}
 
-The goal is to build a tool that makes navigation of the kernel commit information
-simpler, while providing an explanation of the changes occurring at each commit and
-merge. We designed the tool with two use-cases in mind, though a user may switch
+The goal is to build a tool that simplifies  navigation of the kernel commit information, while providing an explanation of the changes occurring at each commit
+and, more importantly, indicates how commits are merged in their way to master.
+We designed the tool with two use-cases in mind, though a user may switch
 between the cases as they work.
 
-\textit{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
+\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
 These are users that are maintaining a section of the kernel and would like to pick
-a merge and merge it directly into their current repository, which is useful for
-reducing the amount of re-implementation work. For them, it is important to have the
+a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
+reducing the amount of re-implementation work. For these users, it is important to have the
 ability to aggregate metadata about files and modules being effected by the merge.
-It is also important for them to be able to navigate from the root toward the
+\dmg{what is root?, what is leaves? perhaps remove unless you can be specific}
+It is also important for these users to be able to navigate from the root toward the
 leaves.
 
-\textit{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
+\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
 These are users that start with a known merge or commit and would like to see what
 other changes are being made in commits that are in the same merge. This is useful
-to see how changes in a given file will break other changes made in the surrounding
-commits in the same merge. This is primarily for maintainers that need to perform
-some implementation work in order to correctly merge patches from the future. We
+to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
+ This is primarily for maintainers that need to perform
+some specific cherry picking of commits. We
 must provide these users a mechanism for navigating from a single commit toward the
-root, allowing them to see the bigger picture.
+master, allowing them to see the bigger picture.
 
@@
 
-We root the tree with the merge made by Linus. Each merge may contain zero or more
-children, which are other merges and commits. If a merge is cherry-picked or rebased
-there will be no children as this information is lost when the rebase occurs.
-Commits have no children and are always leaf nodes. The commits contains the
+In our visualizations, each commit is either made directly to Master, or is part of a Merge-tree, with its root in Master (a merge made by Linus Torvalds). Each merge may contain zero or more
+children, which are other merges and commits.
+
+To improve the visualization of the merge-tree, we transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master. In
+this model, commits have no children and are always leaf nodes. This allows the clustering of commits along their next merge, and avoid the space needed to
+present long sequences of commits without merges in between.
+
+Commits contain the
 metadata for the changes made. This metadata includes the files changed, the lines
@@
 the author of the merge, the log, the commit date, and the author date. The details
-of the model are outlined by German \cite{German2015}.
+of the model are outlined in \cite{German2015}.
 

### 2016-04-25 linvis.tex
minor changes

@@
 These methods and visualizations have been implemented in a web-based
-tool\footnote{It is currently available at \url{http://li.turingmachine.org}}
+tool\footnote{Our visualization tool is currently available at \url{http://li.turingmachine.org}}
 instead of the DAG. Our visualizations and tool provide information about the location of any given
@@
 
-\subsection{Computing the tree-model of the DAG of Linux}
+\subsection{Computing the merge-tree of the DAG of Linux}
 
@@
 \begin{algorithm}
-\caption{Computing the tree-model of Linux Git's DAG}\label{fig:alg}
+\caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
 \begin{algorithmic}
@@
 \section{Design}
+
+\dmg{I would remove the command line shell, and use the space for related work}
 

### 2016-04-25 linvis.tex
manymore changes

@@
 
-\newcommand{\TheTitle}{Visualizing Release Information of Linux}
+\newcommand{\TheTitle}{Merge-tree: Visualizing the integration of commits into Linux}
 \newcommand{\TheAuthors}{Evan Wilde, Daniel German}
@@
 
-        In this paper, we present a tool that uses a tree-based model, as described
-        by German et al., to construct a clearer summary of the changes made. This
-        tool is designed to allow users to explore the topology of the repository
-        for a given release of the kernel and view an aggregated summary of the
-        changes made.
+        In this paper, we present \tool, a tool that uses a tree-based model to construct a clearer summary of the changes made. \tool is designed to allow
+        users to explore the topology of the repository for a given release of the kernel and view an aggregated summary of the changes made.
 
@@
 \ref{fig:gitfail}).
-Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}pdfl, 
-a commit takes a median of 30 days (from the moment it is authored) to the time it arrives to the official repository
+Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}, 
+a commit takes a median of 30 days from the moment it is authored to the time it arrives to the official repository
 This means that, if a person inspected the commits in this repository today, tomorrow it might contain  many more
-of commits that were authored long before. However, distinguishing these new commits is not trivial.
+commits that were authored long before. However, distinguishing these new commits is not trivial.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/gitk.png}
-        \caption{A sample DAG view of commits from Linux in Gitk}
+        \caption{A sample DAG view of commits from Linux in gitk. Commits point towards their ancestors, and there is no clear indication of the way the travel
+          to reach Master. In fact, neither gitk nor git is capable of showing what commits were done in master.}
         \label{fig:gitk}
@@
         \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project}
+        \caption{Github Failure Message showing the DAG of the Linux Project and its relationship to other forks.}
         \label{fig:gitfail}
@@
 To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
-would prefer to know the path that a commit followed to reach the master repository. \dmg{perhaps a figure is good here}.
+would prefer to know the path that a commit followed to reach the master repository.
 Hence, tracing the path that any commit followed to master would imply that for any given merge, it would be possible to know what commits were merged. A user could
@@
 
- The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{tree-model of the
+ The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{merge-tree of the
  repository}) that represents the path used by
@@
 These methods and visualizations have been implemented in a web-based
-tool\footnote{Our visualization tool is currently available at \url{http://li.turingmachine.org}}
-instead of the DAG. Our visualizations and tool provide information about the location of any given
-commit in the respective merge tree, the files edited, the modules edited, and the
-commit message for a given commit or merge. Our tool allows users to apply various
+tool called \tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}. Our visualizations and tool provide information about the location of any given
+commit in its respective merge-tree, the files edited, the modules edited, and the
+commit message for a given commit or merge. \tool allows users to apply various
 filters, including the release version, along with a keyword or phrase from the log
@@
 
-\evantodo{move to discussion}
-
-Using the tree provides many advantages to the DAG model, but there are two
-drawbacks. The first drawback is that a server must query the state of the master
-repository since time information is not stored by git. The second drawback is that
-the tree construction is a heuristic approach, and in some cases may break.
-
-Our core model provides advantages to the DAG model used by other tools. We separate
-the commits and merges into sub-trees and focus on a single sub-tree at a time. This
-enables us to filter out commits and merges that do not pertain to the component of
-the kernel that we are interested in, providing a clearer picture of what is
-happening in a given merge. The tree model also enables us to aggregate metadata
-easily. With the addition of the continuous polling of the repository, the model
-provides us with a mechanism of performing date-range queries with meaningful
-results.
-
 \section{Merge-Tree model}
+\label{sec:mergetree}
 
@@
 Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it feasible to
-compute the its merge-tree model. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
+compute its merge-tree. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
 assertion if previous research~\cite{German2015}. We have developed a heuristic that is presented in Algorithm~\ref{fig:alg}. In a nutshell, first identify the commits directly made
@@
 \begin{algorithmic}
-\Function{ComputeTreeModel}{DAG}: tree 
-\State {\# Compute the tree from the DAG of a git repository.}
+\Function{ComputeMergeTree}{DAG}: tree 
+\State {\# Compute the tree from the DAG of Linux repository.}
 \State {\#  Returns $Tree$, a graph containing every commit }
@@
 \State \Function{distance2Master}{cid} : seconds 
-\State {\# Compute the shortest distance to master setting}
-\State {\# cid's successor (next) in its way to master.}
+\State {\# Helper function}
+\State {\# Recursively compute shortest distance to master}
+\State {\#  setting cid's successor (next) in its way to master.}
 \State {\# This function should be memoized. Otherwise it}
@@
 to fast-forward commits performed in Linus Torvalds repository, this heuristic breaks apart on Oct 15, 2009, and no longer works for older merges.
+
+\section{Visualizing the merge-tree of Linux}
+
+The goal of \tool was to simplify the  navigation of the kernel commit information, specifically by focusing on merges.
+This is done by leveraging the merge-tree view of Linux to inspect how commits are merged in their way to master.
+
+\subsection{Use cases}
+
+We designed \tool with two use-cases in mind, though a user may switch
+between the cases as they work.
+
+\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
+These are users that are maintaining a section of the kernel and would like to pick
+a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
+reducing the amount of re-implementation work. For these users, it is important to have the
+ability to aggregate metadata about files and modules being effected by the merge.
+It is also important for these users to be able to navigate from the root of the merge-tree toward the
+leaves.
+
+\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
+These are users that start with a known merge or commit and would like to see what
+other changes are being made in commits that are in the same merge, including knowing the merge-tree they belong to. This is useful
+to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
+ This is primarily for maintainers that need to perform
+some specific cherry picking of commits. We
+must provide these users a mechanism for navigating from a single commit toward the
+master, allowing them to see other commits that might be related to it.
+
+\subsection{Data Model}
+
+In our visualizations, we leverage the Merge-Tree model described in Section~\ref{sec:mergetree}.
+In this model each commit is either already in Master, or is part of a tree which is rooted in the merge that merged it into Master.
+In that case, each
+commit (whether merge or non-merge) has only one successor at most (the root of each tree has none, as it was made by Linus Torvalds directly into Master).
+Commits contain the
+metadata for the changes made. This metadata includes the files changed, the lines
+added and removed from each file, the author, the date the commit was merged into
+the merge that led to being merged into the kernel, and the date the commit was
+authored, the patch, and the commit log. Merges contain less metadata, only storing
+the author of the merge, the log, the commit date, and the author date, and potentially, changes necessary to address conflicts during the merge. The details
+of the model are outlined in \cite{German2015}.
+
+\section{Design and Implementation}
+
+% \dmg{I would remove the command line shell, and use the space for related work}
+
+% We came up with two designs. We recognized that the structure of the repository was
+% similar to the wide and shallow tree-structure of a file system. We began with a
+% command-line shell interface. Then transitioned to a web-based tool where we looked
+% beyond the file system structure. In this section, we describe the progression of
+% our tool design, showing the benefits and drawbacks of each component.
+
+% \subsection{Command-line Shell}
+
+% Our initial intuition was to work with the commits and merges as directory trees
+% where the commits are files and the merges are directories. There are many
+% similarities between the tree structure of a file system and the tree structure of
+% the repository. Both trees are generally shallow, but very wide. File systems have
+% been studied for quite some time and there are many tools capable of visualizing and
+% working with the file structure.
+
+% The first tool we implemented was a small command-line shell (figure
+% \ref{fig:shell}) that allows users to navigate through the commits and merges as one
+% might navigate through files and directories. A user could navigate toward the
+% leaves of the tree by using the \verb|cd| command followed by the commit ID of the
+% next merge. The user could then navigate toward the root by issuing \verb|cd ../|
+% as they would in a file system. Users can use \verb|ls| to view what commits and
+% merges exist in the current working merge, then use \verb|cat| to view the log
+% preview of a commit or \verb|more| to see the preview and how many lines were added
+% and removed from the files in the commit.
+
+% \begin{figure}
+%         \centering
+%         \includegraphics[width=0.47\textwidth]{figures/shell.png}
+%         \caption{Shell Design}
+%         \label{fig:shell}
+% \end{figure}
+
+% The shell does not meet our design goals. It only provides users with the ability to
+% navigate from the root. If there is a commit or merge that they would like to
+% investigate, they must find it with another tool and navigate through the tree to
+% find it. Furthermore, it does not aggregate information in the merges, so users must
+% manually aggregate the metadata. Another issue with this design is that it presents
+% the users with the commit hashes, which are quite long and unpleasant to remember
+% and type into the shell, but also do not provide any quick explanation of what the
+% commit or merge may contain.
+
+% \subsection{Web-based Tool}
+
+To navigate and inspect the merge-tree view of the Kernel we
+created a web-based tool called \tool. Creating a web-based
+tool enables users to use the system without having to install additional software
+or store a large database, making it more accessible, more easily maintainable, and
+platform independent. 
+% A web-based tool can leverage interactive means of
+% navigation and provide better explanations of commits and merges than the
+% command-line interface.
+\tool uses the following mechanisms to reach our goals of better navigation and
+better explanation of the changes.
+
+\begin{itemize}
+        \item Filter by searching commits by when they are merged into Master.
+        \item View files edited
+        \item View modules
+        \item Tree viewer
+\end{itemize}
+
+\subsection{Searching by when they are merged into Master}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/search.png}
+        \caption{Search View. It allows finding commits that were \textbf{merged} in a given period or by commit it.}
+        \label{fig:search}
+\end{figure}
+
+Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and merges that are irrelevant. The search mechanism breaks down the results by
+release version.  A user can further narrow down the search by \textbf{specifying a range of dates} in which such commits \textbf{were merged by Linus into
+  Master, not
+  when the commits were created.} This distinction is important. We have observed commits that take years
+to arrive into Master after they were originally committed.
+
+A user may then provide a search text, filtering by the author name,
+the commit ID, or keyword from the log. Any part of the author name may show up in
+the results, including searching by email address.
+
+\dmg{What is this commit? any commit id, or a merge into Master? Unclear.}
+If providing commit ID must be specified
+from the first character to the last character. For example, the commit
+`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
+commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
+each of the features of \tool.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
+        \caption{Search Results. Each is a commit merged in the desired merged window.}
+        \label{fig:results}
+\end{figure}
+
+In the search results, the user is presented with the one-line log message preview,
+the author's name and email, the date the commit was authored, and the date the
+commit was last committed. The commit date and author date will usually be the same;
+however, if the commit has been rebased or cherry-picked into a merge that
+eventually is merged into the kernel, the commit date and authored date will be
+different. We can see this in the first and third entries in figure
+\ref{fig:results}.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
+        \caption{Panel showing the Commit Message of a commit.}
+        \label{fig:message}
+\end{figure}
+
+Once a user selects a commit or merge to investigate, they are presented with a
+tabbed pane allowing them to view the full commit log, the files edited, the modules
+involved, and the merge-tree view.
+
+The first tab displays the full commit log (figure \ref{fig:message}). From this, a
+user is able to see what they would see had they searched for the commit using git
+log. This doesn't provide additional information to the other tools, but helps to
+complete the functionality of \tool. The commit log provides a user with the
+information about the content of the commit and who has signed-off on the commit to
+ensure that it is good quality. The message for merges may contain a summary of the
+commits being merged.  The information within these messages is highly variable, and
+is completely dependent on the author's style. As the user moves toward the
+top-level merge, the quality of these messages generally improves.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
+        \caption{Panels showing the files modified by all the commits that are part of this merge.}
+        \label{fig:files}
+\end{figure}
+
+The second tab is the files tab (figure \ref{fig:files}). This tab provides information on what files have been edited, how many lines were added, and how many
+lines were removed in a given commit. For non-merges, this functionality is similar to the other tools available. Our tree-based design model allows us to
+extend this functionality to merges by aggregating information about all the commits that are children of the merge in the merge-tree (which the other tools are
+unable to show). To find the number of lines added to a file in a merge, we take the sum of the lines added to that file in each of the children of that
+merge. We do the same for calculating the number of lines removed. 
+% This could be extended to use the patch information to determine when a line has been
+% replaced rather than incrementing the lines added and removed.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
+        \caption{Panel showing the modules changed by all the commits in this merge.}
+        \label{fig:modules}
+\end{figure}
+
+The modules tab (figure \ref{fig:modules}) shows the modules that are contained
+within the commit. Modules are not natively recognized by git, and are not going to
+be present in all repositories.  In the Linux repository, authors put the module
+they are working on in the one-line preview of the log-message; for example: the log \emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module
+(the coverage testing tool of the kernel). We heuristically
+extract the module by taking all text in the log preview of commits until the first
+colon.  Modules are logical partitions of the information in the kernel. Depending
+on where the author was working, modules can be general, such as ``bluetooth'' and
+``wireless'', or can be quite specific for individual hardware, such as
+``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
+correctly follow this format and the hueristic approach fails. As with the Files panels,
+non-merge commits show their corresponding information, but for merge commits we aggregate all the modules changed
+in all the commits that are part of that merge.
+The output of this view is shown in Figure~\ref{fig:modules}.
+
+Finally, we have the Tree view tab. The tree view is designed for providing easy
+navigation of the commits within the merge-tree that is rooted in the current merge.
+It also provides a clear topological view of the merge and the submerges it includes.We have experimented with various tree designs to find a design that
+allows for easy navigation and visualization of both large and small trees.  We discuss this panel in the next subsection.
+
+\subsection{Merge-Tree Views} \label{treeview_section}
+
+The merge-tree view is what makes \tool unique to other tools that inspect the DAG of a git repository. With it, a user can inspect how commits are merged
+on their way to Master. We have experimented with various types of trees:
+
+\begin{enumerate}
+\item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
+\item Reingold-Tilford trees provide a clear visual represent the tree structure of the merge-tree.
+\item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
+clearly showing the parent-child relationships between commits and merges.
+\end{enumerate}
+
+\subsubsection{List Tree}
+
+The list tree viewer (figure \ref{fig:list_tree}) is in the form of nested lists,
+and is designed to more closely model the tree view found in file browsers. This
+tree only contains the commits and merges that are within the subtrees of the
+current merge. A commit will never have any items in this tree as it is a leaf. To
+accompany the tree, we include breadcrumbs at the top of the page to enable a user
+to navigate both from the root to the leaves and from a leaf to the root. The last
+item in the breadcrumb list is the current commit, the previous item is the parent
+of the current commit, and the first item is the root merge into the kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
+        \caption{List Tree}
+        \label{fig:list_tree}
+\end{figure}
+
+\subsubsection{Reingold-Tilford Tree}
+
+\evantodo{we need a reference to these trees}
+
+To improve the visualization of the merge-tree, we first transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master.
+However, we keep the topological order of the commits in the tree-merge as the order of the
+leaves for a given merge (e.g. if merge-tree contains two commits A, B merged into C, then the simplified tree contains links B and C directly to A, but it
+preserves the order of B with respect to C by showing B above C.
+This allows the clustering of commits along their next merge, and avoid the space needed to
+present long sequences of commits without merges in between.
+
+The Reingold-Tilford tree (Figures \ref{fig:zoomed_tree} and \ref{fig:tree}) allows the visualization and navigation
+of the entire merge-tree in an intuitive representation of the tree. This illustrates a clear notion of
+root and leaves, and how to navigate in either direction. Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
+thousands of commits and merges. While the tree is capable of producing a
+visualization, it becomes much more difficult to understand. For example the merge 73287a43cc79ca06629a88d1a199cd283f42456a performed by Linus Torvalds in May
+1, 2013, merged 1722 non-merge commits and 91 merges. footnote{This commit can be inspected at \url{http://li.turingmachine.org/commits/73287a43cc79ca06629a88d1a199cd283f42456a}}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
+        \caption{Entire Reingold-Tilford Tree of a merge. This merge by Linus (a992bf836f9c3039a16f4bd068d161c86c6c3e2c) contains 1722 includes non-merge commits
+          touching 239 modules and 1525 files.}
+        \label{fig:zoomed_tree}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
+        \caption{Detail of the Reingold-Tilford Tree in Figure~\ref{fig:zoomed_tree}.}
+        \label{fig:tree}
+\end{figure}
+
+Clicking on subtrees will cause them to collapse, only showing parts of the tree
+that are of interest to the user. Furthermore, the user can zoom the tree by
+scrolling the mouse wheel to see more of the topology of the tree, or to see the
+details of a given part of the tree.
+
+\dmg{do you have a figure that shows the collapsed subtrees? Try to use the same HUGE commit}
+
+\subsubsection{Bubble Tree}
+
+\evantodo{we need reference to bubble trees}
+
+The bubble tree (figure \ref{fig:bubble_tree}) provides a very clear picture of
+where a commit is located in the tree hierarchy. The currently selected commit or merge
+is colored red, while non-selected merges use a shade of blue and commits use white.
+The depth of the blue colouring is dependent on how deep in the hierarchy that node
+is. The root is the lightest blue, and the contained merges are progressively
+darker.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
+        \caption{Bubble Tree of a large merge, showing in red (the tiny dot close to the middle) the location of a specific commit of interest.}
+        \label{fig:bubble_tree}
+\end{figure}
+
+The bubble tree doesn't have an implicit way of providing additional information,
+placing any text near the nodes makes the tree impossible to read, so we include a
+separate pane in the web page. When a user hovers over a node, the pane shows
+additional information about the author and the commit message and a link to the
+detailed page for that commit or merge. If the user clicks on a node, the tree will
+zoom to that node and the information in the info pane will persist, enabling the
+user to click the link. This tree provides an easy mechanism for users to navigate
+from the root to the leaves and vice versa.
+
+\section{Implementation Details}
+
+\tool was built using Nginx as the main web server, postgresql as the dbms, and Flask as the framework for generating the dynamic content.
+
+Nginx is a high-performance, open source, scalable HTTP and mail server. It is designed with massive
+scalability and multi-processing in mind. Instead of using request-driven threading,
+Nginx uses an event-driven architecture, making it more scalable, proving itself as
+being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
+other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.
+
+Postgresql is our DBMS of choice, providing fast results.
+
+We chose the python micro-framework
+Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the dynamic
+content. Working with flask is simple and easy, allowing us to work in Python.
+
+\subsection{Extendability}
+
+The front-end uses asynchronous requests to gather the information for the trees and
+tables. This enables third-parties to implement new front-ends, trying other
+designs, though this interface may change in the future. The tree, file, and module
+information is accessible through
+http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
+http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
+http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.
+
+The response for the tree is a single object. This object is the root node of the
+tree, and may contain an object of children objects in the \verb|children| field,
+which are in the same format as the root object. The children object uses the commit
+ID as the key.
+
+The tree node objects have the following fields:
+
+\begin{tabular}{ccl}
+        Field name      & Data type & Description\\\hline
+        \verb|cid|      & string    & Commit ID\\
+        \verb|name|     & string    & One-line log preview\\
+        \verb|mlinus|   & string    & Root merge commit ID\\
+        \verb|author|   & string    & Author name and email\\
+        \verb|mnext|    & string    & Parent merge commit ID\\
+        \verb|children| & object    & Object of tree objects\\
+\end{tabular}
+
+The file responses contain only the files that the selected merge or commit works
+with. The response is a single object in the form of a tree. This object is the root
+of the tree, and represents the current merge or commit. If the current position in
+the tree is an inner node, the response will contain the child nodes in the
+``children'' field, otherwise ``children'' will be an empty list. The format of file
+response is as follows,
+
+\begin{tabular}{lll}
+        Field name & Data type & Description\\\hline
+        \verb|cid| & string & Commit ID\\
+        \verb|mnext| & string & Parent merge commit ID\\
+        \verb|children| & list & List of tree objects\\
+        \verb|files| & list of tuples&
+        \footnotesize{
+        \begin{tabular}{lc}
+                \verb|Filename| & string\\
+                \verb|Lines added| & uint\\
+                \verb|Lines removed| & uint\\
+        \end{tabular}}\\
+\end{tabular}
+
+\section{Discussion}
+
+From a developers point of view, the are two major disadvantages of the DAG model of git: a) its edges point backwards, i.e. commits point to their ancestor,
+not to the commit that succeeds them in their way to be integrated; and b) the DAG contains many more edges that are necessary to understand how integration
+happened. We addressed these two issues with the creation of the merge-tree model of the DAG of Linux. Effectively the merge-tree view recoveres the details of each merged
+branch into master.
+
+In our experience, no other git repository reaches the level of DAG complexity that Linux has. In most git repositories  every a merge into master does not
+merge other merges, and most merges bring very few new commits with them. However, even in those cases, the merge-tree will provide a valuable summary of how
+commits have been integrated into master, specially since the time of integration might be very different from the time the commits were created.
+
+The biggest challenge, however, is computing the merge-tree. Our heuristic is applicable \textbf{only} to the Linux kernel, and only since July 2009 and it is
+very likely it will not work for other repositories, specially those that have fast-forward commits into Master (as the majority of repositories). One way to
+address this is lack of data is by monitoring at short intervals the arrival of commits into Master (as we did in \cite{German2015}, i.e.
+\emph{continuously mining} a git repo). This however, requires planning.
+
+Regarding \tool. Because it leverages the merge-tree of the kernel, it provides mechanisms and visualizations that other tools are unable to
+produce.  It gives us the ability to see how commits are being merged
+into the kernel, allowing us to better understand how code is integrated into the kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
+        \caption{Merge Dag View}
+        \label{fig:dag_view}
+\end{figure}
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.97\textwidth]{figures/042dd_tree.png}
+        \caption{Merge Tree View \dmg{this figure is unreadable. Any chance you can make the text bigger?}}
+        \label{fig:tree_view}
+\end{figure*}
+
+Our goal was to build a tool to enable maintainers to effectively navigate and browse
+the changes performed to the kernel over the period of a release.
+Achieving this goal includes removing information that does not pertain to the area
+of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
+figure \ref{fig:tree_view} we can visually compare the results returned from gitk
+and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
+from kernel version 3.10. Note first that gitk has no way to show that this is a merge into Master.
+
+The log information to this merge-commit in both of these
+figures is identical, but the presentation drastically changes our ability to
+comprehend what we are seeing. The primary difference is the removal of irrelevant
+information. In gitk (\ref{fig:dag_view}), the visualization results include commits
+and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
+only includes the results that are specific to the component of the kernel that we
+are interested in.
+
+This merge is relatively small, containing only four sub-merges, three of which
+only contain a single commit, and the last containing two commits. With our
+tool, we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
+furthermore, users must work to determine where this merge ends and the next one
+begins.
+
+\tool further enhances our understanding by summarizing the files
+and modules that were edited in the entire branch being merged. We are able to determine that three
+files, ``palamas-regulator.c'' had two lines added and two lines removed,
+``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
+lines added and 2 lines removed. Finally, we are able to determine that only the
+``regulator'' module was modified in this merge and was modified by 5 commits.
+
+Using \tool many advantages over the DAG model, but there are two
+drawbacks. The first drawback is that a server must query the state of the master
+repository since time information is not stored by git. The second drawback is the
+need to pre-compute the merge-tree of the repository.
+
+% Our core model provides advantages to the DAG model used by other tools. We separate
+% the commits and merges into sub-trees and focus on a single sub-tree at a time. This
+% enables us to filter out commits and merges that do not pertain to the component of
+% the kernel that we are interested in, providing a clearer picture of what is
+% happening in a given merge. The tree model also enables us to aggregate metadata
+% easily. With the addition of the continuous polling of the repository, the model
+% provides us with a mechanism of performing date-range queries with meaningful
+% results.
+
+\section{Future Work}
+
+There are still many areas that can be improved before the full potential of our
+model can be realized. Here, we outline various areas of the tool that still need
+more attention.
+
+\subsection{Files}
+
+There currently is no functionality surrounding searching by filename. Users falling
+under use-case 2 may know what file they are editing and try to determine how it may
+work with the other commits and merges in the module. It is possible that a third
+use-case may arise, where the user wants to determine all commits that effect a
+given file. In both cases, bottom-to-top approach is applied.
+
+There is limited functionality to the presentation of the file information. At a
+minimum, the patches for the commits can be displayed. From there, the patches can
+be used to piece together parts of the file to generate a single patch at a given
+merge rather than displaying each patch individually. The patches can also be used
+for determining what kinds of changes were made, if the lines are being added,
+removed, or being replaced.
+
+\subsection{Authorship}
+
+Our model can aggregate more information than what we have implemented. The
+authorship information is important for licensing purposes, and we can show all
+authors contributing to commits in a  merge and how many commits they contributed to
+the merge. We could go further, providing information about what files they edited,
+how many lines they added and removed, and more.
+
+\subsection{Evaluation}
+
+At this time, we have no evidence that our tool is able to improve the work-flow of
+maintainers. We believe that the tool is able to improve the work-flow and
+performance of maintainers because it provides cleaner mechanisms of visualizing and
+presenting the commit information. It is able to provide more relevant information,
+while removing information that is irrelevant to a given module or set of merges.
+
+We can either perform user-testing to show how a users workflow changes, or we could
+have maintainers evaluate and critique the tool and use their feedback to determine
+if the tool meets our goals.
 
@@
 
-\section{Visualizing the merge-tree of Linux}
-
-The goal is to build a tool that simplifies  navigation of the kernel commit information, while providing an explanation of the changes occurring at each commit
-and, more importantly, indicates how commits are merged in their way to master.
-We designed the tool with two use-cases in mind, though a user may switch
-between the cases as they work.
-
-\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
-These are users that are maintaining a section of the kernel and would like to pick
-a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
-reducing the amount of re-implementation work. For these users, it is important to have the
-ability to aggregate metadata about files and modules being effected by the merge.
-\dmg{what is root?, what is leaves? perhaps remove unless you can be specific}
-It is also important for these users to be able to navigate from the root toward the
-leaves.
-
-\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
-These are users that start with a known merge or commit and would like to see what
-other changes are being made in commits that are in the same merge. This is useful
-to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
- This is primarily for maintainers that need to perform
-some specific cherry picking of commits. We
-must provide these users a mechanism for navigating from a single commit toward the
-master, allowing them to see the bigger picture.
-
-\subsection{Core Model}
-
-In our visualizations, each commit is either made directly to Master, or is part of a Merge-tree, with its root in Master (a merge made by Linus Torvalds). Each merge may contain zero or more
-children, which are other merges and commits.
-
-To improve the visualization of the merge-tree, we transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master. In
-this model, commits have no children and are always leaf nodes. This allows the clustering of commits along their next merge, and avoid the space needed to
-present long sequences of commits without merges in between.
-
-Commits contain the
-metadata for the changes made. This metadata includes the files changed, the lines
-added and removed from each file, the author, the date the commit was merged into
-the merge that led to being merged into the kernel, and the date the commit was
-authored, the patch, and the commit log. Merges contain less metadata, only storing
-the author of the merge, the log, the commit date, and the author date. The details
-of the model are outlined in \cite{German2015}.
-
-\section{Design}
-
-\dmg{I would remove the command line shell, and use the space for related work}
-
-We came up with two designs. We recognized that the structure of the repository was
-similar to the wide and shallow tree-structure of a file system. We began with a
-command-line shell interface. Then transitioned to a web-based tool where we looked
-beyond the file system structure. In this section, we describe the progression of
-our tool design, showing the benefits and drawbacks of each component.
-
-\subsection{Command-line Shell}
-
-Our initial intuition was to work with the commits and merges as directory trees
-where the commits are files and the merges are directories. There are many
-similarities between the tree structure of a file system and the tree structure of
-the repository. Both trees are generally shallow, but very wide. File systems have
-been studied for quite some time and there are many tools capable of visualizing and
-working with the file structure.
-
-The first tool we implemented was a small command-line shell (figure
-\ref{fig:shell}) that allows users to navigate through the commits and merges as one
-might navigate through files and directories. A user could navigate toward the
-leaves of the tree by using the \verb|cd| command followed by the commit ID of the
-next merge. The user could then navigate toward the root by issuing \verb|cd ../|
-as they would in a file system. Users can use \verb|ls| to view what commits and
-merges exist in the current working merge, then use \verb|cat| to view the log
-preview of a commit or \verb|more| to see the preview and how many lines were added
-and removed from the files in the commit.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/shell.png}
-        \caption{Shell Design}
-        \label{fig:shell}
-\end{figure}
-
-The shell does not meet our design goals. It only provides users with the ability to
-navigate from the root. If there is a commit or merge that they would like to
-investigate, they must find it with another tool and navigate through the tree to
-find it. Furthermore, it does not aggregate information in the merges, so users must
-manually aggregate the metadata. Another issue with this design is that it presents
-the users with the commit hashes, which are quite long and unpleasant to remember
-and type into the shell, but also do not provide any quick explanation of what the
-commit or merge may contain.
-
-\subsection{Web-based Tool}
-
-We replace the command-line interface with a web-based tool. Creating a web-based
-tool enables users to use the system without having to install additional software
-or store a large database, making it more accessible, more easily maintainable, and
-platform independent. A web-based tool can utilize more interactive means of
-navigation and provide better explanations of commits and merges than the
-command-line interface.
-
-Our tool uses the following mechanisms to reach our goals of better navigation and
-better explanation of the changes.
-
-\begin{itemize}
-        \item Filter by search
-        \item View files edited
-        \item View modules
-        \item Tree viewer
-\end{itemize}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View}
-        \label{fig:search}
-\end{figure}
-
-Searching (figure \ref{fig:search}) allows a user to filter commits and merges that
-are irrelevant. The search mechanism breaks down the results by release version. A
-user can further narrow down the search range using the commit date range. The
-commit date refers to the date of the creation of the top-level merge. The commits
-and merges returned may have a commit date beyond the queried dates; the date-range
-query refers to the date that the commit is finally merged into the kernel by Linus.
-A commit may not be merged immediately, and in some cases may take years before
-being applied. A user may then provide a search text, filtering by the author name,
-the commit ID, or keyword from the log. Any part of the author name may show up in
-the results, including searching by email address. The commit ID must be specified
-from the first character to the last character. For example, the commit
-`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
-commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
-each of the features of the tool.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search_results.png}
-        \caption{Search Results}
-        \label{fig:results}
-\end{figure}
-
-In the search results, the user is presented with the one-line log message preview,
-the author's name and email, the date the commit was authored, and the date the
-commit was last committed. The commit date and author date will usually be the same;
-however, if the commit has been rebased or cherry-picked into a merge that
-eventually is merged into the kernel, the commit date and authored date will be
-different. We can see this in the first and third entries in figure
-\ref{fig:results}.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/message_view.png}
-        \caption{Commit Message}
-        \label{fig:message}
-\end{figure}
-
-Once a user selects a commit or merge to investigate, they are presented with a
-tabbed pane allowing them to view the full commit log, the files edited, the modules
-involved, and the tree views.
-
-The first tab displays the full commit log (figure \ref{fig:message}). From this, a
-user is able to see what they would see had they searched for the commit using git
-log. This doesn't provide additional information to the other tools, but helps to
-complete the functionality of our tool. The commit log provides a user with the
-information about the content of the commit and who has signed-off on the commit to
-ensure that it is good quality. The message for merges may contain a summary of the
-commits being merged.  The information within these messages is highly variable, and
-is completely dependent on the author's style. As the user moves toward the
-top-level merge, the quality of these messages generally improves.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/file_view.png}
-        \caption{Commit Files}
-        \label{fig:files}
-\end{figure}
-
-The second tab is the files tab (figure \ref{fig:files}). This tab provides
-information on what files have been edited, how many lines were added, and how many
-lines were removed in a given commit. This functionality is similar to the other
-tools available. Our tree-based design model allows us to extend this functionality
-to merges, which the other tools are unable to show. To find the number of lines
-added to a file in a merge, we take the sum of the lines added to that file in each
-of the children of that merge. We do the same for calculating the number of lines
-removed. This could be extended to use the patch information to determine when a
-line has been replaced rather than incrementing the lines added and removed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/modules.png}
-        \caption{Commit Modules}
-        \label{fig:modules}
-\end{figure}
-
-The modules tab (figure \ref{fig:modules}) shows the modules that are contained
-within the commit. Modules are not natively recognized by git, and are not going to
-be present in all repositories.  In the Linux repository, authors put the module
-they are working on in the one-line preview of the log-message. We heuristically
-extract the module by taking all text in the log preview of commits until the first
-colon.  Modules are logical partitions of the information in the kernel. Depending
-on where the author was working, modules can be general, such as ``bluetooth'' and
-``wireless'', or can be quite specific for individual hardware, such as
-``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
-correctly follow this format and the hueristic approach fails. Like files, we are
-able to show what modules are edited in merges and commits. The image in figure
-\ref{fig:modules} shows the modules for a commit.  As commits have no children, they
-will only have a single module; merges with more than one child will have more.
-
-Finally, we have the tree view tab. The tree view is designed for providing easy
-navigation within a tree and providing a clear topological view of the branch of the
-main kernel. We have experimented with various tree designs to find a design that
-allows for easy navigation and visualization of both large and small trees. We
-discuss the various tree designs in section \ref{treeview_section}.
-
-\subsection{Tree Views} \label{treeview_section}
-
-The tree view is what makes this tool unique to other tools, providing a full view
-of the entire merge tree. We have experimented with various types of trees; list
-trees are similar to the representation used in many file system explorers,
-Reingold-Tilford trees clearly represent the tree structure, and bubble trees
-organize the data hierarchically by having the parent node contain the child nodes
-similarly to tree maps, but clearly showing the parent-child relationships between
-nodes.
-
-\subsubsection{List Tree}
-
-The list tree viewer (figure \ref{fig:list_tree}) is in the form of nested lists,
-and is designed to more closely model the tree view found in file browsers. This
-tree only contains the commits and merges that are within the subtrees of the
-current merge. A commit will never have any items in this tree as it is a leaf. To
-accompany the tree, we include breadcrumbs at the top of the page to enable a user
-to navigate both from the root to the leaves and from a leaf to the root. The last
-item in the breadcrumb list is the current commit, the previous item is the parent
-of the current commit, and the first item is the root merge into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List Tree}
-        \label{fig:list_tree}
-\end{figure}
-
-\subsubsection{Reingold-Tilford Tree}
-
-The Reingold-Tilford tree (figure \ref{fig:tree})produces very nice results,
-creating an intuitive representation of the tree. This illustrates a clear notion of
-root and leaves, and how to navigate in either direction.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
-        \caption{Reingold-Tilford Tree}
-        \label{fig:tree}
-\end{figure}
-
-Clicking on subtrees will cause them to minimize, only showing parts of the tree
-that are of interest to the user. Furthermore, the user can zoom the tree by
-scrolling the mouse wheel to see more of the topology of the tree, or to see the
-details of a given part of the tree.
-
-Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
-thousands of commits and merges. While the tree is capable of producing a
-visualization, it becomes much more difficult to understand.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
-        \caption{Reingold-Tilford Tree zoomed out}
-        \label{fig:zoomed_tree}
-\end{figure}
-
-\subsubsection{Bubble Tree}
-
-The bubble tree (figure \ref{fig:bubble_tree}) provides a very clear picture of
-where a commit is located in the hierarchy. The currently selected commit or merge
-is colored red, while non-selected merges use a shade of blue and commits use white.
-The depth of the blue colouring is dependent on how deep in the hierarchy that node
-is. The root is the lightest blue, and the contained merges are progressively
-darker.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree}
-        \label{fig:bubble_tree}
-\end{figure}
-
-The bubble tree doesn't have an implicit way of providing additional information,
-placing any text near the nodes makes the tree impossible to read, so we include a
-separate pane in the web page. When a user hovers over a node, the pane shows
-additional information about the author and the commit message and a link to the
-detailed page for that commit or merge. If the user clicks on a node, the tree will
-zoom to that node and the information in the info pane will persist, enabling the
-user to click the link. This tree provides an easy mechanism for users to navigate
-from the root to the leaves and vice versa.
-
-\section{Implementation Details}
-
-\evan{Is this section necessary? The tables are probably useful, but the rest... not
-sure on}
-
-We built a web-based tool to provide information about the kernel.  We chose to use
-Nginx as the main web server, postgresql as the dbms, and Flask as the framework for
-generating the dynamic content.
-
-Nginx is a high-performance, open source, scalable HTTP and mail server. Started in
-2002, the project is much younger than the Apache project, and designed with massive
-scalability and multi-processing in mind. Instead of using request-driven threading,
-Nginx uses an event-driven architecture, making it more scalable, proving itself as
-being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
-other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.
-
-Postgresql is our dbms of choice, providing fast and reliable results. There are
-multiple reasons behind choosing postgresql over another dbms. The main reason being
-that the dataset built by German et al. was generated from a postgresql database,
-allowing us to import the data directly without any further conversions.
-
-We chose the python micro-framework
-Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the dynamic
-content. Working with flask is simple and easy, allowing us to work in Python.
-
-The database itself is broken down into 5 main tables; commits, filesmod, logs,
-pathtomerge, and releases. Commits contains most of the metadata for each commit; it
-contains the author, the authored date, and committer, the commit date, a boolean
-saying what is a merge, and the patch. Filesmod contains the information on what
-files were modified for the commits. Since a file may be modified by multiple
-commits, and a commit may be modifying multiple files, we can do one of two things
-for the primary key. We could use both the filename and the commit ID as the primary
-key, but we just use an index. The filesmod table contains the filename, the commit
-ID, the number of lines added, and the number of lines removed. Trivially, it
-contains the index, but that isn't necessary for anything beyond the primary key.
-Logs contains the commit log preview and full commit log. Pathtomerge is the
-important table.  Pathtomerge contains the commit ID, the number of merges between
-the commit and the top-level merge with the kernel, the next merge in the merge path
-up to the top-level merge with the kernel, the top-level merge into the kernel, and
-the date it is merged into the kernel. The information in this table must be
-generated periodically by the server because we are unable to generate the tree
-structure correctly straight from the information stored by git. Pathtomerge does
-not contain any top-level merges, as these will never have a next merge in the tree;
-since we are only storing the parents and not the children, storing the top-level
-merge is redundant and unnecessary. The releases table stores the release
-information.
-
-It contains
-\begin{itemize}
-        \item Version name
-        \item Candidate releases versus real releases
-        \item Previous release
-        \item Previous real release
-        \item Previous release commit ID
-        \item Previous real release commit ID
-        \item Current version commit ID
-\end{itemize}
-The current version commit ID represents the last top-level merge into the version
-of the kernel. Using this and the commits table, we are able to determine the
-release date of any version of the kernel. Then using the previous release date, we
-can find the range of dates where valid top-level merges are found for that version.
-
-\section{Discussion}
-
-Our tool provides mechanisms and visualizations that other tools are unable to
-produce. Continuous mining of the repository provides more information about how the
-repository is changing. It gives us the ability to see how commits are being merged
-into the kernel, allowing us to more reliably build trees from the snapshots of the
-repository.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
-        \caption{Merge Dag View}
-        \label{fig:dag_view}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.png}
-        \caption{Merge Tree View}
-        \label{fig:tree_view}
-\end{figure}
-
-Our goal was to build a tool to enable maintainers to effectively navigate and pull
-an explanation of the changes performed to the kernel over the period of a release.
-Achieving this goal includes removing information that does not pertain to the area
-of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
-figure \ref{fig:tree_view} we can visually compare the results returned from gitk
-and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. The relevant information to this commit in both of these
-figures is identical, but the presentation drastically changes our ability to
-comprehend what we are seeing. The primary difference is the removal of irrelevant
-information. In gitk (\ref{fig:dag_view}), the visualization results include commits
-and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
-only includes the results that are specific to the component of the kernel that we
-are interested in.
-
-This merge is relatively small, containing only four merges, three of which
-only contain a single commit, and the last containing two commits. With our
-tool, we are able to immediately see what commits and merges pertain to our
-section of the kernel. The DAG view of this provides almost no explanation,
-furthermore, users must work to determine where this merge ends and the next one
-begins.
-
-Our tool further enhances our understanding by immediately giving us the files
-and modules that were edited in the branch. We are able to determine that three
-files, ``palamas-regulator.c'' had two lines added and two lines removed,
-``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
-lines added and 2 lines removed. Finally, we are able to determine that only the
-``regulator'' module was modified in this merge and was modified by 5 commits.
-
-\subsection{Extendability}
-
-The front-end uses asynchronous requests to gather the information for the trees and
-tables. This enables third-parties to implement new front-ends, trying other
-designs, though this interface may change in the future. The tree, file, and module
-information is accessible through
-http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.
-
-The response for the tree is a single object. This object is the root node of the
-tree, and may contain an object of children objects in the \verb|children| field,
-which are in the same format as the root object. The children object uses the commit
-ID as the key.
-
-The tree node objects have the following fields:
-
-\begin{tabular}{ccl}
-        Field name      & Data type & Description\\\hline
-        \verb|cid|      & string    & Commit ID\\
-        \verb|name|     & string    & One-line log preview\\
-        \verb|mlinus|   & string    & Root merge commit ID\\
-        \verb|author|   & string    & Author name and email\\
-        \verb|mnext|    & string    & Parent merge commit ID\\
-        \verb|children| & object    & Object of tree objects\\
-\end{tabular}
-
-The file responses contain only the files that the selected merge or commit works
-with. The response is a single object in the form of a tree. This object is the root
-of the tree, and represents the current merge or commit. If the current position in
-the tree is an inner node, the response will contain the child nodes in the
-``children'' field, otherwise ``children'' will be an empty list. The format of file
-response is as follows,
-
-\begin{tabular}{lll}
-        Field name & Data type & Description\\\hline
-        \verb|cid| & string & Commit ID\\
-        \verb|mnext| & string & Parent merge commit ID\\
-        \verb|children| & list & List of tree objects\\
-        \verb|files| & list of tuples&
-        \footnotesize{
-        \begin{tabular}{lc}
-                \verb|Filename| & string\\
-                \verb|Lines added| & uint\\
-                \verb|Lines removed| & uint\\
-        \end{tabular}}\\
-\end{tabular}
-
-\section{Future Work}
-
-There are still many areas that can be improved before the full potential of our
-model can be realized. Here, we outline various areas of the tool that still need
-more attention.
-
-\subsection{Files}
-
-There currently is no functionality surrounding searching by filename. Users falling
-under use-case 2 may know what file they are editing and try to determine how it may
-work with the other commits and merges in the module. It is possible that a third
-use-case may arise, where the user wants to determine all commits that effect a
-given file. In both cases, bottom-to-top approach is applied.
-
-There is limited functionality to the presentation of the file information. At a
-minimum, the patches for the commits can be displayed. From there, the patches can
-be used to piece together parts of the file to generate a single patch at a given
-merge rather than displaying each patch individually. The patches can also be used
-for determining what kinds of changes were made, if the lines are being added,
-removed, or being replaced.
-
-\subsection{Authorship}
-
-Our model can aggregate more information than what we have implemented. The
-authorship information is important for licensing purposes, and we can show all
-authors contributing to commits in a  merge and how many commits they contributed to
-the merge. We could go further, providing information about what files they edited,
-how many lines they added and removed, and more.
-
-\subsection{Evaluation}
-
-At this time, we have no evidence that our tool is able to improve the work-flow of
-maintainers. We believe that the tool is able to improve the work-flow and
-performance of maintainers because it provides cleaner mechanisms of visualizing and
-presenting the commit information. It is able to provide more relevant information,
-while removing information that is irrelevant to a given module or set of merges.
-
-We can either perform user-testing to show how a users workflow changes, or we could
-have maintainers evaluate and critique the tool and use their feedback to determine
-if the tool meets our goals.
-
 \section{Conclusion}

### 2016-04-25 linvis.tex
Adding Related works

--- 2016-04-25 linvis.tex
@@
 
-\evantodo{Move related work to the end, try to match the text of the paper to my intro}
+\evantodo{pick 'gitk' or 'Gitk'}
 
-To our knowledge, we know of no git repository visualization tool that builds
-a tree that maps the path commits follow to master
-from the DAG provided by git. This may be because more information is required
-to generate the tree-model than what is stored by git. % Git uses a DAG model, which
-% enables history-altering actions, like rebases to occur, whereas older tools like
-% CVS and svn construct trees\cite{CVS2008} but do not work with history-altering
-% actions.
+Version control systems monitor the development lifetime of software projects. This
+makes the version control system vital in providing information about how a software
+project is being developed. To our knowledge, we know of no git repository
+visualization tool that builds a tree that maps the path a commit follows to the
+master branch from the DAG provided by git. This may be because more information is
+required to generate the tree-model than what is stored by git. However, there has
+been a lot of work in providing visualizations of various repositories.
 
-% A server must sit and monitor the git repository as it changes over time to record
-% the changes to history. Many of the visualization tools are designed for a
-% general-purpose approach, where any repository is directly usable without additional
-% information. Polling for this additional information makes the visualizer specific
-to a repository or project.
+Many of the tools work to address the issues of communication between developers and
+inter-team collaboration. Codebook uses a mining technique to determine the
+developer of a piece of code, the program manager who wrote the specification for
+the code, and the program managers and developers on the team who were working
+together\cite{Begel2010}. This is designed to help collaboration between developers
+on a project while the project is being developed.  In maintaining the code, the
+code has already been written, the original developer many not have seen this piece
+of code in years, and will most likely have moved to another problem. If the
+maintainer is pulling from a part of the kernel that has been recently updated, in
+the event of a zero-day attack for instance, the information about who is developing
+that portion of the kernel is of use; however, if the maintainer is pulling from
+sections that are older, then the maintainer must be able to understand the code
+using the provided tools, not the original developer.
 
-\subsection{Other Visualizers}
+Hipikat is another tool with a focus on communication issues between developers. It
+is designed for non-collocated teams, helping new developers understand design
+decisions of the system \cite{Cubranic2005}. It records artifacts about the project
+to build a "memory" of the project development that new users can work with to
+understand where and how to make modifications in large projects. The artifacts are
+recorded using tools like bugzilla and CVS. Such a tool could prove useful in
+understanding the changes between a maintainer's current version of an older kernel
+and the changes being made to the current master version of the kernel.
 
-Gitk, Gitg, and other similar git repository visualization tools use the DAG as the
-central organization model. This model is sufficient for smaller projects, providing
-meaningful information; however, in larger projects, like the Linux kernel, the
-branches become too tangled and too deep to derive any meaning from the DAG. Users
-can interact with the DAG by going to the previous commit in the chain or jumping to
-the parent merge. Given the direction that each node in the DAG has (pointing to its parent)
-it can be assumed that tools that exploit the DAG (such as gitk and similar tools) are designed for users
-working from a given commit towards the past. Working in the opposite direction is hard, since merge commits provide multiple paths towards the present, with no
-way to distinguish which path is the one used by the current commit to be merged into master.
+These tools work toward assisting the communications between developers and
+providing a summary of the project. Our goal includes providing a visual summary of
+the topology of the repository, as well as providing a summary of the changes made
+at a location within the repository. EPOSee, Gource, and Gitk are all tools designed
+for visualizing various aspects of repositories.
 
-Furthermore, the DAG model is only able to provide file-level information for
-commits. It is unable to quickly aggregate the file data over all commits that
-are part of a merge. If a user would like to view what files were edited at a merge,
-they must manually aggregate this information by hand. While some merges are
-relatively small, containing less than 10 commits, other merges are large,
-containing over 1500 edited files, as in the case of commit \textit{73287a4}. This
-issue is important for authorship and licensing. A merge can only show one author,
-the contributor making the merge, but does not show who wrote the commits contained
-within that merge.
+EPOSee is a tool for visualizing changes that often happen simultaneously in a
+commit or as a sequence of events \cite{Burch2005}. It produces visualizations about
+the confidence that changes are made either together or sequentially. For instance,
+changes in the header file may also indicate that a change was made in the
+implementation file. The visualization tool is designed to visualize implicit rules
+about the order of operations on various files, if too many outliers are detected,
+the project manager may want to restructure the project.  While this is useful for
+new projects, the development model of the Linux kernel is already structured and is
+unlikely to change without good reasons behind it.
 
-%% Check for papers if they exist
-%% This tool isn't specifically designed for showing Linux
+Gource is a tool for providing an interactive timelapse of the state of a
+repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
+type of contribution a developer is making. These contribution types are one of,
+adding a file, removing a file, and changing a file. While the timelapse is
+interesting to watch, it does not provide any additional explanation of the changes
+actually being made, only the frequency that they are being made and who is making
+them.
 
-\subsection{Treemap visualizer}
-
-University of Maryland presented a treemap visualization
-tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
-pictured in figure \ref{fig:treemap}, for displaying the file structure of a given
-version of the kernel. Treemaps are good for visualizing the topology of shallow,
-wide trees, making it a good candidate for displaying file systems. Treemaps are
-limited in what information can be displayed. The information within the treemap
-does not provide clear insight into what changes were made to the files in the given
-release. It only provides information on where the file is located within the file
-system; they do not provide temporal information. File systems and the merge
-tree of the kernel share many similarities, in both cases, they represent trees that
-are wide and shallow, suggesting that a treemap could potentially work to visualize
-the commit and merge structure of the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
-        \caption{Treemap of Linux 2.5.33 file structure}
-        \label{fig:treemap}
-\end{figure}
-
-The treemap is unable to take into account multiple commits behind the same merge
-with the same log preview message. Commits and merges have an ordering in which they
-were added to the kernel, which cannot be represented by a treemap, but is important
-and can be used to assist in building the tree.
-
-The treemap is also designed for working with data of differing types. In the
-filesystem, there are many kinds of files. There are text files, scripts, images,
-binary files, and other types. These can be given colours making differences stand
-out. There is no obvious way to assign meaningful types to the commits such that all
-the commits within a given merge do not have the same colour, although a user
-could be given a controller to colour by author or number of files edited.
-
-Finally, the treemap does not provide any additional information about the
-contents of each cell, only that the cell exists and where it exists in the
-tree. The user is unable to see their current position within the map, or if we
-assume that their current position is the largest surrounding box, they can
-no longer see how their current position interacts with the rest of the
-system.
-%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
-%% Find paper for this maybe?
-
-\subsection{List viewer}
-
-Moghaddam designed a web-based visualization
-tool\footnote{\url{http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html}}
-displaying the chain of merges into the kernel. There is no information on what
-commits are within a given merge, or what files were edited. The visualization
-tool provides information about who authored the commit and when it was
-authored. We can see the visualization tool in figure \ref{fig:listviewer}
-displaying a part of the kernel commit information, from commit ``26b23ac''
-according to the tool.
-
-%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
-%% Find paper for this maybe?
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/gitvis.png}
-        \caption{List Viewer}
-        \label{fig:listviewer}
-\end{figure}
+The current industry standard tool for git repository visualization is Gitk. The
+Gitk interface is built around the central DAG viewer. The DAG view displays the
+commits on their respective branches, the author of the commit or merge, and the
+date the commit or merge was authored. A user can select a single merge or commit
+to view additional information. For both merges and commits, gitk will present the
+user with the git log. If the user selected a commit, gitk will also present the
+user with the patch information and the names of the files edited. gitk is unable to
+provide the patches and filenames information in merges, as it is unable to
+aggregate commit information in merges.
 

### 2016-04-25 linvis.tex

--- 2016-04-25 linvis.tex
@@
         \includegraphics[width=0.47\textwidth]{figures/gitk.png}
-        \caption{A sample DAG view of commits from Linux in gitk. Commits point towards their ancestors, and there is no clear indication of the way the travel
-          to reach Master. In fact, neither gitk nor git is capable of showing what commits were done in master.}
+        \caption{A sample DAG view of commits from Linux in Gitk. Commits point towards their ancestors, and there is no clear indication of the way the travel
+          to reach Master. In fact, neither Gitk nor git is capable of showing what commits were done in master.}
         \label{fig:gitk}
@@
 
-We designed \tool with two use-cases in mind, though a user may switch
-between the cases as they work.
+We designed \tool with two use-cases in mind, though a user may switch between the
+cases as they work.
 
@@
 of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
-figure \ref{fig:tree_view} we can visually compare the results returned from gitk
+figure \ref{fig:tree_view} we can visually compare the results returned from Gitk
 and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. Note first that gitk has no way to show that this is a merge into Master.
+from kernel version 3.10. Note first that Gitk has no way to show that this is a merge into Master.
 
@@
 comprehend what we are seeing. The primary difference is the removal of irrelevant
-information. In gitk (\ref{fig:dag_view}), the visualization results include commits
+information. In Gitk (\ref{fig:dag_view}), the visualization results include commits
 and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
@@
 
-\evantodo{Move related work to the end, try to match the text of the paper to my intro}
+Version control systems monitor the development lifetime of software projects. This
+makes the version control system vital in providing information about how a software
+project is being developed. To our knowledge, we know of no git repository
+visualization tool that builds a tree that maps the path a commit follows to the
+master branch from the DAG provided by git. This may be because more information is
+required to generate the tree-model than what is stored by git. However, there has
+been a lot of work in providing visualizations of various repositories.
 
-To our knowledge, we know of no git repository visualization tool that builds
-a tree that maps the path commits follow to master
-from the DAG provided by git. This may be because more information is required
-to generate the tree-model than what is stored by git. % Git uses a DAG model, which
-% enables history-altering actions, like rebases to occur, whereas older tools like
-% CVS and svn construct trees\cite{CVS2008} but do not work with history-altering
-% actions.
+Many of the tools work to address the issues of communication between developers and
+inter-team collaboration. Codebook uses a mining technique to determine the
+developer of a piece of code, the program manager who wrote the specification for
+the code, and the program managers and developers on the team who were working
+together\cite{Begel2010}. This is designed to help collaboration between developers
+on a project while the project is being developed.  In maintaining the code, the
+code has already been written, the original developer many not have seen this piece
+of code in years, and will most likely have moved to another problem. If the
+maintainer is pulling from a part of the kernel that has been recently updated, in
+the event of a zero-day attack for instance, the information about who is developing
+that portion of the kernel is of use; however, if the maintainer is pulling from
+sections that are older, then the maintainer must be able to understand the code
+using the provided tools, not the original developer.
 
-% A server must sit and monitor the git repository as it changes over time to record
-% the changes to history. Many of the visualization tools are designed for a
-% general-purpose approach, where any repository is directly usable without additional
-% information. Polling for this additional information makes the visualizer specific
-to a repository or project.
+Hipikat is another tool with a focus on communication issues between developers. It
+is designed for non-collocated teams, helping new developers understand design
+decisions of the system \cite{Cubranic2005}. It records artifacts about the project
+to build a "memory" of the project development that new users can work with to
+understand where and how to make modifications in large projects. The artifacts are
+recorded using tools like bugzilla and CVS. Such a tool could prove useful in
+understanding the changes between a maintainer's current version of an older kernel
+and the changes being made to the current master version of the kernel.
 
-\subsection{Other Visualizers}
+These tools work toward assisting the communications between developers and
+providing a summary of the project. Our goal includes providing a visual summary of
+the topology of the repository, as well as providing a summary of the changes made
+at a location within the repository. EPOSee, Gource, and Gitk are all tools designed
+for visualizing various aspects of repositories.
 
-Gitk, Gitg, and other similar git repository visualization tools use the DAG as the
-central organization model. This model is sufficient for smaller projects, providing
-meaningful information; however, in larger projects, like the Linux kernel, the
-branches become too tangled and too deep to derive any meaning from the DAG. Users
-can interact with the DAG by going to the previous commit in the chain or jumping to
-the parent merge. Given the direction that each node in the DAG has (pointing to its parent)
-it can be assumed that tools that exploit the DAG (such as gitk and similar tools) are designed for users
-working from a given commit towards the past. Working in the opposite direction is hard, since merge commits provide multiple paths towards the present, with no
-way to distinguish which path is the one used by the current commit to be merged into master.
+EPOSee is a tool for visualizing changes that often happen simultaneously in a
+commit or as a sequence of events \cite{Burch2005}. It produces visualizations about
+the confidence that changes are made either together or sequentially. For instance,
+changes in the header file may also indicate that a change was made in the
+implementation file. The visualization tool is designed to visualize implicit rules
+about the order of operations on various files, if too many outliers are detected,
+the project manager may want to restructure the project.  While this is useful for
+new projects, the development model of the Linux kernel is already structured and is
+unlikely to change without good reasons behind it.
 
-Furthermore, the DAG model is only able to provide file-level information for
-commits. It is unable to quickly aggregate the file data over all commits that
-are part of a merge. If a user would like to view what files were edited at a merge,
-they must manually aggregate this information by hand. While some merges are
-relatively small, containing less than 10 commits, other merges are large,
-containing over 1500 edited files, as in the case of commit \textit{73287a4}. This
-issue is important for authorship and licensing. A merge can only show one author,
-the contributor making the merge, but does not show who wrote the commits contained
-within that merge.
+Gource is a tool for providing an interactive timelapse of the state of a
+repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
+type of contribution a developer is making. These contribution types are one of,
+adding a file, removing a file, and changing a file. While the timelapse is
+interesting to watch, it does not provide any additional explanation of the changes
+actually being made, only the frequency that they are being made and who is making
+them.
 
-%% Check for papers if they exist
-%% This tool isn't specifically designed for showing Linux
-
-\subsection{Treemap visualizer}
-
-University of Maryland presented a treemap visualization
-tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
-pictured in figure \ref{fig:treemap}, for displaying the file structure of a given
-version of the kernel. Treemaps are good for visualizing the topology of shallow,
-wide trees, making it a good candidate for displaying file systems. Treemaps are
-limited in what information can be displayed. The information within the treemap
-does not provide clear insight into what changes were made to the files in the given
-release. It only provides information on where the file is located within the file
-system; they do not provide temporal information. File systems and the merge
-tree of the kernel share many similarities, in both cases, they represent trees that
-are wide and shallow, suggesting that a treemap could potentially work to visualize
-the commit and merge structure of the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
-        \caption{Treemap of Linux 2.5.33 file structure}
-        \label{fig:treemap}
-\end{figure}
-
-The treemap is unable to take into account multiple commits behind the same merge
-with the same log preview message. Commits and merges have an ordering in which they
-were added to the kernel, which cannot be represented by a treemap, but is important
-and can be used to assist in building the tree.
-
-The treemap is also designed for working with data of differing types. In the
-filesystem, there are many kinds of files. There are text files, scripts, images,
-binary files, and other types. These can be given colours making differences stand
-out. There is no obvious way to assign meaningful types to the commits such that all
-the commits within a given merge do not have the same colour, although a user
-could be given a controller to colour by author or number of files edited.
-
-Finally, the treemap does not provide any additional information about the
-contents of each cell, only that the cell exists and where it exists in the
-tree. The user is unable to see their current position within the map, or if we
-assume that their current position is the largest surrounding box, they can
-no longer see how their current position interacts with the rest of the
-system.
-%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
-%% Find paper for this maybe?
-
-\subsection{List viewer}
-
-Moghaddam designed a web-based visualization
-tool\footnote{\url{http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html}}
-displaying the chain of merges into the kernel. There is no information on what
-commits are within a given merge, or what files were edited. The visualization
-tool provides information about who authored the commit and when it was
-authored. We can see the visualization tool in figure \ref{fig:listviewer}
-displaying a part of the kernel commit information, from commit ``26b23ac''
-according to the tool.
-
-%% http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html
-%% Find paper for this maybe?
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/gitvis.png}
-        \caption{List Viewer}
-        \label{fig:listviewer}
-\end{figure}
+The current industry standard tool for git repository visualization is Gitk. The
+Gitk interface is built around the central DAG viewer. The DAG view displays the
+commits on their respective branches, the author of the commit or merge, and the
+date the commit or merge was authored. A user can select a single merge or commit
+to view additional information. For both merges and commits, Gitk will present the
+user with the git log. If the user selected a commit, Gitk will also present the
+user with the patch information and the names of the files edited. Gitk is unable to
+provide the patches and filenames information in merges, as it is unable to
+aggregate commit information in merges.
 

### 2016-04-25 linvis.tex
Adding citation to Reingold Tilford trees

@@
 
-\evantodo{we need a reference to these trees}
-
 To improve the visualization of the merge-tree, we first transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master.
@@
 
-The Reingold-Tilford tree (Figures \ref{fig:zoomed_tree} and \ref{fig:tree}) allows the visualization and navigation
+The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:zoomed_tree} and \ref{fig:tree}) allows the visualization and navigation
 of the entire merge-tree in an intuitive representation of the tree. This illustrates a clear notion of

### 2016-04-26 linvis.tex
Updating Related Work section and adding citation
Condensing the related work section, still more to be done for it.
Adding citations for bubble tree.

@@
 
+\evan{pdfs are not included in the repository}
 \begin{figure}[htbp]
@@
 
-The bubble tree (figure \ref{fig:bubble_tree}) provides a very clear picture of
-where a commit is located in the tree hierarchy. The currently selected commit or merge
-is colored red, while non-selected merges use a shade of blue and commits use white.
-The depth of the blue colouring is dependent on how deep in the hierarchy that node
-is. The root is the lightest blue, and the contained merges are progressively
-darker.
+Bubble trees are useful for providing a clear visualization of hierarchal
+data\cite{Boardman2000}. The bubble tree (figure \ref{fig:bubble_tree}) provides a
+very clear picture of where a commit is located in the tree hierarchy. The currently
+selected commit or merge is colored red, while non-selected merges use a shade of
+blue and commits use white.  The depth of the blue colouring is dependent on how
+deep in the hierarchy that node is. The root is the lightest blue, and the contained
+merges are progressively darker.
 
@@
 
-Many of the tools work to address the issues of communication between developers and
-inter-team collaboration. Codebook uses a mining technique to determine the
-developer of a piece of code, the program manager who wrote the specification for
-the code, and the program managers and developers on the team who were working
-together\cite{Begel2010}. This is designed to help collaboration between developers
-on a project while the project is being developed.  In maintaining the code, the
-code has already been written, the original developer many not have seen this piece
-of code in years, and will most likely have moved to another problem. If the
-maintainer is pulling from a part of the kernel that has been recently updated, in
-the event of a zero-day attack for instance, the information about who is developing
-that portion of the kernel is of use; however, if the maintainer is pulling from
-sections that are older, then the maintainer must be able to understand the code
-using the provided tools, not the original developer.
+Many tools work to address the issues in communication between developers in
+inter-team collaboration work. Codebook \cite{Begel2010} uses a data mining
+technique  to determine the developer of a piece of code, the program manager who
+wrote the specification for the code, and the program managers and developers on the
+team who were working together. Hipikat\cite{Cubranic2005} is another tool with a
+focus on communication. Where codebook focuses on developers working on a project,
+Hipikat is focused on enabling easier integration of new developers to a project by
+providing them with easily-searchable artifacts of the changes made. Codebook is
+useful for pairing a contributor with the original developer; however, the developer
+may not have worked with the piece of code in years. A Hipikat program may provide
+more information to the maintainer as it records the artifacts of why certain
+design decisions were made when they were made using other tools like Bugzilla and
+CVS. In both cases, our goal includes providing a summary of the topology of the
+kernel repository through a visual tree.
 
-Hipikat is another tool with a focus on communication issues between developers. It
-is designed for non-collocated teams, helping new developers understand design
-decisions of the system \cite{Cubranic2005}. It records artifacts about the project
-to build a "memory" of the project development that new users can work with to
-understand where and how to make modifications in large projects. The artifacts are
-recorded using tools like bugzilla and CVS. Such a tool could prove useful in
-understanding the changes between a maintainer's current version of an older kernel
-and the changes being made to the current master version of the kernel.
+Most visualizers provide a visual presentation of a certain aspect of a repository.
+Fractal Figures \cite{Ambros2005} uses a unit square to represent a portion of a
+project, then partitions the square based on the proportion of an author's
+contributions to that portion of the project. EPOSee\cite{Burch2005} and Evolution
+Radar\cite{Ambros2009} perform further analysis, determine which files are made
+together, and what changes are made over a sequence of commits, though the goals
+behind these projects is different.
 
-These tools work toward assisting the communications between developers and
-providing a summary of the project. Our goal includes providing a visual summary of
-the topology of the repository, as well as providing a summary of the changes made
-at a location within the repository. EPOSee, Gource, and Gitk are all tools designed
-for visualizing various aspects of repositories.
-
-EPOSee is a tool for visualizing changes that often happen simultaneously in a
-commit or as a sequence of events \cite{Burch2005}. It produces visualizations about
-the confidence that changes are made either together or sequentially. For instance,
-changes in the header file may also indicate that a change was made in the
-implementation file. The visualization tool is designed to visualize implicit rules
-about the order of operations on various files, if too many outliers are detected,
-the project manager may want to restructure the project.  While this is useful for
-new projects, the development model of the Linux kernel is already structured and is
-unlikely to change without good reasons behind it.
+Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work with data
+from CVS repositories. Our goal is to provide information about git repositories,
+specifically the Linux master repository. Fewer tools are available for generating
+visualizations and summaries of git repositories, potentially due to the DAG model
+used by git.
 
@@
 aggregate commit information in merges.
+
+Our tool is primarily aimed at presenting the hierarchical structure of the Linux
+git repository. We use tables for presenting the summarized information of the
+commits and merges, but this information could also be presented in a graphical
+form. Various graphical forms for displaying file and authorship data exist, the
+principal forms being matrix views, city scapes, bar and pie charts, and networks
+\cite{Eick2002}. Any of these data visualization metaphors are applicable to our
+system.
 
@@
 
-\nocite{*}
+% \nocite{*}
 

### 2016-04-26 linvis.tex
Cleaning up a bit
Editing Introduction

@@
 
-Maintainers of older versions of the Linux kernel (such as those used in embedded systems and mobile phones) must keep track of changes to the current version
-of the current in order to determine what commits are relevant to their versions.
-Due to the ratio of growth of the kernel (every year, between 50 and 70 thousand commits get added to it)
-this process requires
-sifting through thousands of
-commits with tools that are unable to filter and effectively visualize projects at
-the scale of the kernel. Tools like Gitk use a directed acyclic graph (DAG) model,
-showing all commits and merges in chronological order by when the commit was
-authored, and not by then it arrives in the official repository.
+Between 50 and 70 thousand commits are added to the Linux kernel per version
+requiring maintainers of older versions of the Linux kernel to sift through
+thousands of commits and merges with tools that are unable to filter and effectively
+visualize projects at the scale of the kernel. Older versions of the kernel are used
+in embedded systems and mobile phones; for security purposes, performance needs, and
+changing hardware requirements, maintainers must be able to understand the changes
+being made in the current version of the kernel in order to produce the necessary
+patches for the older versions of the kernel. Tools like Gitk use a directed acyclic
+graph (DAG) model of the repository, showing all commits and merges in chronological
+order by when the commit was authored, not by when it arrived in the official Linux
+repository.
 
-This representation works in smaller projects; it enables users to see
-when changes are made, when these changes are merged, and how each branch is
-interacting.  
-In large modular projects, like the Linux kernel, the DAG becomes a
-mess of merges and commits (figure \ref{fig:gitk}) losing its visual meaning.  In
-some cases, the Linux kernel is simply too large for the system to generate a
-visualization (eg. Github provides a DAG view for many projects, but is unable to
-display the visualization for projects as large as the Linux kernel (figure
-\ref{fig:gitfail}).
-Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}, 
-a commit takes a median of 30 days from the moment it is authored to the time it arrives to the official repository
-This means that, if a person inspected the commits in this repository today, tomorrow it might contain  many more
-commits that were authored long before. However, distinguishing these new commits is not trivial.
+This representation works in smaller projects; it enables users to see when changes
+are made, when these changes are merged, how each branch is interacting, and the
+point where a branch forks from the master branch. In large modular projects, like
+the Linux kernel, the DAG becomes a mess of merges and commits (figure
+\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel is simply
+too large for the system to generate a visualization; Github provides a DAG view for
+many projects, but is unable to display the visualization for projects at the scale
+of the Linux of the kernel (figure \ref{fig:gitfail}).  Between 60 and 70k new
+commits are created for the Linux project every year; According to our previous
+work\cite{German2015}, a commit takes a median of 30 days from the time it is
+authored until it arrives in the official repository. The snapshot of the kernel
+tomorrow may be different than the snapshot from today, containing new commits
+authored in the past; distinguishing these new commits from the commits in the
+snapshot from today is not trivial.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/gitk.png}
-        \caption{A sample DAG view of commits from Linux in Gitk. Commits point towards their ancestors, and there is no clear indication of the way the travel
-          to reach Master. In fact, neither Gitk nor git is capable of showing what commits were done in master.}
+        \caption{A sample DAG view of commits from Linux in Gitk. Commits point
+                towards their ancestors, and there is no clear indication of the way
+                the travel to reach Master. In fact, neither Gitk nor git is capable 
+                of showing what commits were done in master.}
         \label{fig:gitk}
@@
         \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its relationship to other forks.}
+        \caption{Github Failure Message showing the DAG of the Linux Project and its
+                relationship to other forks.}
         \label{fig:gitfail}
@@
 
-One of the challenges with visualizing the arrival of commits to a repository is that
-in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master).
-To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
-would prefer to know the path that a commit followed to reach the master repository.
-Hence, tracing the path that any commit followed to master would imply that for any given merge, it would be possible to know what commits were merged. A user could
-inspect the commits that have arrived into master within a given time frame by simply checking which commits where merged during that time frame.
+One major challenge with visualizing the arrival of commits to a repository is that
+Git does not store the date that a commit was merged into another branch (including
+master). To complicate the problem, the DAG has reference to the ancestors of a
+commit (a model necessary for the operation of Git), but developers would prefer
+knowing the path a commit followed to reach the master repository. Tracing a path
+that any commit followed to the master repository would imply that for any given
+merge, it would be possible to know which commits were merged. A user could inspect
+the commits that arrived into master within a given time-frame by checking which
+commits were merged during that time-frame.
 
-% \dmg{I am going to describe the method we used here... }
+This paper makes two contributions; first, we describe a method of converting the
+DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
+that represents the path used by a commit to reach master; second, we present a
+method to inspect and visualize the history of merges in the Linux project using the
+merge-tree model.
 
-% German et al. recorded the metadata of every commit and merge in the Linux git
-% repository starting in 2011 \cite{German2015}, including chronological data that is
-% not stored by git. We are able to use this information to determine when a commit
-% has been merged into the kernel, regardless of rebases preceding and proceeding that
-% point in time. This information is then converted by a heuristic method, described
-% by German, to generate sets of trees, which we use as our core model instead of
-% using the DAG. Each tree is rooted by the top-level merge performed by Linus, which
-% merges the branches or a set of commits into the kernel.
-
- The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{merge-tree of the
- repository}) that represents the path used by
- any commit to reach master. Second, 
- we present a method to inspect and visualize the history of merges in the  Linux project using the tree-model of the repository
-
-These methods and visualizations have been implemented in a web-based
-tool called \tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}. Our visualizations and tool provide information about the location of any given
+These methods and visualizations are implemented in a web-based tool called
+\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
+Our visualizations and tool provide information about the location of any given
 commit in its respective merge-tree, the files edited, the modules edited, and the
@@
 
-\evan{pdfs are not included in the repository}
 \begin{figure}[htbp]
@@
 
-The results of this verification show that this heuristic detected no incorrect merges since 2013, two incorrect merges in 2012, and two in 2011. However, due
-to fast-forward commits performed in Linus Torvalds repository, this heuristic breaks apart on Oct 15, 2009, and no longer works for older merges.
+The results of this verification show that this heuristic detected no incorrect merges since 2013, two incorrect merges in 2012, and two in 2011. However, this
+heuristic breaks apart on Oct 15, 2009 due
+to the presence of so-called foxtrot commits that confound the true Master branch of a repository \footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and no longer works for older merges.
 
@@
 
-% \dmg{I would remove the command line shell, and use the space for related work}
-
-% We came up with two designs. We recognized that the structure of the repository was
-% similar to the wide and shallow tree-structure of a file system. We began with a
-% command-line shell interface. Then transitioned to a web-based tool where we looked
-% beyond the file system structure. In this section, we describe the progression of
-% our tool design, showing the benefits and drawbacks of each component.
-
-% \subsection{Command-line Shell}
-
-% Our initial intuition was to work with the commits and merges as directory trees
-% where the commits are files and the merges are directories. There are many
-% similarities between the tree structure of a file system and the tree structure of
-% the repository. Both trees are generally shallow, but very wide. File systems have
-% been studied for quite some time and there are many tools capable of visualizing and
-% working with the file structure.
-
-% The first tool we implemented was a small command-line shell (figure
-% \ref{fig:shell}) that allows users to navigate through the commits and merges as one
-% might navigate through files and directories. A user could navigate toward the
-% leaves of the tree by using the \verb|cd| command followed by the commit ID of the
-% next merge. The user could then navigate toward the root by issuing \verb|cd ../|
-% as they would in a file system. Users can use \verb|ls| to view what commits and
-% merges exist in the current working merge, then use \verb|cat| to view the log
-% preview of a commit or \verb|more| to see the preview and how many lines were added
-% and removed from the files in the commit.
-
-% \begin{figure}
-%         \centering
-%         \includegraphics[width=0.47\textwidth]{figures/shell.png}
-%         \caption{Shell Design}
-%         \label{fig:shell}
-% \end{figure}
-
-% The shell does not meet our design goals. It only provides users with the ability to
-% navigate from the root. If there is a commit or merge that they would like to
-% investigate, they must find it with another tool and navigate through the tree to
-% find it. Furthermore, it does not aggregate information in the merges, so users must
-% manually aggregate the metadata. Another issue with this design is that it presents
-% the users with the commit hashes, which are quite long and unpleasant to remember
-% and type into the shell, but also do not provide any quick explanation of what the
-% commit or merge may contain.
-
-% \subsection{Web-based Tool}
-
-To navigate and inspect the merge-tree view of the Kernel we
-created a web-based tool called \tool. Creating a web-based
-tool enables users to use the system without having to install additional software
-or store a large database, making it more accessible, more easily maintainable, and
-platform independent. 
-% A web-based tool can leverage interactive means of
-% navigation and provide better explanations of commits and merges than the
-% command-line interface.
-\tool uses the following mechanisms to reach our goals of better navigation and
-better explanation of the changes.
+To navigate and inspect the merge-tree view of the Kernel we created a web-based
+tool called \tool. Creating a web-based tool enables users to use the system without
+having to install additional software or store a large database, making it more
+accessible, more easily maintainable, and platform independent. \tool uses the
+following mechanisms to reach our goals of better navigation and better explanation
+of the changes.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
-        \caption{Entire Reingold-Tilford Tree of a merge. This merge by Linus (a992bf836f9c3039a16f4bd068d161c86c6c3e2c) contains 1722 includes non-merge commits
+        \caption{Entire Reingold-Tilford Tree of a merge. This merge by Linus (73287a43cc79ca06629a88d1a199cd283f42456a) contains 1722 includes non-merge commits
                 touching 239 modules and 1525 files.}
@@
 \subsubsection{Bubble Tree}
-
-\evantodo{we need reference to bubble trees}
 

### 2016-04-26 linvis.tex
Merge-Tree model edited

--- 2016-04-26 linvis.tex
@@
 \newif\ifdraft
-%\drafttrue
+\drafttrue
 
@@
 
+\evantodo {
+        \begin{enumerate}
+                \item Add simplified abstract -- From email
+                \item replace figure reference with Figure~\\ref
+                \item Check captions
+                \item gitk to Gitk
+                \item git to Git
+                \item diff to patch
+                \item id to ID
+        \end{enumerate}
+}
+
 Between 50 and 70 thousand commits are added to the Linux kernel per version
@@
 
-Git uses a Directed Acyclic Graph (DAG) as its main data model. In this model, a commit has one or more ordered parents (ancestors), except the root commits of a repository which
-do not have ancestors--Linux git repository has two such commits. Commits are divided into merge commits (they merge two or more branches) and non-merge
-commits. Non-merge commits have only one parent, while merge commits have two or more. The order of the parents matter: the first parent is the branch in which
-the merge is being done; while the rest indicate the branches being merged (git allows any number of branches to be merged at the same time). Their order in the
-parents relationship indicate the order in which the branches are being merged (second parent first, third parent second, etc.).
+Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
+commit has one or more ordered parents (ancestors), except the root commits of a
+repository which do not have ancestors--Linux git repository has two such commits.
+Commits are divided into merge commits, merging two or more branches, and non-merge
+commits. Non-merge commits have only one parent, while merge commits have two or
+more. The order of the parents matter: the first parent is the branch in which the
+merge is being done, while the rest indicate the branches being merged (git allows
+merging multiple branches simultaneously). The second parent is the first branch
+merged, the third parent is the second branched merged.
 
-In Git, commits are immutable. Once they are created, they never change. Git allows operations to alter commits or to reorder them. However, in such case, the
-commit changes its commit id, effectively becoming a new commit. This means that commits do not record the branches or repositories they have traversed in their
-way to master.
-For instance, assume the sequence of commits  exemplified in Figure~\ref{fig:repoEvents}. In this case commits are performed in various repos and branches. The
-DAG representation of these commits is shown in Figure~\ref{fig:repoDAG}. Notice that this DAG loses information about what repository is Master (in fact, it
-loses all information about what repositories any commit has passed through). The merge-tree view of this DAG is shown in Figure~
-\ref{fig:repoTree}. Note that the direction of the edges of the DAG has been inverted. Instead of pointing to the ancestors, it points to its successors (the path to
-Master). Also, the DAG has been simplified, and it only shows for any commit, one edge (its path to Master).
+Once a commit is created, it is never changed. Git allows operations to alter
+commits or reorder them, but it changes the commit id in the process, effectively
+making a new commit. This makes commits unable to record the traversal of merges
+from the commit to the merge into the master branch.
+
+A short example: assume the of commits represented in figure \ref{fig:repoEvents}
+shows the sequence of events in a repository. In this case, commits are performed in
+various repositories and branches. The DAG representation of the commits is shown in
+figure \ref{fig:repoDAG}. Notice that the DAG loses information about the master
+branch and the repository that the master branch is part of. The merge-tree view of
+this DAG is visible in figure \ref{fig:repoTree}. Note that the direction of the
+edges of the DAG has been inverted, instead of pointing from the child to the
+ancestors, it points from the ancestor to its successors, forming a path to the
+master branch. Also note that the DAG has been simplified, showing only a single
+edge on the path to master for any commit.
 
@@
         \includegraphics[width=\columnwidth]{figures/events.pdf}
-        \caption{Example of a sequence of events performed in different repositories. The horizontal axis represents time. Each horizontal section represents a
-                different branch and/or repository. Each commit points to its ancestor.}
+        \caption{Example of a sequence of events performed in different
+                repositories. The horizontal axis represents time. Each horizontal
+                section represents a different branch and/or repository. Each commit
+                points to its ancestor.}
         \label{fig:repoEvents}
@@
         \includegraphics[width=.8\columnwidth]{figures/dag.pdf}
-        \caption{DAG representation of the commits represented in Figure~\ref{fig:repoEvents}. It loses information about what repository the commit is performed and
-                through which repositories it has passed in its way to Master. In fact, the DAG does not even know what Master is.}
+        \caption{DAG representation of the commits represented in
+                Figure~\ref{fig:repoEvents}. It loses information about what
+                repository the commit is performed and through which repositories it
+                has passed in its way to Master. In fact, the DAG does not even know
+                what Master is.}
         \label{fig:repoDAG}
@@
         \includegraphics[width=\columnwidth]{figures/tree.pdf}
-        \caption{Merge-tree view of the commits  represented in Figure~\ref{fig:repoEvents} showing the path they followed to reach master. In this model the successors of
-                each commit represents the path followed by that commit to reach Master.}
+        \caption{Merge-tree view of the commits  represented in
+                Figure~\ref{fig:repoEvents} showing the path they followed to reach
+                master. In this model the successors of each commit represents the
+                path followed by that commit to reach Master.}
         \label{fig:repoTree}
@@
 
-Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it feasible to
-compute its merge-tree. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
-assertion if previous research~\cite{German2015}. We have developed a heuristic that is presented in Algorithm~\ref{fig:alg}. In a nutshell, first identify the commits directly made
-to the master branch. Then,
-recursively, for
-every commit, using the DAG, determine the shortest path (in terms of time) from such commit to master using the inverted DAG.
+Computing the merge-tree from a DAG for any repository may not be possible; however,
+certain features of the development process of Linux make it feasible to compute the
+merge-tree for the Linux repository. First, the master branch of Linux is maintained
+by Linus Torvalds, and only Linus has write access to it. We have verified this
+assertion in previous research~\cite{German2015}. We have developed a heuristic that
+is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first identifies
+the commits made directly to the master branch; whereafter it recursively determines
+the shortest path (in terms of time), using the DAG, from each commit to master
+using the inverted DAG.
 
@@
 
-Merges that do not have conflicts provide information to verify this heuristic. If a merge does not contain a conflict, it records a summary of the commits that
-it merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a list of the first 20 non-merge commits in the merge, including their one-line
-log description, the full logs of the merge commits that merge this subset, and the total number of non-merge commits in the merge.
+Merges that do not have conflicts provide information to verify this heuristic. If a
+merge does not contain a conflict, it records a summary of the commits that it
+merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a
+list of the first 20 non-merge commits in the merge, including their one-line log
+description, the full logs of the merge commits that merge this subset, and the
+total number of non-merge commits in the merge.
 
@@
         \end{verbatim}}
-        \caption{Example of how merges record a subset of commits being merged. The commit only shows the first 20 one line log messages for the 24 non-merge commits
-                it merged. The ending ... is part of the log and represents that other commits were merged.}
+        \caption{Example of how merges record a subset of commits being merged. The
+                commit only shows the first 20 one line log messages for the 24
+                non-merge commits it merged. The ending ... is part of the log and
+                represents that other commits were merged.}
         \label{fig:sampleMerge}
@@
 
-We use this information to evaluate the accuracy of the merge tree-model extracted from the DAG. We therefore compare, for commits with no conflict: a) the number of
-non-merge commits, b) the summaries of the commits available in the log, and c) the summaries of the merges included.
+We use this information to evaluate the accuracy of the merge-tree model extracted
+from the DAG. For commits with no conflicts, we compare: 1) the number of non-merge
+commits, 2) the summaries of the commits available in the log, and 3) the summaries
+of the merges included.
 
-The results of this verification show that this heuristic detected no incorrect merges since 2013, two incorrect merges in 2012, and two in 2011. However, this
-heuristic breaks apart on Oct 15, 2009 due
-to the presence of so-called foxtrot commits that confound the true Master branch of a repository \footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and no longer works for older merges.
+The verification detected no incorrect merges since 2013, two incorrect merges in
+2012, and two in 2011; however, the heuristic breaks in October 15 2009 due to the
+presence of \textit{foxtrot} commit, which confounds the true master branch of a
+repository \footnote{See
+        \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a
+        description of the issue}, and will not work for older merges.
 
@@
 ``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
-correctly follow this format and the hueristic approach fails. As with the Files panels,
+correctly follow this format and the heuristic approach fails. As with the Files panels,
 non-merge commits show their corresponding information, but for merge commits we aggregate all the modules changed
@@
 master branch from the DAG provided by git. This may be because more information is
-required to generate the tree-model than what is stored by git. However, there has
-been a lot of work in providing visualizations of various repositories.
+required to generate the merge-tree model than what is stored by git. However, there
+has been a lot of work in providing visualizations of various repositories.
 

### 2016-04-26 linvis.tex
Visualizing Merge Tree Section

@@
                 \item id to ID
+                \item Master to master
         \end{enumerate}
@@
 2012, and two in 2011; however, the heuristic breaks in October 15 2009 due to the
-presence of \textit{foxtrot} commit, which confounds the true master branch of a
-repository \footnote{See
-        \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a
-        description of the issue}, and will not work for older merges.
+presence of \textit{foxtrot} commits, which confounds the true master branch of a
+repository
+\footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and will not work for older merges.
 
@@
 
-The goal of \tool was to simplify the  navigation of the kernel commit information, specifically by focusing on merges.
-This is done by leveraging the merge-tree view of Linux to inspect how commits are merged in their way to master.
+The goal of \tool is to simplify the navigation of the kernel commit information,
+specifically focusing on merges. This is done by leveraging the merge-tree view to
+inspect how commits are merged on the path to the master repository.
 
@@
 These are users that are maintaining a section of the kernel and would like to pick
-a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
-reducing the amount of re-implementation work. For these users, it is important to have the
-ability to aggregate metadata about files and modules being effected by the merge.
-It is also important for these users to be able to navigate from the root of the merge-tree toward the
-leaves.
+a merge (including all the commits that it merges) and merge it directly into their
+current repository. This is useful for reducing the amount of re-implementation
+work. For these users, it is important to have the ability to aggregate metadata
+about files and modules being effected by the merge.  It is also important for these
+users to be able to navigate from the root of the merge-tree toward the leaves.
 
@@
 These are users that start with a known merge or commit and would like to see what
-other changes are being made in commits that are in the same merge, including knowing the merge-tree they belong to. This is useful
-to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
-This is primarily for maintainers that need to perform
-some specific cherry picking of commits. We
-must provide these users a mechanism for navigating from a single commit toward the
-master, allowing them to see other commits that might be related to it.
+other changes are being made in commits that are in the same merge, including
+knowing the merge-tree they belong to. This is useful to see what other commits are
+related to the current commit and how they get collated into merges that eventually
+end in master.  This is primarily for maintainers that need to perform some specific
+cherry picking of commits. We must provide these users a mechanism for navigating
+from a single commit toward the master, allowing them to see other commits that
+might be related to it.
 
@@
 
-In our visualizations, we leverage the Merge-Tree model described in Section~\ref{sec:mergetree}.
-In this model each commit is either already in Master, or is part of a tree which is rooted in the merge that merged it into Master.
-In that case, each
-commit (whether merge or non-merge) has only one successor at most (the root of each tree has none, as it was made by Linus Torvalds directly into Master).
-Commits contain the
-metadata for the changes made. This metadata includes the files changed, the lines
-added and removed from each file, the author, the date the commit was merged into
-the merge that led to being merged into the kernel, and the date the commit was
-authored, the patch, and the commit log. Merges contain less metadata, only storing
-the author of the merge, the log, the commit date, and the author date, and potentially, changes necessary to address conflicts during the merge. The details
-of the model are outlined in \cite{German2015}.
+In our visualizations, we leverage the Merge-Tree model described in
+Section~\ref{sec:mergetree}. In this model each commit is either already in Master,
+or is part of a tree which is rooted in the merge that merged it into Master.  Each
+commit, whether a merge or non-merge, has only one successor; the root of each tree
+has none as it was made by Linus Torvalds directly into the master branch. Non-merge
+Commits contain the metadata for the changes made. This metadata includes the files
+changed, the lines added and removed from each file, the author, the date the commit
+was merged into the merge that led to being merged into the kernel, the date the
+commit was authored, the patch, and the commit log. Merges contain less metadata,
+only storing the author of the merge, the log, the commit date, and the author date,
+and potentially, changes necessary to address conflicts during the merge. The
+details of the model are outlined in \cite{German2015}.
 

### 2016-04-26 linvis.tex
Minor capitalization changes

@@
         versions of the kernel. Due to security, performance, and changing hardware
-        needs, maintainers must understand what changes (commits) are added the current version of the kernel
-        since the last time they inspected it in order to make the necessary
-        patches.
+        needs, maintainers must understand what changes (commits) are added to the
+        current version of the kernel since the last time they inspected it in order
+        to make the necessary patches.
 
         Current tools provide information about repositories through the directed
-        acyclic graph (DAG) of the repostory, which is helpful for smaller projects. However, with
-        the scale and number of branches in the kernel the DAG becomes overwhelming
-        very quickly. Furthermore, the DAG contains every ancestor of every commit, while maintainers are
-        more interested in how and when a commmit arrives to Linux official repository.
+        acyclic graph (DAG) of the repository, which is helpful for smaller
+        projects. However, with the scale and number of branches in the kernel the
+        DAG becomes overwhelming very quickly. Furthermore, the DAG contains every
+        ancestor of every commit, while maintainers are more interested in how and
+        when a commit arrives to the official Linux repository.
 
-        In this paper, we present \tool, a tool that uses a tree-based model to construct a clearer summary of the changes made. \tool is designed to allow
-        users to explore the topology of the repository for a given release of the kernel and view an aggregated summary of the changes made.
+        In this paper, we propose the merge-tree view, a simplified transformation
+        of the DAG of the Linux git repository that shows the way in which commits
+        are merged into the master branch of Linux. Using the merge-tree view we
+        build \tool, a tool that is designed to allow users to explore how commits
+        are merged into the Linux kernel.
 
@@
         \begin{enumerate}
-                \item Add simplified abstract -- From email
-                \item replace figure reference with Figure~\\ref
                 \item Check captions
-                \item gitk to Gitk
-                \item git to Git
-                \item diff to patch
-                \item id to ID
-                \item Master to master
         \end{enumerate}
@@
 
-Between 50 and 70 thousand commits are added to the Linux kernel per version
+Between 50k and 70k commits are added to the Linux kernel per version
 requiring maintainers of older versions of the Linux kernel to sift through
@@
 point where a branch forks from the master branch. In large modular projects, like
-the Linux kernel, the DAG becomes a mess of merges and commits (figure
-\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel is simply
-too large for the system to generate a visualization; Github provides a DAG view for
-many projects, but is unable to display the visualization for projects at the scale
-of the Linux of the kernel (figure \ref{fig:gitfail}).  Between 60 and 70k new
-commits are created for the Linux project every year; According to our previous
-work\cite{German2015}, a commit takes a median of 30 days from the time it is
-authored until it arrives in the official repository. The snapshot of the kernel
+the Linux kernel, the DAG becomes a mess of merges and commits
+(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel
+is simply too large for the system to generate a visualization; Github provides a
+DAG view for many projects, but is unable to display the visualization for projects
+at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60 and
+70k new commits are created for the Linux project every year; According to our
+previous work\cite{German2015}, a commit takes a median of 30 days from the time it
+is authored until it arrives in the official repository. The snapshot of the kernel
 tomorrow may be different than the snapshot from today, containing new commits
@@
                 towards their ancestors, and there is no clear indication of the way
-                the travel to reach Master. In fact, neither Gitk nor git is capable
+                the travel to reach master. In fact, neither Gitk nor Git is capable
                 of showing what commits were done in master.}
@@
 more. The order of the parents matter: the first parent is the branch in which the
-merge is being done, while the rest indicate the branches being merged (git allows
+merge is being done, while the rest indicate the branches being merged (Git allows
 merging multiple branches simultaneously). The second parent is the first branch
@@
 Once a commit is created, it is never changed. Git allows operations to alter
-commits or reorder them, but it changes the commit id in the process, effectively
+commits or reorder them, but it changes the commit ID in the process, effectively
 making a new commit. This makes commits unable to record the traversal of merges
@@
 
-A short example: assume the of commits represented in figure \ref{fig:repoEvents}
+A short example: assume the of commits represented in Figure~\ref{fig:repoEvents}
 shows the sequence of events in a repository. In this case, commits are performed in
 various repositories and branches. The DAG representation of the commits is shown in
-figure \ref{fig:repoDAG}. Notice that the DAG loses information about the master
+Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about the master
 branch and the repository that the master branch is part of. The merge-tree view of
-this DAG is visible in figure \ref{fig:repoTree}. Note that the direction of the
+this DAG is visible in Figure~\ref{fig:repoTree}. Note that the direction of the
 edges of the DAG has been inverted, instead of pointing from the child to the
@@
                 repository the commit is performed and through which repositories it
-                has passed in its way to Master. In fact, the DAG does not even know
-                what Master is.}
+                has passed in its way to master. In fact, the DAG does not even know
+                what master is.}
         \label{fig:repoDAG}
@@
                 master. In this model the successors of each commit represents the
-                path followed by that commit to reach Master.}
+                path followed by that commit to reach master.}
         \label{fig:repoTree}
@@
 In our visualizations, we leverage the Merge-Tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already in Master,
-or is part of a tree which is rooted in the merge that merged it into Master.  Each
+Section~\ref{sec:mergetree}. In this model each commit is either already in master,
+or is part of a tree which is rooted in the merge that merged it into master.  Each
 commit, whether a merge or non-merge, has only one successor; the root of each tree
@@
 \begin{itemize}
-        \item Filter by searching commits by when they are merged into Master.
+        \item Filter by searching
         \item View files edited
@@
 
-\subsection{Searching by when they are merged into Master}
+\subsection{Searching}
 
@@
         \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View. It allows finding commits that were \textbf{merged} in a given period or by commit it.}
+        \caption{Search View. It allows finding commits that were \textbf{merged} in
+                a given period or by commit it.}
         \label{fig:search}
@@
 
-Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and merges that are irrelevant. The search mechanism breaks down the results by
-release version.  A user can further narrow down the search by \textbf{specifying a range of dates} in which such commits \textbf{were merged by Linus into
-        Master, not
-        when the commits were created.} This distinction is important. We have observed commits that take years
-to arrive into Master after they were originally committed.
+Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
+merges that are irrelevant. The search mechanism breaks down the results by release
+version. A user can further narrow down the search by \textbf{specifying a range of
+dates} in which such commits \textbf{were merged by Linus into master, not when the
+commits were created.} This distinction is important. We have observed commits that
+have taken years to arrive into master after they were originally committed.
 
-A user may then provide a search text, filtering by the author name,
-the commit ID, or keyword from the log. Any part of the author name may show up in
-the results, including searching by email address.
+A user may then provide a search text, filtering by the author name, the commit ID,
+or keyword from the log. Any part of the author name may show up in the results,
+including searching by email address.
 
-\dmg{What is this commit? any commit id, or a merge into Master? Unclear.}
-If providing commit ID must be specified
-from the first character to the last character. For example, the commit
-`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
-commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
-each of the features of \tool.
+If providing commit ID must be specified from the first character to the last
+character. For example, the commit `c267548755a184ef97301071300c1739a564e135' is
+returned when the user searches for a commit ID of `c26' in the 3.10 Linux kernel.
+This commit is a non-merge commit performed, in one of the largest merge-trees of
+the kernel; however, this feature will work with any commit ID.  We will use this
+commit for showing each of the features of \tool.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
-        \caption{Search Results. Each is a commit merged in the desired merged window.}
+        \caption{Search Results. Each is a commit merged in the desired merged
+                window.}
         \label{fig:results}
@@
 eventually is merged into the kernel, the commit date and authored date will be
-different. We can see this in the first and third entries in figure
-\ref{fig:results}.
+different. We can see this in the first entry in Figure~\ref{fig:results}.
 
@@
 
-The first tab displays the full commit log (figure \ref{fig:message}). From this, a
-user is able to see what they would see had they searched for the commit using git
+The first tab displays the full commit log (Figure~\ref{fig:message}). From this, a
+user is able to see what they would see had they searched for the commit using Git
 log. This doesn't provide additional information to the other tools, but helps to
@@
         \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
-        \caption{Panels showing the files modified by all the commits that are part of this merge.}
+        \caption{Panels showing the files modified by all the commits that are part
+                of this merge.}
         \label{fig:files}
@@
 
-The second tab is the files tab (figure \ref{fig:files}). This tab provides information on what files have been edited, how many lines were added, and how many
-lines were removed in a given commit. For non-merges, this functionality is similar to the other tools available. Our tree-based design model allows us to
-extend this functionality to merges by aggregating information about all the commits that are children of the merge in the merge-tree (which the other tools are
-unable to show). To find the number of lines added to a file in a merge, we take the sum of the lines added to that file in each of the children of that
-merge. We do the same for calculating the number of lines removed.
+The second tab is the files tab (Figure~\ref{fig:files}). This tab provides
+information on what files have been edited, how many lines were added, and how many
+lines were removed in a given commit. For non-merges, this functionality is similar
+to the other tools available. Our tree-based design model allows us to extend this
+functionality to merges by aggregating information about all the commits that are
+children of the merge in the merge-tree, which other tools are unable to show. To
+find the number of lines added to a file in a merge, we take the sum of the lines
+added to that file in each of the children of that merge. We do the same for
+calculating the number of lines removed.
+
 % This could be extended to use the patch information to determine when a line has been
@@
 
-The modules tab (figure \ref{fig:modules}) shows the modules that are contained
-within the commit. Modules are not natively recognized by git, and are not going to
-be present in all repositories.  In the Linux repository, authors put the module
-they are working on in the one-line preview of the log-message; for example: the log \emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module
-(the coverage testing tool of the kernel). We heuristically
-extract the module by taking all text in the log preview of commits until the first
-colon.  Modules are logical partitions of the information in the kernel. Depending
-on where the author was working, modules can be general, such as ``bluetooth'' and
-``wireless'', or can be quite specific for individual hardware, such as
-``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
-correctly follow this format and the heuristic approach fails. As with the Files panels,
-non-merge commits show their corresponding information, but for merge commits we aggregate all the modules changed
-in all the commits that are part of that merge.
-The output of this view is shown in Figure~\ref{fig:modules}.
+The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
+within the commit. Modules are not natively recognized by Git, and are not going to
+be present in all repositories. In the Linux repository, authors put the module they
+are working on in the one-line preview of the log-message; for example: the log
+\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
+coverage testing tool of the kernel). We heuristically extract the module by taking
+all text in the log preview of commits until the first colon.  Modules are logical
+partitions of the information in the kernel. Depending on where the author was
+working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
+quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
+few cases, the author of a commit does not correctly follow this format and the
+heuristic approach fails. As with the Files panels, non-merge commits show their
+corresponding information, but for merge commits we aggregate all the modules
+changed in all the commits that are part of that merge.  The output of this view is
+shown in Figure~\ref{fig:modules}.
 
@@
 navigation of the commits within the merge-tree that is rooted in the current merge.
-It also provides a clear topological view of the merge and the submerges it includes.We have experimented with various tree designs to find a design that
-allows for easy navigation and visualization of both large and small trees.  We discuss this panel in the next subsection.
+It also provides a clear topological view of the merge and the submerges it
+includes. We have experimented with various tree designs to find a design that
+allows for easy navigation and visualization of both large and small trees. We
+discuss this panel in the next subsection.
 
@@
 
-The merge-tree view is what makes \tool unique to other tools that inspect the DAG of a git repository. With it, a user can inspect how commits are merged
-on their way to Master. We have experimented with various types of trees:
+The merge-tree view is what makes \tool unique to other tools that inspect the DAG
+of a Git repository. With it, a user can inspect how commits are merged on their way
+to the master branch. We have experimented with various types of trees:
 
@@
 
-The list tree viewer (figure \ref{fig:list_tree}) is in the form of nested lists,
+The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of nested lists,
 and is designed to more closely model the tree view found in file browsers. This
@@
 
-To improve the visualization of the merge-tree, we first transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master.
-However, we keep the topological order of the commits in the tree-merge as the order of the
-leaves for a given merge (e.g. if merge-tree contains two commits A, B merged into C, then the simplified tree contains links B and C directly to A, but it
-preserves the order of B with respect to C by showing B above C.
-This allows the clustering of commits along their next merge, and avoid the space needed to
-present long sequences of commits without merges in between.
-
-The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:zoomed_tree} and \ref{fig:tree}) allows the visualization and navigation
-of the entire merge-tree in an intuitive representation of the tree. This illustrates a clear notion of
-root and leaves, and how to navigate in either direction. Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
-thousands of commits and merges. While the tree is capable of producing a
-visualization, it becomes much more difficult to understand. For example the merge 73287a43cc79ca06629a88d1a199cd283f42456a performed by Linus Torvalds in May
-1, 2013, merged 1722 non-merge commits and 91 merges. footnote{This commit can be inspected at \url{http://li.turingmachine.org/commits/73287a43cc79ca06629a88d1a199cd283f42456a}}
+The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:zoomed_tree} and
+\ref{fig:tree}) allows the visualization and navigation of the entire merge-tree in
+an intuitive representation of the tree.  This illustrates a clear notion of root
+and leaves, and how to navigate in either direction. Some of the merge trees are
+very large (Figure~\ref{fig:zoomed_tree}), containing thousands of commits and
+merges. While the tree is capable of producing a visualization, it becomes much more
+difficult to understand. For example the merge
+73287a43cc79ca06629a88d1a199cd283f42456a performed by Linus Torvalds in May 1, 2013,
+merged 1722 non-merge commits and 91 merges. \footnote{
+        This commit can be inspected at
+        \url{http://li.turingmachine.org/commits/73287a43cc79ca06629a88d1a199cd283f42456a}}
 
@@
 
-Clicking on subtrees will cause them to collapse, only showing parts of the tree
-that are of interest to the user. Furthermore, the user can zoom the tree by
-scrolling the mouse wheel to see more of the topology of the tree, or to see the
-details of a given part of the tree.
-
-\dmg{do you have a figure that shows the collapsed subtrees? Try to use the same HUGE commit}
+The tree can be further extended to include the collapsing of merge nodes. Clicking
+on merge nodes will cause them to collapse, only showing parts of the tree that are
+of interest to the user.Furthermore, the user can zoom the tree by scrolling the
+mouse wheel to see more of the topology of the tree, or to see the details of a
+given part of the tree.
 
@@
 Bubble trees are useful for providing a clear visualization of hierarchal
-data\cite{Boardman2000}. The bubble tree (figure \ref{fig:bubble_tree}) provides a
+data\cite{Boardman2000}. The bubble tree (Figure~\ref{fig:bubble_tree}) provides a
 very clear picture of where a commit is located in the tree hierarchy. The currently
@@
 
-\tool was built using Nginx as the main web server, postgresql as the dbms, and Flask as the framework for generating the dynamic content.
+\tool was built using Nginx as the main web server, postgresql as the DBMS, and
+Flask as the framework for generating the dynamic content.
 
@@
 
-From a developers point of view, the are two major disadvantages of the DAG model of git: a) its edges point backwards, i.e. commits point to their ancestor,
-not to the commit that succeeds them in their way to be integrated; and b) the DAG contains many more edges that are necessary to understand how integration
-happened. We addressed these two issues with the creation of the merge-tree model of the DAG of Linux. Effectively the merge-tree view recoveres the details of each merged
-branch into master.
+From a developers point of view, the are two major disadvantages of the DAG model of
+Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
+commit that succeeds them on the path to being integrated; and b) the DAG contains
+many more edges that are necessary to understand how integration occurred. We
+addressed these two issues with the creation of the merge-tree model of the DAG of
+Linux. Effectively the merge-tree view recovers the details of each merged branch
+into master.
 
-In our experience, no other git repository reaches the level of DAG complexity that Linux has. In most git repositories  every a merge into master does not
-merge other merges, and most merges bring very few new commits with them. However, even in those cases, the merge-tree will provide a valuable summary of how
-commits have been integrated into master, specially since the time of integration might be very different from the time the commits were created.
+In our experience, no other Git repository reaches the level of DAG complexity that
+Linux has. In most Git repositories, merges are not nested, all merges merge
+directly into the master repository, and these merges only merge few commits. Even
+in these simplified cases, the merge-tree can provide a valuable summary of how
+commits are integrated into the master branch, specifically since the time of
+integration may be very different from the time the commits were authored.
 
-The biggest challenge, however, is computing the merge-tree. Our heuristic is applicable \textbf{only} to the Linux kernel, and only since July 2009 and it is
-very likely it will not work for other repositories, specially those that have fast-forward commits into Master (as the majority of repositories). One way to
-address this is lack of data is by monitoring at short intervals the arrival of commits into Master (as we did in \cite{German2015}, i.e.
-\emph{continuously mining} a git repo). This however, requires planning.
+The biggest challenge, however, is computing the merge-tree. Our heuristic
+\textbf{only} applies to the Linux kernel repository, and only since July 2009. It
+is very likely that the heuristic will not work with other repository, specially
+those with fast-forward commits into the master branch (a majority of repositories).
+We can address the lack of information by monitoring the state of the repository at
+short intervals as we did in \cite{German2015}. This continuous mining of the
+repository requires foresight and planning.
 
-Regarding \tool. Because it leverages the merge-tree of the kernel, it provides mechanisms and visualizations that other tools are unable to
-produce.  It gives us the ability to see how commits are being merged
-into the kernel, allowing us to better understand how code is integrated into the kernel.
+Because \tool leverages the merge-tree model of the kernel, it provides mechanisms
+and visualizations that other tools are unable to produce.It gives us the ability to
+see how commits are being merged into the kernel, allowing us to better understand
+how code is integrated into the kernel.
 
@@
         \includegraphics[width=0.97\textwidth]{figures/042dd_tree.png}
-        \caption{Merge Tree View \dmg{this figure is unreadable. Any chance you can make the text bigger?}}
+        \caption{Merge Tree View \dmg{this Figure is unreadable. Any chance you can make the text bigger?}}
         \label{fig:tree_view}
@@
 
-Our goal was to build a tool to enable maintainers to effectively navigate and browse
-the changes performed to the kernel over the period of a release.
-Achieving this goal includes removing information that does not pertain to the area
-of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
-figure \ref{fig:tree_view} we can visually compare the results returned from Gitk
+Our goal was to build a tool to enable maintainers to effectively navigate and
+browse the changes performed to the kernel over the period of a release.  Achieving
+this goal includes removing information that does not pertain to the area of the
+kernel that the user is interested in. In Figure~\ref{fig:dag_view} and
+Figure~\ref{fig:tree_view} we can visually compare the results returned from Gitk
 and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. Note first that Gitk has no way to show that this is a merge into Master.
+from kernel version 3.10. Note that Gitk has no way to show or recognize that this
+is a merge into the master branch.
 
-The log information to this merge-commit in both of these
-figures is identical, but the presentation drastically changes our ability to
-comprehend what we are seeing. The primary difference is the removal of irrelevant
-information. In Gitk (\ref{fig:dag_view}), the visualization results include commits
-and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
+The log information to this merge-commit in both of these figures is identical, but
+the presentation drastically changes our ability to comprehend what we are seeing.
+The primary difference is the removal of irrelevant information. In Gitk
+(Figure~\ref{fig:dag_view}), the visualization results include commits and merges
+from other components of the kernel, while our tool (Figure~\ref{fig:tree_view})
 only includes the results that are specific to the component of the kernel that we
@@
 
-Using \tool many advantages over the DAG model, but there are two
-drawbacks. The first drawback is that a server must query the state of the master
-repository since time information is not stored by git. The second drawback is the
-need to pre-compute the merge-tree of the repository.
+The merge-tree model used in \tool has many advantages to the DAG model, but there
+are two drawbacks. The first drawback is that a server must query the state of the
+master repository since time information is not stored by Git in order to properly
+construct the merge tree. The second drawback is the need to pre-compute the
+merge-tree of the repository.
 
@@
 makes the version control system vital in providing information about how a software
-project is being developed. To our knowledge, we know of no git repository
+project is being developed. To our knowledge, we know of no Git repository
 visualization tool that builds a tree that maps the path a commit follows to the
-master branch from the DAG provided by git. This may be because more information is
-required to generate the merge-tree model than what is stored by git. However, there
+master branch from the DAG provided by Git. This may be because more information is
+required to generate the merge-tree model than what is stored by Git. However, there
 has been a lot of work in providing visualizations of various repositories.
@@
 Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work with data
-from CVS repositories. Our goal is to provide information about git repositories,
+from CVS repositories. Our goal is to provide information about Git repositories,
 specifically the Linux master repository. Fewer tools are available for generating
-visualizations and summaries of git repositories, potentially due to the DAG model
-used by git.
+visualizations and summaries of Git repositories, potentially due to the DAG model
+used by Git.
 
@@
 
-The current industry standard tool for git repository visualization is Gitk. The
+The current industry standard tool for Git repository visualization is Gitk. The
 Gitk interface is built around the central DAG viewer. The DAG view displays the
@@
 to view additional information. For both merges and commits, Gitk will present the
-user with the git log. If the user selected a commit, Gitk will also present the
+user with the Git log. If the user selected a commit, Gitk will also present the
 user with the patch information and the names of the files edited. Gitk is unable to
@@
 
-The tree model requires more information than what is provided by git, requiring a
+The tree model requires more information than what is provided by Git, requiring a
 central server to monitor changes to the project. This breaks the design goals of

### 2016-04-26 linvis.tex
Final Edits

@@
 
-\evantodo {
-        \begin{enumerate}
-                \item Check captions
-        \end{enumerate}
-}
-
-Between 50k and 70k commits are added to the Linux kernel per version
-requiring maintainers of older versions of the Linux kernel to sift through
-thousands of commits and merges with tools that are unable to filter and effectively
-visualize projects at the scale of the kernel. Older versions of the kernel are used
-in embedded systems and mobile phones; for security purposes, performance needs, and
+Between 50k and 70k commits are added to the Linux kernel per version requiring
+maintainers of older versions of the Linux kernel to sift through thousands of
+commits and merges with tools that are unable to filter and effectively visualize
+projects at the scale of the kernel. Older versions of the kernel are used in
+embedded systems and mobile phones; for security purposes, performance needs, and
 changing hardware requirements, maintainers must be able to understand the changes
@@
 DAG view for many projects, but is unable to display the visualization for projects
-at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60 and
-70k new commits are created for the Linux project every year; According to our
+at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and
+70k new commits are created for the Linux project every year; according to our
 previous work\cite{German2015}, a commit takes a median of 30 days from the time it
@@
         \caption{A sample DAG view of commits from Linux in Gitk. Commits point
-                towards their ancestors, and there is no clear indication of the way
-                the travel to reach master. In fact, neither Gitk nor Git is capable
-                of showing what commits were done in master.}
+                towards their ancestors and there is no clear path from the commit
+                to the merge with the master branch. In fact, neither Gitk nor Git
+                are capable of showing what commits were done in master.}
         \label{fig:gitk}
@@
 One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch (including
-master). To complicate the problem, the DAG has reference to the ancestors of a
-commit (a model necessary for the operation of Git), but developers would prefer
-knowing the path a commit followed to reach the master repository. Tracing a path
-that any commit followed to the master repository would imply that for any given
-merge, it would be possible to know which commits were merged. A user could inspect
-the commits that arrived into master within a given time-frame by checking which
-commits were merged during that time-frame.
+Git does not store the date that a commit was merged into another branch, including
+the master branch. To complicate the problem, the DAG only has references to the
+ancestors of a commit (a model necessary for the operation of Git), but maintainers
+would prefer knowing the path a commit followed to reach the master repository.
+Tracing a path that any commit followed to the master repository would imply that
+for any given merge, it would be possible to know which commits were merged. A user
+could inspect the commits that arrived into the master branch within a given
+time-frame by checking which commits were merged during that time-frame.
 
@@
 DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach master; second, we present a
-method to inspect and visualize the history of merges in the Linux project using the
-merge-tree model.
+that represents the path used by a commit to reach the master branch; second, we
+present a method to inspect and visualize the history of merges in the Linux project
+using the merge-tree model.
 
@@
 filters, including the release version, along with a keyword or phrase from the log
-preview, the name of the author, or the commit ID. The user can request all the
-root-level merges containing a commit or merge that matches the query, or all
-commits and merges that match the query.
+preview, the name of the author, or the commit ID. The user can request all merges
+made by Linus that contain a commit or inner merge that matches the search query, or
+just the commits and merges that match the query.
 
@@
 merging multiple branches simultaneously). The second parent is the first branch
-merged, the third parent is the second branched merged.
+merged, the third parent is the second branched merged, etc.
 
@@
 
-A short example: assume the of commits represented in Figure~\ref{fig:repoEvents}
-shows the sequence of events in a repository. In this case, commits are performed in
+A short example: assume the commits represented in Figure~\ref{fig:repoEvents} shows
+the sequence of events in a repository. In this case, commits are performed in
 various repositories and branches. The DAG representation of the commits is shown in
@@
 this DAG is visible in Figure~\ref{fig:repoTree}. Note that the direction of the
-edges of the DAG has been inverted, instead of pointing from the child to the
+edges of the DAG have been inverted, instead of pointing from the child to the
 ancestors, it points from the ancestor to its successors, forming a path to the
@@
         \caption{DAG representation of the commits represented in
-                Figure~\ref{fig:repoEvents}. It loses information about what
-                repository the commit is performed and through which repositories it
-                has passed in its way to master. In fact, the DAG does not even know
-                what master is.}
+                Figure~\ref{fig:repoEvents}. The DAG loses information about which
+                repository the commit is performed in and through which merges it
+                has passed on its way to the master branch. The DAG does not even
+                distinguish the master branch from other branches.}
         \label{fig:repoDAG}
@@
                 Figure~\ref{fig:repoEvents} showing the path they followed to reach
-                master. In this model the successors of each commit represents the
-                path followed by that commit to reach master.}
+                the master branch. In this model the successors of each commit
+                represents the path followed by that commit to reach the master
+                branch.}
         \label{fig:repoTree}
@@
 the commits made directly to the master branch; whereafter it recursively determines
-the shortest path (in terms of time), using the DAG, from each commit to master
-using the inverted DAG.
+the shortest path (in terms of time), using the DAG, from each commit to the master
+branch using the inverted DAG.
 
@@
         \caption{Example of how merges record a subset of commits being merged. The
-                commit only shows the first 20 one line log messages for the 24
-                non-merge commits it merged. The ending ... is part of the log and
-                represents that other commits were merged.}
+                commit only shows the first 20 one-line log messages for the 24
+                non-merge commits it merged. The ending ``...'' is part of the log
+                and represents that other commits were merged.}
         \label{fig:sampleMerge}
@@
 The verification detected no incorrect merges since 2013, two incorrect merges in
-2012, and two in 2011; however, the heuristic breaks in October 15 2009 due to the
+2012, and two in 2011; however, the heuristic breaks in October, 15, 2009 due to the
 presence of \textit{foxtrot} commits, which confounds the true master branch of a
@@
 
-\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
-These are users that are maintaining a section of the kernel and would like to pick
-a merge (including all the commits that it merges) and merge it directly into their
+\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\ These
+are users that are maintaining a section of the kernel and would like to pick a
+merge (including all the commits that it merges) and merge it directly into their
 current repository. This is useful for reducing the amount of re-implementation
 work. For these users, it is important to have the ability to aggregate metadata
-about files and modules being effected by the merge.  It is also important for these
+about files and modules being effected by the merge. Also, it is important for these
 users to be able to navigate from the root of the merge-tree toward the leaves.
@@
 related to the current commit and how they get collated into merges that eventually
-end in master.  This is primarily for maintainers that need to perform some specific
-cherry picking of commits. We must provide these users a mechanism for navigating
-from a single commit toward the master, allowing them to see other commits that
-might be related to it.
+end in the master branch. This is primarily for maintainers that need to perform
+some specific cherry picking of commits. We must provide these users a mechanism for
+navigating from a single commit toward the master branch, allowing them to see other
+commits that might be related to their original commit.
 
@@
 In our visualizations, we leverage the Merge-Tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already in master,
-or is part of a tree which is rooted in the merge that merged it into master.  Each
-commit, whether a merge or non-merge, has only one successor; the root of each tree
-has none as it was made by Linus Torvalds directly into the master branch. Non-merge
-Commits contain the metadata for the changes made. This metadata includes the files
-changed, the lines added and removed from each file, the author, the date the commit
-was merged into the merge that led to being merged into the kernel, the date the
-commit was authored, the patch, and the commit log. Merges contain less metadata,
-only storing the author of the merge, the log, the commit date, and the author date,
-and potentially, changes necessary to address conflicts during the merge. The
-details of the model are outlined in \cite{German2015}.
+Section~\ref{sec:mergetree}. In this model each commit is either already in the
+master branch, or is part of a tree which is rooted in the merge that merged it into
+the master branch.  Each commit, whether a merge or non-merge, has only one
+successor; the root of each tree has none as it was made by Linus Torvalds directly
+into the master branch. Non-merge commits contain the metadata for the changes made.
+This metadata includes the files changed, the lines added and removed from each
+file, the author, the date the commit was merged into the merge that led to being
+merged into the kernel, the date the commit was authored, the patch, and the commit
+log. Merges contain less metadata, only storing the author of the merge, the log,
+the commit date, and the author date, and potentially, changes necessary to address
+conflicts during the merge. The details of the model are outlined in
+\cite{German2015}.
 
@@
 
-To navigate and inspect the merge-tree view of the Kernel we created a web-based
+To navigate and inspect the merge-tree view of the kernel we created a web-based
 tool called \tool. Creating a web-based tool enables users to use the system without
@@
         \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View. It allows finding commits that were \textbf{merged} in
-                a given period or by commit it.}
+        \caption{Search View allows filtering commits that were merged in a given
+                period, filtering by author, keyword, or commit ID.}
         \label{fig:search}
@@
 merges that are irrelevant. The search mechanism breaks down the results by release
-version. A user can further narrow down the search by \textbf{specifying a range of
-dates} in which such commits \textbf{were merged by Linus into master, not when the
-commits were created.} This distinction is important. We have observed commits that
-have taken years to arrive into master after they were originally committed.
+version. A user can further narrow down the search by specifying a range of dates in
+which such commits were merged by Linus into the master branch, not when the commits
+were created. This distinction is important. We have observed commits that have
+taken years to arrive into the master repository after they were originally
+committed.
 
@@
 
-If providing commit ID must be specified from the first character to the last
-character. For example, the commit `c267548755a184ef97301071300c1739a564e135' is
-returned when the user searches for a commit ID of `c26' in the 3.10 Linux kernel.
-This commit is a non-merge commit performed, in one of the largest merge-trees of
-the kernel; however, this feature will work with any commit ID.  We will use this
-commit for showing each of the features of \tool.
+If the user is searching by commit ID, ID must be specified from the first character
+to the last character. For example, the commit
+`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
+commit ID of `c26' in the 3.10 Linux kernel. This commit is a non-merge commit
+performed in one of the largest merge-trees of the kernel; however, this feature
+will work with any commit ID.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
-        \caption{Search Results. Each is a commit merged in the desired merged
-                window.}
+        \caption{Search Results. Each table entry is a commit merged in the desired
+                merged window.}
         \label{fig:results}
@@
 information about the content of the commit and who has signed-off on the commit to
-ensure that it is good quality. The message for merges may contain a summary of the
-commits being merged.  The information within these messages is highly variable, and
-is completely dependent on the author's style. As the user moves toward the
+ensure that it is of good quality. The message for merges may contain a summary of
+the commits being merged.  The information within these messages is highly variable,
+and is completely dependent on the author's style. As the user moves toward the
 top-level merge, the quality of these messages generally improves.
@@
         \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
-        \caption{Panels showing the files modified by all the commits that are part
+        \caption{Panel showing the files modified by all the commits that are part
                 of this merge.}
@@
 few cases, the author of a commit does not correctly follow this format and the
-heuristic approach fails. As with the Files panels, non-merge commits show their
+heuristic approach fails. As with the Files panel, non-merge commits show their
 corresponding information, but for merge commits we aggregate all the modules
@@
         \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
-        \item Reingold-Tilford trees provide a clear visual represent the tree structure of the merge-tree.
+        \item Reingold-Tilford trees provide a clear visual representation of the tree structure of the merge-tree.
         \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
@@
         \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List Tree}
+        \caption{List tree view of merge 73287a43cc79ca06629a88d1a199cd283f42456a}
         \label{fig:list_tree}
@@
 
-The tree can be further extended to include the collapsing of merge nodes. Clicking
-on merge nodes will cause them to collapse, only showing parts of the tree that are
-of interest to the user.Furthermore, the user can zoom the tree by scrolling the
-mouse wheel to see more of the topology of the tree, or to see the details of a
-given part of the tree.
+The tree can be further extended to include the collapsing of merge nodes, only
+showing parts of the tree that are of interest to the user. Furthermore, the user can
+zoom the tree by scrolling the mouse wheel to see more of the topology of the tree,
+or to see the details of a given part of the tree.
 
@@
 
-Bubble trees are useful for providing a clear visualization of hierarchal
+Bubble trees are useful for providing a clear visualization of hierarchical
 data\cite{Boardman2000}. The bubble tree (Figure~\ref{fig:bubble_tree}) provides a
@@
         \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree of a large merge, showing in red (the tiny dot close to the middle) the location of a specific commit of interest.}
+        \caption{Bubble Tree of merge 73287a43cc79ca06629a88d1a199cd283f42456a,
+                showing in red (the tiny dot close to the middle) the location of
+                the commit highlighted in Figure~\ref{fig:zoomed_tree}.}
         \label{fig:bubble_tree}
@@
 The response for the tree is a single object. This object is the root node of the
-tree, and may contain an object of children objects in the \verb|children| field,
-which are in the same format as the root object. The children object uses the commit
-ID as the key.
+tree, and may contain an object of children sub-tree objects in the \verb|children|
+field, which are in the same format as the root object. The children object uses the
+commit ID as the key.
 
@@
 the tree is an inner node, the response will contain the child nodes in the
-``children'' field, otherwise ``children'' will be an empty list. The format of file
+\verb|children| field, otherwise \verb|children| will be an empty list. The format of file
 response is as follows,
@@
 Linux. Effectively the merge-tree view recovers the details of each merged branch
-into master.
+into the master branch.
 
@@
 
-The biggest challenge, however, is computing the merge-tree. Our heuristic
-\textbf{only} applies to the Linux kernel repository, and only since July 2009. It
-is very likely that the heuristic will not work with other repository, specially
-those with fast-forward commits into the master branch (a majority of repositories).
-We can address the lack of information by monitoring the state of the repository at
-short intervals as we did in \cite{German2015}. This continuous mining of the
-repository requires foresight and planning.
+The biggest challenge, however, is computing the merge-tree. Our heuristic only
+applies to the Linux kernel repository, and only since July 2009. It is very likely
+that the heuristic will not work with other repositories, specifically those with
+fast-forward commits into the master branch (a majority of repositories). We can
+address the lack of information by monitoring the state of the repository at short
+intervals as we did in \cite{German2015}. This continuous mining of the repository
+requires foresight and planning.
 
 Because \tool leverages the merge-tree model of the kernel, it provides mechanisms
-and visualizations that other tools are unable to produce.It gives us the ability to
+and visualizations that other tools are unable to produce. It gives us the ability to
 see how commits are being merged into the kernel, allowing us to better understand
@@
         \includegraphics[width=0.97\textwidth]{figures/042dd_tree.png}
-        \caption{Merge Tree View \dmg{this Figure is unreadable. Any chance you can make the text bigger?}}
+        \caption{Merge tree topology for the same merge as in
+                Figure~\ref{fig:dag_view}.}
         \label{fig:tree_view}
@@
 this goal includes removing information that does not pertain to the area of the
-kernel that the user is interested in. In Figure~\ref{fig:dag_view} and
-Figure~\ref{fig:tree_view} we can visually compare the results returned from Gitk
+kernel that the user is interested in. In Figures~\ref{fig:dag_view} and
+\ref{fig:tree_view} we can visually compare the results returned from Gitk
 and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
@@
 team who were working together. Hipikat\cite{Cubranic2005} is another tool with a
-focus on communication. Where codebook focuses on developers working on a project,
+focus on communication. Where Codebook focuses on developers working on a project,
 Hipikat is focused on enabling easier integration of new developers to a project by
@@
 design decisions were made when they were made using other tools like Bugzilla and
-CVS. In both cases, our goal includes providing a summary of the topology of the
-kernel repository through a visual tree.
+CVS. Neither tool is sufficient in meeting our goals to provide a summary of the
+topology of the kernel repository through a visual tree.
 

### 2016-06-21 linvis.tex
Expand Figure 1
More Fixes
 - Under drawbacks, specifies that the continuous mining is only for validation, not
   required. (I've left notes where I made changes, accuracy should be checked)
 - Removed some of the implementation details about flask, python, and postgresql.
   I've left the api information, though it could probably be moved more into
   discussion to say that other developers can use what we've done via the api
 - Changed gitk picture

@@
 
-\begin{figure}
+\begin{figure*}
         \centering
-        \includegraphics[width=0.47\textwidth]{figures/gitk.png}
-        \caption{A sample DAG view of commits from Linux in Gitk. Commits point
-                towards their ancestors and there is no clear path from the commit
-                to the merge with the master branch. In fact, neither Gitk nor Git
-                are capable of showing what commits were done in master.}
+        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
+        \caption{A view of the Gitk interface centered on merge
+          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
+          their ancestors and there is no clear path from the commit to the merge
+          with the master branch. Neither Gitk nor Git are capable of showing the
+          commits in master.}
         \label{fig:gitk}
-\end{figure}
+\end{figure*}
 
@@
         \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
-        \caption{Entire Reingold-Tilford Tree of a merge. This merge by Linus (73287a43cc79ca06629a88d1a199cd283f42456a) contains 1722 includes non-merge commits
-                touching 239 modules and 1525 files.}
+        \caption{The complete Reingold-Tilford tree of a merge by Linus
+          (73287a43cc79ca06629a88d1a199cd283f42456a) containing 1722 non-merge
+          commits touching 239 modules and 1525 files.}
         \label{fig:zoomed_tree}
@@
 
-Bubble trees are useful for providing a clear visualization of hierarchical
-data\cite{Boardman2000}. The bubble tree (Figure~\ref{fig:bubble_tree}) provides a
-very clear picture of where a commit is located in the tree hierarchy. The currently
-selected commit or merge is colored red, while non-selected merges use a shade of
-blue and commits use white.  The depth of the blue colouring is dependent on how
-deep in the hierarchy that node is. The root is the lightest blue, and the contained
-merges are progressively darker.
+Bubble trees are useful for providing a clear visualization of wide, hierarchical
+data\cite{Boardman2000}. The tree structure is represented by the nesting of nodes;
+the largest circle is the root, containing all the other nodes. The smallest circles
+do not contain any nodes and therefore represent the leaf nodes.
+
+Our implementation of the bubble tree (Figure~\ref{fig:bubble_tree}) provides the
+user with a clear picture of where a commit is located in the merge tree. We
+highlight the selected commit or merge in red. Non-selected merges are in a shade of
+blue determined by the depth of a node in the tree. The root is the lightest shade
+of blue, while the contained merges are progressively darker; the commits are white.
 
@@
 
-\tool was built using Nginx as the main web server, postgresql as the DBMS, and
-Flask as the framework for generating the dynamic content.
-
-Nginx is a high-performance, open source, scalable HTTP and mail server. It is designed with massive
-scalability and multi-processing in mind. Instead of using request-driven threading,
-Nginx uses an event-driven architecture, making it more scalable, proving itself as
-being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
-other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.
-
-Postgresql is our DBMS of choice, providing fast results.
-
-We chose the python micro-framework
-Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the dynamic
-content. Working with flask is simple and easy, allowing us to work in Python.
-
-\subsection{Extendability}
+\evantodo{See if we can move API section to be part of another section}
 
@@
 
-The biggest challenge, however, is computing the merge-tree. Our heuristic only
-applies to the Linux kernel repository, and only since July 2009. It is very likely
-that the heuristic will not work with other repositories, specifically those with
-fast-forward commits into the master branch (a majority of repositories). We can
-address the lack of information by monitoring the state of the repository at short
-intervals as we did in \cite{German2015}. This continuous mining of the repository
-requires foresight and planning.
+\evan{Check accuracy of this paragraph}
+
+The biggest challenge is computing the merge-tree of the Linux kernel repository. It
+is likely that our heuristic for computing the merge-tree will not work with other
+repositories. The heuristic requires that there are no fast-forward merges into the
+master branch of the repository (a majority of repositories), and that there are no
+foxtrot merges. We are able to use our heuristic with the Linux kernel because of
+the strict integration model imposed by Linus. To validate the results of the
+heuristic, we use the continuous mining technique described in \cite{German2015}.
+The continuous mining of the repository requiring foresight and planning.
 
@@
 
-The tree model requires more information than what is provided by Git, requiring a
-central server to monitor changes to the project. This breaks the design goals of
-decentralized version control systems and makes the visualization tool specific to
-the project.
+\evan{Check the validity of this paragraph}
+We use a heuristic method to generate the tree from the repository. The heuristic
+requires that the repository does not have fast-forward merges into the master
+branch or foxtrot merges, which obfuscate the master branch. In either case, the
+heuristic will break and result in a tree that does not accurately represent the
+repository. Validation of the results from the heuristic requires continuous mining
+of the repository.
 

### 2016-06-22 linvis.tex
Replaced DAG drawings with tikzpictures
Removed the hand-drawn pictures of the conversion of the DAG to merge tree and
replaced them with tikzpictures.

@@
 \usepackage{algpseudocode}
+\usepackage{tikz}
+\usepackage{xcolor}
+\usepackage{color}
 
@@
 
+% Tikz libraries
+\usetikzlibrary{positioning}
+
+\input{colors.tex}
 \input{macros.tex}
@@
   \centering
-        \includegraphics[width=\columnwidth]{figures/events.pdf}
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
+    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
+
+    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
+    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
+    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
+    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
+
+        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
+    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
+  \end{tikzpicture}
   \caption{Example of a sequence of events performed in different
@@
   \centering
-        \includegraphics[width=.8\columnwidth]{figures/dag.pdf}
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
+  \end{tikzpicture}
   \caption{DAG representation of the commits represented in
@@
   \centering
-        \includegraphics[width=\columnwidth]{figures/tree.pdf}
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+
+    \draw[chartblue]
+      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (2) edge[chartyellow, -stealth] (5)
+          (3) edge[chartmagenta, -stealth](5)
+          (5) edge[chartyellow, -stealth] (7)
+          (7) edge[chartyellow, -stealth] (9)
+          (6) edge[chartred, -stealth] (9)
+          (8) edge[chartmagenta, -stealth] (9)
+          (9) edge[chartyellow, -stealth] (11)
+          (10) edge[chartyellow, -stealth] (12);
+
+  \end{tikzpicture}
   \caption{Merge-tree view of the commits  represented in

### 2016-07-25 linvis.tex
Small eidtorial changes in many places.
- Removed empty space in some of the images
- Edited text in few locations
- updated makefile to make diffs

@@
 
-\newcommand{\TheTitle}{Merge-tree: Visualizing the integration of commits into Linux}
+\newcommand{\TheTitle}{Merge-tree: Visualizing the Integration of Commits into Linux}
 \newcommand{\TheAuthors}{Evan Wilde, Daniel German}
@@
 commit has one or more ordered parents (ancestors), except the root commits of a
-repository which do not have ancestors--Linux git repository has two such commits.
+repository which do not have ancestors--Linux's git repository has two such commits.
 Commits are divided into merge commits, merging two or more branches, and non-merge
@@
 commits or reorder them, but it changes the commit ID in the process, effectively
-making a new commit. This makes commits unable to record the traversal of merges
+replacing it with a new commit. This makes commits unable to record the traversal of merges
 from the commit to the merge into the master branch.
@@
         \centering
-        {\tiny
+        {\fontsize{7}{9}
         \begin{verbatim}
@@
 md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache when mddev has no gendisk
+md/raid5: ensure we create a unique name for kmem_cache...
 ...
@@
 \end{figure}
+
+\textbf{rewrite}
 
@@
 following mechanisms to reach our goals of better navigation and better explanation
-of the changes.
+of the selected changes.
 
@@
 version. A user can further narrow down the search by specifying a range of dates in
-which such commits were merged by Linus into the master branch, not when the commits
-were created. This distinction is important. We have observed commits that have
+which such commits were merged by Linus into the master branch---not when the commits
+were created (author date and/or commit date). This distinction is important. We have observed commits that have
 taken years to arrive into the master repository after they were originally
-committed.
+created.
 
@@
 
-If the user is searching by commit ID, ID must be specified from the first character
-to the last character. For example, the commit
-`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
-commit ID of `c26' in the 3.10 Linux kernel. This commit is a non-merge commit
-performed in one of the largest merge-trees of the kernel; however, this feature
-will work with any commit ID.
+If the user is searching by commit ID, the ID can be specified by using any of its unique prefixes. For example, the commit
+\mycode{c267548755a184ef97301071300c1739a564e135} is returned when the user searches for a
+commit ID of \mycode{c26} in the 3.10 Linux kernel. % This commit is a non-merge commit
+% performed in one of the largest merge-trees of the kernel; however, this feature
+% will work with any commit ID.
 
@@
 
-In the search results, the user is presented with the one-line log message preview,
-the author's name and email, the date the commit was authored, and the date the
-commit was last committed. The commit date and author date will usually be the same;
-however, if the commit has been rebased or cherry-picked into a merge that
-eventually is merged into the kernel, the commit date and authored date will be
-different. We can see this in the first entry in Figure~\ref{fig:results}.
+In the search results (seen in Figure~\ref{fig:results}), the user is presented with the one-line log message preview,
+the author's name and email, the date the commit was authored, and the date the commit was last committed.
 
@@
 be present in all repositories. In the Linux repository, authors put the module they
-are working on in the one-line preview of the log-message; for example: the log
+are working on in the one-line summary of the log-message; for example: the log
 \emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
 coverage testing tool of the kernel). We heuristically extract the module by taking
-all text in the log preview of commits until the first colon.  Modules are logical
+all text in the log summary of commits until the first colon.  Modules are logical
 partitions of the information in the kernel. Depending on where the author was
@@
 
-The tree can be further extended to include the collapsing of merge nodes, only
-showing parts of the tree that are of interest to the user. Furthermore, the user can
+The tree can be displayed with some merges nodes collapsed, hence
+showing only parts of the tree that are of interest to the user. Furthermore, the user can
 zoom the tree by scrolling the mouse wheel to see more of the topology of the tree,
@@
         \verb|cid|      & string    & Commit ID\\
-        \verb|name|     & string    & One-line log preview\\
+        \verb|name|     & string    & One-line log summary\\
         \verb|mlinus|   & string    & Root merge commit ID\\
@@
 addressed these two issues with the creation of the merge-tree model of the DAG of
-Linux. Effectively the merge-tree view recovers the details of each merged branch
+Linux. Effectively the merge-tree view recovers the details of how each branch---and the commits they contain---was merged
 into the master branch.
@@
 In our experience, no other Git repository reaches the level of DAG complexity that
-Linux has. In most Git repositories, merges are not nested, all merges merge
+Linux has. In most Git repositories, most merges are not nested---most merges merge
 directly into the master repository, and these merges only merge few commits. Even
@@
 
-The biggest challenge is computing the merge-tree of the Linux kernel repository. It
+The biggest challenge is computing the merge-tree of any given repository. It
 is likely that our heuristic for computing the merge-tree will not work with other
@@
 master branch of the repository (a majority of repositories), and that there are no
-foxtrot merges. We are able to use our heuristic with the Linux kernel because of
+foxtrot merges (this is practice that starts be considered desirable by git users\footnote{See
+  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git} and
+  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
+We are able to use our heuristic with the Linux kernel because of
 the strict integration model imposed by Linus. To validate the results of the
@@
 
-This merge is relatively small, containing only four sub-merges, three of which
-only contain a single commit, and the last containing two commits. With our
-tool, we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
+Even though this merge is relatively small, containing only four sub-merges, three of which
+only contain a single commit, and the last containing two commits, it is already hard to visualize in Gitk. 
+With \tool we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
 furthermore, users must work to determine where this merge ends and the next one
@@
 
-Our tool shows promising results in working toward our goals, building a tool that
+Our tool, \tool, shows promising results in working toward our goals, building a tool that
 can more easily navigate the repository, and provide clearer explanations of the

### 2016-07-25 eval/closest.pl
Added all the code to compute the merge trees and verify them.
See readme.org for details

@@ -1,0 +1,703 @@
+#!/usr/bin/perl
+
+###
+#
+#drop table closest; create table closest (cid char(40), mnext char(40), mdist integer, mwhen timestamp,  tnext char(40), tdist interval, twhen timestamp); 
+#gitlinux=# \d closest
+#                Table "public.closest"
+#   Column   |            Type             | Modifiers 
+#------------+-----------------------------+-----------
+# cid        | character(40)               | not null
+# mnext      | character(40)               | 
+# mdist      | integer                     | 
+# mwhen      | timestamp without time zone | 
+# mnextmerge | character(40)               | 
+# mcidlinus  | character(40)               | 
+#Indexes:
+#    "closest_pkey" PRIMARY KEY, btree (cid)
+#    "closestmnext" btree (mnext)
+#
+##
+##
+
+sub Find_Children($);
+sub Follow_The_Rabbit($$);
+sub Distance($$);
+sub Set_Distance($$$$$$$);
+sub Find_Children($);
+sub Find_Parents($);
+sub Find_First_Parent($);
+sub Is_Merge($);
+sub Get_Commit_Date($);
+
+my $verbose = 1;
+
+####################################################33
+
+my $rip = '5c2a5ce689c99037771a6c110374461781a6f042';
+
+my $tableClosest = "closest";
+my $tableParents = "parents";
+
+$0 =~ m@/([^/]+)$@;
+my $xxdir = $`; #'
+
+push @INC, $xxdir ;
+
+use POSIX 'strftime';
+#require 'gitCommon.pm';
+
+use strict;
+use DBI;
+
+# latest commit 
+# 2145199c4f0db7c517dd788abec301dc84b91bd0',
+
+#my %headCommit = (
+#    "gitlinux" => '2d534926205db9ffce4bbbde67cb9b2cee4b835c', # just before aug1, 2012
+#    "gitlinux" => '4a490b78cb7e0e5efa44425df72a9fedc1c36366', # just before end of 2012
+#    "gitlinux" =>  '722aacb28588c7d0326493d1a0700d6a886be7b9', # april 10, 2013
+#    'gitlinux' => 'bf81710c4b6e2df2cc047f7c8e1f342511904b74',
+#    "gitlinux" =>  'd9a5c0a4e0b4c84850a1a5bbacba3f7858b67037', # March 24, 2007
+#    );
+
+#7f80850d3f9fd8fda23a317044aef3a6bafab06b',
+
[…]

### 2016-07-25 eval/commitsAtMerge.pl

@@ -1,0 +1,137 @@
+#!/usr/bin/perl
+
+# This one dumps the commits that are merged at a given commit 
+# basically the tree backwards
+
+$0 =~ m@/([^/]+)$@;
+my $xxdir = $`; #'
+
+push @INC, $xxdir ;
+
+use POSIX 'strftime';
+
+
+use strict;
+use DBI;
+
+
+
+
+my $dbName = $ARGV[0];
+my $distanceType = $ARGV[1];
+my $cid = $ARGV[2];
+my $parm = $ARGV[3];
+
+
+my @temp = split(',', $parm);
+my %options;
+
+
+foreach my $a (@temp) {
+    $options{$a} = $a;
+}
+
+
+my %savedDistance; # saved the distance to linus, so we don't have to query it again
+my %unmerged; # commits that have not been merged yet
+
+my %savedIsMerge;
+my %savedFirstParent;
+
+my $userName = 'dmg';
+
+die "usage $0 <db>  <merges|time> <cid> <force|init|print>"  if $dbName eq "" or not ($distanceType eq "merges" or $distanceType eq "time");
+
+die "not implemented" if $distanceType ne 'merges';
+
+my $dbh = DBI->connect("dbi:Pg:dbname=$dbName", "$userName", "",
+		       { RaiseError => 1});
+
+my $merge = $dbh->prepare("select ismerge from commits where cid = ?;");
+
+#print "New: $cid\n";
+my $is = Simple_Query("select mcidlinus from closest where mcidlinus = ?", $cid);
+if ($is ne $cid) {
+    print "Total;$cid;-1\n";
+    exit(0);
+}
+
+my @t = Simple_Query("select ismerge,comdatereal,committer,autdatereal,author,'' as rip,log from commits c natural join logs  where c.cid = ?", $cid);
+my $c = Print_Commit($cid, 0, 0, @t);
+print "Total;$cid;$c\n";
+
+sub Print_Commit
+{
+    my ($cid, $indent, $distance, @info) = @_;
+    
+    print " " x $indent, "$cid;$distance;", join(';', @info), "\n" if $options{"print"};
+    print " " x $indent, "",  $info[6], "\n" if $options{"logs"};
+    
+    # find its predecessors
+
+    my $q = $dbh->prepare("select cid,mdist, 'repo',ismerge,comdatereal,committer,autdatereal,author, log from closest natural join commits natural join logs where mnext = ? order by mdist desc;");
+    
+    $q->execute($cid);
+
[…]

### 2016-07-25 eval/insertCommits.pl

@@ -1,0 +1,246 @@
+#!/usr/bin/perl
+
+
+$0 =~ m@/([^/]+)$@;
+my $xxdir = $`; #'
+
+push @INC, $xxdir ;
+
+# this program inserts the metadata of commits
+
+
+use strict;
+use POSIX 'strftime';
+use Digest::SHA  qw(sha1 sha1_hex sha1_base64);
+
+
+###
+# logRepo.pl
+
+# first we need to find the branches that are here...
+
+# then we have to checkout each of them
+
+# we start with the assumption we are in a gip repo
+
+# get branches
+
+use strict;
+use DBI;
+
+my $dryRun = 0;
+
+
+# find the last-modified for the pack files directory
+# which is the most reliable way to know when the files were last changed
+# i think...
+
+my $dbName = $ARGV[0];
+my $repoDir = $ARGV[1];
+my $parm = $ARGV[2];
+
+my @temp = split(',', $parm);
+my %options;
+
+foreach my $a (@temp) {
+    $options{$a} = $a;
+}
+
+die "usage $0 <db> <repoDirectory> <force?>"  unless -d $repoDir ;
+
+die "$repoDir not a git repo"  unless -d "$repoDir/.git" ;
+
+
+my $userName = 'dmg';
+
+my $dbh = DBI->connect("dbi:Pg:dbname=$dbName", "$userName", "",
+		       { RaiseError => 1,
+                         AutoCommit => 0});
+
+my $temp = $dbh->prepare("set client_encoding to latin1");
+$temp->execute();
+
+my $create = $dbh->prepare("
+CREATE TABLE commits (
+    cid character(40),
+    author character varying(200),
+    committer character varying(200),
+    ismerge boolean,
+    contents character(40),
+    autdatereal timestamp with time zone,
+    comdatereal timestamp with time zone,
+    codecontents character(40),
+    reporeal character varying(80)
+);
+");
+$create->execute();
+
+$create = $dbh->prepare("
+CREATE TABLE parents (
+    cid character(40),
+    index integer,
+    parent character(40)
+);
+");
+$create->execute();
+
+my $insC = $dbh->prepare("insert into commits(author, autdatereal, committer, comdatereal, ismerge, contents, cid) values (?, ?,?,?,?,?,?);");
+my $insP = $dbh->prepare("insert into parents(cid, index, parent) values (?, ?, ?);");
+
+#----------------------------------------------------------------------
+
[…]

### 2016-07-25 eval/insertLogs.pl

@@ -1,0 +1,59 @@
+#!/usr/bin/perl
+
+
+$0 =~ m@/([^/]+)$@;
+my $xxdir = $`; #'
+
+push @INC, $xxdir ;
+
+# this program inserts the metadata of commits
+
+
+use strict;
+use POSIX 'strftime';
+use Digest::SHA  qw(sha1 sha1_hex sha1_base64);
+use DBI;
+
+my $dbName = $ARGV[0];
+my $repoDir = $ARGV[1];
+my $userName = "dmg";
+
+die "usage $0 <db> <repoDirectory> <force?>"  unless -d $repoDir ;
+
+die "$repoDir not a git repo"  unless -d "$repoDir/.git" ;
+
+my $dbh = DBI->connect("dbi:Pg:dbname=$dbName", "$userName", "",
+		       { RaiseError => 1, AutoCommit=>0});
+
+my $create = $dbh->prepare("drop table logs;");
+        
+$create->execute();
+
+my $create = $dbh->prepare("CREATE TABLE logs (
+    cid character(40) NOT NULL,
+    log character varying(10000),
+    primary key (cid)
+);");
+$create->execute();
+
+my $ins = $dbh->prepare("insert into logs(cid, log) values (?, ?); ");
+
+open(IN, "git -C $repoDir log  --pretty=oneline |");
+
+my $i =0;
+while(<IN>) {
+    chomp;
+    my $cid = substr($_, 0,40);
+    my $log = substr($_, 41);
+    $ins->execute($cid, $log);
+    if ($i++ % 10000 == 0) {
+        print STDERR "$cid;[$log]\n";
+        print STDERR "$i\n";
+    }
+}
+print "Done $i commits\n";
+
+close(IN);
+
+$dbh->commit();
+exit (0);

### 2016-07-25 linvis.tex
Updated the text of the paper with the explanation of the evaluation.
I also trimmed the paper to 10 pages by playing a bit with the spacing.

@@
 repository.
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
+        \caption{A view of the Gitk interface centered on merge
+          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
+          their ancestors and there is no clear path from the commit to the merge
+          with the master branch. Neither Gitk nor Git are capable of showing the
+          commits in master.}
+        \label{fig:gitk}
+%\vspace{-4mm}
+\end{figure*}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
+        \caption{Github Failure Message showing the DAG of the Linux Project and its
+                relationship to other forks.}
+        \label{fig:gitfail}
+%\vspace{-2mm}
+\end{figure}
 
@@
 
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
-        \caption{A view of the Gitk interface centered on merge
-          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
-          their ancestors and there is no clear path from the commit to the merge
-          with the master branch. Neither Gitk nor Git are capable of showing the
-          commits in master.}
-        \label{fig:gitk}
-\end{figure*}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its
-                relationship to other forks.}
-        \label{fig:gitfail}
-\end{figure}
-
 One major challenge with visualizing the arrival of commits to a repository is that
@@
   \label{fig:repoEvents}
+%\vspace{-3mm}
 \end{figure}
@@
   \label{fig:repoDAG}
+%\vspace{-3mm}
 \end{figure}
@@
   \label{fig:repoTree}
+\vspace{-3mm}
 \end{figure}
@@
                 \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.\\}
+                \State {\# in DAG with the path it followed to master.}
 
@@
 ...
-        \end{verbatim}}
+        \end{verbatim}}\vspace{-5mm}
         \caption{Example of how merges record a subset of commits being merged. The
-                commit only shows the first 20 one-line log messages for the 24
+                commit only shows the first 20 one-line summaries messages for the 24
                 non-merge commits it merged. The ending ``...'' is part of the log
@@
 
-\textbf{rewrite}
-
-We use this information to evaluate the accuracy of the merge-tree model extracted
-from the DAG. For commits with no conflicts, we compare: 1) the number of non-merge
-commits, 2) the summaries of the commits available in the log, and 3) the summaries
-of the merges included.
+We used this information to evaluate the accuracy of the merge-tree model extracted from the DAG. The method we followed
+started with the extraction of the commit history up to July 20, 2016. We computed the merge tree of every commit until
+then. Since Linus Torvalds mostly does merging directly into master, we assumed that every merge by him is the root of a
+merge-tree. As described above, the log of a commit usually contains the number of commits in the merge the first 20
+summaries of commits being merged. We extracted merges by Linus Torvalds using the command \mycode{log --merges --author='Torvalds} and compared
+the number of commits in according to the log with the number of commits in the  merge-tree rooted in this commit. We
+also used the summaries of the commits found in the merge (not necessarily all---see above) to make sure those 
+commits were in their corresponding merge tree. For example, for the merge in Figure~\ref{sampleMerge} we would expect
+that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line summaries corresponds to commits in
+that merge-tree. We also inspected those with differences to make sure they were true errors.
+
+The results can be summarized as follows:
+
+\begin{itemize}
+\item 5 merges were false-erros because their logs did not contain accurate information (were probably edited by
+  hand). For example in \mycode{42a579a0f...}) one commit summary was missing (the line was empty), in another  \mycode
+  {c55d267}) the summaries were reordered.
+\item The heuristic works correctly for all merges except between Jun 7, 2014 and Jun 2, 2014 (92 merges) and before
+  Sept 4, 2007. Before this date merges did not include a summary of the commits they included, hence making it
+  impossible to verify; however, we correctly identified the merges by Linus into master. 
+\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
+presence of \textit{foxtrot} commits, which confound the true master branch of a
+repository
+\footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and
+is not capable of identifying most of the merges before this date.
+\end{itemize}
+
+Of the merges after Sept 4, 2007, in 99.4\% fo them (16,768 out of 16,860)  our heuristic was correct. Overall, we
+correctly identified the merge trees of 87.15\% of the merges into master in Linux.
 
 The verification detected no incorrect merges since 2013, two incorrect merges in
-2012, and two in 2011; however, the heuristic breaks in October, 15, 2009 due to the
-presence of \textit{foxtrot} commits, which confounds the true master branch of a
-repository
-\footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and will not work for older merges.
+2012, and two in 2011; however, the heuristic breaks in October, 15, 2009
 

### 2016-07-25 linvis.tex
added a minor clarification

@@
 \item The heuristic works correctly for all merges except between Jun 7, 2014 and Jun 2, 2014 (92 merges) and before
-  Sept 4, 2007. Before this date merges did not include a summary of the commits they included, hence making it
+  Sept 4, 2007. Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
   impossible to verify; however, we correctly identified the merges by Linus into master. 
 \item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-presence of \textit{foxtrot} commits, which confound the true master branch of a
+presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
 repository

### 2016-07-25 linvis.tex
In 2004 we are correct, not wrong
Looked more deeply into the issue and found that it is not an error,
it is a HUGE branch created by linus. he committed to a branch on this
with this root 3f17ea6dea8ba5668873afa54628a91aaa3fb1c0.

Fixed grammar issues too.

--- 2016-07-25 linvis.tex
@@
 that merge-tree. We also inspected those with differences to make sure they were true errors.
-
 The results can be summarized as follows:
@@
 \begin{itemize}
-\item 5 merges were false-erros because their logs did not contain accurate information (were probably edited by
-  hand). For example in \mycode{42a579a0f...}) one commit summary was missing (the line was empty), in another  \mycode
-  {c55d267}) the summaries were reordered.
-\item The heuristic works correctly for all merges except between Jun 7, 2014 and Jun 2, 2014 (92 merges) and before
-  Sept 4, 2007. Before this date merges did not include a summary of the commits they included, hence making it
-  impossible to verify; however, we correctly identified the merges by Linus into master. 
+\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
+  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
+  {c55d267} the summaries were reordered.
+\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
+  branch (not master). This branch was merged at (\mycode{3f17ea6d..} which contained 6809 commits.
+\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
+  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
+  impossible to verify; during this period, however, we correctly identified the merges by Linus into master. 
 \item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-presence of \textit{foxtrot} commits, which confound the true master branch of a
-repository
-\footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and
-is not capable of identifying most of the merges before this date.
+presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
+repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
 \end{itemize}
 
-Of the merges after Sept 4, 2007, in 99.4\% fo them (16,768 out of 16,860)  our heuristic was correct. Overall, we
-correctly identified the merge trees of 87.15\% of the merges into master in Linux.
-
-The verification detected no incorrect merges since 2013, two incorrect merges in
-2012, and two in 2011; however, the heuristic breaks in October, 15, 2009
+In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860) our heuristic was correct. It failed in 1,542
+commits before Dec. 12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
 

### 2016-07-25 eval/closest.pl

@@ -692,7 +692,7 @@ sub Get_Baseline
     }
     print STDERR "End Reading baseline [$i]\n";
 }
-g
+
 
 sub Simple_Query
 {

### 2016-08-02 linvis.tex
Minor changes
Just making some minor changes and grammar fixes.

@@
 
-In this paper, we propose the merge-tree view, a simplified transformation
+In this paper, we propose the merge-tree, a simplified transformation
 of the DAG of the Linux git repository that shows the way in which commits
-are merged into the master branch of Linux. Using the merge-tree view we
+are merged into the master branch of Linux. Using the merge-tree, we
 build \tool, a tool that is designed to allow users to explore how commits
@@
 Between 50k and 70k commits are added to the Linux kernel per version requiring
-maintainers of older versions of the Linux kernel to sift through thousands of
+maintainers of older versions of the kernel to sift through thousands of
 commits and merges with tools that are unable to filter and effectively visualize
@@
 Our visualizations and tool provide information about the location of any given
-commit in its respective merge-tree, the files edited, the modules edited, and the
-commit message for a given commit or merge. \tool allows users to apply various
-filters, including the release version, along with a keyword or phrase from the log
-preview, the name of the author, or the commit ID. The user can request all merges
-made by Linus that contain a commit or inner merge that matches the search query, or
-just the commits and merges that match the query.
+commit or merge in its respective merge-tree, the files edited, the modules edited,
+and the commit message. \tool allows users to apply various filters, including the
+release version, along with a keyword or phrase from the log preview, the name of
+the author, or the commit ID. The user can request all merges made by Linus that
+contain a commit or inner merge that matches the search query, or just the commits
+and merges that match the query.
 
@@
 
-A short example: assume the commits represented in Figure~\ref{fig:repoEvents} shows
+A short example: assume the commits represented in Figure~\ref{fig:repoEvents} show
 the sequence of events in a repository. In this case, commits are performed in
@@
 then. Since Linus Torvalds mostly does merging directly into master, we assumed that every merge by him is the root of a
-merge-tree. As described above, the log of a commit usually contains the number of commits in the merge the first 20
+merge-tree. As described above, the log of a merge-commit usually contains the number of commits in the merge the first 20
 summaries of commits being merged. We extracted merges by Linus Torvalds using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits in according to the log with the number of commits in the  merge-tree rooted in this commit. We
+the number of commits according to the log with the number of commits in the  merge-tree rooted in this commit. We
 also used the summaries of the commits found in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the merge in Figure~\ref{sampleMerge} we would expect
+commits were in their corresponding merge tree. For example, for the merge in
+Figure~\ref{fig:sampleMerge} we would expect
 that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line summaries corresponds to commits in
@@
 The goal of \tool is to simplify the navigation of the kernel commit information,
-specifically focusing on merges. This is done by leveraging the merge-tree view to
+specifically focusing on merges. This is done by leveraging the merge-tree to
 inspect how commits are merged on the path to the master repository.
@@
 
-In our visualizations, we leverage the Merge-Tree model described in
+In our visualizations, we leverage the merge-tree model described in
 Section~\ref{sec:mergetree}. In this model each commit is either already in the
@@
 and is completely dependent on the author's style. As the user moves toward the
-top-level merge, the quality of these messages generally improves.
+root-level merge, the quality of these messages generally improves.
 
@@
 many more edges that are necessary to understand how integration occurred. We
-addressed these two issues with the creation of the merge-tree model of the DAG of
-Linux. Effectively the merge-tree view recovers the details of how each branch---and the commits they contain---was merged
+addressed these two issues with the creation of the merge-tree model from the DAG of
+Linux. Effectively, the merge-tree recovers the details of how each branch---and the commits they contain---was merged
 into the master branch.
@@
 Linux has. In most Git repositories, most merges are not nested---most merges merge
-directly into the master repository, and these merges only merge few commits. Even
+directly into the master repository, and these merges only merge a few commits. Even
 in these simplified cases, the merge-tree can provide a valuable summary of how
@@
 master branch of the repository (a majority of repositories), and that there are no
-foxtrot merges (this is practice that starts be considered desirable by git users\footnote{See
+foxtrot merges (a practice that is starting to be considered desirable by git users\footnote{See
   \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git} and
@@
 heuristic, we use the continuous mining technique described in \cite{German2015}.
-The continuous mining of the repository requiring foresight and planning.
+Continuous mining of the repository requires foresight and planning.
 
@@
 and visualizations that other tools are unable to produce. It gives us the ability to
-see how commits are being merged into the kernel, allowing us to better understand
+see how commits are being merged into the master branch, allowing us to better understand
 how code is integrated into the kernel.
@@
 ``regulator'' module was modified in this merge and was modified by 5 commits.
-
-The merge-tree model used in \tool has many advantages to the DAG model, but there
-are two drawbacks. The first drawback is that a server must query the state of the
-master repository since time information is not stored by Git in order to properly
-construct the merge tree. The second drawback is the need to pre-compute the
-merge-tree of the repository.
-
-% Our core model provides advantages to the DAG model used by other tools. We separate
-% the commits and merges into sub-trees and focus on a single sub-tree at a time. This
-% enables us to filter out commits and merges that do not pertain to the component of
-% the kernel that we are interested in, providing a clearer picture of what is
-% happening in a given merge. The tree model also enables us to aggregate metadata
-% easily. With the addition of the continuous polling of the repository, the model
-% provides us with a mechanism of performing date-range queries with meaningful
-% results.
 
@@
 heuristic will break and result in a tree that does not accurately represent the
-repository. Validation of the results from the heuristic requires continuous mining
-of the repository.
-
-While these are significant drawbacks for small projects, large projects will likely
-have a more centralized work flow with a master repository where contributions will
-eventually land, along with various servers that can watch this master repository
-for changes. Given that the importance of clear visualizations of the repository
-outweighs the cost of having a server monitor the state of the repository, the
-tree-based model appears to provide clean visualizations and explanations of the
-changes in the Linux kernel.
+repository.
+
+These may pose as significant drawbacks to smaller projects; however, with a proper
+merge discipline, the heuristic is able to correctly convert the DAG into a
+merge-tree. The importance of clear visualizations in large projects outweighs
+the costs of enforcing a clean merge discipline. The tree-based model appears to
+have the ability to provide a clean visualization and more detailed explanations of
+the changes in the Linux kernel.
 

### 2016-09-18 linvis.tex

@@
 If the user is searching by commit ID, the ID can be specified by using any of its unique prefixes. For example, the commit
-\mycode{c267548755a184ef97301071300c1739a564e135} is returned when the user searches for a
-commit ID of \mycode{c26} in the 3.10 Linux kernel. % This commit is a non-merge commit
-% performed in one of the largest merge-trees of the kernel; however, this feature
-% will work with any commit ID.
+\mycode{3f17ea6dea8ba5668873afa54628a91aaa3fb1c0} is returned when the user searches for a
+commit ID of \mycode{3f17e} in the 3.16 Linux kernel.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List tree view of merge 73287a43cc79ca06629a88d1a199cd283f42456a}
+        \caption{List tree view of merge 3f17ea6}
         \label{fig:list_tree}
@@
 
-The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:zoomed_tree} and
-\ref{fig:tree}) allows the visualization and navigation of the entire merge-tree in
+The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:reingold_tree} and
+\ref{fig:reingold_tree_zoom}) allows the visualization and navigation of the entire merge-tree in
 an intuitive representation of the tree.  This illustrates a clear notion of root
-and leaves, and how to navigate in either direction. Some of the merge trees are
-very large (Figure~\ref{fig:zoomed_tree}), containing thousands of commits and
-merges. While the tree is capable of producing a visualization, it becomes much more
-difficult to understand. For example the merge
-73287a43cc79ca06629a88d1a199cd283f42456a performed by Linus Torvalds in May 1, 2013,
-merged 1722 non-merge commits and 91 merges. \footnote{
-        This commit can be inspected at
-        \url{http://li.turingmachine.org/commits/73287a43cc79ca06629a88d1a199cd283f42456a}}
+and leaves, and how to navigate in either direction. Some merge trees are very
+large, containing thousands of commits and merged. While the tree is capable of
+producing a visualization, it becomes far more difficult to understand. For example,
+the merge, 3f17ea6 performed by Linus Torvalds June 8 2014, contains 7217 commits
+and merges. \footnote{This commit can be inspected at
+  \url{http://li.turingmachine/org/commits/3f17ea6dea8ba5668873afa54628a91aaa3fb1c0}}
 
@@
         \centering
-        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
-        \caption{The complete Reingold-Tilford tree of a merge by Linus
-          (73287a43cc79ca06629a88d1a199cd283f42456a) containing 1722 non-merge
-          commits touching 239 modules and 1525 files.}
-        \label{fig:zoomed_tree}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
-        \caption{Detail of the Reingold-Tilford Tree in Figure~\ref{fig:zoomed_tree}.}
-        \label{fig:tree}
-\end{figure}
-
-The tree can be displayed with some merges nodes collapsed, hence
-showing only parts of the tree that are of interest to the user. Furthermore, the user can
-zoom the tree by scrolling the mouse wheel to see more of the topology of the tree,
-or to see the details of a given part of the tree.
+        \includegraphics[width=0.47\textwidth]{figures/4dc4_tree.pdf}
+        \caption{Merge 4dc4226 is a subtree of 3f17ea6, for the power-management
+          module of the kernel}
+        \label{fig:reingold_tree}
+      \end{figure}
+
+      \begin{figure*}
+        \centering
+        \includegraphics[width=0.98\textwidth]{figures/4dc4_zoom_tree.pdf}
+        \caption{Zoomed out view of the Reingold-Tilford tree for merge 4dc4226}
+        \label{fig:reingold_tree_zoom}
+      \end{figure*}
+
+      The user is initially greeted with their current node centered on the screen.
+      They are able to zoom the tree by scrolling the mouse, and clicking and
+      dragging to pan the tree. They can see more details about a node by clicking
+      on it, which will provide them a link to the specific page for that commit.
 
@@
         \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree of merge 73287a43cc79ca06629a88d1a199cd283f42456a,
-                showing in red (the tiny dot close to the middle) the location of
-                the commit highlighted in Figure~\ref{fig:zoomed_tree}.}
+        \caption{Bubble Tree of merge 3f17ea6, merge 4dc4226 is the currently
+          selected merge and is highlighted in orange.}
         \label{fig:bubble_tree}
@@
 
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.97\textwidth]{figures/042dd_tree.png}
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
         \caption{Merge tree topology for the same merge as in
@@
         \label{fig:tree_view}
-\end{figure*}
+\end{figure}
 

### 2017-06-09 paper/lineval.tex
Initial paper layout

--- 2016-09-18 linvis.tex
@@
-% vim: set fenc=utf-8 ft=latex encoding=utf-8
-% -*- mode: latex; coding: UTF-8; -*-
+% vim:set et fenc=utf-8 ft=tex sw=2 ts=4 tw=84:
 
 \newif\ifdraft
-% \drafttrue
+\drafttrue
 
@@
 
-\newcommand{\TheTitle}{Merge-tree: Visualizing the Integration of Commits into Linux}
-\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
-\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Digesting large amounts of commit data}
-\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}
+\usepackage[hyphens]{url} \urlstyle{same}
 
-\synctex=1
-
-\usepackage[hyphens]{url}
-\urlstyle{same}
-
-\ifdraft
-        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
-                pdfborder={0 0 0},backref=none,colorlinks=true]{hyperref}
-
-\else
-        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
-                pdfborder={0 0 0},backref=none,colorlinks=false]{hyperref}
-\fi
-
+%% Citations
 \usepackage[nospace]{cite}
 
-% Table Support
+%% Table Support
+
 \usepackage{dcolumn}
 \usepackage{longtable}
+\usepackage{multirow}
 
+%% Extra support
+\usepackage{xspace}
+\usepackage{amsmath}
+\usepackage{algorithm}
 \usepackage{balance}
 \usepackage{placeins}
-\usepackage{multirow}
 
-% Extra support
-\usepackage{xspace}
-\usepackage{caption}
-\usepackage[nospace]{cite}
+%% Graphics
 
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
 \usepackage{tikz}
@@
 
-% Fix any bad-hyphenations here
-\hyphenation{contained}
+%% Tikz
+\usetikzlibrary{positioning, arrows}
 
-% Tikz libraries
-\usetikzlibrary{positioning}
+\input{color.tex}
+\input{macros.tex}
 
-\input{colors.tex}
-\input{macros.tex}
+\newcommand{\TheTitle}{An evaluation and generalization of Merge-Tree}
+\newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
+\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
+\newcommand{\TheSubject}{Testing the comprehension of users in large git repositories}
+\newcommand{\TheKeywords}{Linux, git, data structures}
+
+\synctex=1
 
@@
 \title{\TheTitle}
+
 \author{
@@
 \maketitle
+
 \begin{abstract}
 
-With an average of more than 900 top-level merges into the Linux kernel per
-release, many containing hundreds of commits and some containing thousands,
-maintenance of older versions of the kernel becomes nearly impossible.
-Various commercial products, such as the Android platform, run older
-versions of the kernel. Due to security, performance, and changing hardware
-needs, maintainers must understand what changes (commits) are added to the
-current version of the kernel since the last time they inspected it in order
-to make the necessary patches.
-
-Current tools provide information about repositories through the directed
-acyclic graph (DAG) of the repository, which is helpful for smaller
-projects. However, with the scale and number of branches in the kernel the
-DAG becomes overwhelming very quickly. Furthermore, the DAG contains every
-ancestor of every commit, while maintainers are more interested in how and
-when a commit arrives to the official Linux repository.
-
-In this paper, we propose the merge-tree, a simplified transformation
-of the DAG of the Linux git repository that shows the way in which commits
-are merged into the master branch of Linux. Using the merge-tree, we
-build \tool, a tool that is designed to allow users to explore how commits
-are merged into the Linux kernel.
+  \input{abstract.tex}
 
@@
 
-\begin{IEEEkeywords}
-        \TheKeywords
-\end{IEEEkeywords}
-
-\section{Introduction}
-
-Between 50k and 70k commits are added to the Linux kernel per version requiring
-maintainers of older versions of the kernel to sift through thousands of
-commits and merges with tools that are unable to filter and effectively visualize
-projects at the scale of the kernel. Older versions of the kernel are used in
-embedded systems and mobile phones; for security purposes, performance needs, and
-changing hardware requirements, maintainers must be able to understand the changes
-being made in the current version of the kernel in order to produce the necessary
-patches for the older versions of the kernel. Tools like Gitk use a directed acyclic
-graph (DAG) model of the repository, showing all commits and merges in chronological
-order by when the commit was authored, not by when it arrived in the official Linux
-repository.
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
-        \caption{A view of the Gitk interface centered on merge
-          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
-          their ancestors and there is no clear path from the commit to the merge
-          with the master branch. Neither Gitk nor Git are capable of showing the
-          commits in master.}
-        \label{fig:gitk}
-%\vspace{-4mm}
-\end{figure*}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its
-                relationship to other forks.}
-        \label{fig:gitfail}
-%\vspace{-2mm}
-\end{figure}
-
-This representation works in smaller projects; it enables users to see when changes
-are made, when these changes are merged, how each branch is interacting, and the
-point where a branch forks from the master branch. In large modular projects, like
-the Linux kernel, the DAG becomes a mess of merges and commits
-(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel
-is simply too large for the system to generate a visualization; Github provides a
-DAG view for many projects, but is unable to display the visualization for projects
-at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and
-70k new commits are created for the Linux project every year; according to our
-previous work\cite{German2015}, a commit takes a median of 30 days from the time it
-is authored until it arrives in the official repository. The snapshot of the kernel
-tomorrow may be different than the snapshot from today, containing new commits
-authored in the past; distinguishing these new commits from the commits in the
-snapshot from today is not trivial.
-
-One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch, including
-the master branch. To complicate the problem, the DAG only has references to the
-ancestors of a commit (a model necessary for the operation of Git), but maintainers
-would prefer knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would imply that
-for any given merge, it would be possible to know which commits were merged. A user
-could inspect the commits that arrived into the master branch within a given
-time-frame by checking which commits were merged during that time-frame.
-
-This paper makes two contributions; first, we describe a method of converting the
-DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach the master branch; second, we
-present a method to inspect and visualize the history of merges in the Linux project
-using the merge-tree model.
-
-These methods and visualizations are implemented in a web-based tool called
-\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
-Our visualizations and tool provide information about the location of any given
-commit or merge in its respective merge-tree, the files edited, the modules edited,
-and the commit message. \tool allows users to apply various filters, including the
-release version, along with a keyword or phrase from the log preview, the name of
-the author, or the commit ID. The user can request all merges made by Linus that
-contain a commit or inner merge that matches the search query, or just the commits
-and merges that match the query.
-
-\section{Merge-Tree model}
-\label{sec:mergetree}
-
-Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
-commit has one or more ordered parents (ancestors), except the root commits of a
-repository which do not have ancestors--Linux's git repository has two such commits.
-Commits are divided into merge commits, merging two or more branches, and non-merge
-commits. Non-merge commits have only one parent, while merge commits have two or
-more. The order of the parents matter: the first parent is the branch in which the
-merge is being done, while the rest indicate the branches being merged (Git allows
-merging multiple branches simultaneously). The second parent is the first branch
-merged, the third parent is the second branched merged, etc.
-
-Once a commit is created, it is never changed. Git allows operations to alter
-commits or reorder them, but it changes the commit ID in the process, effectively
-replacing it with a new commit. This makes commits unable to record the traversal of merges
-from the commit to the merge into the master branch.
-
-A short example: assume the commits represented in Figure~\ref{fig:repoEvents} show
-the sequence of events in a repository. In this case, commits are performed in
-various repositories and branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about the master
-branch and the repository that the master branch is part of. The merge-tree view of
-this DAG is visible in Figure~\ref{fig:repoTree}. Note that the direction of the
-edges of the DAG have been inverted, instead of pointing from the child to the
-ancestors, it points from the ancestor to its successors, forming a path to the
-master branch. Also note that the DAG has been simplified, showing only a single
-edge on the path to master for any commit.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
-    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-
-    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
-    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
-    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
-    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
-
-        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
-    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
-  \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
-  \label{fig:repoEvents}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG loses information about which
-    repository the commit is performed in and through which merges it
-    has passed on its way to the master branch. The DAG does not even
-    distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (2) edge[chartyellow, -stealth] (5)
-          (3) edge[chartmagenta, -stealth](5)
-          (5) edge[chartyellow, -stealth] (7)
-          (7) edge[chartyellow, -stealth] (9)
-          (6) edge[chartred, -stealth] (9)
-          (8) edge[chartmagenta, -stealth] (9)
-          (9) edge[chartyellow, -stealth] (11)
-          (10) edge[chartyellow, -stealth] (12);
-
-  \end{tikzpicture}
-  \caption{Merge-tree view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
-\vspace{-3mm}
-\end{figure}
-
-\subsection{Computing the merge-tree of the DAG of Linux}
-
-Computing the merge-tree from a DAG for any repository may not be possible; however,
-certain features of the development process of Linux make it feasible to compute the
-merge-tree for the Linux repository. First, the master branch of Linux is maintained
-by Linus Torvalds, and only Linus has write access to it. We have verified this
-assertion in previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first identifies
-the commits made directly to the master branch; whereafter it recursively determines
-the shortest path (in terms of time), using the DAG, from each commit to the master
-branch using the inverted DAG.
-
-\begin{algorithm}
-        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
-        \begin{algorithmic}
-                \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
-                \State $head \gets \textit{Head of master of git repository}$
-                \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
-                \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
-                \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
-                \If {\textit{cid in master}}
-                \State \Return 0
-                \EndIf
-                \State    $d \gets 	\infty$
-                \State {\# Traverse the inverted DAG}
-                \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
-                \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
-                \EndIf
-                \EndFor
-                \State {\# $c$ is the commit that follows $cid$}
-                \State {\# in its way to master}
-                \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
-                \EndFunction
-
-                \State {\# Compute the distance for each commit}
-                \State {\# discarding result}
-                \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
-                \EndFor
-                \State \Return $Tree$
-                \EndFunction
-        \end{algorithmic}
-\end{algorithm}
-
-\subsection{Evaluation}
-
-Merges that do not have conflicts provide information to verify this heuristic. If a
-merge does not contain a conflict, it records a summary of the commits that it
-merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a
-list of the first 20 non-merge commits in the merge, including their one-line log
-description, the full logs of the merge commits that merge this subset, and the
-total number of non-merge commits in the merge.
-
-\begin{figure}[htbp]
-        \centering
-        {\fontsize{7}{9}
-        \begin{verbatim}
-Merge: 8cbd84f fd8aa2c
-Author: Linus Torvalds <torvalds@linux-foundation.org>
-Date:   Tue Aug 10 15:38:19 2010 -0700
-
-Merge branch 'for-linus' of git://neil.brown.name/md
-
-* 'for-linus' of git://neil.brown.name/md: (24 commits)
-md: clean up do_md_stop
-[... edited for the sake of space]
-md: split out md_rdev_init
-md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache...
-...
-        \end{verbatim}}\vspace{-5mm}
-        \caption{Example of how merges record a subset of commits being merged. The
-                commit only shows the first 20 one-line summaries messages for the 24
-                non-merge commits it merged. The ending ``...'' is part of the log
-                and represents that other commits were merged.}
-        \label{fig:sampleMerge}
-\end{figure}
-
-We used this information to evaluate the accuracy of the merge-tree model extracted from the DAG. The method we followed
-started with the extraction of the commit history up to July 20, 2016. We computed the merge tree of every commit until
-then. Since Linus Torvalds mostly does merging directly into master, we assumed that every merge by him is the root of a
-merge-tree. As described above, the log of a merge-commit usually contains the number of commits in the merge the first 20
-summaries of commits being merged. We extracted merges by Linus Torvalds using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits according to the log with the number of commits in the  merge-tree rooted in this commit. We
-also used the summaries of the commits found in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the merge in
-Figure~\ref{fig:sampleMerge} we would expect
-that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line summaries corresponds to commits in
-that merge-tree. We also inspected those with differences to make sure they were true errors.
-The results can be summarized as follows:
-
-\begin{itemize}
-\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
-  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
-  {c55d267} the summaries were reordered.
-\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
-  branch (not master). This branch was merged at (\mycode{3f17ea6d..} which contained 6809 commits.
-\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
-  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
-  impossible to verify; during this period, however, we correctly identified the merges by Linus into master.
-\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
-repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
-\end{itemize}
-
-In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860) our heuristic was correct. It failed in 1,542
-commits before Dec. 12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
-
-\section{Visualizing the merge-tree of Linux}
-
-The goal of \tool is to simplify the navigation of the kernel commit information,
-specifically focusing on merges. This is done by leveraging the merge-tree to
-inspect how commits are merged on the path to the master repository.
-
-\subsection{Use cases}
-
-We designed \tool with two use-cases in mind, though a user may switch between the
-cases as they work.
-
-\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\ These
-are users that are maintaining a section of the kernel and would like to pick a
-merge (including all the commits that it merges) and merge it directly into their
-current repository. This is useful for reducing the amount of re-implementation
-work. For these users, it is important to have the ability to aggregate metadata
-about files and modules being effected by the merge. Also, it is important for these
-users to be able to navigate from the root of the merge-tree toward the leaves.
-
-\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
-These are users that start with a known merge or commit and would like to see what
-other changes are being made in commits that are in the same merge, including
-knowing the merge-tree they belong to. This is useful to see what other commits are
-related to the current commit and how they get collated into merges that eventually
-end in the master branch. This is primarily for maintainers that need to perform
-some specific cherry picking of commits. We must provide these users a mechanism for
-navigating from a single commit toward the master branch, allowing them to see other
-commits that might be related to their original commit.
-
-\subsection{Data Model}
-
-In our visualizations, we leverage the merge-tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already in the
-master branch, or is part of a tree which is rooted in the merge that merged it into
-the master branch.  Each commit, whether a merge or non-merge, has only one
-successor; the root of each tree has none as it was made by Linus Torvalds directly
-into the master branch. Non-merge commits contain the metadata for the changes made.
-This metadata includes the files changed, the lines added and removed from each
-file, the author, the date the commit was merged into the merge that led to being
-merged into the kernel, the date the commit was authored, the patch, and the commit
-log. Merges contain less metadata, only storing the author of the merge, the log,
-the commit date, and the author date, and potentially, changes necessary to address
-conflicts during the merge. The details of the model are outlined in
-\cite{German2015}.
-
-\section{Design and Implementation}
-
-To navigate and inspect the merge-tree view of the kernel we created a web-based
-tool called \tool. Creating a web-based tool enables users to use the system without
-having to install additional software or store a large database, making it more
-accessible, more easily maintainable, and platform independent. \tool uses the
-following mechanisms to reach our goals of better navigation and better explanation
-of the selected changes.
-
-\begin{itemize}
-        \item Filter by searching
-        \item View files edited
-        \item View modules
-        \item Tree viewer
-\end{itemize}
-
-\subsection{Searching}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View allows filtering commits that were merged in a given
-                period, filtering by author, keyword, or commit ID.}
-        \label{fig:search}
-\end{figure}
-
-Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
-merges that are irrelevant. The search mechanism breaks down the results by release
-version. A user can further narrow down the search by specifying a range of dates in
-which such commits were merged by Linus into the master branch---not when the commits
-were created (author date and/or commit date). This distinction is important. We have observed commits that have
-taken years to arrive into the master repository after they were originally
-created.
-
-A user may then provide a search text, filtering by the author name, the commit ID,
-or keyword from the log. Any part of the author name may show up in the results,
-including searching by email address.
-
-If the user is searching by commit ID, the ID can be specified by using any of its unique prefixes. For example, the commit
-\mycode{3f17ea6dea8ba5668873afa54628a91aaa3fb1c0} is returned when the user searches for a
-commit ID of \mycode{3f17e} in the 3.16 Linux kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
-        \caption{Search Results. Each table entry is a commit merged in the desired
-                merged window.}
-        \label{fig:results}
-\end{figure}
-
-In the search results (seen in Figure~\ref{fig:results}), the user is presented with the one-line log message preview,
-the author's name and email, the date the commit was authored, and the date the commit was last committed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
-        \caption{Panel showing the Commit Message of a commit.}
-        \label{fig:message}
-\end{figure}
-
-Once a user selects a commit or merge to investigate, they are presented with a
-tabbed pane allowing them to view the full commit log, the files edited, the modules
-involved, and the merge-tree view.
-
-The first tab displays the full commit log (Figure~\ref{fig:message}). From this, a
-user is able to see what they would see had they searched for the commit using Git
-log. This doesn't provide additional information to the other tools, but helps to
-complete the functionality of \tool. The commit log provides a user with the
-information about the content of the commit and who has signed-off on the commit to
-ensure that it is of good quality. The message for merges may contain a summary of
-the commits being merged.  The information within these messages is highly variable,
-and is completely dependent on the author's style. As the user moves toward the
-root-level merge, the quality of these messages generally improves.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
-        \caption{Panel showing the files modified by all the commits that are part
-                of this merge.}
-        \label{fig:files}
-\end{figure}
-
-The second tab is the files tab (Figure~\ref{fig:files}). This tab provides
-information on what files have been edited, how many lines were added, and how many
-lines were removed in a given commit. For non-merges, this functionality is similar
-to the other tools available. Our tree-based design model allows us to extend this
-functionality to merges by aggregating information about all the commits that are
-children of the merge in the merge-tree, which other tools are unable to show. To
-find the number of lines added to a file in a merge, we take the sum of the lines
-added to that file in each of the children of that merge. We do the same for
-calculating the number of lines removed.
-
-% This could be extended to use the patch information to determine when a line has been
-% replaced rather than incrementing the lines added and removed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
-        \caption{Panel showing the modules changed by all the commits in this merge.}
-        \label{fig:modules}
-\end{figure}
-
-The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
-within the commit. Modules are not natively recognized by Git, and are not going to
-be present in all repositories. In the Linux repository, authors put the module they
-are working on in the one-line summary of the log-message; for example: the log
-\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
-coverage testing tool of the kernel). We heuristically extract the module by taking
-all text in the log summary of commits until the first colon.  Modules are logical
-partitions of the information in the kernel. Depending on where the author was
-working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
-quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
-few cases, the author of a commit does not correctly follow this format and the
-heuristic approach fails. As with the Files panel, non-merge commits show their
-corresponding information, but for merge commits we aggregate all the modules
-changed in all the commits that are part of that merge.  The output of this view is
-shown in Figure~\ref{fig:modules}.
-
-Finally, we have the Tree view tab. The tree view is designed for providing easy
-navigation of the commits within the merge-tree that is rooted in the current merge.
-It also provides a clear topological view of the merge and the submerges it
-includes. We have experimented with various tree designs to find a design that
-allows for easy navigation and visualization of both large and small trees. We
-discuss this panel in the next subsection.
-
-\subsection{Merge-Tree Views} \label{treeview_section}
-
-The merge-tree view is what makes \tool unique to other tools that inspect the DAG
-of a Git repository. With it, a user can inspect how commits are merged on their way
-to the master branch. We have experimented with various types of trees:
-
-\begin{enumerate}
-        \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
-        \item Reingold-Tilford trees provide a clear visual representation of the tree structure of the merge-tree.
-        \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
-                clearly showing the parent-child relationships between commits and merges.
-\end{enumerate}
-
-\subsubsection{List Tree}
-
-The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of nested lists,
-and is designed to more closely model the tree view found in file browsers. This
-tree only contains the commits and merges that are within the subtrees of the
-current merge. A commit will never have any items in this tree as it is a leaf. To
-accompany the tree, we include breadcrumbs at the top of the page to enable a user
-to navigate both from the root to the leaves and from a leaf to the root. The last
-item in the breadcrumb list is the current commit, the previous item is the parent
-of the current commit, and the first item is the root merge into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List tree view of merge 3f17ea6}
-        \label{fig:list_tree}
-\end{figure}
-
-\subsubsection{Reingold-Tilford Tree}
-
-The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:reingold_tree} and
-\ref{fig:reingold_tree_zoom}) allows the visualization and navigation of the entire merge-tree in
-an intuitive representation of the tree.  This illustrates a clear notion of root
-and leaves, and how to navigate in either direction. Some merge trees are very
-large, containing thousands of commits and merged. While the tree is capable of
-producing a visualization, it becomes far more difficult to understand. For example,
-the merge, 3f17ea6 performed by Linus Torvalds June 8 2014, contains 7217 commits
-and merges. \footnote{This commit can be inspected at
-  \url{http://li.turingmachine/org/commits/3f17ea6dea8ba5668873afa54628a91aaa3fb1c0}}
-
-      \begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/4dc4_tree.pdf}
-        \caption{Merge 4dc4226 is a subtree of 3f17ea6, for the power-management
-          module of the kernel}
-        \label{fig:reingold_tree}
-      \end{figure}
-
-      \begin{figure*}
-        \centering
-        \includegraphics[width=0.98\textwidth]{figures/4dc4_zoom_tree.pdf}
-        \caption{Zoomed out view of the Reingold-Tilford tree for merge 4dc4226}
-        \label{fig:reingold_tree_zoom}
-      \end{figure*}
-
-      The user is initially greeted with their current node centered on the screen.
-      They are able to zoom the tree by scrolling the mouse, and clicking and
-      dragging to pan the tree. They can see more details about a node by clicking
-      on it, which will provide them a link to the specific page for that commit.
-
-\subsubsection{Bubble Tree}
-
-Bubble trees are useful for providing a clear visualization of wide, hierarchical
-data\cite{Boardman2000}. The tree structure is represented by the nesting of nodes;
-the largest circle is the root, containing all the other nodes. The smallest circles
-do not contain any nodes and therefore represent the leaf nodes.
-
-Our implementation of the bubble tree (Figure~\ref{fig:bubble_tree}) provides the
-user with a clear picture of where a commit is located in the merge tree. We
-highlight the selected commit or merge in red. Non-selected merges are in a shade of
-blue determined by the depth of a node in the tree. The root is the lightest shade
-of blue, while the contained merges are progressively darker; the commits are white.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree of merge 3f17ea6, merge 4dc4226 is the currently
-          selected merge and is highlighted in orange.}
-        \label{fig:bubble_tree}
-\end{figure}
-
-The bubble tree doesn't have an implicit way of providing additional information,
-placing any text near the nodes makes the tree impossible to read, so we include a
-separate pane in the web page. When a user hovers over a node, the pane shows
-additional information about the author and the commit message and a link to the
-detailed page for that commit or merge. If the user clicks on a node, the tree will
-zoom to that node and the information in the info pane will persist, enabling the
-user to click the link. This tree provides an easy mechanism for users to navigate
-from the root to the leaves and vice versa.
-
-\section{Implementation Details}
-
-\evantodo{See if we can move API section to be part of another section}
-
-The front-end uses asynchronous requests to gather the information for the trees and
-tables. This enables third-parties to implement new front-ends, trying other
-designs, though this interface may change in the future. The tree, file, and module
-information is accessible through
-http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.
-
-The response for the tree is a single object. This object is the root node of the
-tree, and may contain an object of children sub-tree objects in the \verb|children|
-field, which are in the same format as the root object. The children object uses the
-commit ID as the key.
-
-The tree node objects have the following fields:
-
-\begin{tabular}{ccl}
-        Field name      & Data type & Description\\\hline
-        \verb|cid|      & string    & Commit ID\\
-        \verb|name|     & string    & One-line log summary\\
-        \verb|mlinus|   & string    & Root merge commit ID\\
-        \verb|author|   & string    & Author name and email\\
-        \verb|mnext|    & string    & Parent merge commit ID\\
-        \verb|children| & object    & Object of tree objects\\
-\end{tabular}
-
-The file responses contain only the files that the selected merge or commit works
-with. The response is a single object in the form of a tree. This object is the root
-of the tree, and represents the current merge or commit. If the current position in
-the tree is an inner node, the response will contain the child nodes in the
-\verb|children| field, otherwise \verb|children| will be an empty list. The format of file
-response is as follows,
-
-\begin{tabular}{lll}
-        Field name & Data type & Description\\\hline
-        \verb|cid| & string & Commit ID\\
-        \verb|mnext| & string & Parent merge commit ID\\
-        \verb|children| & list & List of tree objects\\
-        \verb|files| & list of tuples&
-        \footnotesize{
-                \begin{tabular}{lc}
-                        \verb|Filename| & string\\
-                        \verb|Lines added| & uint\\
-                        \verb|Lines removed| & uint\\
-                \end{tabular}}\\
-\end{tabular}
-
-\section{Discussion}
-
-From a developers point of view, the are two major disadvantages of the DAG model of
-Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
-commit that succeeds them on the path to being integrated; and b) the DAG contains
-many more edges that are necessary to understand how integration occurred. We
-addressed these two issues with the creation of the merge-tree model from the DAG of
-Linux. Effectively, the merge-tree recovers the details of how each branch---and the commits they contain---was merged
-into the master branch.
-
-In our experience, no other Git repository reaches the level of DAG complexity that
-Linux has. In most Git repositories, most merges are not nested---most merges merge
-directly into the master repository, and these merges only merge a few commits. Even
-in these simplified cases, the merge-tree can provide a valuable summary of how
-commits are integrated into the master branch, specifically since the time of
-integration may be very different from the time the commits were authored.
-
-\evan{Check accuracy of this paragraph}
-
-The biggest challenge is computing the merge-tree of any given repository. It
-is likely that our heuristic for computing the merge-tree will not work with other
-repositories. The heuristic requires that there are no fast-forward merges into the
-master branch of the repository (a majority of repositories), and that there are no
-foxtrot merges (a practice that is starting to be considered desirable by git users\footnote{See
-  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git} and
-  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
-We are able to use our heuristic with the Linux kernel because of
-the strict integration model imposed by Linus. To validate the results of the
-heuristic, we use the continuous mining technique described in \cite{German2015}.
-Continuous mining of the repository requires foresight and planning.
-
-Because \tool leverages the merge-tree model of the kernel, it provides mechanisms
-and visualizations that other tools are unable to produce. It gives us the ability to
-see how commits are being merged into the master branch, allowing us to better understand
-how code is integrated into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
-        \caption{Merge Dag View}
-        \label{fig:dag_view}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
-        \caption{Merge tree topology for the same merge as in
-                Figure~\ref{fig:dag_view}.}
-        \label{fig:tree_view}
-\end{figure}
-
-Our goal was to build a tool to enable maintainers to effectively navigate and
-browse the changes performed to the kernel over the period of a release. Achieving
-this goal includes removing information that does not pertain to the area of the
-kernel that the user is interested in. In Figures~\ref{fig:dag_view} and
-\ref{fig:tree_view} we can visually compare the results returned from Gitk
-and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. Note that Gitk has no way to show or recognize that this
-is a merge into the master branch.
-
-The log information to this merge-commit in both of these figures is identical, but
-the presentation drastically changes our ability to comprehend what we are seeing.
-The primary difference is the removal of irrelevant information. In Gitk
-(Figure~\ref{fig:dag_view}), the visualization results include commits and merges
-from other components of the kernel, while our tool (Figure~\ref{fig:tree_view})
-only includes the results that are specific to the component of the kernel that we
-are interested in.
-
-Even though this merge is relatively small, containing only four sub-merges, three of which
-only contain a single commit, and the last containing two commits, it is already hard to visualize in Gitk.
-With \tool we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
-furthermore, users must work to determine where this merge ends and the next one
-begins.
-
-\tool further enhances our understanding by summarizing the files
-and modules that were edited in the entire branch being merged. We are able to determine that three
-files, ``palamas-regulator.c'' had two lines added and two lines removed,
-``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
-lines added and 2 lines removed. Finally, we are able to determine that only the
-``regulator'' module was modified in this merge and was modified by 5 commits.
-
-\section{Future Work}
-
-There are still many areas that can be improved before the full potential of our
-model can be realized. Here, we outline various areas of the tool that still need
-more attention.
-
-\subsection{Files}
-
-There currently is no functionality surrounding searching by filename. Users falling
-under use-case 2 may know what file they are editing and try to determine how it may
-work with the other commits and merges in the module. It is possible that a third
-use-case may arise, where the user wants to determine all commits that effect a
-given file. In both cases, bottom-to-top approach is applied.
-
-There is limited functionality to the presentation of the file information. At a
-minimum, the patches for the commits can be displayed. From there, the patches can
-be used to piece together parts of the file to generate a single patch at a given
-merge rather than displaying each patch individually. The patches can also be used
-for determining what kinds of changes were made, if the lines are being added,
-removed, or being replaced.
-
-\subsection{Authorship}
-
-Our model can aggregate more information than what we have implemented. The
-authorship information is important for licensing purposes, and we can show all
-authors contributing to commits in a  merge and how many commits they contributed to
-the merge. We could go further, providing information about what files they edited,
-how many lines they added and removed, and more.
-
-\subsection{Evaluation}
-
-At this time, we have no evidence that our tool is able to improve the work-flow of
-maintainers. We believe that the tool is able to improve the work-flow and
-performance of maintainers because it provides cleaner mechanisms of visualizing and
-presenting the commit information. It is able to provide more relevant information,
-while removing information that is irrelevant to a given module or set of merges.
-
-We can either perform user-testing to show how a users workflow changes, or we could
-have maintainers evaluate and critique the tool and use their feedback to determine
-if the tool meets our goals.
-
-\section{Related Work}
-
-Version control systems monitor the development lifetime of software projects. This
-makes the version control system vital in providing information about how a software
-project is being developed. To our knowledge, we know of no Git repository
-visualization tool that builds a tree that maps the path a commit follows to the
-master branch from the DAG provided by Git. This may be because more information is
-required to generate the merge-tree model than what is stored by Git. However, there
-has been a lot of work in providing visualizations of various repositories.
-
-Many tools work to address the issues in communication between developers in
-inter-team collaboration work. Codebook \cite{Begel2010} uses a data mining
-technique  to determine the developer of a piece of code, the program manager who
-wrote the specification for the code, and the program managers and developers on the
-team who were working together. Hipikat\cite{Cubranic2005} is another tool with a
-focus on communication. Where Codebook focuses on developers working on a project,
-Hipikat is focused on enabling easier integration of new developers to a project by
-providing them with easily-searchable artifacts of the changes made. Codebook is
-useful for pairing a contributor with the original developer; however, the developer
-may not have worked with the piece of code in years. A Hipikat program may provide
-more information to the maintainer as it records the artifacts of why certain
-design decisions were made when they were made using other tools like Bugzilla and
-CVS. Neither tool is sufficient in meeting our goals to provide a summary of the
-topology of the kernel repository through a visual tree.
-
-Most visualizers provide a visual presentation of a certain aspect of a repository.
-Fractal Figures \cite{Ambros2005} uses a unit square to represent a portion of a
-project, then partitions the square based on the proportion of an author's
-contributions to that portion of the project. EPOSee\cite{Burch2005} and Evolution
-Radar\cite{Ambros2009} perform further analysis, determine which files are made
-together, and what changes are made over a sequence of commits, though the goals
-behind these projects is different.
-
-Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work with data
-from CVS repositories. Our goal is to provide information about Git repositories,
-specifically the Linux master repository. Fewer tools are available for generating
-visualizations and summaries of Git repositories, potentially due to the DAG model
-used by Git.
-
-Gource is a tool for providing an interactive timelapse of the state of a
-repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
-type of contribution a developer is making. These contribution types are one of,
-adding a file, removing a file, and changing a file. While the timelapse is
-interesting to watch, it does not provide any additional explanation of the changes
-actually being made, only the frequency that they are being made and who is making
-them.
-
-The current industry standard tool for Git repository visualization is Gitk. The
-Gitk interface is built around the central DAG viewer. The DAG view displays the
-commits on their respective branches, the author of the commit or merge, and the
-date the commit or merge was authored. A user can select a single merge or commit
-to view additional information. For both merges and commits, Gitk will present the
-user with the Git log. If the user selected a commit, Gitk will also present the
-user with the patch information and the names of the files edited. Gitk is unable to
-provide the patches and filenames information in merges, as it is unable to
-aggregate commit information in merges.
-
-Our tool is primarily aimed at presenting the hierarchical structure of the Linux
-git repository. We use tables for presenting the summarized information of the
-commits and merges, but this information could also be presented in a graphical
-form. Various graphical forms for displaying file and authorship data exist, the
-principal forms being matrix views, city scapes, bar and pie charts, and networks
-\cite{Eick2002}. Any of these data visualization metaphors are applicable to our
-system.
-
-\section{Conclusion}
-
-Our tool, \tool, shows promising results in working toward our goals, building a tool that
-can more easily navigate the repository, and provide clearer explanations of the
-changes made. The filtering by searches and the various tree visualizations enable
-users to easily navigate through the kernel, finding the commits and merges that
-pertain to what they are interested in. The tree visualizations are able to both
-serve as a navigation piece, providing simpler navigation through the kernel, but
-also demonstrate which commits contribute to which merges, providing further
-explanation over the DAG. The tree model further improves the explanation by enabling
-the tool to aggregate metadata at the merges instead of requiring a user to manually
-aggregate the information. Our tool currently aggregates file information and the
-modules edited, but can be extended to include authorship information and aggregated
-patch information.
-
-\evan{Check the validity of this paragraph}
-We use a heuristic method to generate the tree from the repository. The heuristic
-requires that the repository does not have fast-forward merges into the master
-branch or foxtrot merges, which obfuscate the master branch. In either case, the
-heuristic will break and result in a tree that does not accurately represent the
-repository.
-
-These may pose as significant drawbacks to smaller projects; however, with a proper
-merge discipline, the heuristic is able to correctly convert the DAG into a
-merge-tree. The importance of clear visualizations in large projects outweighs
-the costs of enforcing a clean merge discipline. The tree-based model appears to
-have the ability to provide a clean visualization and more detailed explanations of
-the changes in the Linux kernel.
-
-% \nocite{*}
-
-\bibliographystyle{IEEEtran}
-\bibliography{citations}
-
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2017-06-11 paper/introduction.tex
Adding files for introduction and references

Merge-trees, as presented by Wilde and German\cite{Wilde2016}, are a data structure
designed to make the navigation and summarization of git merges simpler than using
the directed acyclic graph (DAG), as done in gitk among other tools\evan{does this
  need a citation?}. Git uses the DAG model internally, making it simple to
construct tools directly on the DAG;\@ however, this model does not provide a clear
summarization of the events occuring in the repository. Gitk and the git command
line interface (cli) both provide the user with a visualization of the DAG of the
repository.

### 2017-06-11 paper/generalization.tex
Adding content for generalization and userstudy
This is just preliminary content for these sections, but it is heading
in the right direction.

The original algorithm uses a stack-like approach, which results in a depth-first
search of the tree. In the first pass, all of the commits that are in the
fist-parent position of the previous commit will be analyzed. Then following the
depth-first tactic, the tree that was earliest in the history of the repository will
be constructed, followed by the next earliest. This will be repeated until all of
the trees have been constructed and we reach the head of the master branch.

\evan{Provide a description of the new algorithm, and an DAG, and finally the
  resulting merge tree.}

### 2017-06-11 paper/userstudy.tex
Adding content for generalization and userstudy
This is just preliminary content for these sections, but it is heading
in the right direction.

We conducted a user study of \evan{NUMBER} participants to determine if there is a
statistically significant difference among users in performing certain tasks on
repositories between \textit{Linvis} and either gitk or the command-line git
interface. We were looking for improvements in accuracy, time performance, and
overall enjoyment among users when performing summarization and intra-tree
navigation, as these are the tasks that merge-trees are design to simplify.

\evan{still need to figure out who our intended user is... ``git user'' is a bit too
  broad.}

% TODO: Determine who the intended user is. "generic git user" is kind of broad.
% TODO: Add a note stating that we will refer to the primary branch as being the master branch

\subsubsection{Methodology}
\label{ssub:methodology}

\subsubsection{Commit Selection}
\label{ssub:commit_selection}

We selected two commits for the participant to draw and summarize by first
determining the size of tree in the first and second quartiles. We found that
between Linux 3.1 and 3.16, the size of the trees in the first quartile only
contained a single node, and that the size of the trees in the second quartile
resulted in trees up to seven nodes. The third quartile contained trees up to 51
nodes, and the fourth quartile contained one tree of 7217 nodes. It is worth
mentioning that the tree sizes in the fourth quartile increase very quickly, as the
next largest tree only contains 4708 nodes, and the third largest tree contains 2349
commits.

% TODO: Finish the paragraph on complexity
The complexity of the merge trees varies greatly.

The complexity of these trees varies greatly as well, . % TODO: Finish this sentence

We decided to work with one tree of one node, and one tree of seven nodes.

\subsubsection{Questions}
\label{ssub:questions}

We chose questions that were designed to test the two primary goals of merge trees,
compared with the results from gitk or command line git.

\subsubsection{Results}
\label{ssub:results}

\subsubsection{Discussion}
\label{ssub:discussion}

### 2017-06-11 paper/abstract.tex
Re-wording abstract

Natively, git uses a directed acyclic grph mode (DAG) with references pointing from
the (TODO: most recent commit?, it's not the HEAD because that moves around) toward
the initial commit in the repository. This model is useful and necessary to the
internal functionality of the version control system, but does not provide users
with a clear nderstand of the events occuring in the repository. Due to the
direction of the link in the DAG, visualziation tools, like gitk, are unable to
provide summarized aggregation of the commits at a given merge. Furthermore, the
visual DAG does not provide a clear explanation of the events taking place in the
repository.

Merge-Trees are a data structure derived from the DAG, that could potentially
alleiveate the issues with visualization built directly from the DAG. The trees are
defined such that the root is the merge into the master branch, and the leaves are
the individual commits. The references point in both directed, from each node to the
parent, and from the parent to the children. Depending on the complexity of the
merge, the path from leaf to root may pass through multiple internal nodes.

In this paper, we verify that the Merge-tree model is able to simplify the
construction of a conceptual image of the parts of a repository, and simiplify the
summarization tasks through a user study. We will continue by providing a more
generalized approach to constructing merge-tress, which will prune components of the
DAG that are not necessary.

% \evan{Change this when we get the actual results, this is from the pilot study} We
% find that although the DAG view used by Gitk provides more information about the
% repository, in many cases it povides too much information and overwhelms the user.

% \evan{Move this}
% In some cases, where the structure of the merge tree is flat, that is to say that it
% has no inner merges, users are able to easily determine which commits are merged
% together, but are unable to determine which merge is the final merge is the root of
% the tree.

### 2017-06-12 paper/userstudy.tex
Adding Tree size figure
Adds figure to paper, and to the makefile.

@@
 next largest tree only contains 4708 nodes, and the third largest tree contains 2349
-commits.
+commits. We see in figure~\ref{fig:tree_size} how the number of trees decreases as
+the size of the tree increases.
+
+\begin{figure}[htpb]
+\begin{center}
+\caption{Number of merge tree occurrences over size of the merge tree}
+\label{fig:tree_size}
+\begin{tikzpicture}
+  \begin{axis} [
+    ylabel=Occurences,
+    xlabel=Tree Size,
+    grid=both,
+    minor x tick num = 1,
+    minor y tick num = 1,
+    ymode=log
+    ]
+    \addplot[chartred] table[col sep=comma, x index={1}, y index={0}]{data/tree_size.csv};
+  \end{axis}
+\end{tikzpicture}
+\end{center}
+\end{figure}
 

### 2017-06-12 paper/abstract.tex
Resizing text width
84-wide text doesn't fit split-screen on my little monitor.
Resizing it down to 72-wide.

@@
-Natively, git uses a directed acyclic grph mode (DAG) with references pointing from
-the (TODO: most recent commit?, it's not the HEAD because that moves around) toward
-the initial commit in the repository. This model is useful and necessary to the
-internal functionality of the version control system, but does not provide users
-with a clear nderstand of the events occuring in the repository. Due to the
-direction of the link in the DAG, visualziation tools, like gitk, are unable to
-provide summarized aggregation of the commits at a given merge. Furthermore, the
-visual DAG does not provide a clear explanation of the events taking place in the
-repository.
+Natively, git uses a directed acyclic graph model (DAG) with references
+pointing from the most recent commit (TODO: most recent commit?, it's
+not the \emph{HEAD} because that moves around) toward the initial commit
+in the repository. This model is useful and necessary to the internal
+functionality of the version control system, but does not provide users
+with a clear nderstand of the events occuring in the repository. Due to
+the direction of the link in the DAG, visualziation tools, like gitk,
+are unable to provide summarized aggregation of the commits at a given
+merge.  Furthermore, the visual DAG does not provide a clear explanation
+of the events taking place in the repository.
 
-Merge-Trees are a data structure derived from the DAG, that could potentially
-alleiveate the issues with visualization built directly from the DAG. The trees are
-defined such that the root is the merge into the master branch, and the leaves are
-the individual commits. The references point in both directed, from each node to the
-parent, and from the parent to the children. Depending on the complexity of the
-merge, the path from leaf to root may pass through multiple internal nodes.
+Merge-Trees are a data structure derived from the DAG, that could
+potentially alleiveate the issues with visualization built directly from
+the DAG. The trees are defined such that the root is the merge into the
+master branch, and the leaves are the individual commits. The references
+point in both directed, from each node to the parent, and from the
+parent to the children. Depending on the complexity of the merge, the
+path from leaf to root may pass through multiple internal nodes.
 
-In this paper, we verify that the Merge-tree model is able to simplify the
-construction of a conceptual image of the parts of a repository, and simiplify the
-summarization tasks through a user study. We will continue by providing a more
-generalized approach to constructing merge-tress, which will prune components of the
-DAG that are not necessary.
+In this paper, we verify that the Merge-tree model is able to simplify
+the construction of a conceptual image of the parts of a repository, and
+simiplify the summarization tasks through a user study. We will continue
+by providing a more generalized approach to constructing merge-tress,
+which will prune components of the DAG that are not necessary.
 
-% \evan{Change this when we get the actual results, this is from the pilot study} We
-% find that although the DAG view used by Gitk provides more information about the
-% repository, in many cases it povides too much information and overwhelms the user.
+% \evan{Change this when we get the actual results, this is from the
+% pilot study} We find that although the DAG view used by Gitk provides
+% more information about the repository, in many cases it povides too
+% much information and overwhelms the user.
 
-% \evan{Move this}
-% In some cases, where the structure of the merge tree is flat, that is to say that it
-% has no inner merges, users are able to easily determine which commits are merged
-% together, but are unable to determine which merge is the final merge is the root of
-% the tree.
+% \evan{Move this} In some cases, where the structure of the merge tree
+% is flat, that is to say that it has no inner merges, users are able to
+% easily determine which commits are merged together, but are unable to
+% determine which merge is the final merge is the root of the tree.

### 2017-06-12 paper/introduction.tex
Resizing text width
84-wide text doesn't fit split-screen on my little monitor.
Resizing it down to 72-wide.

@@
-Merge-trees, as presented by Wilde and German\cite{Wilde2016}, are a data structure
-designed to make the navigation and summarization of git merges simpler than using
-the directed acyclic graph (DAG), as done in gitk among other tools\evan{does this
-  need a citation?}. Git uses the DAG model internally, making it simple to
-construct tools directly on the DAG;\@ however, this model does not provide a clear
-summarization of the events occuring in the repository. Gitk and the git command
-line interface (cli) both provide the user with a visualization of the DAG of the
-repository.
+Merge-trees, as presented by Wilde and German\cite{Wilde2016}, are a
+data structure designed to make the navigation and summarization of git
+merges simpler than using the directed acyclic graph (DAG), as done in
+gitk among other tools\evan{does this need a citation?}. Git uses the
+DAG model internally, making it simple to construct visualization tools
+directly using the DAG of the repository;\@ however, this model does not
+provide a clear summarization of the events occuring in the repository.
+Gitk and the git command line interface (cli) both provide the user with
+a visualization of the DAG of the repository.
+
+Merge-trees have two primary issues; first, the merge-tree model has not
+been evaulated, and cannot therefor claim to simplify the summarization
+and navigation tasks. Second, merge-trees have only been implemented for
+the Linux repository, which follows a strict merge-policy where only
+Linus can merge into the master branch.
+
+The contributions of this paper tackle both of these issues. We perform
+a user-study, to verify that the merge-tree model is able to improve the
+accuracy and speed of summarization tasks. Second, we generalize the
+algorithm, creating an upper bound on the number of commits visisted in
+the construction of the merge-trees.
+
+\subsection{Merge-Trees}
+\label{sub:merge_trees}
+
+Internally, git uses a DAG as the primary data structure for managing
+commits in a repository. An individual commit will have an ordered list
+of parents, where the parent in the first position is on the same branch
+as the commit, and the other parents are in the same order as they were
+merged. In the event that a merge commit contains a list of parents with
+more than two parents, this merge is an octopus merge. The commits may
+be divided into two groups, commits with only a single parent, and
+commits with at least two commits. Commits with with many parents are
+merge commits, combining the effects of potentially many individual
+commits.

### 2017-06-12 paper/userstudy.tex
Resizing text width
84-wide text doesn't fit split-screen on my little monitor.
Resizing it down to 72-wide.
Figure: Merges over total commits
Shows the nearly linear relationship between commits and merges.
Was not actually expecting this.

@@
-We conducted a user study of \evan{NUMBER} participants to determine if there is a
-statistically significant difference among users in performing certain tasks on
-repositories between \textit{Linvis} and either gitk or the command-line git
-interface. We were looking for improvements in accuracy, time performance, and
-overall enjoyment among users when performing summarization and intra-tree
-navigation, as these are the tasks that merge-trees are design to simplify.
+We conducted a user study of \evan{NUMBER} participants to determine if
+there is a statistically significant difference among users in
+performing certain tasks on repositories between \textit{Linvis} and
+either gitk or the command-line git interface. We were looking for
+improvements in accuracy, time performance, and overall enjoyment among
+users when performing summarization and intra-tree navigation, as these
+are the tasks that merge-trees are design to simplify.
 
-\evan{still need to figure out who our intended user is... ``git user'' is a bit too
-  broad.}
+\evan{still need to figure out who our intended user is... ``git user''
+  is a bit too broad.}
 
@@
 
-We selected two commits for the participant to draw and summarize by first
-determining the size of tree in the first and second quartiles. We found that
-between Linux 3.1 and 3.16, the size of the trees in the first quartile only
-contained a single node, and that the size of the trees in the second quartile
-resulted in trees up to seven nodes. The third quartile contained trees up to 51
-nodes, and the fourth quartile contained one tree of 7217 nodes. It is worth
-mentioning that the tree sizes in the fourth quartile increase very quickly, as the
-next largest tree only contains 4708 nodes, and the third largest tree contains 2349
-commits. We see in figure~\ref{fig:tree_size} how the number of trees decreases as
-the size of the tree increases.
+We selected two commits for the participant to draw and summarize by
+first determining the size of tree in the first and second quartiles. We
+found that between Linux 3.1 and 3.16, the size of the trees in the
+first quartile only contained a single node, and that the size of the
+trees in the second quartile resulted in trees up to seven nodes. The
+third quartile contained trees up to 51 nodes, and the fourth quartile
+contained one tree of 7217 nodes. It is worth mentioning that the tree
+sizes in the fourth quartile increase very quickly, as the next largest
+tree only contains 4708 nodes, and the third largest tree contains 2349
+commits. We see in figure~\ref{fig:tree_size} how the number of trees
+decreases as the size of the tree increases.
 
@@
 % TODO: Finish the paragraph on complexity
-The complexity of the merge trees varies greatly.
+The complexity of the merge trees varies greatly as shown in
+figure~\ref{fig:merge_vs_commits}, which shows that as the number of
+commits increases in a merge-tree, the number of merges also increases
+nearly linearly.
 
-The complexity of these trees varies greatly as well, . % TODO: Finish this sentence
+\begin{figure}[htpb]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis} [
+      ylabel=Total Merges,
+      xlabel=Total Commits,
+      grid=both
+      ]
+      \addplot[chartblue] table[col sep=comma, x index={1}, y index={2}] {data/merge_counts.csv};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Comparison between the number commits and the number of merge commits}
+  \label{fig:merge_vs_commits}
+\end{figure}
 
-We decided to work with one tree of one node, and one tree of seven nodes.
+We decided to work with one tree of one node, and one tree of seven
+nodes.
 
@@
 
-We chose questions that were designed to test the two primary goals of merge trees,
-compared with the results from gitk or command line git.
+We chose questions that were designed to test the two primary goals of
+merge trees, compared with the results from gitk or command line git.
 

### 2017-06-12 paper/userstudy.tex
Adding Figures for userstudy
Added figures showing how the number of merges increase as the size of
the tree increases and how the percentage of commits that are
merge-commits change as the size of the tree increases.

--- 2017-06-11 paper/userstudy.tex
@@
-We conducted a user study of \evan{NUMBER} participants to determine if there is a
-statistically significant difference among users in performing certain tasks on
-repositories between \textit{Linvis} and either gitk or the command-line git
-interface. We were looking for improvements in accuracy, time performance, and
-overall enjoyment among users when performing summarization and intra-tree
-navigation, as these are the tasks that merge-trees are design to simplify.
+We conducted a user study of \evantodo{NUMBER} participants to determine if
+there is a statistically significant difference among users in
+performing certain tasks on repositories between \textit{Linvis} and
+either gitk or the command-line git interface. We were looking for
+improvements in accuracy, time performance, and overall enjoyment among
+users when performing summarization and intra-tree navigation, as these
+are the tasks that merge-trees are design to simplify.
 
-\evan{still need to figure out who our intended user is... ``git user'' is a bit too
-  broad.}
+\evan{still need to figure out who our intended user is... ``git user''
+  is a bit too broad.}
 
@@
 
-We selected two commits for the participant to draw and summarize by first
-determining the size of tree in the first and second quartiles. We found that
-between Linux 3.1 and 3.16, the size of the trees in the first quartile only
-contained a single node, and that the size of the trees in the second quartile
-resulted in trees up to seven nodes. The third quartile contained trees up to 51
-nodes, and the fourth quartile contained one tree of 7217 nodes. It is worth
-mentioning that the tree sizes in the fourth quartile increase very quickly, as the
-next largest tree only contains 4708 nodes, and the third largest tree contains 2349
-commits.
+We selected two commits for the participant to draw and summarize by
+first determining the size of tree in the first and second quartiles. We
+found that between Linux 3.1 and 3.16, the size of the trees in the
+first quartile only contained a single node, and that the size of the
+trees in the second quartile resulted in trees up to seven nodes. The
+third quartile contained trees up to 51 nodes, and the fourth quartile
+contained one tree of 7217 nodes. It is worth mentioning that the tree
+sizes in the fourth quartile increase very quickly, as the next largest
+tree only contains 4708 nodes, and the third largest tree contains 2349
+commits. We see in figure~\ref{fig:tree_size} how the number of trees
+decreases as the size of the tree increases. The results for the other
+plots become very noisy, since there is only one occurrence for any tree
+with more than 947 commits, and only two occurrences of any trees with
+size greater than 336. Due to the noise,
+figures~\ref{fig:tree_size_filter},~\ref{fig:merge_count_filter},
+and~\ref{fig:percentage_filter} are only on the sizes of trees where
+there were at least three occurrences of trees of that size.
 
-% TODO: Finish the paragraph on complexity
-The complexity of the merge trees varies greatly.
+\pgfplotstableread[col sep=comma]{data/merge_counts.csv}\mergetable
 
-The complexity of these trees varies greatly as well, . % TODO: Finish this sentence
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+      ylabel=Occurences,
+      xlabel=Total Commits,
+      grid=both,
+      minor x tick num = 1,
+      minor y tick num = 1
+      ]
+      \addplot[chartblue] table[col sep=comma, x index=1, y index=0]{data/tree_size.csv};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Frequency of trees of varying sizes}
+  \label{fig:tree_size}
+\end{figure}
 
-We decided to work with one tree of one node, and one tree of seven nodes.
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+      ylabel=Occurences,
+      xlabel=Total Commits,
+      grid=both,
+      minor x tick num = 1,
+      minor y tick num = 1
+      ]
+      \addplot[chartblue] table[x index=0, y index=1]\mergetable;
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Frequency of trees of varying sizes}
+  \label{fig:tree_size_filter}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis} [
+      ylabel=Merges,
+      xlabel=Total Commits,
+      grid=both,
+      legend pos = north west
+      ]
+      \addplot[chartred] table[x index=0, y index=5]\mergetable;
+      \addlegendentry{Maximum}
+      \addplot[chartyellow] table[x index=0, y index=6]\mergetable;
+      \addlegendentry{Mean}
+      \addplot[chartgreen] table[x index=0, y index=4]\mergetable;
+      \addlegendentry{Median}
+      \addplot[chartblue] table[x index=0, y index=3]\mergetable;
+      \addlegendentry{Minimum}
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Number of merges versus the size of the merge tree}
+  \label{fig:merge_count_filter}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis} [
+      ylabel=Merge Percentage,
+      xlabel=Total Commits,
+      grid=both
+      ]
+      \addplot[chartpurple] table[x index=0, y index=7]\mergetable;
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Percentage of commits in a merge-tree that are merge commits}
+  \label{fig:percentage_filter}
+\end{figure}
+
+We decided to work with one tree of one node, and one tree of seven
+nodes.
 

### 2017-06-15 paper/futurework.tex
Preliminary future work section

% Inter Navigation
% TODO: verify this with the study
While the merge-tree model provides users with a better conceptual image
of what is occuring within a single merge tree, we have not exprimented
or tested the potential gains of navigation between trees. At this time,
the Linvis search provides results grouped by merge-tree root instead of
as a single block of results. The results are generally similar, with
each merge tree being merged at a different point in time, or to a
different version of the kernel. A future extension of this may
incorporate some information about how a given cluster changes over
time. To do this, a technique would need to be developed to cluster
merge-trees into modules, and then ordering the merge-trees based on
time or release.

% TODO: ensure that linvis has been referenced/defined

### 2017-06-15 paper/userstudy.tex
Commit selection description and figures

@@
       xlabel=Total Commits,
+      minor x tick num = 1,
+      minor y tick num = 1,
       grid=both,
@@
 
-We decided to work with one tree of one node, and one tree of seven
-nodes.
+As 25\% of the trees consist of a single commit, we selected one tree
+consisting of a single commit for the study. 50\% of the trees contained
+up to seven nodes, so we selected another tree of size seven.  Of these
+trees, we restricted the selection to trees consisting of at least one
+merge, not including the merge into the master branch.
+
+From the list of trees consisting of a single node, we selected one at
+random using a simple python script and the \verb|random.choice()|
+function from python 3.6.1, which returned the merge tree rooted at
+\emph{11df5864075f763ec0d1fdecd6a3f0af7d09a553}. As we are not testing
+the inter-tree and tree-search capabilities of \emph{Linvis}, we start
+the user at the corresponding commit. In this case, there is only a
+single option, commit \emph{a3c1239eb59c0a907f8be5587d42e950f44543f8},
+shown in Figure~\ref{fig:commit_1}.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.2\linewidth]{figures/commits/1-commit.pdf}
+  \caption{The first merge tree used in the user study, a merge tree
+    containing a single commit}
+  \label{fig:commit_1}
+\end{figure}
+
+Using the same merge selection technique, we selected the tree rooted at
+\emph{8eb88c80d444fd249edaa7d895666cde79e7b3b8}. To select the starting
+commit, we use the python random choice function on the list of commit
+hashes contained in this tree. This resulted in the selection of commit
+\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, shown in
+Figure~\ref{fig:commit_2}.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
+  \caption{The second merge tree used in the user study, a merge tree
+    containing seven commits}
+  \label{fig:commit_2}
+\end{figure}
 

### 2017-06-19 paper/abstract.tex
Minor rewording in abstract
Natively -> Internally; There are no other options

Fixes spelling of alleviate

Places TODO in a notice command.

@@
-Natively, git uses a directed acyclic graph model (DAG) with references
-pointing from the most recent commit (TODO: most recent commit?, it's
-not the \emph{HEAD} because that moves around) toward the initial commit
-in the repository. This model is useful and necessary to the internal
-functionality of the version control system, but does not provide users
-with a clear nderstand of the events occuring in the repository. Due to
-the direction of the link in the DAG, visualziation tools, like gitk,
-are unable to provide summarized aggregation of the commits at a given
-merge.  Furthermore, the visual DAG does not provide a clear explanation
-of the events taking place in the repository.
+Internally, git uses a directed acyclic graph model (DAG) with references
+pointing from the most recent commit \notice{(TODO: most
+    recent commit?, it's not the \emph{HEAD} because that moves
+    around)} toward the initial commit in the repository. This model is
+useful and necessary to the internal functionality of the version
+control system, but does not provide users with a clear nderstand of the
+events occuring in the repository. Due to the direction of the link in
+the DAG, visualziation tools, like gitk, are unable to provide
+summarized aggregation of the commits at a given merge.  Furthermore,
+the visual DAG does not provide a clear explanation of the events taking
+place in the repository.
 
 Merge-Trees are a data structure derived from the DAG, that could
-potentially alleiveate the issues with visualization built directly from
+potentially alleviate the issues with visualization built directly from
 the DAG. The trees are defined such that the root is the merge into the

### 2017-06-19 paper/userstudy.tex
User study
Adding task types to user study methodology

@@
 there were at least three occurrences of trees of that size.
+
+\pgfplotstableread[col sep=comma]{data/comdate_merge_size.csv}\comtable
+
+\begin{figure}[htpb]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis} [
+      ylabel=Tree size,
+      xlabel=Date,
+      grid=both
+      ]
+      \addplot[chartpurple] table[x index=1, y index=2]\comtable;
+    \end{axis}
+  \end{tikzpicture}
+
+  \caption{Merge-tree size by date}
+  \label{fig:}
+\end{figure}
 
@@
 From the list of trees consisting of a single node, we selected one at
-random using a simple python script and the \verb|random.choice()|
-function from python 3.6.1, which returned the merge tree rooted at
+random using the \verb|random.choice()| function in python 3.6.1, which
+returned the merge tree rooted at
 \emph{11df5864075f763ec0d1fdecd6a3f0af7d09a553}. As we are not testing
@@
 
-\subsubsection{Results}
-\label{ssub:results}
+The primary study questions are broken into two sections, one for
+tree comprehension, and another section for summarization. A final
+section is added, in order to have some information behind the
+demographic of the participant.
 
-\subsubsection{Discussion}
-\label{ssub:discussion}
+\textbf{Conceptual Tasks}
+
+As this information is provided directly by the \emph{Linvis} tree
+visualizations, we use Gitk or git commandline for performing these
+tasks.
+
+\evan{Draw both trees, or just one tree?}
+\begin{enumerate}
+  \item Draw a diagram of the merge tree showing how the commit
+    cdbdd1676a5379f1d5cbd4d476f5e349f445befe is merged into the master
+    branch of the repository.
+  \item How many individual commits are related to this commit?
+  \item How many merges are involved with merging these commits into the
+    master branch?
+\end{enumerate}
+
+We provide the user with 15 minutes to complete the first task, which is
+to draw the diagram shown in Figure~\ref{fig:commit_2}. The other parts
+of this task can be derived directly from the image drawn from the first
+task. For this reason, we keep the ordering of these tasks consistent
+between participants.
+
+\textbf{Summarization Tasks}
+
+We randomize the order of tool use, choosing to start with either
+\emph{Linvis} or Gitk to ensure that information cannot be carried
+forward from the previous tool across all participants.
+
+The questions in this portion of the study are shuffled and presented in
+random order, except where one task follows from another. We use the
+\verb|random.shuffle()| function from python 3.6.1 for performing the
+shuffling and presentation of the tasks.
+
+We asked the following questions in randomized order
+\begin{enumerate}
+  \item What other commits are merged in the same merge
+  \item What is the series of merges involved with merging this commit
+  \item \label{it:aut1} How many authors are involved with this merge
+  \item \label{it:aut2}Who contributed the most changes to this merge
+  \item \label{it:f1}How many files were modified in this merge?
+  \item \label{it:f2}Which files had the most changes in this merge?
+  \item Which modules does this merge tree involve?
+\end{enumerate}
+
+We identify items \ref{it:aut1} and \ref{it:aut2} to be related,
+concerning the authorship, and items \ref{it:f1} and \ref{it:f2} to be
+related, concerning the files of a merge tree. The related items will be
+in the same order consistently between participants, but the order of
+tasks will be changed between participants.
+
+\textbf{User Demographic}
+
+This section of the study provides us with some information about our
+participant, and how they felt about using gitk versus \emph{Linvis}.
+
+We asked the following questions in this order
+\begin{enumerate}
+  \item Given these tasks again, which tool would you prefer to use?
+  \item Which aspects of each tool did you like and why?
+  \item How long have you used git?
+  \item If you have used git, for what kind of projects? (personal,
+    school courses, professional?)
+  \item If you have used git, how many commits, files, and contributors
+    were involved with the largest repository you have worked with?
+\end{enumerate}
+
+These questions will provide us with some information about how the user
+felt, and their experience with git.
+
+\subsubsection{Study}
+\label{ssub:study}
+
+The null hypothesis is that using merge trees or using the DAG directly
+does not affect the performance of users.

### 2017-06-19 paper/lineval.tex
Format of the paper
Adding content for generalization and userstudy
This is just preliminary content for these sections, but it is heading
in the right direction.
Adding Tree size figure
Adds figure to paper, and to the makefile.
Preliminary future work section
Breaking User study section
Integrating the user study section of the paper into the rest of the
paper more cleanly.

The methodology section is nearly done, including information on user
selection, commit selection, and the questions for the study.

Updating the makefile to represent the changes.

@@
 \usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \usepackage{xcolor}
 \usepackage{color}
+\usepackage{listings}
 
@@
 
-\newcommand{\TheTitle}{An evaluation and generalization of Merge-Tree}
+\newcommand{\TheTitle}{An Evaluation and Generalization of Merge-Tree}
 \newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
@@
 \newcommand{\TheSubject}{Testing the comprehension of users in large git repositories}
-\newcommand{\TheKeywords}{Linux, git, data structures}
+\newcommand{\TheKeywords}{Linux, git, visualizations}
+
+\lstset{frame=tb,
+  language=python,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{chartblue},
+  commentstyle=\color{chartred},
+  stringstyle=\color{chartgreen},
+  breaklines=true,
+  breakatwhitespace=true,
+  tabsize=3
+}
 
@@
 
+\section{Introduction}
+\label{sec:introduction}
+
+\input{introduction.tex}
+
+\section{Related Works}
+\label{sec:related_works}
+
+\section{Methodology}
+\label{sec:methodology}
+
+\input{methodology.tex}
+
+\section{Generalization}
+\label{sec:generalization}
+
+\input{generalization.tex}
+
+\section{Results}
+\label{sec:results}
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Future Work}
+\label{sub:future_work}
+
+\input{futurework.tex}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\nocite{*}
+
+\bibliographystyle{IEEEtran}
+\bibliography{references}
+
 \end{document}

### 2017-06-19 paper/methodology.tex
Breaking User study section
Integrating the user study section of the paper into the rest of the
paper more cleanly.

The methodology section is nearly done, including information on user
selection, commit selection, and the questions for the study.

Updating the makefile to represent the changes.

% vim:set et sw=2 ts=4 tw=72:
% Jun 18, 2017

In order to determine the effectiveness of the merge-tree model on
comprehension and summarization of git repositories, we conducted a user
study on \evantodo{NUMBER} participants. The user study is a
mixed-methods study, but with more focused on the quantitative aspects.
We chose to use a mixed-methods approach to determine both if the model
is able to improve the time performance and accuracy of the
participants, and if the participants found the merge trees more
enjoyable for the given tasks.

In this section, we describe the describe the methods used in the user
study. This includes the participant criteria, the questions asked, the
method for selecting commits, and a breakdown of the methods used for
analyzing the results.

\subsection{User Selection}
\label{sub:user_selection}

Our target audience are git users, particularly users with who are
working on understanding what changes are being made, where they are
being made, and who is making the changes. The merge-tree model is also
intended for users who are trying to learn the structure of a
repository, how modules of the repository are broken into groups, if the

We have software engineering and computer science students at the
Masters, Doctorate, and Post-Doc levels available for participation in
this study. \evan{This feels like the weakest part of the paper, is
  there any way to make it sound stronger?}

\subsection{Commit Selection}
\label{sub:commit_selection}

We chose two commits for use in the study. The order that the commits
are presented is randomized between participants, but the order is kept
consistent through the tasks described in the next subsection.

We use the \tool database, containing commits and merge-tree information
from April 16th 2005 to October 14th, 2014, which corresponds to being
between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
  database even more?} We chose the commits based on tree sizes. We
found that a majority of the trees contain at most seven inner commits,
while more than 25\% of the trees contain a single commit. 75\% of the
tree contained up to 51 inner commits and merges, and finally the
largest tree contained 7217 nodes. We chose to work with the trees in
the first and second quartile, as merge trees of sizes between one and
seven, not including the merge into the master branch, make up the
majority of the trees in the database.

From here, we selected one tree from the trees of a single commit at
random. Selecting a commit from that tree is trivial, as there is only
one to choose.

We selected one tree of size seven at random from all trees of size
seven. We placed a restriction the trees, the selected tree must include
at least one internal merge to increase the complexity of the trees
tested. After randomly selecting a tree, we chose in internal commit at
random, using the \verb|random.choice| function in python 3.6.1.

% TODO: Show this in the introduction maybe

% It is worth noting that in the fourth
% quartile, the size of the tree drops off quickly from the top tree, with
% the next tree containing only 4708 nodes, and the third largest tree
% containing only 2349 commits.

Using this technique, we selected commit
\emph{8eb88c80d444fd249edaa7d895666cde79e7b3b8} from the tree containing
a single node (Figure~\ref{fig:commit_1}), and commit
\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from the tree
containing seven nodes (Figure~\ref{fig:commit_2}).

\begin{figure}[bpt]
  \centering
  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
  \caption{The first merge tree used in the user study, a merge tree
    containing a single commit}
  \label{fig:commit_1}
\end{figure}

\begin{figure}[bpt]
  \centering
  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
  \caption{The second merge tree used in the user study, a merge tree
    containing seven commits}
  \label{fig:commit_2}
\end{figure}

\subsection{Questions}
\label{sub:questions}

The questions and tasks are broken down into three groups. The first two
groups are purely quantitative in nature, being conceptual tasks, and
summarization tasks. The third group of questions is to determine the
demographic of our participant.

\textbf{Conceptual Tasks}

The participants were allowed to use Gitk or the git command line tools
to perform these tasks. These tasks are asked in this order, allowing
the user to use their diagram to answer the second two questions. It is
not necessary to use \tool in this portion of the study, as it provides
the answers to these questions directly.

\begin{enumerate}
  \item Draw a diagram of the merge tree, showing how the commit was
    merged into the master branch of the repository
  \item How many individual commits are related to this commit?
  \item How many merges are involved with merging this commit into the
    master branch?
\end{enumerate}

We provide the user with 10 minutes per commit to complete the first
task. The answers to the two other questions in this section are drawn
from the answer in the first task.

\textbf{Summarization Tasks}

For these questions, we randomize the order of the tools, choosing to
either start with \tool or Gitk, to ensure that we do not place a bias
on one tool or another through the experiment. The tasks are broken
into task sets related to the overall goal of the task. The task sets
are shuffled, and the tasks within the task sets are shuffled as well.
Finally, the order of the commits is shuffled. We use the
\verb|random.shuffle| function from python 3.6.1 to perform the
shuffling and presentation of the tasks.

\begin{itemize}
  \item Merge task set
    \begin{itemize}
      \item What is the series of merges involved with merging this
        commit?
      \item What other commits are merged in the same merge tree?
    \end{itemize}

  \item Authors task set
    \begin{itemize}
      \item How many authors are involved with this merge tree?
      \item Who contributed the most changes to this merge tree?
    \end{itemize}

  \item Files task set
    \begin{itemize}
      \item How many files were modified in this merge tree?
      \item Which file had the most changes in this merge tree?
    \end{itemize}

  \item Modules task set
    \begin{itemize}
      \item Which modules doe this merge tree involve?
    \end{itemize}
\end{itemize}

As understanding where the merge begins and ends, the merge task set
will always be the first task set presented; however, the questions will
still be re-ordered within the task set.

% TODO: provide the script for question selection
% \lstinputlistings{../selector.py}

\textbf{User Opinion}

Finally, we ask the user for their opinions on the tools and for some
information about their level of experience with git. This portion of
the study is qualitative and does not rely on either of the tool,
instead attempting to determine which model provided the user with a
better experience.

We ask the following questions in this order;

\begin{enumerate}
  \item Given these tasks again, which tool would you prefer to use?
  \item Which aspects of each tool did you like and why?
  \item How long have you used git?
  \item If you have used git, for what kind of projects? (personal,
    school courses, professional?)
  \item If you have used git, how many commits, files, and contributors
    were involved with the largest repository you have worked with?
\end{enumerate}

### 2017-06-28 paper/methodology.tex
Methodology participants and question answers
Adding the number of participants and the answers to the study
questions.

@@
 comprehension and summarization of git repositories, we conducted a user
-study on \evantodo{NUMBER} participants. The user study is a
+study on 12 participants. The user study is a
 mixed-methods study, but with more focused on the quantitative aspects.
@@
 % TODO: Show this in the introduction maybe
-
 % It is worth noting that in the fourth
@@
 
+% TODO: Fix this, the commit hash is wrong.
 Using this technique, we selected commit
-\emph{8eb88c80d444fd249edaa7d895666cde79e7b3b8} from the tree containing
-a single node (Figure~\ref{fig:commit_1}), and commit
-\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from the tree
-containing seven nodes (Figure~\ref{fig:commit_2}).
+\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
+a single node (Figure~\ref{fig:commit_1}), which we will refer to as
+\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
+the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
+will refer to as \comB.
 
@@
 \end{enumerate}
+
+\subsection{Data Extraction}
+\label{sub:data_extraction}
+
+The study is recorded using screen capture and audio capture software
+for further analysis. From the resulting videos, we are able to extract
+the timing and accuracy of each participant over each task. We record
+the timing, accuracy, task, commit, and tool used for each part. We also
+record some addition information about whether the participant was
+correct, guessing, partially correct, or completely wrong, as this may
+provide some interesting insights.
+
+The answers to each question is relatively straight-forward, as we can
+extract the information directly from gitk and Linvis. Measuring the
+accuracy of the questions is more challenging. In the question regarding
+the merges leading to the merging of a given commit, both the merges and
+the ordering of the merges matter.
+
+The correct answers for \comA are shown in
+Table~\ref{tab:coma_study_answers}, and the correct answers for \comB are
+shown in Table~\ref{tab:comb_study_answers}.
+
+\begin{table}[htpb]
+  \centering
+  \caption{Answers to the question from the conceptual and summarization portions of the user study for \comA}
+  \label{tab:coma_study_answers}
+  \begin{tabular}{l|l}
+    Question                                              & Answer \\\hline\hline
+    Number of commits in the merge tree                   & 1\\
+    Number of merges in the merge tree                    & 1\\\hline
+    Merges merging this commit into the master branch     & 1 (11df586407)\\
+    Other commits that are merged in the same tree        & None\\
+    How many authors are involved with this tree?         & 1\\
+    Who contributed the most changes to this merge tree   & David\\
+    How many files are modified in this merge tree        & 2\\
+    Which file(s) had the most changes in this merge tree & both (pzl.c, asl.c)\\
+    Which modules does this merge tree involve & wusb
+  \end{tabular}
+\end{table}
+
+\begin{table}[htpb]
+  \centering
+  \caption{Answers to the question from the conceptual and summarization portions of the user study for \comB}
+  \label{tab:comb_study_answers}
+  \begin{tabular}{l|l}
+    Question & Answer \\\hline\hline
+    Number of commits in the merge tree & 5 \\
+    Number of merges in the merge tree & 3\\\hline
+    Merges merging this commit into the master branch     & (219b22b245
+    \\
+    & $\rightarrow$ \\
+    & 8eb88c80d4)\\
+    Other commits that are merged in the same tree        & ec4e86ba06\\
+    & 9a3f371e99\\
+    & 1c85cc6445\\
+    & 7c2dfee848\\
+    How many authors are involved with this tree?         & 4\\
+    Who contributed the most changes to this merge tree   & Randy Dunlap\\
+    How many files are modified in this merge tree        & 6\\
+    Which file(s) had the most changes in this merge tree & pcm\_lib.c\\
+    Which modules does this merge tree involve            & ALSA\\
+  \end{tabular}
+\end{table}
+
+Extracting the timing information is prone to some issues with where the
+duration of answering should start and stop. We try to eliminate these
+issues by maintaining the following conventions; the answer duration
+begins at the end of the last word of the prompt. The duration ends
+immediately before the first word of the final modification to the
+answer. Many answers are able to be answered in a single word or phrase,
+but some may require multiple sentences that are broken by additional
+processing by the participant. An Example of questions where the answer
+may be interleaved with solving the task is the task asking the user to
+provide the merges on the path from the commit into the master branch.
+The goal of this is to eliminate the duration required to provide the
+prompt, and eliminate the time required to provide the answer.

### 2017-06-28 paper/results.tex
Adding results section

% vim:set et sw=2 ts=4 tw=72:
% Jun 24, 2017

In this section, we provide the raw aggregated results from the user
study and analysis on those results.

\subsection{Conceptual Questions Results}

% TODO: make plots showing distribution of times and responses

% Number of commits in the merge tree - Make a plot of this
% | text              | hash      | median |    mean | variance | Actual |
% |-------------------+-----------+--------+---------+----------+--------|
% | Number of commits | a3c1239eb |      4 | 19.1111 | 735.1111 |      1 |
% | Number of commits | cdbdd1676 |      4 |     7.8 | 136.8444 |      5 |
% | Number of merges  | a3c1239eb |      5 |  8.2727 |  53.6182 |      1 |
% | Number of merges  | cdbdd1676 |    3.5 |     5.4 |  50.2667 |      3 |
% Time taken to arrive at these results
% | text              | hash      | median(s) |  mean(s) | variance(s) |
% |-------------------+-----------+-----------+----------+-------------|
% | Number of commits | a3c1239eb |      10.0 |  49.9167 |   5952.0833 |
% | Number of merges  | a3c1239eb |       7.5 |  24.6667 |    884.4242 |
% | Number of commits | cdbdd1676 |      31.5 | 106.8333 |  54123.4242 |
% | Number of merges  | cdbdd1676 |      11.0 |     65.5 |  29798.8182 |

In Table~\ref{tab:conceptual_results}, we are able to see an overview of
the results from the conceptual questions.

\begin{table*}[htpb]
  \centering
  \caption{Results from the conceptual questions}
  \label{tab:conceptual_results}
  \begin{tabular}{ll|r|lrr|rrr}
    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
  \end{tabular}
\end{table*}

Users were able to more closely estimate the number of commits and
merges in the larger tree, but generally took longer than the smaller
tree. The tree with a single node resulted in more variability in the
estimate of number of commits .

It should be noted that these questions were answered after spending
roughly ten minutes attesting to draw a picture that held the answers to
these questions.

\subsection{Summarization Question Results}

### 2017-06-29 paper/methodology.tex
Methodology: Analysis Section
Adding some information on the analysis.

@@
 
-% TODO: Show this in the introduction maybe
+% TODO: Show this paragraph in the introduction
 % It is worth noting that in the fourth
@@
 
-% TODO: Fix this, the commit hash is wrong.
 Using this technique, we selected commit
@@
 the merges leading to the merging of a given commit, both the merges and
-the ordering of the merges matter.
+the ordering of the merges matter. In the cases where the answer is not
+a numerical value, we use the edit distance to determine the accuracy of
+an answer. An edit distance of zero indicates a correct answer, while
+any number other than zero indicates an incorrect answer. Adding,
+reordering (where important), and removing commits or merges all incur a
+penalty of 1 added to the edit distance.
 
@@
 prompt, and eliminate the time required to provide the answer.
+
+The information is then inserted into a sql database for further
+analysis.
+
+\subsection{Analysis}
+
+As the goals of each section are different, the techniques for analyzing
+the results are different. The conceptual questions are to demonstrate
+the difficulty in comprehending the DAG model, so we simply provide the
+metrics, showing how well the participants did in answering the
+questions asked. The summarization tasks are meant for comparing the
+ability of the participants to summarize two commits from the Linux
+repository. As this is a comparison, further analysis is necessary to
+determine the significance of our results. The final portion of the
+study is designed to help us understand the group of users that we are
+working with.
+
+As the goals of the conceptual questions are to demonstrate the
+difficulty in comprehending the DAG model, and determine how closely
+users are able to make the mapping from the DAG to showing how a commit
+was merged into the master branch, no further statistical analysis will
+be performed on this data. We will provide the metrics response,
+accuracy, and time metrics for these questions. No comparisons are made
+between the models in these tasks, so no further analysis is necessary.
+
+The summarization tasks are designed to compare the two models when the
+participant is summarizing various aspects of the repository, including
+the files modified and authorship information. We will use three terms
+to describe the performance of a user; correctness, which indicates
+simply whether the participant was correct or not, making no indication
+toward how far from the correct answer they were; accuracy, indicating
+how far from the correct answer the participant was; timing, which
+indicates how long a participant took to answer the question. We begin
+the analysis of these questions using the McNemar test, with $\alpha =
+0.005$, measuring the correctness of the results provided by the
+participant when using \tool versus gitk.
+% TODO: cite the McNemar test maybe?
+If the McNemar test indicates that we do not reject the null hypothesis,
+we continue analyzing the difference in accuracy and timing metrics for
+the merge-tree model versus the DAG.
+
+For each test, we use a confidence level $\alpha = 0.005$, resulting in
+a $\chi^2$ test statistic of 7.879.
+
+% TODO: Add the null hypothesis and type of metrics analyzed for each summarization question

### 2017-06-29 paper/methodology.tex
Methodlogy: Fixing sentence
sentence didn't make sense, but now it does

@@
 
-As understanding where the merge begins and ends, the merge task set
-will always be the first task set presented; however, the questions will
-still be re-ordered within the task set.
+As understanding where the merge begins and ends is vital to answering
+the other question, the merge task set will always be the first task set
+presented; however, the questions will still be re-ordered within the
+task set.
 

### 2017-06-29 paper/results.tex
Results: Adding plots to summarization questions

@@
 
-In this section, we provide the raw aggregated results from the user
-study and analysis on those results.
+In this section, we look at the aggreated results from the user study,
+and perform some analysis on those results. In the discussion section,
+we will further investigate potential reasons behind why various tests
+passed or failed.
 
@@
 \subsection{Summarization Question Results}
+
+We will look at the results from each question, and then look at the
+combined results to determine the correctness of the answers provided by
+the participants.
+
+\subsubsection{What is the series of merges involved with merging this commit}
+\label{ssub:what_is_the_series_of_merges_involved_with_merging_this_commit}
+
+	% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 8, df = 1, p-value = 0.004678
+
+   %         Linvis
+% Gitk        Correct Incorrect
+  % Correct         2         0
+  % Incorrect       8         2
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-3.pdf}
+  \caption{Correctness of participants using \tool versus gitk
+    when determining the series of merges involved with merging a given
+  commit}
+  \label{fig:figures/userstudy/correct_question-3}
+\end{figure}
+
+\subsubsection{What other commits are merged in this same merge tree}
+\label{ssub:what_other_commits_are_merged_in_this_same_merge_tree}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 7, df = 1, p-value = 0.008151
+
+   %         Linvis
+% Gitk        Correct Incorrect
+  % Correct         1         0
+  % Incorrect       7         4
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-4.pdf}
+  \caption{Correctness of participants using \tool versus gitk when
+    determining which other commits are merged into the master branch
+    with this commit}
+  \label{fig:figures/userstudy/correct_question-4}
+\end{figure}
+
+\subsubsection{How many authors are involved with this merge tree}
+\label{ssub:how_many_authors_are_involved_with_this_merge_tree}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 9, df = 1, p-value = 0.0027
+
+   %         Linvis
+% Gitk        Correct Incorrect
+  % Correct         1         0
+  % Incorrect       9         2
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-5.pdf}
+  \caption{Correctness of participants using \tool versus gitk when
+    determining the number of authors involved in the cluster of commits}
+  \label{fig:figures/userstudy/correct_question-5}
+\end{figure}
+
+\subsubsection{Who contributed the most changes to this merge tree}
+\label{ssub:who_contributed_the_most_changes_to_this_merge_tree}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 8, df = 1, p-value = 0.004678
+
+   %         Linvis
+% Gitk        Correct Incorrect
+  % Correct         2         0
+  % Incorrect       8         2
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-6.pdf}
+  \caption{Correctness of participants using \tool versus gitk when
+    determining who contributed the most changes to a merge tree}
+  \label{fig:figures/userstudy/correct_question-6}
+\end{figure}
+
+\subsubsection{How many files were modified in this merge tree}
+\label{ssub:how_many_files_were_modified_in_this_merge_tree}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 9, df = 1, p-value = 0.0027
+
+   %         Linvis
+% Gitk        Correct Incorrect
+  % Correct         1         0
+  % Incorrect       9         2
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-7.pdf}
+  \caption{Correctness of participants using \tool versus gitk when
+    determining how many files were modified in a merge tree}
+  \label{fig:figures/userstudy/correct_question-7}
+\end{figure}
+
+\subsubsection{Which file had the most changes in this merge tree}
+\label{ssub:which_file_had_the_most_changes_in_this_merge_tree}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 8, df = 1, p-value = 0.004678
+
+   %         Linvis
+% Gitk        Correct Incorrect
+  % Correct         1         0
+  % Incorrect       8         3
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-8.pdf}
+  \caption{Correctness of participants using \tool versus gitk when
+    determining which file(s) had the most changes a merge tree}
+  \label{fig:figures/userstudy/correct_question-8}
+\end{figure}
+
+\subsubsection{Which modules were involved with this merge tree}
+\label{ssub:which_modules_were_involved_with_this_merge_tree}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 1, df = 1, p-value = 0.3173
+
+%             Linvis
+% Gitk        Correct Incorrect
+  % Correct         8         1
+  % Incorrect       3         0
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-9.pdf}
+  \caption{Correctness of participants using \tool versus gitk when
+    determining which modules were involved with a merge tree}
+  \label{fig:figures/userstudy/correct_question-9}
+\end{figure}
+
+\subsubsection{Combined Results}
+\label{ssub:combined_results}
+
+% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 49.075, df = 1, p-value = 2.463e-12
+
+%             Linvis
+% Gitk        Correct Incorrect
+% Correct     16      1
+% Incorrect   52      15
+%
+
+For each of the studies, we also tracked the number of people who gave
+up on the question, stated that they made a random guess, or replied
+with infinity. In other plots, all of these responses are either placed
+in the correct or incorrect categories depending on the correctness of
+the answer. In Figure~\ref{fig:summarization_correctness}, we show the
+combined metrics across all questions, also showing the other
+categories.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/summarize_tool_outcomes.pdf}
+  \caption{Correctness of participants across all summarization
+    questions of participants using \tool versus gitk, including the
+    number of instances of giving up on a task, making a random guess,
+    or replying with infinity.}
+  \label{fig:summarization_correctness}
+\end{figure}

### 2017-06-29 paper/results.tex
Results: Conclusions
Fixes the tables, which were wrong. The data only included results from
a single commit instead of both. We include the data for the other
commit, and provide the conclusion for the McNemar test on each
summarization question.

@@
 
+With the McNemar's tests, we analyze the participants that change from
+being correct with one tool to being incorrect with the other. This does
+not take into account the number of people who were able to provide the
+correct answer for both tools, nor the people who were incorrect for both
+tools. The results are an aggregation over all participants and both
+commits analyzing the difference between \tool and gitk in the ability
+to assist user comprehension.
+
+All questions from the study, with the exception of the questions
+regarding which modules are changed in a merge tree, show conclude with
+rejecting the null hypothesis that there is no difference in correctness
+when using gitk versus \tool. Further explanations of the results are
+found in the following subsections.
+
 \subsubsection{What is the series of merges involved with merging this commit}
 \label{ssub:what_is_the_series_of_merges_involved_with_merging_this_commit}
-
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 8, df = 1, p-value = 0.004678
-
-   %         Linvis
-% Gitk        Correct Incorrect
-  % Correct         2         0
-  % Incorrect       8         2
 
@@
   commit}
-  \label{fig:figures/userstudy/correct_question-3}
+  \label{fig:q_3_correctness}
 \end{figure}
 
-\subsubsection{What other commits are merged in this same merge tree}
-\label{ssub:what_other_commits_are_merged_in_this_same_merge_tree}
+In Figure~\ref{fig:q_3_correctness}, we analyze the number of instances
+where people were correct and incorrect with \tool and with gitk, given
+the task of providing the series of merges involved with merging a given
+commit.
 
@@
 % data:  mat
-% McNemar's chi-squared = 7, df = 1, p-value = 0.008151
+% McNemar's chi-squared = 15, df = 1, p-value = 0.0001075
 
@@
 % Gitk        Correct Incorrect
-  % Correct         1         0
-  % Incorrect       7         4
+  % Correct         6         0
+  % Incorrect      15         2
+
+We further break the correctness information to show the instances where
+the participant was correct with both tools, incorrect with both tools,
+and went from being correct to incorrect, or incorrect to correct with a
+given tool for the purpose of applying the McNemar test.
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
+                           & Incorrect & 15                           & 2\\
+  \end{tabular}
+\end{center}
+
+We can see from these results that in a majority of the test instances,
+the participant provided an incorrect answer with Gitk, and a correct
+answer with \tool. The resulting $\chi^2$ value is 15. $15 > 7.879$,
+therefore we reject the null hypothesis that \tool does not make a
+difference in the correctness, with $99.5\%$ confidence.
+
+\subsubsection{What other commits are merged in this same merge tree}
+\label{ssub:what_other_commits_are_merged_in_this_same_merge_tree}
 
@@
     with this commit}
-  \label{fig:figures/userstudy/correct_question-4}
+  \label{fig:q_4_correctness}
 \end{figure}
 
-\subsubsection{How many authors are involved with this merge tree}
-\label{ssub:how_many_authors_are_involved_with_this_merge_tree}
+In Figure~\ref{fig:q_4_correctness}, we can see that there are more
+instances of correct answers from the participants when they use \tool
+than when they use Gitk, when determining other commits that are merged
+in the same merge tree.
+
+We further break the results, looking at the instances where
+participants provided correct answers using both tools, incorrect
+answers using both tools, or provided a correct and an incorrect answer.
 
@@
 % data:  mat
-% McNemar's chi-squared = 9, df = 1, p-value = 0.0027
+% McNemar's chi-squared = 14, df = 1, p-value = 0.0001828
 
@@
 % Gitk        Correct Incorrect
-  % Correct         1         0
-  % Incorrect       9         2
+  % Correct         5         0
+  % Incorrect      14         4
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
+                           & Incorrect & 14                           & 4\\
+  \end{tabular}
+\end{center}
+
+A majority of the test instances provided an incorrect response with
+Gitk, but corrected the response with \tool. Conversely, no incorrect
+responses with \tool were corrected with the use of Gitk. The resulting
+$\chi^2$ value is 14; $14 > 7.879$, therefore we reject the null
+hypothesis that \tool does not make a difference in the correctness,
+with $99.5\%$ confidence.
+
+\subsubsection{How many authors are involved with this merge tree}
+\label{ssub:how_many_authors_are_involved_with_this_merge_tree}
 
@@
   \caption{Correctness of participants using \tool versus gitk when
-    determining the number of authors involved in the cluster of commits}
-  \label{fig:figures/userstudy/correct_question-5}
+    determining the number of authors involved in a merge tree}
+  \label{fig:q_5_correctness}
 \end{figure}
 
-\subsubsection{Who contributed the most changes to this merge tree}
-\label{ssub:who_contributed_the_most_changes_to_this_merge_tree}
+Looking at Figure~\ref{fig:q_5_correctness}, we are able to see that
+there were more instances where participants provided the correct answer
+to the number of authors involved in a merge tree when they were using
+\tool versus gitk.
+
+Further breaking this information we get the following results;
 
@@
 % data:  mat
-% McNemar's chi-squared = 8, df = 1, p-value = 0.004678
+% McNemar's chi-squared = 16, df = 1, p-value = 6.334e-05
 
@@
 % Gitk        Correct Incorrect
-  % Correct         2         0
-  % Incorrect       8         2
+  % Correct         5         0
+  % Incorrect      16         2
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
+                           & Incorrect & 16                           & 2\\
+  \end{tabular}
+\end{center}
+
+The majority of the test instances provide an incorrect response with
+Gitk, but a correct response with \tool. No incorrect responses were
+reported using \tool that were corrected with Gitk. The resulting
+$\chi^2$ value is 16; $16 > 7.879$, therefore we reject the null
+hypothesis that \tool does not make a difference in the correctness of
+determining the number of authors involved in summarizing the number of
+authors in a merge tree.
+
+\subsubsection{Who contributed the most changes to this merge tree}
+\label{ssub:who_contributed_the_most_changes_to_this_merge_tree}
 
@@
     determining who contributed the most changes to a merge tree}
-  \label{fig:figures/userstudy/correct_question-6}
+  \label{fig:q_6_correctness}
 \end{figure}
+
+We are able to determine, in Figure~\ref{fig:q_6_correctness}, that
+there were more instances of correct answers using \tool versus gitk
+when answering who made the most changes in a merge tree.
+
+Further breakdown of this information yielded the following results;
+	% McNemar's Chi-squared test
+
+% data:  mat
+% McNemar's chi-squared = 13, df = 1, p-value = 0.0003115
+
+           % Linvis
+% Gitk        Correct Incorrect
+  % Correct         8         0
+  % Incorrect      13         2
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 8                            & 0\\
+                           & Incorrect & 13                           & 2\\
+  \end{tabular}
+\end{center}
+
+The majority of the test instances provide an incorrect response with
+Gitk, but a correct response with \tool. No incorrect responses were
+reported using \tool that were corrected with Gitk. The resulting
+$\chi^2$ value is 13; $13 > 7.879$, therefore we reject the null
+hypothesis that \tool does not make a difference in the correctness of
+determining the author that contributed the most changes to files in a
+merge tree.
 
@@
 \label{ssub:how_many_files_were_modified_in_this_merge_tree}
-
-% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 9, df = 1, p-value = 0.0027
-
-   %         Linvis
-% Gitk        Correct Incorrect
-  % Correct         1         0
-  % Incorrect       9         2
 
@@
     determining how many files were modified in a merge tree}
-  \label{fig:figures/userstudy/correct_question-7}
+  \label{fig:q_7_correctness}
 \end{figure}
 
-\subsubsection{Which file had the most changes in this merge tree}
-\label{ssub:which_file_had_the_most_changes_in_this_merge_tree}
+In Figure~\ref{fig:q_7_correctness}, we see that there were more
+instance of correct answers using \tool versus gitk when determining how
+many files were modified in a merge tree.
 
@@
 % data:  mat
-% McNemar's chi-squared = 8, df = 1, p-value = 0.004678
+% McNemar's chi-squared = 15, df = 1, p-value = 0.0001075
 
@@
 % Gitk        Correct Incorrect
-  % Correct         1         0
-  % Incorrect       8         3
+  % Correct         6         0
+  % Incorrect      15         2
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
+                           & Incorrect & 15                           & 2\\
+  \end{tabular}
+\end{center}
+
+The majority of the test instances provide an incorrect response with
+Gitk, but a correct response with \tool. No incorrect responses were
+reported using \tool that were corrected with Gitk. The resulting
+$\chi^2$ value is 15; $15 > 7.879$, therefore we reject the null
+hypothesis that \tool does not make a difference in the correctness of
+in determining the number of files modified in a merge tree.
+
+\subsubsection{Which file had the most changes in this merge tree}
+\label{ssub:which_file_had_the_most_changes_in_this_merge_tree}
 
@@
     determining which file(s) had the most changes a merge tree}
-  \label{fig:figures/userstudy/correct_question-8}
+  \label{fig:q_8_correctness}
 \end{figure}
 
-\subsubsection{Which modules were involved with this merge tree}
-\label{ssub:which_modules_were_involved_with_this_merge_tree}
+In Figure~\ref{fig:q_8_correctness}, we see that there were more
+instances of correct answers using \tool versus gitk when determining
+which file had the most changes.
 
@@
 % data:  mat
-% McNemar's chi-squared = 1, df = 1, p-value = 0.3173
+% McNemar's chi-squared = 14, df = 1, p-value = 0.0001828
 
@@
 % Gitk        Correct Incorrect
-  % Correct         8         1
-  % Incorrect       3         0
+  % Correct         6         0
+  % Incorrect      14         3
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
+                           & Incorrect & 14                           & 3\\
+  \end{tabular}
+\end{center}
+
+The resulting $\chi^2$ test value is 14; $14 > 7.879$, therefore we
+reject the null hypothesis that \tool does not make a different in the
+correctness of determining which files has the most changes in a merge
+tree.
+
+\subsubsection{Which modules were involved with this merge tree}
+\label{ssub:which_modules_were_involved_with_this_merge_tree}
 
@@
     determining which modules were involved with a merge tree}
-  \label{fig:figures/userstudy/correct_question-9}
+  \label{fig:q_9_correctness}
 \end{figure}
 
-\subsubsection{Combined Results}
-\label{ssub:combined_results}
+In Figure~\ref{fig:q_9_correctness}, we see that while the number of
+instances where the use of \tool resulted in correct results was higher
+than the number of instances where the use of gitk resulted in correct
+results, this is the only test where Gitk had more instances of a
+correct result than instances of an incorrect result.
 
@@
 % data:  mat
-% McNemar's chi-squared = 49.075, df = 1, p-value = 2.463e-12
+% McNemar's chi-squared = 5.4444, df = 1, p-value = 0.01963
 
@@
 % Gitk        Correct Incorrect
-% Correct     16      1
-% Incorrect   52      15
-%
+  % Correct        14         1
+  % Incorrect       8         0
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                            &           & \multicolumn{2}{c}{Linvis}\\
+                            &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
+                            & Incorrect & 8                            & 0\\
+  \end{tabular}
+\end{center}
+
+The resulting $\chi^2$ test value is $5.44$; $5.44 < 7.879$, therefore
+we do not reject the null hypothesis, suggesting that \tool does not
+make a difference in determining which modules are involved with a merge
+tree.
+
+% TODO: perform further analysis on this, looking at distance from the
+% correct answer, and duration of time taken to answer
+
+\subsubsection{Combined Results}
+\label{ssub:combined_results}
 
@@
 \end{figure}
+
+We further break the results into the number of instances where Gitk and
+and \tool resulted in the correct result, where Gitk resulted in a
+correct result while \tool resulted in an incorrect result, the
+instances where \tool resulted in a correct result while Gitk resulted
+in an incorrect result, and the instances where both resulted in an
+incorrect result. This breakdown yields;
+
+\begin{center}
+  \begin{tabular}{cc|rr}
+                          &           & \multicolumn{2}{c}{Linvis}\\
+                          &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk} & Correct   & 50                           & 1\\
+                          & Incorrect & 95                           & 15\\
+  \end{tabular}
+\end{center}
+
+The resulting $\chi^2$ statistic is $92.042$; $92.042 > 7.879$,
+therefore we should reject the null hypothesis with $99.5\%$ confidence.

### 2017-06-30 paper/discussion.tex
Adding some threats to validity

% vim:set et sw=2 ts=4 tw=72:
% Jun 24, 2017
\subsection{Threats to Validity}
\label{sub:threats_to_validity}

While git comes bundled with gitk and the \verb|git log --graph|
command, many of the users in our study were unfamiliar with the DAG
shown by these tools. A potential reason for this may be that unless a
repository reaches sufficient complexity, both in branching and number
of commits, these tools are not necessary for the normal functionality
of git. However, while they were unfamiliar with either merge trees or
the DAG, these users were still able to better understand and summarize
the structure of the repository using merge trees.

Each question was evaluated independently of the results from the
previous question in the summarization tasks. Had the correct answers
been adjusted to match the participants' understanding of which commits
were involved in a merge tree, the results of the summarization would
have likely been very different. Another approach would to have provided
the participants with the correct tree once they had completed the
conceptual tasks, and the merge task set in the summarization task set.

\subsection{Future Work}
\label{sub:future_work}

% Inter Navigation
% TODO: verify this with the study

While the merge-tree model provides users with a better conceptual image
of what is occuring within a single merge tree, we have not exprimented
or tested the potential gains of navigation between trees. At this time,
the Linvis search provides results grouped by merge-tree root instead of
as a single block of results. The results are generally similar, with
each merge tree being merged at a different point in time, or to a
different version of the kernel. A future extension of this may
incorporate some information about how a given cluster changes over
time. To do this, a technique would need to be developed to cluster
merge-trees into modules, and then ordering the merge-trees based on
time or release.

% TODO: ensure that linvis has been referenced/defined

Perform a user study over multiple repositories using different merge
structures to see how quickly new members of a team can understand the
merge procedure. % Yeah, this is a big one, actually.

### 2017-06-30 paper/results.tex
moving figures
Moving correctness figures to different directory

@@
 rejecting the null hypothesis that there is no difference in correctness
-when using gitk versus \tool. Further explanations of the results are
-found in the following subsections.
+when using gitk versus \tool. We do not see a change in these
+conclusions if we include continuity correction. Further explanations of
+the results are found in the following subsections.
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-3.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-3.pdf}
   \caption{Correctness of participants using \tool versus gitk
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-4.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-4.pdf}
   \caption{Correctness of participants using \tool versus gitk when
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-5.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-5.pdf}
   \caption{Correctness of participants using \tool versus gitk when
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-6.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-6.pdf}
   \caption{Correctness of participants using \tool versus gitk when
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-7.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-7.pdf}
   \caption{Correctness of participants using \tool versus gitk when
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-8.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-8.pdf}
   \caption{Correctness of participants using \tool versus gitk when
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/correct_question-9.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/question-9.pdf}
   \caption{Correctness of participants using \tool versus gitk when
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/summarize_tool_outcomes.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/correctness/summarize_tool_outcomes.pdf}
   \caption{Correctness of participants across all summarization

### 2017-07-02 paper/results.tex
Results: boxplots for accuracy and time information
Adds boxplots for showing the difference between gitk and Linvis in
timing and accuracy metrics.

@@
 when using gitk versus \tool. We do not see a change in these
-conclusions if we include continuity correction. Further explanations of
-the results are found in the following subsections.
+conclusions if we include continuity correction in the McNemar test.
+Further explanations of the correctness results are found in the
+following subsections.
+
+In Table~\ref{tab:summarization_timing}, we see the timing metrics taken
+across all users, and both commits. We then further break the results by
+separating the commits by which commit tree it came from in
+Table~\ref{tab:summarization_timing_cid}, but do not see tremendous
+differences between the commits.
+
+\begin{table}[htpb]
+  \centering
+  \caption{Summarization task timing in seconds, aggregated over all study participants and both commits}
+  \label{tab:summarization_timing}
+  \begin{tabular}{r|c|rrr}
+    Question                             & Tool   & Median(s) & Mean(s) & Variance(s) \\\hline
+    \multirow{2}{*}{Series of Merges}    & Linvis & 8         & 18.1    & 617.5\\
+                                         & Gitk   & 37        & 64.4    & 4349.8\\
+    \multirow{2}{*}{Other Commits}       & Linvis & 14        & 19.8    & 349.5\\
+                                         & Gitk   & 36        & 44.5    & 1471.3\\
+    \multirow{2}{*}{Number of Authors}   & Linvis & 1         & 5.7     & 163.3\\
+                                         & Gitk   & 11        & 41.5    & 5838.1\\
+    \multirow{2}{*}{Biggest Contributor} & Linvis & 3         & 9.4     & 293.0\\
+                                         & Gitk   & 11        & 40.9    & 3423.7\\
+    \multirow{2}{*}{Number of Files}     & Linvis & 1         & 3.1     & 32.4\\
+                                         & Gitk   & 15        & 34.9    & 4143.6\\
+    \multirow{2}{*}{Most changed File}   & Linvis & 4         & 15.0    & 959.4\\
+                                         & Gitk   & 14        & 48.4    & 3862.0\\
+    \multirow{2}{*}{Involved Modules}    & Linvis & 4         & 5.6     & 19.4\\
+                                         & Gitk   & 34        & 45.0    & 1703.4
+  \end{tabular}
+\end{table}
+
+\begin{table*}[htpb]
+  \centering
+  \caption{Summarization task timings broken down by commit, aggregated over all study participants}
+  \label{tab:summarization_timing_cid}
+  \begin{tabular}{l|l|rrrr}
+    Question                            & Commit & Linvis Median(s) & Gitk Median(s) & Linvis Variance(s) & Gitk Variance(s)\\\hline
+    \multirow{2}{*}{Series of Merges}   & \comA  & 8.0              & 37.0           & 437.8              & 3144.5\\
+                                        & \comB  & 9.0              & 48.5           & 802.8              & 5707.5\\
+    \multirow{2}{*}{Other Commits}      & \comA  & 10.0             & 36.0           & 419.9              & 1513.6\\
+                                        & \comB  & 14.5             & 31.0           & 316.3              & 1563.3\\
+    \multirow{2}{*}{Number of Authors}  & \comA  & 2.0              & 11.0           & 23.2               & 1668.5\\
+                                        & \comB  & 1.0              & 12.0           & 293.0              & 9789.8\\
+    \multirow{2}{*}{Biggest Contributor}& \comA  & 2.0              & 6.0            & 1.9                & 2853.2\\
+                                        & \comB  & 6.5              & 18.0           & 494.4              & 4154.6\\
+    \multirow{2}{*}{Number of Files}    & \comA  & 1.0              & 19.0           & 31.0               & 489.7\\
+                                        & \comB  & 1.5              & 8.5            & 36.4               & 7588.3\\
+    \multirow{2}{*}{Most changed File}  & \comA  & 3.0              & 21.0           & 5.9                & 5343.4\\
+                                        & \comB  & 4.0              & 10.0           & 1652.0             & 2696.0\\
+    \multirow{2}{*}{Involved Modules}   & \comA  & 3.0              & 19.0           & 29.5               & 1946.2\\
+                                        & \comB  & 4.5              & 39.5           & 12.1               & 1622.8\\
+  \end{tabular}
+\end{table*}
+
+% We attempt to fit a normal distribution to both the timing and accuracy
+% information, but there are not enough samples to accurately assess the
+% parameterization of this data; we can see a strong shifted normal
+% distribution forming for some questions
+% (Figure~\ref{fig:q8_accuracy_density}), but others do not express this
+% behaviour (Figure~\ref{fig:q3_accuracy_density}).
+
+% \begin{figure}[htpb]
+%   \centering
+%   \includegraphics[page=2, width=0.8\linewidth]{figures/userstudy/accuracy/q8.pdf}
+%   \caption{Density function compared to the fitted normal distribution,
+%     showing the difference in correctness when using Gitk versus \tool
+%     counting the number of files involved.
+%     Negative values imply that participants were more accurate with
+%     Gitk, positive values imply better accuracy with \tool.
+%     }
+%   \label{fig:q8_accuracy_density}
+% \end{figure}
+
+% \begin{figure}[htpb]
+%   \centering
+%   \includegraphics[page=2, width=0.8\linewidth]{figures/userstudy/accuracy/q3.pdf}
+%   \caption{Comparing the density function, plotting the difference in
+%     accuracy between responses provided by Gitk and Linvis. Responses
+%     are answering the series of merges involved with merging a commit.}
+%   \label{fig:q3_accuracy_density}
+% \end{figure}
 
@@
 commit.
-
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 15, df = 1, p-value = 0.0001075
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct         6         0
-  % Incorrect      15         2
 
@@
 difference in the correctness, with $99.5\%$ confidence.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/accuracy/q3.pdf}
+  \caption{Accuracy comparison between Gitk and \tool, for determining
+    the series of merges into the master branch of the repository.}
+  \label{fig:q3_accuracy}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q3.pdf}
+  \caption{Time comparison between Gitk and \tool, for determining
+    the series of merges into the master branch of the repository.}
+  \label{fig:q3_timing}
+\end{figure}
 
@@
 
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 14, df = 1, p-value = 0.0001828
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct         5         0
-  % Incorrect      14         4
-
 \begin{center}
@@
 with $99.5\%$ confidence.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/accuracy/q4.pdf}
+  \caption{Accuracy comparison between Gitk and \tool, for determining
+    other commits in the same merge tree.}
+  \label{fig:q4_accuracy}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q4.pdf}
+  \caption{Time comparison between Gitk and \tool, for determining
+    other commits in the same merge tree.}
+  \label{fig:q4_timing}
+\end{figure}
 
@@
 
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 16, df = 1, p-value = 6.334e-05
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct         5         0
-  % Incorrect      16         2
-
 \begin{center}
@@
 
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/accuracy/q5.pdf}
+  \caption{Accuracy comparison between Gitk and \tool, for answering how
+  many authors were involved with a merge tree.}
+  \label{fig:q5_accuracy}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q5.pdf}
+  \caption{Time comparison between Gitk and \tool, for answering how
+  many authors were involved with a merge tree.}
+  \label{fig:q5_timing}
+\end{figure}
+
 \subsubsection{Who contributed the most changes to this merge tree}
@@
 Further breakdown of this information yielded the following results;
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 13, df = 1, p-value = 0.0003115
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct         8         0
-  % Incorrect      13         2
 
@@
 
+There are no accuracy metrics for this question, as the answer is either
+correct or incorrect.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q6.pdf}
+  \caption{Timing comparison between Gitk and \tool, for answering who
+    contributed the most changes.}
+  \label{fig:q6_timing}
+\end{figure}
+
 \subsubsection{How many files were modified in this merge tree}
@@
 many files were modified in a merge tree.
-
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 15, df = 1, p-value = 0.0001075
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct         6         0
-  % Incorrect      15         2
 
@@
 
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/accuracy/q7.pdf}
+  \caption{Accuracy comparison between Gitk and \tool, for answering how
+  many files were involved with the merge tree.}
+  \label{fig:q7_accuracy}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q7.pdf}
+  \caption{Timing comparison between Gitk and \tool, for answering how
+    many files were involved with the merge tree.}
+  \label{fig:q7_timing}
+\end{figure}
+
 \subsubsection{Which file had the most changes in this merge tree}
@@
 
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 14, df = 1, p-value = 0.0001828
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct         6         0
-  % Incorrect      14         3
-
 \begin{center}
@@
 tree.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/accuracy/q8.pdf}
+  \caption{Accuracy comparison between Gitk and \tool, for answering
+    which file had the most changes.}
+  \label{fig:q8_accuracy}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q8.pdf}
+  \caption{Time comparison between Gitk and \tool, for answering which
+    file had the most changes.}
+  \label{fig:q8_timing}
+\end{figure}
 
@@
 
-	% McNemar's Chi-squared test
-
-% data:  mat
-% McNemar's chi-squared = 5.4444, df = 1, p-value = 0.01963
-
-           % Linvis
-% Gitk        Correct Incorrect
-  % Correct        14         1
-  % Incorrect       8         0
-
 \begin{center}
@@
 
-% TODO: perform further analysis on this, looking at distance from the
-% correct answer, and duration of time taken to answer
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/accuracy/q9.pdf}
+  \caption{Accuracy comparison between Gitk and \tool, for answering
+    which modules were involved with a merge.}
+  \label{fig:q9_accuracy}
+\end{figure}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/timing/q9.pdf}
+  \caption{Time comparison between Gitk and \tool, for answering which
+    modules were involved with a merge.}
+  \label{fig:q9_timing}
+\end{figure}
+
+% Figure~\ref{fig:q9_accuracy}
 

### 2017-07-03 paper/results.tex
Results: Description of accuracy and timing

@@
 
+There is almost no variation in accuracy between participants using
+\tool, while there is more variance in Gitk. There were two outliers
+among the instances using \tool, which lie within the second and third
+quartiles for the accuracy measure of Gitk. The furthest outlier for
+\tool lies at a distance that is slightly greater than the median for
+Gitk. We see these results in Figure~\ref{fig:q3_accuracy}.
+
+We similar results for the timing metric in Figure~\ref{fig:q3_timing}.
+All significant data occurs within less time than the second, third, and
+fourth quartile of the Gitk results. Nearly all outlier instances in
+\tool are within the third quartile group of the instances of Gitk with
+the exception of the most extreme, which lies in the fourth quartile
+group.
+
 \begin{figure}[htpb]
@@
 \end{figure}
+
+Overall, we see that \tool enables users to correctly identify the
+series of merges required to merge a given commit more quickly than with
+Gitk. Furthermore, the accuracy of incorrect results are closer to being
+correct than the incorrect answers for Gitk.
 
@@
 
+In Figure~\ref{fig:q4_accuracy}, we can see that there is again, very
+little variance in the accuracy of results for \tool compared to the
+accuracy of the results in Gitk.
+
+In Figure~\ref{fig:q4_timing} there is less variance in the time taken
+to come to an answer with \tool than with Gitk. Furthermore, we see that
+a majority of the participants were able to complete the task in \tool
+in much less time than 50\% of the participants using Gitk.
+
 \begin{figure}[htpb]
@@
 \end{figure}
+
+Overall, participants are able to determine other commits that are
+related to a given commit more quickly and more accurately with \tool
+than with Gitk.
 
@@
   \caption{Accuracy comparison between Gitk and \tool, for answering how
-  many authors were involved with a merge tree.}
+  many authors are involved with a merge tree.}
   \label{fig:q5_accuracy}
@@
   \caption{Time comparison between Gitk and \tool, for answering how
-  many authors were involved with a merge tree.}
+  many authors are involved with a merge tree.}
   \label{fig:q5_timing}
 \end{figure}
+
+Again, we see very similar results in Figure~\ref{fig:q5_accuracy} and
+Figure~\ref{fig:q5_timing}, to the results found in the previous two
+tasks. The variance in accuracy and time taken is much greater with
+Gitk than with \tool. This suggests that the participants were able to
+determine the number of authors more quickly and with better accuracy
+with \tool than Gitk.
 
@@
 
-\subsubsection{How many files were modified in this merge tree}
+We similar results in Figure~\ref{fig:q6_timing} as with the previous
+questions. There is greater variance in the time taken for participants
+to determine who contributed the most changes to a group of commits
+using Gitk than with \tool.
+
+\subsubsection{How many files are modified in this merge tree}
 \label{ssub:how_many_files_were_modified_in_this_merge_tree}
@@
 \end{figure}
+
+Figure~\ref{fig:q7_accuracy} and Figure~\ref{fig:q7_timing} show similar
+results, showing that the accuracy and time for the participants to
+produce a response was more variable with Gitk than with \tool.
 
@@
   \caption{Accuracy comparison between Gitk and \tool, for answering
-    which file had the most changes.}
+    which file(s) had the most changes.}
   \label{fig:q8_accuracy}
 \end{figure}
+
+It was possible for this question to have multiple answers. If there
+were multiple files with the same number of modifications, in lines of
+code changed, then the correct response would be a list of all files
+with those changes. Ultimately, we see a fairly clean distribution of
+accuracies using Gitk in Figure~\ref{fig:q8_accuracy}, with the median
+being one file away from the correct answer. Aside from two outliers,
+all responses from the participants using \tool were correct.
+
+We see the same behaviour in Figure~\ref{fig:q8_timing}, showing that
+there is more variance in the time taken to respond to the question when
+the participant is using Gitk versus when they are using \tool.
 
@@
   \caption{Time comparison between Gitk and \tool, for answering which
-    file had the most changes.}
+    file(s) had the most changes.}
   \label{fig:q8_timing}
 \end{figure}
+
+Overall, we see an improvement in correctness, accuracy, and time taken
+to determine which file or files had the most changes in the group of
+commits when the participants use \tool over Gitk.
 
@@
 
-% Figure~\ref{fig:q9_accuracy}
+We don't see a tremendous difference in the accuracy performance of the
+participants between Gitk and \tool in Figure~\ref{fig:q9_accuracy}.
+This is consistent with the results for the correctness metric. Unlike
+accuracy and correctness, we see a noticeable difference in the time
+performance of the participants between \tool and Gitk
+(Figure~\ref{fig:q9_timing}).
 

### 2017-07-03 paper/results.tex
Results: Starting user opinion

@@
 tree. The tree with a single node resulted in more variability in the
-estimate of number of commits .
+estimate of number of commits.
 
@@
 therefore we should reject the null hypothesis with $99.5\%$ confidence.
+
+\subsection{User Opinions}
+\label{sub:user_opinions}
+
+There was nearly unanimous preference for \tool to the representation
+provided by Gitk, citing the ability to abstract information about the
+repository from the simple summarization tables and clean visualization.
+Three of the twelve participants suggested that with better
+understanding of the Gitk interface and of the DAG, that an expert may
+be able to extract the summarization information with relative ease. Of
+these three, one participant said that they would prefer to use both
+tools side-by side. Further discussion on this is available in section
+... %TODO Provide section.

### 2017-07-05 paper/introduction.tex
Introduction: Linvis Introduction
Replacing introduction with the introduction from original paper.

--- 2017-06-11 paper/introduction.tex
@@
-Merge-trees, as presented by Wilde and German\cite{Wilde2016}, are a data structure
-designed to make the navigation and summarization of git merges simpler than using
-the directed acyclic graph (DAG), as done in gitk among other tools\evan{does this
-  need a citation?}. Git uses the DAG model internally, making it simple to
-construct tools directly on the DAG;\@ however, this model does not provide a clear
-summarization of the events occuring in the repository. Gitk and the git command
-line interface (cli) both provide the user with a visualization of the DAG of the
-repository.
+% vim:set et sts=2 sw=2 ts=2 tw=72:
+
+Git is a distributed version control system that is used in many
+opensource and closedsource projects. Git enables many developers to
+write code simultaneously, working on various components known as
+branches. Git uses a directed acyclic graph model (DAG) for storing the
+necessary information regarding the state of the code at any point in
+time. Git refers to nodes of the DAG as comits, not distinguishing
+between the code-carrying commits and the structural commits that bring
+two branches together. In this paper, we will refer to the individual
+commits as being commits, the structural merge commits as merges, and
+refereing to all as repository events.
+
+Between 50k and 70k commits are added to the Linux kernel per version
+requiring maintainers of older versions of the kernel to sift through
+thousands of commits and merges with tools that are unable to filter and
+effectively visualize projects at the scale of the kernel. Older
+versions of the kernel are used in embedded systems and mobile phones;
+for security purposes, performance needs, and changing hardware
+requirements, maintainers must be able to understand the changes being
+made in the current version of the kernel in order to produce the
+necessary patches for the older versions of the kernel. Current tools
+and visualizations are based on the directed acyclic graph (DAG) of the
+repository, showing all commits and merges in chronological order by
+when the commit was authored, not by when it arrived in the official
+Linux repository.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/gitk.png}
+  \caption{
+    The Gitk interface, centered at commit
+    cdbdd1676a5379f1d5cbd4d476f5e349f445befe. The top left pane showing
+    the DAG, the top center pane showing the authors, and top right pane
+    showing the commit dates. The bottom-leftpane shows the full
+    commit message and diff of the selected commit, and the bottom-right
+    pane shows the list of the files modified by the selected commit.
+  }
+  \label{fig:gitk}
+  %\vspace{-4mm}
+\end{figure}
+
+The DAG representation works in smaller projects; it enables users to
+see when changes are made, when these changes are merged, how each
+branch is interacting, and the point where a branch forks from the
+master branch. In large modular projects, like the Linux kernel, the DAG
+becomes a mess of merges and commits (Figure~\ref{fig:gitk}) losing its
+visual meaning. In some cases, the repository is simply too large for
+systems to generate a visualization; repositories stored on GitHub will
+have a DAG visualzation to show the branches and forks of a
+repositority, but is unable to display the visualization for projects at
+the scale of Linux or even the Git Git
+repository(Figure~\ref{fig:gitfail}).  Between 60k and 70k new commits
+are created for the Linux project every year; according to previous
+work\cite{German2015}, a commit takes a median of 30 days from the time
+it is authored until it arrives in the official repository. The snapshot
+of the kernel tomorrow may be different than the snapshot from today,
+containing new commits authored in the past; distinguishing these new
+commits from the commits in the snapshot from today is not trivial.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
+  \caption{Github is unable to display the DAG of the Linux
+    Project due to the size and complexity, which would normally
+    show the relationship to other branches and forks.}
+  \label{fig:gitfail}
+  %\vspace{-2mm}
+\end{figure}
+
+This paper investigates the construction, implementation, user-based
+evaluation, and generalization of a new model, the \mt model. We
+begin with a description and algorithm for converting the DAG into a set
+of merge trees. We continue by providing some details about the
+web-based tool, called \tool\footnote{\tool is currently available at
+  \url{http://li.turingmachine.org}}, implemented to test the
+feasibility and potential capabilities of the \mt model. We
+continue with a user study, to evaluate first, if the DAG is capable of
+providing the users with a conceptual understanding of the repository
+events that are related to a given commit, and second, determine if
+there is a difference between the DAG and \mt model in performing
+summarization tasks.
+
+\subsection{Directed Acyclic Graph Model}
+\label{sec:directed_acyclic_graph_model}
+
+Git uses the DAG as the internal model for representing repositories.
+This model functions perfectly, and is necessary, for the internal
+functionality of Git. The DAG model allows for the reconstruction of
+files and the working directory by recursively applying the changes, or
+the diff, made in each repository event that is a parent of the current
+event. We see an example DAG repository Figure~\ref{fig:repoDAG}.
+
+Each repository event contains an ordered list of one or more parents,
+except the initial commit, which has no parents. Commits have a single
+parent, while merges will have at least two parents, the current branch,
+and the branch being merged into the current branch. Git allows for the
+merging of multiple branches simultaneously in octopus merges, where the
+order of the parents represents the order that the branches were
+specified for being merged.
+
+The repository events in the DAG are immutable; once an event is
+created, it can't be changed. Git allows operations to alter commits and
+re-order them, but this will create a new commit with a new commit hash.
+This property makes a repository event unable to record the traversal of
+merges into the master branch of the repository. A commit cannot have
+references in the direction of the merge that merges the commit, as this
+reference will change, requiring the commit to be changed. While this
+provides some challenge, git eliminates this issue of the DAG with the
+\verb|git log --children| command, which traverses the DAG and inverts
+the edges to provide the reference from parent to child. This
+terminology will prove to be confusing through the rest of the paper,
+when we are referring to the \mt form, the parent is the next node
+toward the root. When referring to the DAG, the parent relationship is
+in the opposite direction, from the root toward the branch-point.
+
+Many current visualization systems base the visualization on the DAG,
+because it is made readily available by git. The primary issue being
+that the DAG was not designed with visualization and summarization tasks
+in mind. Due to the direction of the links, the DAG model is unable to
+provide a summarization of the contents of a merge; the DAG is unable to
+show which branch is the master branch, and is unable to filter the
+irrelevant information from the view.
+
+\subsection{\mt Model}
+\label{sub:merge_tree_model}
+
+We use the \mt model to address these issues in the DAG\@. We
+will describe the \mt model in detail in
+Section~\ref{sec:mergetree}. The \mt reverses the direction of
+the references of the DAG.\@ Starting with each merge into the master
+branch, the algorithm recursively aggregates the statistics of each
+commit. The tree is rooted at the merge into the master branch, and the
+commits are the leaves.
+
+\subsection{Implementation}
+\label{sub:implementation}
+
+We implement a tool, \tool, to test and verify the capabilities of the
+\mt model. \tool provides search, summarization, and
+visualization using the \mt model.
+
+We provide users with search-based navigation. The search engine takes a
+plain text query from the user and returns results based on similarity
+to the query. The engine weighs various fields, including the commit
+log, the author name, the commit hash, and the dates that the commit was
+authored and committed. The results are aggregated by the \mt
+with a link to the root at the top, and the repository events that match
+the search query in a table below.  An example of a single \mt is shown
+in Figure~\ref{fig:linvis_search_results}.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=1.0\linewidth]{figures/linvis/search_results.png}
+  \caption{\tool search results for `net-next 2011', showing a single
+    \mt group.}
+  \label{fig:linvis_search_results}
+\end{figure}
+
+\tool uses seven tabs to show the information and visualizations for a
+selected commit or merge. The message tab shows the full commit log
+message. This does not include the diff. The files tab shows an
+aggregated table of all files that were modified in a merge. It includes
+metrics like the number of lines added, lines removed, total lines
+modified, and the delta, summed across all commits in the merge tree
+that modify this file. A small details drop-down button allows a user to
+see exactly which commit makes the changes, as shown in
+Figure~\ref{fig:linvis_files}. If the current repository event being
+viewed is a commit, the aggregate views will only show modifications
+made by the commit. The modules tab shows the modules modified in the
+merge tree. Like the files tab, the modules tab uses a table to show the
+name of the module, the number of commits that are in the merge tree
+that work with the module, and a details button to provide the links to
+those commits.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_files.png}
+  \caption{Table showing the modified files in a merge, with the second
+  entry expanded to show the commit that makes the changes.}
+  \label{fig:linvis_files}
+\end{figure}
+
+The authorship tab is very similar to the files tab, but shows the
+authorship information. It shows the sum of the number of lines added,
+removed, modified, and the delta within the \mt. It also shows
+the number of files that were modified by the author. The details tab is
+organized slightly differently. Instead of organizing these results by
+commit, the details are organized by file, showing which files were
+modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_authors.png}
+  \caption{Table showing the authors who made changes in a merge. The
+    entry for Randy Dunlap is expanded, showing the modifications to
+    each file that were made by Randy.}
+  \label{fig:linvis_authors}
+\end{figure}
+
+\tool implements three visualizations, a list tree, pack tree, and
+Reingold-Tilford tree. The list tree is a series of indented textual
+lists that can be easily searched with the built-in search in most
+web-browsers. The list tree is rooted at the current event, if the event
+is a commit, only that commit will be shown; conversely, if the selected
+event is the root, the entire merge tree will be shown. The Pack-Tree or
+Bubble tree, uses a set-based approach for showing the parent
+relationship. The root node aggregates all of the events within the \mt
+into the master branch. Thus, the root is the superset of all the other
+nodes in the tree. The commits are shown as individual white circles,
+containing no other events, as these are the leaves. This is shown in
+Figure~\ref{fig:linvis_pack}. The selected event will be shown in bright
+orange, commits in white, and merges in shades of blue. Darker shades of
+blue indicate that the merge is deeper in the tree.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
+  \caption{The pack tree visualization for a selected commit, shown in
+    orange.}
+  \label{fig:linvis_pack}
+\end{figure}
+
+The Reingold-Tilford tree is the classic tree visualization, with the
+root at the top, leaves at the bottom, and the edges between them
+showing the parent relationship. An example of the Reingold-Tilford tree
+visualization is shown in Figure~\ref{fig:commit_1} and
+Figure~\ref{fig:commit_2}.
+
+\section{\mt Model}
+\label{sec:mergetree}
+
+The \mt model inverts the DAG, so that the reference point from
+parent to child instead of child to parent. This is to enable
+aggregation at merges. Given a merge, we are able to recursively
+traverse all children of that merge, collecting the statistics at each
+commit. The inverted DAG is then traversed starting at each merge into
+the master branch. The merges into the master branch is the root of each
+tree. The commits are the leaves of the tree. Starting at the merge into
+the master branch, we recursively traverse each child of the current
+node until we reach the branch point from the master branch, or a commit
+that has a possible root merge that is closer in commit time to the
+commit. As a consequence, commits will always take the shortest path, in
+terms of time, to the master branch.
+
+A short example: assume the commits represented in
+Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
+In this case, commits are performed in various repositories and
+branches. The DAG representation of the commits is shown in
+Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
+the master branch and the repository that the master branch is part of.
+The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
+Note that the direction of the edges of the DAG have been inverted,
+instead of pointing from the child to the ancestors, it points from the
+ancestor to its successors, forming a path to the master branch. Also
+note that the DAG has been simplified, showing only a single edge on the
+path to master for any commit.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
+    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
+
+    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
+    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
+    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
+    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
+
+        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
+    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
+  \end{tikzpicture}
+  \caption{Example of a sequence of events performed in different
+    repositories. The horizontal axis represents time. Each horizontal
+    section represents a different branch and/or repository. Each commit
+    points to its ancestor.}
+  \label{fig:repoEvents}
+%\vspace{-3mm}
+\end{figure}
+
+% TODO: make everything one color
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
+  \end{tikzpicture}
+  \caption{DAG representation of the commits represented in
+    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
+    information about which repository a commit comes from. Furthermore,
+    the DAG does not distinguish the master branch from other branches.}
+  \label{fig:repoDAG}
+\vspace{-3mm}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+
+    \draw[chartblue]
+      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (2) edge[chartyellow, -stealth] (5)
+          (3) edge[chartmagenta, -stealth](5)
+          (5) edge[chartyellow, -stealth] (7)
+          (7) edge[chartyellow, -stealth] (9)
+          (6) edge[chartred, -stealth] (9)
+          (8) edge[chartmagenta, -stealth] (9)
+          (9) edge[chartyellow, -stealth] (11)
+          (10) edge[chartyellow, -stealth] (12);
+
+  \end{tikzpicture}
+  \caption{\mt view of the commits  represented in
+    Figure~\ref{fig:repoEvents} showing the path they followed to reach
+    the master branch. In this model the successors of each commit
+    represents the path followed by that commit to reach the master
+    branch.}
+  \label{fig:repoTree}
+\vspace{-3mm}
+\end{figure}
+
+\subsection{Computing the \mt from the DAG of Linux}
+
+Computing the \mt from a DAG for any repository may not be
+possible; however, certain features of the development process of Linux
+make it feasible to compute the \mt for the Linux repository.
+First, the master branch of Linux is maintained by Linus Torvalds, and
+only Linus has write access to it. We have verified this assertion in
+previous research~\cite{German2015}. We have developed a heuristic that
+is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
+identifies the commits made directly to the master branch; whereafter it
+recursively determines the shortest path (in terms of time), using the
+DAG, from each commit to the master branch using the inverted DAG.
+
+\begin{algorithm}
+        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
+        \begin{algorithmic}
+                \Function{ComputeMergeTree}{DAG}: tree
+                \State {\# Compute the tree from the DAG of Linux repository.}
+                \State {\# Returns $Tree$, a graph containing every commit }
+                \State {\# in DAG with the path it followed to master.}
+
+                \State $head \gets \textit{Head of master of git repository}$
+                \State $master \gets \textit{traverse DAG from head using }$
+                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
+                \State $nodes(Tree) \gets nodes(DAG)$
+                \State \Function{distance2Master}{cid} : seconds
+                \State {\# Helper function}
+                \State {\# Recursively compute shortest distance to master}
+                \State {\# setting cid's successor (next) in its way to master.}
+                \State {\# This function should be memoized. Otherwise it}
+                \State {\# would run in exponential time.}
+                \If {\textit{cid in master}}
+                \State \Return 0
+                \EndIf
+                \State    $d \gets 	\infty$
+                \State {\# Traverse the inverted DAG}
+                \For{$c \in children(cid, DAG)$}
+                \If {$c \in master$}
+                \State $d_1 \gets commitTime(c)-commitTime(cid)$
+                \Else
+                \State {$d_1 \gets distance2Master(c)$}
+                \EndIf
+                \If {$d_1 < d $}
+                \State $next \gets c$
+                \State  $d \gets d_1$
+                \EndIf
+                \EndFor
+                \State {\# $c$ is the commit that follows $cid$}
+                \State {\# in its way to master}
+                \State add edge $(cid, next)$ to $Tree$
+                \State \Return $d$
+                \EndFunction
+
+                \State {\# Compute the distance for each commit}
+                \State {\# discarding result}
+                \For{$c \in nodes(DAG)$}
+                \State $distance2Master(c)$
+                \EndFor
+                \State \Return $Tree$
+                \EndFunction
+        \end{algorithmic}
+\end{algorithm}
+
+\subsection{Evaluation}
+
+Merges that do not have merge conflicts provide information to verify
+this heuristic. If a merge does not contain a conflict, it records a
+summary of the commits that it merges. See Figure~\ref{fig:sampleMerge}
+for an example. This summary contains a list of the first 20 non-merge
+commits in the merge, including their one-line log description, the full
+logs of the merge commits that merge this subset, and the total number
+of non-merge commits in the merge.
+
+\begin{figure}[htbp]
+  \centering
+  {\fontsize{7}{9}
+    \begin{verbatim}
+Merge: 8cbd84f fd8aa2c
+Author: Linus Torvalds <torvalds@linux-foundation.org>
+Date:   Tue Aug 10 15:38:19 2010 -0700
+
+Merge branch 'for-linus' of git://neil.brown.name/md
+
+* 'for-linus' of git://neil.brown.name/md: (24 commits)
+md: clean up do_md_stop
+[... edited for the sake of space]
+md: split out md_rdev_init
+md: be more careful setting MD_CHANGE_CLEAN
+md/raid5: ensure we create a unique name for kmem_cache...
+...
+    \end{verbatim}}\vspace{-5mm}
+  \caption{Example of how merges record a subset of commits being
+    merged. The commit only shows the first 20 one-line summaries
+    messages for the 24 non-merge commits it merged. The ending
+    ``\ldots'' is part of the log and represents that other
+    commits were merged.}
+  \label{fig:sampleMerge}
+\end{figure}
+
+We used this information to evaluate the accuracy of the \mt
+model extracted from the DAG\@. The method we followed started with the
+extraction of the commit history up to July 20, 2016. We computed the
+merge tree of every commit until then. Since Linus Torvalds mostly does
+merging directly into master, we assumed that every merge by him is the
+root of a \mt. As described above, the log of a merge-commit
+usually contains the number of commits in the merge the first 20
+summaries of commits being merged. We extracted merges by Linus Torvalds
+using the command \mycode{log --merges --author='Torvalds} and compared
+the number of commits according to the log with the number of commits in
+the  \mt rooted in this commit. We also used the summaries of the
+commits found in the merge (not necessarily all---see above) to make
+sure those commits were in their corresponding merge tree. For example,
+for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
+merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
+one-line summaries corresponds to commits in that \mt. We also
+inspected those with differences to make sure they were true errors.
+The results can be summarized as follows:
+
+\begin{itemize}
+
+  \item
+
+    Five merges were false-errors because their logs did not contain
+    accurate information (were probably edited by hand). For example in
+    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
+    empty), in \mycode {c55d267\ldots} the summaries were reordered.
+
+  \item
+
+    The heuristic correctly identified that 79 of Linus merges (between
+    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
+    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
+    commits.
+
+  \item
+
+    The heuristic worked perfectly until Sept 4, 2007, the earliest date
+    that it could be verified.  Before this date, and until Dec 12,
+    2006, merges did not include a summary of the commits they included,
+    hence making it impossible to verify; during this period, however,
+    we correctly identified the merges by Linus into master.
+
+  \item
+
+    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
+    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
+    which confounded the true master branch of a repository (see
+    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
+    for a description of the issue).
+
+\end{itemize}
+
+In summary, of the merges after Sept. 4, 2007, our heuristic was correct
+in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
+12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
+2007).

### 2017-07-05 paper/lineval.tex
Adding results section
Adding hyperref package
Renaming futurework.tex to discussion.tex
It's the whole discussion, not just futurework subsection now.
Related Works: Adding some basic related works

@@
 %% Table Support
-
 \usepackage{dcolumn}
@@
 
+%% Algorithms
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+
 %% Graphics
-
 \usepackage{tikz}
@@
 
-\newcommand{\TheTitle}{An Evaluation and Generalization of Merge-Tree}
+\newcommand{\TheTitle}{Merge-Tree: Visualizing the Integration of Commits into Linux}
 \newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
 \newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Testing the comprehension of users in large git repositories}
+\newcommand{\TheSubject}{Understanding a large number of commits and merges}
 \newcommand{\TheKeywords}{Linux, git, visualizations}
+
+%% Referencing
+\usepackage[unicode=true,
+          pdftitle={\TheTitle},
+          pdfauthor={\TheAuthors},
+          colorlinks=false]{hyperref}
 
@@
 \begin{abstract}
+  \input{abstract.tex}
+\end{abstract}
 
-  \input{abstract.tex}
-
-\end{abstract}
+\begin{IEEEkeywords}
+  \TheKeywords
+\end{IEEEkeywords}
 
@@
 
+\input{relatedwork.tex}
+
 \section{Methodology}
@@
 \input{methodology.tex}
+
+\section{Results}
+\label{sec:results}
+
+\input{results.tex}
 
@@
 
-\section{Results}
-\label{sec:results}
-
 \section{Discussion}
@@
 
-\subsection{Future Work}
-\label{sub:future_work}
-
-\input{futurework.tex}
+\input{discussion.tex}
 

### 2017-07-05 paper/relatedwork.tex
Related Works: Adding some basic related works

--- 2016-09-18 linvis.tex
@@
-% vim: set fenc=utf-8 ft=latex encoding=utf-8
-% -*- mode: latex; coding: UTF-8; -*-
+% vim:set et sw=2 ts=4 tw=72:
+% Jul 04, 2017
+% Evan Wilde
 
-\newif\ifdraft
-% \drafttrue
+version control systems monitor the development lifetime of software
+projects. This makes the version control system vital in providing
+information about how a software project is being developed. To do this,
+the user must be able to gain a clear understanding of how commits are
+being integrated into the project, along with understanding what changes
+these commits are bringing with them. These are the tasks that we set
+out to solve with the \mt model. Ideally, we could compare our model
+against another that is designed with these goals. To our knowledge, we
+know of no git repository visualzation tool that has these specific
+goals. This may be due to issues with finding the master branch of the
+repository which is a non-trivial task, or due to other factors.
 
-\ifdraft
-        \documentclass[conference, draftclsnofoot, draft]{IEEEtran}
-        \def\baselinestretch{1}
-        \setlength{\marginparwidth}{2cm}
-\else
-        \documentclass[conference]{IEEEtran}
-\fi
+While we have found no tool with the express goal of showing how a
+commit is integrated, and providing a summarization of a merge, there
+has been a lot of work done in providing visualizations of various
+aspects of a repository.
 
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
+Many tools work to address the issues in communication between
+developers in inter-team collaboration work. Codebook\cite{Begel2010}
+uses a data mining technique to determine the developer of a piece of
+code, the program manager who wrote the specification for the code, and
+the program managers and developers on the team who were working
+together. Hipikat\cite{Cubranic2005} is another tool with a focus on
+communication. Where Codebook focuses on developers working on a
+project, Hipikat is focused on enabling easier integration of new
+developers to a project by providing them with easily-searchable
+artifacts of the changes made. Codebook is useful for pairing a
+contributor with the original developer; however, the developer may not
+have worked with the piece of code in years. A Hipikat program may
+provide more information to the maintainer as it records the artifacts
+of why certain design decisions were made when they were made using
+other tools like Bugzilla and CVS.\@ Neither tool is sufficient in
+meeting our goals to provide a summary of the topology of the kernel
+repository through a visual tree.
 
-\newcommand{\TheTitle}{Merge-tree: Visualizing the Integration of Commits into Linux}
-\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
-\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Digesting large amounts of commit data}
-\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}
-
-\synctex=1
-
-\usepackage[hyphens]{url}
-\urlstyle{same}
-
-\ifdraft
-        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
-                pdfborder={0 0 0},backref=none,colorlinks=true]{hyperref}
-
-\else
-        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
-                pdfborder={0 0 0},backref=none,colorlinks=false]{hyperref}
-\fi
-
-\usepackage[nospace]{cite}
-
-% Table Support
-\usepackage{dcolumn}
-\usepackage{longtable}
-
-\usepackage{balance}
-\usepackage{placeins}
-\usepackage{multirow}
-
-% Extra support
-\usepackage{xspace}
-\usepackage{caption}
-\usepackage[nospace]{cite}
-
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{tikz}
-\usepackage{xcolor}
-\usepackage{color}
-
-% Fix any bad-hyphenations here
-\hyphenation{contained}
-
-% Tikz libraries
-\usetikzlibrary{positioning}
-
-\input{colors.tex}
-\input{macros.tex}
-
-\begin{document}
-
-\title{\TheTitle}
-\author{
-        \IEEEauthorblockA{\TheAuthors}
-        \IEEEauthorblockN{Department of Computer Science,
-                University of Victoria, Canada.}
-        \IEEEauthorblockA{Email: \TheEmails}
-}
-\maketitle
-\begin{abstract}
-
-With an average of more than 900 top-level merges into the Linux kernel per
-release, many containing hundreds of commits and some containing thousands,
-maintenance of older versions of the kernel becomes nearly impossible.
-Various commercial products, such as the Android platform, run older
-versions of the kernel. Due to security, performance, and changing hardware
-needs, maintainers must understand what changes (commits) are added to the
-current version of the kernel since the last time they inspected it in order
-to make the necessary patches.
-
-Current tools provide information about repositories through the directed
-acyclic graph (DAG) of the repository, which is helpful for smaller
-projects. However, with the scale and number of branches in the kernel the
-DAG becomes overwhelming very quickly. Furthermore, the DAG contains every
-ancestor of every commit, while maintainers are more interested in how and
-when a commit arrives to the official Linux repository.
-
-In this paper, we propose the merge-tree, a simplified transformation
-of the DAG of the Linux git repository that shows the way in which commits
-are merged into the master branch of Linux. Using the merge-tree, we
-build \tool, a tool that is designed to allow users to explore how commits
-are merged into the Linux kernel.
-
-\end{abstract}
-
-\begin{IEEEkeywords}
-        \TheKeywords
-\end{IEEEkeywords}
-
-\section{Introduction}
-
-Between 50k and 70k commits are added to the Linux kernel per version requiring
-maintainers of older versions of the kernel to sift through thousands of
-commits and merges with tools that are unable to filter and effectively visualize
-projects at the scale of the kernel. Older versions of the kernel are used in
-embedded systems and mobile phones; for security purposes, performance needs, and
-changing hardware requirements, maintainers must be able to understand the changes
-being made in the current version of the kernel in order to produce the necessary
-patches for the older versions of the kernel. Tools like Gitk use a directed acyclic
-graph (DAG) model of the repository, showing all commits and merges in chronological
-order by when the commit was authored, not by when it arrived in the official Linux
-repository.
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
-        \caption{A view of the Gitk interface centered on merge
-          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
-          their ancestors and there is no clear path from the commit to the merge
-          with the master branch. Neither Gitk nor Git are capable of showing the
-          commits in master.}
-        \label{fig:gitk}
-%\vspace{-4mm}
-\end{figure*}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its
-                relationship to other forks.}
-        \label{fig:gitfail}
-%\vspace{-2mm}
-\end{figure}
-
-This representation works in smaller projects; it enables users to see when changes
-are made, when these changes are merged, how each branch is interacting, and the
-point where a branch forks from the master branch. In large modular projects, like
-the Linux kernel, the DAG becomes a mess of merges and commits
-(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel
-is simply too large for the system to generate a visualization; Github provides a
-DAG view for many projects, but is unable to display the visualization for projects
-at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and
-70k new commits are created for the Linux project every year; according to our
-previous work\cite{German2015}, a commit takes a median of 30 days from the time it
-is authored until it arrives in the official repository. The snapshot of the kernel
-tomorrow may be different than the snapshot from today, containing new commits
-authored in the past; distinguishing these new commits from the commits in the
-snapshot from today is not trivial.
-
-One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch, including
-the master branch. To complicate the problem, the DAG only has references to the
-ancestors of a commit (a model necessary for the operation of Git), but maintainers
-would prefer knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would imply that
-for any given merge, it would be possible to know which commits were merged. A user
-could inspect the commits that arrived into the master branch within a given
-time-frame by checking which commits were merged during that time-frame.
-
-This paper makes two contributions; first, we describe a method of converting the
-DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach the master branch; second, we
-present a method to inspect and visualize the history of merges in the Linux project
-using the merge-tree model.
-
-These methods and visualizations are implemented in a web-based tool called
-\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
-Our visualizations and tool provide information about the location of any given
-commit or merge in its respective merge-tree, the files edited, the modules edited,
-and the commit message. \tool allows users to apply various filters, including the
-release version, along with a keyword or phrase from the log preview, the name of
-the author, or the commit ID. The user can request all merges made by Linus that
-contain a commit or inner merge that matches the search query, or just the commits
-and merges that match the query.
-
-\section{Merge-Tree model}
-\label{sec:mergetree}
-
-Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
-commit has one or more ordered parents (ancestors), except the root commits of a
-repository which do not have ancestors--Linux's git repository has two such commits.
-Commits are divided into merge commits, merging two or more branches, and non-merge
-commits. Non-merge commits have only one parent, while merge commits have two or
-more. The order of the parents matter: the first parent is the branch in which the
-merge is being done, while the rest indicate the branches being merged (Git allows
-merging multiple branches simultaneously). The second parent is the first branch
-merged, the third parent is the second branched merged, etc.
-
-Once a commit is created, it is never changed. Git allows operations to alter
-commits or reorder them, but it changes the commit ID in the process, effectively
-replacing it with a new commit. This makes commits unable to record the traversal of merges
-from the commit to the merge into the master branch.
-
-A short example: assume the commits represented in Figure~\ref{fig:repoEvents} show
-the sequence of events in a repository. In this case, commits are performed in
-various repositories and branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about the master
-branch and the repository that the master branch is part of. The merge-tree view of
-this DAG is visible in Figure~\ref{fig:repoTree}. Note that the direction of the
-edges of the DAG have been inverted, instead of pointing from the child to the
-ancestors, it points from the ancestor to its successors, forming a path to the
-master branch. Also note that the DAG has been simplified, showing only a single
-edge on the path to master for any commit.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
-    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-
-    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
-    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
-    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
-    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
-
-        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
-    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
-  \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
-  \label{fig:repoEvents}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG loses information about which
-    repository the commit is performed in and through which merges it
-    has passed on its way to the master branch. The DAG does not even
-    distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (2) edge[chartyellow, -stealth] (5)
-          (3) edge[chartmagenta, -stealth](5)
-          (5) edge[chartyellow, -stealth] (7)
-          (7) edge[chartyellow, -stealth] (9)
-          (6) edge[chartred, -stealth] (9)
-          (8) edge[chartmagenta, -stealth] (9)
-          (9) edge[chartyellow, -stealth] (11)
-          (10) edge[chartyellow, -stealth] (12);
-
-  \end{tikzpicture}
-  \caption{Merge-tree view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
-\vspace{-3mm}
-\end{figure}
-
-\subsection{Computing the merge-tree of the DAG of Linux}
-
-Computing the merge-tree from a DAG for any repository may not be possible; however,
-certain features of the development process of Linux make it feasible to compute the
-merge-tree for the Linux repository. First, the master branch of Linux is maintained
-by Linus Torvalds, and only Linus has write access to it. We have verified this
-assertion in previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first identifies
-the commits made directly to the master branch; whereafter it recursively determines
-the shortest path (in terms of time), using the DAG, from each commit to the master
-branch using the inverted DAG.
-
-\begin{algorithm}
-        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
-        \begin{algorithmic}
-                \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
-                \State $head \gets \textit{Head of master of git repository}$
-                \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
-                \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
-                \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
-                \If {\textit{cid in master}}
-                \State \Return 0
-                \EndIf
-                \State    $d \gets 	\infty$
-                \State {\# Traverse the inverted DAG}
-                \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
-                \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
-                \EndIf
-                \EndFor
-                \State {\# $c$ is the commit that follows $cid$}
-                \State {\# in its way to master}
-                \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
-                \EndFunction
-
-                \State {\# Compute the distance for each commit}
-                \State {\# discarding result}
-                \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
-                \EndFor
-                \State \Return $Tree$
-                \EndFunction
-        \end{algorithmic}
-\end{algorithm}
-
-\subsection{Evaluation}
-
-Merges that do not have conflicts provide information to verify this heuristic. If a
-merge does not contain a conflict, it records a summary of the commits that it
-merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a
-list of the first 20 non-merge commits in the merge, including their one-line log
-description, the full logs of the merge commits that merge this subset, and the
-total number of non-merge commits in the merge.
-
-\begin{figure}[htbp]
-        \centering
-        {\fontsize{7}{9}
-        \begin{verbatim}
-Merge: 8cbd84f fd8aa2c
-Author: Linus Torvalds <torvalds@linux-foundation.org>
-Date:   Tue Aug 10 15:38:19 2010 -0700
-
-Merge branch 'for-linus' of git://neil.brown.name/md
-
-* 'for-linus' of git://neil.brown.name/md: (24 commits)
-md: clean up do_md_stop
-[... edited for the sake of space]
-md: split out md_rdev_init
-md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache...
-...
-        \end{verbatim}}\vspace{-5mm}
-        \caption{Example of how merges record a subset of commits being merged. The
-                commit only shows the first 20 one-line summaries messages for the 24
-                non-merge commits it merged. The ending ``...'' is part of the log
-                and represents that other commits were merged.}
-        \label{fig:sampleMerge}
-\end{figure}
-
-We used this information to evaluate the accuracy of the merge-tree model extracted from the DAG. The method we followed
-started with the extraction of the commit history up to July 20, 2016. We computed the merge tree of every commit until
-then. Since Linus Torvalds mostly does merging directly into master, we assumed that every merge by him is the root of a
-merge-tree. As described above, the log of a merge-commit usually contains the number of commits in the merge the first 20
-summaries of commits being merged. We extracted merges by Linus Torvalds using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits according to the log with the number of commits in the  merge-tree rooted in this commit. We
-also used the summaries of the commits found in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the merge in
-Figure~\ref{fig:sampleMerge} we would expect
-that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line summaries corresponds to commits in
-that merge-tree. We also inspected those with differences to make sure they were true errors.
-The results can be summarized as follows:
-
-\begin{itemize}
-\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
-  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
-  {c55d267} the summaries were reordered.
-\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
-  branch (not master). This branch was merged at (\mycode{3f17ea6d..} which contained 6809 commits.
-\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
-  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
-  impossible to verify; during this period, however, we correctly identified the merges by Linus into master.
-\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
-repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
-\end{itemize}
-
-In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860) our heuristic was correct. It failed in 1,542
-commits before Dec. 12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
-
-\section{Visualizing the merge-tree of Linux}
-
-The goal of \tool is to simplify the navigation of the kernel commit information,
-specifically focusing on merges. This is done by leveraging the merge-tree to
-inspect how commits are merged on the path to the master repository.
-
-\subsection{Use cases}
-
-We designed \tool with two use-cases in mind, though a user may switch between the
-cases as they work.
-
-\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\ These
-are users that are maintaining a section of the kernel and would like to pick a
-merge (including all the commits that it merges) and merge it directly into their
-current repository. This is useful for reducing the amount of re-implementation
-work. For these users, it is important to have the ability to aggregate metadata
-about files and modules being effected by the merge. Also, it is important for these
-users to be able to navigate from the root of the merge-tree toward the leaves.
-
-\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
-These are users that start with a known merge or commit and would like to see what
-other changes are being made in commits that are in the same merge, including
-knowing the merge-tree they belong to. This is useful to see what other commits are
-related to the current commit and how they get collated into merges that eventually
-end in the master branch. This is primarily for maintainers that need to perform
-some specific cherry picking of commits. We must provide these users a mechanism for
-navigating from a single commit toward the master branch, allowing them to see other
-commits that might be related to their original commit.
-
-\subsection{Data Model}
-
-In our visualizations, we leverage the merge-tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already in the
-master branch, or is part of a tree which is rooted in the merge that merged it into
-the master branch.  Each commit, whether a merge or non-merge, has only one
-successor; the root of each tree has none as it was made by Linus Torvalds directly
-into the master branch. Non-merge commits contain the metadata for the changes made.
-This metadata includes the files changed, the lines added and removed from each
-file, the author, the date the commit was merged into the merge that led to being
-merged into the kernel, the date the commit was authored, the patch, and the commit
-log. Merges contain less metadata, only storing the author of the merge, the log,
-the commit date, and the author date, and potentially, changes necessary to address
-conflicts during the merge. The details of the model are outlined in
-\cite{German2015}.
-
-\section{Design and Implementation}
-
-To navigate and inspect the merge-tree view of the kernel we created a web-based
-tool called \tool. Creating a web-based tool enables users to use the system without
-having to install additional software or store a large database, making it more
-accessible, more easily maintainable, and platform independent. \tool uses the
-following mechanisms to reach our goals of better navigation and better explanation
-of the selected changes.
-
-\begin{itemize}
-        \item Filter by searching
-        \item View files edited
-        \item View modules
-        \item Tree viewer
-\end{itemize}
-
-\subsection{Searching}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View allows filtering commits that were merged in a given
-                period, filtering by author, keyword, or commit ID.}
-        \label{fig:search}
-\end{figure}
-
-Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
-merges that are irrelevant. The search mechanism breaks down the results by release
-version. A user can further narrow down the search by specifying a range of dates in
-which such commits were merged by Linus into the master branch---not when the commits
-were created (author date and/or commit date). This distinction is important. We have observed commits that have
-taken years to arrive into the master repository after they were originally
-created.
-
-A user may then provide a search text, filtering by the author name, the commit ID,
-or keyword from the log. Any part of the author name may show up in the results,
-including searching by email address.
-
-If the user is searching by commit ID, the ID can be specified by using any of its unique prefixes. For example, the commit
-\mycode{3f17ea6dea8ba5668873afa54628a91aaa3fb1c0} is returned when the user searches for a
-commit ID of \mycode{3f17e} in the 3.16 Linux kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
-        \caption{Search Results. Each table entry is a commit merged in the desired
-                merged window.}
-        \label{fig:results}
-\end{figure}
-
-In the search results (seen in Figure~\ref{fig:results}), the user is presented with the one-line log message preview,
-the author's name and email, the date the commit was authored, and the date the commit was last committed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
-        \caption{Panel showing the Commit Message of a commit.}
-        \label{fig:message}
-\end{figure}
-
-Once a user selects a commit or merge to investigate, they are presented with a
-tabbed pane allowing them to view the full commit log, the files edited, the modules
-involved, and the merge-tree view.
-
-The first tab displays the full commit log (Figure~\ref{fig:message}). From this, a
-user is able to see what they would see had they searched for the commit using Git
-log. This doesn't provide additional information to the other tools, but helps to
-complete the functionality of \tool. The commit log provides a user with the
-information about the content of the commit and who has signed-off on the commit to
-ensure that it is of good quality. The message for merges may contain a summary of
-the commits being merged.  The information within these messages is highly variable,
-and is completely dependent on the author's style. As the user moves toward the
-root-level merge, the quality of these messages generally improves.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
-        \caption{Panel showing the files modified by all the commits that are part
-                of this merge.}
-        \label{fig:files}
-\end{figure}
-
-The second tab is the files tab (Figure~\ref{fig:files}). This tab provides
-information on what files have been edited, how many lines were added, and how many
-lines were removed in a given commit. For non-merges, this functionality is similar
-to the other tools available. Our tree-based design model allows us to extend this
-functionality to merges by aggregating information about all the commits that are
-children of the merge in the merge-tree, which other tools are unable to show. To
-find the number of lines added to a file in a merge, we take the sum of the lines
-added to that file in each of the children of that merge. We do the same for
-calculating the number of lines removed.
-
-% This could be extended to use the patch information to determine when a line has been
-% replaced rather than incrementing the lines added and removed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
-        \caption{Panel showing the modules changed by all the commits in this merge.}
-        \label{fig:modules}
-\end{figure}
-
-The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
-within the commit. Modules are not natively recognized by Git, and are not going to
-be present in all repositories. In the Linux repository, authors put the module they
-are working on in the one-line summary of the log-message; for example: the log
-\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
-coverage testing tool of the kernel). We heuristically extract the module by taking
-all text in the log summary of commits until the first colon.  Modules are logical
-partitions of the information in the kernel. Depending on where the author was
-working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
-quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
-few cases, the author of a commit does not correctly follow this format and the
-heuristic approach fails. As with the Files panel, non-merge commits show their
-corresponding information, but for merge commits we aggregate all the modules
-changed in all the commits that are part of that merge.  The output of this view is
-shown in Figure~\ref{fig:modules}.
-
-Finally, we have the Tree view tab. The tree view is designed for providing easy
-navigation of the commits within the merge-tree that is rooted in the current merge.
-It also provides a clear topological view of the merge and the submerges it
-includes. We have experimented with various tree designs to find a design that
-allows for easy navigation and visualization of both large and small trees. We
-discuss this panel in the next subsection.
-
-\subsection{Merge-Tree Views} \label{treeview_section}
-
-The merge-tree view is what makes \tool unique to other tools that inspect the DAG
-of a Git repository. With it, a user can inspect how commits are merged on their way
-to the master branch. We have experimented with various types of trees:
-
-\begin{enumerate}
-        \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
-        \item Reingold-Tilford trees provide a clear visual representation of the tree structure of the merge-tree.
-        \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
-                clearly showing the parent-child relationships between commits and merges.
-\end{enumerate}
-
-\subsubsection{List Tree}
-
-The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of nested lists,
-and is designed to more closely model the tree view found in file browsers. This
-tree only contains the commits and merges that are within the subtrees of the
-current merge. A commit will never have any items in this tree as it is a leaf. To
-accompany the tree, we include breadcrumbs at the top of the page to enable a user
-to navigate both from the root to the leaves and from a leaf to the root. The last
-item in the breadcrumb list is the current commit, the previous item is the parent
-of the current commit, and the first item is the root merge into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List tree view of merge 3f17ea6}
-        \label{fig:list_tree}
-\end{figure}
-
-\subsubsection{Reingold-Tilford Tree}
-
-The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:reingold_tree} and
-\ref{fig:reingold_tree_zoom}) allows the visualization and navigation of the entire merge-tree in
-an intuitive representation of the tree.  This illustrates a clear notion of root
-and leaves, and how to navigate in either direction. Some merge trees are very
-large, containing thousands of commits and merged. While the tree is capable of
-producing a visualization, it becomes far more difficult to understand. For example,
-the merge, 3f17ea6 performed by Linus Torvalds June 8 2014, contains 7217 commits
-and merges. \footnote{This commit can be inspected at
-  \url{http://li.turingmachine/org/commits/3f17ea6dea8ba5668873afa54628a91aaa3fb1c0}}
-
-      \begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/4dc4_tree.pdf}
-        \caption{Merge 4dc4226 is a subtree of 3f17ea6, for the power-management
-          module of the kernel}
-        \label{fig:reingold_tree}
-      \end{figure}
-
-      \begin{figure*}
-        \centering
-        \includegraphics[width=0.98\textwidth]{figures/4dc4_zoom_tree.pdf}
-        \caption{Zoomed out view of the Reingold-Tilford tree for merge 4dc4226}
-        \label{fig:reingold_tree_zoom}
-      \end{figure*}
-
-      The user is initially greeted with their current node centered on the screen.
-      They are able to zoom the tree by scrolling the mouse, and clicking and
-      dragging to pan the tree. They can see more details about a node by clicking
-      on it, which will provide them a link to the specific page for that commit.
-
-\subsubsection{Bubble Tree}
-
-Bubble trees are useful for providing a clear visualization of wide, hierarchical
-data\cite{Boardman2000}. The tree structure is represented by the nesting of nodes;
-the largest circle is the root, containing all the other nodes. The smallest circles
-do not contain any nodes and therefore represent the leaf nodes.
-
-Our implementation of the bubble tree (Figure~\ref{fig:bubble_tree}) provides the
-user with a clear picture of where a commit is located in the merge tree. We
-highlight the selected commit or merge in red. Non-selected merges are in a shade of
-blue determined by the depth of a node in the tree. The root is the lightest shade
-of blue, while the contained merges are progressively darker; the commits are white.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree of merge 3f17ea6, merge 4dc4226 is the currently
-          selected merge and is highlighted in orange.}
-        \label{fig:bubble_tree}
-\end{figure}
-
-The bubble tree doesn't have an implicit way of providing additional information,
-placing any text near the nodes makes the tree impossible to read, so we include a
-separate pane in the web page. When a user hovers over a node, the pane shows
-additional information about the author and the commit message and a link to the
-detailed page for that commit or merge. If the user clicks on a node, the tree will
-zoom to that node and the information in the info pane will persist, enabling the
-user to click the link. This tree provides an easy mechanism for users to navigate
-from the root to the leaves and vice versa.
-
-\section{Implementation Details}
-
-\evantodo{See if we can move API section to be part of another section}
-
-The front-end uses asynchronous requests to gather the information for the trees and
-tables. This enables third-parties to implement new front-ends, trying other
-designs, though this interface may change in the future. The tree, file, and module
-information is accessible through
-http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.
-
-The response for the tree is a single object. This object is the root node of the
-tree, and may contain an object of children sub-tree objects in the \verb|children|
-field, which are in the same format as the root object. The children object uses the
-commit ID as the key.
-
-The tree node objects have the following fields:
-
-\begin{tabular}{ccl}
-        Field name      & Data type & Description\\\hline
-        \verb|cid|      & string    & Commit ID\\
-        \verb|name|     & string    & One-line log summary\\
-        \verb|mlinus|   & string    & Root merge commit ID\\
-        \verb|author|   & string    & Author name and email\\
-        \verb|mnext|    & string    & Parent merge commit ID\\
-        \verb|children| & object    & Object of tree objects\\
-\end{tabular}
-
-The file responses contain only the files that the selected merge or commit works
-with. The response is a single object in the form of a tree. This object is the root
-of the tree, and represents the current merge or commit. If the current position in
-the tree is an inner node, the response will contain the child nodes in the
-\verb|children| field, otherwise \verb|children| will be an empty list. The format of file
-response is as follows,
-
-\begin{tabular}{lll}
-        Field name & Data type & Description\\\hline
-        \verb|cid| & string & Commit ID\\
-        \verb|mnext| & string & Parent merge commit ID\\
-        \verb|children| & list & List of tree objects\\
-        \verb|files| & list of tuples&
-        \footnotesize{
-                \begin{tabular}{lc}
-                        \verb|Filename| & string\\
-                        \verb|Lines added| & uint\\
-                        \verb|Lines removed| & uint\\
-                \end{tabular}}\\
-\end{tabular}
-
-\section{Discussion}
-
-From a developers point of view, the are two major disadvantages of the DAG model of
-Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
-commit that succeeds them on the path to being integrated; and b) the DAG contains
-many more edges that are necessary to understand how integration occurred. We
-addressed these two issues with the creation of the merge-tree model from the DAG of
-Linux. Effectively, the merge-tree recovers the details of how each branch---and the commits they contain---was merged
-into the master branch.
-
-In our experience, no other Git repository reaches the level of DAG complexity that
-Linux has. In most Git repositories, most merges are not nested---most merges merge
-directly into the master repository, and these merges only merge a few commits. Even
-in these simplified cases, the merge-tree can provide a valuable summary of how
-commits are integrated into the master branch, specifically since the time of
-integration may be very different from the time the commits were authored.
-
-\evan{Check accuracy of this paragraph}
-
-The biggest challenge is computing the merge-tree of any given repository. It
-is likely that our heuristic for computing the merge-tree will not work with other
-repositories. The heuristic requires that there are no fast-forward merges into the
-master branch of the repository (a majority of repositories), and that there are no
-foxtrot merges (a practice that is starting to be considered desirable by git users\footnote{See
-  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git} and
-  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
-We are able to use our heuristic with the Linux kernel because of
-the strict integration model imposed by Linus. To validate the results of the
-heuristic, we use the continuous mining technique described in \cite{German2015}.
-Continuous mining of the repository requires foresight and planning.
-
-Because \tool leverages the merge-tree model of the kernel, it provides mechanisms
-and visualizations that other tools are unable to produce. It gives us the ability to
-see how commits are being merged into the master branch, allowing us to better understand
-how code is integrated into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
-        \caption{Merge Dag View}
-        \label{fig:dag_view}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
-        \caption{Merge tree topology for the same merge as in
-                Figure~\ref{fig:dag_view}.}
-        \label{fig:tree_view}
-\end{figure}
-
-Our goal was to build a tool to enable maintainers to effectively navigate and
-browse the changes performed to the kernel over the period of a release. Achieving
-this goal includes removing information that does not pertain to the area of the
-kernel that the user is interested in. In Figures~\ref{fig:dag_view} and
-\ref{fig:tree_view} we can visually compare the results returned from Gitk
-and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. Note that Gitk has no way to show or recognize that this
-is a merge into the master branch.
-
-The log information to this merge-commit in both of these figures is identical, but
-the presentation drastically changes our ability to comprehend what we are seeing.
-The primary difference is the removal of irrelevant information. In Gitk
-(Figure~\ref{fig:dag_view}), the visualization results include commits and merges
-from other components of the kernel, while our tool (Figure~\ref{fig:tree_view})
-only includes the results that are specific to the component of the kernel that we
-are interested in.
-
-Even though this merge is relatively small, containing only four sub-merges, three of which
-only contain a single commit, and the last containing two commits, it is already hard to visualize in Gitk.
-With \tool we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
-furthermore, users must work to determine where this merge ends and the next one
-begins.
-
-\tool further enhances our understanding by summarizing the files
-and modules that were edited in the entire branch being merged. We are able to determine that three
-files, ``palamas-regulator.c'' had two lines added and two lines removed,
-``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
-lines added and 2 lines removed. Finally, we are able to determine that only the
-``regulator'' module was modified in this merge and was modified by 5 commits.
-
-\section{Future Work}
-
-There are still many areas that can be improved before the full potential of our
-model can be realized. Here, we outline various areas of the tool that still need
-more attention.
-
-\subsection{Files}
-
-There currently is no functionality surrounding searching by filename. Users falling
-under use-case 2 may know what file they are editing and try to determine how it may
-work with the other commits and merges in the module. It is possible that a third
-use-case may arise, where the user wants to determine all commits that effect a
-given file. In both cases, bottom-to-top approach is applied.
-
-There is limited functionality to the presentation of the file information. At a
-minimum, the patches for the commits can be displayed. From there, the patches can
-be used to piece together parts of the file to generate a single patch at a given
-merge rather than displaying each patch individually. The patches can also be used
-for determining what kinds of changes were made, if the lines are being added,
-removed, or being replaced.
-
-\subsection{Authorship}
-
-Our model can aggregate more information than what we have implemented. The
-authorship information is important for licensing purposes, and we can show all
-authors contributing to commits in a  merge and how many commits they contributed to
-the merge. We could go further, providing information about what files they edited,
-how many lines they added and removed, and more.
-
-\subsection{Evaluation}
-
-At this time, we have no evidence that our tool is able to improve the work-flow of
-maintainers. We believe that the tool is able to improve the work-flow and
-performance of maintainers because it provides cleaner mechanisms of visualizing and
-presenting the commit information. It is able to provide more relevant information,
-while removing information that is irrelevant to a given module or set of merges.
-
-We can either perform user-testing to show how a users workflow changes, or we could
-have maintainers evaluate and critique the tool and use their feedback to determine
-if the tool meets our goals.
-
-\section{Related Work}
-
-Version control systems monitor the development lifetime of software projects. This
-makes the version control system vital in providing information about how a software
-project is being developed. To our knowledge, we know of no Git repository
-visualization tool that builds a tree that maps the path a commit follows to the
-master branch from the DAG provided by Git. This may be because more information is
-required to generate the merge-tree model than what is stored by Git. However, there
-has been a lot of work in providing visualizations of various repositories.
-
-Many tools work to address the issues in communication between developers in
-inter-team collaboration work. Codebook \cite{Begel2010} uses a data mining
-technique  to determine the developer of a piece of code, the program manager who
-wrote the specification for the code, and the program managers and developers on the
-team who were working together. Hipikat\cite{Cubranic2005} is another tool with a
-focus on communication. Where Codebook focuses on developers working on a project,
-Hipikat is focused on enabling easier integration of new developers to a project by
-providing them with easily-searchable artifacts of the changes made. Codebook is
-useful for pairing a contributor with the original developer; however, the developer
-may not have worked with the piece of code in years. A Hipikat program may provide
-more information to the maintainer as it records the artifacts of why certain
-design decisions were made when they were made using other tools like Bugzilla and
-CVS. Neither tool is sufficient in meeting our goals to provide a summary of the
-topology of the kernel repository through a visual tree.
-
-Most visualizers provide a visual presentation of a certain aspect of a repository.
-Fractal Figures \cite{Ambros2005} uses a unit square to represent a portion of a
-project, then partitions the square based on the proportion of an author's
-contributions to that portion of the project. EPOSee\cite{Burch2005} and Evolution
-Radar\cite{Ambros2009} perform further analysis, determine which files are made
-together, and what changes are made over a sequence of commits, though the goals
+Most visualizers provide a visual presentation of a certain aspect of a
+repository. Fractal Figures\cite{Ambros2005} uses a unit square to
+represent a portion of a project, then partitions the square based on
+the proportion of an author's contributions to that portion of the
+project. EPOSee\cite{Burch2005} and Evolution Radar\cite{Ambros2009}
+perform further analysis, determine which files are made together, and
+what changes are made over a sequence of commits, though the goals
 behind these projects is different.
 
-Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work with data
-from CVS repositories. Our goal is to provide information about Git repositories,
-specifically the Linux master repository. Fewer tools are available for generating
-visualizations and summaries of Git repositories, potentially due to the DAG model
-used by Git.
+Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work
+with data from CVS repositories. Our goal is to provide information
+about Git repositories, specifically the Linux master repository. Fewer
+tools are available for generating visualizations and summaries of Git
+repositories, potentially due to the DAG model used by Git.
 
-Gource is a tool for providing an interactive timelapse of the state of a
-repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
-type of contribution a developer is making. These contribution types are one of,
-adding a file, removing a file, and changing a file. While the timelapse is
-interesting to watch, it does not provide any additional explanation of the changes
-actually being made, only the frequency that they are being made and who is making
+Gource is a tool for providing an interactive timelapse of the state of
+a repository\cite{Caudwell2010}. In the timelapse, it shows who
+contributes and what type of contribution a developer is making. These
+contribution types are one of, adding a file, removing a file, and
+changing a file. While the timelapse is interesting to watch, it does
+not provide any additional explanation of the changes actually being
+made, only the frequency that they are being made and who is making
 them.
 
-The current industry standard tool for Git repository visualization is Gitk. The
-Gitk interface is built around the central DAG viewer. The DAG view displays the
-commits on their respective branches, the author of the commit or merge, and the
-date the commit or merge was authored. A user can select a single merge or commit
-to view additional information. For both merges and commits, Gitk will present the
-user with the Git log. If the user selected a commit, Gitk will also present the
-user with the patch information and the names of the files edited. Gitk is unable to
-provide the patches and filenames information in merges, as it is unable to
-aggregate commit information in merges.
+% TODO: add code_swamp or something like that
 
-Our tool is primarily aimed at presenting the hierarchical structure of the Linux
-git repository. We use tables for presenting the summarized information of the
-commits and merges, but this information could also be presented in a graphical
-form. Various graphical forms for displaying file and authorship data exist, the
-principal forms being matrix views, city scapes, bar and pie charts, and networks
-\cite{Eick2002}. Any of these data visualization metaphors are applicable to our
-system.
+% TODO: Check this -- Look into source tree
+Not included in an academic project are both the git commandline with
+the \verb|git log| command, and Gitk, which is a graphical tool shipped
+with git. Many other tools are designed with a similar visual metaphore
+as what is presented in Gitk, including Git kraken, Github Desktop and
+web interfaces, the Git Lab web intereface, the Bitbucket web interface,
+and many others. The Gitk interface is built around the central DAG
+viewer. The DAG displays the repository events on their respective
+branches, the author, and the date that the event was authored. A user
+can select an event to view addition information about it, including the
+full commit log, the parents and children of the event, and the diff
+generated if modifications were made. In the case of the Linux kernel,
+merges are not resolved as merge conflicts in the master repository, but
+by the developers prior to merging. With this development model, no
+merges will have modified any files, and will not show a diff. While the
+merge does not make changes directly, the effects of merging the
+commits persist, but Gitk is unable to provide an aggregated view of the
+files modified, the patches for those files, and the authors who
+contributed commits to that merge.
 
-\section{Conclusion}
-
-Our tool, \tool, shows promising results in working toward our goals, building a tool that
-can more easily navigate the repository, and provide clearer explanations of the
-changes made. The filtering by searches and the various tree visualizations enable
-users to easily navigate through the kernel, finding the commits and merges that
-pertain to what they are interested in. The tree visualizations are able to both
-serve as a navigation piece, providing simpler navigation through the kernel, but
-also demonstrate which commits contribute to which merges, providing further
-explanation over the DAG. The tree model further improves the explanation by enabling
-the tool to aggregate metadata at the merges instead of requiring a user to manually
-aggregate the information. Our tool currently aggregates file information and the
-modules edited, but can be extended to include authorship information and aggregated
-patch information.
-
-\evan{Check the validity of this paragraph}
-We use a heuristic method to generate the tree from the repository. The heuristic
-requires that the repository does not have fast-forward merges into the master
-branch or foxtrot merges, which obfuscate the master branch. In either case, the
-heuristic will break and result in a tree that does not accurately represent the
-repository.
-
-These may pose as significant drawbacks to smaller projects; however, with a proper
-merge discipline, the heuristic is able to correctly convert the DAG into a
-merge-tree. The importance of clear visualizations in large projects outweighs
-the costs of enforcing a clean merge discipline. The tree-based model appears to
-have the ability to provide a clean visualization and more detailed explanations of
-the changes in the Linux kernel.
-
-% \nocite{*}
-
-\bibliographystyle{IEEEtran}
-\bibliography{citations}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:
+Our tool is primarily aimed at presenting the hierarchical structure of
+the Linux git repository. We use tables for presenting the summarized
+information of the commits and merges, but this information could also
+be presented in a graphical form. Various graphical forms for displaying
+file and authorship data exist, the principal forms being matrix views,
+city scapes, bar and pie charts, and networks \cite{Eick2002}. Any of
+these data visualization metaphors are applicable to our system.

### 2017-07-05 paper/methodology.tex
Methodology: Changing order of subsections
New ordering is :

 - Tasks
 - Commit selection
 - User selection
 - Data Extraction
 - Analysis

@@
 comprehension and summarization of git repositories, we conducted a user
-study on 12 participants. The user study is a
-mixed-methods study, but with more focused on the quantitative aspects.
-We chose to use a mixed-methods approach to determine both if the model
-is able to improve the time performance and accuracy of the
-participants, and if the participants found the merge trees more
-enjoyable for the given tasks.
+study on 12 participants. The user study is a mixed-methods study, with
+focus on the quantitative aspects in order to capture how effective the
+\mt model, if it is, at improving the correctness, accuracy, and time
+performance metrics of users when performing summarization tasks. We use
+the qualitative questions as a means of determine the preference of the
+participants, and what aspects they prefered from each tool.
 
-In this section, we describe the describe the methods used in the user
-study. This includes the participant criteria, the questions asked, the
-method for selecting commits, and a breakdown of the methods used for
-analyzing the results.
-
-\subsection{User Selection}
-\label{sub:user_selection}
-
-Our target audience are git users, particularly users with who are
-working on understanding what changes are being made, where they are
-being made, and who is making the changes. The merge-tree model is also
-intended for users who are trying to learn the structure of a
-repository, how modules of the repository are broken into groups, if the
-
-We have software engineering and computer science students at the
-Masters, Doctorate, and Post-Doc levels available for participation in
-this study. \evan{This feels like the weakest part of the paper, is
-  there any way to make it sound stronger?}
-
-\subsection{Commit Selection}
-\label{sub:commit_selection}
-
-We chose two commits for use in the study. The order that the commits
-are presented is randomized between participants, but the order is kept
-consistent through the tasks described in the next subsection.
-
-We use the \tool database, containing commits and merge-tree information
-from April 16th 2005 to October 14th, 2014, which corresponds to being
-between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
-  database even more?} We chose the commits based on tree sizes. We
-found that a majority of the trees contain at most seven inner commits,
-while more than 25\% of the trees contain a single commit. 75\% of the
-tree contained up to 51 inner commits and merges, and finally the
-largest tree contained 7217 nodes. We chose to work with the trees in
-the first and second quartile, as merge trees of sizes between one and
-seven, not including the merge into the master branch, make up the
-majority of the trees in the database.
-
-From here, we selected one tree from the trees of a single commit at
-random. Selecting a commit from that tree is trivial, as there is only
-one to choose.
-
-We selected one tree of size seven at random from all trees of size
-seven. We placed a restriction the trees, the selected tree must include
-at least one internal merge to increase the complexity of the trees
-tested. After randomly selecting a tree, we chose in internal commit at
-random, using the \verb|random.choice| function in python 3.6.1.
-
-% TODO: Show this paragraph in the introduction
-% It is worth noting that in the fourth
-% quartile, the size of the tree drops off quickly from the top tree, with
-% the next tree containing only 4708 nodes, and the third largest tree
-% containing only 2349 commits.
-
-Using this technique, we selected commit
-\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
-a single node (Figure~\ref{fig:commit_1}), which we will refer to as
-\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
-the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
-will refer to as \comB.
-
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
-  \caption{The first merge tree used in the user study, a merge tree
-    containing a single commit}
-  \label{fig:commit_1}
-\end{figure}
-
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
-  \caption{The second merge tree used in the user study, a merge tree
-    containing seven commits}
-  \label{fig:commit_2}
-\end{figure}
+In this section, we describe the methods used in the user study. This
+includes an outline of the tasks, the commit selection, and the user
+selection.
 
@@
 
-The questions and tasks are broken down into three groups. The first two
-groups are purely quantitative in nature, being conceptual tasks, and
-summarization tasks. The third group of questions is to determine the
-demographic of our participant.
+The questions and tasks are broken down into four groups. The first two
+groups are purely quantitative in nature; we begin with conceptual tasks
+to understand whether the DAG is capable of providing users with an
+understanding of how a commit is merged into the master branch, and
+other commits that it is merged with. We continue with summarization
+tasks, to quantitatively understand if the \mt model makes a difference
+in summarization tasks. The third portion is purely qualitative, to gain
+better understanding of whether participants preferred working with the
+\mt model in \tool, or working with the DAG in Gitk, and what aspects
+they preferred from each tool. The final group of questions is simply to
+gain a better understanding of the demographic of our participants.
+
+%TODO: Look at how to form research questions in LaTeX
+
+We have two primary research questions;
+\begin{enumerate}
+  \item Is the DAG capable of providing the participants with an
+    understanding of the repository events that are related to a commit?
+  \item Does the \mt model make a difference in summarization tasks?
+\end{enumerate}
 
@@
 
-% TODO: provide the script for question selection
-% \lstinputlistings{../selector.py}
-
 \textbf{User Opinion}
@@
 \end{enumerate}
+
+\subsection{Commit Selection}
+\label{sub:commit_selection}
+
+We chose two commits for use in the study. The order that the commits
+are presented is randomized between participants, but the order is kept
+consistent through the tasks described in the next subsection.
+
+We use the \tool database, containing commits and merge-tree information
+from April 16th 2005 to October 14th, 2014, which corresponds to being
+between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
+  database even more?} We chose the commits based on tree sizes. We
+found that a majority of the trees contain at most seven inner commits,
+while more than 25\% of the trees contain a single commit. 75\% of the
+tree contained up to 51 inner commits and merges, and finally the
+largest tree contained 7217 nodes. We chose to work with the trees in
+the first and second quartile, as merge trees of sizes between one and
+seven, not including the merge into the master branch, make up the
+majority of the trees in the database.
+
+From here, we selected one tree from the trees of a single commit at
+random. Selecting a commit from that tree is trivial, as there is only
+one to choose.
+
+We selected one tree of size seven at random from all trees of size
+seven. We placed a restriction the trees, the selected tree must include
+at least one internal merge to increase the complexity of the trees
+tested. After randomly selecting a tree, we chose in internal commit at
+random, using the \verb|random.choice| function in python 3.6.1.
+
+% TODO: Show this paragraph in the introduction
+% It is worth noting that in the fourth
+% quartile, the size of the tree drops off quickly from the top tree, with
+% the next tree containing only 4708 nodes, and the third largest tree
+% containing only 2349 commits.
+
+Using this technique, we selected commit
+\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
+a single node (Figure~\ref{fig:commit_1}), which we will refer to as
+\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
+the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
+will refer to as \comB.
+
+\begin{figure}[bpt]
+  \centering
+  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
+  \caption{The first merge tree used in the user study, a merge tree
+    containing a single commit}
+  \label{fig:commit_1}
+\end{figure}
+
+\begin{figure}[bpt]
+  \centering
+  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
+  \caption{The second merge tree used in the user study, a merge tree
+    containing seven commits}
+  \label{fig:commit_2}
+\end{figure}
+
+\subsection{User Selection}
+\label{sub:user_selection}
+
+Our target audience are git users, particularly users with who are
+working on understanding what changes are being made, where they are
+being made, and who is making the changes. The merge-tree model is also
+intended for users who are trying to learn the structure of a
+repository, how modules of the repository are broken into groups, if the
+
+We have software engineering and computer science students at the
+Masters, Doctorate, and Post-Doc levels available for participation in
+this study. \evan{This feels like the weakest part of the paper, is
+  there any way to make it sound stronger?}
 

### 2017-07-05 paper/methodology.tex
Methodology: Removing tables answer tables

@@
 
-%TODO: Look at how to form research questions in LaTeX
+We have two primary research questions;
+\begin{RQ}
+  \item
 
-We have two primary research questions;
-\begin{enumerate}
-  \item Is the DAG capable of providing the participants with an
-    understanding of the repository events that are related to a commit?
-  \item Does the \mt model make a difference in summarization tasks?
-\end{enumerate}
+    Is the DAG capable of providing the users with an understanding of
+    the repository events that are related to a commit?
+
+  \item
+
+    Does the \mt model make a difference in summarization tasks?
+
+\end{RQ}
 
@@
   there any way to make it sound stronger?}
-
-\subsection{Data Extraction}
-\label{sub:data_extraction}
-
-The study is recorded using screen capture and audio capture software
-for further analysis. From the resulting videos, we are able to extract
-the timing and accuracy of each participant over each task. We record
-the timing, accuracy, task, commit, and tool used for each part. We also
-record some addition information about whether the participant was
-correct, guessing, partially correct, or completely wrong, as this may
-provide some interesting insights.
-
-The answers to each question is relatively straight-forward, as we can
-extract the information directly from gitk and Linvis. Measuring the
-accuracy of the questions is more challenging. In the question regarding
-the merges leading to the merging of a given commit, both the merges and
-the ordering of the merges matter. In the cases where the answer is not
-a numerical value, we use the edit distance to determine the accuracy of
-an answer. An edit distance of zero indicates a correct answer, while
-any number other than zero indicates an incorrect answer. Adding,
-reordering (where important), and removing commits or merges all incur a
-penalty of 1 added to the edit distance.
-
-The correct answers for \comA are shown in
-Table~\ref{tab:coma_study_answers}, and the correct answers for \comB are
-shown in Table~\ref{tab:comb_study_answers}.
-
-\begin{table}[htpb]
-  \centering
-  \caption{Answers to the question from the conceptual and summarization portions of the user study for \comA}
-  \label{tab:coma_study_answers}
-  \begin{tabular}{l|l}
-    Question                                              & Answer \\\hline\hline
-    Number of commits in the merge tree                   & 1\\
-    Number of merges in the merge tree                    & 1\\\hline
-    Merges merging this commit into the master branch     & 1 (11df586407)\\
-    Other commits that are merged in the same tree        & None\\
-    How many authors are involved with this tree?         & 1\\
-    Who contributed the most changes to this merge tree   & David\\
-    How many files are modified in this merge tree        & 2\\
-    Which file(s) had the most changes in this merge tree & both (pzl.c, asl.c)\\
-    Which modules does this merge tree involve & wusb
-  \end{tabular}
-\end{table}
-
-\begin{table}[htpb]
-  \centering
-  \caption{Answers to the question from the conceptual and summarization portions of the user study for \comB}
-  \label{tab:comb_study_answers}
-  \begin{tabular}{l|l}
-    Question & Answer \\\hline\hline
-    Number of commits in the merge tree & 5 \\
-    Number of merges in the merge tree & 3\\\hline
-    Merges merging this commit into the master branch     & (219b22b245
-    \\
-    & $\rightarrow$ \\
-    & 8eb88c80d4)\\
-    Other commits that are merged in the same tree        & ec4e86ba06\\
-    & 9a3f371e99\\
-    & 1c85cc6445\\
-    & 7c2dfee848\\
-    How many authors are involved with this tree?         & 4\\
-    Who contributed the most changes to this merge tree   & Randy Dunlap\\
-    How many files are modified in this merge tree        & 6\\
-    Which file(s) had the most changes in this merge tree & pcm\_lib.c\\
-    Which modules does this merge tree involve            & ALSA\\
-  \end{tabular}
-\end{table}
-
-Extracting the timing information is prone to some issues with where the
-duration of answering should start and stop. We try to eliminate these
-issues by maintaining the following conventions; the answer duration
-begins at the end of the last word of the prompt. The duration ends
-immediately before the first word of the final modification to the
-answer. Many answers are able to be answered in a single word or phrase,
-but some may require multiple sentences that are broken by additional
-processing by the participant. An Example of questions where the answer
-may be interleaved with solving the task is the task asking the user to
-provide the merges on the path from the commit into the master branch.
-The goal of this is to eliminate the duration required to provide the
-prompt, and eliminate the time required to provide the answer.
-
-The information is then inserted into a sql database for further
-analysis.
-
-\subsection{Analysis}
-
-As the goals of each section are different, the techniques for analyzing
-the results are different. The conceptual questions are to demonstrate
-the difficulty in comprehending the DAG model, so we simply provide the
-metrics, showing how well the participants did in answering the
-questions asked. The summarization tasks are meant for comparing the
-ability of the participants to summarize two commits from the Linux
-repository. As this is a comparison, further analysis is necessary to
-determine the significance of our results. The final portion of the
-study is designed to help us understand the group of users that we are
-working with.
-
-As the goals of the conceptual questions are to demonstrate the
-difficulty in comprehending the DAG model, and determine how closely
-users are able to make the mapping from the DAG to showing how a commit
-was merged into the master branch, no further statistical analysis will
-be performed on this data. We will provide the metrics response,
-accuracy, and time metrics for these questions. No comparisons are made
-between the models in these tasks, so no further analysis is necessary.
-
-The summarization tasks are designed to compare the two models when the
-participant is summarizing various aspects of the repository, including
-the files modified and authorship information. We will use three terms
-to describe the performance of a user; correctness, which indicates
-simply whether the participant was correct or not, making no indication
-toward how far from the correct answer they were; accuracy, indicating
-how far from the correct answer the participant was; timing, which
-indicates how long a participant took to answer the question. We begin
-the analysis of these questions using the McNemar test, with $\alpha =
-0.005$, measuring the correctness of the results provided by the
-participant when using \tool versus gitk.
-% TODO: cite the McNemar test maybe?
-If the McNemar test indicates that we do not reject the null hypothesis,
-we continue analyzing the difference in accuracy and timing metrics for
-the merge-tree model versus the DAG.
-
-For each test, we use a confidence level $\alpha = 0.005$, resulting in
-a $\chi^2$ test statistic of 7.879.
-
-% TODO: Add the null hypothesis and type of metrics analyzed for each summarization question

### 2017-07-05 paper/lineval.tex
Final Descent: Removing section files
Set up sections for the final version of the paper.
Introduction: DAG, Merge Tree, Evaluation, and Implementation

@@
-% vim:set et fenc=utf-8 ft=tex sw=2 ts=4 tw=84:
+% vim:set et fenc=utf-8 ft=tex sw=2 ts=2 tw=72:
 
@@
 \begin{abstract}
-  \input{abstract.tex}
 \end{abstract}
@@
 
-\input{introduction.tex}
+Git is a distributed version control system that is used in many
+open-source and closed-source projects. Git enables many developers to
+write code simultaneously, working on various components known as
+branches. Git uses a directed acyclic graph model (DAG) for storing the
+necessary information regarding the state of the code at any point in
+time. Git refers to nodes of the DAG as commits, not distinguishing
+between the code-carrying commits and the structural commits that bring
+two branches together. In this paper, we will refer to the individual
+commits as being commits, the structural merge commits as merges, and
+refereeing to all as repository events.
+
+Between 50k and 70k commits are added to the Linux kernel per version
+requiring maintainers of older versions of the kernel to sift through
+thousands of commits and merges with tools that are unable to filter and
+effectively visualize projects at the scale of the kernel. Older
+versions of the kernel are used in embedded systems and mobile phones;
+for security purposes, performance needs, and changing hardware
+requirements, maintainers must be able to understand the changes being
+made in the current version of the kernel in order to produce the
+necessary patches for the older versions of the kernel. Current tools
+and visualizations are based on the directed acyclic graph (DAG) of the
+repository, showing all commits and merges in chronological order by
+when the commit was authored, not by when it arrived in the official
+Linux repository.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/gitk.png}
+  \caption{
+    The Gitk interface, centered at commit
+    cdbdd1676a5379f1d5cbd4d476f5e349f445befe. The top left pane showing
+    the DAG, the top center pane showing the authors, and top right pane
+    showing the commit dates. The bottom-left pane shows the full
+    commit message and diff of the selected commit, and the bottom-right
+    pane shows the list of the files modified by the selected commit.
+  }
+  \label{fig:gitk}
+  %\vspace{-4mm}
+\end{figure}
+
+% DAG model
+
+The DAG representation works in smaller projects; it enables users to
+see when changes are made, when these changes are merged, how each
+branch is interacting, and the point where a branch forks from the
+master branch. In large modular projects, like the Linux kernel, the DAG
+becomes a mess of merges and commits (Figure~\ref{fig:gitk}) losing its
+visual meaning. In some cases, the repository is simply too large for
+systems to generate a visualization; repositories stored on GitHub will
+have a DAG visualization to show the branches and forks of a
+repository, but is unable to display the visualization for projects at
+the scale of Linux or even the Git Git
+repository(Figure~\ref{fig:gitfail}). Between 60k and 70k new commits
+are created for the Linux project every year; according to previous
+work\cite{German2015}, a commit takes a median of 30 days from the time
+it is authored until it arrives in the official repository. The snapshot
+of the kernel tomorrow may be different than the snapshot from today,
+containing new commits authored in the past; distinguishing these new
+commits from the commits in the snapshot from today is not trivial.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
+  \caption{Github is unable to display the DAG of the Linux
+    Project due to the size and complexity, which would normally
+    show the relationship to other branches and forks.}
+  \label{fig:gitfail}
+  %\vspace{-2mm}
+\end{figure}
+
+% Merge tree model
+
+We propose a new model, built from the DAG, with the purpose of
+providing a clearer conceptual image of how a commit is merged, other
+commits that are merged with a commit, and a summarization of the entire
+group of commits that are merged with a commit. We call our model the
+\mt model. The model inverts the DAG, so that reference point from
+parent to child instead of from child to parent. This enables the
+aggregation of commit metrics at merges. Given a merge, we are able to
+recursively traverse the children of the merge, collecting the
+statistics at each commit. The commits become the leaves of the tree,
+and internal nodes are merges that do not merge into the master branch
+of the repository. We terminate the further traversal of an edge under
+two conditions; first, if the next event is on the master branch of the
+repository, second, if the next event has a shorter path into the master
+branch of the repository. As a consequence, commits will always take the
+shortest path, in terms of commit time, to the master branch.
+
+A short example: assume the commits represented in
+Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
+In this case, commits are performed in various repositories and
+branches. The DAG representation of the commits is shown in
+Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
+the master branch and the repository that the master branch is part of.
+The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
+Note that the direction of the edges of the DAG have been inverted,
+instead of pointing from the child to the ancestors, it points from the
+ancestor to its successors, forming a path to the master branch. Also
+note that the DAG has been simplified, showing only a single edge on the
+path to master for any commit.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
+    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
+
+    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
+    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
+    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
+    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
+
+        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
+    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
+  \end{tikzpicture}
+  \caption{Example of a sequence of events performed in different
+    repositories. The horizontal axis represents time. Each horizontal
+    section represents a different branch and/or repository. Each commit
+    points to its ancestor.}
+  \label{fig:repoEvents}
+%\vspace{-3mm}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \node[state] (1) {1};
+    \node[state, above right= of 1] (2) {2};
+    \node[state, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, right= 2cm of 1] (4) {4};
+    \node[state, above right=of 4] (5) {5};
+    \node[state, above right=of 5] (6) {6};
+    \node[state, right=of 5] (7) {7};
+    \node[state, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, right= 2cm of 7] (9) {9};
+    \node[state, right=of 9] (10) {10};
+    \node[state, below right=of 9] (11) {11};
+    \node[state, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[-stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[-stealth] (9);
+    \draw (10) edge[-stealth] (9);
+    \draw (9) edge[-stealth] (8) edge[-stealth] (6)
+              edge[-stealth] (7);
+    \draw (8) edge[-stealth] (7);
+    \draw (7) edge[-stealth] (5);
+    \draw (6) edge[-stealth] (5);
+    \draw (5) edge[-stealth] (3) edge[-stealth] (2)
+              edge[-stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[-stealth] (2);
+    \draw (2) edge[-stealth] (1);
+  \end{tikzpicture}
+  \caption{DAG representation of the commits represented in
+    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
+    information about which repository a commit comes from. Furthermore,
+    the DAG does not distinguish the master branch from other branches.}
+  \label{fig:repoDAG}
+\vspace{-3mm}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+
+    \draw[chartblue]
+      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (2) edge[chartyellow, -stealth] (5)
+          (3) edge[chartmagenta, -stealth](5)
+          (5) edge[chartyellow, -stealth] (7)
+          (7) edge[chartyellow, -stealth] (9)
+          (6) edge[chartred, -stealth] (9)
+          (8) edge[chartmagenta, -stealth] (9)
+          (9) edge[chartyellow, -stealth] (11)
+          (10) edge[chartyellow, -stealth] (12);
+
+  \end{tikzpicture}
+  \caption{\mt view of the commits  represented in
+    Figure~\ref{fig:repoEvents} showing the path they followed to reach
+    the master branch. In this model the successors of each commit
+    represents the path followed by that commit to reach the master
+    branch.}
+  \label{fig:repoTree}
+\vspace{-3mm}
+\end{figure}
+
+Computing the \mt from a DAG for any repository may not be
+possible; however, certain features of the development process of Linux
+make it feasible to compute the \mt for the Linux repository.
+First, the master branch of Linux is maintained by Linus Torvalds, and
+only Linus has write access to it. We have verified this assertion in
+previous research~\cite{German2015}. We have developed a heuristic that
+is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
+identifies the commits made directly to the master branch; whereafter it
+recursively determines the shortest path (in terms of time), using the
+DAG, from each commit to the master branch using the inverted DAG.
+
+\begin{algorithm}
+        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
+        \begin{algorithmic}
+                \Function{ComputeMergeTree}{DAG}: tree
+                \State {\# Compute the tree from the DAG of Linux repository.}
+                \State {\# Returns $Tree$, a graph containing every commit }
+                \State {\# in DAG with the path it followed to master.}
+
+                \State $head \gets \textit{Head of master of git repository}$
+                \State $master \gets \textit{traverse DAG from head using }$
+                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
+                \State $nodes(Tree) \gets nodes(DAG)$
+                \State \Function{distance2Master}{cid} : seconds
+                \State {\# Helper function}
+                \State {\# Recursively compute shortest distance to master}
+                \State {\# setting cid's successor (next) in its way to master.}
+                \State {\# This function should be memoized. Otherwise it}
+                \State {\# would run in exponential time.}
+                \If {\textit{cid in master}}
+                \State \Return 0
+                \EndIf
+                \State    $d \gets 	\infty$
+                \State {\# Traverse the inverted DAG}
+                \For{$c \in children(cid, DAG)$}
+                \If {$c \in master$}
+                \State $d_1 \gets commitTime(c)-commitTime(cid)$
+                \Else
+                \State {$d_1 \gets distance2Master(c)$}
+                \EndIf
+                \If {$d_1 < d $}
+                \State $next \gets c$
+                \State  $d \gets d_1$
+                \EndIf
+                \EndFor
+                \State {\# $c$ is the commit that follows $cid$}
+                \State {\# in its way to master}
+                \State add edge $(cid, next)$ to $Tree$
+                \State \Return $d$
+                \EndFunction
+
+                \State {\# Compute the distance for each commit}
+                \State {\# discarding result}
+                \For{$c \in nodes(DAG)$}
+                \State $distance2Master(c)$
+                \EndFor
+                \State \Return $Tree$
+                \EndFunction
+        \end{algorithmic}
+\end{algorithm}
+
+We evaluate the performance of the heuristic using the merge log made by
+Linus Torvalds, the original creator of Linux, and primary merger of all
+other repository events into the master branch of the repository. In
+Figure~\ref{fig:sampleMerge}, we see an example of the merge log.
+
+\begin{figure}[htbp]
+  \centering
+  {\fontsize{7}{9}
+    \begin{verbatim}
+Merge: 8cbd84f fd8aa2c
+Author: Linus Torvalds <torvalds@linux-foundation.org>
+Date:   Tue Aug 10 15:38:19 2010 -0700
+
+Merge branch 'for-linus' of git://neil.brown.name/md
+
+* 'for-linus' of git://neil.brown.name/md: (24 commits)
+md: clean up do_md_stop
+[... edited for the sake of space]
+md: split out md_rdev_init
+md: be more careful setting MD_CHANGE_CLEAN
+md/raid5: ensure we create a unique name for kmem_cache...
+...
+    \end{verbatim}}\vspace{-5mm}
+  \caption{Example of how merges record a subset of commits being
+    merged. The commit only shows the first 20 one-line summaries
+    messages for the 24 non-merge commits it merged. The ending
+    ``\ldots'' is part of the log and represents that other
+    commits were merged.}
+  \label{fig:sampleMerge}
+\end{figure}
+
+We use this information to evaluate the accuracy of the \mt model
+extracted from the DAG\@. The method we followed started with the
+extraction of the commit history up to July 20, 2016. We computed the
+merge tree of every commit until then. Since Linus Torvalds mostly does
+merging directly into master, we assumed that every merge by him is the
+root of a \mt. As described above, the log of a merge-commit usually
+contains the number of commits in the merge the first 20 summaries of
+commits being merged. We extracted merges by Linus Torvalds using the
+command \mycode{log --merges --author='Torvalds} and compared the number
+of commits according to the log with the number of commits in the  \mt
+rooted in this commit. We also used the summaries of the commits found
+in the merge (not necessarily all---see above) to make sure those
+commits were in their corresponding merge tree. For example, for the
+merge in Figure~\ref{fig:sampleMerge} we would expect that the merge
+tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
+summaries corresponds to commits in that \mt. We also inspected those
+with differences to make sure they were true errors.  The results can be
+summarized as follows:
+
+\begin{itemize}
+
+  \item
+
+    Five merges were false-errors because their logs did not contain
+    accurate information (were probably edited by hand). For example in
+    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
+    empty), in \mycode {c55d267\ldots} the summaries were reordered.
+
+  \item
+
+    The heuristic correctly identified that 79 of Linus merges (between
+    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
+    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
+    commits.
+
+  \item
+
+    The heuristic worked perfectly until Sept 4, 2007, the earliest date
+    that it could be verified.  Before this date, and until Dec 12,
+    2006, merges did not include a summary of the commits they included,
+    hence making it impossible to verify; during this period, however,
+    we correctly identified the merges by Linus into master.
+
+  \item
+
+    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
+    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
+    which confounded the true master branch of a repository (see
+    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
+    for a description of the issue).
+
+\end{itemize}
+
+In summary, of the merges after Sept. 4, 2007, our heuristic was correct
+in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
+12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
+2007).
+
+% Evaluation
+
+In order to determine the effectiveness of the merge-tree model on
+comprehension and summarization of git repositories, we conducted a
+controlled user study on 12 participants. The user study is a
+mixed-methods study, with focus on the quantitative aspects in order to
+capture how effective the \mt model, if it is, at improving the
+correctness, accuracy, and time performance metrics of users when
+performing summarization tasks. We use the qualitative questions as a
+means of determine the preference of the participants, and what aspects
+they preferred from each tool.
+
+We formulate three research questions that this study is attempting to
+answer;
+
+\begin{RQ}
+  \item
+
+    Is the DAG able to provide users with an understanding of the events
+    in a repository?
+
+  \item
+
+    Do \mt make a difference in the ability to summarize basic
+    aggregated information?
+
+  \item
+
+    Do users prefer the \mt model to the DAG for summarization and
+    comprehension tasks?
+
+\end{RQ}
+
+% Implementation
+
+We implement a tool, \tool, to test and verify the capabilities of the
+\mt model. \tool provides search, summarization, and
+visualization using the \mt model.
+
+We provide users with search-based navigation. The search engine takes a
+plain text query from the user and returns results based on similarity
+to the query. The engine weighs various fields, including the commit
+log, the author name, the commit hash, and the dates that the commit was
+authored and committed. The results are aggregated by the \mt
+with a link to the root at the top, and the repository events that match
+the search query in a table below.  An example of a single \mt is shown
+in Figure~\ref{fig:linvis_search_results}.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=1.0\linewidth]{figures/linvis/search_results.png}
+  \caption{\tool search results for `net-next 2011', showing a single
+    \mt group.}
+  \label{fig:linvis_search_results}
+\end{figure}
+
+\tool uses seven tabs to show the information and visualizations for a
+selected commit or merge. The message tab shows the full commit log
+message. This does not include the diff. The files tab shows an
+aggregated table of all files that were modified in a merge. It includes
+metrics like the number of lines added, lines removed, total lines
+modified, and the delta, summed across all commits in the merge tree
+that modify this file. A small details drop-down button allows a user to
+see exactly which commit makes the changes, as shown in
+Figure~\ref{fig:linvis_files}. If the current repository event being
+viewed is a commit, the aggregate views will only show modifications
+made by the commit. The modules tab shows the modules modified in the
+merge tree. Like the files tab, the modules tab uses a table to show the
+name of the module, the number of commits that are in the merge tree
+that work with the module, and a details button to provide the links to
+those commits.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_files.png}
+  \caption{Table showing the modified files in a merge, with the second
+  entry expanded to show the commit that makes the changes.}
+  \label{fig:linvis_files}
+\end{figure}
+
+The authorship tab is very similar to the files tab, but shows the
+authorship information. It shows the sum of the number of lines added,
+removed, modified, and the delta within the \mt. It also shows
+the number of files that were modified by the author. The details tab is
+organized slightly differently. Instead of organizing these results by
+commit, the details are organized by file, showing which files were
+modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_authors.png}
+  \caption{Table showing the authors who made changes in a merge. The
+    entry for Randy Dunlap is expanded, showing the modifications to
+    each file that were made by Randy.}
+  \label{fig:linvis_authors}
+\end{figure}
+
+\tool implements three visualizations, a list tree, pack tree, and
+Reingold-Tilford tree. The list tree is a series of indented textual
+lists that can be easily searched with the built-in search in most
+web-browsers. The list tree is rooted at the current event, if the event
+is a commit, only that commit will be shown; conversely, if the selected
+event is the root, the entire merge tree will be shown. The Pack-Tree or
+Bubble tree, uses a set-based approach for showing the parent
+relationship. The root node aggregates all of the events within the \mt
+into the master branch. Thus, the root is the superset of all the other
+nodes in the tree. The commits are shown as individual white circles,
+containing no other events, as these are the leaves. This is shown in
+Figure~\ref{fig:linvis_pack}. The selected event will be shown in bright
+orange, commits in white, and merges in shades of blue. Darker shades of
+blue indicate that the merge is deeper in the tree.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
+  \caption{The pack tree visualization for a selected commit, shown in
+    orange.}
+  \label{fig:linvis_pack}
+\end{figure}
+
+The Reingold-Tilford tree is the classic tree visualization, with the
+root at the top, leaves at the bottom, and the edges between them
+showing the parent relationship. An example of the Reingold-Tilford tree
+visualization is shown in Figure~\ref{fig:commit_1} and
+Figure~\ref{fig:commit_2}.
 
@@
 
-\input{relatedwork.tex}
-
 \section{Methodology}
 \label{sec:methodology}
-
-\input{methodology.tex}
 
@@
 
-\input{results.tex}
-
-\section{Generalization}
-\label{sec:generalization}
-
-\input{generalization.tex}
-
 \section{Discussion}
 \label{sec:discussion}
-
-\input{discussion.tex}
 

### 2017-07-05 userstudy/analysis/plots.R
Plots: Generates the summary plots

@@ -1,0 +1,97 @@
+#!/usr/bin/Rscript
+# Jul 01, 2017
+# Evan Wilde
+library(RSQLite)
+library(fitdistrplus)
+
+pdf("figures/results.pdf", width=17, height=12)
+par(mfrow=c(3, 7))
+
+colorScheme <- c("#3366CC", "#DC3912", "#FF9900", "#109618", "#990099", "#3B3EAC")
+colorSchemeTrans <- c("#3366CC99", "#DC391299", "#FF990099", "#10961899", "#99009999", "#3B3EAC99")
+con <- dbConnect(drv=dbDriver('SQLite'), dbname='data/data.db')
+
+query_response_correct <- "
+SELECT question, 'Correct' AS result, tools.name AS tool FROM summarize
+JOIN outcomes ON summarize.outcome = outcomes.id
+JOIN tools ON summarize.tool = tools.id
+WHERE summarize.outcome = 1
+UNION ALL
+SELECT question, 'Incorrect' AS result, tools.name AS tool FROM summarize
+JOIN outcomes ON summarize.outcome = outcomes.id
+JOIN tools ON summarize.tool = tools.id
+WHERE summarize.outcome != 1
+"
+data <- dbGetQuery(con, query_response_correct)
+questions <- sort(unique(data$question))
+
+for (question in questions) {
+  d <- data[(data$question == question),]
+  counts <- table(d$tool, d$result)
+  barplot(counts, legend=rownames(counts), beside=TRUE,
+          ylim=c(0,25),
+          main=paste(c('T', question + 1), sep='', collapse=''),
+          col=colorSchemeTrans[1:2],
+          border=colorScheme[1:2])
+}
+
+
+accuracy_query <- "
+SELECT lt.question AS question,
+lt.distance AS lindist,
+gt.distance AS gitdist,
+(gt.distance - lt.distance) AS 'diff'
+FROM
+(SELECT user, question, distance FROM summarize WHERE tool = 1) AS lt
+JOIN
+(SELECT user, question, distance FROM summarize WHERE tool = 2) AS gt
+ON
+lt.user = gt.user AND
+lt.question = gt.question
+WHERE
+diff IS NOT null;
+"
+
+data <- dbGetQuery(con, accuracy_query)
+eps <- 0.000001
+
+for (q in questions) {
+  d <- data[(data$question == q),]
+
+  if (length(unlist(d)) == 0) {
[…]

### 2017-07-05 paper/lineval.tex
Placing related works back
The original related works section was fine, putting that back in
directly.
User Study: rename methodology -> Controlled User Study
Most of the methodology

@@
 \usepackage{multirow}
+\usepackage{booktabs}
+\usepackage{tabulary}
 
@@
 
-\section{Methodology}
-\label{sec:methodology}
+\input{relatedwork.tex}
+
+\section{Controlled User Study}
+\label{sec:study}
+
+We use a controlled mixed-methods users study in order to evaluate the
+ability for the DAG to provide users with a conceptual image of how a
+commit is integrated into a project, to compare the DAG and \mt models,
+and determine which model the users preferred.
+
+In this section, we describe the methods used to for designing and
+implementing the user study.
+
+\subsection{Questions and Tasks}
+\label{sub:questions}
+
+\begin{table*}[htpb]
+  \centering
+  \caption{Conceptual Tasks }
+  \label{tab:conceptual_tasks}
+  \begin{tabulary}{0.9\textwidth}{LL}
+    \toprule
+    Task & Description\\
+    \midrule
+    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
+    T2 & How many individual commits are related to this commit?\\
+    T3 & How many merges are involved with merging this commit into the master branch?\\
+    \bottomrule
+  \end{tabulary}
+\end{table*}
+
+\begin{table}[htpb]
+  \centering
+  \caption{Summarization Tasks}
+  \label{tab:summarization_tasks}
+  \begin{tabulary}{\linewidth}{LLL}
+    \toprule
+    Task Set   & Task & Description\\\midrule
+    Merge      & T4   & What is the series of merges involved with merging this
+    commit?\\
+               & T5   & What other commits are merged?\\
+    Authorship & T6   & How many authors are involved?\\
+               & T7   & Who contributed the most changes?\\
+    Files      & T8   & How many files were modified?\\
+               & T9   & Which file had the most changes?\\
+    Modules    & T10  & Which modules does this merge tree involve?\\
+    \bottomrule
+  \end{tabulary}
+\end{table}
+
+The questions and tasks are broken down into four groups. The first two
+groups are purely quantitative in nature; we begin with conceptual tasks
+to understand whether the DAG is capable of providing users with an
+understanding of how a commit is merged into the master branch, and
+other commits that it is merged with. We continue with summarization
+tasks, to quantitatively understand if the \mt model makes a difference
+in summarization tasks.
+
+The third task set is purely qualitative, to gain better understanding
+of whether participants preferred working with the \mt model in \tool,
+or working with the DAG in Gitk, and what aspects they preferred from
+each tool. The final group of questions is simply to gain a better
+understanding of the demographic of our participants.
+
+\textbf{Conceptual Tasks}
+
+The participants were allowed to use Gitk or the git command line tools
+to perform these tasks. These tasks are asked in the specified order,
+allowing tabularr to first, create a conceptual understanding of the
+commits and merges that will be involved with the rest of the tasks that
+follow.tabularmay then use their diagrams to answer T2 and T3. We do not
+collect metrics from \tool for the conceptual tasks, opting instead to
+determine if the DAG is sufficient for understanding how a commit is
+integrated. \tool uses the \mt model show a tree visualization of how
+commits are merged, until the commit reaches the master branch.
+Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.
+
+We provide the participant with 10 minutes per commit to complete the
+first task. The answers to the two other questions in this section are
+drawn from the answer in the first task.
+
+\textbf{Summarization Tasks}
+
+The summarization tasks (Table~\ref{tab:summarization_tasks}) are
+grouped into task sets by theme. The tasks within the task set are
+shuffled, and the task sets are shuffled, to ensure that similar
+questions are together, but not in the same order, and to ensure that
+the themes are in randomized order between participants. The order of
+the commits is shuffled, as well. The order of the tool is randomized,
+choosing to start with either \tool or Gitk, to ensure that we do not
+place a bias on one tool through the experiment.
+
+As understanding where the merge begins and ends is vital to answering
+the other question, the merge task set will always be the first task set
+presented; however, the questions will still be re-ordered within the
+task set.
+
+\textbf{User Opinion and Profile}
+
+Finally, we ask the user for their opinions on the tools and for some
+information about their level of experience with git. This portion of
+the study is qualitative and does not rely on either of the tool,
+instead attempting to determine which model provided the user with a
+better experience.
+
+We do not shuffle the questions in this set of questions, as this should
+not have an impact on the opinions of the participant at this point.
+These questions are designed to provide us with a better understanding
+of preference, and insight on why, the participant had that preference.
+
+We ask the following questions in this order;
+
+\begin{enumerate}
+  \item Given these tasks again, which tool would you prefer to use?
+  \item Which aspects of each tool did you like and why?
+  \item How long have you used git?
+  \item If you have used git, for what kind of projects? (personal,
+    school courses, professional?)
+  \item If you have used git, how many commits, files, and contributors
+    were involved with the largest repository you have worked with?
+\end{enumerate}
+
+The last three questions in this section are aimed at getting a better
+understanding of the demographic of our participants.
+
+\textbf{Procedure}
+
+For each participant, we run a simple python script to generate the
+script for the study, eliminating the possibility of bias from us. The
+script randomizes the order of the tools, the order of the commits, the
+order of the task sets, and the order of the tasks within each task set,
+where appropriate.
+
+\subsection{Commit Selection}
+\label{sub:commit_selection}
+
+We chose two commits for use in the study. The order that the commits
+are presented is randomized between participants, but the order is kept
+consistent through the tasks described in the next subsection.
+
+We use the \tool database, containing commits and merge-tree information
+from April 16th 2005 to October 14th, 2014, which corresponds to being
+between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
+  database even more?} We chose the commits based on tree sizes. We
+found that a majority of the trees contain at most seven inner commits,
+while more than 25\% of the trees contain a single commit. 75\% of the
+tree contained up to 51 inner commits and merges, and finally the
+largest tree contained 7217 nodes. We chose to work with the trees in
+the first and second quartile, as merge trees of sizes between one and
+seven, not including the merge into the master branch, make up the
+majority of the trees in the database.
+
+From here, we selected one tree from the trees of a single commit at
+random. Selecting a commit from that tree is trivial, as there is only
+one to choose.
+
+We selected one tree of size seven at random from all trees of size
+seven. We placed a restriction the trees, the selected tree must include
+at least one internal merge to increase the complexity of the trees
+tested. After randomly selecting a tree, we chose in internal commit at
+random, using the \verb|random.choice| function in python 3.6.1.
+
+Using this technique, we selected commit
+\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
+a single node (Figure~\ref{fig:commit_1}), which we will refer to as
+\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
+the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
+will refer to as \comB.
+
+\begin{figure}[bpt]
+  \centering
+  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
+  \caption{The first merge tree used in the user study, a merge tree
+    containing a single commit}
+  \label{fig:commit_1}
+\end{figure}
+
+\begin{figure}[bpt]
+  \centering
+  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
+  \caption{The second merge tree used in the user study, a merge tree
+    containing seven commits}
+  \label{fig:commit_2}
+\end{figure}
 

### 2017-07-05 paper/lineval.tex
Results section

@@
 
+We recorded the audio and screen capture video for each participant in
+the study for further analysis. After capturing the information, we
+extracted various metrics from the videos. From the conceptual tasks, we
+recorded the time in seconds, and the answer for the question. We
+extracted three metrics for the summarization tasks, the time in
+seconds, the correctness, and the accuracy. Correctness is a binary
+value, it is true if the response was correct and false if the response
+was incorrect. The accuracy is measures in distance from the correct
+value. An accuracy of zero indicates a correct answer. All tasks in the
+summarization tasks include an accuracy metric except T7, which will
+only be right or wrong, and therefore cannot meaningfully be measured by
+a distance.
+
+\subsection{Conceptual Tasks}
+\label{sub:conceptual_tasks}
+
+In Table~\ref{tab:conceptual_results}, we are able to see an overview of
+the results from the conceptual questions.
+
+\begin{table*}[htpb]
+  \centering
+  \caption{Results from the conceptual questions}
+  \label{tab:conceptual_results}
+  \begin{tabular}{ll|r|lrr|rrr}
+    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
+    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
+    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
+    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
+    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
+  \end{tabular}
+\end{table*}
+
+Users were able to more closely estimate the number of commits and
+merges in the larger tree, but generally took longer than the smaller
+tree. The tree with a single node resulted in more variability in the
+estimate of number of commits.
+
+It should be noted that these questions were answered after spending
+roughly ten minutes attesting to draw a picture that held the answers to
+these questions.
+
+\subsection{Summarization Tasks}
+\label{sub:summarization_tasks}
+
+In Figure~\ref{fig:summarization_results}, we see all of the
+correctness, accuracy, and timing results. The first row of plots are
+the correctness results. We see in all cases except for T10, that the
+number of instances of a correct result with \tool is roughly the same
+as the number of incorrect instances with Gitk. We test each task with
+the McNemar Chi-square test~\cite{McNemar1947}.
+
+\begin{figure*}[htpb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
+  \caption{Results from the summarization tasks. The first row plots the
+    correctness metric for each task. The second row plots the
+    accuracy metric for each task. The third row plots the time metric
+    for each task.}
+  \label{fig:summarization_results}
+\end{figure*}
+
+With the McNemar's tests, we analyze the participants that change from
+being correct with one tool to being incorrect with the other. This does
+not take into account the number of people who were able to provide the
+correct answer for both tools, nor the people who were incorrect for both
+tools. The results show whether providing the participants with \tool
+over the current standard, gitk, improves the correctness of the
+summarization tasks among the participants. We show the results for the
+number of people who were, with both tools, correct, incorrect, and
+changed in Table~\ref{tab:correctness_table}.
+
+In all cases, except for task T10, we saw statistically significant
+results suggesting that we reject the null hypothesis. We tested each
+correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
+degree of freedom is $7.879$.
+
+\begin{table*}[htpb]
+  \centering
+  \caption{Correctness tables showing how many people stayed correct,
+    incorrect, and how many people changed between tool}
+  \label{tab:correctness_table}
+  \begin{tabular}{c|c|c}
+
+    T4 & T5 & T6\\
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
+                           & Incorrect & 15                           & 2\\
+  \end{tabular} &
+
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
+                           & Incorrect & 14                           & 4\\
+  \end{tabular}
+
+  &
+
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
+                           & Incorrect & 16                           & 2\\
+  \end{tabular}
+
+  \\\hline
+  T7 & T8 & T9\\
+
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 8                            & 0\\
+                           & Incorrect & 13                           & 2\\
+  \end{tabular}  &
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
+                           & Incorrect & 15                           & 2\\
+  \end{tabular} &
+  \begin{tabular}{cc|rr}
+                           &           & \multicolumn{2}{c}{Linvis}\\
+                           &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
+                           & Incorrect & 14                           & 3\\
+  \end{tabular}\\\hline
+
+  & T10 & \\
+  & \begin{tabular}{cc|rr}
+                            &           & \multicolumn{2}{c}{Linvis}\\
+                            &           & Correct                      & Incorrect\\\hline
+    \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
+                            & Incorrect & 8                            & 0\\
+  \end{tabular} & \\
+
+  \end{tabular}
+\end{table*}
+
+The correctness test statistics and conclusion for each summarization
+task;
+
+The correctness test statistics, conclusions, and a description of the
+accuracy and timing results for each summarization task follow:
+
+\begin{itemize}
+
+  \item
+
+    T4; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
+    reject the null hypothesis that \tool does not make a difference in
+    the correctness when determining the series of merges that a commit
+    takes, with $99.5\%$ confidence.
+
+    There is almost no variation in accuracy between participants using
+    \tool, while there is more variance in Gitk. There were two outliers
+    among the instances using \tool, which lie within the second and
+    third quartiles for the accuracy measure of Gitk. The furthest
+    outlier for \tool lies at a distance that is slightly greater than
+    the median for Gitk.
+
+    We see similar results in the timing metric; all significant data
+    occurs within less time than the second, third, and fourth quartile
+    of the Gitk results. Nearly all outlier instances in \tool are
+    within the third quartile group of the instances of Gitk with the
+    exception of the most extreme, which lies in the fourth quartile
+    group.
+
+    The results indicate that \tool is able to assist users with
+    determining how a commit is merged more accurately, and more
+    quickly.
+
+  \item
+
+    T5; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
+    reject the null hypothesis that \tool does not make a difference in
+    the correctness when determining the other related commits, with
+    $99.5\%$ confidence.
+
+    There is very little variance in the accuracy of results for \tool
+    compared to the accuracy of the results in Gitk. The median distance
+    is smaller, indicating that the participants performed more
+    accurately using \tool.
+
+    There is less variance in the time taken to come to an answer with
+    \tool than with Gitk. Furthermore, we see that a majority of the
+    participants were able to complete the task in \tool in much less
+    time than 50\% of the participants using Gitk.
+
+    The results indicate that \tool is able to assist users in
+    understanding which commits are related to a given commit.
+
+  \item
+
+    T6; the resulting $\chi^2$ value is $16$. $16 > 7.879$, therefore we
+    reject the null hypothesis that \tool does not make a difference in
+    the correctness when determining the number of authors involved with
+    the merge tree, with $99.5\%$ confidence.
+
+    We see the same results from this task as with the previous tasks,
+    suggesting that \tool assists users determine the number of authors
+    more quickly and more accurately.
+
+  \item
+
+    T7; the resulting $\chi^2$ value is $13$. $13 > 7.879$, therefore we
+    reject the null hypothesis that \tool does not make a difference in
+    the correctness when determining who contributed the most changes,
+    with $99.5\%$ confidence.
+
+    This question does not include an accuracy metric, as there are only
+    two possible outcomes. We see the same result in the timing metric
+    as we did from the previous questions.
+
+    The results suggest that \tool is able to help users correctly
+    identify the person who contributed the most lines of code to a
+    merge tree in less time.
+
+  \item
+
+    T8; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
+    reject the null hypothesis that \tool does not make a difference in
+    the correctness when determining how many files were contributed,
+    with $99.5\%$ confidence.
+
+    The results of this task are the same as in the previous tasks.
+    This suggest that \tool is able to assist users determine how many
+    files were modified, more quickly and more accurately.
+
+  \item
+
+    T9; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
+    reject the null hypothesis that \tool does not make a difference in
+    the correctness when determining which file had the most changes,
+    with $99.5\%$ confidence.
+
+    While this question appears to have a similar answer to the correct
+    answer for T7, the answer \comA had two files with the same number
+    of midifications, thus the participant must correctly identify both
+    of these files to be correct. For this reason, we include an
+    accuracy metric.
+
+    The results of this task are the same as the previous tasks.
+    This suggests that \tool is able to assist users to determine which
+    file had the most changes.
+
+  \item
+
+    T10; the resulting $\chi^2$ value is $5.44$. $5.44 < 7.879$,
+    therefore we do not reject the null hypothesis, suggesting that
+    \tool does not make a difference in determining which modules are
+    involved with a merge tree.
+
+    The accuracy and timing results for this task are the same as the
+    others, showing that there is less variability in both accuracy and
+    the time taken to have an answer between participants using \tool
+    than using Gitk. Looking Table~\ref{tab:correctness_table}, we can
+    see that, unlike the other questions, a majority of the participants
+    were correct using both tools.
+
+\end{itemize}
+
+\subsection{User Opinions}
+\label{sub:user_opinions}
+
+Among the 12 pariticipants, there was nearly unanimous agreement that
+for conceptual understanding and summarization tasks, \tool was easier
+to work with than Gitk. The participants cited the ability to abstract
+information about the merge from the clean summarization tables and
+simple visualizations. Three participants suggested that someone with a
+professional understanding of Gitk and the git commandline may be to
+extract the a good conceptual understanding from the DAG, and perform
+the summarization tasks from the commandline. One of these three
+participants said that they would prefer to have both tools available,
+as they are able to complement eachother.
+
 \section{Discussion}

### 2017-07-06 paper/lineval.tex
Discussion and Conclusion sections

@@
 \usepackage{amsmath}
-\usepackage{algorithm}
 \usepackage{balance}
@@
           colorlinks=false]{hyperref}
+
+\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
 
@@
         \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
+        \label{alg:original}
         \begin{algorithmic}
@@
 
+% TODO: Go back to the conceptual task results and make them stronger
+Overall, the results show fairly conclusive results; the DAG is unable
+to adequately provide users with an explanation of the changes being
+made in the repository, the DAG is not easily understood, and the DAG is
+unable to provide summarizations of the changes being made at a merge.
+Meanwhile, the \mt model is able to provide a better explanation of how
+a commit is integrated, along with better summarization capabilities. In
+this section, we discuss some observations made during the study,
+threats to the validity of the study, comments made by the participants
+during the study, modifications to the original algorithm to help
+generalize it for more repositories, and future work.
+
+\subsection{Observations}
+\label{sub:observations}
+
+% TODO: include info on merge tree sizes
+
+During the course of the study, we noticed a consistent theme among the
+participants. The most difficult challenge was identifying the master
+branch of the repository. Given that single-commit merge trees are the
+most common tree, and that they are relatively simple, we were
+interested in seeing if the participants would be able to identify the
+tree easily, but the results suggest that this was not the case. The
+issue primarily stemmed from not being able to identify the master
+branch. Most of the participants realized that they should not be going
+backward, toward the initial commit, very early on. Furthermore, many
+would identify the correct commit; however, they would start listing
+every merge along the master branch as being a merge in the tree, and
+including the commits from that merge. In the case of the single commit,
+simply identifying the master branch would most likely be sufficient in
+providing a clearer understanding of how a commit or flat tree is
+integrated into the master branch of the repository.
+
+While, in general, the participants of the study preferred using \tool
+to Gitk, there were some noticeable usability issues. The most prominent
+issue being the navigation within the tree. The participants expected
+that the tabs would update when they clicked on a node in the tree,
+instead of having to click the link to the commit after clicking the
+node. This was true in both the Reingold-Tilford tree and the Pack tree.
+The second major usability issue was the inconsistency between the list
+tree and the other tree visualizations. The list tree is rooted at the
+repository event that the user is current looking at, while the other
+tree visualizations show the entire tree, and use bright orange to
+highlight the current node.
+
+\subsection{Comments From a Release Manager}
+\label{sub:comments_from_a_release_manager}
+
+One of the participants of the study had worked as a release manager for
+more than three years, working with both SVN and CVS repositories, but
+had little experience with git.
+
+Contributors making merges need to understand more than just what merges
+a commit was collected into before reaching the repository of the
+contributor. It is also important to understand order that the related
+commits were made, as they tell the story of what the developer was
+thinking as they were writing the changes. The current merge-tree model
+does not order the commits, simply placing them randomly. This is the
+primary reason behind this participant's request to use both
+simultaneously. \tool is able to help with the aggregation of the
+information, and provide a better understanding of the next merge
+involved in integrating this commit, but the DAG in gitk provides the
+full story of the commit instead of hiding it behind a layer of
+abstraction.
+
+The comments from this participant were very insightful, and will assist
+in improving the model in the future.
+
+\subsection{Threats to Validity}
+\label{sub:threats}
+
+While git includes gitk and the git log provides a view of the DAG with
+\verb|git log --graph|, many of the participants in our study were
+unfamiliar with the DAG visualizations shown by these tools. A possible
+reason for this may be that unless a repository reaches a sufficient
+complexity, both in branching and number of commits, these tools are
+necessary for the normal operation of git. Furthermore, fear of the
+complexity caused by branching may cause personal and academic
+repositories to have relatively simple structures. While they were
+unfamiliar with the DAG structure, these users were also unfamiliar with
+the \mt model, and were still able to better understand and summarize
+the repository using \tool.
+
+Each question was evaluated independently of the results from the
+previous question in the summarization tasks. Had we modified the
+correct answer to fit the conceptual understanding of which commits were
+included in the tree, we may have witnessed different results. A
+modification to the study may include providing the participants with
+the correct set of commits prior to the summarization tasks, after the
+merge task set.
+
+\subsection{Generalization}
+\label{sub:generalization}
+
+Seeing that the \mt model is able to assist users with understanding how
+commits are integrated, we continue by looking at ways of improving the
+algorithm for faster operations, and making the conversion from DAG to
+\mt feasible over the internet for smaller repositories.
+
+The problem of finding the master branch is the most difficult, and is
+not possible without heuristic approaches. The master branch can be
+confounded by \textit{foxtrot} merges, which change the order of the
+parents in the parent list of a merge. Other than the ordering of the
+parent list, git has no other way to determine which commits are on the
+master branch of the repository. For the purpose of this discussion, we
+will assume that we can rely on the order of the parent list to be
+correct, and that foxtrots do not occur.
+
+\begin{algorithm}
+  \caption{Computing the generalized Merge Tree}
+  \label{alg:generalized}
+  \begin{algorithmic}[1]
+    \Function{Phase 1}{initial commit $root$} : updated tree
+    \State $depth \gets 0$
+    \State $Q \gets root$
+    \Do
+    \State $cur \gets Q.dequeue$
+    \State $parents \gets cur.parents$
+    \State $depth \gets depth + parents.length - 1$
+    \For{$index, parent \in parents$}
+    \If{$parent.depth$ is $undefined$}
+    \State $parent.depth \gets cur.depth + index$
+    \ElsIf {$cur.depth > parent.depth$}
+    \State $depth \gets depth - 1$
+    \ElsIf {$cur.depth < parent.depth$}
+    \State $depth \gets depth - 1$
+    \State $parent.depth \gets cur.depth$
+    \EndIf
+    \State $parent.children \gets cur$
+    \If{$parent$ has not been visited}
+    \State $Q \gets parent$
+    \State $parent.visited \gets True$
+    \EndIf
+    \EndFor
+    \doWhile{$Q$ not empty and $depth \ne 0$ }
+    \EndFunction
+
+    \Function{Phase 2}{initial commit $root$} : Merge tree
+    \State $tree.root \gets root$
+    \State $Q$ // Empty Queue
+
+    \State $parents \gets root.parents$
+    \State $parents \gets tail\ parents$
+    \For {$parent \in parents$}
+    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
+    \If {$realParent$ is $root$}
+    \State $root.children \gets parent$
+    \State $parent.parent \gets root$
+    \State $Q \gets parent$
+    \EndIf
+    \EndFor
+
+    \While{$Q$ not empty}
+    \State $cur \gets Q.dequeue$
+    \State $parents \gets cur.parents$
+    \For{$parent \in parents$}
+    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
+    \If {$realParent$ is $cur$}
+    \State $cur.children \gets parent$
+    \State $parent.parent \gets cur$
+    \State $Q \gets parent$
+    \EndIf
+
+    \EndFor
+    \EndWhile
+
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+The primary difference between the original algorithm and the
+generalization is the main data structure. Algorithm~\ref{alg:original}
+is centered around the use of the stack, making it akin to a depth-first
+traversal of the DAG.\@ The generalization in
+Algorithm~\ref{alg:generalized} is based on the queue, making it akin to
+the breadth-first traversal of the DAG.\@ While the final results of the
+two traversals should not be different, the runtime properties are
+different.
+
+We will begin with a simplified description of the original algorithm.
+We start at the master branch pointer. We will use the term master
+branch loosely here, meaning any branch that has been labeled in some
+way as being the main branch, this branch may have a different name, but
+standard naming conventions call it the master branch. Starting from the
+master branch label, we traverse the entire master branch to the initial
+commit, recording the commit hash of every repository event encountered.
+
+The second phase iterates over every commit, computing the shortest
+distance to the master branch. For this reason, the entire repository
+must be available to the algorithm. This is a dynamic programming
+solution, centered around the call stack. The next node on the shortest
+path to the master branch will become the parent of this node in the
+tree.
+
+If this algorithm is to be generalized into a dynamic, web-based, tool,
+it may not be feasible to download every commit in the repository in a
+reasonable period of time. We instead look toward a breadth-first
+approach at limiting the number of repository events that must be
+downloaded.
+
+Focusing on the generalized algorithm~\ref{alg:generalized}, the
+algorithm is broken into two phases. The first phase is where the
+improvements take place. The first phase has three main goals, the first
+is to track the master branch. The second goal labels the depth of each
+repository event that is traverses. The final goal is to record the
+children of each commit, reversing the DAG.\@ The first phase also
+begins the process of reversing the DAG, by providing the candidate
+commits that could possibly be the real parent to this node. The first
+phase can terminate early, if it reaches a point where all branches that
+it is following terminate. The depth metric on line 7 indicates how many
+branches are currently being traversed. In line 11 through 15, we
+decrement the depth, as this is the point where a branch occurs. The
+changing of the depth occurs if there are more events on the same branch
+than between the merge and branch point of a separate branch.
+
+\begin{table}[htpb]
+  \centering
+  \caption{Example traversal of the example DAG in
+    Figure~\ref{fig:repoDAG} starting at merge 11.}
+  \label{tab:ex_traversal}
+  \begin{tabular}{cccc|cc}
+    \toprule
+    Current & Parents & Depth & Q       & Visited & Children\\\midrule
+    -       & -       & 0     & 11      & 11:0    & \\
+    11      & 4,9     & 1     & 4,9     & 4:0     & 11, 5\\
+    4       & 1       & 1     & 9,1     & 9:1     & 11\\
+    9       & 7,6,8   & 3     & 1,7,6,8 & 7:1     & 9, 8\\
+    1       &         & 3     & 7,6,8   & 6:2     & 9\\
+    7       & 5       & 3     & 6,8,5   & 8:3     & 9\\
+    6       & 5       & 2     & 8,5     & 5:1     & 7, 6\\
+    8       & 7       & 1     & 5       & 2:1     & 5, 3\\
+    5       & 4,2,3   & 2     & 2,3     & 3:2     & 5\\
+    2       & 1       & 1     & 3       &         & \\
+    3       & 2       & 0     &         &         & \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+In Table~\ref{tab:ex_traversal}, we follow the algorithm starting at
+merge commit 11 from the example in Figure~\ref{fig:repoDAG}, recording
+the current node, the parents of the current node, the depth after
+updating, and the queue structure after updating. The visited column is
+not dependant on the current step, but keeps track of the nodes that
+have been visited, and the depth at that node. When we reverse the DAG,
+the children in this table will become the candidates for being the
+parent of the corresponding node. For example, the parent of node 7 may
+be resolved to either node 9 or node 8. In this case, the shortest path
+to the master branch is through node 9.
+
+The terminology surrounding the parent/child relationship gets very
+confusing at the change between phase one and two because this is where
+the roles switch. In phase 1, we are working with the DAG; in phase 2,
+we are working with a structure that is neither a tree nor a DAG, but
+can only be described as a directed graph.
+
+The goal of the second phase is to resolve the tree from the directed
+graph. We will refer to parents in the same sense as we would refer to
+them in the tree form of the repository for the remainder of this
+section. We have gathered the depths and possible parents (line 17), but
+we need to remove the extra parents. Extra parents are culled in line
+42 of the algorithm. Of the possible parents, we take the parent that is
+closest to the root, but not beyond. This forces the real parent to
+either be deeper in the tree, or on the same branch. Line 30 to 37 are
+for bootstrapping the queue, ensuring that we are not starting on an
+empty queue.
+
+The construction of the tree becomes trivial now that we are able to
+determine the parents and children for each node. Lines 44 and 45
+perform this construction, adding the DAG parent to the tree children of the
+current node, and the DAG parent's parent becomes the current node,
+which was the DAG parent's child. We continue following the path until
+we hit the root node, merging the commit into the master branch.
+
+This algorithm relies heavily on the order of parents in the parent
+list, but is able to trim the number of nodes that must be downloaded.
+Furthermore, the resulting tree is able to show the order of commits,
+telling the story of how a developer was working, before the set of
+commits was merged. We developed a small bitbucket plugin to test the
+design and determine whether it was capable of producing merge trees,
+shown in Figure~\ref{fig:b_plugin}.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/plugin.png}
+  \caption{Screen shot of the Bitbucket plugin showing the online tree
+    computation of the repository shown in~\ref{fig:repoDAG}.}
+  \label{fig:b_plugin}
+\end{figure}
+
 \section{Conclusion}
@@
 
+In this paper, we develop a tree-based model designed to provide git
+users with a better explanation of the events occurring in a git
+repository. The \mt model is rooted at the merge into the master branch
+of a repository, where the leaves are the individual commits.
+
+We implement a tool, \tool, which allows a user to filter the commits
+with a simple search interface. Once the user finds the commit that they
+are interested in, they are provided simple aggregated tables showing
+the files that were modified, the authors that made the modifications,
+among other information. They are also provided with three
+visualizations, a list tree, a pack tree, and a Reingold-Tilford tree
+view. The list tree enables easy searching for textual features, while
+the pack tree is good for visualizing the structure of large merge
+trees, and the Reingold-Tilford tree is better at visualizing the
+structure of small merge trees.
+
+To ensure that our model is necessary, we evaluate the ability of users
+to perform simple conceptual tasks, finding how a commit was merged into
+the master branch of the Linux repository, and determining other commits
+that were merged with the given merge. Finding that the participants in
+our study were unable to perform this task, we compared the ability of
+the participants to summarize key pieces of information aggregated in
+these groups of commits. We asked for information about what files were
+modified, the authors making the changes, among other tasks. Finally, we
+asked the participants about their preference in tools, and which
+aspects of each they preferred. We found that the participants in our
+study were able to correctly identify the correct answer more quickly
+using the \mt model than they were able to with the DAG.
+
+Given that the model provides users with a better understanding of
+events in the repository, we describe some issues with identifying the
+master branch of the repository and a more generalized approach to
+constructing merge trees. We implement the generalized approach as a
+bitbucket plugin to verify that it is able to quickly produce results
+online, unlike in the implementation of \tool.
+
+% TODO: You can tell that I am tired. This conclusion will need to be re-worked.
+
+\balance
+
 % \nocite{*}

### 2017-07-06 paper/lineval.tex
Breaking introduction into sections
Breaking the massive introduction into sections to match original paper

@@
 \ifdraft
-        \documentclass[conference, draftclsnofoot, draft]{IEEEtran}
+        \documentclass[draft]{IEEEtran}
         \def\baselinestretch{1}
-        \setlength{\marginparwidth}{2cm}
+        % \setlength{\marginparwidth}{1.5cm}
 \else
@@
 
-\usepackage[hyphens]{url} \urlstyle{same}
+\usepackage[hyphens]{url}
+\urlstyle{same}
 
@@
 
-Git is a distributed version control system that is used in many
-open-source and closed-source projects. Git enables many developers to
-write code simultaneously, working on various components known as
-branches. Git uses a directed acyclic graph model (DAG) for storing the
-necessary information regarding the state of the code at any point in
-time. Git refers to nodes of the DAG as commits, not distinguishing
-between the code-carrying commits and the structural commits that bring
-two branches together. In this paper, we will refer to the individual
-commits as being commits, the structural merge commits as merges, and
-refereeing to all as repository events.
-
 Between 50k and 70k commits are added to the Linux kernel per version
@@
 
-% DAG model
+Git is a distributed version control system that is used in many
+open-source and closed-source projects. Git enables many developers to
+write code simultaneously, working on various components known as
+branches. Git uses a directed acyclic graph model (DAG) for storing the
+necessary information regarding the state of the code at any point in
+time. Git refers to nodes of the DAG as commits, not distinguishing
+between the code-carrying commits and the structural commits that bring
+two branches together. In this paper, we will refer to the individual
+commits as being commits, the structural merge commits as merges, and
+refereeing to all as repository events.
 
@@
 
-% Merge tree model
+One major challenge with visualizing the arrival of commits to a repository is that
+Git does not store the date that a commit was merged into another branch, including
+the master branch. To complicate the problem, the DAG only has references to the
+ancestors of a commit (a model necessary for the operation of Git), but maintainers
+would prefer knowing the path a commit followed to reach the master repository.
+Tracing a path that any commit followed to the master repository would imply that
+for any given merge, it would be possible to know which commits were merged. A user
+could inspect the commits that arrived into the master branch within a given
+time-frame by checking which commits were merged during that time-frame.
+
+This paper makes two contributions; first, we describe a method of converting the
+DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
+that represents the path used by a commit to reach the master branch; second, we
+present a method to inspect and visualize the history of merges in the Linux project
+using the merge-tree model.
+
+These methods and visualizations are implemented in a web-based tool called
+\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
+Our visualizations and tool provide information about the location of any given
+commit or merge in its respective merge-tree, the files edited, the modules edited,
+and the commit message. \tool allows users to apply various filters, including the
+release version, along with a keyword or phrase from the log preview, the name of
+the author, or the commit ID. The user can request all merges made by Linus that
+contain a commit or inner merge that matches the search query, or just the commits
+and merges that match the query.
+
+\section{\mt Model}
+\label{sec:mt_model}
 
@@
 
-% Evaluation
-
-In order to determine the effectiveness of the merge-tree model on
-comprehension and summarization of git repositories, we conducted a
-controlled user study on 12 participants. The user study is a
-mixed-methods study, with focus on the quantitative aspects in order to
-capture how effective the \mt model, if it is, at improving the
-correctness, accuracy, and time performance metrics of users when
-performing summarization tasks. We use the qualitative questions as a
-means of determine the preference of the participants, and what aspects
-they preferred from each tool.
-
-We formulate three research questions that this study is attempting to
-answer;
-
-\begin{RQ}
-  \item
-
-    Is the DAG able to provide users with an understanding of the events
-    in a repository?
-
-  \item
-
-    Do \mt make a difference in the ability to summarize basic
-    aggregated information?
-
-  \item
-
-    Do users prefer the \mt model to the DAG for summarization and
-    comprehension tasks?
-
-\end{RQ}
-
-% Implementation
+\section{Design and Implementation}
+\label{sec:design_and_implementation}
 
@@
 
-\section{Results}
+\subsection{Results}
 \label{sec:results}

### 2017-07-06 userstudy/analysis/plots.R
Changing separation of plots, to fit them better

@@ -5,7 +5,8 @@ library(RSQLite)
 library(fitdistrplus)
 
 pdf("figures/results.pdf", width=17, height=12)
-par(mfrow=c(3, 7))
+par(mfrow=c(3, 7), mai=c(0.4, 0.3, 0.3, 0.2))
+print(par('mai'))
 
 colorScheme <- c("#3366CC", "#DC3912", "#FF9900", "#109618", "#990099", "#3B3EAC")
 colorSchemeTrans <- c("#3366CC99", "#DC391299", "#FF990099", "#10961899", "#99009999", "#3B3EAC99")

### 2017-07-06 paper/introduction.tex
Adding most recent version of introduction
This is the mongo introduction with everything in it

@@
 Git is a distributed version control system that is used in many
-opensource and closedsource projects. Git enables many developers to
+open-source and closed-source projects. Git enables many developers to
 write code simultaneously, working on various components known as
@@
 necessary information regarding the state of the code at any point in
-time. Git refers to nodes of the DAG as comits, not distinguishing
+time. Git refers to nodes of the DAG as commits, not distinguishing
 between the code-carrying commits and the structural commits that bring
@@
 commits as being commits, the structural merge commits as merges, and
-refereing to all as repository events.
+refereeing to all as repository events.
 
@@
     the DAG, the top center pane showing the authors, and top right pane
-    showing the commit dates. The bottom-leftpane shows the full
+    showing the commit dates. The bottom-left pane shows the full
     commit message and diff of the selected commit, and the bottom-right
@@
 \end{figure}
+
+% DAG model
 
@@
 systems to generate a visualization; repositories stored on GitHub will
-have a DAG visualzation to show the branches and forks of a
-repositority, but is unable to display the visualization for projects at
+have a DAG visualization to show the branches and forks of a
+repository, but is unable to display the visualization for projects at
 the scale of Linux or even the Git Git
@@
 
-This paper investigates the construction, implementation, user-based
-evaluation, and generalization of a new model, the \mt model. We
-begin with a description and algorithm for converting the DAG into a set
-of merge trees. We continue by providing some details about the
-web-based tool, called \tool\footnote{\tool is currently available at
-  \url{http://li.turingmachine.org}}, implemented to test the
-feasibility and potential capabilities of the \mt model. We
-continue with a user study, to evaluate first, if the DAG is capable of
-providing the users with a conceptual understanding of the repository
-events that are related to a given commit, and second, determine if
-there is a difference between the DAG and \mt model in performing
-summarization tasks.
+% Merge tree model
 
-\subsection{Directed Acyclic Graph Model}
-\label{sec:directed_acyclic_graph_model}
+We propose a new model, built from the DAG, with the purpose of
+providing a clearer conceptual image of how a commit is merged, other
+commits that are merged with a commit, and a summarization of the entire
+group of commits that are merged with a commit. We call our model the
+\mt model. The model inverts the DAG, so that reference point from
+parent to child instead of from child to parent. This enables the
+aggregation of commit metrics at merges. Given a merge, we are able to
+recursively traverse the children of the merge, collecting the
+statistics at each commit. The commits become the leaves of the tree,
+and internal nodes are merges that do not merge into the master branch
+of the repository. We terminate the further traversal of an edge under
+two conditions; first, if the next event is on the master branch of the
+repository, second, if the next event has a shorter path into the master
+branch of the repository. As a consequence, commits will always take the
+shortest path, in terms of commit time, to the master branch.
 
-Git uses the DAG as the internal model for representing repositories.
-This model functions perfectly, and is necessary, for the internal
-functionality of Git. The DAG model allows for the reconstruction of
-files and the working directory by recursively applying the changes, or
-the diff, made in each repository event that is a parent of the current
-event. We see an example DAG repository Figure~\ref{fig:repoDAG}.
+A short example: assume the commits represented in
+Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
+In this case, commits are performed in various repositories and
+branches. The DAG representation of the commits is shown in
+Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
+the master branch and the repository that the master branch is part of.
+The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
+Note that the direction of the edges of the DAG have been inverted,
+instead of pointing from the child to the ancestors, it points from the
+ancestor to its successors, forming a path to the master branch. Also
+note that the DAG has been simplified, showing only a single edge on the
+path to master for any commit.
 
-Each repository event contains an ordered list of one or more parents,
-except the initial commit, which has no parents. Commits have a single
-parent, while merges will have at least two parents, the current branch,
-and the branch being merged into the current branch. Git allows for the
-merging of multiple branches simultaneously in octopus merges, where the
-order of the parents represents the order that the branches were
-specified for being merged.
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
+    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
 
-The repository events in the DAG are immutable; once an event is
-created, it can't be changed. Git allows operations to alter commits and
-re-order them, but this will create a new commit with a new commit hash.
-This property makes a repository event unable to record the traversal of
-merges into the master branch of the repository. A commit cannot have
-references in the direction of the merge that merges the commit, as this
-reference will change, requiring the commit to be changed. While this
-provides some challenge, git eliminates this issue of the DAG with the
-\verb|git log --children| command, which traverses the DAG and inverts
-the edges to provide the reference from parent to child. This
-terminology will prove to be confusing through the rest of the paper,
-when we are referring to the \mt form, the parent is the next node
-toward the root. When referring to the DAG, the parent relationship is
-in the opposite direction, from the root toward the branch-point.
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
 
-Many current visualization systems base the visualization on the DAG,
-because it is made readily available by git. The primary issue being
-that the DAG was not designed with visualization and summarization tasks
-in mind. Due to the direction of the links, the DAG model is unable to
-provide a summarization of the contents of a merge; the DAG is unable to
-show which branch is the master branch, and is unable to filter the
-irrelevant information from the view.
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
 
-\subsection{\mt Model}
-\label{sub:merge_tree_model}
+    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
+    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
+    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
+    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
 
-We use the \mt model to address these issues in the DAG\@. We
-will describe the \mt model in detail in
-Section~\ref{sec:mergetree}. The \mt reverses the direction of
-the references of the DAG.\@ Starting with each merge into the master
-branch, the algorithm recursively aggregates the statistics of each
-commit. The tree is rooted at the merge into the master branch, and the
-commits are the leaves.
+        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
+    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
+  \end{tikzpicture}
+  \caption{Example of a sequence of events performed in different
+    repositories. The horizontal axis represents time. Each horizontal
+    section represents a different branch and/or repository. Each commit
+    points to its ancestor.}
+  \label{fig:repoEvents}
+%\vspace{-3mm}
+\end{figure}
 
-\subsection{Implementation}
-\label{sub:implementation}
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+    \node[state] (1) {1};
+    \node[state, above right= of 1] (2) {2};
+    \node[state, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, right= 2cm of 1] (4) {4};
+    \node[state, above right=of 4] (5) {5};
+    \node[state, above right=of 5] (6) {6};
+    \node[state, right=of 5] (7) {7};
+    \node[state, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, right= 2cm of 7] (9) {9};
+    \node[state, right=of 9] (10) {10};
+    \node[state, below right=of 9] (11) {11};
+    \node[state, below right=of 10] (12) {12};
+
+    \draw (12) edge[-stealth] (11) edge[-stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[-stealth] (9);
+    \draw (10) edge[-stealth] (9);
+    \draw (9) edge[-stealth] (8) edge[-stealth] (6)
+              edge[-stealth] (7);
+    \draw (8) edge[-stealth] (7);
+    \draw (7) edge[-stealth] (5);
+    \draw (6) edge[-stealth] (5);
+    \draw (5) edge[-stealth] (3) edge[-stealth] (2)
+              edge[-stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[-stealth] (2);
+    \draw (2) edge[-stealth] (1);
+  \end{tikzpicture}
+  \caption{DAG representation of the commits represented in
+    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
+    information about which repository a commit comes from. Furthermore,
+    the DAG does not distinguish the master branch from other branches.}
+  \label{fig:repoDAG}
+\vspace{-3mm}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+
+    \draw[chartblue]
+      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
+    \draw (2) edge[chartyellow, -stealth] (5)
+          (3) edge[chartmagenta, -stealth](5)
+          (5) edge[chartyellow, -stealth] (7)
+          (7) edge[chartyellow, -stealth] (9)
+          (6) edge[chartred, -stealth] (9)
+          (8) edge[chartmagenta, -stealth] (9)
+          (9) edge[chartyellow, -stealth] (11)
+          (10) edge[chartyellow, -stealth] (12);
+
+  \end{tikzpicture}
+  \caption{\mt view of the commits  represented in
+    Figure~\ref{fig:repoEvents} showing the path they followed to reach
+    the master branch. In this model the successors of each commit
+    represents the path followed by that commit to reach the master
+    branch.}
+  \label{fig:repoTree}
+\vspace{-3mm}
+\end{figure}
+
+Computing the \mt from a DAG for any repository may not be
+possible; however, certain features of the development process of Linux
+make it feasible to compute the \mt for the Linux repository.
+First, the master branch of Linux is maintained by Linus Torvalds, and
+only Linus has write access to it. We have verified this assertion in
+previous research~\cite{German2015}. We have developed a heuristic that
+is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
+identifies the commits made directly to the master branch; whereafter it
+recursively determines the shortest path (in terms of time), using the
+DAG, from each commit to the master branch using the inverted DAG.
+
+\begin{algorithm}
+        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
+        \label{alg:original}
+        \begin{algorithmic}
+                \Function{ComputeMergeTree}{DAG}: tree
+                \State {\# Compute the tree from the DAG of Linux repository.}
+                \State {\# Returns $Tree$, a graph containing every commit }
+                \State {\# in DAG with the path it followed to master.}
+
+                \State $head \gets \textit{Head of master of git repository}$
+                \State $master \gets \textit{traverse DAG from head using }$
+                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
+                \State $nodes(Tree) \gets nodes(DAG)$
+                \State \Function{distance2Master}{cid} : seconds
+                \State {\# Helper function}
+                \State {\# Recursively compute shortest distance to master}
+                \State {\# setting cid's successor (next) in its way to master.}
+                \State {\# This function should be memoized. Otherwise it}
+                \State {\# would run in exponential time.}
+                \If {\textit{cid in master}}
+                \State \Return 0
+                \EndIf
+                \State    $d \gets 	\infty$
+                \State {\# Traverse the inverted DAG}
+                \For{$c \in children(cid, DAG)$}
+                \If {$c \in master$}
+                \State $d_1 \gets commitTime(c)-commitTime(cid)$
+                \Else
+                \State {$d_1 \gets distance2Master(c)$}
+                \EndIf
+                \If {$d_1 < d $}
+                \State $next \gets c$
+                \State  $d \gets d_1$
+                \EndIf
+                \EndFor
+                \State {\# $c$ is the commit that follows $cid$}
+                \State {\# in its way to master}
+                \State add edge $(cid, next)$ to $Tree$
+                \State \Return $d$
+                \EndFunction
+
+                \State {\# Compute the distance for each commit}
+                \State {\# discarding result}
+                \For{$c \in nodes(DAG)$}
+                \State $distance2Master(c)$
+                \EndFor
+                \State \Return $Tree$
+                \EndFunction
+        \end{algorithmic}
+\end{algorithm}
+
+We evaluate the performance of the heuristic using the merge log made by
+Linus Torvalds, the original creator of Linux, and primary merger of all
+other repository events into the master branch of the repository. In
+Figure~\ref{fig:sampleMerge}, we see an example of the merge log.
+
+\begin{figure}[htbp]
+  \centering
+  {\fontsize{7}{9}
+    \begin{verbatim}
+Merge: 8cbd84f fd8aa2c
+Author: Linus Torvalds <torvalds@linux-foundation.org>
+Date:   Tue Aug 10 15:38:19 2010 -0700
+
+Merge branch 'for-linus' of git://neil.brown.name/md
+
+* 'for-linus' of git://neil.brown.name/md: (24 commits)
+md: clean up do_md_stop
+[... edited for the sake of space]
+md: split out md_rdev_init
+md: be more careful setting MD_CHANGE_CLEAN
+md/raid5: ensure we create a unique name for kmem_cache...
+...
+    \end{verbatim}}\vspace{-5mm}
+  \caption{Example of how merges record a subset of commits being
+    merged. The commit only shows the first 20 one-line summaries
+    messages for the 24 non-merge commits it merged. The ending
+    ``\ldots'' is part of the log and represents that other
+    commits were merged.}
+  \label{fig:sampleMerge}
+\end{figure}
+
+We use this information to evaluate the accuracy of the \mt model
+extracted from the DAG\@. The method we followed started with the
+extraction of the commit history up to July 20, 2016. We computed the
+merge tree of every commit until then. Since Linus Torvalds mostly does
+merging directly into master, we assumed that every merge by him is the
+root of a \mt. As described above, the log of a merge-commit usually
+contains the number of commits in the merge the first 20 summaries of
+commits being merged. We extracted merges by Linus Torvalds using the
+command \mycode{log --merges --author='Torvalds} and compared the number
+of commits according to the log with the number of commits in the  \mt
+rooted in this commit. We also used the summaries of the commits found
+in the merge (not necessarily all---see above) to make sure those
+commits were in their corresponding merge tree. For example, for the
+merge in Figure~\ref{fig:sampleMerge} we would expect that the merge
+tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
+summaries corresponds to commits in that \mt. We also inspected those
+with differences to make sure they were true errors.  The results can be
+summarized as follows:
+
+\begin{itemize}
+
+  \item
+
+    Five merges were false-errors because their logs did not contain
+    accurate information (were probably edited by hand). For example in
+    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
+    empty), in \mycode {c55d267\ldots} the summaries were reordered.
+
+  \item
+
+    The heuristic correctly identified that 79 of Linus merges (between
+    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
+    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
+    commits.
+
+  \item
+
+    The heuristic worked perfectly until Sept 4, 2007, the earliest date
+    that it could be verified.  Before this date, and until Dec 12,
+    2006, merges did not include a summary of the commits they included,
+    hence making it impossible to verify; during this period, however,
+    we correctly identified the merges by Linus into master.
+
+  \item
+
+    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
+    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
+    which confounded the true master branch of a repository (see
+    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
+    for a description of the issue).
+
+\end{itemize}
+
+In summary, of the merges after Sept. 4, 2007, our heuristic was correct
+in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
+12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
+2007).
+
+% Evaluation
+
+In order to determine the effectiveness of the merge-tree model on
+comprehension and summarization of git repositories, we conducted a
+controlled user study on 12 participants. The user study is a
+mixed-methods study, with focus on the quantitative aspects in order to
+capture how effective the \mt model, if it is, at improving the
+correctness, accuracy, and time performance metrics of users when
+performing summarization tasks. We use the qualitative questions as a
+means of determine the preference of the participants, and what aspects
+they preferred from each tool.
+
+We formulate three research questions that this study is attempting to
+answer;
+
+\begin{RQ}
+  \item
+
+    Is the DAG able to provide users with an understanding of the events
+    in a repository?
+
+  \item
+
+    Do \mt make a difference in the ability to summarize basic
+    aggregated information?
+
+  \item
+
+    Do users prefer the \mt model to the DAG for summarization and
+    comprehension tasks?
+
+\end{RQ}
+
+% Implementation
 
@@
 Figure~\ref{fig:commit_2}.
-
-\section{\mt Model}
-\label{sec:mergetree}
-
-The \mt model inverts the DAG, so that the reference point from
-parent to child instead of child to parent. This is to enable
-aggregation at merges. Given a merge, we are able to recursively
-traverse all children of that merge, collecting the statistics at each
-commit. The inverted DAG is then traversed starting at each merge into
-the master branch. The merges into the master branch is the root of each
-tree. The commits are the leaves of the tree. Starting at the merge into
-the master branch, we recursively traverse each child of the current
-node until we reach the branch point from the master branch, or a commit
-that has a possible root merge that is closer in commit time to the
-commit. As a consequence, commits will always take the shortest path, in
-terms of time, to the master branch.
-
-A short example: assume the commits represented in
-Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
-In this case, commits are performed in various repositories and
-branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
-the master branch and the repository that the master branch is part of.
-The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
-Note that the direction of the edges of the DAG have been inverted,
-instead of pointing from the child to the ancestors, it points from the
-ancestor to its successors, forming a path to the master branch. Also
-note that the DAG has been simplified, showing only a single edge on the
-path to master for any commit.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
-    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-
-    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
-    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
-    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
-    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
-
-        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
-    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
-  \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
-  \label{fig:repoEvents}
-%\vspace{-3mm}
-\end{figure}
-
-% TODO: make everything one color
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
-    information about which repository a commit comes from. Furthermore,
-    the DAG does not distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (2) edge[chartyellow, -stealth] (5)
-          (3) edge[chartmagenta, -stealth](5)
-          (5) edge[chartyellow, -stealth] (7)
-          (7) edge[chartyellow, -stealth] (9)
-          (6) edge[chartred, -stealth] (9)
-          (8) edge[chartmagenta, -stealth] (9)
-          (9) edge[chartyellow, -stealth] (11)
-          (10) edge[chartyellow, -stealth] (12);
-
-  \end{tikzpicture}
-  \caption{\mt view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
-\vspace{-3mm}
-\end{figure}
-
-\subsection{Computing the \mt from the DAG of Linux}
-
-Computing the \mt from a DAG for any repository may not be
-possible; however, certain features of the development process of Linux
-make it feasible to compute the \mt for the Linux repository.
-First, the master branch of Linux is maintained by Linus Torvalds, and
-only Linus has write access to it. We have verified this assertion in
-previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
-identifies the commits made directly to the master branch; whereafter it
-recursively determines the shortest path (in terms of time), using the
-DAG, from each commit to the master branch using the inverted DAG.
-
-\begin{algorithm}
-        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
-        \begin{algorithmic}
-                \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
-                \State $head \gets \textit{Head of master of git repository}$
-                \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
-                \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
-                \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
-                \If {\textit{cid in master}}
-                \State \Return 0
-                \EndIf
-                \State    $d \gets 	\infty$
-                \State {\# Traverse the inverted DAG}
-                \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
-                \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
-                \EndIf
-                \EndFor
-                \State {\# $c$ is the commit that follows $cid$}
-                \State {\# in its way to master}
-                \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
-                \EndFunction
-
-                \State {\# Compute the distance for each commit}
-                \State {\# discarding result}
-                \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
-                \EndFor
-                \State \Return $Tree$
-                \EndFunction
-        \end{algorithmic}
-\end{algorithm}
-
-\subsection{Evaluation}
-
-Merges that do not have merge conflicts provide information to verify
-this heuristic. If a merge does not contain a conflict, it records a
-summary of the commits that it merges. See Figure~\ref{fig:sampleMerge}
-for an example. This summary contains a list of the first 20 non-merge
-commits in the merge, including their one-line log description, the full
-logs of the merge commits that merge this subset, and the total number
-of non-merge commits in the merge.
-
-\begin{figure}[htbp]
-  \centering
-  {\fontsize{7}{9}
-    \begin{verbatim}
-Merge: 8cbd84f fd8aa2c
-Author: Linus Torvalds <torvalds@linux-foundation.org>
-Date:   Tue Aug 10 15:38:19 2010 -0700
-
-Merge branch 'for-linus' of git://neil.brown.name/md
-
-* 'for-linus' of git://neil.brown.name/md: (24 commits)
-md: clean up do_md_stop
-[... edited for the sake of space]
-md: split out md_rdev_init
-md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache...
-...
-    \end{verbatim}}\vspace{-5mm}
-  \caption{Example of how merges record a subset of commits being
-    merged. The commit only shows the first 20 one-line summaries
-    messages for the 24 non-merge commits it merged. The ending
-    ``\ldots'' is part of the log and represents that other
-    commits were merged.}
-  \label{fig:sampleMerge}
-\end{figure}
-
-We used this information to evaluate the accuracy of the \mt
-model extracted from the DAG\@. The method we followed started with the
-extraction of the commit history up to July 20, 2016. We computed the
-merge tree of every commit until then. Since Linus Torvalds mostly does
-merging directly into master, we assumed that every merge by him is the
-root of a \mt. As described above, the log of a merge-commit
-usually contains the number of commits in the merge the first 20
-summaries of commits being merged. We extracted merges by Linus Torvalds
-using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits according to the log with the number of commits in
-the  \mt rooted in this commit. We also used the summaries of the
-commits found in the merge (not necessarily all---see above) to make
-sure those commits were in their corresponding merge tree. For example,
-for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
-merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
-one-line summaries corresponds to commits in that \mt. We also
-inspected those with differences to make sure they were true errors.
-The results can be summarized as follows:
-
-\begin{itemize}
-
-  \item
-
-    Five merges were false-errors because their logs did not contain
-    accurate information (were probably edited by hand). For example in
-    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
-    empty), in \mycode {c55d267\ldots} the summaries were reordered.
-
-  \item
-
-    The heuristic correctly identified that 79 of Linus merges (between
-    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
-    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
-    commits.
-
-  \item
-
-    The heuristic worked perfectly until Sept 4, 2007, the earliest date
-    that it could be verified.  Before this date, and until Dec 12,
-    2006, merges did not include a summary of the commits they included,
-    hence making it impossible to verify; during this period, however,
-    we correctly identified the merges by Linus into master.
-
-  \item
-
-    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
-    which confounded the true master branch of a repository (see
-    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
-    for a description of the issue).
-
-\end{itemize}
-
-In summary, of the merges after Sept. 4, 2007, our heuristic was correct
-in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
-12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
-2007).

### 2017-07-07 paper/lineval.tex
many comments

@@
 \usepackage[utf8]{inputenc}
-
 \usepackage[hyphens]{url}
@@
 %% Table Support
+\usepackage{array}
 \usepackage{dcolumn}
@@
 
-%% Tikz
 \usetikzlibrary{positioning, arrows}
@@
 
+\dmg{start by describing what you mean by user study... to evaluate linvis we performed a user study...}
+
+\dmg{and you are evaluating the visualizations, NOT the DAG}
+
 We use a controlled mixed-methods users study in order to evaluate the
@@
 implementing the user study.
+\dmg{no, you describe the study, which includes the methods}
+
+\dmg{Start by giving an overall overview of the study: the study consists of a set of tasks that users are expected to
+  perform... do a top down description of what users will do
+}
 
@@
 \label{sub:questions}
+
+\dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}
 
@@
 
+\dmg{avoid editorializing your text: remove purely}
+
+\dmg{watch for tense... we don't beging... the tasks begin... or better, the first tasks...}
+
 The questions and tasks are broken down into four groups. The first two
@@
 
-The third task set is purely qualitative, to gain better understanding
+The third task set is purely qualitative, \dmg{and are their goal is ... don't imply content in your sentences... be
+  precise; this applies to the paragraph before} to gain better understanding
 of whether participants preferred working with the \mt model in \tool,
@@
 \textbf{Conceptual Tasks}
+
+\dmg{given the choice, not allowed. 
+  I don't understand this section.
+  You need to be more top-down in your description. you need to provide context on what the
+  users are expected to do, and why. So what are the tasks? You never explain.
+what is the goal of the tasks? how many? 
+what were the tasks? what is expected from each? you need to answer these questions, not necessarily in the same order.
+also, remember this is a comparative study, so I presume the users were asked to do both with linvis and without it
+(choice of gitk/git command line)
+}
 
@@
 
+\dmg{I don't like the present tense, but that is your option}
+
 We provide the participant with 10 minutes per commit to complete the
@@
 
+\dmg{again, you are going into details first. First the tasks --see above--
+  including their grouping.
+  then at the end, describe the randomization and why it is done (which you did)
+}
+
 The summarization tasks (Table~\ref{tab:summarization_tasks}) are
-grouped into task sets by theme. The tasks within the task set are
+grouped into \dmg{how many?} set of tasks according to their theme. The tasks within a given task set are
 shuffled, and the task sets are shuffled, to ensure that similar
@@
 Finally, we ask the user for their opinions on the tools and for some
-information about their level of experience with git. This portion of
-the study is qualitative and does not rely on either of the tool,
-instead attempting to determine which model provided the user with a
-better experience.
-
-We do not shuffle the questions in this set of questions, as this should
-not have an impact on the opinions of the participant at this point.
-These questions are designed to provide us with a better understanding
+information about their level of experience with git.
+\dmg{be more precise: understanding or preference is too broad. What you want to know 
+answers to specific questions regarding how the tools compare and which ones they would prefer}
+The first three questions are intended to provide us with a better understanding
 of preference, and insight on why, the participant had that preference.  
-
-We ask the following questions in this order;
+The last three questions in this section are aimed at getting a better
+understanding of the demographic of our participants.
 
@@
 
-The last three questions in this section are aimed at getting a better
-understanding of the demographic of our participants.
-
 \textbf{Procedure}
+
+\dmg{tell me first why you need to create a script, so I know what you do something:
+we did random assignment of tasks (explain what you did) and randomized the order in which the tasks were performed.
+ Then explain why you need a script}
+
+\dmg{it does not randomize the order the tools. It randomizes which tool is going to be used on each tasks.
+be precise with your language}
+
+\dmg{what you explained how commits relate to tasks and how they get assigned? think as if you are explaining this to
+  somebody who has to do this: you are the designer, explaining the why and the how }
 
@@
 
+\dmg{see my comments in previous section. it is ok to explain them here, but you need to explain what they are briefly
+  in the previous section}
+
 We chose two commits for use in the study. The order that the commits
@@
 
+\dmg{this place does not require a description of the database. This requires just to say
+the commits/merge set (dates and numbers are ok) only}
 We use the \tool database, containing commits and merge-tree information
@@
 between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
-  database even more?} We chose the commits based on tree sizes. We
+  database even more?}
+\dmg{again, i have problems with your writing: ``we wanted to use in the study commits of different size, to determine
+  if the size had an impact in the completion of the tasks. We chose commits of three different sizes...}
+We chose the commits based on tree sizes. We
+\dmg{we didn't find: this is a fact, not a discovery; what is a majority? 50\%+1? be precise}
+\dmg{drop the word inner, and do it in order: 25\% of trees have only one commit, 50\%.., 75\%...}
 found that a majority of the trees contain at most seven inner commits,
@@
 largest tree contained 7217 nodes. We chose to work with the trees in
+\dmg{if you are going to use the words quartile, just use them before in your description eg: the distribution of the
+  sizes of the trees is x, y merges, whose size corresponded to the 25 ,50 quartiles, simple and precisee}
 the first and second quartile, as merge trees of sizes between one and
@@
 
+\dmg{remove from here, what is here? this is not a road trip :)}
 From here, we selected one tree from the trees of a single commit at
 random. Selecting a commit from that tree is trivial, as there is only
-one to choose.
+one to choose.\dmg{remove the obvious, instead concentrate on why the others are not trivial}
 
 We selected one tree of size seven at random from all trees of size
-seven. We placed a restriction the trees, the selected tree must include
+seven\dmg{how many where there?}. We placed a restriction the trees, the selected tree must include
 at least one internal merge to increase the complexity of the trees
-tested. After randomly selecting a tree, we chose in internal commit at
-random, using the \verb|random.choice| function in python 3.6.1.
+tested. For each randomly selecting a tree, we chose one of its internal commits at
+random (using the \verb|random.choice| function in python 3.6.1).
+
+\dmg{question: so you chose only one commit of each size, and the randomization was which tool to use
+for each question? or did they have to do the same task for two commits?}
 
@@
   \centering
-  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
-  \caption{The first merge tree used in the user study, a merge tree
-    containing a single commit}
+\begin{tabular}{ m{1.5cm} m{3cm} }
+  \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
+  \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
+\end{tabular}
+  \caption{The merge trees used in the user study, 
+    containing one and seven commits, respectively.}
   \label{fig:commit_1}
@@
 
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
-  \caption{The second merge tree used in the user study, a merge tree
-    containing seven commits}
-  \label{fig:commit_2}
-\end{figure}
-
 \subsection{Results}
@@
 
+\dmg{this is still methodology.}
+
 We recorded the audio and screen capture video for each participant in
-the study for further analysis. After capturing the information, we
-extracted various metrics from the videos. From the conceptual tasks, we
+the study for further analysis.
+\dmg{you don't need this sentence, it is redundant}
+After capturing the information, we
+extracted various metrics from the videos.
+\dmg{ you don't record, you measure}
+From the conceptual tasks, we
 recorded the time in seconds, and the answer for the question. We
 extracted three metrics for the summarization tasks, the time in
-seconds, the correctness, and the accuracy. Correctness is a binary
+seconds, the correctness, and the accuracy.
+\dmg{and then somethings you are over precise :): correctness is simply whether their answer to the task was correct or not}
+Correctness is a binary
 value, it is true if the response was correct and false if the response
-was incorrect. The accuracy is measures in distance from the correct
+was incorrect.
+\dmg{if you are going to use this metric, you need to define precisely. remember, you are giving this to somebody else
+  to implement}
+The accuracy is measures in distance from the correct
 value. An accuracy of zero indicates a correct answer. All tasks in the
@@
 
+\dmg{i don't like the expression: we are able to see... too long.. be to the point: Table X shows/depicts/ etc ... they
+  are not overviews, they are a summary}
 In Table~\ref{tab:conceptual_results}, we are able to see an overview of
@@
   \centering
-  \caption{Results from the conceptual questions}
+  \caption{Results from the conceptual questions\dmg{what are the columns? they need captions}}
   \label{tab:conceptual_results}

### 2017-07-07 paper/lineval.tex
Mashing the two papers together

@@
 \begin{abstract}
+
+  With an average of more than 900 top-level merges into the Linux
+  kernel per release, many containing hundreds of commits and some
+  containing thousands, maintenance of older versions of the kernel
+  becomes nearly impossible.  Various commercial products, such as the
+  Android platform, run older versions of the kernel. Due to security,
+  performance, and changing hardware needs, maintainers must understand
+  what changes (commits) are added to the current version of the kernel
+  since the last time they inspected it in order to make the necessary
+  patches.
+
+  Current tools provide information about repositories through the
+  directed acyclic graph (DAG) of the repository, which is helpful for
+  smaller projects. However, with the scale and number of branches in
+  the kernel the DAG becomes overwhelming very quickly. Furthermore, the
+  DAG contains every ancestor of every commit, while maintainers are
+  more interested in how and when a commit arrives to the official Linux
+  repository.
+
+  In this paper, we propose the merge-tree, a simplified transformation
+  of the DAG of the Linux git repository that shows the way in which
+  commits are merged into the master branch of Linux. Using the
+  merge-tree, we build \tool, a tool that is designed to allow users to
+  explore how commits are merged into the Linux kernel.
+
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:introduction}
 
@@
 made in the current version of the kernel in order to produce the
-necessary patches for the older versions of the kernel. Current tools
-and visualizations are based on the directed acyclic graph (DAG) of the
-repository, showing all commits and merges in chronological order by
-when the commit was authored, not by when it arrived in the official
-Linux repository.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/gitk.png}
-  \caption{
-    The Gitk interface, centered at commit
-    cdbdd1676a5379f1d5cbd4d476f5e349f445befe. The top left pane showing
-    the DAG, the top center pane showing the authors, and top right pane
-    showing the commit dates. The bottom-left pane shows the full
-    commit message and diff of the selected commit, and the bottom-right
-    pane shows the list of the files modified by the selected commit.
-  }
+necessary patches for the older versions of the kernel. Tools like Gitk
+use a directed acyclic graph (DAG) model of the repository, showing all
+commits and merges in chronological order by when the commit was
+authored, not by when it arrived in the official Linux repository.
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
+        \caption{A view of the Gitk interface centered on merge
+          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
+          their ancestors and there is no clear path from the commit to the merge
+          with the master branch. Neither Gitk nor Git are capable of showing the
+          commits in master.}
         \label{fig:gitk}
 %\vspace{-4mm}
-\end{figure}
-
-Git is a distributed version control system that is used in many
-open-source and closed-source projects. Git enables many developers to
-write code simultaneously, working on various components known as
-branches. Git uses a directed acyclic graph model (DAG) for storing the
-necessary information regarding the state of the code at any point in
-time. Git refers to nodes of the DAG as commits, not distinguishing
-between the code-carrying commits and the structural commits that bring
-two branches together. In this paper, we will refer to the individual
-commits as being commits, the structural merge commits as merges, and
-refereeing to all as repository events.
-
-The DAG representation works in smaller projects; it enables users to
-see when changes are made, when these changes are merged, how each
-branch is interacting, and the point where a branch forks from the
-master branch. In large modular projects, like the Linux kernel, the DAG
-becomes a mess of merges and commits (Figure~\ref{fig:gitk}) losing its
-visual meaning. In some cases, the repository is simply too large for
-systems to generate a visualization; repositories stored on GitHub will
-have a DAG visualization to show the branches and forks of a
-repository, but is unable to display the visualization for projects at
-the scale of Linux or even the Git Git
-repository(Figure~\ref{fig:gitfail}). Between 60k and 70k new commits
-are created for the Linux project every year; according to previous
-work\cite{German2015}, a commit takes a median of 30 days from the time
-it is authored until it arrives in the official repository. The snapshot
-of the kernel tomorrow may be different than the snapshot from today,
-containing new commits authored in the past; distinguishing these new
-commits from the commits in the snapshot from today is not trivial.
+\end{figure*}
 
@@
         \centering
-  \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
-  \caption{Github is unable to display the DAG of the Linux
-    Project due to the size and complexity, which would normally
-    show the relationship to other branches and forks.}
+        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
+        \caption{Github Failure Message showing the DAG of the Linux Project and its
+                relationship to other forks.}
         \label{fig:gitfail}
@@
 
-One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch, including
-the master branch. To complicate the problem, the DAG only has references to the
-ancestors of a commit (a model necessary for the operation of Git), but maintainers
-would prefer knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would imply that
-for any given merge, it would be possible to know which commits were merged. A user
-could inspect the commits that arrived into the master branch within a given
-time-frame by checking which commits were merged during that time-frame.
-
-This paper makes two contributions; first, we describe a method of converting the
-DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach the master branch; second, we
-present a method to inspect and visualize the history of merges in the Linux project
-using the merge-tree model.
-
-These methods and visualizations are implemented in a web-based tool called
-\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
-Our visualizations and tool provide information about the location of any given
-commit or merge in its respective merge-tree, the files edited, the modules edited,
-and the commit message. \tool allows users to apply various filters, including the
-release version, along with a keyword or phrase from the log preview, the name of
-the author, or the commit ID. The user can request all merges made by Linus that
-contain a commit or inner merge that matches the search query, or just the commits
-and merges that match the query.
-
-\section{\mt Model}
-\label{sec:mt_model}
-
-We propose a new model, built from the DAG, with the purpose of
-providing a clearer conceptual image of how a commit is merged, other
-commits that are merged with a commit, and a summarization of the entire
-group of commits that are merged with a commit. We call our model the
-\mt model. The model inverts the DAG, so that reference point from
-parent to child instead of from child to parent. This enables the
-aggregation of commit metrics at merges. Given a merge, we are able to
-recursively traverse the children of the merge, collecting the
-statistics at each commit. The commits become the leaves of the tree,
-and internal nodes are merges that do not merge into the master branch
-of the repository. We terminate the further traversal of an edge under
-two conditions; first, if the next event is on the master branch of the
-repository, second, if the next event has a shorter path into the master
-branch of the repository. As a consequence, commits will always take the
-shortest path, in terms of commit time, to the master branch.
+This representation works in smaller projects; it enables users to see
+when changes are made, when these changes are merged, how each branch is
+interacting, and the point where a branch forks from the master branch.
+In large modular projects, like the Linux kernel, the DAG becomes a mess
+of merges and commits (Figure~\ref{fig:gitk}) losing its visual meaning.
+In some cases, the Linux kernel is simply too large for the system to
+generate a visualization; Github provides a DAG view for many projects,
+but is unable to display the visualization for projects at the scale of
+the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and 70k
+new commits are created for the Linux project every year; according to
+our previous work\cite{German2015}, a commit takes a median of 30 days
+from the time it is authored until it arrives in the official
+repository. The snapshot of the kernel tomorrow may be different than
+the snapshot from today, containing new commits authored in the past;
+distinguishing these new commits from the commits in the snapshot from
+today is not trivial.
+
+One major challenge with visualizing the arrival of commits to a
+repository is that Git does not store the date that a commit was merged
+into another branch, including the master branch. To complicate the
+problem, the DAG only has references to the ancestors of a commit (a
+model necessary for the operation of Git), but maintainers would prefer
+knowing the path a commit followed to reach the master repository.
+Tracing a path that any commit followed to the master repository would
+imply that for any given merge, it would be possible to know which
+commits were merged. A user could inspect the commits that arrived into
+the master branch within a given time-frame by checking which commits
+were merged during that time-frame.
+
+This paper makes two contributions; first, we describe a method of
+converting the DAG of the Linux repository into a tree, or
+\emph{merge-tree} of the repository, that represents the path used by a
+commit to reach the master branch; second, we present a method to
+inspect and visualize the history of merges in the Linux project using
+the merge-tree model.
+
+These methods and visualizations are implemented in a web-based tool
+called \tool\footnote{\tool is currently available at
+  \url{http://li.turingmachine.org}}.  Our visualizations and tool
+provide information about the location of any given commit or merge in
+its respective merge-tree, the files edited, the modules edited, and the
+commit message. \tool allows users to apply various filters, including
+the release version, along with a keyword or phrase from the log
+preview, the name of the author, or the commit ID. The user can request
+all merges made by Linus that contain a commit or inner merge that
+matches the search query, or just the commits and merges that match the
+query.
+
+\section{Merge-Tree model}
+\label{sec:mergetree}
+
+Git uses a directed acyclic graph (DAG) as its main data model. In this
+model, a commit has one or more ordered parents (ancestors), except the
+root commits of a repository which do not have ancestors--Linux's git
+repository has two such commits.  Commits are divided into merge
+commits, merging two or more branches, and non-merge commits. Non-merge
+commits have only one parent, while merge commits have two or more. The
+order of the parents matter: the first parent is the branch in which the
+merge is being done, while the rest indicate the branches being merged
+(Git allows merging multiple branches simultaneously). The second parent
+is the first branch merged, the third parent is the second branched
+merged, etc.
+
+Once a commit is created, it is never changed. Git allows operations to
+alter commits or reorder them, but it changes the commit ID in the
+process, effectively replacing it with a new commit. This makes commits
+unable to record the traversal of merges from the commit to the merge
+into the master branch.
 
@@
 the master branch and the repository that the master branch is part of.
-The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
+The merge-tree view of this DAG is visible in Figure~\ref{fig:repoTree}.
 Note that the direction of the edges of the DAG have been inverted,
@@
   \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state] (1) {1};
-    \node[state, above right= of 1] (2) {2};
-    \node[state, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, right= 2cm of 1] (4) {4};
-    \node[state, above right=of 4] (5) {5};
-    \node[state, above right=of 5] (6) {6};
-    \node[state, right=of 5] (7) {7};
-    \node[state, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, right= 2cm of 7] (9) {9};
-    \node[state, right=of 9] (10) {10};
-    \node[state, below right=of 9] (11) {11};
-    \node[state, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[-stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[-stealth] (9);
-    \draw (10) edge[-stealth] (9);
-    \draw (9) edge[-stealth] (8) edge[-stealth] (6)
-              edge[-stealth] (7);
-    \draw (8) edge[-stealth] (7);
-    \draw (7) edge[-stealth] (5);
-    \draw (6) edge[-stealth] (5);
-    \draw (5) edge[-stealth] (3) edge[-stealth] (2)
-              edge[-stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[-stealth] (2);
-    \draw (2) edge[-stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
-    information about which repository a commit comes from. Furthermore,
-    the DAG does not distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
     \node[state, draw=chartblue] (1) {1};
@@
 
+    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
+    \draw (10) edge[chartyellow, -stealth] (9);
+    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
+              edge[chartyellow, -stealth] (7);
+    \draw (8) edge[chartmagenta, -stealth] (7);
+    \draw (7) edge[chartyellow, -stealth] (5);
+    \draw (6) edge[chartred, -stealth] (5);
+    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
+              edge[chartyellow, -stealth] (4);
+    \draw (4) edge[-stealth] (1);
+    \draw (3) edge[chartmagenta, -stealth] (2);
+    \draw (2) edge[chartyellow, -stealth] (1);
+  \end{tikzpicture}
+  \caption{DAG representation of the commits represented in
+    Figure~\ref{fig:repoEvents}. The DAG loses information about which
+    repository the commit is performed in and through which merges it
+    has passed on its way to the master branch. The DAG does not even
+    distinguish the master branch from other branches.}
+  \label{fig:repoDAG}
+%\vspace{-3mm}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
+
+    \draw[chartblue]
+      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
+
+    \node[state, draw=chartblue] (1) {1};
+    \node[state, draw=chartyellow, above right= of 1] (2) {2};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
+    \node[state, draw=chartyellow, above right=of 4] (5) {5};
+    \node[state, draw=chartred, above right=of 5] (6) {6};
+    \node[state, draw=chartyellow, right=of 5] (7) {7};
+    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
+    \node[state, draw=chartyellow, right=of 9] (10) {10};
+    \node[state, draw=chartblue, below right=of 9] (11) {11};
+    \node[state, draw=chartblue, below right=of 10] (12) {12};
+
     \draw (2) edge[chartyellow, -stealth] (5)
@@
   \end{tikzpicture}
-  \caption{\mt view of the commits  represented in
+  \caption{Merge-tree view of the commits  represented in
     Figure~\ref{fig:repoEvents} showing the path they followed to reach
@@
 
-Computing the \mt from a DAG for any repository may not be
+\subsection{Computing the merge-tree of the DAG of Linux}
+
+Computing the merge-tree from a DAG for any repository may not be
 possible; however, certain features of the development process of Linux
-make it feasible to compute the \mt for the Linux repository.
+make it feasible to compute the merge-tree for the Linux repository.
 First, the master branch of Linux is maintained by Linus Torvalds, and
@@
 \begin{algorithm}
-        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
-        \label{alg:original}
+        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
         \begin{algorithmic}
@@
 
-We evaluate the performance of the heuristic using the merge log made by
-Linus Torvalds, the original creator of Linux, and primary merger of all
-other repository events into the master branch of the repository. In
-Figure~\ref{fig:sampleMerge}, we see an example of the merge log.
+\subsection{Evaluation}
+
+Merges that do not have conflicts provide information to verify this
+heuristic. If a merge does not contain a conflict, it records a summary
+of the commits that it merges. See Figure~\ref{fig:sampleMerge} for an
+example. This summary contains a list of the first 20 non-merge commits
+in the merge, including their one-line log description, the full logs of
+the merge commits that merge this subset, and the total number of
+non-merge commits in the merge.
 
@@
         \end{verbatim}}\vspace{-5mm}
-  \caption{Example of how merges record a subset of commits being
-    merged. The commit only shows the first 20 one-line summaries
-    messages for the 24 non-merge commits it merged. The ending
-    ``\ldots'' is part of the log and represents that other
-    commits were merged.}
+        \caption{Example of how merges record a subset of commits being merged. The
+                commit only shows the first 20 one-line summaries messages for the 24
+                non-merge commits it merged. The ending ``...'' is part of the log
+                and represents that other commits were merged.}
         \label{fig:sampleMerge}
@@
 
-We use this information to evaluate the accuracy of the \mt model
-extracted from the DAG\@. The method we followed started with the
+We used this information to evaluate the accuracy of the merge-tree
+model extracted from the DAG. The method we followed started with the
 extraction of the commit history up to July 20, 2016. We computed the
@@
 merging directly into master, we assumed that every merge by him is the
-root of a \mt. As described above, the log of a merge-commit usually
-contains the number of commits in the merge the first 20 summaries of
-commits being merged. We extracted merges by Linus Torvalds using the
-command \mycode{log --merges --author='Torvalds} and compared the number
-of commits according to the log with the number of commits in the  \mt
-rooted in this commit. We also used the summaries of the commits found
-in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the
-merge in Figure~\ref{fig:sampleMerge} we would expect that the merge
-tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
-summaries corresponds to commits in that \mt. We also inspected those
-with differences to make sure they were true errors.  The results can be
-summarized as follows:
+root of a merge-tree. As described above, the log of a merge-commit
+usually contains the number of commits in the merge the first 20
+summaries of commits being merged. We extracted merges by Linus Torvalds
+using the command \mycode{log --merges --author='Torvalds} and compared
+the number of commits according to the log with the number of commits in
+the  merge-tree rooted in this commit. We also used the summaries of the
+commits found in the merge (not necessarily all---see above) to make
+sure those commits were in their corresponding merge tree. For example,
+for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
+merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
+one-line summaries corresponds to commits in that merge-tree. We also
+inspected those with differences to make sure they were true errors.
+The results can be summarized as follows:
 
 \begin{itemize}
-
-  \item
-
-    Five merges were false-errors because their logs did not contain
-    accurate information (were probably edited by hand). For example in
-    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
-    empty), in \mycode {c55d267\ldots} the summaries were reordered.
-
-  \item
-
-    The heuristic correctly identified that 79 of Linus merges (between
-    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
-    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
-    commits.
-
-  \item
-
-    The heuristic worked perfectly until Sept 4, 2007, the earliest date
-    that it could be verified.  Before this date, and until Dec 12,
-    2006, merges did not include a summary of the commits they included,
-    hence making it impossible to verify; during this period, however,
-    we correctly identified the merges by Linus into master.
-
-  \item
-
-    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
-    which confounded the true master branch of a repository (see
-    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
-    for a description of the issue).
-
+\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
+  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
+  {c55d267} the summaries were reordered.
+\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
+  branch (not master). This branch was merged at \mycode{3f17ea6d..} which contained 6809 commits.
+\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
+  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
+  impossible to verify; during this period, however, we correctly identified the merges by Linus into master.
+\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
+presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
+repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
 \end{itemize}
 
-In summary, of the merges after Sept. 4, 2007, our heuristic was correct
-in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
-12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
-2007).
+In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860)
+our heuristic was correct. It failed in 1,542 commits before Dec. 12,
+2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
+
+\section{Visualizing the merge-tree of Linux}
+
+The goal of \tool is to simplify the navigation of the kernel commit
+information, specifically focusing on merges. This is done by leveraging
+the merge-tree to inspect how commits are merged on the path to the
+master repository.
+
+\subsection{Use cases}
+
+We designed \tool with two use-cases in mind, though a user may switch
+between the cases as they work.
+
+\noindent \textbf{Use-case 1: top-to-bottom
+  approach}\label{sec:usecase1}\\ These are users that are maintaining a
+section of the kernel and would like to pick a merge (including all the
+commits that it merges) and merge it directly into their current
+repository. This is useful for reducing the amount of re-implementation
+work. For these users, it is important to have the ability to aggregate
+metadata about files and modules being effected by the merge. Also, it
+is important for these users to be able to navigate from the root of the
+merge-tree toward the leaves.
+
+\noindent \textbf{Use-case 2: bottom-to-top
+  approach}\label{sec:usecase2}\\ These are users that start with a
+known merge or commit and would like to see what other changes are being
+made in commits that are in the same merge, including knowing the
+merge-tree they belong to. This is useful to see what other commits are
+related to the current commit and how they get collated into merges that
+eventually end in the master branch. This is primarily for maintainers
+that need to perform some specific cherry picking of commits. We must
+provide these users a mechanism for navigating from a single commit
+toward the master branch, allowing them to see other commits that might
+be related to their original commit.
+
+\subsection{Data Model}
+
+In our visualizations, we leverage the merge-tree model described in
+Section~\ref{sec:mergetree}. In this model each commit is either already
+in the master branch, or is part of a tree which is rooted in the merge
+that merged it into the master branch.  Each commit, whether a merge or
+non-merge, has only one successor; the root of each tree has none as it
+was made by Linus Torvalds directly into the master branch. Non-merge
+commits contain the metadata for the changes made.  This metadata
+includes the files changed, the lines added and removed from each file,
+the author, the date the commit was merged into the merge that led to
+being merged into the kernel, the date the commit was authored, the
+patch, and the commit log. Merges contain less metadata, only storing
+the author of the merge, the log, the commit date, and the author date,
+and potentially, changes necessary to address conflicts during the
+merge. The details of the model are outlined in \cite{German2015}.
 
 \section{Design and Implementation}
-\label{sec:design_and_implementation}
-
-We implement a tool, \tool, to test and verify the capabilities of the
-\mt model. \tool provides search, summarization, and
-visualization using the \mt model.
-
-We provide users with search-based navigation. The search engine takes a
-plain text query from the user and returns results based on similarity
-to the query. The engine weighs various fields, including the commit
-log, the author name, the commit hash, and the dates that the commit was
-authored and committed. The results are aggregated by the \mt
-with a link to the root at the top, and the repository events that match
-the search query in a table below.  An example of a single \mt is shown
-in Figure~\ref{fig:linvis_search_results}.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/search_results.png}
-  \caption{\tool search results for `net-next 2011', showing a single
-    \mt group.}
-  \label{fig:linvis_search_results}
+
+To navigate and inspect the merge-tree view of the kernel we created a
+web-based tool called \tool. Creating a web-based tool enables users to
+use the system without having to install additional software or store a
+large database, making it more accessible, more easily maintainable, and
+platform independent. \tool uses the following mechanisms to reach our
+goals of better navigation and better explanation of the selected
+changes.
+
+\begin{itemize}
+        \item Filter by searching
+        \item View files edited
+        \item View modules
+        \item Tree viewer
+\end{itemize}
+
+\subsection{Searching}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/search.png}
+        \caption{Search View allows filtering commits that were merged in a given
+                period, filtering by author, keyword, or commit ID.}
+        \label{fig:search}
 \end{figure}
 
-\tool uses seven tabs to show the information and visualizations for a
-selected commit or merge. The message tab shows the full commit log
-message. This does not include the diff. The files tab shows an
-aggregated table of all files that were modified in a merge. It includes
-metrics like the number of lines added, lines removed, total lines
-modified, and the delta, summed across all commits in the merge tree
-that modify this file. A small details drop-down button allows a user to
-see exactly which commit makes the changes, as shown in
-Figure~\ref{fig:linvis_files}. If the current repository event being
-viewed is a commit, the aggregate views will only show modifications
-made by the commit. The modules tab shows the modules modified in the
-merge tree. Like the files tab, the modules tab uses a table to show the
-name of the module, the number of commits that are in the merge tree
-that work with the module, and a details button to provide the links to
-those commits.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_files.png}
-  \caption{Table showing the modified files in a merge, with the second
-  entry expanded to show the commit that makes the changes.}
-  \label{fig:linvis_files}
+Searching (depicted in Figure \ref{fig:search}) allows a user to filter
+commits and merges that are irrelevant. The search mechanism breaks down
+the results by release version. A user can further narrow down the
+search by specifying a range of dates in which such commits were merged
+by Linus into the master branch---not when the commits were created
+(author date and/or commit date). This distinction is important. We have
+observed commits that have taken years to arrive into the master
+repository after they were originally created.
+
+A user may then provide a search text, filtering by the author name, the
+commit ID, or keyword from the log. Any part of the author name may show
+up in the results, including searching by email address.
+
+If the user is searching by commit ID, the ID can be specified by using
+any of its unique prefixes. For example, the commit
+\mycode{3f17ea6dea8ba5668873afa54628a91aaa3fb1c0} is returned when the
+user searches for a commit ID of \mycode{3f17e} in the 3.16 Linux
+kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
+        \caption{Search Results. Each table entry is a commit merged in the desired
+                merged window.}
+        \label{fig:results}
 \end{figure}
 
-The authorship tab is very similar to the files tab, but shows the
-authorship information. It shows the sum of the number of lines added,
-removed, modified, and the delta within the \mt. It also shows
-the number of files that were modified by the author. The details tab is
-organized slightly differently. Instead of organizing these results by
-commit, the details are organized by file, showing which files were
-modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_authors.png}
-  \caption{Table showing the authors who made changes in a merge. The
-    entry for Randy Dunlap is expanded, showing the modifications to
-    each file that were made by Randy.}
-  \label{fig:linvis_authors}
+In the search results (seen in Figure~\ref{fig:results}), the user is
+presented with the one-line log message preview, the author's name and
+email, the date the commit was authored, and the date the commit was
+last committed.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
+        \caption{Panel showing the Commit Message of a commit.}
+        \label{fig:message}
 \end{figure}
 
-\tool implements three visualizations, a list tree, pack tree, and
-Reingold-Tilford tree. The list tree is a series of indented textual
-lists that can be easily searched with the built-in search in most
-web-browsers. The list tree is rooted at the current event, if the event
-is a commit, only that commit will be shown; conversely, if the selected
-event is the root, the entire merge tree will be shown. The Pack-Tree or
-Bubble tree, uses a set-based approach for showing the parent
-relationship. The root node aggregates all of the events within the \mt
-into the master branch. Thus, the root is the superset of all the other
-nodes in the tree. The commits are shown as individual white circles,
-containing no other events, as these are the leaves. This is shown in
-Figure~\ref{fig:linvis_pack}. The selected event will be shown in bright
-orange, commits in white, and merges in shades of blue. Darker shades of
-blue indicate that the merge is deeper in the tree.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
-  \caption{The pack tree visualization for a selected commit, shown in
-    orange.}
-  \label{fig:linvis_pack}
+Once a user selects a commit or merge to investigate, they are presented
+with a tabbed pane allowing them to view the full commit log, the files
+edited, the modules involved, and the merge-tree view.
+
+The first tab displays the full commit log (Figure~\ref{fig:message}).
+From this, a user is able to see what they would see had they searched
+for the commit using Git log. This doesn't provide additional
+information to the other tools, but helps to complete the functionality
+of \tool. The commit log provides a user with the information about the
+content of the commit and who has signed-off on the commit to ensure
+that it is of good quality. The message for merges may contain a summary
+of the commits being merged.  The information within these messages is
+highly variable, and is completely dependent on the author's style. As
+the user moves toward the root-level merge, the quality of these
+messages generally improves.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
+        \caption{Panel showing the files modified by all the commits that are part
+                of this merge.}
+        \label{fig:files}
 \end{figure}
 
-The Reingold-Tilford tree is the classic tree visualization, with the
-root at the top, leaves at the bottom, and the edges between them
-showing the parent relationship. An example of the Reingold-Tilford tree
-visualization is shown in Figure~\ref{fig:commit_1} and
-Figure~\ref{fig:commit_2}.
+The second tab is the files tab (Figure~\ref{fig:files}). This tab
+provides information on what files have been edited, how many lines were
+added, and how many lines were removed in a given commit. For
+non-merges, this functionality is similar to the other tools available.
+Our tree-based design model allows us to extend this functionality to
+merges by aggregating information about all the commits that are
+children of the merge in the merge-tree, which other tools are unable to
+show. To find the number of lines added to a file in a merge, we take
+the sum of the lines added to that file in each of the children of that
+merge. We do the same for calculating the number of lines removed.
+
+% This could be extended to use the patch information to determine when a line has been
+% replaced rather than incrementing the lines added and removed.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
+        \caption{Panel showing the modules changed by all the commits in this merge.}
+        \label{fig:modules}
+\end{figure}
+
+The modules tab (Figure~\ref{fig:modules}) shows the modules that are
+contained within the commit. Modules are not natively recognized by Git,
+and are not going to be present in all repositories. In the Linux
+repository, authors put the module they are working on in the one-line
+summary of the log-message; for example: the log \emph{gcov: add support
+  for FCC 4.9} has updated the \emph{gcov} module (the coverage testing
+tool of the kernel). We heuristically extract the module by taking all
+text in the log summary of commits until the first colon.  Modules are
+logical partitions of the information in the kernel. Depending on where
+the author was working, modules can be general, such as ``bluetooth''
+and ``wireless'', or can be quite specific for individual hardware, such
+as ``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit
+does not correctly follow this format and the heuristic approach fails.
+As with the Files panel, non-merge commits show their corresponding
+information, but for merge commits we aggregate all the modules changed
+in all the commits that are part of that merge.  The output of this view
+is shown in Figure~\ref{fig:modules}.
+
+Finally, we have the Tree view tab. The tree view is designed for
+providing easy navigation of the commits within the merge-tree that is
+rooted in the current merge.  It also provides a clear topological view
+of the merge and the submerges it includes. We have experimented with
+various tree designs to find a design that allows for easy navigation
+and visualization of both large and small trees. We discuss this panel
+in the next subsection.
+
+\subsection{Merge-Tree Views} \label{treeview_section}
+
+The merge-tree view is what makes \tool unique to other tools that
+inspect the DAG of a Git repository. With it, a user can inspect how
+commits are merged on their way to the master branch. We have
+experimented with various types of trees:
+
+\begin{enumerate}
+        \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
+        \item Reingold-Tilford trees provide a clear visual representation of the tree structure of the merge-tree.
+        \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
+                clearly showing the parent-child relationships between commits and merges.
+\end{enumerate}
+
+\subsubsection{List Tree}
+
+The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of
+nested lists, and is designed to more closely model the tree view found
+in file browsers. This tree only contains the commits and merges that
+are within the subtrees of the current merge. A commit will never have
+any items in this tree as it is a leaf. To accompany the tree, we
+include breadcrumbs at the top of the page to enable a user to navigate
+both from the root to the leaves and from a leaf to the root. The last
+item in the breadcrumb list is the current commit, the previous item is
+the parent of the current commit, and the first item is the root merge
+into the kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
+        \caption{List tree view of merge 3f17ea6}
+        \label{fig:list_tree}
+\end{figure}
+
+\subsubsection{Reingold-Tilford Tree}
+
+The Reingold-Tilford tree\cite{Reingold1981} (Figures
+\ref{fig:reingold_tree} and \ref{fig:reingold_tree_zoom}) allows the
+visualization and navigation of the entire merge-tree in an intuitive
+representation of the tree.  This illustrates a clear notion of root and
+leaves, and how to navigate in either direction. Some merge trees are
+very large, containing thousands of commits and merged. While the tree
+is capable of producing a visualization, it becomes far more difficult
+to understand. For example, the merge, 3f17ea6 performed by Linus
+Torvalds June 8 2014, contains 7217 commits and merges. \footnote{This
+  commit can be inspected at
+  \url{http://li.turingmachine/org/commits/3f17ea6dea8ba5668873afa54628a91aaa3fb1c0}}
+
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/4dc4_tree.pdf}
+        \caption{Merge 4dc4226 is a subtree of 3f17ea6, for the power-management
+          module of the kernel}
+        \label{fig:reingold_tree}
+      \end{figure}
+
+      \begin{figure*}
+        \centering
+        \includegraphics[width=0.98\textwidth]{figures/4dc4_zoom_tree.pdf}
+        \caption{Zoomed out view of the Reingold-Tilford tree for merge 4dc4226}
+        \label{fig:reingold_tree_zoom}
+      \end{figure*}
+
+      The user is initially greeted with their current node centered on
+      the screen.  They are able to zoom the tree by scrolling the
+      mouse, and clicking and dragging to pan the tree. They can see
+      more details about a node by clicking on it, which will provide
+      them a link to the specific page for that commit.
+
+\subsubsection{Bubble Tree}
+
+Bubble trees are useful for providing a clear visualization of wide,
+hierarchical data\cite{Boardman2000}. The tree structure is represented
+by the nesting of nodes; the largest circle is the root, containing all
+the other nodes. The smallest circles do not contain any nodes and
+therefore represent the leaf nodes.
+
+Our implementation of the bubble tree (Figure~\ref{fig:bubble_tree})
+provides the user with a clear picture of where a commit is located in
+the merge tree. We highlight the selected commit or merge in red.
+Non-selected merges are in a shade of blue determined by the depth of a
+node in the tree. The root is the lightest shade of blue, while the
+contained merges are progressively darker; the commits are white.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
+        \caption{Bubble Tree of merge 3f17ea6, merge 4dc4226 is the currently
+          selected merge and is highlighted in orange.}
+        \label{fig:bubble_tree}
+\end{figure}
+
+The bubble tree doesn't have an implicit way of providing additional
+information, placing any text near the nodes makes the tree impossible
+to read, so we include a separate pane in the web page. When a user
+hovers over a node, the pane shows additional information about the
+author and the commit message and a link to the detailed page for that
+commit or merge. If the user clicks on a node, the tree will zoom to
+that node and the information in the info pane will persist, enabling
+the user to click the link. This tree provides an easy mechanism for
+users to navigate from the root to the leaves and vice versa.
+
+\section{Discussion}
+
+From a developers point of view, the are two major disadvantages of the
+DAG model of Git: a) its edges point backwards, i.e. commits point to
+their ancestor, not to the commit that succeeds them on the path to
+being integrated; and b) the DAG contains many more edges that are
+necessary to understand how integration occurred. We addressed these two
+issues with the creation of the merge-tree model from the DAG of Linux.
+Effectively, the merge-tree recovers the details of how each
+branch---and the commits they contain---was merged into the master
+branch.
+
+In our experience, no other Git repository reaches the level of DAG
+complexity that Linux has. In most Git repositories, most merges are not
+nested---most merges merge directly into the master repository, and
+these merges only merge a few commits. Even in these simplified cases,
+the merge-tree can provide a valuable summary of how commits are
+integrated into the master branch, specifically since the time of
+integration may be very different from the time the commits were
+authored.
+
+\evan{Check accuracy of this paragraph}
+
+The biggest challenge is computing the merge-tree of any given
+repository. It is likely that our heuristic for computing the merge-tree
+will not work with other repositories. The heuristic requires that there
+are no fast-forward merges into the master branch of the repository (a
+majority of repositories), and that there are no foxtrot merges (a
+practice that is starting to be considered desirable by git
+users\footnote{See
+  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git}
+  and
+  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
+We are able to use our heuristic with the Linux kernel because of the
+strict integration model imposed by Linus. To validate the results of
+the heuristic, we use the continuous mining technique described in
+\cite{German2015}.  Continuous mining of the repository requires
+foresight and planning.
+
+Because \tool leverages the merge-tree model of the kernel, it provides
+mechanisms and visualizations that other tools are unable to produce. It
+gives us the ability to see how commits are being merged into the master
+branch, allowing us to better understand how code is integrated into the
+kernel.
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
+        \caption{Merge Dag View}
+        \label{fig:dag_view}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
+        \caption{Merge tree topology for the same merge as in
+                Figure~\ref{fig:dag_view}.}
+        \label{fig:tree_view}
+\end{figure}
+
+Our goal was to build a tool to enable maintainers to effectively
+navigate and browse the changes performed to the kernel over the period
+of a release. Achieving this goal includes removing information that
+does not pertain to the area of the kernel that the user is interested
+in. In Figures~\ref{fig:dag_view} and \ref{fig:tree_view} we can
+visually compare the results returned from Gitk and our tool for the
+top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6'' from kernel
+version 3.10. Note that Gitk has no way to show or recognize that this
+is a merge into the master branch.
+
+The log information to this merge-commit in both of these figures is
+identical, but the presentation drastically changes our ability to
+comprehend what we are seeing.  The primary difference is the removal of
+irrelevant information. In Gitk (Figure~\ref{fig:dag_view}), the
+visualization results include commits and merges from other components
+of the kernel, while our tool (Figure~\ref{fig:tree_view}) only includes
+the results that are specific to the component of the kernel that we are
+interested in.
+
+Even though this merge is relatively small, containing only four
+sub-merges, three of which only contain a single commit, and the last
+containing two commits, it is already hard to visualize in Gitk.  With
+\tool we are able to immediately see what section of the kernel these
+commits and merges pertain to. The DAG view of this provides almost no
+explanation, furthermore, users must work to determine where this merge
+ends and the next one begins.
+
+\tool further enhances our understanding by summarizing the files and
+modules that were edited in the entire branch being merged. We are able
+to determine that three files, ``palamas-regulator.c'' had two lines
+added and two lines removed, ``dbx500-prcmu.c'' had 12 lines added and
+12 lines removed, and ``core.c'' had 5 lines added and 2 lines removed.
+Finally, we are able to determine that only the ``regulator'' module was
+modified in this merge and was modified by 5 commits.
 
@@
 
-\dmg{watch for tense... we don't beging... the tasks begin... or better, the first tasks...}
+\dmg{watch for tense... we don't begin... the tasks begin... or better, the first tasks...}
 
 The questions and tasks are broken down into four groups. The first two
-groups are purely quantitative in nature; we begin with conceptual tasks
-to understand whether the DAG is capable of providing users with an
-understanding of how a commit is merged into the master branch, and
-other commits that it is merged with. We continue with summarization
-tasks, to quantitatively understand if the \mt model makes a difference
-in summarization tasks.
+groups are purely quantitative in nature; the study begins with
+conceptual tasks to understand whether the DAG is capable of providing
+users with an understanding of how a commit is merged into the master
+branch, and other commits that it is merged with. We continue with
+summarization tasks, to quantitatively understand if the \mt model makes
+a difference in summarization tasks.
 
@@
 
-\dmg{given the choice, not allowed. 
-  I don't understand this section.
-  You need to be more top-down in your description. you need to provide context on what the
-  users are expected to do, and why. So what are the tasks? You never explain.
-what is the goal of the tasks? how many? 
-what were the tasks? what is expected from each? you need to answer these questions, not necessarily in the same order.
-also, remember this is a comparative study, so I presume the users were asked to do both with linvis and without it
-(choice of gitk/git command line)
-}
+\dmg{given the choice, not allowed.  I don't understand this section.
+  You need to be more top-down in your description. you need to provide
+  context on what the users are expected to do, and why. So what are the
+  tasks? You never explain.  what is the goal of the tasks? how many?
+  what were the tasks? what is expected from each? you need to answer
+  these questions, not necessarily in the same order.  also, remember
+  this is a comparative study, so I presume the users were asked to do
+  both with linvis and without it (choice of gitk/git command line) }
 
@@
 
-\dmg{again, you are going into details first. First the tasks --see above--
-  including their grouping.
-  then at the end, describe the randomization and why it is done (which you did)
-}
+\dmg{again, you are going into details first. First the tasks --see
+  above-- including their grouping.  then at the end, describe the
+  randomization and why it is done (which you did) }
 
 The summarization tasks (Table~\ref{tab:summarization_tasks}) are
-grouped into \dmg{how many?} set of tasks according to their theme. The tasks within a given task set are
-shuffled, and the task sets are shuffled, to ensure that similar
-questions are together, but not in the same order, and to ensure that
-the themes are in randomized order between participants. The order of
-the commits is shuffled, as well. The order of the tool is randomized,
-choosing to start with either \tool or Gitk, to ensure that we do not
-place a bias on one tool through the experiment.
+grouped into \dmg{how many?} set of tasks according to their theme. The
+tasks within a given task set are shuffled, and the task sets are
+shuffled, to ensure that similar questions are together, but not in the
+same order, and to ensure that the themes are in randomized order
+between participants. The order of the commits is shuffled, as well. The
+order of the tool is randomized, choosing to start with either \tool or
+Gitk, to ensure that we do not place a bias on one tool through the
+experiment.
 
@@
 Finally, we ask the user for their opinions on the tools and for some
-information about their level of experience with git.
-\dmg{be more precise: understanding or preference is too broad. What you want to know 
-answers to specific questions regarding how the tools compare and which ones they would prefer}
-The first three questions are intended to provide us with a better understanding
-of preference, and insight on why, the participant had that preference.  
-The last three questions in this section are aimed at getting a better
-understanding of the demographic of our participants.
+information about their level of experience with git.  \dmg{be more
+  precise: understanding or preference is too broad. What you want to
+  know answers to specific questions regarding how the tools compare and
+  which ones they would prefer} The first three questions are intended
+to provide us with a better understanding of preference, and insight on
+why, the participant had that preference.  The last three questions in
+this section are aimed at getting a better understanding of the
+demographic of our participants.
 
@@
 
-\dmg{tell me first why you need to create a script, so I know what you do something:
-we did random assignment of tasks (explain what you did) and randomized the order in which the tasks were performed.
- Then explain why you need a script}
-
-\dmg{it does not randomize the order the tools. It randomizes which tool is going to be used on each tasks.
-be precise with your language}
-
-\dmg{what you explained how commits relate to tasks and how they get assigned? think as if you are explaining this to
-  somebody who has to do this: you are the designer, explaining the why and the how }
+\dmg{tell me first why you need to create a script, so I know what you
+  do something: we did random assignment of tasks (explain what you did)
+  and randomized the order in which the tasks were performed.  Then
+  explain why you need a script}
+
+\dmg{it does not randomize the order the tools. It randomizes which tool
+  is going to be used on each tasks.  be precise with your language}
+
+\dmg{what you explained how commits relate to tasks and how they get
+  assigned? think as if you are explaining this to somebody who has to
+  do this: you are the designer, explaining the why and the how }
 
@@
 
-\dmg{see my comments in previous section. it is ok to explain them here, but you need to explain what they are briefly
-  in the previous section}
+\dmg{see my comments in previous section. it is ok to explain them here,
+  but you need to explain what they are briefly in the previous section}
 
@@
 
-\dmg{this place does not require a description of the database. This requires just to say
-the commits/merge set (dates and numbers are ok) only}
-We use the \tool database, containing commits and merge-tree information
-from April 16th 2005 to October 14th, 2014, which corresponds to being
-between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
-  database even more?}
-\dmg{again, i have problems with your writing: ``we wanted to use in the study commits of different size, to determine
-  if the size had an impact in the completion of the tasks. We chose commits of three different sizes...}
-We chose the commits based on tree sizes. We
-\dmg{we didn't find: this is a fact, not a discovery; what is a majority? 50\%+1? be precise}
-\dmg{drop the word inner, and do it in order: 25\% of trees have only one commit, 50\%.., 75\%...}
-found that a majority of the trees contain at most seven inner commits,
-while more than 25\% of the trees contain a single commit. 75\% of the
-tree contained up to 51 inner commits and merges, and finally the
-largest tree contained 7217 nodes. We chose to work with the trees in
-\dmg{if you are going to use the words quartile, just use them before in your description eg: the distribution of the
-  sizes of the trees is x, y merges, whose size corresponded to the 25 ,50 quartiles, simple and precisee}
-the first and second quartile, as merge trees of sizes between one and
-seven, not including the merge into the master branch, make up the
-majority of the trees in the database.
-
-\dmg{remove from here, what is here? this is not a road trip :)}
-From here, we selected one tree from the trees of a single commit at
-random. Selecting a commit from that tree is trivial, as there is only
-one to choose.\dmg{remove the obvious, instead concentrate on why the others are not trivial}
+\dmg{this place does not require a description of the database. This
+  requires just to say the commits/merge set (dates and numbers are ok)
+  only} We use the \tool database, containing commits and merge-tree
+information from April 16th 2005 to October 14th, 2014, which
+corresponds to being between Linux release 2.6.12-rc3 and Linux
+3.17-rc1. \evan{Cite the database even more?} \dmg{again, i have
+  problems with your writing: ``we wanted to use in the study commits of
+  different size, to determine if the size had an impact in the
+  completion of the tasks. We chose commits of three different sizes...}
+We chose the commits based on tree sizes. We \dmg{we didn't find: this
+  is a fact, not a discovery; what is a majority? 50\%+1? be precise}
+\dmg{drop the word inner, and do it in order: 25\% of trees have only
+  one commit, 50\%.., 75\%...} found that a majority of the trees
+contain at most seven inner commits, while more than 25\% of the trees
+contain a single commit. 75\% of the tree contained up to 51 inner
+commits and merges, and finally the largest tree contained 7217 nodes.
+We chose to work with the trees in \dmg{if you are going to use the
+  words quartile, just use them before in your description eg: the
+  distribution of the sizes of the trees is x, y merges, whose size
+  corresponded to the 25 ,50 quartiles, simple and precisee} the first
+and second quartile, as merge trees of sizes between one and seven, not
+including the merge into the master branch, make up the majority of the
+trees in the database.
+
+\dmg{remove from here, what is here? this is not a road trip :)} From
+here, we selected one tree from the trees of a single commit at random.
+Selecting a commit from that tree is trivial, as there is only one to
+choose.\dmg{remove the obvious, instead concentrate on why the others
+  are not trivial}
 
 We selected one tree of size seven at random from all trees of size
-seven\dmg{how many where there?}. We placed a restriction the trees, the selected tree must include
-at least one internal merge to increase the complexity of the trees
-tested. For each randomly selecting a tree, we chose one of its internal commits at
-random (using the \verb|random.choice| function in python 3.6.1).
-
-\dmg{question: so you chose only one commit of each size, and the randomization was which tool to use
-for each question? or did they have to do the same task for two commits?}
+seven\dmg{how many where there?}. We placed a restriction the trees, the
+selected tree must include at least one internal merge to increase the
+complexity of the trees tested. For each randomly selecting a tree, we
+chose one of its internal commits at random (using the
+\verb|random.choice| function in python 3.6.1).
+
+\dmg{question: so you chose only one commit of each size, and the
+  randomization was which tool to use for each question? or did they
+  have to do the same task for two commits?}
 

### 2017-07-07 paper/abstract.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

--- 2017-06-11 paper/abstract.tex
@@
-Natively, git uses a directed acyclic grph mode (DAG) with references pointing from
-the (TODO: most recent commit?, it's not the HEAD because that moves around) toward
-the initial commit in the repository. This model is useful and necessary to the
-internal functionality of the version control system, but does not provide users
-with a clear nderstand of the events occuring in the repository. Due to the
-direction of the link in the DAG, visualziation tools, like gitk, are unable to
-provide summarized aggregation of the commits at a given merge. Furthermore, the
-visual DAG does not provide a clear explanation of the events taking place in the
+% vim:set et sw=2 ts=4 tw=72:
+With an average of more than 900 top-level merges into the Linux kernel
+per release, many containing hundreds of commits and some containing
+thousands, maintenance of older versions of the kernel becomes nearly
+impossible.  Various commercial products, such as the Android platform,
+run older versions of the kernel. Due to security, performance, and
+changing hardware needs, maintainers must understand what changes
+(commits) are added to the current version of the kernel since the last
+time they inspected it in order to make the necessary patches.
+
+Current tools provide information about repositories through the
+directed acyclic graph (DAG) of the repository, which is helpful for
+smaller projects. However, with the scale and number of branches in the
+kernel the DAG becomes overwhelming very quickly. Furthermore, the DAG
+contains every ancestor of every commit, while maintainers are more
+interested in how and when a commit arrives to the official Linux
 repository.
 
-Merge-Trees are a data structure derived from the DAG, that could potentially
-alleiveate the issues with visualization built directly from the DAG. The trees are
-defined such that the root is the merge into the master branch, and the leaves are
-the individual commits. The references point in both directed, from each node to the
-parent, and from the parent to the children. Depending on the complexity of the
-merge, the path from leaf to root may pass through multiple internal nodes.
-
-In this paper, we verify that the Merge-tree model is able to simplify the
-construction of a conceptual image of the parts of a repository, and simiplify the
-summarization tasks through a user study. We will continue by providing a more
-generalized approach to constructing merge-tress, which will prune components of the
-DAG that are not necessary.
-
-% \evan{Change this when we get the actual results, this is from the pilot study} We
-% find that although the DAG view used by Gitk provides more information about the
-% repository, in many cases it povides too much information and overwhelms the user.
-
-% \evan{Move this}
-% In some cases, where the structure of the merge tree is flat, that is to say that it
-% has no inner merges, users are able to easily determine which commits are merged
-% together, but are unable to determine which merge is the final merge is the root of
-% the tree.
+In this paper, we propose the merge-tree, a simplified transformation of
+the DAG of the Linux git repository that shows the way in which commits
+are merged into the master branch of Linux. Using the merge-tree, we
+build \tool, a tool that is designed to allow users to explore how
+commits are merged into the Linux kernel.

### 2017-07-07 paper/conclusion.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

--- 2017-07-06 paper/lineval.tex
@@
-% vim:set et fenc=utf-8 ft=tex sw=2 ts=2 tw=72:
-
-\newif\ifdraft
-% \drafttrue
-
-\ifdraft
-        \documentclass[draft]{IEEEtran}
-        \def\baselinestretch{1}
-        % \setlength{\marginparwidth}{1.5cm}
-\else
-        \documentclass[conference]{IEEEtran}
-\fi
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage[hyphens]{url}
-\urlstyle{same}
-
-%% Citations
-\usepackage[nospace]{cite}
-
-%% Table Support
-\usepackage{dcolumn}
-\usepackage{longtable}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage{tabulary}
-
-%% Extra support
-\usepackage{xspace}
-\usepackage{amsmath}
-\usepackage{balance}
-\usepackage{placeins}
-
-%% Algorithms
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-
-%% Graphics
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{xcolor}
-\usepackage{color}
-\usepackage{listings}
-
-%% Tikz
-\usetikzlibrary{positioning, arrows}
-
-\input{color.tex}
-\input{macros.tex}
-
-\newcommand{\TheTitle}{Merge-Tree: Visualizing the Integration of Commits into Linux}
-\newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
-\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Understanding a large number of commits and merges}
-\newcommand{\TheKeywords}{Linux, git, visualizations}
-
-%% Referencing
-\usepackage[unicode=true,
-          pdftitle={\TheTitle},
-          pdfauthor={\TheAuthors},
-          colorlinks=false]{hyperref}
-
-\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
-
-\lstset{frame=tb,
-  language=python,
-  aboveskip=3mm,
-  belowskip=3mm,
-  showstringspaces=false,
-  columns=flexible,
-  basicstyle={\small\ttfamily},
-  numbers=none,
-  numberstyle=\tiny\color{gray},
-  keywordstyle=\color{chartblue},
-  commentstyle=\color{chartred},
-  stringstyle=\color{chartgreen},
-  breaklines=true,
-  breakatwhitespace=true,
-  tabsize=3
-}
-
-\synctex=1
-
-\begin{document}
-
-\title{\TheTitle}
-
-\author{
-        \IEEEauthorblockA{\TheAuthors}
-        \IEEEauthorblockN{Department of Computer Science,
-                University of Victoria, Canada.}
-        \IEEEauthorblockA{Email: \TheEmails}
-}
-\maketitle
-
-\begin{abstract}
-\end{abstract}
-
-\begin{IEEEkeywords}
-  \TheKeywords
-\end{IEEEkeywords}
-
-\section{Introduction}
-\label{sec:introduction}
-
-Between 50k and 70k commits are added to the Linux kernel per version
-requiring maintainers of older versions of the kernel to sift through
-thousands of commits and merges with tools that are unable to filter and
-effectively visualize projects at the scale of the kernel. Older
-versions of the kernel are used in embedded systems and mobile phones;
-for security purposes, performance needs, and changing hardware
-requirements, maintainers must be able to understand the changes being
-made in the current version of the kernel in order to produce the
-necessary patches for the older versions of the kernel. Current tools
-and visualizations are based on the directed acyclic graph (DAG) of the
-repository, showing all commits and merges in chronological order by
-when the commit was authored, not by when it arrived in the official
-Linux repository.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/gitk.png}
-  \caption{
-    The Gitk interface, centered at commit
-    cdbdd1676a5379f1d5cbd4d476f5e349f445befe. The top left pane showing
-    the DAG, the top center pane showing the authors, and top right pane
-    showing the commit dates. The bottom-left pane shows the full
-    commit message and diff of the selected commit, and the bottom-right
-    pane shows the list of the files modified by the selected commit.
-  }
-  \label{fig:gitk}
-  %\vspace{-4mm}
-\end{figure}
-
-Git is a distributed version control system that is used in many
-open-source and closed-source projects. Git enables many developers to
-write code simultaneously, working on various components known as
-branches. Git uses a directed acyclic graph model (DAG) for storing the
-necessary information regarding the state of the code at any point in
-time. Git refers to nodes of the DAG as commits, not distinguishing
-between the code-carrying commits and the structural commits that bring
-two branches together. In this paper, we will refer to the individual
-commits as being commits, the structural merge commits as merges, and
-refereeing to all as repository events.
-
-The DAG representation works in smaller projects; it enables users to
-see when changes are made, when these changes are merged, how each
-branch is interacting, and the point where a branch forks from the
-master branch. In large modular projects, like the Linux kernel, the DAG
-becomes a mess of merges and commits (Figure~\ref{fig:gitk}) losing its
-visual meaning. In some cases, the repository is simply too large for
-systems to generate a visualization; repositories stored on GitHub will
-have a DAG visualization to show the branches and forks of a
-repository, but is unable to display the visualization for projects at
-the scale of Linux or even the Git Git
-repository(Figure~\ref{fig:gitfail}). Between 60k and 70k new commits
-are created for the Linux project every year; according to previous
-work\cite{German2015}, a commit takes a median of 30 days from the time
-it is authored until it arrives in the official repository. The snapshot
-of the kernel tomorrow may be different than the snapshot from today,
-containing new commits authored in the past; distinguishing these new
-commits from the commits in the snapshot from today is not trivial.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
-  \caption{Github is unable to display the DAG of the Linux
-    Project due to the size and complexity, which would normally
-    show the relationship to other branches and forks.}
-  \label{fig:gitfail}
-  %\vspace{-2mm}
-\end{figure}
-
-One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch, including
-the master branch. To complicate the problem, the DAG only has references to the
-ancestors of a commit (a model necessary for the operation of Git), but maintainers
-would prefer knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would imply that
-for any given merge, it would be possible to know which commits were merged. A user
-could inspect the commits that arrived into the master branch within a given
-time-frame by checking which commits were merged during that time-frame.
-
-This paper makes two contributions; first, we describe a method of converting the
-DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach the master branch; second, we
-present a method to inspect and visualize the history of merges in the Linux project
-using the merge-tree model.
-
-These methods and visualizations are implemented in a web-based tool called
-\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
-Our visualizations and tool provide information about the location of any given
-commit or merge in its respective merge-tree, the files edited, the modules edited,
-and the commit message. \tool allows users to apply various filters, including the
-release version, along with a keyword or phrase from the log preview, the name of
-the author, or the commit ID. The user can request all merges made by Linus that
-contain a commit or inner merge that matches the search query, or just the commits
-and merges that match the query.
-
-\section{\mt Model}
-\label{sec:mt_model}
-
-We propose a new model, built from the DAG, with the purpose of
-providing a clearer conceptual image of how a commit is merged, other
-commits that are merged with a commit, and a summarization of the entire
-group of commits that are merged with a commit. We call our model the
-\mt model. The model inverts the DAG, so that reference point from
-parent to child instead of from child to parent. This enables the
-aggregation of commit metrics at merges. Given a merge, we are able to
-recursively traverse the children of the merge, collecting the
-statistics at each commit. The commits become the leaves of the tree,
-and internal nodes are merges that do not merge into the master branch
-of the repository. We terminate the further traversal of an edge under
-two conditions; first, if the next event is on the master branch of the
-repository, second, if the next event has a shorter path into the master
-branch of the repository. As a consequence, commits will always take the
-shortest path, in terms of commit time, to the master branch.
-
-A short example: assume the commits represented in
-Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
-In this case, commits are performed in various repositories and
-branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
-the master branch and the repository that the master branch is part of.
-The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
-Note that the direction of the edges of the DAG have been inverted,
-instead of pointing from the child to the ancestors, it points from the
-ancestor to its successors, forming a path to the master branch. Also
-note that the DAG has been simplified, showing only a single edge on the
-path to master for any commit.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
-    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-
-    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
-    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
-    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
-    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
-
-        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
-    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
-  \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
-  \label{fig:repoEvents}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state] (1) {1};
-    \node[state, above right= of 1] (2) {2};
-    \node[state, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, right= 2cm of 1] (4) {4};
-    \node[state, above right=of 4] (5) {5};
-    \node[state, above right=of 5] (6) {6};
-    \node[state, right=of 5] (7) {7};
-    \node[state, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, right= 2cm of 7] (9) {9};
-    \node[state, right=of 9] (10) {10};
-    \node[state, below right=of 9] (11) {11};
-    \node[state, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[-stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[-stealth] (9);
-    \draw (10) edge[-stealth] (9);
-    \draw (9) edge[-stealth] (8) edge[-stealth] (6)
-              edge[-stealth] (7);
-    \draw (8) edge[-stealth] (7);
-    \draw (7) edge[-stealth] (5);
-    \draw (6) edge[-stealth] (5);
-    \draw (5) edge[-stealth] (3) edge[-stealth] (2)
-              edge[-stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[-stealth] (2);
-    \draw (2) edge[-stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
-    information about which repository a commit comes from. Furthermore,
-    the DAG does not distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (2) edge[chartyellow, -stealth] (5)
-          (3) edge[chartmagenta, -stealth](5)
-          (5) edge[chartyellow, -stealth] (7)
-          (7) edge[chartyellow, -stealth] (9)
-          (6) edge[chartred, -stealth] (9)
-          (8) edge[chartmagenta, -stealth] (9)
-          (9) edge[chartyellow, -stealth] (11)
-          (10) edge[chartyellow, -stealth] (12);
-
-  \end{tikzpicture}
-  \caption{\mt view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
-\vspace{-3mm}
-\end{figure}
-
-Computing the \mt from a DAG for any repository may not be
-possible; however, certain features of the development process of Linux
-make it feasible to compute the \mt for the Linux repository.
-First, the master branch of Linux is maintained by Linus Torvalds, and
-only Linus has write access to it. We have verified this assertion in
-previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
-identifies the commits made directly to the master branch; whereafter it
-recursively determines the shortest path (in terms of time), using the
-DAG, from each commit to the master branch using the inverted DAG.
-
-\begin{algorithm}
-        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
-        \label{alg:original}
-        \begin{algorithmic}
-                \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
-                \State $head \gets \textit{Head of master of git repository}$
-                \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
-                \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
-                \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
-                \If {\textit{cid in master}}
-                \State \Return 0
-                \EndIf
-                \State    $d \gets 	\infty$
-                \State {\# Traverse the inverted DAG}
-                \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
-                \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
-                \EndIf
-                \EndFor
-                \State {\# $c$ is the commit that follows $cid$}
-                \State {\# in its way to master}
-                \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
-                \EndFunction
-
-                \State {\# Compute the distance for each commit}
-                \State {\# discarding result}
-                \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
-                \EndFor
-                \State \Return $Tree$
-                \EndFunction
-        \end{algorithmic}
-\end{algorithm}
-
-We evaluate the performance of the heuristic using the merge log made by
-Linus Torvalds, the original creator of Linux, and primary merger of all
-other repository events into the master branch of the repository. In
-Figure~\ref{fig:sampleMerge}, we see an example of the merge log.
-
-\begin{figure}[htbp]
-  \centering
-  {\fontsize{7}{9}
-    \begin{verbatim}
-Merge: 8cbd84f fd8aa2c
-Author: Linus Torvalds <torvalds@linux-foundation.org>
-Date:   Tue Aug 10 15:38:19 2010 -0700
-
-Merge branch 'for-linus' of git://neil.brown.name/md
-
-* 'for-linus' of git://neil.brown.name/md: (24 commits)
-md: clean up do_md_stop
-[... edited for the sake of space]
-md: split out md_rdev_init
-md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache...
-...
-    \end{verbatim}}\vspace{-5mm}
-  \caption{Example of how merges record a subset of commits being
-    merged. The commit only shows the first 20 one-line summaries
-    messages for the 24 non-merge commits it merged. The ending
-    ``\ldots'' is part of the log and represents that other
-    commits were merged.}
-  \label{fig:sampleMerge}
-\end{figure}
-
-We use this information to evaluate the accuracy of the \mt model
-extracted from the DAG\@. The method we followed started with the
-extraction of the commit history up to July 20, 2016. We computed the
-merge tree of every commit until then. Since Linus Torvalds mostly does
-merging directly into master, we assumed that every merge by him is the
-root of a \mt. As described above, the log of a merge-commit usually
-contains the number of commits in the merge the first 20 summaries of
-commits being merged. We extracted merges by Linus Torvalds using the
-command \mycode{log --merges --author='Torvalds} and compared the number
-of commits according to the log with the number of commits in the  \mt
-rooted in this commit. We also used the summaries of the commits found
-in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the
-merge in Figure~\ref{fig:sampleMerge} we would expect that the merge
-tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
-summaries corresponds to commits in that \mt. We also inspected those
-with differences to make sure they were true errors.  The results can be
-summarized as follows:
-
-\begin{itemize}
-
-  \item
-
-    Five merges were false-errors because their logs did not contain
-    accurate information (were probably edited by hand). For example in
-    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
-    empty), in \mycode {c55d267\ldots} the summaries were reordered.
-
-  \item
-
-    The heuristic correctly identified that 79 of Linus merges (between
-    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
-    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
-    commits.
-
-  \item
-
-    The heuristic worked perfectly until Sept 4, 2007, the earliest date
-    that it could be verified.  Before this date, and until Dec 12,
-    2006, merges did not include a summary of the commits they included,
-    hence making it impossible to verify; during this period, however,
-    we correctly identified the merges by Linus into master.
-
-  \item
-
-    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
-    which confounded the true master branch of a repository (see
-    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
-    for a description of the issue).
-
-\end{itemize}
-
-In summary, of the merges after Sept. 4, 2007, our heuristic was correct
-in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
-12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
-2007).
-
-\section{Design and Implementation}
-\label{sec:design_and_implementation}
-
-We implement a tool, \tool, to test and verify the capabilities of the
-\mt model. \tool provides search, summarization, and
-visualization using the \mt model.
-
-We provide users with search-based navigation. The search engine takes a
-plain text query from the user and returns results based on similarity
-to the query. The engine weighs various fields, including the commit
-log, the author name, the commit hash, and the dates that the commit was
-authored and committed. The results are aggregated by the \mt
-with a link to the root at the top, and the repository events that match
-the search query in a table below.  An example of a single \mt is shown
-in Figure~\ref{fig:linvis_search_results}.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/search_results.png}
-  \caption{\tool search results for `net-next 2011', showing a single
-    \mt group.}
-  \label{fig:linvis_search_results}
-\end{figure}
-
-\tool uses seven tabs to show the information and visualizations for a
-selected commit or merge. The message tab shows the full commit log
-message. This does not include the diff. The files tab shows an
-aggregated table of all files that were modified in a merge. It includes
-metrics like the number of lines added, lines removed, total lines
-modified, and the delta, summed across all commits in the merge tree
-that modify this file. A small details drop-down button allows a user to
-see exactly which commit makes the changes, as shown in
-Figure~\ref{fig:linvis_files}. If the current repository event being
-viewed is a commit, the aggregate views will only show modifications
-made by the commit. The modules tab shows the modules modified in the
-merge tree. Like the files tab, the modules tab uses a table to show the
-name of the module, the number of commits that are in the merge tree
-that work with the module, and a details button to provide the links to
-those commits.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_files.png}
-  \caption{Table showing the modified files in a merge, with the second
-  entry expanded to show the commit that makes the changes.}
-  \label{fig:linvis_files}
-\end{figure}
-
-The authorship tab is very similar to the files tab, but shows the
-authorship information. It shows the sum of the number of lines added,
-removed, modified, and the delta within the \mt. It also shows
-the number of files that were modified by the author. The details tab is
-organized slightly differently. Instead of organizing these results by
-commit, the details are organized by file, showing which files were
-modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_authors.png}
-  \caption{Table showing the authors who made changes in a merge. The
-    entry for Randy Dunlap is expanded, showing the modifications to
-    each file that were made by Randy.}
-  \label{fig:linvis_authors}
-\end{figure}
-
-\tool implements three visualizations, a list tree, pack tree, and
-Reingold-Tilford tree. The list tree is a series of indented textual
-lists that can be easily searched with the built-in search in most
-web-browsers. The list tree is rooted at the current event, if the event
-is a commit, only that commit will be shown; conversely, if the selected
-event is the root, the entire merge tree will be shown. The Pack-Tree or
-Bubble tree, uses a set-based approach for showing the parent
-relationship. The root node aggregates all of the events within the \mt
-into the master branch. Thus, the root is the superset of all the other
-nodes in the tree. The commits are shown as individual white circles,
-containing no other events, as these are the leaves. This is shown in
-Figure~\ref{fig:linvis_pack}. The selected event will be shown in bright
-orange, commits in white, and merges in shades of blue. Darker shades of
-blue indicate that the merge is deeper in the tree.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
-  \caption{The pack tree visualization for a selected commit, shown in
-    orange.}
-  \label{fig:linvis_pack}
-\end{figure}
-
-The Reingold-Tilford tree is the classic tree visualization, with the
-root at the top, leaves at the bottom, and the edges between them
-showing the parent relationship. An example of the Reingold-Tilford tree
-visualization is shown in Figure~\ref{fig:commit_1} and
-Figure~\ref{fig:commit_2}.
-
-\section{Related Works}
-\label{sec:related_works}
-
-\input{relatedwork.tex}
-
-\section{Controlled User Study}
-\label{sec:study}
-
-We use a controlled mixed-methods users study in order to evaluate the
-ability for the DAG to provide users with a conceptual image of how a
-commit is integrated into a project, to compare the DAG and \mt models,
-and determine which model the users preferred.
-
-In this section, we describe the methods used to for designing and
-implementing the user study.
-
-\subsection{Questions and Tasks}
-\label{sub:questions}
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Conceptual Tasks }
-  \label{tab:conceptual_tasks}
-  \begin{tabulary}{0.9\textwidth}{LL}
-    \toprule
-    Task & Description\\
-    \midrule
-    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
-    T2 & How many individual commits are related to this commit?\\
-    T3 & How many merges are involved with merging this commit into the master branch?\\
-    \bottomrule
-  \end{tabulary}
-\end{table*}
-
-\begin{table}[htpb]
-  \centering
-  \caption{Summarization Tasks}
-  \label{tab:summarization_tasks}
-  \begin{tabulary}{\linewidth}{LLL}
-    \toprule
-    Task Set   & Task & Description\\\midrule
-    Merge      & T4   & What is the series of merges involved with merging this
-    commit?\\
-               & T5   & What other commits are merged?\\
-    Authorship & T6   & How many authors are involved?\\
-               & T7   & Who contributed the most changes?\\
-    Files      & T8   & How many files were modified?\\
-               & T9   & Which file had the most changes?\\
-    Modules    & T10  & Which modules does this merge tree involve?\\
-    \bottomrule
-  \end{tabulary}
-\end{table}
-
-The questions and tasks are broken down into four groups. The first two
-groups are purely quantitative in nature; we begin with conceptual tasks
-to understand whether the DAG is capable of providing users with an
-understanding of how a commit is merged into the master branch, and
-other commits that it is merged with. We continue with summarization
-tasks, to quantitatively understand if the \mt model makes a difference
-in summarization tasks.
-
-The third task set is purely qualitative, to gain better understanding
-of whether participants preferred working with the \mt model in \tool,
-or working with the DAG in Gitk, and what aspects they preferred from
-each tool. The final group of questions is simply to gain a better
-understanding of the demographic of our participants.
-
-\textbf{Conceptual Tasks}
-
-The participants were allowed to use Gitk or the git command line tools
-to perform these tasks. These tasks are asked in the specified order,
-allowing tabularr to first, create a conceptual understanding of the
-commits and merges that will be involved with the rest of the tasks that
-follow.tabularmay then use their diagrams to answer T2 and T3. We do not
-collect metrics from \tool for the conceptual tasks, opting instead to
-determine if the DAG is sufficient for understanding how a commit is
-integrated. \tool uses the \mt model show a tree visualization of how
-commits are merged, until the commit reaches the master branch.
-Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.
-
-We provide the participant with 10 minutes per commit to complete the
-first task. The answers to the two other questions in this section are
-drawn from the answer in the first task.
-
-\textbf{Summarization Tasks}
-
-The summarization tasks (Table~\ref{tab:summarization_tasks}) are
-grouped into task sets by theme. The tasks within the task set are
-shuffled, and the task sets are shuffled, to ensure that similar
-questions are together, but not in the same order, and to ensure that
-the themes are in randomized order between participants. The order of
-the commits is shuffled, as well. The order of the tool is randomized,
-choosing to start with either \tool or Gitk, to ensure that we do not
-place a bias on one tool through the experiment.
-
-As understanding where the merge begins and ends is vital to answering
-the other question, the merge task set will always be the first task set
-presented; however, the questions will still be re-ordered within the
-task set.
-
-\textbf{User Opinion and Profile}
-
-Finally, we ask the user for their opinions on the tools and for some
-information about their level of experience with git. This portion of
-the study is qualitative and does not rely on either of the tool,
-instead attempting to determine which model provided the user with a
-better experience.
-
-We do not shuffle the questions in this set of questions, as this should
-not have an impact on the opinions of the participant at this point.
-These questions are designed to provide us with a better understanding
-of preference, and insight on why, the participant had that preference.
-
-We ask the following questions in this order;
-
-\begin{enumerate}
-  \item Given these tasks again, which tool would you prefer to use?
-  \item Which aspects of each tool did you like and why?
-  \item How long have you used git?
-  \item If you have used git, for what kind of projects? (personal,
-    school courses, professional?)
-  \item If you have used git, how many commits, files, and contributors
-    were involved with the largest repository you have worked with?
-\end{enumerate}
-
-The last three questions in this section are aimed at getting a better
-understanding of the demographic of our participants.
-
-\textbf{Procedure}
-
-For each participant, we run a simple python script to generate the
-script for the study, eliminating the possibility of bias from us. The
-script randomizes the order of the tools, the order of the commits, the
-order of the task sets, and the order of the tasks within each task set,
-where appropriate.
-
-\subsection{Commit Selection}
-\label{sub:commit_selection}
-
-We chose two commits for use in the study. The order that the commits
-are presented is randomized between participants, but the order is kept
-consistent through the tasks described in the next subsection.
-
-We use the \tool database, containing commits and merge-tree information
-from April 16th 2005 to October 14th, 2014, which corresponds to being
-between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
-  database even more?} We chose the commits based on tree sizes. We
-found that a majority of the trees contain at most seven inner commits,
-while more than 25\% of the trees contain a single commit. 75\% of the
-tree contained up to 51 inner commits and merges, and finally the
-largest tree contained 7217 nodes. We chose to work with the trees in
-the first and second quartile, as merge trees of sizes between one and
-seven, not including the merge into the master branch, make up the
-majority of the trees in the database.
-
-From here, we selected one tree from the trees of a single commit at
-random. Selecting a commit from that tree is trivial, as there is only
-one to choose.
-
-We selected one tree of size seven at random from all trees of size
-seven. We placed a restriction the trees, the selected tree must include
-at least one internal merge to increase the complexity of the trees
-tested. After randomly selecting a tree, we chose in internal commit at
-random, using the \verb|random.choice| function in python 3.6.1.
-
-Using this technique, we selected commit
-\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
-a single node (Figure~\ref{fig:commit_1}), which we will refer to as
-\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
-the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
-will refer to as \comB.
-
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
-  \caption{The first merge tree used in the user study, a merge tree
-    containing a single commit}
-  \label{fig:commit_1}
-\end{figure}
-
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
-  \caption{The second merge tree used in the user study, a merge tree
-    containing seven commits}
-  \label{fig:commit_2}
-\end{figure}
-
-\subsection{Results}
-\label{sec:results}
-
-We recorded the audio and screen capture video for each participant in
-the study for further analysis. After capturing the information, we
-extracted various metrics from the videos. From the conceptual tasks, we
-recorded the time in seconds, and the answer for the question. We
-extracted three metrics for the summarization tasks, the time in
-seconds, the correctness, and the accuracy. Correctness is a binary
-value, it is true if the response was correct and false if the response
-was incorrect. The accuracy is measures in distance from the correct
-value. An accuracy of zero indicates a correct answer. All tasks in the
-summarization tasks include an accuracy metric except T7, which will
-only be right or wrong, and therefore cannot meaningfully be measured by
-a distance.
-
-\subsection{Conceptual Tasks}
-\label{sub:conceptual_tasks}
-
-In Table~\ref{tab:conceptual_results}, we are able to see an overview of
-the results from the conceptual questions.
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Results from the conceptual questions}
-  \label{tab:conceptual_results}
-  \begin{tabular}{ll|r|lrr|rrr}
-    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
-    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
-    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
-    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
-    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
-  \end{tabular}
-\end{table*}
-
-Users were able to more closely estimate the number of commits and
-merges in the larger tree, but generally took longer than the smaller
-tree. The tree with a single node resulted in more variability in the
-estimate of number of commits.
-
-It should be noted that these questions were answered after spending
-roughly ten minutes attesting to draw a picture that held the answers to
-these questions.
-
-\subsection{Summarization Tasks}
-\label{sub:summarization_tasks}
-
-In Figure~\ref{fig:summarization_results}, we see all of the
-correctness, accuracy, and timing results. The first row of plots are
-the correctness results. We see in all cases except for T10, that the
-number of instances of a correct result with \tool is roughly the same
-as the number of incorrect instances with Gitk. We test each task with
-the McNemar Chi-square test~\cite{McNemar1947}.
-
-\begin{figure*}[htpb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
-  \caption{Results from the summarization tasks. The first row plots the
-    correctness metric for each task. The second row plots the
-    accuracy metric for each task. The third row plots the time metric
-    for each task.}
-  \label{fig:summarization_results}
-\end{figure*}
-
-With the McNemar's tests, we analyze the participants that change from
-being correct with one tool to being incorrect with the other. This does
-not take into account the number of people who were able to provide the
-correct answer for both tools, nor the people who were incorrect for both
-tools. The results show whether providing the participants with \tool
-over the current standard, gitk, improves the correctness of the
-summarization tasks among the participants. We show the results for the
-number of people who were, with both tools, correct, incorrect, and
-changed in Table~\ref{tab:correctness_table}.
-
-In all cases, except for task T10, we saw statistically significant
-results suggesting that we reject the null hypothesis. We tested each
-correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
-degree of freedom is $7.879$.
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Correctness tables showing how many people stayed correct,
-    incorrect, and how many people changed between tool}
-  \label{tab:correctness_table}
-  \begin{tabular}{c|c|c}
-
-    T4 & T5 & T6\\
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 15                           & 2\\
-  \end{tabular} &
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
-                           & Incorrect & 14                           & 4\\
-  \end{tabular}
-
-  &
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
-                           & Incorrect & 16                           & 2\\
-  \end{tabular}
-
-  \\\hline
-  T7 & T8 & T9\\
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 8                            & 0\\
-                           & Incorrect & 13                           & 2\\
-  \end{tabular}  &
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 15                           & 2\\
-  \end{tabular} &
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 14                           & 3\\
-  \end{tabular}\\\hline
-
-  & T10 & \\
-  & \begin{tabular}{cc|rr}
-                            &           & \multicolumn{2}{c}{Linvis}\\
-                            &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
-                            & Incorrect & 8                            & 0\\
-  \end{tabular} & \\
-
-  \end{tabular}
-\end{table*}
-
-The correctness test statistics and conclusion for each summarization
-task;
-
-The correctness test statistics, conclusions, and a description of the
-accuracy and timing results for each summarization task follow:
-
-\begin{itemize}
-
-  \item
-
-    T4; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the series of merges that a commit
-    takes, with $99.5\%$ confidence.
-
-    There is almost no variation in accuracy between participants using
-    \tool, while there is more variance in Gitk. There were two outliers
-    among the instances using \tool, which lie within the second and
-    third quartiles for the accuracy measure of Gitk. The furthest
-    outlier for \tool lies at a distance that is slightly greater than
-    the median for Gitk.
-
-    We see similar results in the timing metric; all significant data
-    occurs within less time than the second, third, and fourth quartile
-    of the Gitk results. Nearly all outlier instances in \tool are
-    within the third quartile group of the instances of Gitk with the
-    exception of the most extreme, which lies in the fourth quartile
-    group.
-
-    The results indicate that \tool is able to assist users with
-    determining how a commit is merged more accurately, and more
-    quickly.
-
-  \item
-
-    T5; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the other related commits, with
-    $99.5\%$ confidence.
-
-    There is very little variance in the accuracy of results for \tool
-    compared to the accuracy of the results in Gitk. The median distance
-    is smaller, indicating that the participants performed more
-    accurately using \tool.
-
-    There is less variance in the time taken to come to an answer with
-    \tool than with Gitk. Furthermore, we see that a majority of the
-    participants were able to complete the task in \tool in much less
-    time than 50\% of the participants using Gitk.
-
-    The results indicate that \tool is able to assist users in
-    understanding which commits are related to a given commit.
-
-  \item
-
-    T6; the resulting $\chi^2$ value is $16$. $16 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the number of authors involved with
-    the merge tree, with $99.5\%$ confidence.
-
-    We see the same results from this task as with the previous tasks,
-    suggesting that \tool assists users determine the number of authors
-    more quickly and more accurately.
-
-  \item
-
-    T7; the resulting $\chi^2$ value is $13$. $13 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining who contributed the most changes,
-    with $99.5\%$ confidence.
-
-    This question does not include an accuracy metric, as there are only
-    two possible outcomes. We see the same result in the timing metric
-    as we did from the previous questions.
-
-    The results suggest that \tool is able to help users correctly
-    identify the person who contributed the most lines of code to a
-    merge tree in less time.
-
-  \item
-
-    T8; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining how many files were contributed,
-    with $99.5\%$ confidence.
-
-    The results of this task are the same as in the previous tasks.
-    This suggest that \tool is able to assist users determine how many
-    files were modified, more quickly and more accurately.
-
-  \item
-
-    T9; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining which file had the most changes,
-    with $99.5\%$ confidence.
-
-    While this question appears to have a similar answer to the correct
-    answer for T7, the answer \comA had two files with the same number
-    of midifications, thus the participant must correctly identify both
-    of these files to be correct. For this reason, we include an
-    accuracy metric.
-
-    The results of this task are the same as the previous tasks.
-    This suggests that \tool is able to assist users to determine which
-    file had the most changes.
-
-  \item
-
-    T10; the resulting $\chi^2$ value is $5.44$. $5.44 < 7.879$,
-    therefore we do not reject the null hypothesis, suggesting that
-    \tool does not make a difference in determining which modules are
-    involved with a merge tree.
-
-    The accuracy and timing results for this task are the same as the
-    others, showing that there is less variability in both accuracy and
-    the time taken to have an answer between participants using \tool
-    than using Gitk. Looking Table~\ref{tab:correctness_table}, we can
-    see that, unlike the other questions, a majority of the participants
-    were correct using both tools.
-
-\end{itemize}
-
-\subsection{User Opinions}
-\label{sub:user_opinions}
-
-Among the 12 pariticipants, there was nearly unanimous agreement that
-for conceptual understanding and summarization tasks, \tool was easier
-to work with than Gitk. The participants cited the ability to abstract
-information about the merge from the clean summarization tables and
-simple visualizations. Three participants suggested that someone with a
-professional understanding of Gitk and the git commandline may be to
-extract the a good conceptual understanding from the DAG, and perform
-the summarization tasks from the commandline. One of these three
-participants said that they would prefer to have both tools available,
-as they are able to complement eachother.
-
-\section{Discussion}
-\label{sec:discussion}
-
-% TODO: Go back to the conceptual task results and make them stronger
-Overall, the results show fairly conclusive results; the DAG is unable
-to adequately provide users with an explanation of the changes being
-made in the repository, the DAG is not easily understood, and the DAG is
-unable to provide summarizations of the changes being made at a merge.
-Meanwhile, the \mt model is able to provide a better explanation of how
-a commit is integrated, along with better summarization capabilities. In
-this section, we discuss some observations made during the study,
-threats to the validity of the study, comments made by the participants
-during the study, modifications to the original algorithm to help
-generalize it for more repositories, and future work.
-
-\subsection{Observations}
-\label{sub:observations}
-
-% TODO: include info on merge tree sizes
-
-During the course of the study, we noticed a consistent theme among the
-participants. The most difficult challenge was identifying the master
-branch of the repository. Given that single-commit merge trees are the
-most common tree, and that they are relatively simple, we were
-interested in seeing if the participants would be able to identify the
-tree easily, but the results suggest that this was not the case. The
-issue primarily stemmed from not being able to identify the master
-branch. Most of the participants realized that they should not be going
-backward, toward the initial commit, very early on. Furthermore, many
-would identify the correct commit; however, they would start listing
-every merge along the master branch as being a merge in the tree, and
-including the commits from that merge. In the case of the single commit,
-simply identifying the master branch would most likely be sufficient in
-providing a clearer understanding of how a commit or flat tree is
-integrated into the master branch of the repository.
-
-While, in general, the participants of the study preferred using \tool
-to Gitk, there were some noticeable usability issues. The most prominent
-issue being the navigation within the tree. The participants expected
-that the tabs would update when they clicked on a node in the tree,
-instead of having to click the link to the commit after clicking the
-node. This was true in both the Reingold-Tilford tree and the Pack tree.
-The second major usability issue was the inconsistency between the list
-tree and the other tree visualizations. The list tree is rooted at the
-repository event that the user is current looking at, while the other
-tree visualizations show the entire tree, and use bright orange to
-highlight the current node.
-
-\subsection{Comments From a Release Manager}
-\label{sub:comments_from_a_release_manager}
-
-One of the participants of the study had worked as a release manager for
-more than three years, working with both SVN and CVS repositories, but
-had little experience with git.
-
-Contributors making merges need to understand more than just what merges
-a commit was collected into before reaching the repository of the
-contributor. It is also important to understand order that the related
-commits were made, as they tell the story of what the developer was
-thinking as they were writing the changes. The current merge-tree model
-does not order the commits, simply placing them randomly. This is the
-primary reason behind this participant's request to use both
-simultaneously. \tool is able to help with the aggregation of the
-information, and provide a better understanding of the next merge
-involved in integrating this commit, but the DAG in gitk provides the
-full story of the commit instead of hiding it behind a layer of
-abstraction.
-
-The comments from this participant were very insightful, and will assist
-in improving the model in the future.
-
-\subsection{Threats to Validity}
-\label{sub:threats}
-
-While git includes gitk and the git log provides a view of the DAG with
-\verb|git log --graph|, many of the participants in our study were
-unfamiliar with the DAG visualizations shown by these tools. A possible
-reason for this may be that unless a repository reaches a sufficient
-complexity, both in branching and number of commits, these tools are
-necessary for the normal operation of git. Furthermore, fear of the
-complexity caused by branching may cause personal and academic
-repositories to have relatively simple structures. While they were
-unfamiliar with the DAG structure, these users were also unfamiliar with
-the \mt model, and were still able to better understand and summarize
-the repository using \tool.
-
-Each question was evaluated independently of the results from the
-previous question in the summarization tasks. Had we modified the
-correct answer to fit the conceptual understanding of which commits were
-included in the tree, we may have witnessed different results. A
-modification to the study may include providing the participants with
-the correct set of commits prior to the summarization tasks, after the
-merge task set.
-
-\subsection{Generalization}
-\label{sub:generalization}
-
-Seeing that the \mt model is able to assist users with understanding how
-commits are integrated, we continue by looking at ways of improving the
-algorithm for faster operations, and making the conversion from DAG to
-\mt feasible over the internet for smaller repositories.
-
-The problem of finding the master branch is the most difficult, and is
-not possible without heuristic approaches. The master branch can be
-confounded by \textit{foxtrot} merges, which change the order of the
-parents in the parent list of a merge. Other than the ordering of the
-parent list, git has no other way to determine which commits are on the
-master branch of the repository. For the purpose of this discussion, we
-will assume that we can rely on the order of the parent list to be
-correct, and that foxtrots do not occur.
-
-\begin{algorithm}
-  \caption{Computing the generalized Merge Tree}
-  \label{alg:generalized}
-  \begin{algorithmic}[1]
-    \Function{Phase 1}{initial commit $root$} : updated tree
-    \State $depth \gets 0$
-    \State $Q \gets root$
-    \Do
-    \State $cur \gets Q.dequeue$
-    \State $parents \gets cur.parents$
-    \State $depth \gets depth + parents.length - 1$
-    \For{$index, parent \in parents$}
-    \If{$parent.depth$ is $undefined$}
-    \State $parent.depth \gets cur.depth + index$
-    \ElsIf {$cur.depth > parent.depth$}
-    \State $depth \gets depth - 1$
-    \ElsIf {$cur.depth < parent.depth$}
-    \State $depth \gets depth - 1$
-    \State $parent.depth \gets cur.depth$
-    \EndIf
-    \State $parent.children \gets cur$
-    \If{$parent$ has not been visited}
-    \State $Q \gets parent$
-    \State $parent.visited \gets True$
-    \EndIf
-    \EndFor
-    \doWhile{$Q$ not empty and $depth \ne 0$ }
-    \EndFunction
-
-    \Function{Phase 2}{initial commit $root$} : Merge tree
-    \State $tree.root \gets root$
-    \State $Q$ // Empty Queue
-
-    \State $parents \gets root.parents$
-    \State $parents \gets tail\ parents$
-    \For {$parent \in parents$}
-    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
-    \If {$realParent$ is $root$}
-    \State $root.children \gets parent$
-    \State $parent.parent \gets root$
-    \State $Q \gets parent$
-    \EndIf
-    \EndFor
-
-    \While{$Q$ not empty}
-    \State $cur \gets Q.dequeue$
-    \State $parents \gets cur.parents$
-    \For{$parent \in parents$}
-    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
-    \If {$realParent$ is $cur$}
-    \State $cur.children \gets parent$
-    \State $parent.parent \gets cur$
-    \State $Q \gets parent$
-    \EndIf
-
-    \EndFor
-    \EndWhile
-
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-The primary difference between the original algorithm and the
-generalization is the main data structure. Algorithm~\ref{alg:original}
-is centered around the use of the stack, making it akin to a depth-first
-traversal of the DAG.\@ The generalization in
-Algorithm~\ref{alg:generalized} is based on the queue, making it akin to
-the breadth-first traversal of the DAG.\@ While the final results of the
-two traversals should not be different, the runtime properties are
-different.
-
-We will begin with a simplified description of the original algorithm.
-We start at the master branch pointer. We will use the term master
-branch loosely here, meaning any branch that has been labeled in some
-way as being the main branch, this branch may have a different name, but
-standard naming conventions call it the master branch. Starting from the
-master branch label, we traverse the entire master branch to the initial
-commit, recording the commit hash of every repository event encountered.
-
-The second phase iterates over every commit, computing the shortest
-distance to the master branch. For this reason, the entire repository
-must be available to the algorithm. This is a dynamic programming
-solution, centered around the call stack. The next node on the shortest
-path to the master branch will become the parent of this node in the
-tree.
-
-If this algorithm is to be generalized into a dynamic, web-based, tool,
-it may not be feasible to download every commit in the repository in a
-reasonable period of time. We instead look toward a breadth-first
-approach at limiting the number of repository events that must be
-downloaded.
-
-Focusing on the generalized algorithm~\ref{alg:generalized}, the
-algorithm is broken into two phases. The first phase is where the
-improvements take place. The first phase has three main goals, the first
-is to track the master branch. The second goal labels the depth of each
-repository event that is traverses. The final goal is to record the
-children of each commit, reversing the DAG.\@ The first phase also
-begins the process of reversing the DAG, by providing the candidate
-commits that could possibly be the real parent to this node. The first
-phase can terminate early, if it reaches a point where all branches that
-it is following terminate. The depth metric on line 7 indicates how many
-branches are currently being traversed. In line 11 through 15, we
-decrement the depth, as this is the point where a branch occurs. The
-changing of the depth occurs if there are more events on the same branch
-than between the merge and branch point of a separate branch.
-
-\begin{table}[htpb]
-  \centering
-  \caption{Example traversal of the example DAG in
-    Figure~\ref{fig:repoDAG} starting at merge 11.}
-  \label{tab:ex_traversal}
-  \begin{tabular}{cccc|cc}
-    \toprule
-    Current & Parents & Depth & Q       & Visited & Children\\\midrule
-    -       & -       & 0     & 11      & 11:0    & \\
-    11      & 4,9     & 1     & 4,9     & 4:0     & 11, 5\\
-    4       & 1       & 1     & 9,1     & 9:1     & 11\\
-    9       & 7,6,8   & 3     & 1,7,6,8 & 7:1     & 9, 8\\
-    1       &         & 3     & 7,6,8   & 6:2     & 9\\
-    7       & 5       & 3     & 6,8,5   & 8:3     & 9\\
-    6       & 5       & 2     & 8,5     & 5:1     & 7, 6\\
-    8       & 7       & 1     & 5       & 2:1     & 5, 3\\
-    5       & 4,2,3   & 2     & 2,3     & 3:2     & 5\\
-    2       & 1       & 1     & 3       &         & \\
-    3       & 2       & 0     &         &         & \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-In Table~\ref{tab:ex_traversal}, we follow the algorithm starting at
-merge commit 11 from the example in Figure~\ref{fig:repoDAG}, recording
-the current node, the parents of the current node, the depth after
-updating, and the queue structure after updating. The visited column is
-not dependant on the current step, but keeps track of the nodes that
-have been visited, and the depth at that node. When we reverse the DAG,
-the children in this table will become the candidates for being the
-parent of the corresponding node. For example, the parent of node 7 may
-be resolved to either node 9 or node 8. In this case, the shortest path
-to the master branch is through node 9.
-
-The terminology surrounding the parent/child relationship gets very
-confusing at the change between phase one and two because this is where
-the roles switch. In phase 1, we are working with the DAG; in phase 2,
-we are working with a structure that is neither a tree nor a DAG, but
-can only be described as a directed graph.
-
-The goal of the second phase is to resolve the tree from the directed
-graph. We will refer to parents in the same sense as we would refer to
-them in the tree form of the repository for the remainder of this
-section. We have gathered the depths and possible parents (line 17), but
-we need to remove the extra parents. Extra parents are culled in line
-42 of the algorithm. Of the possible parents, we take the parent that is
-closest to the root, but not beyond. This forces the real parent to
-either be deeper in the tree, or on the same branch. Line 30 to 37 are
-for bootstrapping the queue, ensuring that we are not starting on an
-empty queue.
-
-The construction of the tree becomes trivial now that we are able to
-determine the parents and children for each node. Lines 44 and 45
-perform this construction, adding the DAG parent to the tree children of the
-current node, and the DAG parent's parent becomes the current node,
-which was the DAG parent's child. We continue following the path until
-we hit the root node, merging the commit into the master branch.
-
-This algorithm relies heavily on the order of parents in the parent
-list, but is able to trim the number of nodes that must be downloaded.
-Furthermore, the resulting tree is able to show the order of commits,
-telling the story of how a developer was working, before the set of
-commits was merged. We developed a small bitbucket plugin to test the
-design and determine whether it was capable of producing merge trees,
-shown in Figure~\ref{fig:b_plugin}.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=\linewidth]{figures/plugin.png}
-  \caption{Screen shot of the Bitbucket plugin showing the online tree
-    computation of the repository shown in~\ref{fig:repoDAG}.}
-  \label{fig:b_plugin}
-\end{figure}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
 In this paper, we develop a tree-based model designed to provide git
@@
 % TODO: You can tell that I am tired. This conclusion will need to be re-worked.
-
-\balance
-
-% \nocite{*}
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\end{document}

### 2017-07-07 paper/design.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

--- 2017-07-07 paper/lineval.tex
@@
-% vim:set et fenc=utf-8 ft=tex sw=2 ts=2 tw=72:
-
-\newif\ifdraft
-%\drafttrue
-
-\ifdraft
-        \documentclass[draft]{IEEEtran}
-        \def\baselinestretch{1}
-        % \setlength{\marginparwidth}{1.5cm}
-\else
-        \documentclass[conference]{IEEEtran}
-\fi
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[hyphens]{url}
-\urlstyle{same}
-
-%% Citations
-\usepackage[nospace]{cite}
-
-%% Table Support
-\usepackage{array}
-\usepackage{dcolumn}
-\usepackage{longtable}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage{tabulary}
-
-%% Extra support
-\usepackage{xspace}
-\usepackage{amsmath}
-\usepackage{balance}
-\usepackage{placeins}
-
-%% Algorithms
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-
-%% Graphics
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{xcolor}
-\usepackage{color}
-\usepackage{listings}
-
-\usetikzlibrary{positioning, arrows}
-
-\input{color.tex}
-\input{macros.tex}
-
-\newcommand{\TheTitle}{Merge-Tree: Visualizing the Integration of Commits into Linux}
-\newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
-\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Understanding a large number of commits and merges}
-\newcommand{\TheKeywords}{Linux, git, visualizations}
-
-%% Referencing
-\usepackage[unicode=true,
-          pdftitle={\TheTitle},
-          pdfauthor={\TheAuthors},
-          colorlinks=false]{hyperref}
-
-\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
-
-\lstset{frame=tb,
-  language=python,
-  aboveskip=3mm,
-  belowskip=3mm,
-  showstringspaces=false,
-  columns=flexible,
-  basicstyle={\small\ttfamily},
-  numbers=none,
-  numberstyle=\tiny\color{gray},
-  keywordstyle=\color{chartblue},
-  commentstyle=\color{chartred},
-  stringstyle=\color{chartgreen},
-  breaklines=true,
-  breakatwhitespace=true,
-  tabsize=3
-}
-
-\synctex=1
-
-\begin{document}
-
-\title{\TheTitle}
-
-\author{
-        \IEEEauthorblockA{\TheAuthors}
-        \IEEEauthorblockN{Department of Computer Science,
-                University of Victoria, Canada.}
-        \IEEEauthorblockA{Email: \TheEmails}
-}
-\maketitle
-
-\begin{abstract}
-
-  With an average of more than 900 top-level merges into the Linux
-  kernel per release, many containing hundreds of commits and some
-  containing thousands, maintenance of older versions of the kernel
-  becomes nearly impossible.  Various commercial products, such as the
-  Android platform, run older versions of the kernel. Due to security,
-  performance, and changing hardware needs, maintainers must understand
-  what changes (commits) are added to the current version of the kernel
-  since the last time they inspected it in order to make the necessary
-  patches.
-
-  Current tools provide information about repositories through the
-  directed acyclic graph (DAG) of the repository, which is helpful for
-  smaller projects. However, with the scale and number of branches in
-  the kernel the DAG becomes overwhelming very quickly. Furthermore, the
-  DAG contains every ancestor of every commit, while maintainers are
-  more interested in how and when a commit arrives to the official Linux
-  repository.
-
-  In this paper, we propose the merge-tree, a simplified transformation
-  of the DAG of the Linux git repository that shows the way in which
-  commits are merged into the master branch of Linux. Using the
-  merge-tree, we build \tool, a tool that is designed to allow users to
-  explore how commits are merged into the Linux kernel.
-
-\end{abstract}
-
-\begin{IEEEkeywords}
-  \TheKeywords
-\end{IEEEkeywords}
-
-\section{Introduction}
-
-Between 50k and 70k commits are added to the Linux kernel per version
-requiring maintainers of older versions of the kernel to sift through
-thousands of commits and merges with tools that are unable to filter and
-effectively visualize projects at the scale of the kernel. Older
-versions of the kernel are used in embedded systems and mobile phones;
-for security purposes, performance needs, and changing hardware
-requirements, maintainers must be able to understand the changes being
-made in the current version of the kernel in order to produce the
-necessary patches for the older versions of the kernel. Tools like Gitk
-use a directed acyclic graph (DAG) model of the repository, showing all
-commits and merges in chronological order by when the commit was
-authored, not by when it arrived in the official Linux repository.
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
-        \caption{A view of the Gitk interface centered on merge
-          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
-          their ancestors and there is no clear path from the commit to the merge
-          with the master branch. Neither Gitk nor Git are capable of showing the
-          commits in master.}
-        \label{fig:gitk}
-%\vspace{-4mm}
-\end{figure*}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its
-                relationship to other forks.}
-        \label{fig:gitfail}
-%\vspace{-2mm}
-\end{figure}
-
-This representation works in smaller projects; it enables users to see
-when changes are made, when these changes are merged, how each branch is
-interacting, and the point where a branch forks from the master branch.
-In large modular projects, like the Linux kernel, the DAG becomes a mess
-of merges and commits (Figure~\ref{fig:gitk}) losing its visual meaning.
-In some cases, the Linux kernel is simply too large for the system to
-generate a visualization; Github provides a DAG view for many projects,
-but is unable to display the visualization for projects at the scale of
-the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and 70k
-new commits are created for the Linux project every year; according to
-our previous work\cite{German2015}, a commit takes a median of 30 days
-from the time it is authored until it arrives in the official
-repository. The snapshot of the kernel tomorrow may be different than
-the snapshot from today, containing new commits authored in the past;
-distinguishing these new commits from the commits in the snapshot from
-today is not trivial.
-
-One major challenge with visualizing the arrival of commits to a
-repository is that Git does not store the date that a commit was merged
-into another branch, including the master branch. To complicate the
-problem, the DAG only has references to the ancestors of a commit (a
-model necessary for the operation of Git), but maintainers would prefer
-knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would
-imply that for any given merge, it would be possible to know which
-commits were merged. A user could inspect the commits that arrived into
-the master branch within a given time-frame by checking which commits
-were merged during that time-frame.
-
-This paper makes two contributions; first, we describe a method of
-converting the DAG of the Linux repository into a tree, or
-\emph{merge-tree} of the repository, that represents the path used by a
-commit to reach the master branch; second, we present a method to
-inspect and visualize the history of merges in the Linux project using
-the merge-tree model.
-
-These methods and visualizations are implemented in a web-based tool
-called \tool\footnote{\tool is currently available at
-  \url{http://li.turingmachine.org}}.  Our visualizations and tool
-provide information about the location of any given commit or merge in
-its respective merge-tree, the files edited, the modules edited, and the
-commit message. \tool allows users to apply various filters, including
-the release version, along with a keyword or phrase from the log
-preview, the name of the author, or the commit ID. The user can request
-all merges made by Linus that contain a commit or inner merge that
-matches the search query, or just the commits and merges that match the
-query.
-
-\section{Merge-Tree model}
-\label{sec:mergetree}
-
-Git uses a directed acyclic graph (DAG) as its main data model. In this
-model, a commit has one or more ordered parents (ancestors), except the
-root commits of a repository which do not have ancestors--Linux's git
-repository has two such commits.  Commits are divided into merge
-commits, merging two or more branches, and non-merge commits. Non-merge
-commits have only one parent, while merge commits have two or more. The
-order of the parents matter: the first parent is the branch in which the
-merge is being done, while the rest indicate the branches being merged
-(Git allows merging multiple branches simultaneously). The second parent
-is the first branch merged, the third parent is the second branched
-merged, etc.
-
-Once a commit is created, it is never changed. Git allows operations to
-alter commits or reorder them, but it changes the commit ID in the
-process, effectively replacing it with a new commit. This makes commits
-unable to record the traversal of merges from the commit to the merge
-into the master branch.
-
-A short example: assume the commits represented in
-Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
-In this case, commits are performed in various repositories and
-branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
-the master branch and the repository that the master branch is part of.
-The merge-tree view of this DAG is visible in Figure~\ref{fig:repoTree}.
-Note that the direction of the edges of the DAG have been inverted,
-instead of pointing from the child to the ancestors, it points from the
-ancestor to its successors, forming a path to the master branch. Also
-note that the DAG has been simplified, showing only a single edge on the
-path to master for any commit.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
-    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-
-    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
-    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
-    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
-    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
-
-        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
-    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
-  \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
-  \label{fig:repoEvents}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG loses information about which
-    repository the commit is performed in and through which merges it
-    has passed on its way to the master branch. The DAG does not even
-    distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (2) edge[chartyellow, -stealth] (5)
-          (3) edge[chartmagenta, -stealth](5)
-          (5) edge[chartyellow, -stealth] (7)
-          (7) edge[chartyellow, -stealth] (9)
-          (6) edge[chartred, -stealth] (9)
-          (8) edge[chartmagenta, -stealth] (9)
-          (9) edge[chartyellow, -stealth] (11)
-          (10) edge[chartyellow, -stealth] (12);
-
-  \end{tikzpicture}
-  \caption{Merge-tree view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
-\vspace{-3mm}
-\end{figure}
-
-\subsection{Computing the merge-tree of the DAG of Linux}
-
-Computing the merge-tree from a DAG for any repository may not be
-possible; however, certain features of the development process of Linux
-make it feasible to compute the merge-tree for the Linux repository.
-First, the master branch of Linux is maintained by Linus Torvalds, and
-only Linus has write access to it. We have verified this assertion in
-previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
-identifies the commits made directly to the master branch; whereafter it
-recursively determines the shortest path (in terms of time), using the
-DAG, from each commit to the master branch using the inverted DAG.
-
-\begin{algorithm}
-        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
-        \begin{algorithmic}
-                \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
-                \State $head \gets \textit{Head of master of git repository}$
-                \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
-                \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
-                \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
-                \If {\textit{cid in master}}
-                \State \Return 0
-                \EndIf
-                \State    $d \gets 	\infty$
-                \State {\# Traverse the inverted DAG}
-                \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
-                \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
-                \EndIf
-                \EndFor
-                \State {\# $c$ is the commit that follows $cid$}
-                \State {\# in its way to master}
-                \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
-                \EndFunction
-
-                \State {\# Compute the distance for each commit}
-                \State {\# discarding result}
-                \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
-                \EndFor
-                \State \Return $Tree$
-                \EndFunction
-        \end{algorithmic}
-\end{algorithm}
-
-\subsection{Evaluation}
-
-Merges that do not have conflicts provide information to verify this
-heuristic. If a merge does not contain a conflict, it records a summary
-of the commits that it merges. See Figure~\ref{fig:sampleMerge} for an
-example. This summary contains a list of the first 20 non-merge commits
-in the merge, including their one-line log description, the full logs of
-the merge commits that merge this subset, and the total number of
-non-merge commits in the merge.
-
-\begin{figure}[htbp]
-        \centering
-        {\fontsize{7}{9}
-        \begin{verbatim}
-Merge: 8cbd84f fd8aa2c
-Author: Linus Torvalds <torvalds@linux-foundation.org>
-Date:   Tue Aug 10 15:38:19 2010 -0700
-
-Merge branch 'for-linus' of git://neil.brown.name/md
-
-* 'for-linus' of git://neil.brown.name/md: (24 commits)
-md: clean up do_md_stop
-[... edited for the sake of space]
-md: split out md_rdev_init
-md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache...
-...
-        \end{verbatim}}\vspace{-5mm}
-        \caption{Example of how merges record a subset of commits being merged. The
-                commit only shows the first 20 one-line summaries messages for the 24
-                non-merge commits it merged. The ending ``...'' is part of the log
-                and represents that other commits were merged.}
-        \label{fig:sampleMerge}
-\end{figure}
-
-We used this information to evaluate the accuracy of the merge-tree
-model extracted from the DAG. The method we followed started with the
-extraction of the commit history up to July 20, 2016. We computed the
-merge tree of every commit until then. Since Linus Torvalds mostly does
-merging directly into master, we assumed that every merge by him is the
-root of a merge-tree. As described above, the log of a merge-commit
-usually contains the number of commits in the merge the first 20
-summaries of commits being merged. We extracted merges by Linus Torvalds
-using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits according to the log with the number of commits in
-the  merge-tree rooted in this commit. We also used the summaries of the
-commits found in the merge (not necessarily all---see above) to make
-sure those commits were in their corresponding merge tree. For example,
-for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
-merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
-one-line summaries corresponds to commits in that merge-tree. We also
-inspected those with differences to make sure they were true errors.
-The results can be summarized as follows:
-
-\begin{itemize}
-\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
-  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
-  {c55d267} the summaries were reordered.
-\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
-  branch (not master). This branch was merged at \mycode{3f17ea6d..} which contained 6809 commits.
-\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
-  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
-  impossible to verify; during this period, however, we correctly identified the merges by Linus into master.
-\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
-repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
-\end{itemize}
-
-In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860)
-our heuristic was correct. It failed in 1,542 commits before Dec. 12,
-2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
-
-\section{Visualizing the merge-tree of Linux}
-
-The goal of \tool is to simplify the navigation of the kernel commit
-information, specifically focusing on merges. This is done by leveraging
-the merge-tree to inspect how commits are merged on the path to the
-master repository.
-
-\subsection{Use cases}
-
-We designed \tool with two use-cases in mind, though a user may switch
-between the cases as they work.
-
-\noindent \textbf{Use-case 1: top-to-bottom
-  approach}\label{sec:usecase1}\\ These are users that are maintaining a
-section of the kernel and would like to pick a merge (including all the
-commits that it merges) and merge it directly into their current
-repository. This is useful for reducing the amount of re-implementation
-work. For these users, it is important to have the ability to aggregate
-metadata about files and modules being effected by the merge. Also, it
-is important for these users to be able to navigate from the root of the
-merge-tree toward the leaves.
-
-\noindent \textbf{Use-case 2: bottom-to-top
-  approach}\label{sec:usecase2}\\ These are users that start with a
-known merge or commit and would like to see what other changes are being
-made in commits that are in the same merge, including knowing the
-merge-tree they belong to. This is useful to see what other commits are
-related to the current commit and how they get collated into merges that
-eventually end in the master branch. This is primarily for maintainers
-that need to perform some specific cherry picking of commits. We must
-provide these users a mechanism for navigating from a single commit
-toward the master branch, allowing them to see other commits that might
-be related to their original commit.
-
-\subsection{Data Model}
-
-In our visualizations, we leverage the merge-tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already
-in the master branch, or is part of a tree which is rooted in the merge
-that merged it into the master branch.  Each commit, whether a merge or
-non-merge, has only one successor; the root of each tree has none as it
-was made by Linus Torvalds directly into the master branch. Non-merge
-commits contain the metadata for the changes made.  This metadata
-includes the files changed, the lines added and removed from each file,
-the author, the date the commit was merged into the merge that led to
-being merged into the kernel, the date the commit was authored, the
-patch, and the commit log. Merges contain less metadata, only storing
-the author of the merge, the log, the commit date, and the author date,
-and potentially, changes necessary to address conflicts during the
-merge. The details of the model are outlined in \cite{German2015}.
-
-\section{Design and Implementation}
+% vim:set et sw=2 ts=4 tw=72:
 
@@
 users to navigate from the root to the leaves and vice versa.
-
-\section{Discussion}
-
-From a developers point of view, the are two major disadvantages of the
-DAG model of Git: a) its edges point backwards, i.e. commits point to
-their ancestor, not to the commit that succeeds them on the path to
-being integrated; and b) the DAG contains many more edges that are
-necessary to understand how integration occurred. We addressed these two
-issues with the creation of the merge-tree model from the DAG of Linux.
-Effectively, the merge-tree recovers the details of how each
-branch---and the commits they contain---was merged into the master
-branch.
-
-In our experience, no other Git repository reaches the level of DAG
-complexity that Linux has. In most Git repositories, most merges are not
-nested---most merges merge directly into the master repository, and
-these merges only merge a few commits. Even in these simplified cases,
-the merge-tree can provide a valuable summary of how commits are
-integrated into the master branch, specifically since the time of
-integration may be very different from the time the commits were
-authored.
-
-\evan{Check accuracy of this paragraph}
-
-The biggest challenge is computing the merge-tree of any given
-repository. It is likely that our heuristic for computing the merge-tree
-will not work with other repositories. The heuristic requires that there
-are no fast-forward merges into the master branch of the repository (a
-majority of repositories), and that there are no foxtrot merges (a
-practice that is starting to be considered desirable by git
-users\footnote{See
-  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git}
-  and
-  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
-We are able to use our heuristic with the Linux kernel because of the
-strict integration model imposed by Linus. To validate the results of
-the heuristic, we use the continuous mining technique described in
-\cite{German2015}.  Continuous mining of the repository requires
-foresight and planning.
-
-Because \tool leverages the merge-tree model of the kernel, it provides
-mechanisms and visualizations that other tools are unable to produce. It
-gives us the ability to see how commits are being merged into the master
-branch, allowing us to better understand how code is integrated into the
-kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
-        \caption{Merge Dag View}
-        \label{fig:dag_view}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
-        \caption{Merge tree topology for the same merge as in
-                Figure~\ref{fig:dag_view}.}
-        \label{fig:tree_view}
-\end{figure}
-
-Our goal was to build a tool to enable maintainers to effectively
-navigate and browse the changes performed to the kernel over the period
-of a release. Achieving this goal includes removing information that
-does not pertain to the area of the kernel that the user is interested
-in. In Figures~\ref{fig:dag_view} and \ref{fig:tree_view} we can
-visually compare the results returned from Gitk and our tool for the
-top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6'' from kernel
-version 3.10. Note that Gitk has no way to show or recognize that this
-is a merge into the master branch.
-
-The log information to this merge-commit in both of these figures is
-identical, but the presentation drastically changes our ability to
-comprehend what we are seeing.  The primary difference is the removal of
-irrelevant information. In Gitk (Figure~\ref{fig:dag_view}), the
-visualization results include commits and merges from other components
-of the kernel, while our tool (Figure~\ref{fig:tree_view}) only includes
-the results that are specific to the component of the kernel that we are
-interested in.
-
-Even though this merge is relatively small, containing only four
-sub-merges, three of which only contain a single commit, and the last
-containing two commits, it is already hard to visualize in Gitk.  With
-\tool we are able to immediately see what section of the kernel these
-commits and merges pertain to. The DAG view of this provides almost no
-explanation, furthermore, users must work to determine where this merge
-ends and the next one begins.
-
-\tool further enhances our understanding by summarizing the files and
-modules that were edited in the entire branch being merged. We are able
-to determine that three files, ``palamas-regulator.c'' had two lines
-added and two lines removed, ``dbx500-prcmu.c'' had 12 lines added and
-12 lines removed, and ``core.c'' had 5 lines added and 2 lines removed.
-Finally, we are able to determine that only the ``regulator'' module was
-modified in this merge and was modified by 5 commits.
-
-\section{Related Works}
-\label{sec:related_works}
-
-\input{relatedwork.tex}
-
-\section{Controlled User Study}
-\label{sec:study}
-
-\dmg{start by describing what you mean by user study... to evaluate linvis we performed a user study...}
-
-\dmg{and you are evaluating the visualizations, NOT the DAG}
-
-We use a controlled mixed-methods users study in order to evaluate the
-ability for the DAG to provide users with a conceptual image of how a
-commit is integrated into a project, to compare the DAG and \mt models,
-and determine which model the users preferred.
-
-In this section, we describe the methods used to for designing and
-implementing the user study.
-\dmg{no, you describe the study, which includes the methods}
-
-\dmg{Start by giving an overall overview of the study: the study consists of a set of tasks that users are expected to
-  perform... do a top down description of what users will do
-}
-
-\subsection{Questions and Tasks}
-\label{sub:questions}
-
-\dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Conceptual Tasks }
-  \label{tab:conceptual_tasks}
-  \begin{tabulary}{0.9\textwidth}{LL}
-    \toprule
-    Task & Description\\
-    \midrule
-    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
-    T2 & How many individual commits are related to this commit?\\
-    T3 & How many merges are involved with merging this commit into the master branch?\\
-    \bottomrule
-  \end{tabulary}
-\end{table*}
-
-\begin{table}[htpb]
-  \centering
-  \caption{Summarization Tasks}
-  \label{tab:summarization_tasks}
-  \begin{tabulary}{\linewidth}{LLL}
-    \toprule
-    Task Set   & Task & Description\\\midrule
-    Merge      & T4   & What is the series of merges involved with merging this
-    commit?\\
-               & T5   & What other commits are merged?\\
-    Authorship & T6   & How many authors are involved?\\
-               & T7   & Who contributed the most changes?\\
-    Files      & T8   & How many files were modified?\\
-               & T9   & Which file had the most changes?\\
-    Modules    & T10  & Which modules does this merge tree involve?\\
-    \bottomrule
-  \end{tabulary}
-\end{table}
-
-\dmg{avoid editorializing your text: remove purely}
-
-\dmg{watch for tense... we don't begin... the tasks begin... or better, the first tasks...}
-
-The questions and tasks are broken down into four groups. The first two
-groups are purely quantitative in nature; the study begins with
-conceptual tasks to understand whether the DAG is capable of providing
-users with an understanding of how a commit is merged into the master
-branch, and other commits that it is merged with. We continue with
-summarization tasks, to quantitatively understand if the \mt model makes
-a difference in summarization tasks.
-
-The third task set is purely qualitative, \dmg{and are their goal is ... don't imply content in your sentences... be
-  precise; this applies to the paragraph before} to gain better understanding
-of whether participants preferred working with the \mt model in \tool,
-or working with the DAG in Gitk, and what aspects they preferred from
-each tool. The final group of questions is simply to gain a better
-understanding of the demographic of our participants.
-
-\textbf{Conceptual Tasks}
-
-\dmg{given the choice, not allowed.  I don't understand this section.
-  You need to be more top-down in your description. you need to provide
-  context on what the users are expected to do, and why. So what are the
-  tasks? You never explain.  what is the goal of the tasks? how many?
-  what were the tasks? what is expected from each? you need to answer
-  these questions, not necessarily in the same order.  also, remember
-  this is a comparative study, so I presume the users were asked to do
-  both with linvis and without it (choice of gitk/git command line) }
-
-The participants were allowed to use Gitk or the git command line tools
-to perform these tasks. These tasks are asked in the specified order,
-allowing tabularr to first, create a conceptual understanding of the
-commits and merges that will be involved with the rest of the tasks that
-follow.tabularmay then use their diagrams to answer T2 and T3. We do not
-collect metrics from \tool for the conceptual tasks, opting instead to
-determine if the DAG is sufficient for understanding how a commit is
-integrated. \tool uses the \mt model show a tree visualization of how
-commits are merged, until the commit reaches the master branch.
-Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.
-
-\dmg{I don't like the present tense, but that is your option}
-
-We provide the participant with 10 minutes per commit to complete the
-first task. The answers to the two other questions in this section are
-drawn from the answer in the first task.
-
-\textbf{Summarization Tasks}
-
-\dmg{again, you are going into details first. First the tasks --see
-  above-- including their grouping.  then at the end, describe the
-  randomization and why it is done (which you did) }
-
-The summarization tasks (Table~\ref{tab:summarization_tasks}) are
-grouped into \dmg{how many?} set of tasks according to their theme. The
-tasks within a given task set are shuffled, and the task sets are
-shuffled, to ensure that similar questions are together, but not in the
-same order, and to ensure that the themes are in randomized order
-between participants. The order of the commits is shuffled, as well. The
-order of the tool is randomized, choosing to start with either \tool or
-Gitk, to ensure that we do not place a bias on one tool through the
-experiment.
-
-As understanding where the merge begins and ends is vital to answering
-the other question, the merge task set will always be the first task set
-presented; however, the questions will still be re-ordered within the
-task set.
-
-\textbf{User Opinion and Profile}
-
-Finally, we ask the user for their opinions on the tools and for some
-information about their level of experience with git.  \dmg{be more
-  precise: understanding or preference is too broad. What you want to
-  know answers to specific questions regarding how the tools compare and
-  which ones they would prefer} The first three questions are intended
-to provide us with a better understanding of preference, and insight on
-why, the participant had that preference.  The last three questions in
-this section are aimed at getting a better understanding of the
-demographic of our participants.
-
-\begin{enumerate}
-  \item Given these tasks again, which tool would you prefer to use?
-  \item Which aspects of each tool did you like and why?
-  \item How long have you used git?
-  \item If you have used git, for what kind of projects? (personal,
-    school courses, professional?)
-  \item If you have used git, how many commits, files, and contributors
-    were involved with the largest repository you have worked with?
-\end{enumerate}
-
-\textbf{Procedure}
-
-\dmg{tell me first why you need to create a script, so I know what you
-  do something: we did random assignment of tasks (explain what you did)
-  and randomized the order in which the tasks were performed.  Then
-  explain why you need a script}
-
-\dmg{it does not randomize the order the tools. It randomizes which tool
-  is going to be used on each tasks.  be precise with your language}
-
-\dmg{what you explained how commits relate to tasks and how they get
-  assigned? think as if you are explaining this to somebody who has to
-  do this: you are the designer, explaining the why and the how }
-
-For each participant, we run a simple python script to generate the
-script for the study, eliminating the possibility of bias from us. The
-script randomizes the order of the tools, the order of the commits, the
-order of the task sets, and the order of the tasks within each task set,
-where appropriate.
-
-\subsection{Commit Selection}
-\label{sub:commit_selection}
-
-\dmg{see my comments in previous section. it is ok to explain them here,
-  but you need to explain what they are briefly in the previous section}
-
-We chose two commits for use in the study. The order that the commits
-are presented is randomized between participants, but the order is kept
-consistent through the tasks described in the next subsection.
-
-\dmg{this place does not require a description of the database. This
-  requires just to say the commits/merge set (dates and numbers are ok)
-  only} We use the \tool database, containing commits and merge-tree
-information from April 16th 2005 to October 14th, 2014, which
-corresponds to being between Linux release 2.6.12-rc3 and Linux
-3.17-rc1. \evan{Cite the database even more?} \dmg{again, i have
-  problems with your writing: ``we wanted to use in the study commits of
-  different size, to determine if the size had an impact in the
-  completion of the tasks. We chose commits of three different sizes...}
-We chose the commits based on tree sizes. We \dmg{we didn't find: this
-  is a fact, not a discovery; what is a majority? 50\%+1? be precise}
-\dmg{drop the word inner, and do it in order: 25\% of trees have only
-  one commit, 50\%.., 75\%...} found that a majority of the trees
-contain at most seven inner commits, while more than 25\% of the trees
-contain a single commit. 75\% of the tree contained up to 51 inner
-commits and merges, and finally the largest tree contained 7217 nodes.
-We chose to work with the trees in \dmg{if you are going to use the
-  words quartile, just use them before in your description eg: the
-  distribution of the sizes of the trees is x, y merges, whose size
-  corresponded to the 25 ,50 quartiles, simple and precisee} the first
-and second quartile, as merge trees of sizes between one and seven, not
-including the merge into the master branch, make up the majority of the
-trees in the database.
-
-\dmg{remove from here, what is here? this is not a road trip :)} From
-here, we selected one tree from the trees of a single commit at random.
-Selecting a commit from that tree is trivial, as there is only one to
-choose.\dmg{remove the obvious, instead concentrate on why the others
-  are not trivial}
-
-We selected one tree of size seven at random from all trees of size
-seven\dmg{how many where there?}. We placed a restriction the trees, the
-selected tree must include at least one internal merge to increase the
-complexity of the trees tested. For each randomly selecting a tree, we
-chose one of its internal commits at random (using the
-\verb|random.choice| function in python 3.6.1).
-
-\dmg{question: so you chose only one commit of each size, and the
-  randomization was which tool to use for each question? or did they
-  have to do the same task for two commits?}
-
-Using this technique, we selected commit
-\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
-a single node (Figure~\ref{fig:commit_1}), which we will refer to as
-\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
-the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
-will refer to as \comB.
-
-\begin{figure}[bpt]
-  \centering
-\begin{tabular}{ m{1.5cm} m{3cm} }
-  \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
-  \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
-\end{tabular}
-  \caption{The merge trees used in the user study, 
-    containing one and seven commits, respectively.}
-  \label{fig:commit_1}
-\end{figure}
-
-\subsection{Results}
-\label{sec:results}
-
-\dmg{this is still methodology.}
-
-We recorded the audio and screen capture video for each participant in
-the study for further analysis.
-\dmg{you don't need this sentence, it is redundant}
-After capturing the information, we
-extracted various metrics from the videos.
-\dmg{ you don't record, you measure}
-From the conceptual tasks, we
-recorded the time in seconds, and the answer for the question. We
-extracted three metrics for the summarization tasks, the time in
-seconds, the correctness, and the accuracy.
-\dmg{and then somethings you are over precise :): correctness is simply whether their answer to the task was correct or not}
-Correctness is a binary
-value, it is true if the response was correct and false if the response
-was incorrect.
-\dmg{if you are going to use this metric, you need to define precisely. remember, you are giving this to somebody else
-  to implement}
-The accuracy is measures in distance from the correct
-value. An accuracy of zero indicates a correct answer. All tasks in the
-summarization tasks include an accuracy metric except T7, which will
-only be right or wrong, and therefore cannot meaningfully be measured by
-a distance.
-
-\subsection{Conceptual Tasks}
-\label{sub:conceptual_tasks}
-
-\dmg{i don't like the expression: we are able to see... too long.. be to the point: Table X shows/depicts/ etc ... they
-  are not overviews, they are a summary}
-In Table~\ref{tab:conceptual_results}, we are able to see an overview of
-the results from the conceptual questions.
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Results from the conceptual questions\dmg{what are the columns? they need captions}}
-  \label{tab:conceptual_results}
-  \begin{tabular}{ll|r|lrr|rrr}
-    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
-    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
-    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
-    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
-    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
-  \end{tabular}
-\end{table*}
-
-Users were able to more closely estimate the number of commits and
-merges in the larger tree, but generally took longer than the smaller
-tree. The tree with a single node resulted in more variability in the
-estimate of number of commits.
-
-It should be noted that these questions were answered after spending
-roughly ten minutes attesting to draw a picture that held the answers to
-these questions.
-
-\subsection{Summarization Tasks}
-\label{sub:summarization_tasks}
-
-In Figure~\ref{fig:summarization_results}, we see all of the
-correctness, accuracy, and timing results. The first row of plots are
-the correctness results. We see in all cases except for T10, that the
-number of instances of a correct result with \tool is roughly the same
-as the number of incorrect instances with Gitk. We test each task with
-the McNemar Chi-square test~\cite{McNemar1947}.
-
-\begin{figure*}[htpb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
-  \caption{Results from the summarization tasks. The first row plots the
-    correctness metric for each task. The second row plots the
-    accuracy metric for each task. The third row plots the time metric
-    for each task.}
-  \label{fig:summarization_results}
-\end{figure*}
-
-With the McNemar's tests, we analyze the participants that change from
-being correct with one tool to being incorrect with the other. This does
-not take into account the number of people who were able to provide the
-correct answer for both tools, nor the people who were incorrect for both
-tools. The results show whether providing the participants with \tool
-over the current standard, gitk, improves the correctness of the
-summarization tasks among the participants. We show the results for the
-number of people who were, with both tools, correct, incorrect, and
-changed in Table~\ref{tab:correctness_table}.
-
-In all cases, except for task T10, we saw statistically significant
-results suggesting that we reject the null hypothesis. We tested each
-correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
-degree of freedom is $7.879$.
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Correctness tables showing how many people stayed correct,
-    incorrect, and how many people changed between tool}
-  \label{tab:correctness_table}
-  \begin{tabular}{c|c|c}
-
-    T4 & T5 & T6\\
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 15                           & 2\\
-  \end{tabular} &
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
-                           & Incorrect & 14                           & 4\\
-  \end{tabular}
-
-  &
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
-                           & Incorrect & 16                           & 2\\
-  \end{tabular}
-
-  \\\hline
-  T7 & T8 & T9\\
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 8                            & 0\\
-                           & Incorrect & 13                           & 2\\
-  \end{tabular}  &
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 15                           & 2\\
-  \end{tabular} &
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 14                           & 3\\
-  \end{tabular}\\\hline
-
-  & T10 & \\
-  & \begin{tabular}{cc|rr}
-                            &           & \multicolumn{2}{c}{Linvis}\\
-                            &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
-                            & Incorrect & 8                            & 0\\
-  \end{tabular} & \\
-
-  \end{tabular}
-\end{table*}
-
-The correctness test statistics and conclusion for each summarization
-task;
-
-The correctness test statistics, conclusions, and a description of the
-accuracy and timing results for each summarization task follow:
-
-\begin{itemize}
-
-  \item
-
-    T4; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the series of merges that a commit
-    takes, with $99.5\%$ confidence.
-
-    There is almost no variation in accuracy between participants using
-    \tool, while there is more variance in Gitk. There were two outliers
-    among the instances using \tool, which lie within the second and
-    third quartiles for the accuracy measure of Gitk. The furthest
-    outlier for \tool lies at a distance that is slightly greater than
-    the median for Gitk.
-
-    We see similar results in the timing metric; all significant data
-    occurs within less time than the second, third, and fourth quartile
-    of the Gitk results. Nearly all outlier instances in \tool are
-    within the third quartile group of the instances of Gitk with the
-    exception of the most extreme, which lies in the fourth quartile
-    group.
-
-    The results indicate that \tool is able to assist users with
-    determining how a commit is merged more accurately, and more
-    quickly.
-
-  \item
-
-    T5; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the other related commits, with
-    $99.5\%$ confidence.
-
-    There is very little variance in the accuracy of results for \tool
-    compared to the accuracy of the results in Gitk. The median distance
-    is smaller, indicating that the participants performed more
-    accurately using \tool.
-
-    There is less variance in the time taken to come to an answer with
-    \tool than with Gitk. Furthermore, we see that a majority of the
-    participants were able to complete the task in \tool in much less
-    time than 50\% of the participants using Gitk.
-
-    The results indicate that \tool is able to assist users in
-    understanding which commits are related to a given commit.
-
-  \item
-
-    T6; the resulting $\chi^2$ value is $16$. $16 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the number of authors involved with
-    the merge tree, with $99.5\%$ confidence.
-
-    We see the same results from this task as with the previous tasks,
-    suggesting that \tool assists users determine the number of authors
-    more quickly and more accurately.
-
-  \item
-
-    T7; the resulting $\chi^2$ value is $13$. $13 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining who contributed the most changes,
-    with $99.5\%$ confidence.
-
-    This question does not include an accuracy metric, as there are only
-    two possible outcomes. We see the same result in the timing metric
-    as we did from the previous questions.
-
-    The results suggest that \tool is able to help users correctly
-    identify the person who contributed the most lines of code to a
-    merge tree in less time.
-
-  \item
-
-    T8; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining how many files were contributed,
-    with $99.5\%$ confidence.
-
-    The results of this task are the same as in the previous tasks.
-    This suggest that \tool is able to assist users determine how many
-    files were modified, more quickly and more accurately.
-
-  \item
-
-    T9; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining which file had the most changes,
-    with $99.5\%$ confidence.
-
-    While this question appears to have a similar answer to the correct
-    answer for T7, the answer \comA had two files with the same number
-    of midifications, thus the participant must correctly identify both
-    of these files to be correct. For this reason, we include an
-    accuracy metric.
-
-    The results of this task are the same as the previous tasks.
-    This suggests that \tool is able to assist users to determine which
-    file had the most changes.
-
-  \item
-
-    T10; the resulting $\chi^2$ value is $5.44$. $5.44 < 7.879$,
-    therefore we do not reject the null hypothesis, suggesting that
-    \tool does not make a difference in determining which modules are
-    involved with a merge tree.
-
-    The accuracy and timing results for this task are the same as the
-    others, showing that there is less variability in both accuracy and
-    the time taken to have an answer between participants using \tool
-    than using Gitk. Looking Table~\ref{tab:correctness_table}, we can
-    see that, unlike the other questions, a majority of the participants
-    were correct using both tools.
-
-\end{itemize}
-
-\subsection{User Opinions}
-\label{sub:user_opinions}
-
-Among the 12 pariticipants, there was nearly unanimous agreement that
-for conceptual understanding and summarization tasks, \tool was easier
-to work with than Gitk. The participants cited the ability to abstract
-information about the merge from the clean summarization tables and
-simple visualizations. Three participants suggested that someone with a
-professional understanding of Gitk and the git commandline may be to
-extract the a good conceptual understanding from the DAG, and perform
-the summarization tasks from the commandline. One of these three
-participants said that they would prefer to have both tools available,
-as they are able to complement eachother.
-
-\section{Discussion}
-\label{sec:discussion}
-
-% TODO: Go back to the conceptual task results and make them stronger
-Overall, the results show fairly conclusive results; the DAG is unable
-to adequately provide users with an explanation of the changes being
-made in the repository, the DAG is not easily understood, and the DAG is
-unable to provide summarizations of the changes being made at a merge.
-Meanwhile, the \mt model is able to provide a better explanation of how
-a commit is integrated, along with better summarization capabilities. In
-this section, we discuss some observations made during the study,
-threats to the validity of the study, comments made by the participants
-during the study, modifications to the original algorithm to help
-generalize it for more repositories, and future work.
-
-\subsection{Observations}
-\label{sub:observations}
-
-% TODO: include info on merge tree sizes
-
-During the course of the study, we noticed a consistent theme among the
-participants. The most difficult challenge was identifying the master
-branch of the repository. Given that single-commit merge trees are the
-most common tree, and that they are relatively simple, we were
-interested in seeing if the participants would be able to identify the
-tree easily, but the results suggest that this was not the case. The
-issue primarily stemmed from not being able to identify the master
-branch. Most of the participants realized that they should not be going
-backward, toward the initial commit, very early on. Furthermore, many
-would identify the correct commit; however, they would start listing
-every merge along the master branch as being a merge in the tree, and
-including the commits from that merge. In the case of the single commit,
-simply identifying the master branch would most likely be sufficient in
-providing a clearer understanding of how a commit or flat tree is
-integrated into the master branch of the repository.
-
-While, in general, the participants of the study preferred using \tool
-to Gitk, there were some noticeable usability issues. The most prominent
-issue being the navigation within the tree. The participants expected
-that the tabs would update when they clicked on a node in the tree,
-instead of having to click the link to the commit after clicking the
-node. This was true in both the Reingold-Tilford tree and the Pack tree.
-The second major usability issue was the inconsistency between the list
-tree and the other tree visualizations. The list tree is rooted at the
-repository event that the user is current looking at, while the other
-tree visualizations show the entire tree, and use bright orange to
-highlight the current node.
-
-\subsection{Comments From a Release Manager}
-\label{sub:comments_from_a_release_manager}
-
-One of the participants of the study had worked as a release manager for
-more than three years, working with both SVN and CVS repositories, but
-had little experience with git.
-
-Contributors making merges need to understand more than just what merges
-a commit was collected into before reaching the repository of the
-contributor. It is also important to understand order that the related
-commits were made, as they tell the story of what the developer was
-thinking as they were writing the changes. The current merge-tree model
-does not order the commits, simply placing them randomly. This is the
-primary reason behind this participant's request to use both
-simultaneously. \tool is able to help with the aggregation of the
-information, and provide a better understanding of the next merge
-involved in integrating this commit, but the DAG in gitk provides the
-full story of the commit instead of hiding it behind a layer of
-abstraction.
-
-The comments from this participant were very insightful, and will assist
-in improving the model in the future.
-
-\subsection{Threats to Validity}
-\label{sub:threats}
-
-While git includes gitk and the git log provides a view of the DAG with
-\verb|git log --graph|, many of the participants in our study were
-unfamiliar with the DAG visualizations shown by these tools. A possible
-reason for this may be that unless a repository reaches a sufficient
-complexity, both in branching and number of commits, these tools are
-necessary for the normal operation of git. Furthermore, fear of the
-complexity caused by branching may cause personal and academic
-repositories to have relatively simple structures. While they were
-unfamiliar with the DAG structure, these users were also unfamiliar with
-the \mt model, and were still able to better understand and summarize
-the repository using \tool.
-
-Each question was evaluated independently of the results from the
-previous question in the summarization tasks. Had we modified the
-correct answer to fit the conceptual understanding of which commits were
-included in the tree, we may have witnessed different results. A
-modification to the study may include providing the participants with
-the correct set of commits prior to the summarization tasks, after the
-merge task set.
-
-\subsection{Generalization}
-\label{sub:generalization}
-
-Seeing that the \mt model is able to assist users with understanding how
-commits are integrated, we continue by looking at ways of improving the
-algorithm for faster operations, and making the conversion from DAG to
-\mt feasible over the internet for smaller repositories.
-
-The problem of finding the master branch is the most difficult, and is
-not possible without heuristic approaches. The master branch can be
-confounded by \textit{foxtrot} merges, which change the order of the
-parents in the parent list of a merge. Other than the ordering of the
-parent list, git has no other way to determine which commits are on the
-master branch of the repository. For the purpose of this discussion, we
-will assume that we can rely on the order of the parent list to be
-correct, and that foxtrots do not occur.
-
-\begin{algorithm}
-  \caption{Computing the generalized Merge Tree}
-  \label{alg:generalized}
-  \begin{algorithmic}[1]
-    \Function{Phase 1}{initial commit $root$} : updated tree
-    \State $depth \gets 0$
-    \State $Q \gets root$
-    \Do
-    \State $cur \gets Q.dequeue$
-    \State $parents \gets cur.parents$
-    \State $depth \gets depth + parents.length - 1$
-    \For{$index, parent \in parents$}
-    \If{$parent.depth$ is $undefined$}
-    \State $parent.depth \gets cur.depth + index$
-    \ElsIf {$cur.depth > parent.depth$}
-    \State $depth \gets depth - 1$
-    \ElsIf {$cur.depth < parent.depth$}
-    \State $depth \gets depth - 1$
-    \State $parent.depth \gets cur.depth$
-    \EndIf
-    \State $parent.children \gets cur$
-    \If{$parent$ has not been visited}
-    \State $Q \gets parent$
-    \State $parent.visited \gets True$
-    \EndIf
-    \EndFor
-    \doWhile{$Q$ not empty and $depth \ne 0$ }
-    \EndFunction
-
-    \Function{Phase 2}{initial commit $root$} : Merge tree
-    \State $tree.root \gets root$
-    \State $Q$ // Empty Queue
-
-    \State $parents \gets root.parents$
-    \State $parents \gets tail\ parents$
-    \For {$parent \in parents$}
-    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
-    \If {$realParent$ is $root$}
-    \State $root.children \gets parent$
-    \State $parent.parent \gets root$
-    \State $Q \gets parent$
-    \EndIf
-    \EndFor
-
-    \While{$Q$ not empty}
-    \State $cur \gets Q.dequeue$
-    \State $parents \gets cur.parents$
-    \For{$parent \in parents$}
-    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
-    \If {$realParent$ is $cur$}
-    \State $cur.children \gets parent$
-    \State $parent.parent \gets cur$
-    \State $Q \gets parent$
-    \EndIf
-
-    \EndFor
-    \EndWhile
-
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-The primary difference between the original algorithm and the
-generalization is the main data structure. Algorithm~\ref{alg:original}
-is centered around the use of the stack, making it akin to a depth-first
-traversal of the DAG.\@ The generalization in
-Algorithm~\ref{alg:generalized} is based on the queue, making it akin to
-the breadth-first traversal of the DAG.\@ While the final results of the
-two traversals should not be different, the runtime properties are
-different.
-
-We will begin with a simplified description of the original algorithm.
-We start at the master branch pointer. We will use the term master
-branch loosely here, meaning any branch that has been labeled in some
-way as being the main branch, this branch may have a different name, but
-standard naming conventions call it the master branch. Starting from the
-master branch label, we traverse the entire master branch to the initial
-commit, recording the commit hash of every repository event encountered.
-
-The second phase iterates over every commit, computing the shortest
-distance to the master branch. For this reason, the entire repository
-must be available to the algorithm. This is a dynamic programming
-solution, centered around the call stack. The next node on the shortest
-path to the master branch will become the parent of this node in the
-tree.
-
-If this algorithm is to be generalized into a dynamic, web-based, tool,
-it may not be feasible to download every commit in the repository in a
-reasonable period of time. We instead look toward a breadth-first
-approach at limiting the number of repository events that must be
-downloaded.
-
-Focusing on the generalized algorithm~\ref{alg:generalized}, the
-algorithm is broken into two phases. The first phase is where the
-improvements take place. The first phase has three main goals, the first
-is to track the master branch. The second goal labels the depth of each
-repository event that is traverses. The final goal is to record the
-children of each commit, reversing the DAG.\@ The first phase also
-begins the process of reversing the DAG, by providing the candidate
-commits that could possibly be the real parent to this node. The first
-phase can terminate early, if it reaches a point where all branches that
-it is following terminate. The depth metric on line 7 indicates how many
-branches are currently being traversed. In line 11 through 15, we
-decrement the depth, as this is the point where a branch occurs. The
-changing of the depth occurs if there are more events on the same branch
-than between the merge and branch point of a separate branch.
-
-\begin{table}[htpb]
-  \centering
-  \caption{Example traversal of the example DAG in
-    Figure~\ref{fig:repoDAG} starting at merge 11.}
-  \label{tab:ex_traversal}
-  \begin{tabular}{cccc|cc}
-    \toprule
-    Current & Parents & Depth & Q       & Visited & Children\\\midrule
-    -       & -       & 0     & 11      & 11:0    & \\
-    11      & 4,9     & 1     & 4,9     & 4:0     & 11, 5\\
-    4       & 1       & 1     & 9,1     & 9:1     & 11\\
-    9       & 7,6,8   & 3     & 1,7,6,8 & 7:1     & 9, 8\\
-    1       &         & 3     & 7,6,8   & 6:2     & 9\\
-    7       & 5       & 3     & 6,8,5   & 8:3     & 9\\
-    6       & 5       & 2     & 8,5     & 5:1     & 7, 6\\
-    8       & 7       & 1     & 5       & 2:1     & 5, 3\\
-    5       & 4,2,3   & 2     & 2,3     & 3:2     & 5\\
-    2       & 1       & 1     & 3       &         & \\
-    3       & 2       & 0     &         &         & \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-In Table~\ref{tab:ex_traversal}, we follow the algorithm starting at
-merge commit 11 from the example in Figure~\ref{fig:repoDAG}, recording
-the current node, the parents of the current node, the depth after
-updating, and the queue structure after updating. The visited column is
-not dependant on the current step, but keeps track of the nodes that
-have been visited, and the depth at that node. When we reverse the DAG,
-the children in this table will become the candidates for being the
-parent of the corresponding node. For example, the parent of node 7 may
-be resolved to either node 9 or node 8. In this case, the shortest path
-to the master branch is through node 9.
-
-The terminology surrounding the parent/child relationship gets very
-confusing at the change between phase one and two because this is where
-the roles switch. In phase 1, we are working with the DAG; in phase 2,
-we are working with a structure that is neither a tree nor a DAG, but
-can only be described as a directed graph.
-
-The goal of the second phase is to resolve the tree from the directed
-graph. We will refer to parents in the same sense as we would refer to
-them in the tree form of the repository for the remainder of this
-section. We have gathered the depths and possible parents (line 17), but
-we need to remove the extra parents. Extra parents are culled in line
-42 of the algorithm. Of the possible parents, we take the parent that is
-closest to the root, but not beyond. This forces the real parent to
-either be deeper in the tree, or on the same branch. Line 30 to 37 are
-for bootstrapping the queue, ensuring that we are not starting on an
-empty queue.
-
-The construction of the tree becomes trivial now that we are able to
-determine the parents and children for each node. Lines 44 and 45
-perform this construction, adding the DAG parent to the tree children of the
-current node, and the DAG parent's parent becomes the current node,
-which was the DAG parent's child. We continue following the path until
-we hit the root node, merging the commit into the master branch.
-
-This algorithm relies heavily on the order of parents in the parent
-list, but is able to trim the number of nodes that must be downloaded.
-Furthermore, the resulting tree is able to show the order of commits,
-telling the story of how a developer was working, before the set of
-commits was merged. We developed a small bitbucket plugin to test the
-design and determine whether it was capable of producing merge trees,
-shown in Figure~\ref{fig:b_plugin}.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=\linewidth]{figures/plugin.png}
-  \caption{Screen shot of the Bitbucket plugin showing the online tree
-    computation of the repository shown in~\ref{fig:repoDAG}.}
-  \label{fig:b_plugin}
-\end{figure}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this paper, we develop a tree-based model designed to provide git
-users with a better explanation of the events occurring in a git
-repository. The \mt model is rooted at the merge into the master branch
-of a repository, where the leaves are the individual commits.
-
-We implement a tool, \tool, which allows a user to filter the commits
-with a simple search interface. Once the user finds the commit that they
-are interested in, they are provided simple aggregated tables showing
-the files that were modified, the authors that made the modifications,
-among other information. They are also provided with three
-visualizations, a list tree, a pack tree, and a Reingold-Tilford tree
-view. The list tree enables easy searching for textual features, while
-the pack tree is good for visualizing the structure of large merge
-trees, and the Reingold-Tilford tree is better at visualizing the
-structure of small merge trees.
-
-To ensure that our model is necessary, we evaluate the ability of users
-to perform simple conceptual tasks, finding how a commit was merged into
-the master branch of the Linux repository, and determining other commits
-that were merged with the given merge. Finding that the participants in
-our study were unable to perform this task, we compared the ability of
-the participants to summarize key pieces of information aggregated in
-these groups of commits. We asked for information about what files were
-modified, the authors making the changes, among other tasks. Finally, we
-asked the participants about their preference in tools, and which
-aspects of each they preferred. We found that the participants in our
-study were able to correctly identify the correct answer more quickly
-using the \mt model than they were able to with the DAG.
-
-Given that the model provides users with a better understanding of
-events in the repository, we describe some issues with identifying the
-master branch of the repository and a more generalized approach to
-constructing merge trees. We implement the generalized approach as a
-bitbucket plugin to verify that it is able to quickly produce results
-online, unlike in the implementation of \tool.
-
-% TODO: You can tell that I am tired. This conclusion will need to be re-worked.
-
-\balance
-
-% \nocite{*}
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2017-07-07 paper/discussion.tex
small comment in discussion, not a major change
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

@@
 % Jun 24, 2017
+% TODO: Go back to the conceptual task results and make them stronger
+Overall, the results show fairly conclusive results; the DAG is unable
+to adequately provide users with an explanation of the changes being
+made in the repository, the DAG is not easily understood, and the DAG is
+unable to provide summarizations of the changes being made at a merge.
+Meanwhile, the \mt model is able to provide a better explanation of how
+a commit is integrated, along with better summarization capabilities. In
+this section, we discuss some observations made during the study,
+threats to the validity of the study, comments made by the participants
+during the study, modifications to the original algorithm to help
+generalize it for more repositories, and future work.
+
+\subsection{Observations}
+\label{sub:observations}
+
+% TODO: include info on merge tree sizes
+
+During the course of the study, we noticed a consistent theme among the
+participants. The most difficult challenge was identifying the master
+branch of the repository. Given that single-commit merge trees are the
+most common tree, and that they are relatively simple, we were
+interested in seeing if the participants would be able to identify the
+tree easily, but the results suggest that this was not the case. The
+issue primarily stemmed from not being able to identify the master
+branch. Most of the participants realized that they should not be going
+backward, toward the initial commit, very early on. Furthermore, many
+would identify the correct commit; however, they would start listing
+every merge along the master branch as being a merge in the tree, and
+including the commits from that merge. In the case of the single commit,
+simply identifying the master branch would most likely be sufficient in
+providing a clearer understanding of how a commit or flat tree is
+integrated into the master branch of the repository.
+
+While, in general, the participants of the study preferred using \tool
+to Gitk, there were some noticeable usability issues. The most prominent
+issue being the navigation within the tree. The participants expected
+that the tabs would update when they clicked on a node in the tree,
+instead of having to click the link to the commit after clicking the
+node. This was true in both the Reingold-Tilford tree and the Pack tree.
+The second major usability issue was the inconsistency between the list
+tree and the other tree visualizations. The list tree is rooted at the
+repository event that the user is current looking at, while the other
+tree visualizations show the entire tree, and use bright orange to
+highlight the current node.
+
+\subsection{Comments From a Release Manager}
+\label{sub:comments_from_a_release_manager}
+
+One of the participants of the study had worked as a release manager for
+more than three years, working with both SVN and CVS repositories, but
+had little experience with git.
+
+Contributors making merges need to understand more than just what merges
+a commit was collected into before reaching the repository of the
+contributor. It is also important to understand order that the related
+commits were made, as they tell the story of what the developer was
+thinking as they were writing the changes. The current merge-tree model
+does not order the commits, simply placing them randomly. This is the
+primary reason behind this participant's request to use both
+simultaneously. \tool is able to help with the aggregation of the
+information, and provide a better understanding of the next merge
+involved in integrating this commit, but the DAG in gitk provides the
+full story of the commit instead of hiding it behind a layer of
+abstraction.
+
+The comments from this participant were very insightful, and will assist
+in improving the model in the future.
+
 \subsection{Threats to Validity}
-\label{sub:threats_to_validity}
+\label{sub:threats}
 
-While git comes bundled with gitk and the \verb|git log --graph|
-command, many of the users in our study were unfamiliar with the DAG
-shown by these tools. A potential reason for this may be that unless a
-repository reaches sufficient complexity, both in branching and number
-of commits, these tools are not necessary for the normal functionality
-of git. However, while they were unfamiliar with either merge trees or
-the DAG, these users were still able to better understand and summarize
-the structure of the repository using merge trees.
+While git includes gitk and the git log provides a view of the DAG with
+\verb|git log --graph|, many of the participants in our study were
+unfamiliar with the DAG visualizations shown by these tools. A possible
+reason for this may be that unless a repository reaches a sufficient
+complexity, both in branching and number of commits, these tools are
+necessary for the normal operation of git. Furthermore, fear of the
+complexity caused by branching may cause personal and academic
+repositories to have relatively simple structures. While they were
+unfamiliar with the DAG structure, these users were also unfamiliar with
+the \mt model, and were still able to better understand and summarize
+the repository using \tool.
 
 Each question was evaluated independently of the results from the
-previous question in the summarization tasks. Had the correct answers
-been adjusted to match the participants' understanding of which commits
-were involved in a merge tree, the results of the summarization would
-have likely been very different. Another approach would to have provided
-the participants with the correct tree once they had completed the
-conceptual tasks, and the merge task set in the summarization task set.
+previous question in the summarization tasks. Had we modified the
+correct answer to fit the conceptual understanding of which commits were
+included in the tree, we may have witnessed different results. A
+modification to the study may include providing the participants with
+the correct set of commits prior to the summarization tasks, after the
+merge task set.
 
-\subsection{Future Work}
-\label{sub:future_work}
+\subsection{Generalization}
+\label{sub:generalization}
 
-% Inter Navigation
-% TODO: verify this with the study
+Seeing that the \mt model is able to assist users with understanding how
+commits are integrated, we continue by looking at ways of improving the
+algorithm for faster operations, and making the conversion from DAG to
+\mt feasible over the internet for smaller repositories.
 
-While the merge-tree model provides users with a better conceptual image
-of what is occuring within a single merge tree, we have not exprimented
-or tested the potential gains of navigation between trees. At this time,
-the Linvis search provides results grouped by merge-tree root instead of
-as a single block of results. The results are generally similar, with
-each merge tree being merged at a different point in time, or to a
-different version of the kernel. A future extension of this may
-incorporate some information about how a given cluster changes over
-time. To do this, a technique would need to be developed to cluster
-merge-trees into modules, and then ordering the merge-trees based on
-time or release.
+The problem of finding the master branch is the most difficult, and is
+not possible without heuristic approaches. The master branch can be
+confounded by \textit{foxtrot} merges, which change the order of the
+parents in the parent list of a merge. Other than the ordering of the
+parent list, git has no other way to determine which commits are on the
+master branch of the repository. For the purpose of this discussion, we
+will assume that we can rely on the order of the parent list to be
+correct, and that foxtrots do not occur.
 
-% TODO: ensure that linvis has been referenced/defined
+\begin{algorithm}
+  \caption{Computing the generalized Merge Tree}
+  \label{alg:generalized}
+  \begin{algorithmic}[1]
+    \Function{Phase 1}{initial commit $root$} : updated tree
+    \State $depth \gets 0$
+    \State $Q \gets root$
+    \Do
+    \State $cur \gets Q.dequeue$
+    \State $parents \gets cur.parents$
+    \State $depth \gets depth + parents.length - 1$
+    \For{$index, parent \in parents$}
+    \If{$parent.depth$ is $undefined$}
+    \State $parent.depth \gets cur.depth + index$
+    \ElsIf {$cur.depth > parent.depth$}
+    \State $depth \gets depth - 1$
+    \ElsIf {$cur.depth < parent.depth$}
+    \State $depth \gets depth - 1$
+    \State $parent.depth \gets cur.depth$
+    \EndIf
+    \State $parent.children \gets cur$
+    \If{$parent$ has not been visited}
+    \State $Q \gets parent$
+    \State $parent.visited \gets True$
+    \EndIf
+    \EndFor
+    \doWhile{$Q$ not empty and $depth \ne 0$ }
+    \EndFunction
 
-Perform a user study over multiple repositories using different merge
-structures to see how quickly new members of a team can understand the
-merge procedure. % Yeah, this is a big one, actually.
+    \Function{Phase 2}{initial commit $root$} : Merge tree
+    \State $tree.root \gets root$
+    \State $Q$ // Empty Queue
+
+    \State $parents \gets root.parents$
+    \State $parents \gets tail\ parents$
+    \For {$parent \in parents$}
+    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
+    \If {$realParent$ is $root$}
+    \State $root.children \gets parent$
+    \State $parent.parent \gets root$
+    \State $Q \gets parent$
+    \EndIf
+    \EndFor
+
+    \While{$Q$ not empty}
+    \State $cur \gets Q.dequeue$
+    \State $parents \gets cur.parents$
+    \For{$parent \in parents$}
+    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
+    \If {$realParent$ is $cur$}
+    \State $cur.children \gets parent$
+    \State $parent.parent \gets cur$
+    \State $Q \gets parent$
+    \EndIf
+
+    \EndFor
+    \EndWhile
+
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+The primary difference between the original algorithm and the
+generalization is the main data structure. Algorithm~\ref{alg:original}
+is centered around the use of the stack, making it akin to a depth-first
+traversal of the DAG.\@ The generalization in
+Algorithm~\ref{alg:generalized} is based on the queue, making it akin to
+the breadth-first traversal of the DAG.\@ While the final results of the
+two traversals should not be different, the runtime properties are
+different.
+
+We will begin with a simplified description of the original algorithm.
+We start at the master branch pointer. We will use the term master
+branch loosely here, meaning any branch that has been labeled in some
+way as being the main branch, this branch may have a different name, but
+standard naming conventions call it the master branch. Starting from the
+master branch label, we traverse the entire master branch to the initial
+commit, recording the commit hash of every repository event encountered.
+
+The second phase iterates over every commit, computing the shortest
+distance to the master branch. For this reason, the entire repository
+must be available to the algorithm. This is a dynamic programming
+solution, centered around the call stack. The next node on the shortest
+path to the master branch will become the parent of this node in the
+tree.
+
+If this algorithm is to be generalized into a dynamic, web-based, tool,
+it may not be feasible to download every commit in the repository in a
+reasonable period of time. We instead look toward a breadth-first
+approach at limiting the number of repository events that must be
+downloaded.
+
+Focusing on the generalized algorithm~\ref{alg:generalized}, the
+algorithm is broken into two phases. The first phase is where the
+improvements take place. The first phase has three main goals, the first
+is to track the master branch. The second goal labels the depth of each
+repository event that is traverses. The final goal is to record the
+children of each commit, reversing the DAG.\@ The first phase also
+begins the process of reversing the DAG, by providing the candidate
+commits that could possibly be the real parent to this node. The first
+phase can terminate early, if it reaches a point where all branches that
+it is following terminate. The depth metric on line 7 indicates how many
+branches are currently being traversed. In line 11 through 15, we
+decrement the depth, as this is the point where a branch occurs. The
+changing of the depth occurs if there are more events on the same branch
+than between the merge and branch point of a separate branch.
+
+\begin{table}[htpb]
+  \centering
+  \caption{Example traversal of the example DAG in
+    Figure~\ref{fig:repoDAG} starting at merge 11.}
+  \label{tab:ex_traversal}
+  \begin{tabular}{cccc|cc}
+    \toprule
+    Current & Parents & Depth & Q       & Visited & Children\\\midrule
+    -       & -       & 0     & 11      & 11:0    & \\
+    11      & 4,9     & 1     & 4,9     & 4:0     & 11, 5\\
+    4       & 1       & 1     & 9,1     & 9:1     & 11\\
+    9       & 7,6,8   & 3     & 1,7,6,8 & 7:1     & 9, 8\\
+    1       &         & 3     & 7,6,8   & 6:2     & 9\\
+    7       & 5       & 3     & 6,8,5   & 8:3     & 9\\
+    6       & 5       & 2     & 8,5     & 5:1     & 7, 6\\
+    8       & 7       & 1     & 5       & 2:1     & 5, 3\\
+    5       & 4,2,3   & 2     & 2,3     & 3:2     & 5\\
+    2       & 1       & 1     & 3       &         & \\
+    3       & 2       & 0     &         &         & \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+In Table~\ref{tab:ex_traversal}, we follow the algorithm starting at
+merge commit 11 from the example in Figure~\ref{fig:repoDAG}, recording
+the current node, the parents of the current node, the depth after
+updating, and the queue structure after updating. The visited column is
+not dependant on the current step, but keeps track of the nodes that
+have been visited, and the depth at that node. When we reverse the DAG,
+the children in this table will become the candidates for being the
+parent of the corresponding node. For example, the parent of node 7 may
+be resolved to either node 9 or node 8. In this case, the shortest path
+to the master branch is through node 9.
+
+The terminology surrounding the parent/child relationship gets very
+confusing at the change between phase one and two because this is where
+the roles switch. In phase 1, we are working with the DAG; in phase 2,
+we are working with a structure that is neither a tree nor a DAG, but
+can only be described as a directed graph.
+
+The goal of the second phase is to resolve the tree from the directed
+graph. We will refer to parents in the same sense as we would refer to
+them in the tree form of the repository for the remainder of this
+section. We have gathered the depths and possible parents (line 17), but
+we need to remove the extra parents. Extra parents are culled in line
+42 of the algorithm. Of the possible parents, we take the parent that is
+closest to the root, but not beyond. This forces the real parent to
+either be deeper in the tree, or on the same branch. Line 30 to 37 are
+for bootstrapping the queue, ensuring that we are not starting on an
+empty queue.
+
+The construction of the tree becomes trivial now that we are able to
+determine the parents and children for each node. Lines 44 and 45
+perform this construction, adding the DAG parent to the tree children of the
+current node, and the DAG parent's parent becomes the current node,
+which was the DAG parent's child. We continue following the path until
+we hit the root node, merging the commit into the master branch.
+
+This algorithm relies heavily on the order of parents in the parent
+list, but is able to trim the number of nodes that must be downloaded.
+Furthermore, the resulting tree is able to show the order of commits,
+telling the story of how a developer was working, before the set of
+commits was merged. We developed a small bitbucket plugin to test the
+design and determine whether it was capable of producing merge trees,
+shown in Figure~\ref{fig:b_plugin}.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/plugin.png}
+  \caption{Screen shot of the Bitbucket plugin showing the online tree
+    computation of the repository shown in~\ref{fig:repoDAG}.}
+  \label{fig:b_plugin}
+\end{figure}

### 2017-07-07 paper/introduction.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

--- 2017-06-12 paper/introduction.tex
@@
-Merge-trees, as presented by Wilde and German\cite{Wilde2016}, are a
-data structure designed to make the navigation and summarization of git
-merges simpler than using the directed acyclic graph (DAG), as done in
-gitk among other tools\evan{does this need a citation?}. Git uses the
-DAG model internally, making it simple to construct visualization tools
-directly using the DAG of the repository;\@ however, this model does not
-provide a clear summarization of the events occuring in the repository.
-Gitk and the git command line interface (cli) both provide the user with
-a visualization of the DAG of the repository.
+% vim:set et sts=2 sw=2 ts=2 tw=72:
+Between 50k and 70k commits are added to the Linux kernel per version
+requiring maintainers of older versions of the kernel to sift through
+thousands of commits and merges with tools that are unable to filter and
+effectively visualize projects at the scale of the kernel. Older
+versions of the kernel are used in embedded systems and mobile phones;
+for security purposes, performance needs, and changing hardware
+requirements, maintainers must be able to understand the changes being
+made in the current version of the kernel in order to produce the
+necessary patches for the older versions of the kernel. Tools like Gitk
+use a directed acyclic graph (DAG) model of the repository, showing all
+commits and merges in chronological order by when the commit was
+authored, not by when it arrived in the official Linux repository.
 
-Merge-trees have two primary issues; first, the merge-tree model has not
-been evaulated, and cannot therefor claim to simplify the summarization
-and navigation tasks. Second, merge-trees have only been implemented for
-the Linux repository, which follows a strict merge-policy where only
-Linus can merge into the master branch.
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
+        \caption{A view of the Gitk interface centered on merge
+          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
+          their ancestors and there is no clear path from the commit to the merge
+          with the master branch. Neither Gitk nor Git are capable of showing the
+          commits in master.}
+        \label{fig:gitk}
+%\vspace{-4mm}
+\end{figure*}
 
-The contributions of this paper tackle both of these issues. We perform
-a user-study, to verify that the merge-tree model is able to improve the
-accuracy and speed of summarization tasks. Second, we generalize the
-algorithm, creating an upper bound on the number of commits visisted in
-the construction of the merge-trees.
+\begin{figure}
+        \centering
+        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
+        \caption{Github Failure Message showing the DAG of the Linux Project and its
+                relationship to other forks.}
+        \label{fig:gitfail}
+%\vspace{-2mm}
+\end{figure}
 
-\subsection{Merge-Trees}
-\label{sub:merge_trees}
+This representation works in smaller projects; it enables users to see
+when changes are made, when these changes are merged, how each branch is
+interacting, and the point where a branch forks from the master branch.
+In large modular projects, like the Linux kernel, the DAG becomes a mess
+of merges and commits (Figure~\ref{fig:gitk}) losing its visual meaning.
+In some cases, the Linux kernel is simply too large for the system to
+generate a visualization; Github provides a DAG view for many projects,
+but is unable to display the visualization for projects at the scale of
+the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and 70k
+new commits are created for the Linux project every year; according to
+our previous work\cite{German2015}, a commit takes a median of 30 days
+from the time it is authored until it arrives in the official
+repository. The snapshot of the kernel tomorrow may be different than
+the snapshot from today, containing new commits authored in the past;
+distinguishing these new commits from the commits in the snapshot from
+today is not trivial.
 
-Internally, git uses a DAG as the primary data structure for managing
-commits in a repository. An individual commit will have an ordered list
-of parents, where the parent in the first position is on the same branch
-as the commit, and the other parents are in the same order as they were
-merged. In the event that a merge commit contains a list of parents with
-more than two parents, this merge is an octopus merge. The commits may
-be divided into two groups, commits with only a single parent, and
-commits with at least two commits. Commits with with many parents are
-merge commits, combining the effects of potentially many individual
-commits.
+One major challenge with visualizing the arrival of commits to a
+repository is that Git does not store the date that a commit was merged
+into another branch, including the master branch. To complicate the
+problem, the DAG only has references to the ancestors of a commit (a
+model necessary for the operation of Git), but maintainers would prefer
+knowing the path a commit followed to reach the master repository.
+Tracing a path that any commit followed to the master repository would
+imply that for any given merge, it would be possible to know which
+commits were merged. A user could inspect the commits that arrived into
+the master branch within a given time-frame by checking which commits
+were merged during that time-frame.
+
+This paper makes two contributions; first, we describe a method of
+converting the DAG of the Linux repository into a tree, or
+\emph{merge-tree} of the repository, that represents the path used by a
+commit to reach the master branch; second, we present a method to
+inspect and visualize the history of merges in the Linux project using
+the merge-tree model.
+
+These methods and visualizations are implemented in a web-based tool
+called \tool\footnote{\tool is currently available at
+  \url{http://li.turingmachine.org}}.  Our visualizations and tool
+provide information about the location of any given commit or merge in
+its respective merge-tree, the files edited, the modules edited, and the
+commit message. \tool allows users to apply various filters, including
+the release version, along with a keyword or phrase from the log
+preview, the name of the author, or the commit ID. The user can request
+all merges made by Linus that contain a commit or inner merge that
+matches the search query, or just the commits and merges that match the
+query.

### 2017-07-07 paper/lineval.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

--- 2017-07-06 paper/lineval.tex
@@
 
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage[hyphens]{url}
-\urlstyle{same}
-
-%% Citations
-\usepackage[nospace]{cite}
-
-%% Table Support
-\usepackage{dcolumn}
-\usepackage{longtable}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage{tabulary}
-
-%% Extra support
-\usepackage{xspace}
-\usepackage{amsmath}
-\usepackage{balance}
-\usepackage{placeins}
-
-%% Algorithms
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-
-%% Graphics
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{xcolor}
-\usepackage{color}
-\usepackage{listings}
-
-%% Tikz
-\usetikzlibrary{positioning, arrows}
-
-\input{color.tex}
-\input{macros.tex}
-
-\newcommand{\TheTitle}{Merge-Tree: Visualizing the Integration of Commits into Linux}
-\newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
-\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Understanding a large number of commits and merges}
-\newcommand{\TheKeywords}{Linux, git, visualizations}
-
-%% Referencing
-\usepackage[unicode=true,
-          pdftitle={\TheTitle},
-          pdfauthor={\TheAuthors},
-          colorlinks=false]{hyperref}
-
-\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
-
-\lstset{frame=tb,
-  language=python,
-  aboveskip=3mm,
-  belowskip=3mm,
-  showstringspaces=false,
-  columns=flexible,
-  basicstyle={\small\ttfamily},
-  numbers=none,
-  numberstyle=\tiny\color{gray},
-  keywordstyle=\color{chartblue},
-  commentstyle=\color{chartred},
-  stringstyle=\color{chartgreen},
-  breaklines=true,
-  breakatwhitespace=true,
-  tabsize=3
-}
-
-\synctex=1
+\input{preamble.tex}
 
@@
 \begin{abstract}
+
+  \input{abstract.tex}
+
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:introduction}
 
-Between 50k and 70k commits are added to the Linux kernel per version
-requiring maintainers of older versions of the kernel to sift through
-thousands of commits and merges with tools that are unable to filter and
-effectively visualize projects at the scale of the kernel. Older
-versions of the kernel are used in embedded systems and mobile phones;
-for security purposes, performance needs, and changing hardware
-requirements, maintainers must be able to understand the changes being
-made in the current version of the kernel in order to produce the
-necessary patches for the older versions of the kernel. Current tools
-and visualizations are based on the directed acyclic graph (DAG) of the
-repository, showing all commits and merges in chronological order by
-when the commit was authored, not by when it arrived in the official
-Linux repository.
+\input{introduction.tex}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/gitk.png}
-  \caption{
-    The Gitk interface, centered at commit
-    cdbdd1676a5379f1d5cbd4d476f5e349f445befe. The top left pane showing
-    the DAG, the top center pane showing the authors, and top right pane
-    showing the commit dates. The bottom-left pane shows the full
-    commit message and diff of the selected commit, and the bottom-right
-    pane shows the list of the files modified by the selected commit.
-  }
-  \label{fig:gitk}
-  %\vspace{-4mm}
-\end{figure}
+\section{Merge-Tree model}
+\label{sec:mergetree}
 
-Git is a distributed version control system that is used in many
-open-source and closed-source projects. Git enables many developers to
-write code simultaneously, working on various components known as
-branches. Git uses a directed acyclic graph model (DAG) for storing the
-necessary information regarding the state of the code at any point in
-time. Git refers to nodes of the DAG as commits, not distinguishing
-between the code-carrying commits and the structural commits that bring
-two branches together. In this paper, we will refer to the individual
-commits as being commits, the structural merge commits as merges, and
-refereeing to all as repository events.
+\input{model.tex}
 
-The DAG representation works in smaller projects; it enables users to
-see when changes are made, when these changes are merged, how each
-branch is interacting, and the point where a branch forks from the
-master branch. In large modular projects, like the Linux kernel, the DAG
-becomes a mess of merges and commits (Figure~\ref{fig:gitk}) losing its
-visual meaning. In some cases, the repository is simply too large for
-systems to generate a visualization; repositories stored on GitHub will
-have a DAG visualization to show the branches and forks of a
-repository, but is unable to display the visualization for projects at
-the scale of Linux or even the Git Git
-repository(Figure~\ref{fig:gitfail}). Between 60k and 70k new commits
-are created for the Linux project every year; according to previous
-work\cite{German2015}, a commit takes a median of 30 days from the time
-it is authored until it arrives in the official repository. The snapshot
-of the kernel tomorrow may be different than the snapshot from today,
-containing new commits authored in the past; distinguishing these new
-commits from the commits in the snapshot from today is not trivial.
+\section{Visualizing the merge-tree of Linux}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
-  \caption{Github is unable to display the DAG of the Linux
-    Project due to the size and complexity, which would normally
-    show the relationship to other branches and forks.}
-  \label{fig:gitfail}
-  %\vspace{-2mm}
-\end{figure}
+The goal of \tool is to simplify the navigation of the kernel commit
+information, specifically focusing on merges. This is done by leveraging
+the merge-tree to inspect how commits are merged on the path to the
+master repository.
 
-One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch, including
-the master branch. To complicate the problem, the DAG only has references to the
-ancestors of a commit (a model necessary for the operation of Git), but maintainers
-would prefer knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would imply that
-for any given merge, it would be possible to know which commits were merged. A user
-could inspect the commits that arrived into the master branch within a given
-time-frame by checking which commits were merged during that time-frame.
+\subsection{Use cases}
 
-This paper makes two contributions; first, we describe a method of converting the
-DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach the master branch; second, we
-present a method to inspect and visualize the history of merges in the Linux project
-using the merge-tree model.
+We designed \tool with two use-cases in mind, though a user may switch
+between the cases as they work.
 
-These methods and visualizations are implemented in a web-based tool called
-\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
-Our visualizations and tool provide information about the location of any given
-commit or merge in its respective merge-tree, the files edited, the modules edited,
-and the commit message. \tool allows users to apply various filters, including the
-release version, along with a keyword or phrase from the log preview, the name of
-the author, or the commit ID. The user can request all merges made by Linus that
-contain a commit or inner merge that matches the search query, or just the commits
-and merges that match the query.
+\noindent \textbf{Use-case 1: top-to-bottom
+  approach}\label{sec:usecase1}\\ These are users that are maintaining a
+section of the kernel and would like to pick a merge (including all the
+commits that it merges) and merge it directly into their current
+repository. This is useful for reducing the amount of re-implementation
+work. For these users, it is important to have the ability to aggregate
+metadata about files and modules being effected by the merge. Also, it
+is important for these users to be able to navigate from the root of the
+merge-tree toward the leaves.
 
-\section{\mt Model}
-\label{sec:mt_model}
+\noindent \textbf{Use-case 2: bottom-to-top
+  approach}\label{sec:usecase2}\\ These are users that start with a
+known merge or commit and would like to see what other changes are being
+made in commits that are in the same merge, including knowing the
+merge-tree they belong to. This is useful to see what other commits are
+related to the current commit and how they get collated into merges that
+eventually end in the master branch. This is primarily for maintainers
+that need to perform some specific cherry picking of commits. We must
+provide these users a mechanism for navigating from a single commit
+toward the master branch, allowing them to see other commits that might
+be related to their original commit.
 
-We propose a new model, built from the DAG, with the purpose of
-providing a clearer conceptual image of how a commit is merged, other
-commits that are merged with a commit, and a summarization of the entire
-group of commits that are merged with a commit. We call our model the
-\mt model. The model inverts the DAG, so that reference point from
-parent to child instead of from child to parent. This enables the
-aggregation of commit metrics at merges. Given a merge, we are able to
-recursively traverse the children of the merge, collecting the
-statistics at each commit. The commits become the leaves of the tree,
-and internal nodes are merges that do not merge into the master branch
-of the repository. We terminate the further traversal of an edge under
-two conditions; first, if the next event is on the master branch of the
-repository, second, if the next event has a shorter path into the master
-branch of the repository. As a consequence, commits will always take the
-shortest path, in terms of commit time, to the master branch.
+\subsection{Data Model}
 
-A short example: assume the commits represented in
-Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
-In this case, commits are performed in various repositories and
-branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
-the master branch and the repository that the master branch is part of.
-The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
-Note that the direction of the edges of the DAG have been inverted,
-instead of pointing from the child to the ancestors, it points from the
-ancestor to its successors, forming a path to the master branch. Also
-note that the DAG has been simplified, showing only a single edge on the
-path to master for any commit.
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
-    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
-
-    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
-    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
-    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
-    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};
-
-        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
-    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
-  \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
-  \label{fig:repoEvents}
-%\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state] (1) {1};
-    \node[state, above right= of 1] (2) {2};
-    \node[state, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, right= 2cm of 1] (4) {4};
-    \node[state, above right=of 4] (5) {5};
-    \node[state, above right=of 5] (6) {6};
-    \node[state, right=of 5] (7) {7};
-    \node[state, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, right= 2cm of 7] (9) {9};
-    \node[state, right=of 9] (10) {10};
-    \node[state, below right=of 9] (11) {11};
-    \node[state, below right=of 10] (12) {12};
-
-    \draw (12) edge[-stealth] (11) edge[-stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[-stealth] (9);
-    \draw (10) edge[-stealth] (9);
-    \draw (9) edge[-stealth] (8) edge[-stealth] (6)
-              edge[-stealth] (7);
-    \draw (8) edge[-stealth] (7);
-    \draw (7) edge[-stealth] (5);
-    \draw (6) edge[-stealth] (5);
-    \draw (5) edge[-stealth] (3) edge[-stealth] (2)
-              edge[-stealth] (4);
-    \draw (4) edge[-stealth] (1);
-    \draw (3) edge[-stealth] (2);
-    \draw (2) edge[-stealth] (1);
-  \end{tikzpicture}
-  \caption{DAG representation of the commits represented in
-    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
-    information about which repository a commit comes from. Furthermore,
-    the DAG does not distinguish the master branch from other branches.}
-  \label{fig:repoDAG}
-\vspace{-3mm}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-
-    \draw[chartblue]
-      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);
-
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
-
-    \draw (2) edge[chartyellow, -stealth] (5)
-          (3) edge[chartmagenta, -stealth](5)
-          (5) edge[chartyellow, -stealth] (7)
-          (7) edge[chartyellow, -stealth] (9)
-          (6) edge[chartred, -stealth] (9)
-          (8) edge[chartmagenta, -stealth] (9)
-          (9) edge[chartyellow, -stealth] (11)
-          (10) edge[chartyellow, -stealth] (12);
-
-  \end{tikzpicture}
-  \caption{\mt view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
-\vspace{-3mm}
-\end{figure}
-
-Computing the \mt from a DAG for any repository may not be
-possible; however, certain features of the development process of Linux
-make it feasible to compute the \mt for the Linux repository.
-First, the master branch of Linux is maintained by Linus Torvalds, and
-only Linus has write access to it. We have verified this assertion in
-previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
-identifies the commits made directly to the master branch; whereafter it
-recursively determines the shortest path (in terms of time), using the
-DAG, from each commit to the master branch using the inverted DAG.
-
-\begin{algorithm}
-        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
-        \label{alg:original}
-        \begin{algorithmic}
-                \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
-                \State $head \gets \textit{Head of master of git repository}$
-                \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
-                \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
-                \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
-                \If {\textit{cid in master}}
-                \State \Return 0
-                \EndIf
-                \State    $d \gets 	\infty$
-                \State {\# Traverse the inverted DAG}
-                \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
-                \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
-                \EndIf
-                \EndFor
-                \State {\# $c$ is the commit that follows $cid$}
-                \State {\# in its way to master}
-                \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
-                \EndFunction
-
-                \State {\# Compute the distance for each commit}
-                \State {\# discarding result}
-                \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
-                \EndFor
-                \State \Return $Tree$
-                \EndFunction
-        \end{algorithmic}
-\end{algorithm}
-
-We evaluate the performance of the heuristic using the merge log made by
-Linus Torvalds, the original creator of Linux, and primary merger of all
-other repository events into the master branch of the repository. In
-Figure~\ref{fig:sampleMerge}, we see an example of the merge log.
-
-\begin{figure}[htbp]
-  \centering
-  {\fontsize{7}{9}
-    \begin{verbatim}
-Merge: 8cbd84f fd8aa2c
-Author: Linus Torvalds <torvalds@linux-foundation.org>
-Date:   Tue Aug 10 15:38:19 2010 -0700
-
-Merge branch 'for-linus' of git://neil.brown.name/md
-
-* 'for-linus' of git://neil.brown.name/md: (24 commits)
-md: clean up do_md_stop
-[... edited for the sake of space]
-md: split out md_rdev_init
-md: be more careful setting MD_CHANGE_CLEAN
-md/raid5: ensure we create a unique name for kmem_cache...
-...
-    \end{verbatim}}\vspace{-5mm}
-  \caption{Example of how merges record a subset of commits being
-    merged. The commit only shows the first 20 one-line summaries
-    messages for the 24 non-merge commits it merged. The ending
-    ``\ldots'' is part of the log and represents that other
-    commits were merged.}
-  \label{fig:sampleMerge}
-\end{figure}
-
-We use this information to evaluate the accuracy of the \mt model
-extracted from the DAG\@. The method we followed started with the
-extraction of the commit history up to July 20, 2016. We computed the
-merge tree of every commit until then. Since Linus Torvalds mostly does
-merging directly into master, we assumed that every merge by him is the
-root of a \mt. As described above, the log of a merge-commit usually
-contains the number of commits in the merge the first 20 summaries of
-commits being merged. We extracted merges by Linus Torvalds using the
-command \mycode{log --merges --author='Torvalds} and compared the number
-of commits according to the log with the number of commits in the  \mt
-rooted in this commit. We also used the summaries of the commits found
-in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the
-merge in Figure~\ref{fig:sampleMerge} we would expect that the merge
-tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
-summaries corresponds to commits in that \mt. We also inspected those
-with differences to make sure they were true errors.  The results can be
-summarized as follows:
-
-\begin{itemize}
-
-  \item
-
-    Five merges were false-errors because their logs did not contain
-    accurate information (were probably edited by hand). For example in
-    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
-    empty), in \mycode {c55d267\ldots} the summaries were reordered.
-
-  \item
-
-    The heuristic correctly identified that 79 of Linus merges (between
-    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
-    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
-    commits.
-
-  \item
-
-    The heuristic worked perfectly until Sept 4, 2007, the earliest date
-    that it could be verified.  Before this date, and until Dec 12,
-    2006, merges did not include a summary of the commits they included,
-    hence making it impossible to verify; during this period, however,
-    we correctly identified the merges by Linus into master.
-
-  \item
-
-    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
-    which confounded the true master branch of a repository (see
-    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
-    for a description of the issue).
-
-\end{itemize}
-
-In summary, of the merges after Sept. 4, 2007, our heuristic was correct
-in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
-12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
-2007).
+In our visualizations, we leverage the merge-tree model described in
+Section~\ref{sec:mergetree}. In this model each commit is either already
+in the master branch, or is part of a tree which is rooted in the merge
+that merged it into the master branch.  Each commit, whether a merge or
+non-merge, has only one successor; the root of each tree has none as it
+was made by Linus Torvalds directly into the master branch. Non-merge
+commits contain the metadata for the changes made.  This metadata
+includes the files changed, the lines added and removed from each file,
+the author, the date the commit was merged into the merge that led to
+being merged into the kernel, the date the commit was authored, the
+patch, and the commit log. Merges contain less metadata, only storing
+the author of the merge, the log, the commit date, and the author date,
+and potentially, changes necessary to address conflicts during the
+merge. The details of the model are outlined in \cite{German2015}.
 
 \section{Design and Implementation}
-\label{sec:design_and_implementation}
 
-We implement a tool, \tool, to test and verify the capabilities of the
-\mt model. \tool provides search, summarization, and
-visualization using the \mt model.
-
-We provide users with search-based navigation. The search engine takes a
-plain text query from the user and returns results based on similarity
-to the query. The engine weighs various fields, including the commit
-log, the author name, the commit hash, and the dates that the commit was
-authored and committed. The results are aggregated by the \mt
-with a link to the root at the top, and the repository events that match
-the search query in a table below.  An example of a single \mt is shown
-in Figure~\ref{fig:linvis_search_results}.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/search_results.png}
-  \caption{\tool search results for `net-next 2011', showing a single
-    \mt group.}
-  \label{fig:linvis_search_results}
-\end{figure}
-
-\tool uses seven tabs to show the information and visualizations for a
-selected commit or merge. The message tab shows the full commit log
-message. This does not include the diff. The files tab shows an
-aggregated table of all files that were modified in a merge. It includes
-metrics like the number of lines added, lines removed, total lines
-modified, and the delta, summed across all commits in the merge tree
-that modify this file. A small details drop-down button allows a user to
-see exactly which commit makes the changes, as shown in
-Figure~\ref{fig:linvis_files}. If the current repository event being
-viewed is a commit, the aggregate views will only show modifications
-made by the commit. The modules tab shows the modules modified in the
-merge tree. Like the files tab, the modules tab uses a table to show the
-name of the module, the number of commits that are in the merge tree
-that work with the module, and a details button to provide the links to
-those commits.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_files.png}
-  \caption{Table showing the modified files in a merge, with the second
-  entry expanded to show the commit that makes the changes.}
-  \label{fig:linvis_files}
-\end{figure}
-
-The authorship tab is very similar to the files tab, but shows the
-authorship information. It shows the sum of the number of lines added,
-removed, modified, and the delta within the \mt. It also shows
-the number of files that were modified by the author. The details tab is
-organized slightly differently. Instead of organizing these results by
-commit, the details are organized by file, showing which files were
-modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_authors.png}
-  \caption{Table showing the authors who made changes in a merge. The
-    entry for Randy Dunlap is expanded, showing the modifications to
-    each file that were made by Randy.}
-  \label{fig:linvis_authors}
-\end{figure}
-
-\tool implements three visualizations, a list tree, pack tree, and
-Reingold-Tilford tree. The list tree is a series of indented textual
-lists that can be easily searched with the built-in search in most
-web-browsers. The list tree is rooted at the current event, if the event
-is a commit, only that commit will be shown; conversely, if the selected
-event is the root, the entire merge tree will be shown. The Pack-Tree or
-Bubble tree, uses a set-based approach for showing the parent
-relationship. The root node aggregates all of the events within the \mt
-into the master branch. Thus, the root is the superset of all the other
-nodes in the tree. The commits are shown as individual white circles,
-containing no other events, as these are the leaves. This is shown in
-Figure~\ref{fig:linvis_pack}. The selected event will be shown in bright
-orange, commits in white, and merges in shades of blue. Darker shades of
-blue indicate that the merge is deeper in the tree.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
-  \caption{The pack tree visualization for a selected commit, shown in
-    orange.}
-  \label{fig:linvis_pack}
-\end{figure}
-
-The Reingold-Tilford tree is the classic tree visualization, with the
-root at the top, leaves at the bottom, and the edges between them
-showing the parent relationship. An example of the Reingold-Tilford tree
-visualization is shown in Figure~\ref{fig:commit_1} and
-Figure~\ref{fig:commit_2}.
+\input{design.tex}
 
@@
 
-We use a controlled mixed-methods users study in order to evaluate the
-ability for the DAG to provide users with a conceptual image of how a
-commit is integrated into a project, to compare the DAG and \mt models,
-and determine which model the users preferred.
-
-In this section, we describe the methods used to for designing and
-implementing the user study.
-
-\subsection{Questions and Tasks}
-\label{sub:questions}
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Conceptual Tasks }
-  \label{tab:conceptual_tasks}
-  \begin{tabulary}{0.9\textwidth}{LL}
-    \toprule
-    Task & Description\\
-    \midrule
-    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
-    T2 & How many individual commits are related to this commit?\\
-    T3 & How many merges are involved with merging this commit into the master branch?\\
-    \bottomrule
-  \end{tabulary}
-\end{table*}
-
-\begin{table}[htpb]
-  \centering
-  \caption{Summarization Tasks}
-  \label{tab:summarization_tasks}
-  \begin{tabulary}{\linewidth}{LLL}
-    \toprule
-    Task Set   & Task & Description\\\midrule
-    Merge      & T4   & What is the series of merges involved with merging this
-    commit?\\
-               & T5   & What other commits are merged?\\
-    Authorship & T6   & How many authors are involved?\\
-               & T7   & Who contributed the most changes?\\
-    Files      & T8   & How many files were modified?\\
-               & T9   & Which file had the most changes?\\
-    Modules    & T10  & Which modules does this merge tree involve?\\
-    \bottomrule
-  \end{tabulary}
-\end{table}
-
-The questions and tasks are broken down into four groups. The first two
-groups are purely quantitative in nature; we begin with conceptual tasks
-to understand whether the DAG is capable of providing users with an
-understanding of how a commit is merged into the master branch, and
-other commits that it is merged with. We continue with summarization
-tasks, to quantitatively understand if the \mt model makes a difference
-in summarization tasks.
-
-The third task set is purely qualitative, to gain better understanding
-of whether participants preferred working with the \mt model in \tool,
-or working with the DAG in Gitk, and what aspects they preferred from
-each tool. The final group of questions is simply to gain a better
-understanding of the demographic of our participants.
-
-\textbf{Conceptual Tasks}
-
-The participants were allowed to use Gitk or the git command line tools
-to perform these tasks. These tasks are asked in the specified order,
-allowing tabularr to first, create a conceptual understanding of the
-commits and merges that will be involved with the rest of the tasks that
-follow.tabularmay then use their diagrams to answer T2 and T3. We do not
-collect metrics from \tool for the conceptual tasks, opting instead to
-determine if the DAG is sufficient for understanding how a commit is
-integrated. \tool uses the \mt model show a tree visualization of how
-commits are merged, until the commit reaches the master branch.
-Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.
-
-We provide the participant with 10 minutes per commit to complete the
-first task. The answers to the two other questions in this section are
-drawn from the answer in the first task.
-
-\textbf{Summarization Tasks}
-
-The summarization tasks (Table~\ref{tab:summarization_tasks}) are
-grouped into task sets by theme. The tasks within the task set are
-shuffled, and the task sets are shuffled, to ensure that similar
-questions are together, but not in the same order, and to ensure that
-the themes are in randomized order between participants. The order of
-the commits is shuffled, as well. The order of the tool is randomized,
-choosing to start with either \tool or Gitk, to ensure that we do not
-place a bias on one tool through the experiment.
-
-As understanding where the merge begins and ends is vital to answering
-the other question, the merge task set will always be the first task set
-presented; however, the questions will still be re-ordered within the
-task set.
-
-\textbf{User Opinion and Profile}
-
-Finally, we ask the user for their opinions on the tools and for some
-information about their level of experience with git. This portion of
-the study is qualitative and does not rely on either of the tool,
-instead attempting to determine which model provided the user with a
-better experience.
-
-We do not shuffle the questions in this set of questions, as this should
-not have an impact on the opinions of the participant at this point.
-These questions are designed to provide us with a better understanding
-of preference, and insight on why, the participant had that preference.
-
-We ask the following questions in this order;
-
-\begin{enumerate}
-  \item Given these tasks again, which tool would you prefer to use?
-  \item Which aspects of each tool did you like and why?
-  \item How long have you used git?
-  \item If you have used git, for what kind of projects? (personal,
-    school courses, professional?)
-  \item If you have used git, how many commits, files, and contributors
-    were involved with the largest repository you have worked with?
-\end{enumerate}
-
-The last three questions in this section are aimed at getting a better
-understanding of the demographic of our participants.
-
-\textbf{Procedure}
-
-For each participant, we run a simple python script to generate the
-script for the study, eliminating the possibility of bias from us. The
-script randomizes the order of the tools, the order of the commits, the
-order of the task sets, and the order of the tasks within each task set,
-where appropriate.
-
-\subsection{Commit Selection}
-\label{sub:commit_selection}
-
-We chose two commits for use in the study. The order that the commits
-are presented is randomized between participants, but the order is kept
-consistent through the tasks described in the next subsection.
-
-We use the \tool database, containing commits and merge-tree information
-from April 16th 2005 to October 14th, 2014, which corresponds to being
-between Linux release 2.6.12-rc3 and Linux 3.17-rc1. \evan{Cite the
-  database even more?} We chose the commits based on tree sizes. We
-found that a majority of the trees contain at most seven inner commits,
-while more than 25\% of the trees contain a single commit. 75\% of the
-tree contained up to 51 inner commits and merges, and finally the
-largest tree contained 7217 nodes. We chose to work with the trees in
-the first and second quartile, as merge trees of sizes between one and
-seven, not including the merge into the master branch, make up the
-majority of the trees in the database.
-
-From here, we selected one tree from the trees of a single commit at
-random. Selecting a commit from that tree is trivial, as there is only
-one to choose.
-
-We selected one tree of size seven at random from all trees of size
-seven. We placed a restriction the trees, the selected tree must include
-at least one internal merge to increase the complexity of the trees
-tested. After randomly selecting a tree, we chose in internal commit at
-random, using the \verb|random.choice| function in python 3.6.1.
-
-Using this technique, we selected commit
-\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
-a single node (Figure~\ref{fig:commit_1}), which we will refer to as
-\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
-the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
-will refer to as \comB.
-
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.08\linewidth]{figures/commits/1-commit.pdf}
-  \caption{The first merge tree used in the user study, a merge tree
-    containing a single commit}
-  \label{fig:commit_1}
-\end{figure}
-
-\begin{figure}[bpt]
-  \centering
-  \includegraphics[width=0.5\linewidth]{figures/commits/7-commits.pdf}
-  \caption{The second merge tree used in the user study, a merge tree
-    containing seven commits}
-  \label{fig:commit_2}
-\end{figure}
-
-\subsection{Results}
-\label{sec:results}
-
-We recorded the audio and screen capture video for each participant in
-the study for further analysis. After capturing the information, we
-extracted various metrics from the videos. From the conceptual tasks, we
-recorded the time in seconds, and the answer for the question. We
-extracted three metrics for the summarization tasks, the time in
-seconds, the correctness, and the accuracy. Correctness is a binary
-value, it is true if the response was correct and false if the response
-was incorrect. The accuracy is measures in distance from the correct
-value. An accuracy of zero indicates a correct answer. All tasks in the
-summarization tasks include an accuracy metric except T7, which will
-only be right or wrong, and therefore cannot meaningfully be measured by
-a distance.
-
-\subsection{Conceptual Tasks}
-\label{sub:conceptual_tasks}
-
-In Table~\ref{tab:conceptual_results}, we are able to see an overview of
-the results from the conceptual questions.
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Results from the conceptual questions}
-  \label{tab:conceptual_results}
-  \begin{tabular}{ll|r|lrr|rrr}
-    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
-    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
-    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
-    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
-    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
-  \end{tabular}
-\end{table*}
-
-Users were able to more closely estimate the number of commits and
-merges in the larger tree, but generally took longer than the smaller
-tree. The tree with a single node resulted in more variability in the
-estimate of number of commits.
-
-It should be noted that these questions were answered after spending
-roughly ten minutes attesting to draw a picture that held the answers to
-these questions.
-
-\subsection{Summarization Tasks}
-\label{sub:summarization_tasks}
-
-In Figure~\ref{fig:summarization_results}, we see all of the
-correctness, accuracy, and timing results. The first row of plots are
-the correctness results. We see in all cases except for T10, that the
-number of instances of a correct result with \tool is roughly the same
-as the number of incorrect instances with Gitk. We test each task with
-the McNemar Chi-square test~\cite{McNemar1947}.
-
-\begin{figure*}[htpb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
-  \caption{Results from the summarization tasks. The first row plots the
-    correctness metric for each task. The second row plots the
-    accuracy metric for each task. The third row plots the time metric
-    for each task.}
-  \label{fig:summarization_results}
-\end{figure*}
-
-With the McNemar's tests, we analyze the participants that change from
-being correct with one tool to being incorrect with the other. This does
-not take into account the number of people who were able to provide the
-correct answer for both tools, nor the people who were incorrect for both
-tools. The results show whether providing the participants with \tool
-over the current standard, gitk, improves the correctness of the
-summarization tasks among the participants. We show the results for the
-number of people who were, with both tools, correct, incorrect, and
-changed in Table~\ref{tab:correctness_table}.
-
-In all cases, except for task T10, we saw statistically significant
-results suggesting that we reject the null hypothesis. We tested each
-correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
-degree of freedom is $7.879$.
-
-\begin{table*}[htpb]
-  \centering
-  \caption{Correctness tables showing how many people stayed correct,
-    incorrect, and how many people changed between tool}
-  \label{tab:correctness_table}
-  \begin{tabular}{c|c|c}
-
-    T4 & T5 & T6\\
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 15                           & 2\\
-  \end{tabular} &
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
-                           & Incorrect & 14                           & 4\\
-  \end{tabular}
-
-  &
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
-                           & Incorrect & 16                           & 2\\
-  \end{tabular}
-
-  \\\hline
-  T7 & T8 & T9\\
-
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 8                            & 0\\
-                           & Incorrect & 13                           & 2\\
-  \end{tabular}  &
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 15                           & 2\\
-  \end{tabular} &
-  \begin{tabular}{cc|rr}
-                           &           & \multicolumn{2}{c}{Linvis}\\
-                           &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-                           & Incorrect & 14                           & 3\\
-  \end{tabular}\\\hline
-
-  & T10 & \\
-  & \begin{tabular}{cc|rr}
-                            &           & \multicolumn{2}{c}{Linvis}\\
-                            &           & Correct                      & Incorrect\\\hline
-    \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
-                            & Incorrect & 8                            & 0\\
-  \end{tabular} & \\
-
-  \end{tabular}
-\end{table*}
-
-The correctness test statistics and conclusion for each summarization
-task;
-
-The correctness test statistics, conclusions, and a description of the
-accuracy and timing results for each summarization task follow:
-
-\begin{itemize}
-
-  \item
-
-    T4; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the series of merges that a commit
-    takes, with $99.5\%$ confidence.
-
-    There is almost no variation in accuracy between participants using
-    \tool, while there is more variance in Gitk. There were two outliers
-    among the instances using \tool, which lie within the second and
-    third quartiles for the accuracy measure of Gitk. The furthest
-    outlier for \tool lies at a distance that is slightly greater than
-    the median for Gitk.
-
-    We see similar results in the timing metric; all significant data
-    occurs within less time than the second, third, and fourth quartile
-    of the Gitk results. Nearly all outlier instances in \tool are
-    within the third quartile group of the instances of Gitk with the
-    exception of the most extreme, which lies in the fourth quartile
-    group.
-
-    The results indicate that \tool is able to assist users with
-    determining how a commit is merged more accurately, and more
-    quickly.
-
-  \item
-
-    T5; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the other related commits, with
-    $99.5\%$ confidence.
-
-    There is very little variance in the accuracy of results for \tool
-    compared to the accuracy of the results in Gitk. The median distance
-    is smaller, indicating that the participants performed more
-    accurately using \tool.
-
-    There is less variance in the time taken to come to an answer with
-    \tool than with Gitk. Furthermore, we see that a majority of the
-    participants were able to complete the task in \tool in much less
-    time than 50\% of the participants using Gitk.
-
-    The results indicate that \tool is able to assist users in
-    understanding which commits are related to a given commit.
-
-  \item
-
-    T6; the resulting $\chi^2$ value is $16$. $16 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the number of authors involved with
-    the merge tree, with $99.5\%$ confidence.
-
-    We see the same results from this task as with the previous tasks,
-    suggesting that \tool assists users determine the number of authors
-    more quickly and more accurately.
-
-  \item
-
-    T7; the resulting $\chi^2$ value is $13$. $13 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining who contributed the most changes,
-    with $99.5\%$ confidence.
-
-    This question does not include an accuracy metric, as there are only
-    two possible outcomes. We see the same result in the timing metric
-    as we did from the previous questions.
-
-    The results suggest that \tool is able to help users correctly
-    identify the person who contributed the most lines of code to a
-    merge tree in less time.
-
-  \item
-
-    T8; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining how many files were contributed,
-    with $99.5\%$ confidence.
-
-    The results of this task are the same as in the previous tasks.
-    This suggest that \tool is able to assist users determine how many
-    files were modified, more quickly and more accurately.
-
-  \item
-
-    T9; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining which file had the most changes,
-    with $99.5\%$ confidence.
-
-    While this question appears to have a similar answer to the correct
-    answer for T7, the answer \comA had two files with the same number
-    of midifications, thus the participant must correctly identify both
-    of these files to be correct. For this reason, we include an
-    accuracy metric.
-
-    The results of this task are the same as the previous tasks.
-    This suggests that \tool is able to assist users to determine which
-    file had the most changes.
-
-  \item
-
-    T10; the resulting $\chi^2$ value is $5.44$. $5.44 < 7.879$,
-    therefore we do not reject the null hypothesis, suggesting that
-    \tool does not make a difference in determining which modules are
-    involved with a merge tree.
-
-    The accuracy and timing results for this task are the same as the
-    others, showing that there is less variability in both accuracy and
-    the time taken to have an answer between participants using \tool
-    than using Gitk. Looking Table~\ref{tab:correctness_table}, we can
-    see that, unlike the other questions, a majority of the participants
-    were correct using both tools.
-
-\end{itemize}
-
-\subsection{User Opinions}
-\label{sub:user_opinions}
-
-Among the 12 pariticipants, there was nearly unanimous agreement that
-for conceptual understanding and summarization tasks, \tool was easier
-to work with than Gitk. The participants cited the ability to abstract
-information about the merge from the clean summarization tables and
-simple visualizations. Three participants suggested that someone with a
-professional understanding of Gitk and the git commandline may be to
-extract the a good conceptual understanding from the DAG, and perform
-the summarization tasks from the commandline. One of these three
-participants said that they would prefer to have both tools available,
-as they are able to complement eachother.
+\input{study.tex}
 
@@
 
-% TODO: Go back to the conceptual task results and make them stronger
-Overall, the results show fairly conclusive results; the DAG is unable
-to adequately provide users with an explanation of the changes being
-made in the repository, the DAG is not easily understood, and the DAG is
-unable to provide summarizations of the changes being made at a merge.
-Meanwhile, the \mt model is able to provide a better explanation of how
-a commit is integrated, along with better summarization capabilities. In
-this section, we discuss some observations made during the study,
-threats to the validity of the study, comments made by the participants
-during the study, modifications to the original algorithm to help
-generalize it for more repositories, and future work.
-
-\subsection{Observations}
-\label{sub:observations}
-
-% TODO: include info on merge tree sizes
-
-During the course of the study, we noticed a consistent theme among the
-participants. The most difficult challenge was identifying the master
-branch of the repository. Given that single-commit merge trees are the
-most common tree, and that they are relatively simple, we were
-interested in seeing if the participants would be able to identify the
-tree easily, but the results suggest that this was not the case. The
-issue primarily stemmed from not being able to identify the master
-branch. Most of the participants realized that they should not be going
-backward, toward the initial commit, very early on. Furthermore, many
-would identify the correct commit; however, they would start listing
-every merge along the master branch as being a merge in the tree, and
-including the commits from that merge. In the case of the single commit,
-simply identifying the master branch would most likely be sufficient in
-providing a clearer understanding of how a commit or flat tree is
-integrated into the master branch of the repository.
-
-While, in general, the participants of the study preferred using \tool
-to Gitk, there were some noticeable usability issues. The most prominent
-issue being the navigation within the tree. The participants expected
-that the tabs would update when they clicked on a node in the tree,
-instead of having to click the link to the commit after clicking the
-node. This was true in both the Reingold-Tilford tree and the Pack tree.
-The second major usability issue was the inconsistency between the list
-tree and the other tree visualizations. The list tree is rooted at the
-repository event that the user is current looking at, while the other
-tree visualizations show the entire tree, and use bright orange to
-highlight the current node.
-
-\subsection{Comments From a Release Manager}
-\label{sub:comments_from_a_release_manager}
-
-One of the participants of the study had worked as a release manager for
-more than three years, working with both SVN and CVS repositories, but
-had little experience with git.
-
-Contributors making merges need to understand more than just what merges
-a commit was collected into before reaching the repository of the
-contributor. It is also important to understand order that the related
-commits were made, as they tell the story of what the developer was
-thinking as they were writing the changes. The current merge-tree model
-does not order the commits, simply placing them randomly. This is the
-primary reason behind this participant's request to use both
-simultaneously. \tool is able to help with the aggregation of the
-information, and provide a better understanding of the next merge
-involved in integrating this commit, but the DAG in gitk provides the
-full story of the commit instead of hiding it behind a layer of
-abstraction.
-
-The comments from this participant were very insightful, and will assist
-in improving the model in the future.
-
-\subsection{Threats to Validity}
-\label{sub:threats}
-
-While git includes gitk and the git log provides a view of the DAG with
-\verb|git log --graph|, many of the participants in our study were
-unfamiliar with the DAG visualizations shown by these tools. A possible
-reason for this may be that unless a repository reaches a sufficient
-complexity, both in branching and number of commits, these tools are
-necessary for the normal operation of git. Furthermore, fear of the
-complexity caused by branching may cause personal and academic
-repositories to have relatively simple structures. While they were
-unfamiliar with the DAG structure, these users were also unfamiliar with
-the \mt model, and were still able to better understand and summarize
-the repository using \tool.
-
-Each question was evaluated independently of the results from the
-previous question in the summarization tasks. Had we modified the
-correct answer to fit the conceptual understanding of which commits were
-included in the tree, we may have witnessed different results. A
-modification to the study may include providing the participants with
-the correct set of commits prior to the summarization tasks, after the
-merge task set.
-
-\subsection{Generalization}
-\label{sub:generalization}
-
-Seeing that the \mt model is able to assist users with understanding how
-commits are integrated, we continue by looking at ways of improving the
-algorithm for faster operations, and making the conversion from DAG to
-\mt feasible over the internet for smaller repositories.
-
-The problem of finding the master branch is the most difficult, and is
-not possible without heuristic approaches. The master branch can be
-confounded by \textit{foxtrot} merges, which change the order of the
-parents in the parent list of a merge. Other than the ordering of the
-parent list, git has no other way to determine which commits are on the
-master branch of the repository. For the purpose of this discussion, we
-will assume that we can rely on the order of the parent list to be
-correct, and that foxtrots do not occur.
-
-\begin{algorithm}
-  \caption{Computing the generalized Merge Tree}
-  \label{alg:generalized}
-  \begin{algorithmic}[1]
-    \Function{Phase 1}{initial commit $root$} : updated tree
-    \State $depth \gets 0$
-    \State $Q \gets root$
-    \Do
-    \State $cur \gets Q.dequeue$
-    \State $parents \gets cur.parents$
-    \State $depth \gets depth + parents.length - 1$
-    \For{$index, parent \in parents$}
-    \If{$parent.depth$ is $undefined$}
-    \State $parent.depth \gets cur.depth + index$
-    \ElsIf {$cur.depth > parent.depth$}
-    \State $depth \gets depth - 1$
-    \ElsIf {$cur.depth < parent.depth$}
-    \State $depth \gets depth - 1$
-    \State $parent.depth \gets cur.depth$
-    \EndIf
-    \State $parent.children \gets cur$
-    \If{$parent$ has not been visited}
-    \State $Q \gets parent$
-    \State $parent.visited \gets True$
-    \EndIf
-    \EndFor
-    \doWhile{$Q$ not empty and $depth \ne 0$ }
-    \EndFunction
-
-    \Function{Phase 2}{initial commit $root$} : Merge tree
-    \State $tree.root \gets root$
-    \State $Q$ // Empty Queue
-
-    \State $parents \gets root.parents$
-    \State $parents \gets tail\ parents$
-    \For {$parent \in parents$}
-    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
-    \If {$realParent$ is $root$}
-    \State $root.children \gets parent$
-    \State $parent.parent \gets root$
-    \State $Q \gets parent$
-    \EndIf
-    \EndFor
-
-    \While{$Q$ not empty}
-    \State $cur \gets Q.dequeue$
-    \State $parents \gets cur.parents$
-    \For{$parent \in parents$}
-    \State $realParent \gets min(\forall child \in parent.children,\ parent.depth >= child.depth)$
-    \If {$realParent$ is $cur$}
-    \State $cur.children \gets parent$
-    \State $parent.parent \gets cur$
-    \State $Q \gets parent$
-    \EndIf
-
-    \EndFor
-    \EndWhile
-
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
-
-The primary difference between the original algorithm and the
-generalization is the main data structure. Algorithm~\ref{alg:original}
-is centered around the use of the stack, making it akin to a depth-first
-traversal of the DAG.\@ The generalization in
-Algorithm~\ref{alg:generalized} is based on the queue, making it akin to
-the breadth-first traversal of the DAG.\@ While the final results of the
-two traversals should not be different, the runtime properties are
-different.
-
-We will begin with a simplified description of the original algorithm.
-We start at the master branch pointer. We will use the term master
-branch loosely here, meaning any branch that has been labeled in some
-way as being the main branch, this branch may have a different name, but
-standard naming conventions call it the master branch. Starting from the
-master branch label, we traverse the entire master branch to the initial
-commit, recording the commit hash of every repository event encountered.
-
-The second phase iterates over every commit, computing the shortest
-distance to the master branch. For this reason, the entire repository
-must be available to the algorithm. This is a dynamic programming
-solution, centered around the call stack. The next node on the shortest
-path to the master branch will become the parent of this node in the
-tree.
-
-If this algorithm is to be generalized into a dynamic, web-based, tool,
-it may not be feasible to download every commit in the repository in a
-reasonable period of time. We instead look toward a breadth-first
-approach at limiting the number of repository events that must be
-downloaded.
-
-Focusing on the generalized algorithm~\ref{alg:generalized}, the
-algorithm is broken into two phases. The first phase is where the
-improvements take place. The first phase has three main goals, the first
-is to track the master branch. The second goal labels the depth of each
-repository event that is traverses. The final goal is to record the
-children of each commit, reversing the DAG.\@ The first phase also
-begins the process of reversing the DAG, by providing the candidate
-commits that could possibly be the real parent to this node. The first
-phase can terminate early, if it reaches a point where all branches that
-it is following terminate. The depth metric on line 7 indicates how many
-branches are currently being traversed. In line 11 through 15, we
-decrement the depth, as this is the point where a branch occurs. The
-changing of the depth occurs if there are more events on the same branch
-than between the merge and branch point of a separate branch.
-
-\begin{table}[htpb]
-  \centering
-  \caption{Example traversal of the example DAG in
-    Figure~\ref{fig:repoDAG} starting at merge 11.}
-  \label{tab:ex_traversal}
-  \begin{tabular}{cccc|cc}
-    \toprule
-    Current & Parents & Depth & Q       & Visited & Children\\\midrule
-    -       & -       & 0     & 11      & 11:0    & \\
-    11      & 4,9     & 1     & 4,9     & 4:0     & 11, 5\\
-    4       & 1       & 1     & 9,1     & 9:1     & 11\\
-    9       & 7,6,8   & 3     & 1,7,6,8 & 7:1     & 9, 8\\
-    1       &         & 3     & 7,6,8   & 6:2     & 9\\
-    7       & 5       & 3     & 6,8,5   & 8:3     & 9\\
-    6       & 5       & 2     & 8,5     & 5:1     & 7, 6\\
-    8       & 7       & 1     & 5       & 2:1     & 5, 3\\
-    5       & 4,2,3   & 2     & 2,3     & 3:2     & 5\\
-    2       & 1       & 1     & 3       &         & \\
-    3       & 2       & 0     &         &         & \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-In Table~\ref{tab:ex_traversal}, we follow the algorithm starting at
-merge commit 11 from the example in Figure~\ref{fig:repoDAG}, recording
-the current node, the parents of the current node, the depth after
-updating, and the queue structure after updating. The visited column is
-not dependant on the current step, but keeps track of the nodes that
-have been visited, and the depth at that node. When we reverse the DAG,
-the children in this table will become the candidates for being the
-parent of the corresponding node. For example, the parent of node 7 may
-be resolved to either node 9 or node 8. In this case, the shortest path
-to the master branch is through node 9.
-
-The terminology surrounding the parent/child relationship gets very
-confusing at the change between phase one and two because this is where
-the roles switch. In phase 1, we are working with the DAG; in phase 2,
-we are working with a structure that is neither a tree nor a DAG, but
-can only be described as a directed graph.
-
-The goal of the second phase is to resolve the tree from the directed
-graph. We will refer to parents in the same sense as we would refer to
-them in the tree form of the repository for the remainder of this
-section. We have gathered the depths and possible parents (line 17), but
-we need to remove the extra parents. Extra parents are culled in line
-42 of the algorithm. Of the possible parents, we take the parent that is
-closest to the root, but not beyond. This forces the real parent to
-either be deeper in the tree, or on the same branch. Line 30 to 37 are
-for bootstrapping the queue, ensuring that we are not starting on an
-empty queue.
-
-The construction of the tree becomes trivial now that we are able to
-determine the parents and children for each node. Lines 44 and 45
-perform this construction, adding the DAG parent to the tree children of the
-current node, and the DAG parent's parent becomes the current node,
-which was the DAG parent's child. We continue following the path until
-we hit the root node, merging the commit into the master branch.
-
-This algorithm relies heavily on the order of parents in the parent
-list, but is able to trim the number of nodes that must be downloaded.
-Furthermore, the resulting tree is able to show the order of commits,
-telling the story of how a developer was working, before the set of
-commits was merged. We developed a small bitbucket plugin to test the
-design and determine whether it was capable of producing merge trees,
-shown in Figure~\ref{fig:b_plugin}.
-
-\begin{figure}[htpb]
-  \centering
-  \includegraphics[width=\linewidth]{figures/plugin.png}
-  \caption{Screen shot of the Bitbucket plugin showing the online tree
-    computation of the repository shown in~\ref{fig:repoDAG}.}
-  \label{fig:b_plugin}
-\end{figure}
+\input{discussion.tex}
 
@@
 
-In this paper, we develop a tree-based model designed to provide git
-users with a better explanation of the events occurring in a git
-repository. The \mt model is rooted at the merge into the master branch
-of a repository, where the leaves are the individual commits.
-
-We implement a tool, \tool, which allows a user to filter the commits
-with a simple search interface. Once the user finds the commit that they
-are interested in, they are provided simple aggregated tables showing
-the files that were modified, the authors that made the modifications,
-among other information. They are also provided with three
-visualizations, a list tree, a pack tree, and a Reingold-Tilford tree
-view. The list tree enables easy searching for textual features, while
-the pack tree is good for visualizing the structure of large merge
-trees, and the Reingold-Tilford tree is better at visualizing the
-structure of small merge trees.
-
-To ensure that our model is necessary, we evaluate the ability of users
-to perform simple conceptual tasks, finding how a commit was merged into
-the master branch of the Linux repository, and determining other commits
-that were merged with the given merge. Finding that the participants in
-our study were unable to perform this task, we compared the ability of
-the participants to summarize key pieces of information aggregated in
-these groups of commits. We asked for information about what files were
-modified, the authors making the changes, among other tasks. Finally, we
-asked the participants about their preference in tools, and which
-aspects of each they preferred. We found that the participants in our
-study were able to correctly identify the correct answer more quickly
-using the \mt model than they were able to with the DAG.
-
-Given that the model provides users with a better understanding of
-events in the repository, we describe some issues with identifying the
-master branch of the repository and a more generalized approach to
-constructing merge trees. We implement the generalized approach as a
-bitbucket plugin to verify that it is able to quickly produce results
-online, unlike in the implementation of \tool.
-
-% TODO: You can tell that I am tired. This conclusion will need to be re-worked.
+\input{conclusion.tex}
 

### 2017-07-07 paper/model.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

--- 2016-09-18 linvis.tex
@@
-% vim: set fenc=utf-8 ft=latex encoding=utf-8
-% -*- mode: latex; coding: UTF-8; -*-
+% vim:set et sw=2 ts=4 tw=72:
+Git uses a directed acyclic graph (DAG) as its main data model. In this
+model, a commit has one or more ordered parents (ancestors), except the
+root commits of a repository which do not have ancestors--Linux's git
+repository has two such commits.  Commits are divided into merge
+commits, merging two or more branches, and non-merge commits. Non-merge
+commits have only one parent, while merge commits have two or more. The
+order of the parents matter: the first parent is the branch in which the
+merge is being done, while the rest indicate the branches being merged
+(Git allows merging multiple branches simultaneously). The second parent
+is the first branch merged, the third parent is the second branched
+merged, etc.
 
-\newif\ifdraft
-% \drafttrue
+Once a commit is created, it is never changed. Git allows operations to
+alter commits or reorder them, but it changes the commit ID in the
+process, effectively replacing it with a new commit. This makes commits
+unable to record the traversal of merges from the commit to the merge
+into the master branch.
 
-\ifdraft
-        \documentclass[conference, draftclsnofoot, draft]{IEEEtran}
-        \def\baselinestretch{1}
-        \setlength{\marginparwidth}{2cm}
-\else
-        \documentclass[conference]{IEEEtran}
-\fi
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\newcommand{\TheTitle}{Merge-tree: Visualizing the Integration of Commits into Linux}
-\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
-\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
-\newcommand{\TheSubject}{Digesting large amounts of commit data}
-\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}
-
-\synctex=1
-
-\usepackage[hyphens]{url}
-\urlstyle{same}
-
-\ifdraft
-        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
-                pdfborder={0 0 0},backref=none,colorlinks=true]{hyperref}
-
-\else
-        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
-                pdfborder={0 0 0},backref=none,colorlinks=false]{hyperref}
-\fi
-
-\usepackage[nospace]{cite}
-
-% Table Support
-\usepackage{dcolumn}
-\usepackage{longtable}
-
-\usepackage{balance}
-\usepackage{placeins}
-\usepackage{multirow}
-
-% Extra support
-\usepackage{xspace}
-\usepackage{caption}
-\usepackage[nospace]{cite}
-
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{tikz}
-\usepackage{xcolor}
-\usepackage{color}
-
-% Fix any bad-hyphenations here
-\hyphenation{contained}
-
-% Tikz libraries
-\usetikzlibrary{positioning}
-
-\input{colors.tex}
-\input{macros.tex}
-
-\begin{document}
-
-\title{\TheTitle}
-\author{
-        \IEEEauthorblockA{\TheAuthors}
-        \IEEEauthorblockN{Department of Computer Science,
-                University of Victoria, Canada.}
-        \IEEEauthorblockA{Email: \TheEmails}
-}
-\maketitle
-\begin{abstract}
-
-With an average of more than 900 top-level merges into the Linux kernel per
-release, many containing hundreds of commits and some containing thousands,
-maintenance of older versions of the kernel becomes nearly impossible.
-Various commercial products, such as the Android platform, run older
-versions of the kernel. Due to security, performance, and changing hardware
-needs, maintainers must understand what changes (commits) are added to the
-current version of the kernel since the last time they inspected it in order
-to make the necessary patches.
-
-Current tools provide information about repositories through the directed
-acyclic graph (DAG) of the repository, which is helpful for smaller
-projects. However, with the scale and number of branches in the kernel the
-DAG becomes overwhelming very quickly. Furthermore, the DAG contains every
-ancestor of every commit, while maintainers are more interested in how and
-when a commit arrives to the official Linux repository.
-
-In this paper, we propose the merge-tree, a simplified transformation
-of the DAG of the Linux git repository that shows the way in which commits
-are merged into the master branch of Linux. Using the merge-tree, we
-build \tool, a tool that is designed to allow users to explore how commits
-are merged into the Linux kernel.
-
-\end{abstract}
-
-\begin{IEEEkeywords}
-        \TheKeywords
-\end{IEEEkeywords}
-
-\section{Introduction}
-
-Between 50k and 70k commits are added to the Linux kernel per version requiring
-maintainers of older versions of the kernel to sift through thousands of
-commits and merges with tools that are unable to filter and effectively visualize
-projects at the scale of the kernel. Older versions of the kernel are used in
-embedded systems and mobile phones; for security purposes, performance needs, and
-changing hardware requirements, maintainers must be able to understand the changes
-being made in the current version of the kernel in order to produce the necessary
-patches for the older versions of the kernel. Tools like Gitk use a directed acyclic
-graph (DAG) model of the repository, showing all commits and merges in chronological
-order by when the commit was authored, not by when it arrived in the official Linux
-repository.
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
-        \caption{A view of the Gitk interface centered on merge
-          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
-          their ancestors and there is no clear path from the commit to the merge
-          with the master branch. Neither Gitk nor Git are capable of showing the
-          commits in master.}
-        \label{fig:gitk}
-%\vspace{-4mm}
-\end{figure*}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its
-                relationship to other forks.}
-        \label{fig:gitfail}
-%\vspace{-2mm}
-\end{figure}
-
-This representation works in smaller projects; it enables users to see when changes
-are made, when these changes are merged, how each branch is interacting, and the
-point where a branch forks from the master branch. In large modular projects, like
-the Linux kernel, the DAG becomes a mess of merges and commits
-(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel
-is simply too large for the system to generate a visualization; Github provides a
-DAG view for many projects, but is unable to display the visualization for projects
-at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and
-70k new commits are created for the Linux project every year; according to our
-previous work\cite{German2015}, a commit takes a median of 30 days from the time it
-is authored until it arrives in the official repository. The snapshot of the kernel
-tomorrow may be different than the snapshot from today, containing new commits
-authored in the past; distinguishing these new commits from the commits in the
-snapshot from today is not trivial.
-
-One major challenge with visualizing the arrival of commits to a repository is that
-Git does not store the date that a commit was merged into another branch, including
-the master branch. To complicate the problem, the DAG only has references to the
-ancestors of a commit (a model necessary for the operation of Git), but maintainers
-would prefer knowing the path a commit followed to reach the master repository.
-Tracing a path that any commit followed to the master repository would imply that
-for any given merge, it would be possible to know which commits were merged. A user
-could inspect the commits that arrived into the master branch within a given
-time-frame by checking which commits were merged during that time-frame.
-
-This paper makes two contributions; first, we describe a method of converting the
-DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
-that represents the path used by a commit to reach the master branch; second, we
-present a method to inspect and visualize the history of merges in the Linux project
-using the merge-tree model.
-
-These methods and visualizations are implemented in a web-based tool called
-\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
-Our visualizations and tool provide information about the location of any given
-commit or merge in its respective merge-tree, the files edited, the modules edited,
-and the commit message. \tool allows users to apply various filters, including the
-release version, along with a keyword or phrase from the log preview, the name of
-the author, or the commit ID. The user can request all merges made by Linus that
-contain a commit or inner merge that matches the search query, or just the commits
-and merges that match the query.
-
-\section{Merge-Tree model}
-\label{sec:mergetree}
-
-Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
-commit has one or more ordered parents (ancestors), except the root commits of a
-repository which do not have ancestors--Linux's git repository has two such commits.
-Commits are divided into merge commits, merging two or more branches, and non-merge
-commits. Non-merge commits have only one parent, while merge commits have two or
-more. The order of the parents matter: the first parent is the branch in which the
-merge is being done, while the rest indicate the branches being merged (Git allows
-merging multiple branches simultaneously). The second parent is the first branch
-merged, the third parent is the second branched merged, etc.
-
-Once a commit is created, it is never changed. Git allows operations to alter
-commits or reorder them, but it changes the commit ID in the process, effectively
-replacing it with a new commit. This makes commits unable to record the traversal of merges
-from the commit to the merge into the master branch.
-
-A short example: assume the commits represented in Figure~\ref{fig:repoEvents} show
-the sequence of events in a repository. In this case, commits are performed in
-various repositories and branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about the master
-branch and the repository that the master branch is part of. The merge-tree view of
-this DAG is visible in Figure~\ref{fig:repoTree}. Note that the direction of the
-edges of the DAG have been inverted, instead of pointing from the child to the
-ancestors, it points from the ancestor to its successors, forming a path to the
-master branch. Also note that the DAG has been simplified, showing only a single
-edge on the path to master for any commit.
+A short example: assume the commits represented in
+Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
+In this case, commits are performed in various repositories and
+branches. The DAG representation of the commits is shown in
+Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
+the master branch and the repository that the master branch is part of.
+The merge-tree view of this DAG is visible in Figure~\ref{fig:repoTree}.
+Note that the direction of the edges of the DAG have been inverted,
+instead of pointing from the child to the ancestors, it points from the
+ancestor to its successors, forming a path to the master branch. Also
+note that the DAG has been simplified, showing only a single edge on the
+path to master for any commit.
 
@@
 
-Computing the merge-tree from a DAG for any repository may not be possible; however,
-certain features of the development process of Linux make it feasible to compute the
-merge-tree for the Linux repository. First, the master branch of Linux is maintained
-by Linus Torvalds, and only Linus has write access to it. We have verified this
-assertion in previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first identifies
-the commits made directly to the master branch; whereafter it recursively determines
-the shortest path (in terms of time), using the DAG, from each commit to the master
-branch using the inverted DAG.
+Computing the merge-tree from a DAG for any repository may not be
+possible; however, certain features of the development process of Linux
+make it feasible to compute the merge-tree for the Linux repository.
+First, the master branch of Linux is maintained by Linus Torvalds, and
+only Linus has write access to it. We have verified this assertion in
+previous research~\cite{German2015}. We have developed a heuristic that
+is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
+identifies the commits made directly to the master branch; whereafter it
+recursively determines the shortest path (in terms of time), using the
+DAG, from each commit to the master branch using the inverted DAG.
 
@@
 
-Merges that do not have conflicts provide information to verify this heuristic. If a
-merge does not contain a conflict, it records a summary of the commits that it
-merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a
-list of the first 20 non-merge commits in the merge, including their one-line log
-description, the full logs of the merge commits that merge this subset, and the
-total number of non-merge commits in the merge.
+Merges that do not have conflicts provide information to verify this
+heuristic. If a merge does not contain a conflict, it records a summary
+of the commits that it merges. See Figure~\ref{fig:sampleMerge} for an
+example. This summary contains a list of the first 20 non-merge commits
+in the merge, including their one-line log description, the full logs of
+the merge commits that merge this subset, and the total number of
+non-merge commits in the merge.
 
@@
 
-We used this information to evaluate the accuracy of the merge-tree model extracted from the DAG. The method we followed
-started with the extraction of the commit history up to July 20, 2016. We computed the merge tree of every commit until
-then. Since Linus Torvalds mostly does merging directly into master, we assumed that every merge by him is the root of a
-merge-tree. As described above, the log of a merge-commit usually contains the number of commits in the merge the first 20
-summaries of commits being merged. We extracted merges by Linus Torvalds using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits according to the log with the number of commits in the  merge-tree rooted in this commit. We
-also used the summaries of the commits found in the merge (not necessarily all---see above) to make sure those
-commits were in their corresponding merge tree. For example, for the merge in
-Figure~\ref{fig:sampleMerge} we would expect
-that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the one-line summaries corresponds to commits in
-that merge-tree. We also inspected those with differences to make sure they were true errors.
+We used this information to evaluate the accuracy of the merge-tree
+model extracted from the DAG. The method we followed started with the
+extraction of the commit history up to July 20, 2016. We computed the
+merge tree of every commit until then. Since Linus Torvalds mostly does
+merging directly into master, we assumed that every merge by him is the
+root of a merge-tree. As described above, the log of a merge-commit
+usually contains the number of commits in the merge the first 20
+summaries of commits being merged. We extracted merges by Linus Torvalds
+using the command \mycode{log --merges --author='Torvalds} and compared
+the number of commits according to the log with the number of commits in
+the  merge-tree rooted in this commit. We also used the summaries of the
+commits found in the merge (not necessarily all---see above) to make
+sure those commits were in their corresponding merge tree. For example,
+for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
+merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
+one-line summaries corresponds to commits in that merge-tree. We also
+inspected those with differences to make sure they were true errors.
 The results can be summarized as follows:
@@
 \item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
-  branch (not master). This branch was merged at (\mycode{3f17ea6d..} which contained 6809 commits.
+  branch (not master). This branch was merged at \mycode{3f17ea6d..} which contained 6809 commits.
 \item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
@@
 
-In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860) our heuristic was correct. It failed in 1,542
-commits before Dec. 12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
-
-\section{Visualizing the merge-tree of Linux}
-
-The goal of \tool is to simplify the navigation of the kernel commit information,
-specifically focusing on merges. This is done by leveraging the merge-tree to
-inspect how commits are merged on the path to the master repository.
-
-\subsection{Use cases}
-
-We designed \tool with two use-cases in mind, though a user may switch between the
-cases as they work.
-
-\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\ These
-are users that are maintaining a section of the kernel and would like to pick a
-merge (including all the commits that it merges) and merge it directly into their
-current repository. This is useful for reducing the amount of re-implementation
-work. For these users, it is important to have the ability to aggregate metadata
-about files and modules being effected by the merge. Also, it is important for these
-users to be able to navigate from the root of the merge-tree toward the leaves.
-
-\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
-These are users that start with a known merge or commit and would like to see what
-other changes are being made in commits that are in the same merge, including
-knowing the merge-tree they belong to. This is useful to see what other commits are
-related to the current commit and how they get collated into merges that eventually
-end in the master branch. This is primarily for maintainers that need to perform
-some specific cherry picking of commits. We must provide these users a mechanism for
-navigating from a single commit toward the master branch, allowing them to see other
-commits that might be related to their original commit.
-
-\subsection{Data Model}
-
-In our visualizations, we leverage the merge-tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already in the
-master branch, or is part of a tree which is rooted in the merge that merged it into
-the master branch.  Each commit, whether a merge or non-merge, has only one
-successor; the root of each tree has none as it was made by Linus Torvalds directly
-into the master branch. Non-merge commits contain the metadata for the changes made.
-This metadata includes the files changed, the lines added and removed from each
-file, the author, the date the commit was merged into the merge that led to being
-merged into the kernel, the date the commit was authored, the patch, and the commit
-log. Merges contain less metadata, only storing the author of the merge, the log,
-the commit date, and the author date, and potentially, changes necessary to address
-conflicts during the merge. The details of the model are outlined in
-\cite{German2015}.
-
-\section{Design and Implementation}
-
-To navigate and inspect the merge-tree view of the kernel we created a web-based
-tool called \tool. Creating a web-based tool enables users to use the system without
-having to install additional software or store a large database, making it more
-accessible, more easily maintainable, and platform independent. \tool uses the
-following mechanisms to reach our goals of better navigation and better explanation
-of the selected changes.
-
-\begin{itemize}
-        \item Filter by searching
-        \item View files edited
-        \item View modules
-        \item Tree viewer
-\end{itemize}
-
-\subsection{Searching}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View allows filtering commits that were merged in a given
-                period, filtering by author, keyword, or commit ID.}
-        \label{fig:search}
-\end{figure}
-
-Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
-merges that are irrelevant. The search mechanism breaks down the results by release
-version. A user can further narrow down the search by specifying a range of dates in
-which such commits were merged by Linus into the master branch---not when the commits
-were created (author date and/or commit date). This distinction is important. We have observed commits that have
-taken years to arrive into the master repository after they were originally
-created.
-
-A user may then provide a search text, filtering by the author name, the commit ID,
-or keyword from the log. Any part of the author name may show up in the results,
-including searching by email address.
-
-If the user is searching by commit ID, the ID can be specified by using any of its unique prefixes. For example, the commit
-\mycode{3f17ea6dea8ba5668873afa54628a91aaa3fb1c0} is returned when the user searches for a
-commit ID of \mycode{3f17e} in the 3.16 Linux kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
-        \caption{Search Results. Each table entry is a commit merged in the desired
-                merged window.}
-        \label{fig:results}
-\end{figure}
-
-In the search results (seen in Figure~\ref{fig:results}), the user is presented with the one-line log message preview,
-the author's name and email, the date the commit was authored, and the date the commit was last committed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
-        \caption{Panel showing the Commit Message of a commit.}
-        \label{fig:message}
-\end{figure}
-
-Once a user selects a commit or merge to investigate, they are presented with a
-tabbed pane allowing them to view the full commit log, the files edited, the modules
-involved, and the merge-tree view.
-
-The first tab displays the full commit log (Figure~\ref{fig:message}). From this, a
-user is able to see what they would see had they searched for the commit using Git
-log. This doesn't provide additional information to the other tools, but helps to
-complete the functionality of \tool. The commit log provides a user with the
-information about the content of the commit and who has signed-off on the commit to
-ensure that it is of good quality. The message for merges may contain a summary of
-the commits being merged.  The information within these messages is highly variable,
-and is completely dependent on the author's style. As the user moves toward the
-root-level merge, the quality of these messages generally improves.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
-        \caption{Panel showing the files modified by all the commits that are part
-                of this merge.}
-        \label{fig:files}
-\end{figure}
-
-The second tab is the files tab (Figure~\ref{fig:files}). This tab provides
-information on what files have been edited, how many lines were added, and how many
-lines were removed in a given commit. For non-merges, this functionality is similar
-to the other tools available. Our tree-based design model allows us to extend this
-functionality to merges by aggregating information about all the commits that are
-children of the merge in the merge-tree, which other tools are unable to show. To
-find the number of lines added to a file in a merge, we take the sum of the lines
-added to that file in each of the children of that merge. We do the same for
-calculating the number of lines removed.
-
-% This could be extended to use the patch information to determine when a line has been
-% replaced rather than incrementing the lines added and removed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
-        \caption{Panel showing the modules changed by all the commits in this merge.}
-        \label{fig:modules}
-\end{figure}
-
-The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
-within the commit. Modules are not natively recognized by Git, and are not going to
-be present in all repositories. In the Linux repository, authors put the module they
-are working on in the one-line summary of the log-message; for example: the log
-\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
-coverage testing tool of the kernel). We heuristically extract the module by taking
-all text in the log summary of commits until the first colon.  Modules are logical
-partitions of the information in the kernel. Depending on where the author was
-working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
-quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
-few cases, the author of a commit does not correctly follow this format and the
-heuristic approach fails. As with the Files panel, non-merge commits show their
-corresponding information, but for merge commits we aggregate all the modules
-changed in all the commits that are part of that merge.  The output of this view is
-shown in Figure~\ref{fig:modules}.
-
-Finally, we have the Tree view tab. The tree view is designed for providing easy
-navigation of the commits within the merge-tree that is rooted in the current merge.
-It also provides a clear topological view of the merge and the submerges it
-includes. We have experimented with various tree designs to find a design that
-allows for easy navigation and visualization of both large and small trees. We
-discuss this panel in the next subsection.
-
-\subsection{Merge-Tree Views} \label{treeview_section}
-
-The merge-tree view is what makes \tool unique to other tools that inspect the DAG
-of a Git repository. With it, a user can inspect how commits are merged on their way
-to the master branch. We have experimented with various types of trees:
-
-\begin{enumerate}
-        \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
-        \item Reingold-Tilford trees provide a clear visual representation of the tree structure of the merge-tree.
-        \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
-                clearly showing the parent-child relationships between commits and merges.
-\end{enumerate}
-
-\subsubsection{List Tree}
-
-The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of nested lists,
-and is designed to more closely model the tree view found in file browsers. This
-tree only contains the commits and merges that are within the subtrees of the
-current merge. A commit will never have any items in this tree as it is a leaf. To
-accompany the tree, we include breadcrumbs at the top of the page to enable a user
-to navigate both from the root to the leaves and from a leaf to the root. The last
-item in the breadcrumb list is the current commit, the previous item is the parent
-of the current commit, and the first item is the root merge into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List tree view of merge 3f17ea6}
-        \label{fig:list_tree}
-\end{figure}
-
-\subsubsection{Reingold-Tilford Tree}
-
-The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:reingold_tree} and
-\ref{fig:reingold_tree_zoom}) allows the visualization and navigation of the entire merge-tree in
-an intuitive representation of the tree.  This illustrates a clear notion of root
-and leaves, and how to navigate in either direction. Some merge trees are very
-large, containing thousands of commits and merged. While the tree is capable of
-producing a visualization, it becomes far more difficult to understand. For example,
-the merge, 3f17ea6 performed by Linus Torvalds June 8 2014, contains 7217 commits
-and merges. \footnote{This commit can be inspected at
-  \url{http://li.turingmachine/org/commits/3f17ea6dea8ba5668873afa54628a91aaa3fb1c0}}
-
-      \begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/4dc4_tree.pdf}
-        \caption{Merge 4dc4226 is a subtree of 3f17ea6, for the power-management
-          module of the kernel}
-        \label{fig:reingold_tree}
-      \end{figure}
-
-      \begin{figure*}
-        \centering
-        \includegraphics[width=0.98\textwidth]{figures/4dc4_zoom_tree.pdf}
-        \caption{Zoomed out view of the Reingold-Tilford tree for merge 4dc4226}
-        \label{fig:reingold_tree_zoom}
-      \end{figure*}
-
-      The user is initially greeted with their current node centered on the screen.
-      They are able to zoom the tree by scrolling the mouse, and clicking and
-      dragging to pan the tree. They can see more details about a node by clicking
-      on it, which will provide them a link to the specific page for that commit.
-
-\subsubsection{Bubble Tree}
-
-Bubble trees are useful for providing a clear visualization of wide, hierarchical
-data\cite{Boardman2000}. The tree structure is represented by the nesting of nodes;
-the largest circle is the root, containing all the other nodes. The smallest circles
-do not contain any nodes and therefore represent the leaf nodes.
-
-Our implementation of the bubble tree (Figure~\ref{fig:bubble_tree}) provides the
-user with a clear picture of where a commit is located in the merge tree. We
-highlight the selected commit or merge in red. Non-selected merges are in a shade of
-blue determined by the depth of a node in the tree. The root is the lightest shade
-of blue, while the contained merges are progressively darker; the commits are white.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree of merge 3f17ea6, merge 4dc4226 is the currently
-          selected merge and is highlighted in orange.}
-        \label{fig:bubble_tree}
-\end{figure}
-
-The bubble tree doesn't have an implicit way of providing additional information,
-placing any text near the nodes makes the tree impossible to read, so we include a
-separate pane in the web page. When a user hovers over a node, the pane shows
-additional information about the author and the commit message and a link to the
-detailed page for that commit or merge. If the user clicks on a node, the tree will
-zoom to that node and the information in the info pane will persist, enabling the
-user to click the link. This tree provides an easy mechanism for users to navigate
-from the root to the leaves and vice versa.
-
-\section{Implementation Details}
-
-\evantodo{See if we can move API section to be part of another section}
-
-The front-end uses asynchronous requests to gather the information for the trees and
-tables. This enables third-parties to implement new front-ends, trying other
-designs, though this interface may change in the future. The tree, file, and module
-information is accessible through
-http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
-http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.
-
-The response for the tree is a single object. This object is the root node of the
-tree, and may contain an object of children sub-tree objects in the \verb|children|
-field, which are in the same format as the root object. The children object uses the
-commit ID as the key.
-
-The tree node objects have the following fields:
-
-\begin{tabular}{ccl}
-        Field name      & Data type & Description\\\hline
-        \verb|cid|      & string    & Commit ID\\
-        \verb|name|     & string    & One-line log summary\\
-        \verb|mlinus|   & string    & Root merge commit ID\\
-        \verb|author|   & string    & Author name and email\\
-        \verb|mnext|    & string    & Parent merge commit ID\\
-        \verb|children| & object    & Object of tree objects\\
-\end{tabular}
-
-The file responses contain only the files that the selected merge or commit works
-with. The response is a single object in the form of a tree. This object is the root
-of the tree, and represents the current merge or commit. If the current position in
-the tree is an inner node, the response will contain the child nodes in the
-\verb|children| field, otherwise \verb|children| will be an empty list. The format of file
-response is as follows,
-
-\begin{tabular}{lll}
-        Field name & Data type & Description\\\hline
-        \verb|cid| & string & Commit ID\\
-        \verb|mnext| & string & Parent merge commit ID\\
-        \verb|children| & list & List of tree objects\\
-        \verb|files| & list of tuples&
-        \footnotesize{
-                \begin{tabular}{lc}
-                        \verb|Filename| & string\\
-                        \verb|Lines added| & uint\\
-                        \verb|Lines removed| & uint\\
-                \end{tabular}}\\
-\end{tabular}
-
-\section{Discussion}
-
-From a developers point of view, the are two major disadvantages of the DAG model of
-Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
-commit that succeeds them on the path to being integrated; and b) the DAG contains
-many more edges that are necessary to understand how integration occurred. We
-addressed these two issues with the creation of the merge-tree model from the DAG of
-Linux. Effectively, the merge-tree recovers the details of how each branch---and the commits they contain---was merged
-into the master branch.
-
-In our experience, no other Git repository reaches the level of DAG complexity that
-Linux has. In most Git repositories, most merges are not nested---most merges merge
-directly into the master repository, and these merges only merge a few commits. Even
-in these simplified cases, the merge-tree can provide a valuable summary of how
-commits are integrated into the master branch, specifically since the time of
-integration may be very different from the time the commits were authored.
-
-\evan{Check accuracy of this paragraph}
-
-The biggest challenge is computing the merge-tree of any given repository. It
-is likely that our heuristic for computing the merge-tree will not work with other
-repositories. The heuristic requires that there are no fast-forward merges into the
-master branch of the repository (a majority of repositories), and that there are no
-foxtrot merges (a practice that is starting to be considered desirable by git users\footnote{See
-  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git} and
-  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
-We are able to use our heuristic with the Linux kernel because of
-the strict integration model imposed by Linus. To validate the results of the
-heuristic, we use the continuous mining technique described in \cite{German2015}.
-Continuous mining of the repository requires foresight and planning.
-
-Because \tool leverages the merge-tree model of the kernel, it provides mechanisms
-and visualizations that other tools are unable to produce. It gives us the ability to
-see how commits are being merged into the master branch, allowing us to better understand
-how code is integrated into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
-        \caption{Merge Dag View}
-        \label{fig:dag_view}
-\end{figure}
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
-        \caption{Merge tree topology for the same merge as in
-                Figure~\ref{fig:dag_view}.}
-        \label{fig:tree_view}
-\end{figure}
-
-Our goal was to build a tool to enable maintainers to effectively navigate and
-browse the changes performed to the kernel over the period of a release. Achieving
-this goal includes removing information that does not pertain to the area of the
-kernel that the user is interested in. In Figures~\ref{fig:dag_view} and
-\ref{fig:tree_view} we can visually compare the results returned from Gitk
-and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
-from kernel version 3.10. Note that Gitk has no way to show or recognize that this
-is a merge into the master branch.
-
-The log information to this merge-commit in both of these figures is identical, but
-the presentation drastically changes our ability to comprehend what we are seeing.
-The primary difference is the removal of irrelevant information. In Gitk
-(Figure~\ref{fig:dag_view}), the visualization results include commits and merges
-from other components of the kernel, while our tool (Figure~\ref{fig:tree_view})
-only includes the results that are specific to the component of the kernel that we
-are interested in.
-
-Even though this merge is relatively small, containing only four sub-merges, three of which
-only contain a single commit, and the last containing two commits, it is already hard to visualize in Gitk.
-With \tool we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
-furthermore, users must work to determine where this merge ends and the next one
-begins.
-
-\tool further enhances our understanding by summarizing the files
-and modules that were edited in the entire branch being merged. We are able to determine that three
-files, ``palamas-regulator.c'' had two lines added and two lines removed,
-``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
-lines added and 2 lines removed. Finally, we are able to determine that only the
-``regulator'' module was modified in this merge and was modified by 5 commits.
-
-\section{Future Work}
-
-There are still many areas that can be improved before the full potential of our
-model can be realized. Here, we outline various areas of the tool that still need
-more attention.
-
-\subsection{Files}
-
-There currently is no functionality surrounding searching by filename. Users falling
-under use-case 2 may know what file they are editing and try to determine how it may
-work with the other commits and merges in the module. It is possible that a third
-use-case may arise, where the user wants to determine all commits that effect a
-given file. In both cases, bottom-to-top approach is applied.
-
-There is limited functionality to the presentation of the file information. At a
-minimum, the patches for the commits can be displayed. From there, the patches can
-be used to piece together parts of the file to generate a single patch at a given
-merge rather than displaying each patch individually. The patches can also be used
-for determining what kinds of changes were made, if the lines are being added,
-removed, or being replaced.
-
-\subsection{Authorship}
-
-Our model can aggregate more information than what we have implemented. The
-authorship information is important for licensing purposes, and we can show all
-authors contributing to commits in a  merge and how many commits they contributed to
-the merge. We could go further, providing information about what files they edited,
-how many lines they added and removed, and more.
-
-\subsection{Evaluation}
-
-At this time, we have no evidence that our tool is able to improve the work-flow of
-maintainers. We believe that the tool is able to improve the work-flow and
-performance of maintainers because it provides cleaner mechanisms of visualizing and
-presenting the commit information. It is able to provide more relevant information,
-while removing information that is irrelevant to a given module or set of merges.
-
-We can either perform user-testing to show how a users workflow changes, or we could
-have maintainers evaluate and critique the tool and use their feedback to determine
-if the tool meets our goals.
-
-\section{Related Work}
-
-Version control systems monitor the development lifetime of software projects. This
-makes the version control system vital in providing information about how a software
-project is being developed. To our knowledge, we know of no Git repository
-visualization tool that builds a tree that maps the path a commit follows to the
-master branch from the DAG provided by Git. This may be because more information is
-required to generate the merge-tree model than what is stored by Git. However, there
-has been a lot of work in providing visualizations of various repositories.
-
-Many tools work to address the issues in communication between developers in
-inter-team collaboration work. Codebook \cite{Begel2010} uses a data mining
-technique  to determine the developer of a piece of code, the program manager who
-wrote the specification for the code, and the program managers and developers on the
-team who were working together. Hipikat\cite{Cubranic2005} is another tool with a
-focus on communication. Where Codebook focuses on developers working on a project,
-Hipikat is focused on enabling easier integration of new developers to a project by
-providing them with easily-searchable artifacts of the changes made. Codebook is
-useful for pairing a contributor with the original developer; however, the developer
-may not have worked with the piece of code in years. A Hipikat program may provide
-more information to the maintainer as it records the artifacts of why certain
-design decisions were made when they were made using other tools like Bugzilla and
-CVS. Neither tool is sufficient in meeting our goals to provide a summary of the
-topology of the kernel repository through a visual tree.
-
-Most visualizers provide a visual presentation of a certain aspect of a repository.
-Fractal Figures \cite{Ambros2005} uses a unit square to represent a portion of a
-project, then partitions the square based on the proportion of an author's
-contributions to that portion of the project. EPOSee\cite{Burch2005} and Evolution
-Radar\cite{Ambros2009} perform further analysis, determine which files are made
-together, and what changes are made over a sequence of commits, though the goals
-behind these projects is different.
-
-Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work with data
-from CVS repositories. Our goal is to provide information about Git repositories,
-specifically the Linux master repository. Fewer tools are available for generating
-visualizations and summaries of Git repositories, potentially due to the DAG model
-used by Git.
-
-Gource is a tool for providing an interactive timelapse of the state of a
-repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
-type of contribution a developer is making. These contribution types are one of,
-adding a file, removing a file, and changing a file. While the timelapse is
-interesting to watch, it does not provide any additional explanation of the changes
-actually being made, only the frequency that they are being made and who is making
-them.
-
-The current industry standard tool for Git repository visualization is Gitk. The
-Gitk interface is built around the central DAG viewer. The DAG view displays the
-commits on their respective branches, the author of the commit or merge, and the
-date the commit or merge was authored. A user can select a single merge or commit
-to view additional information. For both merges and commits, Gitk will present the
-user with the Git log. If the user selected a commit, Gitk will also present the
-user with the patch information and the names of the files edited. Gitk is unable to
-provide the patches and filenames information in merges, as it is unable to
-aggregate commit information in merges.
-
-Our tool is primarily aimed at presenting the hierarchical structure of the Linux
-git repository. We use tables for presenting the summarized information of the
-commits and merges, but this information could also be presented in a graphical
-form. Various graphical forms for displaying file and authorship data exist, the
-principal forms being matrix views, city scapes, bar and pie charts, and networks
-\cite{Eick2002}. Any of these data visualization metaphors are applicable to our
-system.
-
-\section{Conclusion}
-
-Our tool, \tool, shows promising results in working toward our goals, building a tool that
-can more easily navigate the repository, and provide clearer explanations of the
-changes made. The filtering by searches and the various tree visualizations enable
-users to easily navigate through the kernel, finding the commits and merges that
-pertain to what they are interested in. The tree visualizations are able to both
-serve as a navigation piece, providing simpler navigation through the kernel, but
-also demonstrate which commits contribute to which merges, providing further
-explanation over the DAG. The tree model further improves the explanation by enabling
-the tool to aggregate metadata at the merges instead of requiring a user to manually
-aggregate the information. Our tool currently aggregates file information and the
-modules edited, but can be extended to include authorship information and aggregated
-patch information.
-
-\evan{Check the validity of this paragraph}
-We use a heuristic method to generate the tree from the repository. The heuristic
-requires that the repository does not have fast-forward merges into the master
-branch or foxtrot merges, which obfuscate the master branch. In either case, the
-heuristic will break and result in a tree that does not accurately represent the
-repository.
-
-These may pose as significant drawbacks to smaller projects; however, with a proper
-merge discipline, the heuristic is able to correctly convert the DAG into a
-merge-tree. The importance of clear visualizations in large projects outweighs
-the costs of enforcing a clean merge discipline. The tree-based model appears to
-have the ability to provide a clean visualization and more detailed explanations of
-the changes in the Linux kernel.
-
-% \nocite{*}
-
-\bibliographystyle{IEEEtran}
-\bibliography{citations}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:
+In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860)
+our heuristic was correct. It failed in 1,542 commits before Dec. 12,
+2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).

### 2017-07-07 paper/study.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

\dmg{start by describing what you mean by user study... to evaluate linvis we performed a user study...}

\dmg{and you are evaluating the visualizations, NOT the DAG}

We use a controlled mixed-methods users study in order to evaluate the
ability for the DAG to provide users with a conceptual image of how a
commit is integrated into a project, to compare the DAG and \mt models,
and determine which model the users preferred.

In this section, we describe the methods used to for designing and
implementing the user study.
\dmg{no, you describe the study, which includes the methods}

\dmg{Start by giving an overall overview of the study: the study consists of a set of tasks that users are expected to
  perform... do a top down description of what users will do
}

\subsection{Questions and Tasks}
\label{sub:questions}

\dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}

\begin{table*}[htpb]
  \centering
  \caption{Conceptual Tasks }
  \label{tab:conceptual_tasks}
  \begin{tabulary}{0.9\textwidth}{LL}
    \toprule
    Task & Description\\
    \midrule
    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
    T2 & How many individual commits are related to this commit?\\
    T3 & How many merges are involved with merging this commit into the master branch?\\
    \bottomrule
  \end{tabulary}
\end{table*}

\begin{table}[htpb]
  \centering
  \caption{Summarization Tasks}
  \label{tab:summarization_tasks}
  \begin{tabulary}{\linewidth}{LLL}
    \toprule
    Task Set   & Task & Description\\\midrule
    Merge      & T4   & What is the series of merges involved with merging this
    commit?\\
               & T5   & What other commits are merged?\\
    Authorship & T6   & How many authors are involved?\\
               & T7   & Who contributed the most changes?\\
    Files      & T8   & How many files were modified?\\
               & T9   & Which file had the most changes?\\
    Modules    & T10  & Which modules does this merge tree involve?\\
    \bottomrule
  \end{tabulary}
\end{table}

\dmg{avoid editorializing your text: remove purely}

\dmg{watch for tense... we don't begin... the tasks begin... or better, the first tasks...}

The questions and tasks are broken down into four groups. The first two
groups are purely quantitative in nature; the study begins with
conceptual tasks to understand whether the DAG is capable of providing
users with an understanding of how a commit is merged into the master
branch, and other commits that it is merged with. We continue with
summarization tasks, to quantitatively understand if the \mt model makes
a difference in summarization tasks.

The third task set is purely qualitative, \dmg{and are their goal is ... don't imply content in your sentences... be
  precise; this applies to the paragraph before} to gain better understanding
of whether participants preferred working with the \mt model in \tool,
or working with the DAG in Gitk, and what aspects they preferred from
each tool. The final group of questions is simply to gain a better
understanding of the demographic of our participants.

\textbf{Conceptual Tasks}

\dmg{given the choice, not allowed.  I don't understand this section.
  You need to be more top-down in your description. you need to provide
  context on what the users are expected to do, and why. So what are the
  tasks? You never explain.  what is the goal of the tasks? how many?
  what were the tasks? what is expected from each? you need to answer
  these questions, not necessarily in the same order.  also, remember
  this is a comparative study, so I presume the users were asked to do
  both with linvis and without it (choice of gitk/git command line) }

The participants were allowed to use Gitk or the git command line tools
to perform these tasks. These tasks are asked in the specified order,
allowing tabularr to first, create a conceptual understanding of the
commits and merges that will be involved with the rest of the tasks that
follow.tabularmay then use their diagrams to answer T2 and T3. We do not
collect metrics from \tool for the conceptual tasks, opting instead to
determine if the DAG is sufficient for understanding how a commit is
integrated. \tool uses the \mt model show a tree visualization of how
commits are merged, until the commit reaches the master branch.
Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.

\dmg{I don't like the present tense, but that is your option}

We provide the participant with 10 minutes per commit to complete the
first task. The answers to the two other questions in this section are
drawn from the answer in the first task.

\textbf{Summarization Tasks}

\dmg{again, you are going into details first. First the tasks --see
  above-- including their grouping.  then at the end, describe the
  randomization and why it is done (which you did) }

The summarization tasks (Table~\ref{tab:summarization_tasks}) are
grouped into \dmg{how many?} set of tasks according to their theme. The
tasks within a given task set are shuffled, and the task sets are
shuffled, to ensure that similar questions are together, but not in the
same order, and to ensure that the themes are in randomized order
between participants. The order of the commits is shuffled, as well. The
order of the tool is randomized, choosing to start with either \tool or
Gitk, to ensure that we do not place a bias on one tool through the
experiment.

As understanding where the merge begins and ends is vital to answering
the other question, the merge task set will always be the first task set
presented; however, the questions will still be re-ordered within the
task set.

\textbf{User Opinion and Profile}

Finally, we ask the user for their opinions on the tools and for some
information about their level of experience with git.  \dmg{be more
  precise: understanding or preference is too broad. What you want to
  know answers to specific questions regarding how the tools compare and
  which ones they would prefer} The first three questions are intended
to provide us with a better understanding of preference, and insight on
why, the participant had that preference.  The last three questions in
this section are aimed at getting a better understanding of the
demographic of our participants.

\begin{enumerate}
  \item Given these tasks again, which tool would you prefer to use?
  \item Which aspects of each tool did you like and why?
  \item How long have you used git?
  \item If you have used git, for what kind of projects? (personal,
    school courses, professional?)
  \item If you have used git, how many commits, files, and contributors
    were involved with the largest repository you have worked with?
\end{enumerate}

\textbf{Procedure}

\dmg{tell me first why you need to create a script, so I know what you
  do something: we did random assignment of tasks (explain what you did)
  and randomized the order in which the tasks were performed.  Then
  explain why you need a script}

\dmg{it does not randomize the order the tools. It randomizes which tool
  is going to be used on each tasks.  be precise with your language}

\dmg{what you explained how commits relate to tasks and how they get
  assigned? think as if you are explaining this to somebody who has to
  do this: you are the designer, explaining the why and the how }

For each participant, we run a simple python script to generate the
script for the study, eliminating the possibility of bias from us. The
script randomizes the order of the tools, the order of the commits, the
order of the task sets, and the order of the tasks within each task set,
where appropriate.

\subsection{Commit Selection}
\label{sub:commit_selection}

\dmg{see my comments in previous section. it is ok to explain them here,
  but you need to explain what they are briefly in the previous section}

We chose two commits for use in the study. The order that the commits
are presented is randomized between participants, but the order is kept
consistent through the tasks described in the next subsection.

\dmg{this place does not require a description of the database. This
  requires just to say the commits/merge set (dates and numbers are ok)
  only} We use the \tool database, containing commits and merge-tree
information from April 16th 2005 to October 14th, 2014, which
corresponds to being between Linux release 2.6.12-rc3 and Linux
3.17-rc1. \evan{Cite the database even more?} \dmg{again, i have
  problems with your writing: ``we wanted to use in the study commits of
  different size, to determine if the size had an impact in the
  completion of the tasks. We chose commits of three different sizes...}
We chose the commits based on tree sizes. We \dmg{we didn't find: this
  is a fact, not a discovery; what is a majority? 50\%+1? be precise}
\dmg{drop the word inner, and do it in order: 25\% of trees have only
  one commit, 50\%.., 75\%...} found that a majority of the trees
contain at most seven inner commits, while more than 25\% of the trees
contain a single commit. 75\% of the tree contained up to 51 inner
commits and merges, and finally the largest tree contained 7217 nodes.
We chose to work with the trees in \dmg{if you are going to use the
  words quartile, just use them before in your description eg: the
  distribution of the sizes of the trees is x, y merges, whose size
  corresponded to the 25 ,50 quartiles, simple and precisee} the first
and second quartile, as merge trees of sizes between one and seven, not
including the merge into the master branch, make up the majority of the
trees in the database.

\dmg{remove from here, what is here? this is not a road trip :)} From
here, we selected one tree from the trees of a single commit at random.
Selecting a commit from that tree is trivial, as there is only one to
choose.\dmg{remove the obvious, instead concentrate on why the others
  are not trivial}

We selected one tree of size seven at random from all trees of size
seven\dmg{how many where there?}. We placed a restriction the trees, the
selected tree must include at least one internal merge to increase the
complexity of the trees tested. For each randomly selecting a tree, we
chose one of its internal commits at random (using the
\verb|random.choice| function in python 3.6.1).

\dmg{question: so you chose only one commit of each size, and the
  randomization was which tool to use for each question? or did they
  have to do the same task for two commits?}

Using this technique, we selected commit
\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
a single node (Figure~\ref{fig:commit_1}), which we will refer to as
\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
will refer to as \comB.

\begin{figure}[bpt]
  \centering
\begin{tabular}{ m{1.5cm} m{3cm} }
  \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
  \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
\end{tabular}
  \caption{The merge trees used in the user study, 
    containing one and seven commits, respectively.}
  \label{fig:commit_1}
\end{figure}

\subsection{Results}
\label{sec:results}

\dmg{this is still methodology.}

We recorded the audio and screen capture video for each participant in
the study for further analysis.
\dmg{you don't need this sentence, it is redundant}
After capturing the information, we
extracted various metrics from the videos.
\dmg{ you don't record, you measure}
From the conceptual tasks, we
recorded the time in seconds, and the answer for the question. We
extracted three metrics for the summarization tasks, the time in
seconds, the correctness, and the accuracy.
\dmg{and then somethings you are over precise :): correctness is simply whether their answer to the task was correct or not}
Correctness is a binary
value, it is true if the response was correct and false if the response
was incorrect.
\dmg{if you are going to use this metric, you need to define precisely. remember, you are giving this to somebody else
  to implement}
The accuracy is measures in distance from the correct
value. An accuracy of zero indicates a correct answer. All tasks in the
summarization tasks include an accuracy metric except T7, which will
only be right or wrong, and therefore cannot meaningfully be measured by
a distance.

\subsection{Conceptual Tasks}
\label{sub:conceptual_tasks}

\dmg{i don't like the expression: we are able to see... too long.. be to the point: Table X shows/depicts/ etc ... they
  are not overviews, they are a summary}
In Table~\ref{tab:conceptual_results}, we are able to see an overview of
the results from the conceptual questions.

\begin{table*}[htpb]
  \centering
  \caption{Results from the conceptual questions\dmg{what are the columns? they need captions}}
  \label{tab:conceptual_results}
  \begin{tabular}{ll|r|lrr|rrr}
    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
  \end{tabular}
\end{table*}

Users were able to more closely estimate the number of commits and
merges in the larger tree, but generally took longer than the smaller
tree. The tree with a single node resulted in more variability in the
estimate of number of commits.

It should be noted that these questions were answered after spending
roughly ten minutes attesting to draw a picture that held the answers to
these questions.

\subsection{Summarization Tasks}
\label{sub:summarization_tasks}

In Figure~\ref{fig:summarization_results}, we see all of the
correctness, accuracy, and timing results. The first row of plots are
the correctness results. We see in all cases except for T10, that the
number of instances of a correct result with \tool is roughly the same
as the number of incorrect instances with Gitk. We test each task with
the McNemar Chi-square test~\cite{McNemar1947}.

\begin{figure*}[htpb]
  \centering
  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
  \caption{Results from the summarization tasks. The first row plots the
    correctness metric for each task. The second row plots the
    accuracy metric for each task. The third row plots the time metric
    for each task.}
  \label{fig:summarization_results}
\end{figure*}

With the McNemar's tests, we analyze the participants that change from
being correct with one tool to being incorrect with the other. This does
not take into account the number of people who were able to provide the
correct answer for both tools, nor the people who were incorrect for both
tools. The results show whether providing the participants with \tool
over the current standard, gitk, improves the correctness of the
summarization tasks among the participants. We show the results for the
number of people who were, with both tools, correct, incorrect, and
changed in Table~\ref{tab:correctness_table}.

In all cases, except for task T10, we saw statistically significant
results suggesting that we reject the null hypothesis. We tested each
correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
degree of freedom is $7.879$.

\begin{table*}[htpb]
  \centering
  \caption{Correctness tables showing how many people stayed correct,
    incorrect, and how many people changed between tool}
  \label{tab:correctness_table}
  \begin{tabular}{c|c|c}

    T4 & T5 & T6\\
  \begin{tabular}{cc|rr}
                           &           & \multicolumn{2}{c}{Linvis}\\
                           &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
                           & Incorrect & 15                           & 2\\
  \end{tabular} &

  \begin{tabular}{cc|rr}
                           &           & \multicolumn{2}{c}{Linvis}\\
                           &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
                           & Incorrect & 14                           & 4\\
  \end{tabular}

  &

  \begin{tabular}{cc|rr}
                           &           & \multicolumn{2}{c}{Linvis}\\
                           &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
                           & Incorrect & 16                           & 2\\
  \end{tabular}

  \\\hline
  T7 & T8 & T9\\

  \begin{tabular}{cc|rr}
                           &           & \multicolumn{2}{c}{Linvis}\\
                           &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}  & Correct   & 8                            & 0\\
                           & Incorrect & 13                           & 2\\
  \end{tabular}  &
  \begin{tabular}{cc|rr}
                           &           & \multicolumn{2}{c}{Linvis}\\
                           &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
                           & Incorrect & 15                           & 2\\
  \end{tabular} &
  \begin{tabular}{cc|rr}
                           &           & \multicolumn{2}{c}{Linvis}\\
                           &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
                           & Incorrect & 14                           & 3\\
  \end{tabular}\\\hline

  & T10 & \\
  & \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
    \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
                            & Incorrect & 8                            & 0\\
  \end{tabular} & \\

  \end{tabular}
\end{table*}

The correctness test statistics and conclusion for each summarization
task;

The correctness test statistics, conclusions, and a description of the
accuracy and timing results for each summarization task follow:

\begin{itemize}

  \item

    T4; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
    reject the null hypothesis that \tool does not make a difference in
    the correctness when determining the series of merges that a commit
    takes, with $99.5\%$ confidence.

    There is almost no variation in accuracy between participants using
    \tool, while there is more variance in Gitk. There were two outliers
    among the instances using \tool, which lie within the second and
    third quartiles for the accuracy measure of Gitk. The furthest
    outlier for \tool lies at a distance that is slightly greater than
    the median for Gitk.

    We see similar results in the timing metric; all significant data
    occurs within less time than the second, third, and fourth quartile
    of the Gitk results. Nearly all outlier instances in \tool are
    within the third quartile group of the instances of Gitk with the
    exception of the most extreme, which lies in the fourth quartile
    group.

    The results indicate that \tool is able to assist users with
    determining how a commit is merged more accurately, and more
    quickly.

  \item

    T5; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
    reject the null hypothesis that \tool does not make a difference in
    the correctness when determining the other related commits, with
    $99.5\%$ confidence.

    There is very little variance in the accuracy of results for \tool
    compared to the accuracy of the results in Gitk. The median distance
    is smaller, indicating that the participants performed more
    accurately using \tool.

    There is less variance in the time taken to come to an answer with
    \tool than with Gitk. Furthermore, we see that a majority of the
    participants were able to complete the task in \tool in much less
    time than 50\% of the participants using Gitk.

    The results indicate that \tool is able to assist users in
    understanding which commits are related to a given commit.

  \item

    T6; the resulting $\chi^2$ value is $16$. $16 > 7.879$, therefore we
    reject the null hypothesis that \tool does not make a difference in
    the correctness when determining the number of authors involved with
    the merge tree, with $99.5\%$ confidence.

    We see the same results from this task as with the previous tasks,
    suggesting that \tool assists users determine the number of authors
    more quickly and more accurately.

  \item

    T7; the resulting $\chi^2$ value is $13$. $13 > 7.879$, therefore we
    reject the null hypothesis that \tool does not make a difference in
    the correctness when determining who contributed the most changes,
    with $99.5\%$ confidence.

    This question does not include an accuracy metric, as there are only
    two possible outcomes. We see the same result in the timing metric
    as we did from the previous questions.

    The results suggest that \tool is able to help users correctly
    identify the person who contributed the most lines of code to a
    merge tree in less time.

  \item

    T8; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
    reject the null hypothesis that \tool does not make a difference in
    the correctness when determining how many files were contributed,
    with $99.5\%$ confidence.

    The results of this task are the same as in the previous tasks.
    This suggest that \tool is able to assist users determine how many
    files were modified, more quickly and more accurately.

  \item

    T9; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
    reject the null hypothesis that \tool does not make a difference in
    the correctness when determining which file had the most changes,
    with $99.5\%$ confidence.

    While this question appears to have a similar answer to the correct
    answer for T7, the answer \comA had two files with the same number
    of midifications, thus the participant must correctly identify both
    of these files to be correct. For this reason, we include an
    accuracy metric.

    The results of this task are the same as the previous tasks.
    This suggests that \tool is able to assist users to determine which
    file had the most changes.

  \item

    T10; the resulting $\chi^2$ value is $5.44$. $5.44 < 7.879$,
    therefore we do not reject the null hypothesis, suggesting that
    \tool does not make a difference in determining which modules are
    involved with a merge tree.

    The accuracy and timing results for this task are the same as the
    others, showing that there is less variability in both accuracy and
    the time taken to have an answer between participants using \tool
    than using Gitk. Looking Table~\ref{tab:correctness_table}, we can
    see that, unlike the other questions, a majority of the participants
    were correct using both tools.

\end{itemize}

\subsection{User Opinions}
\label{sub:user_opinions}

Among the 12 pariticipants, there was nearly unanimous agreement that
for conceptual understanding and summarization tasks, \tool was easier
to work with than Gitk. The participants cited the ability to abstract
information about the merge from the clean summarization tables and
simple visualizations. Three participants suggested that someone with a
professional understanding of Gitk and the git commandline may be to
extract the a good conceptual understanding from the DAG, and perform
the summarization tasks from the commandline. One of these three
participants said that they would prefer to have both tools available,
as they are able to complement eachother.

### 2017-07-07 paper/introduction.tex
Updates for the introduction
I've fixed the old introduction to match the current paper.

@@
 % vim:set et sts=2 sw=2 ts=2 tw=72:
-Between 50k and 70k commits are added to the Linux kernel per version
+
+Between 50k and 70k commits are added to the Linux kernel per version,
 requiring maintainers of older versions of the kernel to sift through
@@
 
-\begin{figure*}
+\begin{figure}
         \centering
-        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
-        \caption{A view of the Gitk interface centered on merge
-          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
-          their ancestors and there is no clear path from the commit to the merge
-          with the master branch. Neither Gitk nor Git are capable of showing the
-          commits in master.}
+        \includegraphics[width=0.97\linewidth]{figures/gitk.png}
+        \caption{The Gitk interface centered on commit
+          cdbdd1676a5379f1d5cbd4d476f5e349f445befe, \comB from the user
+          study. The top-left pane shows the DAG and commit log preview,
+          the top center pane shows the authors, and the top-right
+          pane shows the commit dates. The bottom-left pane shows the
+          full commit message, the parents and children of the commit,
+          and the changes made to files in this commit. The bottom-right
+          pane shows the list of the files modified by this commit.}
         \label{fig:gitk}
 %\vspace{-4mm}
-\end{figure*}
+\end{figure}
 
@@
         \centering
-        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
+        \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
         \caption{Github Failure Message showing the DAG of the Linux Project and its
@@
 new commits are created for the Linux project every year; according to
-our previous work\cite{German2015}, a commit takes a median of 30 days
-from the time it is authored until it arrives in the official
-repository. The snapshot of the kernel tomorrow may be different than
-the snapshot from today, containing new commits authored in the past;
-distinguishing these new commits from the commits in the snapshot from
-today is not trivial.
+previous work\cite{German2015}, a commit takes a median of 30 days from
+the time it is authored until it arrives in the official repository. The
+snapshot of the kernel tomorrow may be different than the snapshot from
+today, containing new commits authored in the past; distinguishing these
+new commits from the commits in the snapshot from today is not trivial.
 
@@
 
-This paper makes two contributions; first, we describe a method of
-converting the DAG of the Linux repository into a tree, or
-\emph{merge-tree} of the repository, that represents the path used by a
-commit to reach the master branch; second, we present a method to
-inspect and visualize the history of merges in the Linux project using
-the merge-tree model.
+This paper makes three contributions; first, we describe a method of
+converting the DAG of the Linux repository into the \mt of the
+repository, that represents the path used by a commit to reach the
+master branch; second, we present an implementation enabling the
+inspection and visualization of merges in the Linux project using the
+\mt modle; finally, we validate the \mt model and the implementation of
+the visulization through a controlled user study. We further discuss the
+issues of generalizing the model to other repositories, but present an
+updated version that improves the performance of the algorithm by
+pruning the DAG.\@
 
@@
 the release version, along with a keyword or phrase from the log
-preview, the name of the author, or the commit ID. The user can request
-all merges made by Linus that contain a commit or inner merge that
-matches the search query, or just the commits and merges that match the
-query.
+preview, the name of the author, or the commit hash. The user can
+request all merges made by Linus that contain a commit or inner merge
+that matches the search query, or just the commits and merges that match
+the query.

### 2017-07-08 paper/background.tex
Adding background section

% vim:set et sw=2 ts=4 tw=72:
% Jul 08, 2017

Git is a distributed version control system that is used in many
open-source and closed-source projects. It enables many developers to
write code simultaneously, working on separate branches of the same
project. Git uses a directed acyclic graph model (DAG) for storing the
necessary information regarding the state of the code at any point in
time. Git refers to nodes of the DAG as commits, not distinguishing
between the code-carrying commits and the structural commits that bring
two branches together. In this paper, we will refer to the individual
commits code-carrying as commits, the structural merge commits as
merges, and refer to both types without discrimination as repository
events.

Each repository event contains an ordered list of one or more parents,
except the initial commit which has no parents. The parent relationship
points from the current event toward the initial commit.  Commits have a
single parent, while merges will have at least two parents, the parent
on the current branch followed by the branches being merged into the
current branch. Git allows for the merging of multiple branches
simultaneously in octopus merges, where the order of the parents
represents the order that the branches were specified for being merged.
A \foxtrot\footnote{See
  \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
  for a full description of the issue.} merge occurs when the parent on
the current branch is replaced by a parent from a branch being merged
into the current branch, thus confounding the branch relationships.

The repository events in the DAG are immutable; once a repository event
is created, it can't be changed. Git allows operations to alter commits
and re-order them, but this will create a new commit with a new commit
hash.  This property makes a repository event unable to record the
traversal of merges into the master branch of the repository. A commit
cannot have references in the direction of the merge that merges the
commit, as this reference will change, requiring the commit to be
changed. While this provides some challenge, git eliminates this issue
of the DAG with the \verb|git log --children| command, which traverses
the DAG and inverts the edges to provide the reference from parent to
child.

Under perfect conditions, visualizations of the DAG show how a
repository changes over time; where logical branches are being made,
where merges are happening, and who is making the changes. Most
repositories do not portray the perfect conditions, where there are
relatively few branches and the merges are well-behaved, leaving a
well-defined master branch. Most repositories do not mantain the perfect
structure; either introducing \foxtrot merges, containing many commits,
or many branches, making the DAG difficult to interpret.

### 2017-07-08 paper/introduction.tex
Updated intro, model section, and makefile

@@
         \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
-        \caption{Github Failure Message showing the DAG of the Linux Project and its
-                relationship to other forks.}
+        \caption{Github normally shows a visualization of the DAG,
+          showing the commits, branches, and forks, but is unable to
+          generate a visualization for projects at the size of the Linux
+          repository.}
         \label{fig:gitfail}
@@
 
-This representation works in smaller projects; it enables users to see
-when changes are made, when these changes are merged, how each branch is
-interacting, and the point where a branch forks from the master branch.
-In large modular projects, like the Linux kernel, the DAG becomes a mess
-of merges and commits (Figure~\ref{fig:gitk}) losing its visual meaning.
-In some cases, the Linux kernel is simply too large for the system to
-generate a visualization; Github provides a DAG view for many projects,
-but is unable to display the visualization for projects at the scale of
-the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and 70k
-new commits are created for the Linux project every year; according to
+The DAG is able to provide a meaningful visualization in smaller
+projects; it enables users to see when changes are made, when these
+changes are merged, how each branch is interacting, and the point where
+a branch forks from the master branch. In large modular projects, like
+the Linux kernel, the DAG becomes a mess of merges and commits
+(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the
+Linux kernel is simply too large for the system to generate a
+visualization; Github provides a DAG view for many projects, but is
+unable to display the visualization for projects at the scale of the
+Linux of the kernel (Figure~\ref{fig:gitfail}). Between 60k and 70k new
+commits are created for the Linux project every year; according to
 previous work\cite{German2015}, a commit takes a median of 30 days from

### 2017-07-08 paper/model.tex
Updated intro, model section, and makefile

@@
 % vim:set et sw=2 ts=4 tw=72:
-Git uses a directed acyclic graph (DAG) as its main data model. In this
-model, a commit has one or more ordered parents (ancestors), except the
-root commits of a repository which do not have ancestors--Linux's git
-repository has two such commits.  Commits are divided into merge
-commits, merging two or more branches, and non-merge commits. Non-merge
-commits have only one parent, while merge commits have two or more. The
-order of the parents matter: the first parent is the branch in which the
-merge is being done, while the rest indicate the branches being merged
-(Git allows merging multiple branches simultaneously). The second parent
-is the first branch merged, the third parent is the second branched
-merged, etc.
-
-Once a commit is created, it is never changed. Git allows operations to
-alter commits or reorder them, but it changes the commit ID in the
-process, effectively replacing it with a new commit. This makes commits
-unable to record the traversal of merges from the commit to the merge
-into the master branch.
+The \mt model inverts the DAG, so that reference point from the parent
+to the child instead of from child to parent. This provides a clear path
+from a commit to the closest merge into the master branch. We define the
+distance as a measure of time from when the commit was initially
+commited to the date it was merged into the project. The merge into the
+master branch becomes the root of the tree, and the commits are the
+leaves. Any merges that that a commit must pass through to reach the
+master branch become an inner node of the tree. The structure is able to
+help a user identify the path that the commit took to reach the master
+branch, as well as understand how commits were grouped together in order
+to be integrated. In this model, not only has the DAG been inverted, but
+the entire notion of parent-child relationship has been reversed. The
+terminology surrounding the parent-child relationship will prove to be
+confusing through the rest of the paper, when we are referring to the
+\mt, the parent is the next node toward the merge into the master
+branch, or the root of the tree. When referring to the DAG, the parent
+relationship is in the opposite direction, from the root toward the
+branch-point. In addition to identifying the path that a commit took to
+being merged, we are able to aggregate the commit metadata at merges as
+we retain the parent-child and child-parent relationship for each event.
 
@@
 Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
-In this case, commits are performed in various repositories and
-branches. The DAG representation of the commits is shown in
-Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
+The sequence starts with the initiall commit in the master branch of the
+master repository at time $t_0$. Repository event 1 is a commit, which
+gets forked into a separate repository, \textit{Repo A}, where another
+commit is made, event 2. Event 5 is a merge event, merging event 2, 3,
+and 4 into \textit{Repo A}. Event 5 is branched from, event 6 happens in
+the new branch, while event 7 is added simultaneously to the original
+branch in \textit{Repo A}. Events 11 and 12 are both merge events,
+merging changes made in \textit{Repo A} into the master branch of the
+master repository.
+
+As every repository is a first-class repository, including local copies
+and forks, Git does not distinguish between forked repositories and
+branches, and in neither case does it explicitly record where a commit
+was made. In this case, commits are performed in various repositories
+and branches. The DAG representation of these events is shown in
+Figure~\ref{fig:repoDAG}.
+
+ Notice that the DAG loses information about
 the master branch and the repository that the master branch is part of.
@@
 Note that the direction of the edges of the DAG have been inverted,
-instead of pointing from the child to the ancestors, it points from the
-ancestor to its successors, forming a path to the master branch. Also
+instead of pointing from the child to the parent, it points from the
+parent to its children, forming a path to the master branch. Also
 note that the DAG has been simplified, showing only a single edge on the
@@
     section represents a different branch and/or repository. Each commit
-    points to its ancestor.}
+    points to its parent. The intial commit is at time $t_0$, and the
+    head is at $t_8$.}
   \label{fig:repoEvents}
@@
                 commit only shows the first 20 one-line summaries messages for the 24
-                non-merge commits it merged. The ending ``...'' is part of the log
+                non-merge commits it merged. The ending ``\ldots'' is part of the log
                 and represents that other commits were merged.}
@@
 We used this information to evaluate the accuracy of the merge-tree
-model extracted from the DAG. The method we followed started with the
+model extracted from the DAG\@. The method we followed started with the
 extraction of the commit history up to July 20, 2016. We computed the
@@
 \begin{itemize}
-\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
-  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
-  {c55d267} the summaries were reordered.
-\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
-  branch (not master). This branch was merged at \mycode{3f17ea6d..} which contained 6809 commits.
-\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
-  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
-  impossible to verify; during this period, however, we correctly identified the merges by Linus into master.
-\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
-presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
-repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
+
+  \item
+
+    Five merges were false-errors because their logs did not contain
+    accurate information (were probably edited by hand). For example in
+    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
+    empty), in \mycode {c55d267\ldots} the summaries were reordered.
+
+  \item
+
+    The heuristic correctly identified that 79 of Linus merges (between
+    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
+    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
+    commits.
+
+  \item
+
+    The heuristic worked perfectly until Sept 4, 2007, the earliest date
+    that it could be verified.  Before this date, and until Dec 12,
+    2006, merges did not include a summary of the commits they included,
+    hence making it impossible to verify; during this period, however,
+    we correctly identified the merges by Linus into master.
+
+  \item
+
+    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
+    presence of a \foxtrot commit (\mycode{c436688\ldots}),
+    which confounded the true master branch of a repository.
+
 \end{itemize}
 
-In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860)
-our heuristic was correct. It failed in 1,542 commits before Dec. 12,
-2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).
+In summary, of the merges after Sept. 4, 2007, our heuristic was correct
+in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
+12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
+2007).

### 2017-07-09 paper/design.tex
Linvis design and implementation

@@
 % vim:set et sw=2 ts=4 tw=72:
+
+We constructed a web-based tool called \tool in order to navigate and
+inspect the \mt model of the kernel. Creating the web-based tool enables
+users to use the system without having to install additional software or
+store a large database, which makes it more accessible, easily
+maintainable, and platform independent.
 
@@
         \item Filter by searching
-        \item View files edited
-        \item View modules
-        \item Tree viewer
+        \item View aggregated summaries of authors, files, and modules
+        \item Visualize \mt{s}
 \end{itemize}
 
-\subsection{Searching}
+\subsection{Search}
+
+\tool provides a search engine for navigating within the kernel,
+filtering commits that are not relevant. The search engine takes a
+plain text query from the user and returns the results that are relevant
+in the order of relevancy. When computing relevancy, the engine uses the
+commit log, the author name, the filenames, the commit hash, the date
+the commit was authored, and the date the commit was committed.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/linvis/search_results.png}
+  \caption{A single merge tree in the search results of \tool, showing
+    the link to the root at the top and a table of the relevant commits
+    in the bottom, sorted by relevance.}
+  \label{fig:search_results}
+\end{figure}
+
+Before presentation, the results are grouped by merge tree. Each
+merge tree group has a link to the root-node of the tree at the top,
+followed by a table of commits and merges within the tree that match the
+query, shown in Figure~\ref{fig:search_results}. The table includes the
+relevancy rank of the entry, the commit log message preview, author, the
+date the repository event was committed, and the date it was authored.
+The merge tree groups are ordered based on the mean of the relevancy
+scores of the relevant results in that tree.
+
+\subsection{Summarization}
+\label{sub:summarization}
+
+\tool uses seven tabs to show the information and visualizations for a
+selected commit or merge. The message tab shows the full commit log
+message. This does not include the diff, but given the commit hash, this
+information can be found directly from the repository. The files tab
+shows an aggregated table of all files that were modified in a merge. It
+includes metrics like the number of lines added, lines removed, total
+lines modified, and the delta, summed across all commits in the merge
+tree that modify this file. A details drop-down button allows a user to
+see exactly which commit makes the changes, as shown in
+Figure~\ref{fig:linvis_files}. If the current repository event being
+viewed is a commit, the aggregate views will only show modifications
+made by the commit. The modules tab shows the modules modified in the
+merge tree. Like the files tab, the modules tab uses a table to show the
+name of the module, the number of commits that are in the merge tree
+that work with the module, and a details button to provide the links to
+those commits.
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/linvis/linvis_files.png}
+  \caption{Table showing the modified files in a merge, with the second
+  entry expanded to show the commit that makes the changes.}
+  \label{fig:linvis_files}
+\end{figure}
+
+The authorship tab is very similar to the files tab, but shows the
+authorship information. It shows the sum of the number of lines added,
+removed, modified, and the delta within the \mt. It also shows
+the number of files that were modified by the author. The details tab is
+organized slightly differently. Instead of organizing these results by
+commit, the details are organized by file, showing which files were
+modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/linvis/linvis_authors.png}
+  \caption{Table showing the authors who made changes in a merge. The
+    entry for Randy Dunlap is expanded, showing the modifications to
+    each file that were made by Randy.}
+  \label{fig:linvis_authors}
+\end{figure}
+
+\subsection{Visualization}
+\label{sub:visualization}
+
+The ability to easily visualize the integration of commits into a
+project is what makes \tool unique from other tools. \tool implements
+three visualizations, the list tree, pack tree, and \rt tree.
+
+\subsubsection{List Tree}
+
+The list tree, depicted in Figure~\ref{fig:linvis_list_tree}, is
+constructed as a series of nested lists. The nesting indicates the
+parent-child relationship. This visualization is text-based enabling
+fast navigation to commits using the built-in search fin most
+web-browsers. The list tree is rooted at the current repository event;
+if the current event being inspected is a commit, only that commit will
+be shown. Conversely, if the selected event is the root, the entire
+merge tree will be shown.
 
@@
         \centering
-        \includegraphics[width=0.47\textwidth]{figures/search.png}
-        \caption{Search View allows filtering commits that were merged in a given
-                period, filtering by author, keyword, or commit ID.}
-        \label{fig:search}
+        \includegraphics[width=0.9\linewidth]{figures/linvis/linvis_list_tree.png}
+        \caption{The list tree visualization}
+        \label{fig:linvis_list_tree}
 \end{figure}
 
-Searching (depicted in Figure \ref{fig:search}) allows a user to filter
-commits and merges that are irrelevant. The search mechanism breaks down
-the results by release version. A user can further narrow down the
-search by specifying a range of dates in which such commits were merged
-by Linus into the master branch---not when the commits were created
-(author date and/or commit date). This distinction is important. We have
-observed commits that have taken years to arrive into the master
-repository after they were originally created.
+\subsubsection{\rt Tree}
 
-A user may then provide a search text, filtering by the author name, the
-commit ID, or keyword from the log. Any part of the author name may show
-up in the results, including searching by email address.
+The \rt tree\cite{Reingold1981}, depicted for two trees in
+Figure~\ref{fig:study_commits}, is the classic tree visualization, with
+the root at the top, leaves at the bottom, and edges between them
+showing the parent-child relationship. This tree visualization works
+very well in many cases, but does not easily visualize very wide trees,
+with many repository events on a single level.
 
-If the user is searching by commit ID, the ID can be specified by using
-any of its unique prefixes. For example, the commit
-\mycode{3f17ea6dea8ba5668873afa54628a91aaa3fb1c0} is returned when the
-user searches for a commit ID of \mycode{3f17e} in the 3.16 Linux
-kernel.
+When a user first looks at the visualization, the current repository
+event is placed in the center. Furthermore, it is highlighted in bright
+orange. Commits, the leaves, are shown in white, while the merges are
+colored shades of blue where lighter blue indicates fewer child events
+of that node, and darker blue indicates more children.
 
-\begin{figure}
+\subsubsection{Pack Tree}
+
+Pack trees\cite{Wang2006} are useful for displaying an overview of large
+data sets. The initial goals for the pack tree were for visualizing file
+systems, which are similar to git repositories in that they are
+relatively shallow, but very wide trees. The tree is represented by sets
+of nested circles. The largest circle, containing all of the other
+nodes, is the root, while the smallest circles are the leaves. In our
+representation, the root maps to the merge into the master branch, while
+the leaves are the commits.
+
+\begin{figure}[htpb]
   \centering
-        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
-        \caption{Search Results. Each table entry is a commit merged in the desired
-                merged window.}
-        \label{fig:results}
+  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
+  \caption{The pack tree visualization, with the root depicted as the
+    outer-most circle, containing all other nodes, the leaves depicted
+    as white circles containing no other nodes. The currently selected
+    event is shown in orange.}
+  \label{fig:linvis_pack}
 \end{figure}
 
-In the search results (seen in Figure~\ref{fig:results}), the user is
-presented with the one-line log message preview, the author's name and
-email, the date the commit was authored, and the date the commit was
-last committed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
-        \caption{Panel showing the Commit Message of a commit.}
-        \label{fig:message}
-\end{figure}
-
-Once a user selects a commit or merge to investigate, they are presented
-with a tabbed pane allowing them to view the full commit log, the files
-edited, the modules involved, and the merge-tree view.
-
-The first tab displays the full commit log (Figure~\ref{fig:message}).
-From this, a user is able to see what they would see had they searched
-for the commit using Git log. This doesn't provide additional
-information to the other tools, but helps to complete the functionality
-of \tool. The commit log provides a user with the information about the
-content of the commit and who has signed-off on the commit to ensure
-that it is of good quality. The message for merges may contain a summary
-of the commits being merged.  The information within these messages is
-highly variable, and is completely dependent on the author's style. As
-the user moves toward the root-level merge, the quality of these
-messages generally improves.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
-        \caption{Panel showing the files modified by all the commits that are part
-                of this merge.}
-        \label{fig:files}
-\end{figure}
-
-The second tab is the files tab (Figure~\ref{fig:files}). This tab
-provides information on what files have been edited, how many lines were
-added, and how many lines were removed in a given commit. For
-non-merges, this functionality is similar to the other tools available.
-Our tree-based design model allows us to extend this functionality to
-merges by aggregating information about all the commits that are
-children of the merge in the merge-tree, which other tools are unable to
-show. To find the number of lines added to a file in a merge, we take
-the sum of the lines added to that file in each of the children of that
-merge. We do the same for calculating the number of lines removed.
-
-% This could be extended to use the patch information to determine when a line has been
-% replaced rather than incrementing the lines added and removed.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
-        \caption{Panel showing the modules changed by all the commits in this merge.}
-        \label{fig:modules}
-\end{figure}
-
-The modules tab (Figure~\ref{fig:modules}) shows the modules that are
-contained within the commit. Modules are not natively recognized by Git,
-and are not going to be present in all repositories. In the Linux
-repository, authors put the module they are working on in the one-line
-summary of the log-message; for example: the log \emph{gcov: add support
-  for FCC 4.9} has updated the \emph{gcov} module (the coverage testing
-tool of the kernel). We heuristically extract the module by taking all
-text in the log summary of commits until the first colon.  Modules are
-logical partitions of the information in the kernel. Depending on where
-the author was working, modules can be general, such as ``bluetooth''
-and ``wireless'', or can be quite specific for individual hardware, such
-as ``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit
-does not correctly follow this format and the heuristic approach fails.
-As with the Files panel, non-merge commits show their corresponding
-information, but for merge commits we aggregate all the modules changed
-in all the commits that are part of that merge.  The output of this view
-is shown in Figure~\ref{fig:modules}.
-
-Finally, we have the Tree view tab. The tree view is designed for
-providing easy navigation of the commits within the merge-tree that is
-rooted in the current merge.  It also provides a clear topological view
-of the merge and the submerges it includes. We have experimented with
-various tree designs to find a design that allows for easy navigation
-and visualization of both large and small trees. We discuss this panel
-in the next subsection.
-
-\subsection{Merge-Tree Views} \label{treeview_section}
-
-The merge-tree view is what makes \tool unique to other tools that
-inspect the DAG of a Git repository. With it, a user can inspect how
-commits are merged on their way to the master branch. We have
-experimented with various types of trees:
-
-\begin{enumerate}
-        \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
-        \item Reingold-Tilford trees provide a clear visual representation of the tree structure of the merge-tree.
-        \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
-                clearly showing the parent-child relationships between commits and merges.
-\end{enumerate}
-
-\subsubsection{List Tree}
-
-The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of
-nested lists, and is designed to more closely model the tree view found
-in file browsers. This tree only contains the commits and merges that
-are within the subtrees of the current merge. A commit will never have
-any items in this tree as it is a leaf. To accompany the tree, we
-include breadcrumbs at the top of the page to enable a user to navigate
-both from the root to the leaves and from a leaf to the root. The last
-item in the breadcrumb list is the current commit, the previous item is
-the parent of the current commit, and the first item is the root merge
-into the kernel.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
-        \caption{List tree view of merge 3f17ea6}
-        \label{fig:list_tree}
-\end{figure}
-
-\subsubsection{Reingold-Tilford Tree}
-
-The Reingold-Tilford tree\cite{Reingold1981} (Figures
-\ref{fig:reingold_tree} and \ref{fig:reingold_tree_zoom}) allows the
-visualization and navigation of the entire merge-tree in an intuitive
-representation of the tree.  This illustrates a clear notion of root and
-leaves, and how to navigate in either direction. Some merge trees are
-very large, containing thousands of commits and merged. While the tree
-is capable of producing a visualization, it becomes far more difficult
-to understand. For example, the merge, 3f17ea6 performed by Linus
-Torvalds June 8 2014, contains 7217 commits and merges. \footnote{This
-  commit can be inspected at
-  \url{http://li.turingmachine/org/commits/3f17ea6dea8ba5668873afa54628a91aaa3fb1c0}}
-
-      \begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/4dc4_tree.pdf}
-        \caption{Merge 4dc4226 is a subtree of 3f17ea6, for the power-management
-          module of the kernel}
-        \label{fig:reingold_tree}
-      \end{figure}
-
-      \begin{figure*}
-        \centering
-        \includegraphics[width=0.98\textwidth]{figures/4dc4_zoom_tree.pdf}
-        \caption{Zoomed out view of the Reingold-Tilford tree for merge 4dc4226}
-        \label{fig:reingold_tree_zoom}
-      \end{figure*}
-
-      The user is initially greeted with their current node centered on
-      the screen.  They are able to zoom the tree by scrolling the
-      mouse, and clicking and dragging to pan the tree. They can see
-      more details about a node by clicking on it, which will provide
-      them a link to the specific page for that commit.
-
-\subsubsection{Bubble Tree}
-
-Bubble trees are useful for providing a clear visualization of wide,
-hierarchical data\cite{Boardman2000}. The tree structure is represented
-by the nesting of nodes; the largest circle is the root, containing all
-the other nodes. The smallest circles do not contain any nodes and
-therefore represent the leaf nodes.
-
-Our implementation of the bubble tree (Figure~\ref{fig:bubble_tree})
-provides the user with a clear picture of where a commit is located in
-the merge tree. We highlight the selected commit or merge in red.
-Non-selected merges are in a shade of blue determined by the depth of a
-node in the tree. The root is the lightest shade of blue, while the
-contained merges are progressively darker; the commits are white.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
-        \caption{Bubble Tree of merge 3f17ea6, merge 4dc4226 is the currently
-          selected merge and is highlighted in orange.}
-        \label{fig:bubble_tree}
-\end{figure}
-
-The bubble tree doesn't have an implicit way of providing additional
-information, placing any text near the nodes makes the tree impossible
-to read, so we include a separate pane in the web page. When a user
-hovers over a node, the pane shows additional information about the
-author and the commit message and a link to the detailed page for that
-commit or merge. If the user clicks on a node, the tree will zoom to
-that node and the information in the info pane will persist, enabling
-the user to click the link. This tree provides an easy mechanism for
-users to navigate from the root to the leaves and vice versa.
+We depict the pack tree in Figure~\ref{fig:linvis_pack}. Like with the
+\rt tree, \tool uses white to indicate the individual commits, and
+orange to indicate the currently selected node. The inner merges are
+colored with shades of blue to indicate the depth in the tree. Darker
+shades of blue indicate that the merge is deeper in the tree.

### 2017-07-10 paper/study.tex
Linvis design and implementation
Adding method details
Adding accuracy metric details for the summarization tasks. Also more
details for measuring the timing from the videos.

@@
 
-\subsection{Questions and Tasks}
-\label{sub:questions}
+\subsection{Tasks}
+\label{sub:tasks}
 
@@
   both with linvis and without it (choice of gitk/git command line) }
+
+% TODO: Move to methodology
+The participants were given the choice of Gitk or the git command line
+tools to perform the tasks .
 
@@
     containing one and seven commits, respectively.}
-  \label{fig:commit_1}
+  \label{fig:study_commits}
 \end{figure}
@@
 \dmg{this is still methodology.}
-
+% TODO: Move to methodology
 We recorded the audio and screen capture video for each participant in
@@
 seconds, the correctness, and the accuracy.
-\dmg{and then somethings you are over precise :): correctness is simply whether their answer to the task was correct or not}
-Correctness is a binary
-value, it is true if the response was correct and false if the response
-was incorrect.
-\dmg{if you are going to use this metric, you need to define precisely. remember, you are giving this to somebody else
-  to implement}
-The accuracy is measures in distance from the correct
-value. An accuracy of zero indicates a correct answer. All tasks in the
-summarization tasks include an accuracy metric except T7, which will
-only be right or wrong, and therefore cannot meaningfully be measured by
-a distance.
+
+The correctness metric measures whether the answer was correct or
+incorrect.
+
+The accuracy metric measures the distance from the correct
+answer. In task T4, the merges are important, and the order of those
+merges is important to the answer. For this task, we use the edit
+distance, which will increment if adding merges, removing merges, and
+changing the order of the merges is necessary. T5 only asks for the set
+of commits, so the order is not important, but the distance will
+increase if adding or removing commits is necessary for getting the
+correct answer. In T6 and T8 the answer is a single number, so the
+distance is simply the difference between the answer provided and the
+correct answer. In T7, there is only a single answer that is correct, so
+the correctness metric and accuracy are the same, for this reason we do
+not measure the accuracy of this task. The answer for T9 is similar in
+requirements to T5, only asking for the set of files that had the most
+changes. We use the same distance function in T9 and T10 as we did in
+T5, measuring the number of items that must be added and removed from
+the answer in order to change the response into the correct answer. % TODO: reword
+T9 appears to be similar to T7 in that there is only a single file name
+that must be provided, but in \comA, two files are modified the same
+amount and are both the correct answer. In order to provide the correct
+answer, the participant must be able to identify both files. In all
+means of accuracy measurement, a correct answer will have a distance of
+zero, while a non-zero distance indicates an incorrect answer.
+
+The timing metric is measured from the video. We begin timing the task
+at the end of the question, and stop timing before the last modification
+to the answer. For this reason, it is possible for times to overlap
+between tasks if the participant changes their answer after answering
+another question. This will also modify the answer used for measuring
+the accuracy.
 
@@
 correctness, accuracy, and timing results. The first row of plots are
-the correctness results. We see in all cases except for T10, that the
+the correctness results. \dmg{be more direct: we see that linvis outperforms gitk by a large margin (then you can
+  clarify with the statement you have below; otherwise you are asking the reader to make that connection; don't leave
+  that to chance. be explicit on what the results tell you, at the high level.)}
+ We see in all cases except for T10, that the
 number of instances of a correct result with \tool is roughly the same
-as the number of incorrect instances with Gitk. We test each task with
-the McNemar Chi-square test~\cite{McNemar1947}.
+as the number of incorrect instances with Gitk.
+\dmg{you don't test the task. You test what? the statistical significance than one tool is better than the other? Again,
+be explicit}
+We test each task with
+the McNemar Chi-square test~\cite{McNemar1947}.\dmg{don't break the paragraphs here.. you keep talking about the
+  test. If you need to break them, break them before this sentence, not after}
 
@@
 
+\dmg{McNemar is used to test if there was an effect on the tool used in the correctness of the answers of the
+  users. It is not an analysis of the participants. Don't elaborate on what it does or does not. Just leave it at what
+  it does: if you want to be precise you can say it ``marginal homogeneity'' of the answers}
+
 With the McNemar's tests, we analyze the participants that change from
-being correct with one tool to being incorrect with the other. This does
+being correct with one tool to being incorrect with the other. \dmg{Again, never, ever, use a pronom like this to start
+  a stence. State that this is instead: The McNemar test..., but i vote for removing this info} This does
 not take into account the number of people who were able to provide the
 correct answer for both tools, nor the people who were incorrect for both
-tools. The results show whether providing the participants with \tool
+tools. \dmg{this is better, see my comment before this paragraph, you can use this sentence instead, except: not the
+  results: the test is used to test the hypothesis that ...} The results show whether providing the participants with \tool
 over the current standard, gitk, improves the correctness of the
-summarization tasks among the participants. We show the results for the
+summarization tasks among the participants. \dmg{Change it to Table shows..., we don't show, we wrote} We show the results for the
 number of people who were, with both tools, correct, incorrect, and
@@
 
-In all cases, except for task T10, we saw statistically significant
-results suggesting that we reject the null hypothesis. We tested each
+In all cases, except for task T10, \dmg{we don't ``see'', ``these is'' instead: in all cases.. there the differences
+  are statistically significant} we saw statistically significant
+results \dmg{remove suggesting} suggesting that we reject the null hypothesis \dmg{did you tell us what the null hypo
+  is?}. \dmg{say this before making conclusions about the results} We tested each
 correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
-degree of freedom is $7.879$.
+degree of freedom is $7.879$.\dmg{what is this number? what is it used for? where does it come from?}
 
@@
 participants said that they would prefer to have both tools available,
-as they are able to complement eachother.
+as they are able to complement each other.

### 2017-07-10 paper/study.tex
more changes

@@
 
-\subsection{Tasks}
-\label{sub:tasks}
+\subsection{Questions and Tasks}
+\label{sub:questions}
 
@@
   both with linvis and without it (choice of gitk/git command line) }
-
-% TODO: Move to methodology
-The participants were given the choice of Gitk or the git command line
-tools to perform the tasks .
 
@@
 \dmg{this is still methodology.}
-% TODO: Move to methodology
+
 We recorded the audio and screen capture video for each participant in
@@
 seconds, the correctness, and the accuracy.
-
-The correctness metric measures whether the answer was correct or
-incorrect.
-
-The accuracy metric measures the distance from the correct
-answer. In task T4, the merges are important, and the order of those
-merges is important to the answer. For this task, we use the edit
-distance, which will increment if adding merges, removing merges, and
-changing the order of the merges is necessary. T5 only asks for the set
-of commits, so the order is not important, but the distance will
-increase if adding or removing commits is necessary for getting the
-correct answer. In T6 and T8 the answer is a single number, so the
-distance is simply the difference between the answer provided and the
-correct answer. In T7, there is only a single answer that is correct, so
-the correctness metric and accuracy are the same, for this reason we do
-not measure the accuracy of this task. The answer for T9 is similar in
-requirements to T5, only asking for the set of files that had the most
-changes. We use the same distance function in T9 and T10 as we did in
-T5, measuring the number of items that must be added and removed from
-the answer in order to change the response into the correct answer. % TODO: reword
-T9 appears to be similar to T7 in that there is only a single file name
-that must be provided, but in \comA, two files are modified the same
-amount and are both the correct answer. In order to provide the correct
-answer, the participant must be able to identify both files. In all
-means of accuracy measurement, a correct answer will have a distance of
-zero, while a non-zero distance indicates an incorrect answer.
-
-The timing metric is measured from the video. We begin timing the task
-at the end of the question, and stop timing before the last modification
-to the answer. For this reason, it is possible for times to overlap
-between tasks if the participant changes their answer after answering
-another question. This will also modify the answer used for measuring
-the accuracy.
+\dmg{and then somethings you are over precise :): correctness is simply whether their answer to the task was correct or not}
+Correctness is a binary
+value, it is true if the response was correct and false if the response
+was incorrect.
+\dmg{if you are going to use this metric, you need to define precisely. remember, you are giving this to somebody else
+  to implement}
+The accuracy is measures in distance from the correct
+value. An accuracy of zero indicates a correct answer. All tasks in the
+summarization tasks include an accuracy metric except T7, which will
+only be right or wrong, and therefore cannot meaningfully be measured by
+a distance.
 

### 2017-07-10 paper/study.tex
Addressing first set of comments in Sec. VII
Addressing the comments regarding the first few paragraphs up to the
Tasks subsection.

Also fixing the sections placing the results for each portion of the
study in the results section.

--- 2017-07-10 paper/study.tex
@@
-\dmg{start by describing what you mean by user study... to evaluate linvis we performed a user study...}
+The purpose of the study is to quantitatively and qualitatively evaluate
+the effectiveness of \tool compared to the DAG-based visualizations
+found in Gitk or the git command line. The study is broken into three
+task-based sections; conceptual tasks, summarization tasks, and
+participant opinions. The study was performed in June of 2017 in
+Kingston ON, in a controlled environment running Ubuntu 14.04. When
+conducting the study, participants were allowed to use both the git
+command line and Gitk together, and \tool. We will refer to Gitk in the
+study, but this includes the git command line, as we recognize that git
+power-users may be more comfortable with the command line than a GUI.
 
-\dmg{and you are evaluating the visualizations, NOT the DAG}
+The primary goal of the study is to determine if \tool is more capable
+of providing the participants of our study with a better conceptual
+understanding of how a commit is integrated into the kernel, and a
+better summarization of various metrics about the files and authors
+involved with the integration. The conceptual questions are to determine
+if the DAG visualization provides enough information to determine how a
+commit, and any related commits, are merged into a project. \tool
+displays this information directly in the form of a tree, so we will not
+be performing a comparison on this section, only evaluating the results
+from Gitk. The summarization tasks compare \tool and Gitk, with the
+participants using one tool followed by the other on two commits from
+merge-trees of differeing sizes. In both of these sections, the
+participant is given a task that they are to find an answer to; we
+evaluate the time it took, if the answer was correct, and how far from
+the right answer they were. The user opinions provide us with a clear
+comparison between the tools from the point-of-view of the participant,
+providing us information about which aspects of each tool they prefered.
 
-We use a controlled mixed-methods users study in order to evaluate the
-ability for the DAG to provide users with a conceptual image of how a
-commit is integrated into a project, to compare the DAG and \mt models,
-and determine which model the users preferred.
-
-In this section, we describe the methods used to for designing and
-implementing the user study.
-\dmg{no, you describe the study, which includes the methods}
-
-\dmg{Start by giving an overall overview of the study: the study consists of a set of tasks that users are expected to
-  perform... do a top down description of what users will do
-}
+% TODO: Add the participant profile
 
@@
 
+\subsection{Participant Profile}
+\label{sub:participant_profile}
+
 \subsection{Results}
@@
 
-\subsection{Conceptual Tasks}
+\subsubsection{Conceptual Tasks}
 \label{sub:conceptual_tasks}
@@
 
-\subsection{Summarization Tasks}
+\subsubsection{Summarization Tasks}
 \label{sub:summarization_tasks}
@@
 
-\subsection{User Opinions}
+\subsubsection{User Opinions}
 \label{sub:user_opinions}

### 2017-07-11 paper/discussion.tex
more comments

@@
 % TODO: Go back to the conceptual task results and make them stronger
+
+\dmg{don't editorialize: simply say: the results show that Linvis is superior to ... in ....
+as we discussed today, it is about the tools more than the model
+}
+
 Overall, the results show fairly conclusive results; the DAG is unable
@@
 
+\dmg{I think you mean challenge using gitk, right? so be explicit}
+
 During the course of the study, we noticed a consistent theme among the
@@
 interested in seeing if the participants would be able to identify the
-tree easily, but the results suggest that this was not the case. The
+tree easily, but the results suggest that this was not the case.
+\dmg{the following is hard to understand. even i don't quite get what you mean}
+The
 issue primarily stemmed from not being able to identify the master
@@
 
+\dmg{divide it onto internal and external validity: internal: the tasks were only two merges, but they
+were randomnly chosen to be representative; the tasks were randomized to avoid bias, few participants, 
+but statistical analysis used to measure significance}
+
+\dmg{regarding external, as you say, participants might not be experts, since they are students,
+  you compared against git and gitk (part of git) because they are the most default tools everybody uses
+  but other tools or version control systems might provide better summarizations --i don't know of any.
+}
+
 While git includes gitk and the git log provides a view of the DAG with
@@
 
+\dmg{this paragraph i don't get:}
 Each question was evaluated independently of the results from the
@@
 
-The problem of finding the master branch is the most difficult, and is
+The problem of identifying the master branch is the most difficult, and is
 not possible without heuristic approaches. The master branch can be
@@
 \end{figure}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "userstudy"
+%%% End:

### 2017-07-11 paper/study.tex
Conceptual and summarization tasks
I think this covers most of the conceptual and summarization task
methodology.

@@
 participant opinions. The study was performed in June of 2017 in
-Kingston ON, in a controlled environment running Ubuntu 14.04. When
-conducting the study, participants were allowed to use both the git
-command line and Gitk together, and \tool. We will refer to Gitk in the
-study, but this includes the git command line, as we recognize that git
-power-users may be more comfortable with the command line than a GUI.
+Kingston ON, in a controlled environment running Ubuntu 14.04.
+Participants were allowed to use both Gitk and the git command line
+tools together, which we will refer to as the git tools. Then they were
+asked to perform the same tasks using \tool.
 
@@
 
-\dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}
+% \dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}
+
+The conceptual tasks are designed to gain better insight on the
+comprehension of the DAG, and determine if the DAG is sufficient for
+understanding what other commits are related to a given commit, and how
+those commits are merged. The tasks are outlined in
+Table~\ref{tab:conceptual_tasks}. The tasks are performed in order
+using the git tools. T1 has the participant draw a diagram of the merge
+tree, which will be used to answer the other two questions. This enables
+us to visually see issues with comprehending the DAG visualizations.
+Participants are given 10 minutes to perform this task, which also is
+used to familiarize the participant with the git tools interface, if
+they are not already, and with the set of commits they will be working
+with. T2 and T3 are drawn directly from the diagram from T1, giving
+quantitative metrics to measure the comprehension. T2 asks the
+participant to verify the number of commits, and T3 asks for the number
+of merges. The three tasks are performed on both commits before
+continuing to the summarization tasks.
 
@@
 \end{table*}
+
+The summarization tasks, listed in Table~\ref{tab:summarization_tasks},
+are designed to compare the ability of the participants to summarize
+information about the two commits used in the study when using git tools
+verses \tool. The tasks are broken into four sets based on what is being
+summarized: merges, authorship, files, and modules. The order that the
+participants perform each task is randomized within each task set, and
+the order of the task sets are then randomized with the exception of the
+merges task set which will always come first. We measure and evaluate
+three metrics for each task: correctness, accuracy, and timing.
+Correctness measures whether the answer was correct or incorrect.
+Accuracy measures how far from being correct the provided answer was.
+Timing is the duration, in seconds, that the participant took to answer
+the task. We begin timing the task at the end of the question, and stop
+timing before the last modification to the answer. For this reason, it
+is possible for times to overlap between tasks if the participant
+changes their answer after answering another question. This will also
+modify the answer used for measuring the accuracy. We measure the timing
+from the screen capture.
+
+\evan{Is this better, or do you still want me to move the information
+  about the randomization to the bottom?}
 
@@
 
-\dmg{avoid editorializing your text: remove purely}
+While the merge task set is part of the summarization task group, it
+works less with the summarization of the commits and more with
+comprehension of the models. The task is designed with two goals in
+mind; first we want to measure and define which commits and which merges
+the participant has defined to be in the merge tree; second to help
+solidify which commits and merges will be used in the merge tree.
 
-\dmg{watch for tense... we don't begin... the tasks begin... or better, the first tasks...}
+T4 provides a concrete series of merges that the participant believes to
+be integrating the commit into the repository. T5 provides a concrete
+set of commits that are related to the commit. We place these tasks
+first as the answers to these will be necessary for the participant to
+answer the tasks that follow, specifically finding the merge into the
+master branch. T6 and T7 are related to the authorship at merges,
+determining how many authors are involved and who made the most changes.
+T8 and T9 are related to the files that were modified in the merge,
+determining how many files were modified and which file had the most
+changes. T10 is related to the modules that were modified in the merge.
+Modules are not explicitly defined in git, but are a property of the
+Linux repository that we noticed while looking at various commits. To
+find the modules, take the text up to the first colon in the git log
+preview. For the log preview \textit{``ALSA: kernel docs: fix
+  sound/core/ kernel-doc''} the module is \textit{``ALSA''}, e.g.
 
-The questions and tasks are broken down into four groups. The first two
-groups are purely quantitative in nature; the study begins with
-conceptual tasks to understand whether the DAG is capable of providing
-users with an understanding of how a commit is merged into the master
-branch, and other commits that it is merged with. We continue with
-summarization tasks, to quantitatively understand if the \mt model makes
-a difference in summarization tasks.
-
-The third task set is purely qualitative, \dmg{and are their goal is ... don't imply content in your sentences... be
-  precise; this applies to the paragraph before} to gain better understanding
-of whether participants preferred working with the \mt model in \tool,
-or working with the DAG in Gitk, and what aspects they preferred from
-each tool. The final group of questions is simply to gain a better
-understanding of the demographic of our participants.
-
-\textbf{Conceptual Tasks}
-
-\dmg{given the choice, not allowed.  I don't understand this section.
-  You need to be more top-down in your description. you need to provide
-  context on what the users are expected to do, and why. So what are the
-  tasks? You never explain.  what is the goal of the tasks? how many?
-  what were the tasks? what is expected from each? you need to answer
-  these questions, not necessarily in the same order.  also, remember
-  this is a comparative study, so I presume the users were asked to do
-  both with linvis and without it (choice of gitk/git command line) }
-
-% TODO: Move to methodology
-The participants were given the choice of Gitk or the git command line
-tools to perform the tasks .
-
-The participants were allowed to use Gitk or the git command line tools
-to perform these tasks. These tasks are asked in the specified order,
-allowing tabularr to first, create a conceptual understanding of the
-commits and merges that will be involved with the rest of the tasks that
-follow.tabularmay then use their diagrams to answer T2 and T3. We do not
-collect metrics from \tool for the conceptual tasks, opting instead to
-determine if the DAG is sufficient for understanding how a commit is
-integrated. \tool uses the \mt model show a tree visualization of how
-commits are merged, until the commit reaches the master branch.
-Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.
-
-\dmg{I don't like the present tense, but that is your option}
-
-We provide the participant with 10 minutes per commit to complete the
-first task. The answers to the two other questions in this section are
-drawn from the answer in the first task.
-
-\textbf{Summarization Tasks}
-
-\dmg{again, you are going into details first. First the tasks --see
-  above-- including their grouping.  then at the end, describe the
-  randomization and why it is done (which you did) }
-
-The summarization tasks (Table~\ref{tab:summarization_tasks}) are
-grouped into \dmg{how many?} set of tasks according to their theme. The
-tasks within a given task set are shuffled, and the task sets are
-shuffled, to ensure that similar questions are together, but not in the
-same order, and to ensure that the themes are in randomized order
-between participants. The order of the commits is shuffled, as well. The
-order of the tool is randomized, choosing to start with either \tool or
-Gitk, to ensure that we do not place a bias on one tool through the
-experiment.
-
-As understanding where the merge begins and ends is vital to answering
-the other question, the merge task set will always be the first task set
-presented; however, the questions will still be re-ordered within the
-task set.
+The accuracy metrics are measured slightly differently between tasks due
+to different requirements and structures in the answer. In task T4, both
+the order and the merges are important. We define the distance to be the
+edit distance with insertion, deletion, and transposition at unit cost.
+The accuracy is the minimum edit distance necessary to make the answer
+correct. T5, T9, and T10 are concerned with the set of elements in the
+response; the order is not important. We use the Levenshtein distance,
+with insertion, deletion, and substitution of unit cost, from the
+provided answer to the correct answer. The answers to T6 and T8 are
+single number values; the distance is measured as the absolute value of
+the difference between the provided answer and the correct answer. T7
+has only a single correct answer, thus the distance would either be zero
+or one. For this reason, it is acceptable to omit the accuracy metric
+from this task. The correctness metric is simply whether the answer was
+correct or not; an accuracy of zero indicates a correct answer.
 
@@
 % TODO: Move to methodology
-We recorded the audio and screen capture video for each participant in
-the study for further analysis.
-\dmg{you don't need this sentence, it is redundant}
-After capturing the information, we
-extracted various metrics from the videos.
-\dmg{ you don't record, you measure}
-From the conceptual tasks, we
-recorded the time in seconds, and the answer for the question. We
-extracted three metrics for the summarization tasks, the time in
-seconds, the correctness, and the accuracy.
+% We recorded the audio and screen capture video for each participant in
+% the study for further analysis.
+% \dmg{you don't need this sentence, it is redundant}
+% After capturing the information, we
+% extracted various metrics from the videos.
+% \dmg{ you don't record, you measure}
+% From the conceptual tasks, we
+% recorded the time in seconds, and the answer for the question. We
+% extracted three metrics for the summarization tasks, the time in
+% seconds, the correctness, and the accuracy.
 
-The correctness metric measures whether the answer was correct or
-incorrect.
-
-The accuracy metric measures the distance from the correct
-answer. In task T4, the merges are important, and the order of those
+In task T4, the merges are important, and the order of those
 merges is important to the answer. For this task, we use the edit
@@
 zero, while a non-zero distance indicates an incorrect answer.
-
-The timing metric is measured from the video. We begin timing the task
-at the end of the question, and stop timing before the last modification
-to the answer. For this reason, it is possible for times to overlap
-between tasks if the participant changes their answer after answering
-another question. This will also modify the answer used for measuring
-the accuracy.
 

### 2017-07-11 paper/study.tex
Description of the user-opinion questions
I can't think of how to better describe these questions.

@@
 
-The conceptual tasks are designed to gain better insight on the
+\textbf{The conceptual tasks} are designed to gain better insight on the
 comprehension of the DAG, and determine if the DAG is sufficient for
@@
 
-The summarization tasks, listed in Table~\ref{tab:summarization_tasks},
+\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
 are designed to compare the ability of the participants to summarize
@@
 
-\textbf{User Opinion and Profile}
+\textbf{The user-opinion} questions, listed in
+Table~\ref{tab:opinion_questions}, are designed to determine which tool
+provides a better user experience for summarization and comprehension
+tasks, and what aspects of each assisted with their understanding. T11
+asks for the preference of the participant, given that their goal is
+comprehension and summarization. % TODO add T11 results
+We recognize that neither tool is perfect, and participants may have
+complaints about both tools, but are interested in what parts of each
+tool assisted them with understanding the events in the repository. T12
+is meant to address this. % TODO add T12 results
 
-Finally, we ask the user for their opinions on the tools and for some
-information about their level of experience with git.  \dmg{be more
-  precise: understanding or preference is too broad. What you want to
-  know answers to specific questions regarding how the tools compare and
-  which ones they would prefer} The first three questions are intended
-to provide us with a better understanding of preference, and insight on
-why, the participant had that preference.  The last three questions in
-this section are aimed at getting a better understanding of the
-demographic of our participants.
+\begin{table}[htpb]
+  \centering
+  \caption{User-Opinion Questions}
+  \label{tab:opinion_questions}
+  \begin{tabulary}{0.9\linewidth}{LL}
+    \toprule
+    Task & Description\\
+    \midrule
+    T11 & Given these tasks again, which tool would you prefer?\\
+    T12 & Which aspects of each tool did you like and why?\\
+    \bottomrule
+  \end{tabulary}
+\end{table}
 
-\begin{enumerate}
-  \item Given these tasks again, which tool would you prefer to use?
-  \item Which aspects of each tool did you like and why?
-  \item How long have you used git?
-  \item If you have used git, for what kind of projects? (personal,
-    school courses, professional?)
-  \item If you have used git, how many commits, files, and contributors
-    were involved with the largest repository you have worked with?
-\end{enumerate}
+% \textbf{User Opinion and Profile}
+
+% Finally, we ask the user for their opinions on the tools and for some
+% information about their level of experience with git.  \dmg{be more
+%   precise: understanding or preference is too broad. What you want to
+%   know answers to specific questions regarding how the tools compare and
+%   which ones they would prefer} The first three questions are intended
+% to provide us with a better understanding of preference, and insight on
+% why, the participant had that preference.  The last three questions in
+% this section are aimed at getting a better understanding of the
+% demographic of our participants.
+% \begin{enumerate}
+%   \item Given these tasks again, which tool would you prefer to use?
+%   \item Which aspects of each tool did you like and why?
+%   \item How long have you used git?
+%   \item If you have used git, for what kind of projects? (personal,
+%     school courses, professional?)
+%   \item If you have used git, how many commits, files, and contributors
+%     were involved with the largest repository you have worked with?
+% \end{enumerate}
+
+\subsection{Study Procedure}
+\label{sub:study_procedure}
 

### 2017-07-12 paper/study.tex
Study: Adding commit selection

@@
 
-% \textbf{User Opinion and Profile}
+\subsection{Commit Selection}
+\label{sub:commit_selection}
 
-% Finally, we ask the user for their opinions on the tools and for some
-% information about their level of experience with git.  \dmg{be more
-%   precise: understanding or preference is too broad. What you want to
-%   know answers to specific questions regarding how the tools compare and
-%   which ones they would prefer} The first three questions are intended
-% to provide us with a better understanding of preference, and insight on
-% why, the participant had that preference.  The last three questions in
-% this section are aimed at getting a better understanding of the
-% demographic of our participants.
-% \begin{enumerate}
-%   \item Given these tasks again, which tool would you prefer to use?
-%   \item Which aspects of each tool did you like and why?
-%   \item How long have you used git?
-%   \item If you have used git, for what kind of projects? (personal,
-%     school courses, professional?)
-%   \item If you have used git, how many commits, files, and contributors
-%     were involved with the largest repository you have worked with?
-% \end{enumerate}
+We include two commits in the study to detect differences between merge
+tree sizes. The order that the commits are presented to each participant
+is randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
+October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
+and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
+not including the root, while 50\% of the trees contain at most seven
+nodes non-root nodes.
+% TODO: get the number for this if there is time
+75\% of the trees contain at least 51 nodes, and the largest tree
+contains 7217 nodes.
+
+From this information, we selected one tree with a single commit, and
+one tree with seven other nodes to represent a small tree and a
+medium-sized tree. A majority of the trees are flat, where all nodes
+merge directly into the root node. Of the 8031 trees containing at least
+seven nodes, only 593 contained at least a single internal merge node.
+Trivially, trees with a single node cannot have any internal nodes, and
+of the 624 trees with seven nodes, only 135 contained at least one inner
+node. To increase the complexity of the medium-sized tree, we randomly
+selected one of the 135 trees to represent the medium tree.
+
+From the 2008 trees containing only a single node, we selected one at
+random, and trivially selected the only commit in the tree. From the
+trees containing seven nodes, we selected one randomly, then randomly
+selected one of the commits in the tree. We selected commit
+\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
+a single node, which we will refer to as \comA, and commit
+\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe} from the tree containing
+seven nodes, which we will refer to as \comB. A visual representation
+for these trees is shown in Figure~\ref{fig:study_commits}.
+
+\begin{figure}[bpt]
+  \centering
+  \begin{tabular}{ m{1.5cm} m{3cm} }
+    \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
+    \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
+  \end{tabular}
+  \caption{The merge trees used in the user study,
+    containing one and seven commits, respectively.}
+  \label{fig:study_commits}
+\end{figure}
+
+\subsection{Participant Profile}
+\label{sub:participant_profile}
 
@@
 \label{sub:study_procedure}
+
+The study begins with a short introduction to the \mt model, followed by
+an introduction to Gitk and \tool. We provided two examples of
+converting the DAG to the \mt. The first DAG was a flat merge, so the
+corresponding merge tree did not have any internal nodes. The second was
+not a flat merge, so the corresponding merge tree contained a single
+internal node. Any questions about the model were answered at this time.
+Following the introduction to the \mt model, we introduce Gitk and
+\tool, providing information about what each pane shows in Gitk, and
+what each tab does in \tool.
+
+To mitigate the bias between tools, commits, and tasks, we randomize the
+orders.
 
@@
 where appropriate.
-
-\subsection{Commit Selection}
-\label{sub:commit_selection}
-
-\dmg{see my comments in previous section. it is ok to explain them here,
-  but you need to explain what they are briefly in the previous section}
-
-We chose two commits for use in the study. The order that the commits
-are presented is randomized between participants, but the order is kept
-consistent through the tasks described in the next subsection.
-
-\dmg{this place does not require a description of the database. This
-  requires just to say the commits/merge set (dates and numbers are ok)
-  only} We use the \tool database, containing commits and merge-tree
-information from April 16th 2005 to October 14th, 2014, which
-corresponds to being between Linux release 2.6.12-rc3 and Linux
-3.17-rc1. \evan{Cite the database even more?} \dmg{again, i have
-  problems with your writing: ``we wanted to use in the study commits of
-  different size, to determine if the size had an impact in the
-  completion of the tasks. We chose commits of three different sizes...}
-We chose the commits based on tree sizes. We \dmg{we didn't find: this
-  is a fact, not a discovery; what is a majority? 50\%+1? be precise}
-\dmg{drop the word inner, and do it in order: 25\% of trees have only
-  one commit, 50\%.., 75\%...} found that a majority of the trees
-contain at most seven inner commits, while more than 25\% of the trees
-contain a single commit. 75\% of the tree contained up to 51 inner
-commits and merges, and finally the largest tree contained 7217 nodes.
-We chose to work with the trees in \dmg{if you are going to use the
-  words quartile, just use them before in your description eg: the
-  distribution of the sizes of the trees is x, y merges, whose size
-  corresponded to the 25 ,50 quartiles, simple and precisee} the first
-and second quartile, as merge trees of sizes between one and seven, not
-including the merge into the master branch, make up the majority of the
-trees in the database.
-
-\dmg{remove from here, what is here? this is not a road trip :)} From
-here, we selected one tree from the trees of a single commit at random.
-Selecting a commit from that tree is trivial, as there is only one to
-choose.\dmg{remove the obvious, instead concentrate on why the others
-  are not trivial}
-
-We selected one tree of size seven at random from all trees of size
-seven\dmg{how many where there?}. We placed a restriction the trees, the
-selected tree must include at least one internal merge to increase the
-complexity of the trees tested. For each randomly selecting a tree, we
-chose one of its internal commits at random (using the
-\verb|random.choice| function in python 3.6.1).
-
-\dmg{question: so you chose only one commit of each size, and the
-  randomization was which tool to use for each question? or did they
-  have to do the same task for two commits?}
-
-Using this technique, we selected commit
-\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
-a single node (Figure~\ref{fig:commit_1}), which we will refer to as
-\comA, and commit \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe}, from
-the tree containing seven nodes (Figure~\ref{fig:commit_2}), which we
-will refer to as \comB.
-
-\begin{figure}[bpt]
-  \centering
-  \begin{tabular}{ m{1.5cm} m{3cm} }
-    \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
-    \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
-  \end{tabular}
-  \caption{The merge trees used in the user study,
-    containing one and seven commits, respectively.}
-  \label{fig:study_commits}
-\end{figure}
-
-\subsection{Participant Profile}
-\label{sub:participant_profile}
 

### 2017-07-12 paper/study.tex
study: Participant profile
I've added information about the participants and their experience with
git.

@@
 asked to perform the same tasks using \tool.
+
+\evan{Do I need to say where the study took place? I see it showing up
+  in other papers, but it makes it kind of obvious about the lab I'm
+  talking about, doesn't it?}
 
@@
 
+The study was conducted with 12 participants, all of whom were
+masters, PhD, or post-doc researchers in the field of software
+engineering. The participants had between 6 months and 10 years
+experience with git, with the median being 3.5 years. Most of the
+participants had additional experience with SVN and CVS.\@ One of the
+participants in the study worked as a release engineer, studying merge
+practices to determine the best way to merge branches while minimizing
+the number of merge conflicts, on SVN repositories. The participants
+worked with repositories ranging from around 10 commits up to 38000
+commits, with the median being 350 commits. Two of the participants had
+never collaborated with anyone in a repository, while the rest had some
+experience with repositories being modified by multiple people, with the
+most being 219. The median number of collaborators was four.
+
+Five participants had only worked with personal and academic
+repositories. Two participants had only worked with school and
+professional repositories, and two had only worked with professional
+repositories. One participant worked with personal, academic, and
+professional repositories, one participant had only worked with personal
+repositories, and one participant had worked with personal, academic,
+and school repositories.
+
+\evan{Do you want more information about the participants?}
+
+All participants have had at least some experience with the idea of
+version control, branching in repositories, and git. The participants
+are all from the same lab, and each participant worked with both tools
+in the study, thus, keeping the sample populations identical for both
+tools, with some variation between participants in experience with
+repositories.
+
 \subsection{Study Procedure}
@@
 
-To mitigate the bias between tools, commits, and tasks, we randomize the
-orders.
+Each task group is applied to both commits, with the order of the commits
+randomized between participants to mitigate the effects of order-bias.
+
+The conceptual tasks are the second part of the study. These tasks start
+with the participant taking 10 minutes to draw a diagram showing how one
+of the commits
+
+% To mitigate the order bias caused by using one tool before the other,
+% one commit before the other, and the various tasks, we randomize...
 

### 2017-07-12 paper/study.tex
Study procedure
Include the study procedure in the paper.
This is high-level, the depth is in the subsections for the task groups.

@@
 
-Each task group is applied to both commits, with the order of the commits
-randomized between participants to mitigate the effects of order-bias.
+The conceptual tasks follow the introduction, followed by the
+summarization, which are then followed by the user opinion. The study is
+completed with three questions to enable us to know more about the
+participants;
 
-The conceptual tasks are the second part of the study. These tasks start
-with the participant taking 10 minutes to draw a diagram showing how one
-of the commits
+\begin{itemize}
+  \item How long have you used git?
+  \item For what kind of projects have you used git?
+  \item How many commits, files, and collaborators were involved with
+    the largest repository you have worked with?
+\end{itemize}
+
+To mitigate the order bias cased by using one tool before the other or
+one commit before the other, we randomize both the tool and commit
+between participants. The conceptual tasks are run twice, once with
+\comA and once with \comB, in the same order for all participants using
+the git tools.
+
+\evan{This paragraph is kind of tricky to explain well. See if you can
+  understand what I'm trying to say} The summarization tasks are run
+four times, once with \comA and \tool, \comB and \tool, \comA and git
+tools, and \comB and git tools. The participant will start with one
+tool, and complete the tasks for both commits in that tool. The order of
+the commits is the same as in the conceptual tasks. Once they complete
+the summarization tasks for both \comA and \comB, they will switch to
+the other tool and complete the tasks again on the commits in the same
+order. The order of the tasks is consistent for all four runs.
+
+The user opinions do not directly involve working with the tools or the
+commits, so the ordering of the tools
 
@@
 
-\textbf{Procedure}
+We use a script written in python 3.6.1 in order to assure that we do no
+bias the randomization, and ensure that correct ordering is maintained
+between tools, tasks, and commits. The script produces the entire script
+for the study, so we only need to read it to correctly conduct the
+study. Re-running the script generates the study for the next
+participant.
 
-\dmg{tell me first why you need to create a script, so I know what you
-  do something: we did random assignment of tasks (explain what you did)
-  and randomized the order in which the tasks were performed.  Then
-  explain why you need a script}
-
-\dmg{it does not randomize the order the tools. It randomizes which tool
-  is going to be used on each tasks.  be precise with your language}
-
-\dmg{what you explained how commits relate to tasks and how they get
-  assigned? think as if you are explaining this to somebody who has to
-  do this: you are the designer, explaining the why and the how }
-
-For each participant, we run a simple python script to generate the
-script for the study, eliminating the possibility of bias from us. The
-script randomizes the order of the tools, the order of the commits, the
-order of the task sets, and the order of the tasks within each task set,
-where appropriate.
+\evan{Should I include the python script? It would make explaining how
+  the test was conducted 100\% clear.}
 

### 2017-07-13 paper/study.tex
Study procedure: adding recording method
Fixing conceptual questions results section
Fixing the wording in the conceptual questions. More concise, and more
details.

Should probably include some of the drawing from the study though.

@@
 study. Re-running the script generates the study for the next
-participant.
+participant. We use screen capture to record the audio and video. From
+the screen capture, we measure the time, as well as recording the
+responses in the study.
 
@@
 
+% TODO: maybe move the time measuring procedure from summ. to here
+
 \subsection{Results}
@@
 
-\dmg{this is still methodology.}
-% TODO: Move to methodology
-% We recorded the audio and screen capture video for each participant in
-% the study for further analysis.
-% \dmg{you don't need this sentence, it is redundant}
-% After capturing the information, we
-% extracted various metrics from the videos.
-% \dmg{ you don't record, you measure}
-% From the conceptual tasks, we
-% recorded the time in seconds, and the answer for the question. We
-% extracted three metrics for the summarization tasks, the time in
-% seconds, the correctness, and the accuracy.
-
-In task T4, the merges are important, and the order of those
-merges is important to the answer. For this task, we use the edit
-distance, which will increment if adding merges, removing merges, and
-changing the order of the merges is necessary. T5 only asks for the set
-of commits, so the order is not important, but the distance will
-increase if adding or removing commits is necessary for getting the
-correct answer. In T6 and T8 the answer is a single number, so the
-distance is simply the difference between the answer provided and the
-correct answer. In T7, there is only a single answer that is correct, so
-the correctness metric and accuracy are the same, for this reason we do
-not measure the accuracy of this task. The answer for T9 is similar in
-requirements to T5, only asking for the set of files that had the most
-changes. We use the same distance function in T9 and T10 as we did in
-T5, measuring the number of items that must be added and removed from
-the answer in order to change the response into the correct answer. % TODO: reword
-T9 appears to be similar to T7 in that there is only a single file name
-that must be provided, but in \comA, two files are modified the same
-amount and are both the correct answer. In order to provide the correct
-answer, the participant must be able to identify both files. In all
-means of accuracy measurement, a correct answer will have a distance of
-zero, while a non-zero distance indicates an incorrect answer.
+\evan{This looks weird not having text between the subsections... Should
+  I put something here?...}
 
@@
 
-\dmg{i don't like the expression: we are able to see... too long.. be to the point: Table X shows/depicts/ etc ... they
-  are not overviews, they are a summary}
-In Table~\ref{tab:conceptual_results}, we are able to see an overview of
-the results from the conceptual questions.
+Table~\ref{tab:conceptual_results} shows a summary of the results from
+conceptual tasks T2 and T3. The answer column contains the correct
+answer to the corresponding task, while the median, mean, and variance
+columns are with respect to the answers provided by the
+participants.\evan{Is this better for describing the columns in the
+  table?} The median(s), mean(s), and variance(s) columns are with
+respect to the time taken, in seconds, by the participants to respond to
+the tasks.
+
+\evan{Should I add some of the drawings to the paper?}
+The results from T1 did not vary greatly between the participants. The
+drawings generally show something in the form of a list, starting from
+the provided commit and traversing the master branch, indicating that
+the participants were unable to determine which merge was in the master
+branch. Many of the participants identified the next tag, v2.6.29-rc6,
+to be ``the master branch'', however, a tag is not a branch. One
+participant was able to draw the correct diagram for \comA directly
+using the git command line, but rejected their response after inspecting
+the DAG visualization provided by Gitk and proceeded to draw a list of
+merges.
 
@@
   \centering
-  \caption{Results from the conceptual questions\dmg{what are the columns? they need captions}}
+  \caption{Results from the conceptual questions}
   \label{tab:conceptual_results}
@@
     Question & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
-    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
-    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
-    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
-    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
+    T2       & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
+    T3       & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
+    T2       & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
+    T3       & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
   \end{tabular}
@@
 It should be noted that these questions were answered after spending
-roughly ten minutes attesting to draw a picture that held the answers to
-these questions.
+roughly ten minutes attempting to draw a picture that held the answers
+to these questions, so the times indicate how quickly the participant
+was able to interpret their conceptual understanding.
 

### 2017-07-13 paper/study.tex
Correctness results for summarization tasks

@@
 Participants were allowed to use both Gitk and the git command line
-tools together, which we will refer to as the git tools. Then they were
-asked to perform the same tasks using \tool.
+tools together, which we will refer to as the git tools or as Gitk. Then
+they were asked to perform the same tasks using \tool.
 
@@
 
-In Figure~\ref{fig:summarization_results}, we see all of the
-correctness, accuracy, and timing results. The first row of plots are
-the correctness results. \dmg{be more direct: we see that linvis outperforms gitk by a large margin (then you can
-  clarify with the statement you have below; otherwise you are asking the reader to make that connection; don't leave
-  that to chance. be explicit on what the results tell you, at the high level.)}
-We see in all cases except for T10, that the
-number of instances of a correct result with \tool is roughly the same
-as the number of incorrect instances with Gitk.
-\dmg{you don't test the task. You test what? the statistical significance than one tool is better than the other? Again,
-  be explicit}
-We test each task with
-the McNemar Chi-square test~\cite{McNemar1947}.\dmg{don't break the paragraphs here.. you keep talking about the
-  test. If you need to break them, break them before this sentence, not after}
+% TODO: Add the null hypotheses for each question to the procedures
+% TODO: Include the Wilcoxon Test and Cliff effect size test results
+
+The results show ( Figure~\ref{fig:summarization_results} ) that in all
+cases, except for T10, the number of instances of a correct result with
+\tool is roughly the same as the number of incorrect instances with
+gitk. \tool outperforms Gitk in correctness by a large margin. We use
+the McNemar Chi-square test~\cite{McNemar1947} with continuity
+correction to test the statistical significance of the hypothesis that
+\tool improves the correctness of the participants. We use $\alpha =
+0.05$, or a $95\%$ confidence level, therefore the $\chi^2$ test
+statistic is 3.841.
 
@@
 
-\dmg{McNemar is used to test if there was an effect on the tool used in the correctness of the answers of the
-  users. It is not an analysis of the participants. Don't elaborate on what it does or does not. Just leave it at what
-  it does: if you want to be precise you can say it ``marginal homogeneity'' of the answers}
+\begin{table}[htpb]
+  \centering
+  \caption{Correctness Results}
+  \label{tab:correctness_table}
+  \begin{tabular}{crrc}
+    \toprule
+    Task & $\chi^2$ & $p$-value & Conclusion\\\midrule
+    T4   & 13.07    & 0.0003    & Reject $H_0$\\
+    T5   & 12.07    & 0.0005    & Reject $H_0$\\
+    T6   & 14.06    & 0.0002    & Reject $H_0$\\
+    T7   & 11.07    & 0.0009    & Reject $H_0$\\
+    T8   & 13.07    & 0.0003    & Reject $H_0$\\
+    T9   & 12.07    & 0.0005    & Reject $H_0$\\
+    T10  & 4.00     & 0.0455    & Reject $H_0$\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
-With the McNemar's tests, we analyze the participants that change from
-being correct with one tool to being incorrect with the other. \dmg{Again, never, ever, use a pronom like this to start
-  a stence. State that this is instead: The McNemar test..., but i vote for removing this info} This does
-not take into account the number of people who were able to provide the
-correct answer for both tools, nor the people who were incorrect for both
-tools. \dmg{this is better, see my comment before this paragraph, you can use this sentence instead, except: not the
-  results: the test is used to test the hypothesis that ...} The results show whether providing the participants with \tool
-over the current standard, gitk, improves the correctness of the
-summarization tasks among the participants. \dmg{Change it to Table shows..., we don't show, we wrote} We show the results for the
-number of people who were, with both tools, correct, incorrect, and
-changed in Table~\ref{tab:correctness_table}.
+The results for the correctness metric are summarized in
+Table~\ref{tab:correctness_table}; the full results are as follows;
 
-In all cases, except for task T10, \dmg{we don't ``see'', ``these is'' instead: in all cases.. there the differences
-  are statistically significant} we saw statistically significant
-results \dmg{remove suggesting} suggesting that we reject the null hypothesis \dmg{did you tell us what the null hypo
-  is?}. \dmg{say this before making conclusions about the results} We tested each
-correctness with an $\alpha = 0.005$. The $\chi^2$ test value with one
-degree of freedom is $7.879$.\dmg{what is this number? what is it used for? where does it come from?}
+% TODO: add results for other metrics
 
-\begin{table*}[htpb]
-  \centering
-  \caption{Correctness tables showing how many people stayed correct,
-    incorrect, and how many people changed between tool}
-  \label{tab:correctness_table}
-  \begin{tabular}{c|c|c}
+\begin{itemize}
+  \item
 
-    T4 & T5 & T6\\
+    T4; The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
+    $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
+    hypothesis.
+
     \begin{tabular}{cc|rr}
@@
                             & Incorrect & 15                           & 2\\
-    \end{tabular} &
+    \end{tabular}\\
+
+  \item
+
+    T5; The $\chi^2$ statistic is $12.07$ with a p-value of $0.0005$.
+    $12.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
+    hypothesis.
 
@@
                             & Incorrect & 14                           & 4\\
-    \end{tabular}
+    \end{tabular}\\
 
-    &
+  \item
+
+    T6; The $\chi^2$ statistic is $14.06$ with a p-value of $0.0002$.
+    $14.06 > 3.841$ and $0.0002 < 0.05$, therefore we reject the null
+    hypothesis.
 
@@
 
-    \\\hline
-    T7 & T8 & T9\\
+  \item
+
+    T7; The $\chi^2$ statistic is $11.07$ with a p-value of $0.0009$.
+    $11.07 > 3.841$ and $0.0009 < 0.05$, therefore we reject the null
+    hypothesis.
 
@@
                             & Incorrect & 13                           & 2\\
-    \end{tabular}  &
+    \end{tabular}\\
+
+  \item
+
+    T8; The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
+    $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
+    hypothesis.
+
     \begin{tabular}{cc|rr}
@@
                             & Incorrect & 15                           & 2\\
-    \end{tabular} &
+    \end{tabular}\\
+
+  \item
+
+    T9; The $\chi^2$ statistic is $12.07$ with a p-value of $0.0005$.
+    $12.07 > 3.841$ and $0.0005 < 0.05$, therefore we reject the null
+    hypothesis.
+
     \begin{tabular}{cc|rr}
@@
       & Incorrect & 14                           & 3\\
-    \end{tabular}\\\hline
+    \end{tabular}
 
-    & T10 & \\
-    & \begin{tabular}{cc|rr}
+  \item
+
+    T10; The $\chi^2$ statistic is $4$ with a p-value of $0.0455$.
+    $4 > 3.841$ and $0.0455 < 0.05$, therefore we reject the null
+    hypothesis. The p-value is very close to the $\alpha$ value in this
+    case, so the conclusion is not as well defined.
+
+    \begin{tabular}{cc|rr}
       &           & \multicolumn{2}{c}{Linvis}\\
@@
       & Incorrect & 8                            & 0\\
-    \end{tabular} & \\
-
     \end{tabular}
-\end{table*}
-
-The correctness test statistics and conclusion for each summarization
-task;
-
-The correctness test statistics, conclusions, and a description of the
-accuracy and timing results for each summarization task follow:
-
-\begin{itemize}
-
-  \item
-
-    T4; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the series of merges that a commit
-    takes, with $99.5\%$ confidence.
-
-    There is almost no variation in accuracy between participants using
-    \tool, while there is more variance in Gitk. There were two outliers
-    among the instances using \tool, which lie within the second and
-    third quartiles for the accuracy measure of Gitk. The furthest
-    outlier for \tool lies at a distance that is slightly greater than
-    the median for Gitk.
-
-    We see similar results in the timing metric; all significant data
-    occurs within less time than the second, third, and fourth quartile
-    of the Gitk results. Nearly all outlier instances in \tool are
-    within the third quartile group of the instances of Gitk with the
-    exception of the most extreme, which lies in the fourth quartile
-    group.
-
-    The results indicate that \tool is able to assist users with
-    determining how a commit is merged more accurately, and more
-    quickly.
-
-  \item
-
-    T5; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the other related commits, with
-    $99.5\%$ confidence.
-
-    There is very little variance in the accuracy of results for \tool
-    compared to the accuracy of the results in Gitk. The median distance
-    is smaller, indicating that the participants performed more
-    accurately using \tool.
-
-    There is less variance in the time taken to come to an answer with
-    \tool than with Gitk. Furthermore, we see that a majority of the
-    participants were able to complete the task in \tool in much less
-    time than 50\% of the participants using Gitk.
-
-    The results indicate that \tool is able to assist users in
-    understanding which commits are related to a given commit.
-
-  \item
-
-    T6; the resulting $\chi^2$ value is $16$. $16 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining the number of authors involved with
-    the merge tree, with $99.5\%$ confidence.
-
-    We see the same results from this task as with the previous tasks,
-    suggesting that \tool assists users determine the number of authors
-    more quickly and more accurately.
-
-  \item
-
-    T7; the resulting $\chi^2$ value is $13$. $13 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining who contributed the most changes,
-    with $99.5\%$ confidence.
-
-    This question does not include an accuracy metric, as there are only
-    two possible outcomes. We see the same result in the timing metric
-    as we did from the previous questions.
-
-    The results suggest that \tool is able to help users correctly
-    identify the person who contributed the most lines of code to a
-    merge tree in less time.
-
-  \item
-
-    T8; the resulting $\chi^2$ value is $15$. $15 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining how many files were contributed,
-    with $99.5\%$ confidence.
-
-    The results of this task are the same as in the previous tasks.
-    This suggest that \tool is able to assist users determine how many
-    files were modified, more quickly and more accurately.
-
-  \item
-
-    T9; the resulting $\chi^2$ value is $14$. $14 > 7.879$, therefore we
-    reject the null hypothesis that \tool does not make a difference in
-    the correctness when determining which file had the most changes,
-    with $99.5\%$ confidence.
-
-    While this question appears to have a similar answer to the correct
-    answer for T7, the answer \comA had two files with the same number
-    of midifications, thus the participant must correctly identify both
-    of these files to be correct. For this reason, we include an
-    accuracy metric.
-
-    The results of this task are the same as the previous tasks.
-    This suggests that \tool is able to assist users to determine which
-    file had the most changes.
-
-  \item
-
-    T10; the resulting $\chi^2$ value is $5.44$. $5.44 < 7.879$,
-    therefore we do not reject the null hypothesis, suggesting that
-    \tool does not make a difference in determining which modules are
-    involved with a merge tree.
-
-    The accuracy and timing results for this task are the same as the
-    others, showing that there is less variability in both accuracy and
-    the time taken to have an answer between participants using \tool
-    than using Gitk. Looking Table~\ref{tab:correctness_table}, we can
-    see that, unlike the other questions, a majority of the participants
-    were correct using both tools.
 

### 2017-07-13 userstudy/analysis/plots.R
New figures: Splitting commits when necessary
Given that not all tasks can have the commits aggregated, we split the
correct ones. The results are what is expected though.

Also scales the time and accuracy boxplots on the y-axis so results can
be compared between tasks more easily.

@@ -12,21 +12,34 @@ colorScheme <- c("#3366CC", "#DC3912", "#FF9900", "#109618", "#990099", "#3B3EAC
 colorSchemeTrans <- c("#3366CC99", "#DC391299", "#FF990099", "#10961899", "#99009999", "#3B3EAC99")
 con <- dbConnect(drv=dbDriver('SQLite'), dbname='data/data.db')
 
+correctness_ignore <- c(4, 8)
+timing_ignore <- c(6)
+
+# Aggregated Correctness
+
 query_response_correct <- "
-SELECT question, 'Correct' AS result, tools.name AS tool FROM summarize
+SELECT question, cid, 'Correct' AS result, tools.name AS tool FROM summarize
 JOIN outcomes ON summarize.outcome = outcomes.id
 JOIN tools ON summarize.tool = tools.id
 WHERE summarize.outcome = 1
 UNION ALL
-SELECT question, 'Incorrect' AS result, tools.name AS tool FROM summarize
+SELECT question, cid, 'Incorrect' AS result, tools.name AS tool FROM summarize
 JOIN outcomes ON summarize.outcome = outcomes.id
 JOIN tools ON summarize.tool = tools.id
 WHERE summarize.outcome != 1
 "
 data <- dbGetQuery(con, query_response_correct)
 questions <- sort(unique(data$question))
+cids <- sort(unique(data$cid))
+
 
 for (question in questions) {
+  if (question %in% correctness_ignore) {
+    print(paste(c("Skipping T", question+1), collapse=''))
+    plot.new()
+    title(main=paste(c('T', question + 1), collapse=''))
+    next
+  }
   d <- data[(data$question == question),]
   counts <- table(d$tool, d$result)
   barplot(counts, legend=rownames(counts), beside=TRUE,
@@ -36,24 +49,28 @@ for (question in questions) {
           border=colorScheme[1:2])
 }
 
+# Aggregated Accuracy
 
 accuracy_query <- "
 SELECT lt.question AS question,
+lt.cid,
 lt.distance AS lindist,
 gt.distance AS gitdist,
 (gt.distance - lt.distance) AS 'diff'
 FROM
-(SELECT user, question, distance FROM summarize WHERE tool = 1) AS lt
+(SELECT user, cid, question, distance FROM summarize WHERE tool = 1) AS lt
 JOIN
[…]

### 2017-07-13 paper/study.tex
Fixing study parts (part 1 of many)
Adding null hypotheses to summarization study.

Adding statistical results for Cross-commit and Correctness.
Fixing the correctness metrics to be split for T5 and T9.

@@
 response; the order is not important. We use the Levenshtein distance,
-with insertion, deletion, and substitution of unit cost, from the
-provided answer to the correct answer. The answers to T6 and T8 are
+with insertion, deletion, and substitution operations of unit cost, from
+the provided answer to the correct answer. The answers to T6 and T8 are
 single number values; the distance is measured as the absolute value of
@@
 correct or not; an accuracy of zero indicates a correct answer.
+
+Statistical significance testing is performed to verify that the results
+are meaningful. We use $alpha = 0.05$ or a $95\%$ confidence level. The
+McNemar $\chi^2$ test\cite{McNemar1947} with continuity correct is used
+to test the statistical significance of the hypothesis that \tool
+improves correctness of the participants. The $\chi^2$ threshold
+statistic at $\alpha = 0.05$ is 3.841. The Wilcoxon\cite{Wilcoxon45}
+test and Cliff's effect size\cite{Cliff93} are applied to the accuracy
+and timing metrics to determine the significance of the difference
+between \tool and Gitk. The Wilcoxon test is applied between commits to
+determine if the size of the merge-tree effects timing and accuracy. The
+null hypotheses and alternative hypotheses are stated in
+Table~\ref{tab:null_alt_hypoths}.
+
+\begin{table*}[h!]
+  \centering
+  \caption{Null and Alternative Hypotheses for Summarization Tasks\evan{Should I say exactly
+      what the null hypotheses are referring to (counting authors,
+      files, etc... or is this good enough? It's already a huge and
+      redundant table) }}
+  \label{tab:null_alt_hypoths}
+  \begin{tabulary}{0.98\linewidth}{CCLL}
+    \toprule
+    Task & Metric             & Null Hypothesis                                                      & Alternative Hypothesis\\\midrule
+    T4   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T4   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T4   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T4   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+
+    T5   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T5   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T5   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T5   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+
+    T6   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T6   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T6   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T6   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+
+    T7   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T7   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T7   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T7   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+
+    T8   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T8   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T8   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T8   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+
+    T9   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T9   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T9   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T9   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+
+    T10   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T10   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T10   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T10   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+    \bottomrule
+  \end{tabulary}
+\end{table*}
 
@@
 
-% TODO: Add the null hypotheses for each question to the procedures
-% TODO: Include the Wilcoxon Test and Cliff effect size test results
+Table~\ref{tab:cross_commit_results} shows whether the merge-tree size
+has an effect on the three studied metrics. An interesting observation
+from the figure is that the corrected for T5 and T9 are effected by the
+size of the merge-tree, but the accuracy is not. The correctness results
+for T5 and T9 are shown in Figure~\ref{fig:correctness_t5_t9}, and the
+timing results for T7 are shown in Figure~\ref{fig:timing_t7}.
 
-The results show ( Figure~\ref{fig:summarization_results} ) that in all
-cases, except for T10, the number of instances of a correct result with
-\tool is roughly the same as the number of incorrect instances with
-gitk. \tool outperforms Gitk in correctness by a large margin. We use
-the McNemar Chi-square test~\cite{McNemar1947} with continuity
-correction to test the statistical significance of the hypothesis that
-\tool improves the correctness of the participants. We use $\alpha =
-0.05$, or a $95\%$ confidence level, therefore the $\chi^2$ test
-statistic is 3.841.
+\begin{table}[htpb]
+  \centering
+  \caption{Cross-Commit Results\evan{Should I group the results by task
+      or by metric? It's by metric right now}}
+  \label{tab:cross_commit_results}
+  \begin{tabular}{ccrl}
+    \toprule
+    Task & Metric      & $p$-value & Conclusion\\\midrule
+    T4   & Correctness & 0.22      & Do not reject $H_0$\\
+    T5   & Correctness & 0.04      & Reject $H_0$\\
+    T6   & Correctness & 0.13      & Do not reject $H_0$\\
+    T7   & Correctness & 0.06      & Do not reject $H_0$\\
+    T8   & Correctness & 0.07      & Do not reject $H_0$\\
+    T9   & Correctness & 0.04      & Reject $H_0$\\
+    T10  & Correctness & 0.62      & Do not reject $H_0$\\
+
+    T4  & Accuracy & 0.94 & Do not reject $H_0$\\
+    T5  & Accuracy & 0.09 & Do not reject $H_0$\\
+    T6  & Accuracy & 0.19 & Do not reject $H_0$\\
+    T8  & Accuracy & 0.16 & Do not reject $H_0$\\
+    T9  & Accuracy & 0.08 & Do not reject $H_0$\\
+    T10 & Accuracy & 0.37 & Do not reject $H_0$\\
+
+    T4  & Time & 0.77 & Do not reject $H_0$\\
+    T5  & Time & 0.97 & Do not reject $H_0$\\
+    T6  & Time & 0.90 & Do not reject $H_0$\\
+    T7  & Time & 0.01 & Reject $H_0$\\
+    T8  & Time & 0.87 & Do not reject $H_0$\\
+    T9  & Time & 0.99 & Do not reject $H_0$\\
+    T10 & Time & 0.68 & Do not reject $H_0$\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
   \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
-  \caption{Results from the summarization tasks. The first row plots the
+  \caption{Aggregated results from the summarization tasks. The first row plots the
     correctness metric for each task. The second row plots the
     accuracy metric for each task. The third row plots the time metric
-    for each task.}
+    for each task. The correctness results for T5, T9, and the timing
+    results for T7 cannot be aggregated across commits, thus are
+    omitted.}
   \label{fig:summarization_results}
@@
 
+\begin{figure*}[htpb]
+  \centering
+
+  \begin{tabular}{ m{6cm} m{6cm} }
+    \includegraphics[width=6cm]{figures/userstudy/correctness/5.pdf} &
+    \includegraphics[width=6cm]{figures/userstudy/correctness/9.pdf}
+  \end{tabular}
+  \caption{Unaggregated Correctness results for T5 and T9 \evan{Split
+      this? Is it clear that between Commit 2 and Commit 1 is where the
+    separation of the plots is?}}
+  \label{fig:correctness_t5_t9}
+\end{figure*}
+
+\begin{figure}[htpb]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/userstudy/time/7.pdf}
+  \caption{Unaggregated Timing results for T7}
+  \label{fig:timing_t7}
+\end{figure}
+
 \begin{table}[htpb]
   \centering
-  \caption{Correctness Results}
+  \caption{Summarization Correctness Results}
   \label{tab:correctness_table}
@@
     T4                  & 13.07    & 0.0003    & Reject $H_0$\\
-    T5   & 12.07    & 0.0005    & Reject $H_0$\\
+    \multirow{2}{*}{T5} & 5.14     & 0.0233    & Reject $H_0$\\
+                        & 5.14     & 0.0233    & Reject $H_0$\\
     T6                  & 14.06    & 0.0002    & Reject $H_0$\\
@@
     T8                  & 13.07    & 0.0003    & Reject $H_0$\\
-    T9   & 12.07    & 0.0005    & Reject $H_0$\\
+    \multirow{2}{*}{T9} & 4.17     & 0.0412    & Reject $H_0$\\
+                        & 6.13     & 0.0133    & Reject $H_0$\\
     T10                 & 4.00     & 0.0455    & Reject $H_0$\\
@@
 
-The results for the correctness metric are summarized in
-Table~\ref{tab:correctness_table}; the full results are as follows;
+Summaries for correctness, accuracy, and timing are shown in
+tables~\ref{tab:correctness_table}.
 
 % TODO: add results for other metrics
+
+The full results are as follows;
 
@@
     hypothesis.
-
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
       & Incorrect & 15                           & 2\\
-    \end{tabular}\\
+    \end{tabular}
+    \vspace{3mm}
 
@@
 
-    T5; The $\chi^2$ statistic is $12.07$ with a p-value of $0.0005$.
-    $12.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
+    T5; The correctness results are split for this task. For \comA, the
+    $\chi^2$ statistic is $5.14$ with a $p-value$ of $0.02334$. $5.14 >
+    3.841$ and $0.02334 < 0.05$, therefore we reject the null
     hypothesis.
-
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
                              &           & Correct                      & Incorrect\\\hline
-      \multirow{2}{*}{Gitk} & Correct   & 5                            & 0\\
-                            & Incorrect & 14                           & 4\\
-    \end{tabular}\\
+      \multirow{2}{*}{Gitk}  & Correct   & 4                            & 0\\
+                             & Incorrect & 7                            & 0\\
+    \end{tabular}
+    \vspace{3mm}
+
+    The $\chi^2$ statistic for \comB is $5.14$ with a p-value of
+    $0.02334$. $4.14 > 3.841$ and $0.02334 < 0.05$, therefore we reject
+    the null hypothesis.
+    \vspace{2mm}
+    \begin{tabular}{cc|rr}
+                             &           & \multicolumn{2}{c}{Linvis}\\
+                             &           & Correct                      & Incorrect\\\hline
+      \multirow{2}{*}{Gitk}  & Correct   & 1                            & 0\\
+                             & Incorrect & 7                            & 4\\
+    \end{tabular}
+    \vspace{3mm}
 
@@
 
-    T9; The $\chi^2$ statistic is $12.07$ with a p-value of $0.0005$.
-    $12.07 > 3.841$ and $0.0005 < 0.05$, therefore we reject the null
-    hypothesis.
-
+    T9; The correctness results are split for this task. For \comA, the
+    $\chi^2$ statistic is $4.17$ with a p-value of $0.0412$. $4.17 >
+    3.841$ and $0.0412 < 0.05$, therefore we reject the null hypothesis.
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
                              &           & Correct                      & Incorrect\\\hline
-      \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
-      & Incorrect & 14                           & 3\\
+      \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
+                             & Incorrect & 6                            & 0\\
     \end{tabular}
+    \vspace{3mm}
+
+    The $\chi^2$ statistic for \comB is $6.125$ with a p-value of
+    $0.0133$. $6.125 > 3.841$ and $0.0133 < 0.05$, therefore we reject
+    the null hypothesis.
+    \vspace{2mm}
+    \begin{tabular}{cc|rr}
+                             &           & \multicolumn{2}{c}{Linvis}\\
+                             &           & Correct                      & Incorrect\\\hline
+      \multirow{2}{*}{Gitk}  & Correct   & 1                            & 0\\
+                             & Incorrect & 8                            & 3\\
+    \end{tabular}
+    \vspace{3mm}
 

### 2017-07-13 paper/study.tex
Fixing study parts (part 2)
Oh that was painful to type.

@@
 
-\begin{table}[htpb]
+\begin{table*}[htpb]
   \centering
-  \caption{Summarization Correctness Results}
-  \label{tab:correctness_table}
+  \caption{Summarized Correctness, Accuracy, and Timing results for the summarization tasks}
+  \label{tab:summarization_table}
+  \begin{tabular}{ccc}
+    % Correctness
     \begin{tabular}{crrc}
@@
     \end{tabular}
-\end{table}
+    &
+
+    % Accuracy
+    \begin{tabular}{clrc}
+      \toprule
+      Task & $p$-value            & Delta Est. & Conclusion\\\midrule
+      T4   & $7.8 \times 10^{-5}$ & -0.61      & Reject $H_0$\\
+      T5   & $7.2 \times 10^{-5}$ & -0.63      & Reject $H_0$\\
+      T6   & $1.2 \times 10^{-5}$ & -0.69      & Reject $H_0$\\
+      T8   & $5.8 \times 10^{-4}$ & -0.54      & Reject $H_0$\\
+      T9   & $0.009$              & -0.42      & Reject $H_0$\\
+      T10  & $0.002$              & -0.35      & Reject $H_0$\\
+      \bottomrule
+    \end{tabular}
+    &
+    % Timing
+    \begin{tabular}{clrc}
+      \toprule
+      Task                 & $p$-value & Delta Est. & Conclusion\\\midrule
+      T4                   & 0.0007    & -0.58      & Reject $H_0$\\
+      T5                   & 0.0142    & -0.42      & Reject $H_0$\\
+      T6                   & 0.0168    & -0.41      & Reject $H_0$\\
+      \multirow{2}{*}{T7}  & 0.2586    & -0.29      & Do not reject $H_0$\\
+                           & 0.0780    & -0.43      & Do not reject $H_0$\\
+      T8                   & 0.0009    & -0.57      & Reject $H_0$\\
+      T9                   & 0.0002    & -0.63      & Reject $H_0$\\
+      T10                  & 0.0001    & -0.66      & Reject $H_0$\\
+      \bottomrule
+    \end{tabular}
+  \end{tabular}
+\end{table*}
 
@@
 \begin{itemize}
-  \item
+  \item T4
 
-    T4; The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
+    \textbf{Correctness:}
+    The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
     $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
@@
 
-  \item
+    \tool improves the correctness when determine the series of merges
+    involved with merging a commit.
 
-    T5; The correctness results are split for this task. For \comA, the
+    \textbf{Accuracy:} The p-value is $7.8 \times 10^{-5}$, which is
+    less than 0.05; therefore we reject the null hypothesis. The Delta
+    estimate is -0.61, which indicates a large effect size. \tool
+    improves accuracy when determining the series of merges involved
+    with merging a commit.
+
+    \textbf{Timing:}
+    The p-value is $0.0007$, which is less than 0.05; therefore we
+    reject the null hypothesis. The delta estimate is -0.58, which
+    indicates a large effect size. \tool decreases the time taken to
+    determine the series of merges involved with merging a commit.
+
+    Overall, \tool is able to help users determine the series of merges
+    that a commit takes more quickly and more accurately than Gitk.
+
+  \item T5
+
+    \textbf{Correctness:}
+    The correctness results are split for this task. For \comA, the
     $\chi^2$ statistic is $5.14$ with a $p-value$ of $0.02334$. $5.14 >
@@
 
-  \item
+    While there is a difference in correctness between different-sized
+    merge-trees, \tool improves correctness of determining the other
+    commits related to a commit in trees of varying sizes.
 
-    T6; The $\chi^2$ statistic is $14.06$ with a p-value of $0.0002$.
+    \textbf{Accuracy:}
+    The p-value is $7.2 \times 10^{-5}$, which is less than 0.05;
+    therefore we reject the null hypothesis. The Delta estimate is
+    -0.63, which indicates a large effect size. \tool improves
+    accuracy when determining other commits related to a commit.
+
+    \textbf{Timing:}
+    The p-value is $0.0142$, which is less than 0.05; therefore we
+    reject the null hypothesis. The delta estimate is -0.58, which
+    indicates a large effect size. \tool decreases the time taken to
+    determine other commits related to a commit.
+
+    Overall, \tool helps users determine other commits that are merged
+    with another commit more quickly and more accurately.
+
+  \item T6
+
+    \textbf{Correctness:}
+    The $\chi^2$ statistic is $14.06$ with a p-value of $0.0002$.
     $14.06 > 3.841$ and $0.0002 < 0.05$, therefore we reject the null
     hypothesis.
-
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
     \end{tabular}
+    \vspace{3mm}
 
-  \item
+    \tool improves correctness when determining how many authors are
+    involved with a merge.
 
-    T7; The $\chi^2$ statistic is $11.07$ with a p-value of $0.0009$.
+    \textbf{Accuracy:} The p-value is $1.2 \times 10^{-5}$, which is
+    less than 0.05; we reject the null hypothesis. The delta estimate is
+    -0.69, which indicates a large effect size. \tool improves accuracy
+    when determining how many authors are involved with a merge.
+
+    \textbf{Timing:} The p-value is 0.0168, which is less that 0.05; we
+    reject the null hypothesis. The delta estimate is -0.41, which
+    indicates a medium effect size. \tool decreases the time taken to
+    determine the number of authors involved in a commit.
+
+    Overall, \tool assists users with determining the number of authors
+    involved in a merge more quickly and more accurately.
+
+  \item T7
+
+    \textbf{Correctness:}
+    The $\chi^2$ statistic is $11.07$ with a p-value of $0.0009$.
     $11.07 > 3.841$ and $0.0009 < 0.05$, therefore we reject the null
     hypothesis.
-
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
     \end{tabular}\\
+    \vspace{3mm}
 
-  \item
+    \tool improves the correctness of participants when determining the
+    author that contributed the most changes to a merge.
 
-    T8; The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
+    \textbf{Timing:} The null hypothesis is not rejected in the case of
+    \comA, but is rejected in the case of \comB. This indicates that
+    \tool is able to help users determine who made the most
+    contributions to a merge in a large tree than in a small tree.
+
+    Overall, \tool improves the correctness for trees of various sizes,
+    but only has an effect on the amount of time taken when the tree is
+    large.
+
+  \item T8
+
+    \textbf{Correctness:}
+    The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
     $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
     hypothesis.
-
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
     \end{tabular}\\
+    \vspace{3mm}
 
-  \item
+    \textbf{Accuracy:} The p-value is $5.8 \times 10^{-4}$, which is
+    less than 0.05; we reject the null hypothesis. The delta estimate is
+    -0.54, indicating a large effect size. \tool improves accuracy when
+    determining how many files were modified at a merge.
 
-    T9; The correctness results are split for this task. For \comA, the
+    \textbf{Timing:} The p-value is 0.0009, which is less than 0.05; we
+    reject the null hypothesis. The delta estimate is -0.57, indicating
+    a large effect size.
+
+  \item T9
+
+    \textbf{Correctness:}
+    The correctness results are split for this task. For \comA, the
     $\chi^2$ statistic is $4.17$ with a p-value of $0.0412$. $4.17 >
@@
 
-  \item
+    \textbf{Accuracy:} The p-value is $0.009$, which is less than
+    $0.05$; we reject the null hypothesis. The delta estimate is -0.42,
+    indicating a medium effect size. \tool improves accuracy when
+    determining which file had the most changes.
 
-    T10; The $\chi^2$ statistic is $4$ with a p-value of $0.0455$.
+    \textbf{Timing:} We reject the null hypothesis; the delta estimate
+    is -0.63, indicating a large effect size. \tool decreases the time
+    taken to determine which file had the most changes.
+
+    Overall, \tool is able to assist users determine which file had the
+    most changes more quickly and accurately.
+
+  \item T10
+
+    \textbf{Correctness:}
+    The $\chi^2$ statistic is $4$ with a p-value of $0.0455$.
     $4 > 3.841$ and $0.0455 < 0.05$, therefore we reject the null
@@
     case, so the conclusion is not as well defined.
-
+    \vspace{2mm}
     \begin{tabular}{cc|rr}
@@
     \end{tabular}
+    \vspace{3mm}
+
+    \textbf{Accuracy:} The null hypothesis is rejected; the delta
+    estimate is -0.35, indicating a medium effect size. \tool provides
+    an improvement in accuracy when determining which modules are
+    involved with a merge.
+
+    \textbf{Timing:} The null hypothesis is rejected; the delta estimate
+    is -0.66, indicating a large effect size. \tool decreases the time
+    taken to determine the modules involved with a merge.
+
+    Overall, \tool assists users determine which modules are involved
+    with a merge more quickly and accurately.
 
 \end{itemize}
+
+\tool is able to assist users with summarizing various aspects of a
+merge more quickly and accurately than Gitk. In two tasks, T5 and T9,
+correctness was affected by the size of a tree, but in both trees, \tool
+was able to provide a statistically significant improvement to the
+correctness. Only in one task, T7, was timing affected by the size of a
+tree; \tool did not significantly improve the time taken to answer T7 on
+the smaller tree, but did improve the time taken on the larger tree.
 

### 2017-07-13 paper/study.tex
Minor cleanup on user opinions
This section is small, just fixing some typos and wording.

@@
 
-Among the 12 pariticipants, there was nearly unanimous agreement that
+Among the 12 participants, there was nearly unanimous agreement that
 for conceptual understanding and summarization tasks, \tool was easier
@@
 simple visualizations. Three participants suggested that someone with a
-professional understanding of Gitk and the git commandline may be to
-extract the a good conceptual understanding from the DAG, and perform
-the summarization tasks from the commandline. One of these three
-participants said that they would prefer to have both tools available,
-as they are able to complement each other.
+professional understanding of Gitk and the git command-line may be to
+extract a conceptual understanding from the DAG and perform the
+summarization tasks. One of these three participants said that they
+would prefer to have both tools available, as they are able to
+complement each other.

### 2017-07-13 paper/discussion.tex
Cleaning up todos
Removing some TODO's, removing some hyphenations.
Trying to clarify discussion a bit

@@
 % Jun 24, 2017
-% TODO: Go back to the conceptual task results and make them stronger
 
-\dmg{don't editorialize: simply say: the results show that Linvis is superior to ... in ....
-as we discussed today, it is about the tools more than the model
-}
-
-Overall, the results show fairly conclusive results; the DAG is unable
-to adequately provide users with an explanation of the changes being
-made in the repository, the DAG is not easily understood, and the DAG is
-unable to provide summarizations of the changes being made at a merge.
-Meanwhile, the \mt model is able to provide a better explanation of how
-a commit is integrated, along with better summarization capabilities. In
-this section, we discuss some observations made during the study,
-threats to the validity of the study, comments made by the participants
-during the study, modifications to the original algorithm to help
-generalize it for more repositories, and future work.
+The results show conclusive results; the DAG is unable to adequately
+provide users with an explanation of the changes being made in the
+repository, the DAG is not easily understood, and the DAG is unable to
+provide summarizations of the changes being made at a merge. Meanwhile,
+the \mt model is able to provide a better explanation of how a commit is
+integrated, along with better summarization capabilities. In this
+section, we discuss some observations made during the study, threats to
+the validity of the study, comments made by the participants during the
+study, modifications to the original algorithm to help generalize it for
+more repositories, and future work.
 
@@
 
-\dmg{I think you mean challenge using gitk, right? so be explicit}
+% \dmg{I think you mean challenge using gitk, right? so be explicit}
 
-During the course of the study, we noticed a consistent theme among the
-participants. The most difficult challenge was identifying the master
-branch of the repository. Given that single-commit merge trees are the
-most common tree, and that they are relatively simple, we were
-interested in seeing if the participants would be able to identify the
-tree easily, but the results suggest that this was not the case.
-\dmg{the following is hard to understand. even i don't quite get what you mean}
-The
-issue primarily stemmed from not being able to identify the master
-branch. Most of the participants realized that they should not be going
-backward, toward the initial commit, very early on. Furthermore, many
-would identify the correct commit; however, they would start listing
-every merge along the master branch as being a merge in the tree, and
-including the commits from that merge. In the case of the single commit,
-simply identifying the master branch would most likely be sufficient in
-providing a clearer understanding of how a commit or flat tree is
-integrated into the master branch of the repository.
+During the course of the study, we noticed an issue that was consistent
+among the participants when using Gitk; the most difficult challenge was
+identifying the master branch of the repository. Many participants would
+identify the tag as being the ``master branch'', but a tag is not a
+branch. The visualization of the DAG does not provide any notion of
+which branch being shown is the master branch. The default assumption is
+that the first branch on the left is the master branch, but this is not
+a good assumption. Furthermore, the visualization in Gitk is not
+consistent, the ordering of the branches changes between runs, so
+identifying the branch once will not guarantee that it will be
+identifiable after restarting Gitk.
+
+If users were able to easily identify the master branch using the
+visualizations of the DAG, the results of the study would likely be very
+different. This would certainly be the case for the single-commit tree;
+it is trivial to summarize information about a single item, but
+identifying that there was only a single commit in the tree was
+non-trivial. This information is important because many trees contain a
+single commit, and the structure of these trees is identical to the
+structure of flat trees. Flat trees are the most common branching
+structure in the repository, with the root being the parent to all of
+the other nodes in the tree. To improve summarization and comprehension,
+it would likely by sufficient to indicate the master branch in the DAG.
 
@@
 instead of having to click the link to the commit after clicking the
-node. This was true in both the Reingold-Tilford tree and the Pack tree.
-The second major usability issue was the inconsistency between the list
-tree and the other tree visualizations. The list tree is rooted at the
-repository event that the user is current looking at, while the other
-tree visualizations show the entire tree, and use bright orange to
-highlight the current node.
+node. This was true in both the \rt tree and the Pack tree. The second
+major usability issue was the inconsistency between the list tree and
+the other tree visualizations. The list tree is rooted at the repository
+event that the user is current looking at, while the other tree
+visualizations show the entire tree, and use bright orange to highlight
+the current node.
 
@@
 
-While git includes gitk and the git log provides a view of the DAG with
+\textbf{Internal Threats:}
+
+The repository contains many commits, of which we only chose two. The
+commits chosen were randomly chosen to be representative of those in the
+repository. Additional commits could be added, but this would include
+additional time costs.
+
+The answers provided by the participants to earlier tasks were not
+taken into account for determining the accuracy. If the participant was
+unable to determine the correct commits for the tree within the
+conceptual task group, the summarization results would be recorded as
+wrong, even if the summarization they provided was correct for the set
+of commits selected to be part of the tree. A further study could be
+done where the correct answer was provided to the participant after each
+task so that errors would not propagate between tasks.
+
+To mitigate the effects of order-bias, the order that the tasks were
+performed was shuffled between participants. Furthermore, the order that
+the tool the tasks were performed on changed between participants, and
+the order of the commits that the tasks were performed on was changed
+between participants.
+
+We were only able to work with 12 participants. We apply various
+statistical tests to ensure that our findings are reasonable.
+
+\textbf{External Threats:}
+
+While git includes gitk and git log provides a view of the DAG with
 \verb|git log --graph|, many of the participants in our study were
-unfamiliar with the DAG visualizations shown by these tools. A possible
-reason for this may be that unless a repository reaches a sufficient
-complexity, both in branching and number of commits, these tools are
-necessary for the normal operation of git. Furthermore, fear of the
-complexity caused by branching may cause personal and academic
-repositories to have relatively simple structures. While they were
-unfamiliar with the DAG structure, these users were also unfamiliar with
-the \mt model, and were still able to better understand and summarize
-the repository using \tool.
+unfamiliar with the DAG visualizations shown by these tools. Other tools
+may provide better summarizations and visualizations, but we are unaware
+of them. We investigated the gui tools for Linux on the git
+website\footnote{\url{https://git-scm.com/download/gui/linux}}, but none
+of the tools, with the exception of Gitk and git log, were able to
+produce a visualization of the DAG for the Linux repository.
+Investigating the tools more deeply, it appears that GitKraken is the
+most popularly cited git client. The visualization provided by GitKraken
+was centered around the same DAG metaphor used by git log. GitKraken was
+unable to produce a visualization for the Linux repository on our
+system.
 
-\dmg{this paragraph i don't get:}
-Each question was evaluated independently of the results from the
-previous question in the summarization tasks. Had we modified the
-correct answer to fit the conceptual understanding of which commits were
-included in the tree, we may have witnessed different results. A
-modification to the study may include providing the participants with
-the correct set of commits prior to the summarization tasks, after the
-merge task set.
+% A possible
+% reason for this may be that unless a repository reaches a sufficient
+% complexity, both in branching and number of commits, these tools are
+% unnecessary for the normal operation of git. Furthermore, fear of the
+% complexity caused by branching may cause personal and academic
+% repositories to have relatively simple structures. While they were
+% unfamiliar with the DAG structure, these users were also unfamiliar with
+% the \mt model, and were still able to better understand and summarize
+% the repository using \tool.
 

### 2017-07-14 paper/background.tex
many changes, looks very good

@@
 % Jul 08, 2017
+
+\section{Background}
 
@@
 simultaneously in octopus merges, where the order of the parents
-represents the order that the branches were specified for being merged.
+represents the order that the branches were specified for being merged\footnote{One octopus merge in Linux
+  (2cde51fbd0f310c8a2c5f977e665c0ac3945b46d) has 66 parents.}.
 A \foxtrot\footnote{See
@@
 the current branch is replaced by a parent from a branch being merged
-into the current branch, thus confounding the branch relationships.
+into the current branch, thus confounding the branch relationships.\dmg{this would benefit from a diagram, since
+most people will have no idea what you are talking about.}
 
 The repository events in the DAG are immutable; once a repository event
-is created, it can't be changed. Git allows operations to alter commits
+is created, it can't be changed. \dmg{the next sentences should be repository events, not commits}
+Git allows operations to alter commits
 and re-order them, but this will create a new commit with a new commit
@@
 commit, as this reference will change, requiring the commit to be
-changed. While this provides some challenge, git eliminates this issue
-of the DAG with the \verb|git log --children| command, which traverses
-the DAG and inverts the edges to provide the reference from parent to
-child.
+changed. \dmg{BE CONSISTENT, Git or git?} git provides
+the command \verb|git log --children| which traverses
+the DAG and prints the inverted edges relationship between events.
 
-Under perfect conditions, visualizations of the DAG show how a
+Under perfect conditions, visualizations of the DAG should show how a
 repository changes over time; where logical branches are being made,
 where merges are happening, and who is making the changes. Most
-repositories do not portray the perfect conditions, where there are
-relatively few branches and the merges are well-behaved, leaving a
-well-defined master branch. Most repositories do not mantain the perfect
-structure; either introducing \foxtrot merges, containing many commits,
-or many branches, making the DAG difficult to interpret.
+repositories are relatively simple, with
+relatively few, small, branches with a clear, well defined
+master branch. However, in large, active repositories it becomes very difficult for developers to use the DAG
+to identify
+the master branch, and as a consequence, where branches start and are merged back in their
+way to master.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-14 paper/introduction.tex
many changes, looks very good

@@
 % vim:set et sts=2 sw=2 ts=2 tw=72:
+
+\section{Introduction}
 
@@
 Linux kernel is simply too large for the system to generate a
-visualization; Github provides a DAG view for many projects, but is
+visualization; Github provides a DAG view for a repository, but is
 unable to display the visualization for projects at the scale of the
@@
 
+\dmg{add to the abstract an abbreviation of this sentence: see my comment in the abstract }
 This paper makes three contributions; first, we describe a method of

### 2017-07-14 paper/lineval.tex
Adding background section
many changes, looks very good

@@
 
-\section{Introduction}
-
 \input{introduction.tex}
 
-\section{Merge-Tree model}
-\label{sec:mergetree}
+\label{sec:background}
+
+\input{background.tex}
 
@@
 being merged into the kernel, the date the commit was authored, the
-patch, and the commit log. Merges contain less metadata, only storing
-the author of the merge, the log, the commit date, and the author date,
+patch, and the commit log. Merges contain similar metadata,
+the author of the merge, the committer of the merge, if it has been rebased, the log, the commit date, and the author date,
 and potentially, changes necessary to address conflicts during the
@@
 
-\section{Related Works}
-\label{sec:related_works}
-
 \input{relatedwork.tex}
-
-\section{Controlled User Study}
-\label{sec:study}
 

### 2017-07-14 paper/model.tex
many changes, looks very good

@@
 % vim:set et sw=2 ts=4 tw=72:
-The \mt model inverts the DAG, so that reference point from the parent
-to the child instead of from child to parent. This provides a clear path
-from a commit to the closest merge into the master branch. We define the
-distance as a measure of time from when the commit was initially
-commited to the date it was merged into the project. The merge into the
-master branch becomes the root of the tree, and the commits are the
-leaves. Any merges that that a commit must pass through to reach the
-master branch become an inner node of the tree. The structure is able to
-help a user identify the path that the commit took to reach the master
-branch, as well as understand how commits were grouped together in order
-to be integrated. In this model, not only has the DAG been inverted, but
-the entire notion of parent-child relationship has been reversed. The
-terminology surrounding the parent-child relationship will prove to be
-confusing through the rest of the paper, when we are referring to the
+
+\section{Merge-Tree model}
+
+\label{sec:mergetree}
+
+The \mt model abstracts the DAG of the repository into a set of \mt{}s. Each \mt is rooted at a merge into the master
+branch. The children of this tree are the commits and merges that are merged together into master by this merge. A \mt is build
+recursively with every merge in the \mt merging a sub-\mt. Any merges that a commit must pass through to reach the
+master branch become an inner node of the tree. 
+Effectively, a \mt is a tree that
+shows the path that commits follow in their way to the master branch of a repository.
+The \mt{}s also help understand how commits were grouped together in order
+to be integrated. In this model, not only has the DAG been inverted (and simplified), but
+the entire notion of parent-child relationship has been reversed.
+
+Please note that when we are referring to the
 \mt, the parent is the next node toward the merge into the master
@@
 
-A short example: assume the commits represented in
+To illustrate this model we will use a small example:
+assume the commits represented in
 Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
@@
 
+\dmg{I think you abbreviated this section too much. It will be helpful to describe the algorithm
+at a very high level,as in the vissoft paper.}
+
 \begin{algorithm}
@@
 2007).
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-14 paper/relatedwork.tex
many changes, looks very good

@@
 
-version control systems monitor the development lifetime of software
+\section{Related Works}
+\label{sec:related_works}
+
+Version control systems monitor the development lifetime of software
 projects. This makes the version control system vital in providing
@@
 % TODO: Check this -- Look into source tree
-Not included in an academic project are both the git commandline with
-the \verb|git log| command, and Gitk, which is a graphical tool shipped
-with git. Many other tools are designed with a similar visual metaphore
+Git itself is shipped with summarization and visualization tools:
+the \verb|git log| command and Gitk, which is a graphical tool for the purpose of browsing
+the DAG of the repository.
+Many other tools are designed with a similar visual metaphore
 as what is presented in Gitk, including Git kraken, Github Desktop and
@@
 these data visualization metaphors are applicable to our system.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-14 paper/study.tex
Cleaning up todos
Removing some TODO's, removing some hyphenations.
many changes, looks very good

@@
+\section{Controlled User Study}
+\label{sec:study}
+
 The purpose of the study is to quantitatively and qualitatively evaluate
@@
 task-based sections; conceptual tasks, summarization tasks, and
-participant opinions. The study was performed in June of 2017 in
-Kingston ON, in a controlled environment running Ubuntu 14.04.
+participant opinions. The study was performed during June of 2017
+%in Kingston ON,
+in a controlled environment running Ubuntu 14.04.
 Participants were allowed to use both Gitk and the git command line
@@
 
-\evan{Do I need to say where the study took place? I see it showing up
-  in other papers, but it makes it kind of obvious about the lab I'm
-  talking about, doesn't it?}
+\dmg{I thought you had randomized the order of tool use? also clarify that they were done with different commits to
+  avoid bias}
+
+% \evan{Do I need to say where the study took place? I see it showing up
+%   in other papers, but it makes it kind of obvious about the lab I'm
+%   talking about, doesn't it?}
 
@@
 involved with the integration. The conceptual questions are to determine
-if the DAG visualization provides enough information to determine how a
+if the DAG-based visualizations (gitk and git command line) provide enough information to determine how a
 commit, and any related commits, are merged into a project. \tool
@@
 
-% TODO: Add the participant profile
-
 \subsection{Tasks}
@@
 
-\textbf{The conceptual tasks} are designed to gain better insight on the
-comprehension of the DAG, and determine if the DAG is sufficient for
-understanding what other commits are related to a given commit, and how
-those commits are merged. The tasks are outlined in
-Table~\ref{tab:conceptual_tasks}. The tasks are performed in order
-using the git tools. T1 has the participant draw a diagram of the merge
-tree, which will be used to answer the other two questions. This enables
-us to visually see issues with comprehending the DAG visualizations.
-Participants are given 10 minutes to perform this task, which also is
-used to familiarize the participant with the git tools interface, if
-they are not already, and with the set of commits they will be working
-with. T2 and T3 are drawn directly from the diagram from T1, giving
-quantitative metrics to measure the comprehension. T2 asks the
-participant to verify the number of commits, and T3 asks for the number
-of merges. The three tasks are performed on both commits before
-continuing to the summarization tasks.
+\dmg{I don't like how you talk about the DAG comprehension. Few paragraphs above I replaced it
+  with the DAG-based tools, which might be a better description, since you are evaluating tool
+vs tool}
 
@@
 \end{table*}
-
-\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
-are designed to compare the ability of the participants to summarize
-information about the two commits used in the study when using git tools
-verses \tool. The tasks are broken into four sets based on what is being
-summarized: merges, authorship, files, and modules. The order that the
-participants perform each task is randomized within each task set, and
-the order of the task sets are then randomized with the exception of the
-merges task set which will always come first. We measure and evaluate
-three metrics for each task: correctness, accuracy, and timing.
-Correctness measures whether the answer was correct or incorrect.
-Accuracy measures how far from being correct the provided answer was.
-Timing is the duration, in seconds, that the participant took to answer
-the task. We begin timing the task at the end of the question, and stop
-timing before the last modification to the answer. For this reason, it
-is possible for times to overlap between tasks if the participant
-changes their answer after answering another question. This will also
-modify the answer used for measuring the accuracy. We measure the timing
-from the screen capture.
-
-\evan{Is this better, or do you still want me to move the information
-  about the randomization to the bottom?}
 
@@
 \end{table}
+
+\textbf{The conceptual tasks} are designed to gain better insight on the
+comprehension of the DAG, and determine if the DAG is sufficient for
+understanding what other commits are related to a given commit, and how
+those commits are merged. The tasks are outlined in
+Table~\ref{tab:conceptual_tasks}. The tasks are performed in order
+using the git tools. T1 has the participant draw a diagram of the merge
+tree, which will be used to answer the other two questions. This enables
+us to visually see issues with comprehending the DAG visualizations.
+Participants are given 10 minutes to perform this task, which also is
+used to familiarize the participant with the git tools interface, if
+they are not already, and with the set of commits they will be working
+with. T2 and T3 are drawn directly from the diagram from T1, giving
+quantitative metrics to measure the comprehension. T2 asks the
+participant to verify the number of commits, and T3 asks for the number
+of merges. The three tasks are performed on both commits before
+continuing to the summarization tasks.
+
+\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
+are designed to compare the ability of the participants to summarize
+information about the two commits used in the study when using git tools\dmg{be consistent, either you say dag-based
+  tools, or git tools, either one is fine, but be consistent}
+versus \tool. The tasks are broken into four sets based on what is being
+summarized: merges, authorship, files, and modules. The order that the
+participants perform each task is randomized within each task set, and
+the order of the task sets are then randomized with the exception of the
+merges task set which will always come first. We measure and evaluate
+three metrics for each task: correctness, accuracy, and timing.
+Correctness measures whether the answer was correct or incorrect.
+Accuracy measures how far from being correct the provided answer was.
+Timing is the duration, in seconds, that the participant took to answer
+the task. We begin timing the task at the end of the question, and stop
+timing before the last modification to the answer. For this reason, it
+is possible for times to overlap between tasks if the participant
+changes their answer after answering another question. This will also
+modify the answer used for measuring the accuracy. We measure the timing
+from the screen capture.
+
+\evan{Is this better, or do you still want me to move the information
+  about the randomization to the bottom?}
+\dmg{I didn't see anything about choosing what commits to inspect here... but this section looks good}
 
@@
 not including the root, while 50\% of the trees contain at most seven
-nodes non-root nodes.
-% TODO: get the number for this if there is time
-75\% of the trees contain at least 51 nodes, and the largest tree
-contains 7217 nodes.
+nodes non-root nodes. 75\% of the trees contain at least 51 nodes, and
+the largest tree contains 7217 nodes.
 
@@
 Five participants had only worked with personal and academic
-repositories. Two participants had only worked with school and
+repositories. Two participants had worked with school and
 professional repositories, and two had only worked with professional
@@
 and school repositories.
-
+\dmg{you made it confusing: just say: how many had experience with
+  professional repositories and remove the above
+}
 \evan{Do you want more information about the participants?}
@@
 repositories.
+
+\dmg{BIG QUESTION}
+for each task, did they do always git tools then linvis for the same commit?  if so, there is potential for learning
+bias: they learn the structure of the commits with git tools, so therefore it was easier for them to do
+them with linvis.
+\dmg{END QUESTION}
+
+\dmg{I'll stop here today}
 
@@
 Summaries for correctness, accuracy, and timing are shown in
-tables~\ref{tab:correctness_table}.
-
-% TODO: add results for other metrics
-
+tables~\ref{tab:summarization_table}.
 The full results are as follows;
@@
 complement each other.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-14 paper/study.tex
editing: study procedure

@@
 an introduction to Gitk and \tool. We provided two examples of
-converting the DAG to the \mt. The first DAG was a flat merge, so the
-corresponding merge tree did not have any internal nodes. The second was
-not a flat merge, so the corresponding merge tree contained a single
-internal node. Any questions about the model were answered at this time.
-Following the introduction to the \mt model, we introduce Gitk and
-\tool, providing information about what each pane shows in Gitk, and
-what each tab does in \tool.
+converting the DAG to the \mt. The first example DAG was a flat merge,
+so the corresponding \mt did not have any internal nodes. The
+second was not a flat merge, so the corresponding \mt contained a
+single internal node. Any questions about the model were answered at
+this time. Following the introduction to the \mt model, we introduce
+Gitk and \tool, providing information about what each pane shows in
+Gitk, and what each tab does in \tool.
 
@@
 summarization, which are then followed by the user opinion. The study is
-completed with three questions to enable us to know more about the
+concluded with three questions to enable us to learn more about the
 participants;
@@
 To mitigate the order bias cased by using one tool before the other or
-one commit before the other, we randomize both the tool and commit
-between participants. The conceptual tasks are run twice, once with
-\comA and once with \comB, in the same order for all participants using
-the git tools.
+one commit before the other, we randomize both the order that the tools
+are used, and the order that the commits are used, between participants.
+The conceptual tasks are run twice, once with \comA and once with \comB,
+in the same order for all participants using the git tools.
 
@@
 the other tool and complete the tasks again on the commits in the same
-order. The order of the tasks is consistent for all four runs.
+order. The order of the tasks is consistent over all four runs. The
+questions in the user opinions section do not directly involve the tools
+or commits.
 
-The user opinions do not directly involve working with the tools or the
-commits, so the ordering of the tools
-
-% To mitigate the order bias caused by using one tool before the other,
-% one commit before the other, and the various tasks, we randomize...
-
-We use a script written in python 3.6.1 in order to assure that we do no
-bias the randomization, and ensure that correct ordering is maintained
-between tools, tasks, and commits. The script produces the entire script
-for the study, so we only need to read it to correctly conduct the
-study. Re-running the script generates the study for the next
-participant. We use screen capture to record the audio and video. From
-the screen capture, we measure the time, as well as recording the
+We use a script written in python 3.6.1 in order to assure that we do
+not bias the randomization, and ensure that correct ordering is
+maintained between tools, tasks, and commits. The script produces the
+entire script for the study, so we only need to read it to correctly
+conduct the study. Re-running the script generates the study for the
+next participant. We use screen capture to record the audio and video.
+From the screen capture, we measure the time, as well as recording the
 responses in the study.
@@
   the test was conducted 100\% clear.}
-
-% TODO: maybe move the time measuring procedure from summ. to here
+\evan{I think it would be far too long then.}
 

### 2017-07-14 paper/study.tex
editing conceptual task results

@@
 \evan{Should I add some of the drawings to the paper?}
+
 The results from T1 did not vary greatly between the participants. The
@@
 participant was able to draw the correct diagram for \comA directly
-using the git command line, but rejected their response after inspecting
+from the git command line, but rejected their response after inspecting
 the DAG visualization provided by Gitk and proceeded to draw a list of
@@
 tree. The tree with a single node resulted in more variability in the
-estimate of number of commits.
-
-It should be noted that these questions were answered after spending
-roughly ten minutes attempting to draw a picture that held the answers
-to these questions, so the times indicate how quickly the participant
-was able to interpret their conceptual understanding.
+estimate of number of commits. It should be noted that these questions
+were answered after spending roughly ten minutes attempting to draw a
+picture that held the answers to these questions, so the times indicate
+how quickly the participant was able to interpret their conceptual
+understanding.
 

### 2017-07-14 paper/study.tex
editing summarization results

@@
 has an effect on the three studied metrics. An interesting observation
-from the figure is that the corrected for T5 and T9 are effected by the
-size of the merge-tree, but the accuracy is not. The correctness results
-for T5 and T9 are shown in Figure~\ref{fig:correctness_t5_t9}, and the
-timing results for T7 are shown in Figure~\ref{fig:timing_t7}.
+from the figure is that the correctness for T5 and T9 are effected by
+the size of the merge-tree, but the accuracy is not. The correctness
+results for T5 and T9 are shown in Figure~\ref{fig:correctness_t5_t9},
+and the timing results for T7 are shown in Figure~\ref{fig:timing_t7}.
 
@@
 \begin{itemize}
-  \item T4
+  \item \emph{T4}: What is the series of merges involved with merging this
+    commit?
 
@@
     hypothesis.
+
     \vspace{2mm}
@@
 
-  \item T5
+  \item \emph{T5}: What other commits are merged with this commit?
 
@@
     the null hypothesis.
+
     \vspace{2mm}
@@
     merge-trees, \tool improves correctness of determining the other
-    commits related to a commit in trees of varying sizes.
+    commits related to a commit in trees of varying sizes. The
+    difference in the two commits stem from the participants that did
+    not change outcomes between the tools.
 
@@
 
-  \item T6
+  \item \emph{T6}: How many authors are involved in this merge?
 
@@
 
-  \item T7
+  \item \emph{T7}: Who contributed the most changes to this merge?
 
@@
     \tool is able to help users determine who made the most
-    contributions to a merge in a large tree than in a small tree.
+    contributions to a merge in a large tree but not in a small tree.
 
     Overall, \tool improves the correctness for trees of various sizes,
-    but only has an effect on the amount of time taken when the tree is
+    but only has an effect on the length of time taken when the tree is
     large.
 
-  \item T8
+  \item \emph{T8}: How many files were modified in this merge?
 
@@
 
+    \tool improves the correctness of users when determining the number
+    of files that were modified in a merge.
+
     \textbf{Accuracy:} The p-value is $5.8 \times 10^{-4}$, which is
@@
     reject the null hypothesis. The delta estimate is -0.57, indicating
-    a large effect size.
+    a large effect size. \tool decreases the time taken to determine the
+    number files modified in a merge.
 
-  \item T9
+    Overall, \tool assists users to determine the number of files
+    modified in a merge more quickly and accurately.
+
+  \item \emph{T9}: Which file(s) had the most changes in this merge?
 
@@
     3.841$ and $0.0412 < 0.05$, therefore we reject the null hypothesis.
+
     \vspace{2mm}
@@
     \vspace{3mm}
+
+    While there is a different in correctness between different-sized
+    \mt{s}, \tool improves the correctness when determining which files
+    had the most changes in trees of varying sizes. The difference in
+    the two commits stem from the participants that did not change
+    outcomes between tools.
 
@@
 
-  \item T10
+  \item \emph{T10}: Which modules does this \mt involve?
 
@@
     case, so the conclusion is not as well defined.
+
     \vspace{2mm}
@@
 
+    \tool improves correctness when determining the modules involved at
+    a merge.
+
     \textbf{Accuracy:} The null hypothesis is rejected; the delta
-    estimate is -0.35, indicating a medium effect size. \tool provides
-    an improvement in accuracy when determining which modules are
-    involved with a merge.
+    estimate is -0.35, indicating a medium effect size. \tool improves
+    accuracy when determining which modules are involved with a merge.
 

### 2017-07-14 paper/study.tex
editing user opinions results

@@
 
-Among the 12 participants, there was nearly unanimous agreement that
-for conceptual understanding and summarization tasks, \tool was easier
-to work with than Gitk. The participants cited the ability to abstract
+Among the 12 participants, there was nearly unanimous agreement that for
+conceptual understanding and summarization tasks; \tool was easier to
+work with than Gitk. The participants cited the ability to abstract
 information about the merge from the clean summarization tables and
@@
 professional understanding of Gitk and the git command-line may be to
-extract a conceptual understanding from the DAG and perform the
-summarization tasks. One of these three participants said that they
-would prefer to have both tools available, as they are able to
+extract a conceptual understanding from the DAG visualization and
+perform the summarization tasks. One of these three participants said
+that they would prefer to have both tools available, as they are able to
 complement each other.

### 2017-07-14 paper/discussion.tex
editing discussion

@@
 
-The results show conclusive results; the DAG is unable to adequately
-provide users with an explanation of the changes being made in the
-repository, the DAG is not easily understood, and the DAG is unable to
-provide summarizations of the changes being made at a merge. Meanwhile,
-the \mt model is able to provide a better explanation of how a commit is
-integrated, along with better summarization capabilities. In this
-section, we discuss some observations made during the study, threats to
-the validity of the study, comments made by the participants during the
-study, modifications to the original algorithm to help generalize it for
-more repositories, and future work.
+The results show conclusive results; the DAG visualization is unable to
+adequately provide users with an explanation of the changes being made
+in the repository, the DAG visualizatoin is not easily understood, and
+is unable to provide summarizations of the changes being made at a
+merge. Meanwhile, the \mt visualization is better able to provide an
+explanation of how a commit is integrated, along with better
+summarization. In this section, we discuss some observations, threats to
+the validity, and comments made by the participants during the study
+followed by discussion about modifications to the original algorithm to
+help generalize it for more repositories, concluded with future work.
+
+% TODO: Future work?
 
@@
 structure in the repository, with the root being the parent to all of
-the other nodes in the tree. To improve summarization and comprehension,
-it would likely by sufficient to indicate the master branch in the DAG.
+the other nodes in the tree; to improve summarization and comprehension
+of these trees, it would likely by sufficient to indicate the master
+branch in the DAG visualization.
 
 While, in general, the participants of the study preferred using \tool
-to Gitk, there were some noticeable usability issues. The most prominent
-issue being the navigation within the tree. The participants expected
-that the tabs would update when they clicked on a node in the tree,
-instead of having to click the link to the commit after clicking the
-node. This was true in both the \rt tree and the Pack tree. The second
-major usability issue was the inconsistency between the list tree and
-the other tree visualizations. The list tree is rooted at the repository
-event that the user is current looking at, while the other tree
-visualizations show the entire tree, and use bright orange to highlight
-the current node.
+to Gitk, there were some noticeable usability issues in \tool. The most
+prominent issue being the navigation within the tree visualizations. The
+participants expected that the tabs would update when they clicked on a
+node in the tree, instead of having to click the link to the commit
+after clicking the node. This was true in both the \rt tree and the Pack
+tree. The second major usability issue was the inconsistency between the
+list tree and the other tree visualizations. The list tree is rooted at
+the repository event that the user is current looking at, while the
+other tree visualizations show the entire tree, and use bright orange to
+highlight the current node.
 
@@
 
-One of the participants of the study had worked as a release manager for
-more than three years, working with both SVN and CVS repositories, but
-had little experience with git.
+One participant in the study worked as a release manager for more than
+three years, working with both SVN and CVS repositories, but had little
+experience with git. This subsection contains improvements made by this
+participant during the study and afterward.
 
@@
 contributor. It is also important to understand order that the related
-commits were made, as they tell the story of what the developer was
-thinking as they were writing the changes. The current merge-tree model
+commits were made, as the order tells the story of what the developer
+was thinking as they were writing the changes. The current \mt model
 does not order the commits, simply placing them randomly. This is the
@@
 information, and provide a better understanding of the next merge
-involved in integrating this commit, but the DAG in gitk provides the
-full story of the commit instead of hiding it behind a layer of
-abstraction.
+involved in integrating this commit, but the DAG visualization in gitk
+provides the full story of the commit instead of hiding it behind a
+layer of abstraction.
 
@@
 
-\dmg{divide it onto internal and external validity: internal: the tasks were only two merges, but they
-were randomnly chosen to be representative; the tasks were randomized to avoid bias, few participants, 
-but statistical analysis used to measure significance}
-
-\dmg{regarding external, as you say, participants might not be experts, since they are students,
-  you compared against git and gitk (part of git) because they are the most default tools everybody uses
-  but other tools or version control systems might provide better summarizations --i don't know of any.
-}
-
 \textbf{Internal Threats:}
@@
 The repository contains many commits, of which we only chose two. The
-commits chosen were randomly chosen to be representative of those in the
-repository. Additional commits could be added, but this would include
-additional time costs.
+commits chosen were randomly selected to be representative of those in
+the repository. Additional commits could be added, but this would
+include additional time costs during the study.
 
@@
 
-To mitigate the effects of order-bias, the order that the tasks were
-performed was shuffled between participants. Furthermore, the order that
-the tool the tasks were performed on changed between participants, and
-the order of the commits that the tasks were performed on was changed
-between participants.
+To mitigate the effects of order-bias, the order the tasks were
+performed, the tools were used, the commits were studied was shuffled
+between participants. A single participant is affected by the effect,
+but this should be mitigated by the results from other participants.
 
@@
 
-While git includes gitk and git log provides a view of the DAG with
-\verb|git log --graph|, many of the participants in our study were
+While git includes gitk and git log provides a visualization of the DAG
+with \verb|git log --graph|, many of the participants in our study were
 unfamiliar with the DAG visualizations shown by these tools. Other tools
 may provide better summarizations and visualizations, but we are unaware
-of them. We investigated the gui tools for Linux on the git
+of any. We investigated the gui tools for Linux on the git
 website\footnote{\url{https://git-scm.com/download/gui/linux}}, but none
@@
 unable to produce a visualization for the Linux repository on our
-system.
+system. Inspecting these tools on smaller repositories show that the
+visualizations of the DAG were very similar to the visualization in
+Gitk.
 
@@
 
-Seeing that the \mt model is able to assist users with understanding how
-commits are integrated, we continue by looking at ways of improving the
-algorithm for faster operations, and making the conversion from DAG to
-\mt feasible over the internet for smaller repositories.
+Seeing that the visualizations of the \mt model are able to assist users
+with understanding how commits are integrated, we continue by looking at
+ways of improving the algorithm for faster operations, and making the
+conversion from DAG to \mt feasible over the internet.
 
@@
 not possible without heuristic approaches. The master branch can be
-confounded by \textit{foxtrot} merges, which change the order of the
+confounded by \foxtrot merges, which change the order of the
 parents in the parent list of a merge. Other than the ordering of the
@@
 will assume that we can rely on the order of the parent list to be
-correct, and that foxtrots do not occur.
+correct, and that \foxtrot{s} do not occur. Furthermore, merge
+information is lost in the event of fast-forward merges, which collapse
+the branch into the current branch.
 
@@
     \Function{Phase 1}{initial commit $root$} : updated tree
-    \State $depth \gets 0$
+    \State $openBranches \gets 0$
     \State $Q \gets root$
@@
     \State $parents \gets cur.parents$
-    \State $depth \gets depth + parents.length - 1$
+    \State $openBranches \gets openBranches + parents.length - 1$
     \For{$index, parent \in parents$}
@@
     \ElsIf {$cur.depth > parent.depth$}
-    \State $depth \gets depth - 1$
+    \State $openBranches \gets openBranches - 1$
     \ElsIf {$cur.depth < parent.depth$}
-    \State $depth \gets depth - 1$
+    \State $openBranches \gets openBranches - 1$
     \State $parent.depth \gets cur.depth$
@@
     \EndFor
-    \doWhile{$Q$ not empty and $depth \ne 0$ }
+    \doWhile{$Q$ not empty and $openBranches \ne 0$ }
     \EndFunction
@@
 The primary difference between the original algorithm and the
-generalization is the main data structure. Algorithm~\ref{alg:original}
+generalization is the main data structure. Algorithm~\ref{fig:alg}
 is centered around the use of the stack, making it akin to a depth-first
-traversal of the DAG.\@ The generalization in
+traversal of the DAG with dynamic programming. The generalization in
 Algorithm~\ref{alg:generalized} is based on the queue, making it akin to
@@
 We will begin with a simplified description of the original algorithm.
-We start at the master branch pointer. We will use the term master
-branch loosely here, meaning any branch that has been labeled in some
-way as being the main branch, this branch may have a different name, but
-standard naming conventions call it the master branch. Starting from the
-master branch label, we traverse the entire master branch to the initial
-commit, recording the commit hash of every repository event encountered.
+The algorithm starts at the reference to the master branch. We will use
+the term master branch loosely here, meaning any branch that has been
+labeled in some way as being the main branch, this branch may have a
+different name, but standard naming conventions call it the master
+branch. Starting from the master branch label, we traverse the entire
+master branch to the initial commit, recording the commit hash of every
+repository event encountered.
 
-The second phase iterates over every commit, computing the shortest
-distance to the master branch. For this reason, the entire repository
-must be available to the algorithm. This is a dynamic programming
-solution, centered around the call stack. The next node on the shortest
-path to the master branch will become the parent of this node in the
-tree.
+The second phase iterates over every repository event, computing the
+shortest distance to the master branch. For this reason, the entire
+repository must be available to the algorithm. This is a dynamic
+programming solution, centered around the call stack. The next node on
+the shortest path to the master branch will become the parent of this
+node in the tree.
 
 If this algorithm is to be generalized into a dynamic, web-based, tool,
-it may not be feasible to download every commit in the repository in a
-reasonable period of time. We instead look toward a breadth-first
-approach at limiting the number of repository events that must be
-downloaded.
+it may not be feasible to download every every repository event in the
+repository in a reasonable period of time. We instead look toward a
+breadth-first approach at limiting the number of repository events that
+must be downloaded.
 
-Focusing on the generalized algorithm~\ref{alg:generalized}, the
-algorithm is broken into two phases. The first phase is where the
-improvements take place. The first phase has three main goals, the first
-is to track the master branch. The second goal labels the depth of each
-repository event that is traverses. The final goal is to record the
-children of each commit, reversing the DAG.\@ The first phase also
-begins the process of reversing the DAG, by providing the candidate
-commits that could possibly be the real parent to this node. The first
-phase can terminate early, if it reaches a point where all branches that
-it is following terminate. The depth metric on line 7 indicates how many
-branches are currently being traversed. In line 11 through 15, we
-decrement the depth, as this is the point where a branch occurs. The
-changing of the depth occurs if there are more events on the same branch
-than between the merge and branch point of a separate branch.
+Focusing on the generalized algorithm (Algorithm~\ref{alg:generalized}),
+the algorithm is broken into two phases. The first phase is where the
+improvements take place. The first phase has three main goals; the first
+is to track the master branch; the second goal labels the depth of each
+repository event that is traverses, or how many merges the event is from
+being merged; the third goal is to record the children of each
+repository event. The first phase also begins the process of reversing
+the DAG, by providing the candidate events that could possibly be the
+real parent to this node. The first phase can terminate early, if it
+reaches a point where all branches that it is following terminate. The
+$openBranches$ metric on line 7 indicates how many branches are
+currently being traversed. In line 11 through 15, we decrement the
+$openBranches$, as this is the point where a branch occurs. The second
+branch closing clause is necessary to handle the case when the depth was
+incorrectly assigned, which can occur if there are more events on the
+current branch than on the side branch.
 
@@
     \toprule
-    Current & Parents & Depth & Q       & Visited & Children\\\midrule
+    Current & Parents & Open Branches & Q       & Visited & Children\\\midrule
     -       & -       & 0             & 11      & 11:0    & \\
@@
 In Table~\ref{tab:ex_traversal}, we follow the algorithm starting at
-merge commit 11 from the example in Figure~\ref{fig:repoDAG}, recording
-the current node, the parents of the current node, the depth after
-updating, and the queue structure after updating. The visited column is
-not dependant on the current step, but keeps track of the nodes that
-have been visited, and the depth at that node. When we reverse the DAG,
-the children in this table will become the candidates for being the
-parent of the corresponding node. For example, the parent of node 7 may
-be resolved to either node 9 or node 8. In this case, the shortest path
-to the master branch is through node 9.
+merge 11 from the example in Figure~\ref{fig:repoDAG}, recording the
+current node, the parents of the current node, the open branches, and
+the queue structure after updating. The visited column is not dependant
+on the current step, but keeps track of the nodes that have been
+visited, the depth, and the children of the corresponding node. When we
+reverse the DAG, the children in this table will become the candidates
+for being the parent of the corresponding node. For example, the parent
+of node 7 may be resolved to either node 9 or node 8. In this case, the
+shortest path to the master branch is through node 9.
 
@@
 section. We have gathered the depths and possible parents (line 17), but
-we need to remove the extra parents. Extra parents are culled in line
-42 of the algorithm. Of the possible parents, we take the parent that is
+we need to remove the extra parents. Extra parents are culled in line 42
+of the algorithm. Of the possible parents, we take the parent that is
 closest to the root, but not beyond. This forces the real parent to
-either be deeper in the tree, or on the same branch. Line 30 to 37 are
-for bootstrapping the queue, ensuring that we are not starting on an
-empty queue.
+either be deeper in the tree, or on the same branch as the current node.
+Line 30 to 37 are for bootstrapping the queue, ensuring that we are not
+starting on an empty queue and is nearly identical to lines 41 to 48.
 
@@
 telling the story of how a developer was working, before the set of
-commits was merged. We developed a small bitbucket plugin to test the
+commits was merged. We developed a small Bitbucket plugin to test the
 design and determine whether it was capable of producing merge trees,

### 2017-07-14 paper/conclusion.tex
editing conclusion

@@
-In this paper, we develop a tree-based model designed to provide git
+In this paper, we develop a tree-based model, \mt, designed to provide git
 users with a better explanation of the events occurring in a git
@@
 
-We implement a tool, \tool, which allows a user to filter the commits
-with a simple search interface. Once the user finds the commit that they
-are interested in, they are provided simple aggregated tables showing
-the files that were modified, the authors that made the modifications,
-among other information. They are also provided with three
-visualizations, a list tree, a pack tree, and a Reingold-Tilford tree
-view. The list tree enables easy searching for textual features, while
-the pack tree is good for visualizing the structure of large merge
-trees, and the Reingold-Tilford tree is better at visualizing the
-structure of small merge trees.
+We implement a tool, \tool, using the \mt model, which allows a user to
+filter the commits with a simple search interface. Once the user finds
+the commit that they are interested in, they are provided simple
+aggregated tables showing the files that were modified, the authors that
+made the modifications, among other information. They are also provided
+with three visualizations, a list tree, a pack tree, and a
+Reingold-Tilford tree view. The list tree enables easy searching for
+textual features, while the pack tree is good for visualizing the
+structure of large \mt{s}, and the Reingold-Tilford tree is better
+at visualizing the structure of small \mt{s}.
 
@@
 aspects of each they preferred. We found that the participants in our
-study were able to correctly identify the correct answer more quickly
-using the \mt model than they were able to with the DAG.
+study were able to identify the correct answer more quickly and
+accuractly using the \mt visualization than they were able to with
+visualizations of the DAG.
 
-Given that the model provides users with a better understanding of
-events in the repository, we describe some issues with identifying the
-master branch of the repository and a more generalized approach to
-constructing merge trees. We implement the generalized approach as a
-bitbucket plugin to verify that it is able to quickly produce results
-online, unlike in the implementation of \tool.
+Given that visualization of the model provides users with a better
+understanding of events in the repository, we describe some issues with
+identifying the master branch of the repository and a more generalized
+approach to constructing \mt{s}. We implement the generalized approach
+as a Bitbucket plugin to verify that it is able to quickly produce
+results online, unlike in the implementation of \tool.
 
-% TODO: You can tell that I am tired. This conclusion will need to be re-worked.
+Participants in our study found them to be more enjoyable for
+summarization tasks than visualizations of the DAG.\@ While there are
+some issues with identifying the master branch in a robust manner, \mt
+visualizations are an effective means of providing a clear conceptual
+understanding and simple summarization of the events being merged into a
+the master branch of a repository.

### 2017-07-14 paper/abstract.tex
many changes, looks very good
editing abstract

@@
 % vim:set et sw=2 ts=4 tw=72:
-With an average of more than 900 top-level merges into the Linux kernel
-per release, many containing hundreds of commits and some containing
+
+% 654 -- containing hundreds
+% 44 -- containing thousands
+
+With an average of more than 900 merges into the Linux kernel per
+release, many containing hundreds of commits and some containing
 thousands, maintenance of older versions of the kernel becomes nearly
 impossible. Various commercial products, such as the Android platform,
-run older versions of the kernel. Due to security, performance, and
+run older versions of the kernel; due to security, performance, and
 changing hardware needs, maintainers must understand what changes
@@
 kernel the DAG becomes overwhelming very quickly. Furthermore, the DAG
-contains every ancestor of every commit, while maintainers are more
+contains every parents of every commit, while maintainers are more
 interested in how and when a commit arrives to the official Linux
@@
 
-In this paper, we propose the merge-tree, a simplified transformation of
-the DAG of the Linux git repository that shows the way in which commits
-are merged into the master branch of Linux. Using the merge-tree, we
-build \tool, a tool that is designed to allow users to explore how
-commits are merged into the Linux kernel.
+This paper make three contributions; a conversion from DAG to \mt, an
+implemenation of a tool built on the \mt model, and a user study to
+evaluate and validate the implementation and model.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-14 paper/introduction.tex
editing introduction

--- 2017-07-08 paper/introduction.tex
@@
 changes are merged, how each branch is interacting, and the point where
-a branch forks from the master branch. In large modular projects, like
+a branch forks from another branch. In large modular projects, like
 the Linux kernel, the DAG becomes a mess of merges and commits
@@
 into another branch, including the master branch. To complicate the
-problem, the DAG only has references to the ancestors of a commit (a
+problem, the DAG only has references to the parents of a commit (a
 model necessary for the operation of Git), but maintainers would prefer
@@
 Tracing a path that any commit followed to the master repository would
-imply that for any given merge, it would be possible to know which
-commits were merged. A user could inspect the commits that arrived into
+imply that for any given commit, we know the next merge in the direction
+of the master branch. A user could inspect the commits that arrived into
 the master branch within a given time-frame by checking which commits
@@
 inspection and visualization of merges in the Linux project using the
-\mt modle; finally, we validate the \mt model and the implementation of
-the visulization through a controlled user study. We further discuss the
+\mt model; third, we validate the \mt model and the implementation of
+the visualization through a controlled user study. We further discuss the
 issues of generalizing the model to other repositories, but present an
@@
 its respective merge-tree, the files edited, the modules edited, and the
-commit message. \tool allows users to apply various filters, including
-the release version, along with a keyword or phrase from the log
-preview, the name of the author, or the commit hash. The user can
-request all merges made by Linus that contain a commit or inner merge
-that matches the search query, or just the commits and merges that match
-the query.
+commit message. \tool provides a search engine that filters commits by
+commit hash, terms from the commit log, the author, filenames, and other
+information. The results are then grouped together by the root of the
+corresponding \mt.

### 2017-07-14 paper/background.tex
editing background

@@
 The repository events in the DAG are immutable; once a repository event
-is created, it can't be changed. \dmg{the next sentences should be repository events, not commits}
-Git allows operations to alter commits
-and re-order them, but this will create a new commit with a new commit
-hash.  This property makes a repository event unable to record the
-traversal of merges into the master branch of the repository. A commit
-cannot have references in the direction of the merge that merges the
-commit, as this reference will change, requiring the commit to be
-changed. \dmg{BE CONSISTENT, Git or git?} git provides
-the command \verb|git log --children| which traverses
-the DAG and prints the inverted edges relationship between events.
+is created, it can't be changed. \dmg{the next sentences should be
+  repository events, not commits} Git allows operations to alter
+repository events and re-order them, but this will create a new event
+with a new commit hash. This property makes a repository event unable to
+record the traversal of merges into the master branch of the repository.
+Introducing an additional field to track the path to the master branch
+would create a circular dependency between the parent and child. Any
+updates to the DAG would modify the commit hashes for all repository
+events in the repository. To alleviate this, git provides the command
+\verb|git log --children| which traverses the DAG and prints the
+inverted edges relationship between events.
 
-Under perfect conditions, visualizations of the DAG should show how a
+Under perfect conditions, visualizations of the DAG show how a
 repository changes over time; where logical branches are being made,
 where merges are happening, and who is making the changes. Most
-repositories are relatively simple, with
-relatively few, small, branches with a clear, well defined
-master branch. However, in large, active repositories it becomes very difficult for developers to use the DAG
-to identify
-the master branch, and as a consequence, where branches start and are merged back in their
-way to master.
+repositories are relatively simple, with relatively few, small, branches
+with a clear, well defined master branch. However, in large, active
+repositories it becomes very difficult for developers to use the DAG to
+identify the master branch, and as a consequence, where branches start
+and are merged on the path to the master branch. \evan{They do show in
+  perfect conditions. If you don't have many branches, and you don't
+  have any foxtrots, the first branch on the left of the DAG
+  visualization WILL be the master branch}
 

### 2017-07-14 paper/model.tex
editing model section

@@
 \section{Merge-Tree model}
-
 \label{sec:mergetree}
 
-The \mt model abstracts the DAG of the repository into a set of \mt{}s. Each \mt is rooted at a merge into the master
-branch. The children of this tree are the commits and merges that are merged together into master by this merge. A \mt is build
-recursively with every merge in the \mt merging a sub-\mt. Any merges that a commit must pass through to reach the
-master branch become an inner node of the tree. 
-Effectively, a \mt is a tree that
-shows the path that commits follow in their way to the master branch of a repository.
-The \mt{}s also help understand how commits were grouped together in order
-to be integrated. In this model, not only has the DAG been inverted (and simplified), but
-the entire notion of parent-child relationship has been reversed.
+The \mt model abstracts the DAG of the repository into a set of \mt{s}.
+Each \mt is rooted at a merge into the master branch. The leaves of this
+tree are the commits and the merges are the inner nodes. A \mt is built
+recursively with every merge in the \mt merging a sub-\mt. Any merges
+that a commit must pass through to reach the master branch become an
+inner node of the tree. Effectively, a \mt is a tree that shows the path
+that commits follow in their way to the master branch of a repository.
+The \mt{s} also help understand how commits were grouped together in
+order to be integrated. In this model, not only has the DAG been
+inverted (and simplified), but the entire notion of parent-child
+relationship has been reversed. Due to this property, the terminology
+will be different depending on the model we are referring to; when we
+are referring to the \mt, the parent is the next node toward the merge
+into the master branch, or the root of the tree. When referring to the
+DAG, the parent relationship is in the opposite direction, from the root
+toward the branch-point. In addition to identifying the path that a
+commit took to being merged, we are able to aggregate the commit
+metadata at merges as we retain the parent-child and child-parent
+relationship for each event and know which commits belong to the merge.
 
-Please note that when we are referring to the
-\mt, the parent is the next node toward the merge into the master
-branch, or the root of the tree. When referring to the DAG, the parent
-relationship is in the opposite direction, from the root toward the
-branch-point. In addition to identifying the path that a commit took to
-being merged, we are able to aggregate the commit metadata at merges as
-we retain the parent-child and child-parent relationship for each event.
+To illustrate this model we will use a small example: assume the commits
+represented in Figure~\ref{fig:repoEvents} show the sequence of events
+in a repository. The sequence starts with the initial commit in the
+master branch of the master repository at time $t_0$. Repository event 1
+is a commit, which gets forked into a separate repository, \textit{Repo
+  A}, where another commit is made, event 2. Event 5 is a merge event,
+merging event 2, 3, and 4 into \textit{Repo A}. Event 5 is branched
+from, commit 6 happens in the new branch, while commit 7 is added
+simultaneously to the original branch in \textit{Repo A}. Events 11 and
+12 are both merge events, merging changes made in \textit{Repo A} into
+the master branch of the master repository. As every repository is a
+first-class repository, including local copies and forks, Git does not
+distinguish between forked repositories and branches, and in neither
+case does it explicitly record where a commit was made. In this case,
+commits are performed in various repositories and branches. The DAG
+representation of these events is shown in Figure~\ref{fig:repoDAG}.
 
-To illustrate this model we will use a small example:
-assume the commits represented in
-Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
-The sequence starts with the initiall commit in the master branch of the
-master repository at time $t_0$. Repository event 1 is a commit, which
-gets forked into a separate repository, \textit{Repo A}, where another
-commit is made, event 2. Event 5 is a merge event, merging event 2, 3,
-and 4 into \textit{Repo A}. Event 5 is branched from, event 6 happens in
-the new branch, while event 7 is added simultaneously to the original
-branch in \textit{Repo A}. Events 11 and 12 are both merge events,
-merging changes made in \textit{Repo A} into the master branch of the
-master repository.
-
-As every repository is a first-class repository, including local copies
-and forks, Git does not distinguish between forked repositories and
-branches, and in neither case does it explicitly record where a commit
-was made. In this case, commits are performed in various repositories
-and branches. The DAG representation of these events is shown in
-Figure~\ref{fig:repoDAG}.
-
- Notice that the DAG loses information about
-the master branch and the repository that the master branch is part of.
-The merge-tree view of this DAG is visible in Figure~\ref{fig:repoTree}.
-Note that the direction of the edges of the DAG have been inverted,
-instead of pointing from the child to the parent, it points from the
-parent to its children, forming a path to the master branch. Also
-note that the DAG has been simplified, showing only a single edge on the
-path to master for any commit.
+Notice that the DAG loses information about the master branch and the
+repository that the master branch is part of. The \mt view of
+this DAG is visible in Figure~\ref{fig:repoTree}. Note that the
+direction of the edges of the DAG have been inverted, instead of
+pointing from the child to the parent, it points from the parent to its
+children, forming a path to the master branch. Also note that the DAG
+has been simplified, showing only a single edge on the path to master
+for any commit.
 
@@
   \centering
-  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
-    \node[state, draw=chartblue] (1) {1};
-    \node[state, draw=chartyellow, above right= of 1] (2) {2};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
-    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
-    \node[state, draw=chartyellow, above right=of 4] (5) {5};
-    \node[state, draw=chartred, above right=of 5] (6) {6};
-    \node[state, draw=chartyellow, right=of 5] (7) {7};
-    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
-    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
-    \node[state, draw=chartyellow, right=of 9] (10) {10};
-    \node[state, draw=chartblue, below right=of 9] (11) {11};
-    \node[state, draw=chartblue, below right=of 10] (12) {12};
+  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black, black}]
+    \node[state, black] (1) {1};
+    \node[state, black, above right= of 1] (2) {2};
+    \node[state, black, above right= 2cm and 1cm of 2] (3) {3};
+    \node[state, black, right= 2cm of 1] (4) {4};
+    \node[state, black, above right=of 4] (5) {5};
+    \node[state, black, above right=of 5] (6) {6};
+    \node[state, black, right=of 5] (7) {7};
+    \node[state, black, above right= 2cm and 1cm of 7](8) {8};
+    \node[state, black, right= 2cm of 7] (9) {9};
+    \node[state, black, right=of 9] (10) {10};
+    \node[state, black, below right=of 9] (11) {11};
+    \node[state, black, below right=of 10] (12) {12};
 
-    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
-    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
-    \draw (10) edge[chartyellow, -stealth] (9);
-    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
-              edge[chartyellow, -stealth] (7);
-    \draw (8) edge[chartmagenta, -stealth] (7);
-    \draw (7) edge[chartyellow, -stealth] (5);
-    \draw (6) edge[chartred, -stealth] (5);
-    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
-              edge[chartyellow, -stealth] (4);
+    \draw (12) edge[-stealth] (11) edge[-stealth] (10);
+    \draw (11) edge[-stealth] (4) edge[-stealth] (9);
+    \draw (10) edge[-stealth] (9);
+    \draw (9) edge[-stealth] (8) edge[-stealth] (6)
+              edge[-stealth] (7);
+    \draw (8) edge[-stealth] (7);
+    \draw (7) edge[-stealth] (5);
+    \draw (6) edge[-stealth] (5);
+    \draw (5) edge[-stealth] (3) edge[-stealth] (2)
+              edge[-stealth] (4);
     \draw (4) edge[-stealth] (1);
-    \draw (3) edge[chartmagenta, -stealth] (2);
-    \draw (2) edge[chartyellow, -stealth] (1);
+    \draw (3) edge[-stealth] (2);
+    \draw (2) edge[-stealth] (1);
   \end{tikzpicture}
@@
 
-\subsection{Computing the merge-tree of the DAG of Linux}
+\subsection{Computing the \mt of the DAG of Linux}
 
-Computing the merge-tree from a DAG for any repository may not be
+Computing the \mt from a DAG for any repository may not be
 possible; however, certain features of the development process of Linux
-make it feasible to compute the merge-tree for the Linux repository.
+make it feasible to compute the \mt for the Linux repository.
 First, the master branch of Linux is maintained by Linus Torvalds, and
@@
 is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
-identifies the commits made directly to the master branch; whereafter it
+identifies the commits made directly to the master branch, whereafter it
 recursively determines the shortest path (in terms of time), using the
@@
 
-\dmg{I think you abbreviated this section too much. It will be helpful to describe the algorithm
-at a very high level,as in the vissoft paper.}
-
 \begin{algorithm}
-        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
+        \caption{Computing the \mt of Linux Git's DAG}\label{fig:alg}
         \begin{algorithmic}
@@
 example. This summary contains a list of the first 20 non-merge commits
-in the merge, including their one-line log description, the full logs of
-the merge commits that merge this subset, and the total number of
-non-merge commits in the merge.
+in the merge, including their one-line log description, and the total
+number of non-merge commits in the merge.
 
@@
 
-We used this information to evaluate the accuracy of the merge-tree
-model extracted from the DAG\@. The method we followed started with the
-extraction of the commit history up to July 20, 2016. We computed the
-merge tree of every commit until then. Since Linus Torvalds mostly does
-merging directly into master, we assumed that every merge by him is the
-root of a merge-tree. As described above, the log of a merge-commit
-usually contains the number of commits in the merge the first 20
-summaries of commits being merged. We extracted merges by Linus Torvalds
-using the command \mycode{log --merges --author='Torvalds} and compared
-the number of commits according to the log with the number of commits in
-the  merge-tree rooted in this commit. We also used the summaries of the
-commits found in the merge (not necessarily all---see above) to make
-sure those commits were in their corresponding merge tree. For example,
-for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
-merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
-one-line summaries corresponds to commits in that merge-tree. We also
-inspected those with differences to make sure they were true errors.
-The results can be summarized as follows:
+We used this information to evaluate the accuracy of the \mt model
+extracted from the DAG\@. The method we followed started with the
+extraction of the commit history up to July 20, 2016 form the Linux
+repository. We computed the merge tree of every commit until that date.
+Since Linus Torvalds mostly does merging directly into the master
+branch, we assumed that every merge by Linus is the root of a \mt, later
+detecting merges that do not merge into the master branch and removing
+them from the set of root merges. As described above, the log of a
+merge-commit usually contains the number of commits in the merge the
+first 20 summaries of commits being merged. We extracted merges by Linus
+Torvalds using the command \verb|git log --merges --author='Torvalds'|
+and compared the number of commits according to the log with the number
+of commits in the \mt rooted in this commit. We also used the summaries
+of the commits found in the merge (not necessarily all---see above) to
+make sure those commits were in their corresponding merge tree. For
+example, for the merge in Figure~\ref{fig:sampleMerge} we would expect
+that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and
+the one-line summaries corresponds to commits in that \mt. We also
+inspected those with differences to make sure they were true errors. The
+results can be summarized as follows:
 
@@
     hence making it impossible to verify; during this period, however,
-    we correctly identified the merges by Linus into master.
+    we correctly identified the merges by Linus into the master branch.
 
@@
     presence of a \foxtrot commit (\mycode{c436688\ldots}),
-    which confounded the true master branch of a repository.
+    which confounded the true master branch.
 

### 2017-07-14 paper/lineval.tex
editing visualizing section

@@
 
-\section{Visualizing the merge-tree of Linux}
+\section{Visualizing the \mt of Linux}
 
@@
 information, specifically focusing on merges. This is done by leveraging
-the merge-tree to inspect how commits are merged on the path to the
-master repository.
+the \mt to inspect how commits are merged on the path to the
+master branch of the repository.
 
@@
 
-\noindent \textbf{Use-case 1: top-to-bottom
-  approach}\label{sec:usecase1}\\ These are users that are maintaining a
-section of the kernel and would like to pick a merge (including all the
-commits that it merges) and merge it directly into their current
-repository. This is useful for reducing the amount of re-implementation
-work. For these users, it is important to have the ability to aggregate
-metadata about files and modules being effected by the merge. Also, it
-is important for these users to be able to navigate from the root of the
-merge-tree toward the leaves.
+\noindent \textbf{Use-case 1: top-to-bottom approach}
+\label{sec:usecase1}\\ These are users that are maintaining a section of
+the kernel and would like to pick a merge (including all commits that it
+merges) and merge it directly into their current repository. This is
+useful for reducing the amount of re-implementation work necessary for
+integration. For these users, it is important to have the ability to
+aggregate metadata about files and modules being effected by the merge.
+Also, it is important for these users to be able to navigate from the
+root of the \mt toward the leaves.
 
-\noindent \textbf{Use-case 2: bottom-to-top
-  approach}\label{sec:usecase2}\\ These are users that start with a
-known merge or commit and would like to see what other changes are being
-made in commits that are in the same merge, including knowing the
-merge-tree they belong to. This is useful to see what other commits are
-related to the current commit and how they get collated into merges that
-eventually end in the master branch. This is primarily for maintainers
-that need to perform some specific cherry picking of commits. We must
-provide these users a mechanism for navigating from a single commit
-toward the master branch, allowing them to see other commits that might
-be related to their original commit.
+\noindent \textbf{Use-case 2: bottom-to-top approach}
+\label{sec:usecase2}\\ These are users that start with a known merge or
+commit and would like to see what other changes are being made in
+commits that are in the same merge, including knowing the \mt
+they belong to. This is useful to see what other commits are related to
+the current commit and how they get joined into merges that eventually
+end in the master branch. This is primarily for maintainers that need to
+perform some specific cherry picking of commits. We must provide these
+users a mechanism for navigating from a single commit toward the master
+branch, allowing them to see other commits that might be related to
+their original commit.
 
@@
 
-In our visualizations, we leverage the merge-tree model described in
-Section~\ref{sec:mergetree}. In this model each commit is either already
-in the master branch, or is part of a tree which is rooted in the merge
-that merged it into the master branch.  Each commit, whether a merge or
-non-merge, has only one successor; the root of each tree has none as it
-was made by Linus Torvalds directly into the master branch. Non-merge
+In our visualizations, we leverage the \mt model described in
+Section~\ref{sec:mergetree}. In this model, commits are either on the
+master branch, or part of a tree which is rooted at a merge into the
+master branch. Each commit, whether a merge or non-merge, has only one
+parent, with the exception of the root which has no parents. Non-merge
 commits contain the metadata for the changes made. This metadata
@@
 being merged into the kernel, the date the commit was authored, the
-patch, and the commit log. Merges contain similar metadata,
-the author of the merge, the committer of the merge, if it has been rebased, the log, the commit date, and the author date,
-and potentially, changes necessary to address conflicts during the
-merge. The details of the model are outlined in \cite{German2015}.
+patch, and the commit log. Merges contain similar metadata, the author
+of the merge, the committer of the merge, if it has been rebased, the
+log, the commit date, and the author date, and potentially, changes
+necessary to address conflicts during the merge.
 

### 2017-07-14 paper/design.tex
editing design

@@
 We constructed a web-based tool called \tool in order to navigate and
-inspect the \mt model of the kernel. Creating the web-based tool enables
-users to use the system without having to install additional software or
-store a large database, which makes it more accessible, easily
-maintainable, and platform independent.
-
-To navigate and inspect the merge-tree view of the kernel we created a
-web-based tool called \tool. Creating a web-based tool enables users to
-use the system without having to install additional software or store a
-large database, making it more accessible, more easily maintainable, and
-platform independent. \tool uses the following mechanisms to reach our
-goals of better navigation and better explanation of the selected
-changes.
+inspect the \mt visualizations of the kernel. Creating the web-based
+tool enables users to use the system without having to install
+additional software or store a large database, which makes it more
+accessible, easily maintainable, and platform independent. \tool uses
+the following mechanisms to reach our goals of better navigation and
+better explanation of the selected changes.
 
@@
 
-Before presentation, the results are grouped by merge tree. Each
-merge tree group has a link to the root-node of the tree at the top,
-followed by a table of commits and merges within the tree that match the
-query, shown in Figure~\ref{fig:search_results}. The table includes the
-relevancy rank of the entry, the commit log message preview, author, the
-date the repository event was committed, and the date it was authored.
-The merge tree groups are ordered based on the mean of the relevancy
-scores of the relevant results in that tree.
+Before presentation, the results are grouped by \mt root. Each group
+has a link to the root-node of the tree at the top, followed by a table
+of commits and merges within the tree that match the query, shown in
+Figure~\ref{fig:search_results}. The table includes the relevancy rank
+of the entry, the commit log message preview, author, the date the
+repository event was committed, and the date it was authored. The merge
+tree groups are ordered based on the mean of the relevancy scores of the
+relevant results in that tree.
 
@@
 \tool uses seven tabs to show the information and visualizations for a
-selected commit or merge. The message tab shows the full commit log
-message. This does not include the diff, but given the commit hash, this
-information can be found directly from the repository. The files tab
-shows an aggregated table of all files that were modified in a merge. It
-includes metrics like the number of lines added, lines removed, total
-lines modified, and the delta, summed across all commits in the merge
-tree that modify this file. A details drop-down button allows a user to
-see exactly which commit makes the changes, as shown in
+selected repository event. The summarization tabs are; messages, files, modules,
+authors; the visualization tabs are; list tree, pack tree, and \rt tree.
+
+The message tab shows the full commit log message. This does not include
+the diff, but given the commit hash, this information can be found
+directly from the repository.
+
+The files tab shows an aggregated table of all files that were modified
+in a merge. It includes metrics like the number of lines added, lines
+removed, total lines modified, and the delta, summed across all commits
+in the merge tree that modify this file. A details drop-down button
+allows a user to see exactly which commit makes the changes, as shown in
 Figure~\ref{fig:linvis_files}. If the current repository event being
 viewed is a commit, the aggregate views will only show modifications
-made by the commit. The modules tab shows the modules modified in the
-merge tree. Like the files tab, the modules tab uses a table to show the
-name of the module, the number of commits that are in the merge tree
-that work with the module, and a details button to provide the links to
-those commits.
+made by the commit.
+
+The modules tab shows the modules modified in the merge tree. Like the
+files tab, the modules tab uses a table to show the name of the module,
+the number of commits that are in the merge tree that work with the
+module, and a details button to provide the links to those commits.
 
@@
 authorship information. It shows the sum of the number of lines added,
-removed, modified, and the delta within the \mt. It also shows
-the number of files that were modified by the author. The details tab is
-organized slightly differently. Instead of organizing these results by
+removed, modified, and the delta within the \mt. It also shows the
+number of files that were modified by the author. The details tab is
+organized slightly differently. Instead of organizing the results by
 commit, the details are organized by file, showing which files were
@@
 parent-child relationship. This visualization is text-based enabling
-fast navigation to commits using the built-in search fin most
+fast navigation to commits using the built-in search in most
 web-browsers. The list tree is rooted at the current repository event;
@@
 Pack trees\cite{Wang2006} are useful for displaying an overview of large
-data sets. The initial goals for the pack tree were for visualizing file
+data sets. The initial goal for the pack tree was for visualizing file
 systems, which are similar to git repositories in that they are
@@
 colored with shades of blue to indicate the depth in the tree. Darker
-shades of blue indicate that the merge is deeper in the tree.
+shades of blue indicate that the merge is deeper in the tree; the root
+will be the lightest shade.

### 2017-07-14 paper/relatedwork.tex
editing related work and adding code swarm

@@
 against another that is designed with these goals. To our knowledge, we
-know of no git repository visualzation tool that has these specific
+know of no git repository visualization tool that has these specific
 goals. This may be due to issues with finding the master branch of the
-repository which is a non-trivial task, or due to other factors.
-
-While we have found no tool with the express goal of showing how a
-commit is integrated, and providing a summarization of a merge, there
-has been a lot of work done in providing visualizations of various
-aspects of a repository.
+repository which is a non-trivial task, or due to other factors. While
+we have found no tool with the express goal of showing how a commit is
+integrated, and providing a summarization of a merge, there has been a
+lot of work done in providing visualizations of various aspects of a
+repository.
 
@@
 made, only the frequency that they are being made and who is making
-them.
+them. Codeswarm\cite{ogawa09} is similar to Gource, using an organic approach to
+visualizing events in the repository. Unlike Gource, which focuses on
+the files, Codeswarm focuses on developers and the number of commits
+they are making.
 
-% TODO: add code_swamp or something like that
-
-% TODO: Check this -- Look into source tree
-Git itself is shipped with summarization and visualization tools:
-the \verb|git log| command and Gitk, which is a graphical tool for the purpose of browsing
-the DAG of the repository.
-Many other tools are designed with a similar visual metaphore
-as what is presented in Gitk, including Git kraken, Github Desktop and
-web interfaces, the Git Lab web intereface, the Bitbucket web interface,
-and many others. The Gitk interface is built around the central DAG
-viewer. The DAG displays the repository events on their respective
-branches, the author, and the date that the event was authored. A user
-can select an event to view addition information about it, including the
-full commit log, the parents and children of the event, and the diff
-generated if modifications were made. In the case of the Linux kernel,
-merges are not resolved as merge conflicts in the master repository, but
-by the developers prior to merging. With this development model, no
+Git itself is shipped with summarization and visualization tools: the
+\verb|git log| command and Gitk, which is a graphical tool for the
+purpose of browsing the DAG of the repository. Many other tools are
+designed with a similar visual metaphore as what is presented in Gitk,
+including GitKraken, Github Desktop and web interfaces, the Git Lab web
+interface, the Bitbucket web interface, are a few examples. The Gitk
+interface is built around the central DAG viewer. The DAG displays the
+repository events on their respective branches, the author, and the date
+that the event was authored. A user can select an event to view addition
+information about it, including the full commit log, the parents and
+children of the event, and the diff generated if modifications were
+made. In the case of the Linux kernel, merges are not resolved as merge
+conflicts in the master repository, but by the developers prior to
+merging through rebase operations. With this development model, no
 merges will have modified any files, and will not show a diff. While the
-merge does not make changes directly, the effects of merging the
-commits persist, but Gitk is unable to provide an aggregated view of the
-files modified, the patches for those files, and the authors who
-contributed commits to that merge.
+merge does not make changes directly, the effects of merging the commits
+persist, but Gitk is unable to provide an aggregated view of the files
+modified, the patches for those files, and the authors who contributed
+commits to that merge.
 

### 2017-07-14 paper/study.tex
Editing first half of study

@@
 
-The purpose of the study is to quantitatively and qualitatively evaluate
-the effectiveness of \tool compared to the DAG-based visualizations
-found in Gitk or the git command line. The study is broken into three
-task-based sections; conceptual tasks, summarization tasks, and
-participant opinions. The study was performed during June of 2017
-%in Kingston ON,
-in a controlled environment running Ubuntu 14.04.
-Participants were allowed to use both Gitk and the git command line
-tools together, which we will refer to as the git tools or as Gitk. Then
-they were asked to perform the same tasks using \tool.
+We conducted a study to quantitatively and qualitatively evaluate the
+effectiveness of \tool compared to the DAG-based visualizations found in
+Gitk or the git command line. The study is broken into three task-based
+sections; conceptual tasks, summarization tasks, and participant
+opinions. The study was performed during June of 2017 in a controlled
+environment running Ubuntu 14.04. To evaluate the DAG-based
+visualizations, participants were allowed to use both Gitk and the git
+command-line tools together, which we will refer to as Gitk. \tool was
+used for evaluating \mt-based visualizations.
 
-\dmg{I thought you had randomized the order of tool use? also clarify that they were done with different commits to
-  avoid bias}
-
-% \evan{Do I need to say where the study took place? I see it showing up
-%   in other papers, but it makes it kind of obvious about the lab I'm
-%   talking about, doesn't it?}
-
-The primary goal of the study is to determine if \tool is more capable
-of providing the participants of our study with a better conceptual
-understanding of how a commit is integrated into the kernel, and a
-better summarization of various metrics about the files and authors
-involved with the integration. The conceptual questions are to determine
-if the DAG-based visualizations (gitk and git command line) provide enough information to determine how a
-commit, and any related commits, are merged into a project. \tool
-displays this information directly in the form of a tree, so we will not
-be performing a comparison on this section, only evaluating the results
-from Gitk. The summarization tasks compare \tool and Gitk, with the
-participants using one tool followed by the other on two commits from
-merge-trees of differeing sizes. In both of these sections, the
-participant is given a task that they are to find an answer to; we
-evaluate the time it took, if the answer was correct, and how far from
-the right answer they were. The user opinions provide us with a clear
-comparison between the tools from the point-of-view of the participant,
-providing us information about which aspects of each tool they prefered.
+The primary goal of the study is two fold; first determining if the
+DAG-based visualization is sufficient for conceptual understanding;
+second comparing \tool and Gitk to determine which is more capable of
+providing users with a summarization of various metrics involved with
+integrating a commit into the repository. The conceptual questions are
+to determine if the DAG-based visualizations (Gitk and git command line)
+provide enough information to determine how a commit, and any related
+commits, are merged into a project. \tool displays this information
+directly in the form of a tree, so we will not be performing a
+comparison on this section, only evaluating the results from Gitk. The
+summarization tasks compare \tool and Gitk, with the participants using
+one tool followed by the other on two commits from \mt{s} of
+differing sizes. In both of these sections, the participant is given a
+task that they are to find an answer to; we evaluate the time it took,
+if the answer was correct, and how far from the right answer they were.
+The user opinions provide us with a clear comparison between the tools
+from the point-of-view of the participant, providing us information
+about which aspects of each tool they preferred.
 
@@
 \label{sub:tasks}
-
-% \dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}
-
-\dmg{I don't like how you talk about the DAG comprehension. Few paragraphs above I replaced it
-  with the DAG-based tools, which might be a better description, since you are evaluating tool
-vs tool}
 
@@
 \textbf{The conceptual tasks} are designed to gain better insight on the
-comprehension of the DAG, and determine if the DAG is sufficient for
-understanding what other commits are related to a given commit, and how
-those commits are merged. The tasks are outlined in
-Table~\ref{tab:conceptual_tasks}. The tasks are performed in order
-using the git tools. T1 has the participant draw a diagram of the merge
-tree, which will be used to answer the other two questions. This enables
-us to visually see issues with comprehending the DAG visualizations.
-Participants are given 10 minutes to perform this task, which also is
-used to familiarize the participant with the git tools interface, if
-they are not already, and with the set of commits they will be working
-with. T2 and T3 are drawn directly from the diagram from T1, giving
-quantitative metrics to measure the comprehension. T2 asks the
-participant to verify the number of commits, and T3 asks for the number
-of merges. The three tasks are performed on both commits before
-continuing to the summarization tasks.
+comprehension of the visualization in Gitk, and determine if it is
+sufficient for understanding what other commits are related to a given
+commit, and how those commits are merged. The tasks are outlined in
+Table~\ref{tab:conceptual_tasks}. The tasks are performed in order using
+Gitk. T1 has the participant draw a diagram of the merge tree,
+which will be used to answer the other two questions. This enables us to
+see issues with comprehending the DAG visualization. Participants are
+given 10 minutes to perform this task, which also is used to familiarize
+the participant with the Gitk interface, if they are not already, and
+with the set of commits they will be working with. T2 and T3 are drawn
+directly from the diagram from T1, giving quantitative metrics to
+measure the comprehension. T2 asks the participant to verify the number
+of commits, and T3 asks for the number of merges. The three tasks are
+performed on both commits before continuing to the summarization tasks.
 
-\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
-are designed to compare the ability of the participants to summarize
-information about the two commits used in the study when using git tools\dmg{be consistent, either you say dag-based
-  tools, or git tools, either one is fine, but be consistent}
-versus \tool. The tasks are broken into four sets based on what is being
-summarized: merges, authorship, files, and modules. The order that the
-participants perform each task is randomized within each task set, and
-the order of the task sets are then randomized with the exception of the
-merges task set which will always come first. We measure and evaluate
-three metrics for each task: correctness, accuracy, and timing.
+\textbf{The summarization tasks}, listed in
+Table~\ref{tab:summarization_tasks}, are designed to compare the ability
+of the participants to summarize information about the two commits used
+in the study when using Gitk versus \tool. The tasks are broken into
+four sets based on what is being summarized: merges, authorship, files,
+and modules. The order that the participants perform each task is
+randomized within each task set, and the order of the task sets are then
+randomized with the exception of the merge task set which will always
+come first. The order that the tools are used is kept consistent for a
+participant, but randomized between participants. We measure and
+evaluate three metrics for each task: correctness, accuracy, and timing.
 Correctness measures whether the answer was correct or incorrect.
@@
 
-\evan{Is this better, or do you still want me to move the information
-  about the randomization to the bottom?}
-\dmg{I didn't see anything about choosing what commits to inspect here... but this section looks good}
-
 While the merge task set is part of the summarization task group, it
-works less with the summarization of the commits and more with
-comprehension of the models. The task is designed with two goals in
-mind; first we want to measure and define which commits and which merges
-the participant has defined to be in the merge tree; second to help
-solidify which commits and merges will be used in the merge tree.
+focuses less with summarization of the commits and more with
+comprehension of the model visualizations. The task is designed with two
+goals in mind; first we want to measure and define which commits and
+which merges the participant has defined to be in the merge tree; second
+to help solidify which commits and merges will be used in the merge
+tree.
 
@@
 answer the tasks that follow, specifically finding the merge into the
-master branch. T6 and T7 are related to the authorship at merges,
+master branch. T6 and T7 are related to the authorship of a merge,
 determining how many authors are involved and who made the most changes.
@@
 Linux repository that we noticed while looking at various commits. To
-find the modules, take the text up to the first colon in the git log
+find the module name, take the text up to the first colon in the git log
 preview. For the log preview \textit{``ALSA: kernel docs: fix
@@
 
-The accuracy metrics are measured slightly differently between tasks due
-to different requirements and structures in the answer. In task T4, both
-the order and the merges are important. We define the distance to be the
-edit distance with insertion, deletion, and transposition at unit cost.
-The accuracy is the minimum edit distance necessary to make the answer
+The correctness metric is simply whether the answer was correct or not;
+an accuracy of zero indicates a correct answer. The accuracy metrics are
+measured slightly differently between tasks due to different
+requirements and structures in the answer. In task T4, both the order
+and the merges are important. We define the distance to be the edit
+distance with insertion, deletion, and transposition at unit cost. The
+accuracy is the minimum edit distance necessary to make the answer
 correct. T5, T9, and T10 are concerned with the set of elements in the
@@
 or one. For this reason, it is acceptable to omit the accuracy metric
-from this task. The correctness metric is simply whether the answer was
-correct or not; an accuracy of zero indicates a correct answer.
+from this task.
 
@@
 
-We include two commits in the study to detect differences between merge
-tree sizes. The order that the commits are presented to each participant
-is randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
+We include two commits in the study to detect differences between \mt
+sizes. The order that the commits are presented to each participant is
+randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
 October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
@@
 not including the root, while 50\% of the trees contain at most seven
-nodes non-root nodes. 75\% of the trees contain at least 51 nodes, and
-the largest tree contains 7217 nodes.
+non-root nodes. 75\% of the trees contain at least 51 nodes, and the
+largest tree contains 7217 nodes.
 
 From this information, we selected one tree with a single commit, and
-one tree with seven other nodes to represent a small tree and a
+one tree with seven non-root nodes to represent a small tree and a
 medium-sized tree. A majority of the trees are flat, where all nodes
 merge directly into the root node. Of the 8031 trees containing at least
-seven nodes, only 593 contained at least a single internal merge node.
-Trivially, trees with a single node cannot have any internal nodes, and
-of the 624 trees with seven nodes, only 135 contained at least one inner
-node. To increase the complexity of the medium-sized tree, we randomly
-selected one of the 135 trees to represent the medium tree.
+seven non-root nodes, only 593 contained at least a single internal
+merge node. Trivially, trees with a single node cannot have any internal
+nodes, and of the 624 trees with seven non-root nodes, only 135
+contained at least one inner node. To increase the complexity of the
+medium-sized tree, we randomly selected one of the 135 trees to
+represent the medium tree.
 
@@
 random, and trivially selected the only commit in the tree. From the
-trees containing seven nodes, we selected one randomly, then randomly
-selected one of the commits in the tree. We selected commit
-\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
-a single node, which we will refer to as \comA, and commit
+trees containing seven non-root nodes, we selected one tree randomly,
+then randomly selected one of the commits in the tree. We selected
+commit \emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree
+containing a single node, which we will refer to as \comA, and commit
 \emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe} from the tree containing
@@
   \caption{The merge trees used in the user study,
-    containing one and seven commits, respectively.}
+    containing one and seven non-root nodes, respectively.}
   \label{fig:study_commits}
@@
 
-The study was conducted with 12 participants, all of whom were
-masters, PhD, or post-doc researchers in the field of software
-engineering. The participants had between 6 months and 10 years
-experience with git, with the median being 3.5 years. Most of the
-participants had additional experience with SVN and CVS.\@ One of the
-participants in the study worked as a release engineer, studying merge
-practices to determine the best way to merge branches while minimizing
-the number of merge conflicts, on SVN repositories. The participants
-worked with repositories ranging from around 10 commits up to 38000
-commits, with the median being 350 commits. Two of the participants had
-never collaborated with anyone in a repository, while the rest had some
-experience with repositories being modified by multiple people, with the
-most being 219. The median number of collaborators was four.
+The study was conducted with 12 participants, all of whom were masters,
+PhD, or post-doc researchers in the field of software engineering. The
+participants had between 6 months and 10 years experience with git, with
+the median being 3.5 years. Most participants had additional experience
+with SVN and CVS.\@ One of the participants in the study worked as a
+release engineer, studying merge practices to determine the best way to
+merge branches while minimizing the number of merge conflicts, on SVN
+repositories. The participants worked with repositories ranging from
+around 10 commits up to 38000 commits, with the median being 350
+commits. Two of the participants had never collaborated with anyone in a
+repository, while the rest had some experience with repositories being
+modified by multiple people, with the most being 219. The median number
+of collaborators was four. Participants had most experience with
+personal and academic repositories. Three of the twelve participants had
+experience with professional repositories.
 
-Five participants had only worked with personal and academic
-repositories. Two participants had worked with school and
-professional repositories, and two had only worked with professional
-repositories. One participant worked with personal, academic, and
-professional repositories, one participant had only worked with personal
-repositories, and one participant had worked with personal, academic,
-and school repositories.
-\dmg{you made it confusing: just say: how many had experience with
-  professional repositories and remove the above
-}
-\evan{Do you want more information about the participants?}
-
-All participants have had at least some experience with the idea of
-version control, branching in repositories, and git. The participants
-are all from the same lab, and each participant worked with both tools
-in the study, thus, keeping the sample populations identical for both
-tools, with some variation between participants in experience with
-repositories.
-
-\dmg{BIG QUESTION}
-for each task, did they do always git tools then linvis for the same commit?  if so, there is potential for learning
-bias: they learn the structure of the commits with git tools, so therefore it was easier for them to do
-them with linvis.
-\dmg{END QUESTION}
-
-\dmg{I'll stop here today}
+All participants have had at least some experience with version control,
+branching in repositories, and git. The participants are from the same
+lab, and each participant worked with both tools in the study, thus,
+keeping the sample populations identical for both tools, with some
+variation between participants in experience with repositories.
 
@@
 The conceptual tasks are run twice, once with \comA and once with \comB,
-in the same order for all participants using the git tools.
+in the same order for all participants using the Gitk.
 
@@
 four times, once with \comA and \tool, \comB and \tool, \comA and git
-tools, and \comB and git tools. The participant will start with one
+tools, and \comB and Gitk. The participant will start with one
 tool, and complete the tasks for both commits in that tool. The order of

### 2017-07-14 paper/discussion.tex
fixing capitalization

@@
 information, and provide a better understanding of the next merge
-involved in integrating this commit, but the DAG visualization in gitk
+involved in integrating this commit, but the DAG visualization in Gitk
 provides the full story of the commit instead of hiding it behind a
@@
 
-While git includes gitk and git log provides a visualization of the DAG
+While git includes Gitk and git log provides a visualization of the DAG
 with \verb|git log --graph|, many of the participants in our study were

### 2017-07-14 paper/introduction.tex
fixing capitalization

@@
 One major challenge with visualizing the arrival of commits to a
-repository is that Git does not store the date that a commit was merged
+repository is that git does not store the date that a commit was merged
 into another branch, including the master branch. To complicate the
 problem, the DAG only has references to the parents of a commit (a
-model necessary for the operation of Git), but maintainers would prefer
+model necessary for the operation of git), but maintainers would prefer
 knowing the path a commit followed to reach the master repository.

### 2017-07-14 paper/model.tex
fixing capitalization

@@
 the master branch of the master repository. As every repository is a
-first-class repository, including local copies and forks, Git does not
+first-class repository, including local copies and forks, git does not
 distinguish between forked repositories and branches, and in neither
@@
 \begin{algorithm}
-        \caption{Computing the \mt of Linux Git's DAG}\label{fig:alg}
+        \caption{Computing the \mt of Linux from the DAG}\label{fig:alg}
         \begin{algorithmic}

### 2017-07-14 paper/relatedwork.tex
fixing capitalization

@@
 with data from CVS repositories. Our goal is to provide information
-about Git repositories, specifically the Linux master repository. Fewer
-tools are available for generating visualizations and summaries of Git
-repositories, potentially due to the DAG model used by Git.
+about git repositories, specifically the Linux master repository. Fewer
+tools are available for generating visualizations and summaries of git
+repositories, potentially due to the DAG model used by git.
 
@@
 designed with a similar visual metaphore as what is presented in Gitk,
-including GitKraken, Github Desktop and web interfaces, the Git Lab web
+including GitKraken, Github Desktop and web interfaces, the GitLab web
 interface, the Bitbucket web interface, are a few examples. The Gitk

### 2017-07-14 paper/design.tex
make use of "Merge-Tree" consistent

@@
   \includegraphics[width=\linewidth]{figures/linvis/search_results.png}
-  \caption{A single merge tree in the search results of \tool, showing
+  \caption{A single \mt tree in the search results of \tool, showing
     the link to the root at the top and a table of the relevant commits
@@
 removed, total lines modified, and the delta, summed across all commits
-in the merge tree that modify this file. A details drop-down button
+in the \mt tree that modify this file. A details drop-down button
 allows a user to see exactly which commit makes the changes, as shown in
@@
 
-The modules tab shows the modules modified in the merge tree. Like the
+The modules tab shows the modules modified in the \mt. Like the
 files tab, the modules tab uses a table to show the name of the module,
-the number of commits that are in the merge tree that work with the
+the number of commits that are in the \mt tree that work with the
 module, and a details button to provide the links to those commits.
@@
 be shown. Conversely, if the selected event is the root, the entire
-merge tree will be shown.
+\mt tree will be shown.
 

### 2017-07-14 paper/discussion.tex
make use of "Merge-Tree" consistent

@@
 \label{sub:observations}
-
-% TODO: include info on merge tree sizes
 
@@
 commits was merged. We developed a small Bitbucket plugin to test the
-design and determine whether it was capable of producing merge trees,
+design and determine whether it was capable of producing \mt{s},
 shown in Figure~\ref{fig:b_plugin}.

### 2017-07-14 paper/introduction.tex
make use of "Merge-Tree" consistent

@@
 provide information about the location of any given commit or merge in
-its respective merge-tree, the files edited, the modules edited, and the
+its respective \mt, the files edited, the modules edited, and the
 commit message. \tool provides a search engine that filters commits by

### 2017-07-14 paper/model.tex
make use of "Merge-Tree" consistent

@@
   \end{tikzpicture}
-  \caption{Merge-tree view of the commits  represented in
+  \caption{\mt view of the commits  represented in
     Figure~\ref{fig:repoEvents} showing the path they followed to reach
@@
 extraction of the commit history up to July 20, 2016 form the Linux
-repository. We computed the merge tree of every commit until that date.
+repository. We computed the \mtof every commit until that date.
 Since Linus Torvalds mostly does merging directly into the master
@@
 of the commits found in the merge (not necessarily all---see above) to
-make sure those commits were in their corresponding merge tree. For
+make sure those commits were in their corresponding \mt For
 example, for the merge in Figure~\ref{fig:sampleMerge} we would expect
-that the merge tree rooted at \mycode{8cbd84f} contains 24 commits, and
+that the \mt rooted at \mycode{8cbd84f} contains 24 commits, and
 the one-line summaries corresponds to commits in that \mt. We also

### 2017-07-14 paper/study.tex
make use of "Merge-Tree" consistent

@@
     & T9   & Which file had the most changes?\\
-    Modules    & T10  & Which modules does this merge tree involve?\\
+    Modules    & T10  & Which modules does this \mt involve?\\
     \bottomrule
@@
 Table~\ref{tab:conceptual_tasks}. The tasks are performed in order using
-Gitk. T1 has the participant draw a diagram of the merge tree,
+Gitk. T1 has the participant draw a diagram of the \mt,
 which will be used to answer the other two questions. This enables us to
@@
 goals in mind; first we want to measure and define which commits and
-which merges the participant has defined to be in the merge tree; second
+which merges the participant has defined to be in the \mt; second
 to help solidify which commits and merges will be used in the merge
@@
 between \tool and Gitk. The Wilcoxon test is applied between commits to
-determine if the size of the merge-tree effects timing and accuracy. The
+determine if the size of the \mt effects timing and accuracy. The
 null hypotheses and alternative hypotheses are stated in
@@
 sizes. The order that the commits are presented to each participant is
-randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
+randomized. We analyzed the 15096 \mt{s} from April 16th 2005 to
 October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
@@
   \end{tabular}
-  \caption{The merge trees used in the user study,
+  \caption{The \mt{s} used in the user study,
     containing one and seven non-root nodes, respectively.}
@@
 
-Table~\ref{tab:cross_commit_results} shows whether the merge-tree size
+Table~\ref{tab:cross_commit_results} shows whether the \mtsize
 has an effect on the three studied metrics. An interesting observation
 from the figure is that the correctness for T5 and T9 are effected by
-the size of the merge-tree, but the accuracy is not. The correctness
+the size of the \mt, but the accuracy is not. The correctness
 results for T5 and T9 are shown in Figure~\ref{fig:correctness_t5_t9},
@@
     While there is a difference in correctness between different-sized
-    merge-trees, \tool improves correctness of determining the other
+    \mt{s}, \tool improves correctness of determining the other
     commits related to a commit in trees of varying sizes. The

### 2017-07-14 paper/study.tex
Fixing the "commit sizes" sentence

@@
 
-We include two commits in the study to detect differences between \mt
-sizes. The order that the commits are presented to each participant is
-randomized. We analyzed the 15096 \mt{s} from April 16th 2005 to
-October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
-and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
-not including the root, while 50\% of the trees contain at most seven
-non-root nodes. 75\% of the trees contain at least 51 nodes, and the
-largest tree contains 7217 nodes.
+We include two commits in the study to detect differences in the
+correctness, accuracy, and timing between \mt sizes. The order that the
+commits are presented to each participant is randomized. We analyzed the
+15096 \mt{s} from April 16th 2005 to October 14th 2014, which
+corresponds to the releases Linux 2.6.12-rc3 and Linux 3.17-rc1. 25\% of
+the trees contain at most a single commit not including the root, while
+50\% of the trees contain at most seven non-root nodes. 75\% of the
+trees contain at least 51 nodes, and the largest tree contains 7217
+nodes.
 
@@
 
-Table~\ref{tab:cross_commit_results} shows whether the \mtsize
+Table~\ref{tab:cross_commit_results} shows whether the \mt size
 has an effect on the three studied metrics. An interesting observation

### 2017-07-14 paper/conclusion.tex
more comments

@@
+\section{Conclusion}
+\label{sec:conclusion}
+
+\dmg{in this paper we describe, not develop}
+
 In this paper, we develop a tree-based model, \mt, designed to provide git
@@
 We implement a tool, \tool, using the \mt model, which allows a user to
-filter the commits with a simple search interface. Once the user finds
+filter the commits with a simple search interface. Once the user finds \dmg{user => they or he/she or users => they, tense issue?}
 the commit that they are interested in, they are provided simple
 aggregated tables showing the files that were modified, the authors that
-made the modifications, among other information. They are also provided
+made the modifications, among other information. \dmg{they is very ambiguous: are we talking authors, users, tables,
+  files? be precise}
+They are also provided
 with three visualizations, a list tree, a pack tree, and a
@@
 
+\dmg{not necessary: useful, and not only the model, but the implementation}
 To ensure that our model is necessary, we evaluate the ability of users
-to perform simple conceptual tasks, finding how a commit was merged into
+to perform simple conceptual tasks \dmg{using linvis and comparing it with ... be precise}
+, finding how a commit was merged into
 the master branch of the Linux repository, and determining other commits
 that were merged with the given merge. Finding that the participants in
-our study were unable to perform this task, we compared the ability of
+our study were unable to perform this task \dmg{using gitk?, be precise, but you didn't ask them to compare based on
+  their unability to perform the task, they were ask to do it nonetheless}, we compared the ability of
 the participants to summarize key pieces of information aggregated in
 these groups of commits. We asked for information about what files were
-modified, the authors making the changes, among other tasks. Finally, we
+modified, the authors making the changes, among other tasks\dmg{among other tasks? tasks are T1...Tn. Don't overload
+  uses of a word}. Finally, we
 asked the participants about their preference in tools, and which
@@
 identifying the master branch of the repository and a more generalized
-approach to constructing \mt{s}. We implement the generalized approach
+approach to constructing \mt{s}. We implement \dmg{this should be past tense} the generalized approach
 as a Bitbucket plugin to verify that it is able to quickly produce
-results online, unlike in the implementation of \tool.
+results online, unlike in the implementation of \tool.\dmg{the most important aspect is nto that it can quickly produce
+  results, but that it can be applied to any repo in bitbucket}
 
-Participants in our study found them to be more enjoyable for
+Participants in our study found them \dmg{who/what is them?} to be more enjoyable for
 summarization tasks than visualizations of the DAG.\@ While there are
-some issues with identifying the master branch in a robust manner, \mt
+some issues \dmg{with what tool? linvis?, \mt should be in the first phrase, not the ending phrase} with identifying the master branch in a robust manner, \mt
 visualizations are an effective means of providing a clear conceptual
@@
 the master branch of a repository.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-14 paper/discussion.tex
more comments

@@
 
-The results show conclusive results; the DAG visualization is unable to
+\section{Discussion}
+\label{sec:discussion}
+
+The results show conclusive results; the DAG visualization \dmg{provided by gitk...} is unable to \dmg{this is a very
+  long sentence that does not read very well}
 adequately provide users with an explanation of the changes being made
-in the repository, the DAG visualizatoin is not easily understood, and
+in the repository, the DAG visualization is not easily understood, and
 is unable to provide summarizations of the changes being made at a
-merge. Meanwhile, the \mt visualization is better able to provide an
+merge. Meanwhile, the \mt visualization \dmg{of linvis...} is better able to provide an
 explanation of how a commit is integrated, along with better
@@
 During the course of the study, we noticed an issue that was consistent
-among the participants when using Gitk; the most difficult challenge was
+among the participants when using Gitk; \dmg{wrong use of semicolon} the most difficult challenge was
 identifying the master branch of the repository. Many participants would
 identify the tag as being the ``master branch'', but a tag is not a
-branch. The visualization of the DAG does not provide any notion of
+branch (it simply points to a given commit). The visualization of the DAG does not provide any notion of
 which branch being shown is the master branch. The default assumption is
 that the first branch on the left is the master branch, but this is not
-a good assumption. Furthermore, the visualization in Gitk is not
+always correct. Furthermore, the visualization in Gitk is not
 consistent, the ordering of the branches changes between runs, so
 identifying the branch once will not guarantee that it will be
-identifiable after restarting Gitk.
+identifiable after restarting Gitk.\dmg{i didn't know the ordering is non-deterministic, are you sure? make sure you are
+sure, sure? :)}
 
@@
 visualizations of the DAG, the results of the study would likely be very
-different. This would certainly be the case for the single-commit tree;
+different\dmg{be explicit eg.: that there was very little difference, if any, between the use of both tools}. This would
+certainly be the case for the single-commit tree; \dmg{another wrong use, i want you to read the description of colon vs
+semicolon and scan the paper for them to make sure they are used correctly}
 it is trivial to summarize information about a single item, but
@@
 experience with git. This subsection contains improvements made by this
-participant during the study and afterward.
+participant during the study and afterward.\dmg{he made improvements? to what? Isn't this section about his views of
+  what is needed?}
 
@@
 was thinking as they were writing the changes. The current \mt model
-does not order the commits, simply placing them randomly. This is the
-primary reason behind this participant's request to use both
+does not order the commits, simply placing them randomly. \dmg{this surprised me. we order the commits in the tree, you
+  know which is before which, so this is more an issue of you not displaying the data than we not having it} This is the
+primary reason behind this participant's request to use both \dmg{tools}
 simultaneously. \tool is able to help with the aggregation of the
@@
 The answers provided by the participants to earlier tasks were not
-taken into account for determining the accuracy. If the participant was
+taken into account for determining the accuracy\dmg{accuracy of what? the stattement is not clear to me. Oh, i think you
+mean, if they get an answer wrong it might impact the results of some following tasks; you can be more precise in the description}. If the participant was
 unable to determine the correct commits for the tree within the
@@
 We were only able to work with 12 participants. We apply various
-statistical tests to ensure that our findings are reasonable.
+statistical tests to ensure that our findings are statistically significant.
 
@@
 
+\dmg{add a couple of sentences on the fact that these were students, some with industrial experience. People who are
+  professinals might do better}
+
 % A possible
+
 % reason for this may be that unless a repository reaches a sufficient
@@
 % the repository using \tool.
+
+\dmg{the following should be a section on its own, because it is no longer part of the study, also
+use a better title: generalization of the creation of the merge-tree?}
 
@@
 
+\dmg{you don't generalize an alg into a tool; you implement the algorithm in a tool}
 If this algorithm is to be generalized into a dynamic, web-based, tool,
@@
 %%% mode: latex
-%%% TeX-master: "userstudy"
+%%% TeX-master: "lineval.tex"
 %%% End:

### 2017-07-14 paper/study.tex
more comments

@@
   \centering
-  \caption{Cross-Commit Results\evan{Should I group the results by task
-      or by metric? It's by metric right now}}
+  \caption{Cross-Commit Results\dmg{Be more descriptive in the captions.People read the papers via tables and
+      pictures. What does the table say? e.g. as it can be seen.... tell a story with the picture, instead of asking the
+    reader to make up the story... which might be different than the one you want them to have}}
   \label{tab:cross_commit_results}
@@
   \end{tabular}
-  \caption{Unaggregated Correctness results for T5 and T9 \evan{Split
-      this? Is it clear that between Commit 2 and Commit 1 is where the
-    separation of the plots is?}}
+  \caption{Unaggregated Correctness results for T5 and T9.}
   \label{fig:correctness_t5_t9}
@@
 The full results are as follows;
+
+\dmg{you wrongly use ``semicolon'', you mean ``colon''. Same below...}
 
@@
 
+\dmg{Emphasize: To summarize...}
+
 \tool is able to assist users with summarizing various aspects of a
@@
 Among the 12 participants, there was nearly unanimous agreement that for
-conceptual understanding and summarization tasks; \tool was easier to
-work with than Gitk. The participants cited the ability to abstract
+conceptual understanding and summarization tasks; \dmg{this is another wrong use of semicolon}\tool was easier to
+work with \dmg{avoid idiomatic expressions: what does ``work with'' mean? use a precise verb. English is a rich language.} than Gitk. The participants cited the ability to abstract
 information about the merge from the clean summarization tables and
-simple visualizations. Three participants suggested that someone with a
-professional understanding of Gitk and the git command-line may be to
+simple visualizations \dmg{be precise: from Linvis' clean summarization tables and ... }. Three participants suggested that someone with a
+professional understanding of Gitk and the git command-line may be able to
 extract a conceptual understanding from the DAG visualization and

### 2017-07-15 paper/conclusion.tex
Fixing the conclusion

--- 2017-07-14 paper/conclusion.tex
@@
-In this paper, we develop a tree-based model, \mt, designed to provide git
+\section{Conclusion}
+\label{sec:conclusion}
+
+In this paper, we describe a tree-based model, \mt, designed to provide git
 users with a better explanation of the events occurring in a git
@@
 We implement a tool, \tool, using the \mt model, which allows a user to
-filter the commits with a simple search interface. Once the user finds
-the commit that they are interested in, they are provided simple
-aggregated tables showing the files that were modified, the authors that
-made the modifications, among other information. They are also provided
-with three visualizations, a list tree, a pack tree, and a
-Reingold-Tilford tree view. The list tree enables easy searching for
-textual features, while the pack tree is good for visualizing the
-structure of large \mt{s}, and the Reingold-Tilford tree is better
-at visualizing the structure of small \mt{s}.
+filter the commits with a simple search interface. Once the desired
+commit is found, \tool provides tabs for simple aggregated tables
+showing the files that were modified, the authors that made the
+modifications, among other information. \tool also provides three
+visualizations, list tree, pack tree, and a \rt tree view. The list tree
+enables easy searching for textual features, while the pack tree is good
+for visualizing the structure of large \mt{s}, and the \rt tree is
+better at visualizing the structure of small \mt{s}.
 
-To ensure that our model is necessary, we evaluate the ability of users
-to perform simple conceptual tasks, finding how a commit was merged into
-the master branch of the Linux repository, and determining other commits
-that were merged with the given merge. Finding that the participants in
-our study were unable to perform this task, we compared the ability of
-the participants to summarize key pieces of information aggregated in
-these groups of commits. We asked for information about what files were
-modified, the authors making the changes, among other tasks. Finally, we
-asked the participants about their preference in tools, and which
-aspects of each they preferred. We found that the participants in our
-study were able to identify the correct answer more quickly and
-accuractly using the \mt visualization than they were able to with
-visualizations of the DAG.
+To ensure that our model and implementation are useful, we evaluated the
+tool with a three-part user study. We tested the conceptual understanding
+using the Gitk, ability to summarize merges into the master branch, and
+ask for the opinions of the participants. We found that the participants
+were unable to determine the commits and merges that are related to the
+commits we tested with and build a conceptual understanding from the DAG
+visualization. The correctness, accuracy, and timing of summarization
+tasks were improved using \tool compared to Gitk. The participants
+generally enjoyed the clean summarizations and simple visualizations
+provided by the \mt-based visualizations in \tool.
 
@@
 identifying the master branch of the repository and a more generalized
-approach to constructing \mt{s}. We implement the generalized approach
-as a Bitbucket plugin to verify that it is able to quickly produce
-results online, unlike in the implementation of \tool.
+approach to constructing \mt{s}. We implemented the generalized approach
+as a Bitbucket plugin to verify that it is able to produce
+visualizations for arbitrary bitbucket repositories. We found that it
+was able to produce visualizations for many repositories, but was unable
+to produce the visualizations of others due to rate-limiting. An
+interesting side-effect of pruning the DAG is that some \mt{s} in the
+repository can be visualized, while other trees are too large to
+download the commit information for. The original algorithm design would
+be unable to produce a visualization for any of the \mt{s}.
 
-Participants in our study found them to be more enjoyable for
-summarization tasks than visualizations of the DAG.\@ While there are
-some issues with identifying the master branch in a robust manner, \mt
-visualizations are an effective means of providing a clear conceptual
-understanding and simple summarization of the events being merged into a
-the master branch of a repository.
+\mt{s} are a novel means of building a model which can be visualized and
+summarized in an effective way. Participants in our study found
+visualizations of the \mt to be more enjoyable for summarization tasks
+than the visualizations of the DAG\@. While there are some
+implementation issues with \tool, \mt visualizations are an effective
+means of providing a clear conceptual understanding and simple
+summarization of the events being merged into the master branch of a
+repository.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "lineval.tex"
+%%% End:

### 2017-07-15 paper/background.tex
adding background fixes

@@
 the current branch is replaced by a parent from a branch being merged
-into the current branch, thus confounding the branch relationships.\dmg{this would benefit from a diagram, since
-most people will have no idea what you are talking about.}
+into the current branch, thus confounding the branch relationships.
+
+% TODO: Get a diagram of a foxtrot
 
 The repository events in the DAG are immutable; once a repository event
-is created, it can't be changed. \dmg{the next sentences should be
-  repository events, not commits} Git allows operations to alter
+is created, it can't be changed. Git allows operations to alter
 repository events and re-order them, but this will create a new event

### 2017-07-15 paper/discussion.tex
adding fixes for discussion

--- 2017-07-14 paper/discussion.tex
@@
 
-The results show conclusive results; the DAG visualization is unable to
-adequately provide users with an explanation of the changes being made
-in the repository, the DAG visualizatoin is not easily understood, and
-is unable to provide summarizations of the changes being made at a
-merge. Meanwhile, the \mt visualization is better able to provide an
-explanation of how a commit is integrated, along with better
-summarization. In this section, we discuss some observations, threats to
-the validity, and comments made by the participants during the study
-followed by discussion about modifications to the original algorithm to
-help generalize it for more repositories, concluded with future work.
+\section{Discussion}
+\label{sec:discussion}
+
+The results show conclusive results; the DAG visualization provided by
+Gitk is unable to provide an explanation of the changed being made in
+the repository, is not easily understood, and is unable to provide
+summarizations of the changes being made at a merge. Meanwhile, the \mt
+visualization of \tool is better able to provide an explanation of how a
+commit is integrated, along with better summarization. In this section,
+we discuss some observations, threats to the validity, and comments made
+by the participants during the study followed by discussion about
+modifications to the original algorithm to help generalize it for more
+repositories, concluded with future work.
 
@@
 
-% \dmg{I think you mean challenge using gitk, right? so be explicit}
-
 During the course of the study, we noticed an issue that was consistent
@@
 identify the tag as being the ``master branch'', but a tag is not a
-branch. The visualization of the DAG does not provide any notion of
+branch (it simply points to a given commit). The visualization of the DAG does not provide any notion of
 which branch being shown is the master branch. The default assumption is
 that the first branch on the left is the master branch, but this is not
-a good assumption. Furthermore, the visualization in Gitk is not
+always correct. Furthermore, the visualization in Gitk is not
 consistent, the ordering of the branches changes between runs, so
@@
 
+% \evan{I watch the DAG change three times in a row on the visualization for \comB}
+
 If users were able to easily identify the master branch using the
 visualizations of the DAG, the results of the study would likely be very
-different. This would certainly be the case for the single-commit tree;
-it is trivial to summarize information about a single item, but
-identifying that there was only a single commit in the tree was
-non-trivial. This information is important because many trees contain a
-single commit, and the structure of these trees is identical to the
-structure of flat trees. Flat trees are the most common branching
-structure in the repository, with the root being the parent to all of
-the other nodes in the tree; to improve summarization and comprehension
-of these trees, it would likely by sufficient to indicate the master
-branch in the DAG visualization.
+different, making the differences between Gitk and \tool less
+significant. This would help in the case of the single-commit trees, it
+is trivial to summarize information about a single item, but identifying
+that there was only a single commit in the tree was non-trivial. This
+information is important because more than a quarter of the trees in the
+Linux repository contain a single commit, and the structure of these
+trees is identical to the structure of flat trees. Flat trees are the
+most common branching structure in the repository, with the root being
+the parent to all of the other nodes in the tree; to improve
+summarization and comprehension of these trees, it would likely by
+sufficient to indicate the master branch in the DAG visualization.
 
@@
 three years, working with both SVN and CVS repositories, but had little
-experience with git. This subsection contains improvements made by this
-participant during the study and afterward.
+experience with git. This subsection contains improvements proposed by
+this participant during the study and afterward.
 
@@
 commits were made, as the order tells the story of what the developer
-was thinking as they were writing the changes. The current \mt model
-does not order the commits, simply placing them randomly. This is the
-primary reason behind this participant's request to use both
-simultaneously. \tool is able to help with the aggregation of the
-information, and provide a better understanding of the next merge
-involved in integrating this commit, but the DAG visualization in Gitk
-provides the full story of the commit instead of hiding it behind a
-layer of abstraction.
+was thinking as they were writing the changes. The visualization of the
+\mt in \tool does not order the commits, randomly ordering them in each
+level. The \mt model does retain this information, so implementing this
+modification is trivial. This is the primary reason behind this
+participant's request to use both tools simultaneously. \tool is able to
+help with the aggregation of the information, and provide a better
+understanding of the next merge involved in integrating this commit, but
+the DAG visualization in Gitk provides the full story of the commit
+instead of hiding it behind a layer of abstraction.
 
@@
 
-The answers provided by the participants to earlier tasks were not
-taken into account for determining the accuracy. If the participant was
-unable to determine the correct commits for the tree within the
-conceptual task group, the summarization results would be recorded as
-wrong, even if the summarization they provided was correct for the set
-of commits selected to be part of the tree. A further study could be
-done where the correct answer was provided to the participant after each
-task so that errors would not propagate between tasks.
+The answers provided by the participants to earlier tasks were not taken
+into account for determining correctness or accuracy in further
+questions. An incorrect answer may impact the results of the tasks that
+followed. If the participant was unable to determine the correct commits
+for the tree within the conceptual task group, the summarization results
+would be recorded as wrong, even if the summarization they provided was
+correct for the set of commits selected to be part of the tree. A
+further study could be done where the correct answer was provided to the
+participant after each task so that errors would not propagate between
+tasks.
 
@@
 
-% A possible
-% reason for this may be that unless a repository reaches a sufficient
-% complexity, both in branching and number of commits, these tools are
-% unnecessary for the normal operation of git. Furthermore, fear of the
-% complexity caused by branching may cause personal and academic
-% repositories to have relatively simple structures. While they were
-% unfamiliar with the DAG structure, these users were also unfamiliar with
-% the \mt model, and were still able to better understand and summarize
-% the repository using \tool.
+The participants in the study were students, some with industrial
+experience. Most of the participants had worked relatively few
+collaborators, so professional developers may have been able to better
+comprehend the DAG visualizations in Gitk.
 
-\subsection{Generalization}
+\section{Generalizing the Creation of the \mt}
 \label{sub:generalization}
@@
 
-If this algorithm is to be generalized into a dynamic, web-based, tool,
-it may not be feasible to download every every repository event in the
+If this algorithm is to be generalized to work in a dynamic, web-based,
+tool it may not be feasible to download every every repository event in the
 repository in a reasonable period of time. We instead look toward a

### 2017-07-15 paper/study.tex
fixing study

@@
   \centering
-  \caption{Cross-Commit Results\dmg{Be more descriptive in the captions.People read the papers via tables and
-      pictures. What does the table say? e.g. as it can be seen.... tell a story with the picture, instead of asking the
-    reader to make up the story... which might be different than the one you want them to have}}
+  \caption{Cross-Commit Metrics: Showing that tree size make a
+    difference in the correctness metric for T5 and T9, and the timing
+    metric for task T7.}
   \label{tab:cross_commit_results}
@@
   \centering
-  \caption{Summarized Correctness, Accuracy, and Timing results for the summarization tasks}
+  \caption{Summarized Correctness, Accuracy, and Timing results for the
+    summarization tasks. In all cases except for the timing metric of
+    T7, there was a difference between \tool and Gitk.}
   \label{tab:summarization_table}
@@
 tables~\ref{tab:summarization_table}.
-The full results are as follows;
-
-\dmg{you wrongly use ``semicolon'', you mean ``colon''. Same below...}
+The full results are as follows:
 
@@
 
-\dmg{Emphasize: To summarize...}
-
-\tool is able to assist users with summarizing various aspects of a
-merge more quickly and accurately than Gitk. In two tasks, T5 and T9,
-correctness was affected by the size of a tree, but in both trees, \tool
-was able to provide a statistically significant improvement to the
-correctness. Only in one task, T7, was timing affected by the size of a
-tree; \tool did not significantly improve the time taken to answer T7 on
-the smaller tree, but did improve the time taken on the larger tree.
+To summarize, \tool is able to assist users with summarizing various
+aspects of a merge more quickly and accurately than Gitk. In two tasks,
+T5 and T9, correctness was affected by the size of a tree, but in both
+trees, \tool was able to provide a statistically significant improvement
+to the correctness. Only in one task, T7, was timing affected by the
+size of a tree; \tool did not significantly improve the time taken to
+answer T7 on the smaller tree, but did improve the time taken on the
+larger tree.
 
@@
 Among the 12 participants, there was nearly unanimous agreement that for
-conceptual understanding and summarization tasks; \dmg{this is another wrong use of semicolon}\tool was easier to
-work with \dmg{avoid idiomatic expressions: what does ``work with'' mean? use a precise verb. English is a rich language.} than Gitk. The participants cited the ability to abstract
+conceptual understanding and summarization tasks, \tool was easier to
+use than Gitk. The participants cited the ability to abstract
 information about the merge from the clean summarization tables and
-simple visualizations \dmg{be precise: from Linvis' clean summarization tables and ... }. Three participants suggested that someone with a
-professional understanding of Gitk and the git command-line may be able to
-extract a conceptual understanding from the DAG visualization and
-perform the summarization tasks. One of these three participants said
-that they would prefer to have both tools available, as they are able to
+simple visualizations as the primary reasons for preferring \tool to
+Gitk. Three participants suggested that someone with a professional
+understanding of Gitk and the git command-line may be able to extract a
+conceptual understanding from the DAG visualization and perform the
+summarization tasks. One of these three participants said that they
+would prefer to have both tools available, as they are able to
 complement each other.

### 2017-07-15 paper/relatedwork.tex
make use of "Merge-Tree" consistent
fixing one colon issue

@@
 made, only the frequency that they are being made and who is making
-them. Codeswarm\cite{ogawa09} is similar to Gource, using an organic approach to
-visualizing events in the repository. Unlike Gource, which focuses on
-the files, Codeswarm focuses on developers and the number of commits
-they are making.
+them. Codeswarm\cite{ogawa09} is similar to Gource, using an organic
+approach to visualizing events in the repository. Unlike Gource, which
+focuses on the files, Codeswarm focuses on developers and the number of
+commits they are making.
 
-Git itself is shipped with summarization and visualization tools: the
-\verb|git log| command and Gitk, which is a graphical tool for the
-purpose of browsing the DAG of the repository. Many other tools are
+Git itself is shipped with summarization and visualization tools; Gitk
+and \verb|git log --graph| are both built-in tools for the purpose of
+visualizing and browsing the DAG of the repository. Many other tools are
 designed with a similar visual metaphore as what is presented in Gitk,

### 2017-07-15 paper/abstract.tex
Fixing spelling error

@@
 This paper make three contributions; a conversion from DAG to \mt, an
-implemenation of a tool built on the \mt model, and a user study to
+implementation of a tool built on the \mt model, and a user study to
 evaluate and validate the implementation and model.

### 2017-07-19 paper/model.tex
Fixing the "commit sizes" sentence
correcting the algorithm
Improves the description, provides an example, a new figure, and
corrects the algorithm.

@@
 
-Notice that the DAG loses information about the master branch and the
-repository that the master branch is part of. The \mt view of
-this DAG is visible in Figure~\ref{fig:repoTree}. Note that the
-direction of the edges of the DAG have been inverted, instead of
-pointing from the child to the parent, it points from the parent to its
-children, forming a path to the master branch. Also note that the DAG
-has been simplified, showing only a single edge on the path to master
-for any commit.
+The DAG does not retain information about where a commit was originally
+created. The \mt form (Figure~\ref{fig:repoTree}) does not completely
+rebuild the lost information, but is able to re-create the sequence of
+merges required to merge the event, and the depth, in the tree, of that
+event. The \mt does not include information about how the repository
+events that are the children of a merge are related. The order of the
+children does not matter.
+
+Combining the DAG and the \mt results in a hybrid structure
+(Figure~\ref{fig:repoDAGTree}) that no only shows which merges a commit
+passes through, it rebuilds the order of those commits. In the example
+shown, this hybrid structure makes it clear that merge 9 merges the
+changes made in 6 and 8 into the commit 7. Finally, 11 merges merge 9
+into the master branch.
 
@@
   \caption{Example of a sequence of events performed in different
-    repositories. The horizontal axis represents time. Each horizontal
-    section represents a different branch and/or repository. Each commit
-    points to its parent. The intial commit is at time $t_0$, and the
+    repositories. The horizontal axis represents time. The branches and
+    repositories are aligned horizontally, and color-coded. Each commit
+    points to its parent. The initial commit is at time $t_0$, and the
     head is at $t_8$.}
@@
 
+\begin{figure}[htpb]
+  \begin{center}
+    \begin{tikzpicture}[auto, on grid, semithick, node distance=1cm, state/.style={circle, text=black, minimum size=7mm}]
+
+      \node[state, draw=chartblue] (1) {1};
+      \node[state, draw=chartblue, right=of 1] (4) {4};
+      \node[state, draw=chartblue, right=of 4] (11) {11};
+      \node[state, draw=chartblue, right=2.5cm of 11] (12) {12};
+
+      \node[state, draw=chartyellow, above right= 1cm and 0.5cm of 11] (7) {7};
+      \node[state, draw=chartyellow, above left= 1cm and 0.5cm of 11] (5) {5};
+      \node[state, draw=chartyellow, right=of 7] (9) {9};
+      \node[state, draw=chartyellow, left=of 5] (2) {2};
+      \node[state, draw=chartmagenta, above=of 5] (3) {3};
+      \node[state, draw=chartmagenta, above left=1cm and 0.5cm of 9] (6) {6};
+      \node[state, draw=chartmagenta, above right=1cm and 0.5cm of 9] (8){8};
+      \node[state, draw=chartyellow, above=of 12] (10) {10};
+
+      \draw (11) edge[chartyellow, stealth-] (2) edge[chartyellow, stealth-] (5) edge[chartyellow, stealth-] (7) edge[chartyellow, stealth-] (9)
+            (5) edge[chartmagenta, stealth-] (3)
+            (9) edge[chartmagenta, stealth-] (6) edge[chartmagenta, stealth-] (8)
+            (12) edge[chartyellow, stealth-] (10);
+    \end{tikzpicture}
+  \end{center}
+  \caption{The \mt{s} computed for each commit in
+    Figure~\ref{fig:repoDAG} showing the path that each commit takes to
+    be merged into the master branch of the repository. This does not
+    indicate how the events being merged are related. We retained the
+    numerical order of the events, but the order can be arbitrary.}
+  \label{fig:repoTree}
+\end{figure}
+
 \begin{figure}[htbp]
@@
   \end{tikzpicture}
-  \caption{\mt view of the commits  represented in
-    Figure~\ref{fig:repoEvents} showing the path they followed to reach
-    the master branch. In this model the successors of each commit
-    represents the path followed by that commit to reach the master
-    branch.}
-  \label{fig:repoTree}
+  \caption{The \mt applied to the inverted DAG to show the order that
+    commits are created, as well as the merge sequence for the events in
+  Figure~\ref{fig:repoEvents}. This requires a post-processing step
+  involving the \mt produced by Algorithm~\ref{fig:alg} and the DAG.}
+\label{fig:repoDAGTree}
 \vspace{-3mm}
@@
 
-Computing the \mt from a DAG for any repository may not be
-possible; however, certain features of the development process of Linux
-make it feasible to compute the \mt for the Linux repository.
-First, the master branch of Linux is maintained by Linus Torvalds, and
-only Linus has write access to it. We have verified this assertion in
-previous research~\cite{German2015}. We have developed a heuristic that
-is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
+Computing the \mt from a DAG for any repository may not be possible;
+however, certain features of the development process of Linux make it
+feasible to compute the \mt for the Linux repository. First, the master
+branch of Linux is maintained by Linus Torvalds, and only Linus has
+write access to it. We have verified this assertion in previous
+research~\cite{German2015}. We have developed a heuristic that is
+presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
 identifies the commits made directly to the master branch, whereafter it
-recursively determines the shortest path (in terms of time), using the
-DAG, from each commit to the master branch using the inverted DAG.
+recursively determines the shortest path, using the DAG, from each
+commit to the master branch using the inverted DAG.\@
 
@@
         \caption{Computing the \mt of Linux from the DAG}\label{fig:alg}
-        \begin{algorithmic}
+        \begin{algorithmic}[1]
                 \Function{ComputeMergeTree}{DAG}: tree
-                \State {\# Compute the tree from the DAG of Linux repository.}
-                \State {\# Returns $Tree$, a graph containing every commit }
-                \State {\# in DAG with the path it followed to master.}
-
+                % \State {\# Compute the tree from the DAG of Linux repository.}
+                % \State {\# Returns $Tree$, a graph containing every commit }
+                % \State {\# in DAG with the path it followed to master.}
                 \State $head \gets \textit{Head of master of git repository}$
                 \State $master \gets \textit{traverse DAG from head using }$
-                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
+                \State \quad\quad\quad\quad $\textit{first parent until reaching root}$
                 \State $nodes(Tree) \gets nodes(DAG)$
-                \State \Function{distance2Master}{cid} : seconds
+                \State \Function{MergeAtMaster}{cid}
+                \State {\# Returns $(depth, merge, next)$}
                 \State {\# Helper function}
-                \State {\# Recursively compute shortest distance to master}
-                \State {\# setting cid's successor (next) in its way to master.}
-                \State {\# This function should be memoized. Otherwise it}
-                \State {\# would run in exponential time.}
+                \State {\# Compute the closest merge into master, }
+                \State {\# setting the children on the way to master.}
                 \If {\textit{cid in master}}
-                \State \Return 0
+                \State \Return $(0, cid, \varnothing)$
                 \EndIf
-                \State    $d \gets 	\infty$
+                \State {$d \gets \infty$}
                 \State {\# Traverse the inverted DAG}
                 \For{$c \in children(cid, DAG)$}
-                \If {$c \in master$}
-                \State $d_1 \gets commitTime(c)-commitTime(cid)$
-                \Else
-                \State {$d_1 \gets distance2Master(c)$}
+                \State $(d_c, merge_c, next_c) \gets MergeAtMaster(c)$
+                \If {$IsMerge(c)$}
+                \State $fp \gets FindFirstParent(c)$
+                \If {$fp \neq cid$}
+                \State $d_c \gets d_c + 1$
+                \State $next_c \gets c$
                 \EndIf
-                \If {$d_1 < d $}
-                \State $next \gets c$
-                \State  $d \gets d_1$
+                \EndIf
+                \State {\# Find the shortest path}
+                \If {$d_c < d$}
+                \State $(d, m, next) \gets (d_c, merge_c, next_c)$
+                \ElsIf{ $d_c = d$ }
+
+                \State {\# Use the time as a tie-breaker}
+                \If {$ cTime(merge_c) < cTime(m) $}
+                \State $(m, next) \gets (merge_c, next_c)$
+                \EndIf
                 \EndIf
@@
                 \State add edge $(cid, next)$ to $Tree$
-                \State \Return $d$
+                \State \Return $(d, m, next)$
                 \EndFunction
@@
                 \For{$c \in nodes(DAG)$}
-                \State $distance2Master(c)$
+                \State $MergeAtMaster(c)$
                 \EndFor
@@
 \end{algorithm}
+
+The algorithm is broken into two phases. The first is determining which
+repository events are on the master branch. This is done by traversing
+the first parent from the master branch reference to the commit that has
+no parents. The second phase is encompassed by the function
+$MergeAtMaster$ which determines, for each commit, which merge the
+commit is merged at, the depth (as variable $d$ in the algorithm), and
+the next merge on the path to the master branch. The function
+$MergeAtMaster$ has two parts, the first for determining the depth, from
+the master branch, that the repository event is at. The second phase
+determines the merge into the master branch, and the next merge on the
+way to the master branch. The distance is by shortest path, staying as
+close to the master branch as possible. If there is a tie between two
+paths, the path that merges into the master branch sooner is taken.
+
+To demonstrate the behaviour of the algorithm, we use a short example,
+computing the merge at commit 5 in Figure~\ref{fig:repoEvents}.
+$MergeAtMaster$, recurses along the children of the nodes it visits.
+Eventually every child of every node along the path will be visited at
+least once. Without loss of generality, suppose that the path recursed
+along is from node 5 to 6, 9, 10, and finally 12. The depth for each,
+except 12 (a merge into the master branch), is initialized to infinity,
+the merge into master is blank, and the next merge is blank. Merges into
+master trivially have a distance of 0 from the master branch, and it
+merges itself into the master branch. The recursion at 12 returns the
+triple $(0, \varnothing, 12)$ to the call from 10. 12 is a merge commit
+and 10 is not the first parent, so the temporary depth, $d_c$, is
+incremented to 1 and the temporary next merge, $next_c$, is changed to
+12. 1 is less than infinity, so the depth is set to 1, the merge to 12,
+and the next to 12. This returns the triple $(1, 12, 12)$ to the call
+from 9. 9 is the first parent of 10, so no changes are made to the
+temporary variables. The call to 9 recurses to the second child, 11. 11
+is a merge into the master so it returns $(0, \varnothing, 11)$ to the
+call from 9. 9 is not the first parent of 11, so the $d_c$ is
+incremented to 1 and $next_c$ is changed to 11. The distances $d_c$ and
+$d$ are the same, so to break the tie, we use the time. 12 was merged
+after 11, so 11 replaces 12 as the merge into the master branch for 9,
+as well as being the next merge. The call for 9 returns the triple $(1,
+11, 11)$ to the call for 6. 6 is not the first parent of 9, so $d_c$ is
+incremented and $next_c$ is changed to 9, as 9 merges 6. 2 is less than
+infinity, so the $d$ is changed to 2, the merge to 11, and the next
+merge to 9. The call to 6 returns the triple $(2, 11, 9)$ to the call
+for 5. The call for 5 recurses on the second child of 5, calling on 7,
+which calls 8, and then 9. 9 can continue, but if the implementation of
+the algorithm uses memoization, the call to 9 can immediately return the
+triple $(1, 11, 11)$ to the call for 8, and avoid an exponential
+runtime. 8 is not the first parent of 8, so $d_c$ is incremented to 2
+and $next_c$ is changed to 9. 2 is less than infinity, so $d$ is changed
+to 2, $merge$ to 11, and $next$ to 9. The call to 8 returns the triple
+$(2, 11, 9)$ to the call for 7, which recurses on the second child of 7,
+9. 9 returns the triple $(1, 11, 11)$. 7 is the first parent of 9, so
+the depth is not incremented. $d_c$ is less than $d$, so $d$ is changed
+to 1, $m$ to 11, and $next$ to 11, returning $(1, 11, 11)$ to the call
+for 5. $d_c$ is less than $d$, so $d$ is changed to 1, $m$ to 11, and
+$next$ to 11. There are no other children, so the function halts.
 
@@
         \centering
+        \vspace{3mm}
         {\fontsize{7}{9}
@@
         \label{fig:sampleMerge}
+  \vspace{2mm}
 \end{figure}
@@
 extraction of the commit history up to July 20, 2016 form the Linux
-repository. We computed the \mtof every commit until that date.
-Since Linus Torvalds mostly does merging directly into the master
-branch, we assumed that every merge by Linus is the root of a \mt, later
-detecting merges that do not merge into the master branch and removing
-them from the set of root merges. As described above, the log of a
-merge-commit usually contains the number of commits in the merge the
-first 20 summaries of commits being merged. We extracted merges by Linus
-Torvalds using the command \verb|git log --merges --author='Torvalds'|
-and compared the number of commits according to the log with the number
-of commits in the \mt rooted in this commit. We also used the summaries
+repository. We computed the \mt of every commit until that date. Since
+Linus Torvalds mostly does merging directly into the master branch, we
+assumed that every merge by Linus is the root of a \mt, later detecting
+merges that do not merge into the master branch and removing them from
+the set of root merges. As described above, the log of a merge-commit
+usually contains the number of commits in the merge the first 20
+summaries of commits being merged. We extracted merges by Linus Torvalds
+using the command \verb|git log --merges --author='Torvalds'| and
+compared the number of commits stated in the log message with the number
+of commits in the \mt. We also used the summaries
 of the commits found in the merge (not necessarily all---see above) to
-make sure those commits were in their corresponding \mt For
-example, for the merge in Figure~\ref{fig:sampleMerge} we would expect
-that the \mt rooted at \mycode{8cbd84f} contains 24 commits, and
-the one-line summaries corresponds to commits in that \mt. We also
-inspected those with differences to make sure they were true errors. The
-results can be summarized as follows:
+make sure those commits were in their corresponding \mt For example, for
+the merge in Figure~\ref{fig:sampleMerge} we would expect that the \mt
+rooted at \mycode{8cbd84f} contains 24 commits, and the one-line
+summaries corresponds to commits in that \mt. We also inspected those
+with differences to make sure they were true errors. The results can be
+summarized as follows:
 

### 2017-07-19 paper/background.tex
editing descriptions of algorithm

@@
 two branches together. In this paper, we will refer to the individual
-commits code-carrying as commits, the structural merge commits as
+code-carrying commits as commits, the structural merge commits as
 merges, and refer to both types without discrimination as repository

### 2017-07-19 paper/discussion.tex
editing descriptions of algorithm

@@
 
-Seeing that the visualizations of the \mt model are able to assist users
-with understanding how commits are integrated, we continue by looking at
-ways of improving the algorithm for faster operations, and making the
-conversion from DAG to \mt feasible over the internet.
+The visualizations of the \mt model assist users with understanding how
+commits are integrated. We continue by discussing possible methods for
+improving the algorithm to directly produce the combined DAG \mt without
+a post-processing step and pruning the DAG to improve the performance
+and make it feasible to compute over the internet. Furthermore, we look
+to generalize the algorithm to work with arbitrary repositories.
 
-The problem of identifying the master branch is the most difficult, and is
-not possible without heuristic approaches. The master branch can be
-confounded by \foxtrot merges, which change the order of the
-parents in the parent list of a merge. Other than the ordering of the
-parent list, git has no other way to determine which commits are on the
-master branch of the repository. For the purpose of this discussion, we
-will assume that we can rely on the order of the parent list to be
-correct, and that \foxtrot{s} do not occur. Furthermore, merge
-information is lost in the event of fast-forward merges, which collapse
-the branch into the current branch.
+The problem of identifying the master branch is the most difficult, and
+is not possible without heuristic approaches. The master branch can be
+confounded by \foxtrot merges, which change the order of the parents in
+the parent list of a merge. Other than the ordering of the parent list,
+git has no other way to determine which commits are on the master branch
+of the repository. For the purpose of this discussion, we will assume
+that we can rely on the order of the parent list to be correct, and that
+\foxtrot{s} do not occur. Furthermore, merge information is lost in the
+event of fast-forward merges, which collapse the branch into the current
+branch.
 

### 2017-07-19 paper/model.tex
editing descriptions of algorithm

@@
   \end{tikzpicture}
-  \caption{Example of a sequence of events performed in different
+  \caption{An example sequence of events performed in different
     repositories. The horizontal axis represents time. The branches and
@@
 least once. Without loss of generality, suppose that the path recursed
-along is from node 5 to 6, 9, 10, and finally 12. The depth for each,
-except 12 (a merge into the master branch), is initialized to infinity,
-the merge into master is blank, and the next merge is blank. Merges into
-master trivially have a distance of 0 from the master branch, and it
-merges itself into the master branch. The recursion at 12 returns the
-triple $(0, \varnothing, 12)$ to the call from 10. 12 is a merge commit
-and 10 is not the first parent, so the temporary depth, $d_c$, is
-incremented to 1 and the temporary next merge, $next_c$, is changed to
-12. 1 is less than infinity, so the depth is set to 1, the merge to 12,
-and the next to 12. This returns the triple $(1, 12, 12)$ to the call
-from 9. 9 is the first parent of 10, so no changes are made to the
-temporary variables. The call to 9 recurses to the second child, 11. 11
-is a merge into the master so it returns $(0, \varnothing, 11)$ to the
-call from 9. 9 is not the first parent of 11, so the $d_c$ is
-incremented to 1 and $next_c$ is changed to 11. The distances $d_c$ and
-$d$ are the same, so to break the tie, we use the time. 12 was merged
-after 11, so 11 replaces 12 as the merge into the master branch for 9,
-as well as being the next merge. The call for 9 returns the triple $(1,
-11, 11)$ to the call for 6. 6 is not the first parent of 9, so $d_c$ is
-incremented and $next_c$ is changed to 9, as 9 merges 6. 2 is less than
-infinity, so the $d$ is changed to 2, the merge to 11, and the next
-merge to 9. The call to 6 returns the triple $(2, 11, 9)$ to the call
-for 5. The call for 5 recurses on the second child of 5, calling on 7,
-which calls 8, and then 9. 9 can continue, but if the implementation of
-the algorithm uses memoization, the call to 9 can immediately return the
+along is from node 5 to 6, 9, 10, and finally 12.
+
+The depth for each, except 12 (a merge into the master branch), is
+initialized to infinity, the merge into master is blank, and the next
+merge is blank. Merges into master trivially have a distance of 0 from
+the master branch, and it merges itself into the master branch. The
+recursion at 12 returns the triple $(0, \varnothing, 12)$ to the call
+from 10. 12 is a merge commit and 10 is not the first parent, so the
+temporary depth, $d_c$, is incremented to 1 and the temporary next
+merge, $next_c$, is changed to 12. 1 is less than infinity, so the depth
+is set to 1, the merge to 12, and the next to 12. This returns the
+triple $(1, 12, 12)$ to the call from 9. 9 is the first parent of 10, so
+no changes are made to the temporary variables.
+
+The call to 9 recurses to the second child, 11. 11 is a merge into the
+master so it returns $(0, \varnothing, 11)$ to the call from 9. 9 is not
+the first parent of 11, so the $d_c$ is incremented to 1 and $next_c$ is
+changed to 11. The distances $d_c$ and $d$ are the same, so to break the
+tie, we use the time. 12 was merged after 11, so 11 replaces 12 as the
+merge into the master branch for 9, as well as being the next merge. The
+call for 9 returns the triple $(1, 11, 11)$ to the call for 6. 6 is not
+the first parent of 9, so $d_c$ is incremented and $next_c$ is changed
+to 9, as 9 merges 6. 2 is less than infinity, so the $d$ is changed to
+2, the merge to 11, and the next merge to 9. The call to 6 returns the
+triple $(2, 11, 9)$ to the call for 5.
+
+The call for 5 recurses on the second child of 5, calling on 7, which
+calls 8, and then 9. 9 can continue, but if the implementation of the
+algorithm uses memoization, the call to 9 can immediately return the
 triple $(1, 11, 11)$ to the call for 8, and avoid an exponential
@@
         \vspace{3mm}
-        {\fontsize{7}{9}
-        \begin{verbatim}
+        \fontsize{7}{9}
+        \begin{Verbatim}[frame=single]
 Merge: 8cbd84f fd8aa2c
@@
 ...
-        \end{verbatim}}\vspace{-5mm}
+        \end{Verbatim}
         \caption{Example of how merges record a subset of commits being merged. The
@@
 of the commits found in the merge (not necessarily all---see above) to
-make sure those commits were in their corresponding \mt For example, for
+make sure those commits were in their corresponding \mt. For example, for
 the merge in Figure~\ref{fig:sampleMerge} we would expect that the \mt

### 2017-07-19 response/response.txt
adding response letter

We would like to thank the reviewers for their extensive comments and
analysis. The paper has been considerably improved as a results of the
review.

The paper has been heavily modified, shifting focus from the
generalization of the algorithm toward the comparison user study between
Linvis and Gitk.

Linvis has since had minor updates. We removed most of the search
interface, replacing it with a full-text search engine that searches by
author name, filename, commit hash, commit and authorship date, and
keywords from the commit logs. The results of the search engine are
aggregated together by merge-tree.

The responses to Reviewer 1:

      While the technical descriptions of the algorithms are detailed and
      comprehensive, they need to be re-written for better clarification. The
      reading of Section III.B is specially hard. It is a complicated
      algorithm that demands a lot of attention from the reader. As such, in
      order to keep the reader interested, the authors should improve the
      description by adding a diagram or a flowchart. A simple example that
      the reader can follow along with the authors would also be a great
      improvement.

We provided a further description of the original algorithm followed by
a step-by-step example starting at commit 5 from the example in Figure 3
to help clarify things in Section III.A.

The generalized algorithm that was presented in section III.B was moved
to a separate section (Section IX). We re-worded and extended the
description and provided an example starting with the merge commit 11
(the algorithms run in opposing directions on the DAG) from the example
shown in Figure 3.

      The testing and evaluation of the generalized version of the algorithm
      must also be improved. Right now the test suite and the results are
      not concretely described and must be better clarified.

This is not applicable anymore. The generalization has become a
discussion point in Section IX. The proposed algorithm is a possible
means of improving the runtime performance by pruning the DAG to contain
a smaller superset of the repository events necessary for computing the
merge-tree of a given merge. The test suite was not a comprehensive test
of all possible merges, but tested some of the issues that we
encountered while designing it and the original algorithm.

    While the tree-based visualizations are well described, they are not
    novel visualizations and simply use abstractions that are already
    available from previous work. As such, the practical use of these
    visualizations in the context of the paper should be better described.
    The authors are advised to add at least one fully described use case,
    as a step-by-step that can be followed by the reader, showing how the
    visualizations are actually used to obtain useful insight from the
    merge-trees.

The intent is less focused on providing a new metaphor for trees, and
more focused on providing a new visualization of git repositories. The
visualization is focused on showing how a commit is integrated into the
master branch.

We conducted a user study to verify that the goals of the model and tool
are met with the visual metaphors and tables provided. The study is
broken down into what we see as being the two major issues with the DAG;
conceptual understanding and summarization.

See section VII.

Responses to the Attached PDF:

    Introduction
    The last paragraph of the intro is a bit confusing. It is not clear
    whether the authors mean that the algorithm was extended to work with
    more than one repository, or with repositories other than the Linux
    kernel (or both at the same time, which is what I think is the right
    answer). Also, they talk about an extension, explain its goals, but
    not what that extension is, i.e., how the original algorithm is
    actually extended to reach the goals (that should be done here only
    briefly, of course). A clearer explanation here would help motivate
    the reader to continue reading.

The generalization has been moved to the discussion portion of the paper
to signify that it is less of a contribution, and more of an idea. The
extension section is introduced with

"
The visualizations of the Merge-Tree model assist users with
understanding how commits are integrated. We continue by discussing
possible methods for improving the algorithm to directly produce the
combined DAG Merge-Tree without a post-processing step and pruning the
DAG to improve the performance and make it feasible to compute over the
internet. Furthermore, we look to generalize the algorithm to work with
arbitrary repositories.
"

To address the confusion surrounding the type of generalization, we
specify it in the last sentence of the first paragraph of section IX.

"Furthermore, we look to generalize the algorithm to work with arbitrary
repositories."

    In page 2, column 2, line 28, the authors introduce the concept of an
    “event”. By reading further, one realizes that “events” are either
    commits or merges. However, in previous paragraphs, the authors
    regarded this as “merge commits” and ”non-merge commits”. A
    clarification of these two terminologies would be useful to make the
    text clearer.

The terminology surrounding events, commits, and merges has been
clarified in the last sentence of the first paragraph of the Background
Section.

"In this paper, we will refer to the individual code-carrying commits as
commits, the structural merge commits as merges, and refer to both types
without discrimination as repository events."

    In Figure 3, legend: The sentence “Each horizontal section” is not
    totally clear. One way of improving the figure would be to add a
    separator between branches/repos, effectively turning the image into a
    matrix. Or simply explaining more clearly that nodes are arranged in
    rows. Also, the authors should mention that the colors also represent
    branches/repos.

The caption has been clarified and we state that the branches and
repositories are color-coded.

"
An example sequence of events performed in different repositories. The
horizontal axis represents time. The branches and repositories are
aligned horizontally, and color-coded. Each commit points to its parent.
The initial commit is at time t0, and the head is at t8.
"

    In Figure 4: the authors mention that the DAG loses information about
    which repository or branch each commit is performed in, but the figure
    retains the nodes’ colors, which contradicts that statement. While the
    reader comprehends the concept, removing the colors of the figure
    would probably communicate the message better.

The color was removed from figure four to clarify that branch and
repository information is removed in the DAG.

    Figures 3, 4 and 5 represent a sequence of steps in the explanation of
    the algorithm, so it would be significantly better for the reader if
    they were all in the same page.

The figures have been placed on the same page, and are in the same
column.

    Constructing the Merge Tree
    A small description of the children function on Algorithm 1 would be
    useful, as it searches for children in the opposite way than a normal
    depth-search algorithm does. A normal depth-search would traverse the
    nodes to which cid points to (its ancestors in git), but children does
    the opposite: it traverses the nodes which point to cid

The children function is the list of (DAG) children for the commit. I am
unsure of how to make this more explicit.

    There is apparently a problem when calculating d in Algorithm 1. It is
    only updated when a commit is directly connected to a master commit,
    i.e., it has a child c in master. As an example, consider first node 9
    from Fig. 4. It has a child in master (node 11), so its distance is
    d(9) = commitTime(11) – commitTime(9). Now let’s consider node 8. Its
    only child is node 9, which is not in master. Thus, it recursively
    calls distance2Master(9), which returns d(9). Since d(9) is smaller
    than infinity, then d(8) = d(9). But shouldn’t it be d(8) =
    commitTime(8) – d(9)? The problem becomes worse when we consider node
    7. It has two children, nodes 8 and 9. If d(8) = d(9), then how can
    node 7 know which of its children to choose? By looking at Figure 5 it
    obviously chose node 9, which seems to be the right choice, but was it
    an arbitrary choice?

Thank you for identifying this issue. We missed a few details of the
algorithm as implemented when describing it in the paper. We have
carefully followed our implementation to re-describe the algorithm more
accurately. Furthermore, a full example stepping through the algorithm
has been provided, starting at commit 5. We have worked to fully clarify
the algorithm and how it works. We are also making the source code
available in the replication package.

    Section III.B starts abruptly without a clear motivation on what the
    Generalized Merge Tree Construction is supposed to do. The authors
    should better motivate the jump from the Linux kernel tree to the
    generalized one. Also, it seems that the authors have used “branch”
    and “commit” interchangeably, without a clear distinction. The
    sentence “There is still some ambiguity in which child is the correct
    child to link to for a branch after phase 1”, for example, is not
    clear. In the sentence “The order of children at a branch node is
    based on the length of the path to that node, not the order that the
    nodes show up in as parents.”, what is a “branch node”, and what does
    the order of its children impact? Such concepts were not introduced
    before this section, and as such it makes the reading very hard. I
    suggest the authors start this section with a more general and
    “lightweight” introduction of concepts and ideas before going fully
    into technical details.

The section has been removed and replaced with a new section for
discussing the generalization, Section IX.

The first sentence motivates generalizing the algorithm as we have
quantifiable evidence that the merge trees help users. Therefore a
generalization is the next step.

"
The visualizations of the Merge-Tree model assist users with
understanding how commits are integrated into a project. We
continue by discussing possible methods for improving the
algorithm to produce the combined DAG Merge-Tree without
a post-processing step and pruning the DAG to improve the
performance and make it feasible to compute over the internet.
Furthermore, we look to generalize the algorithm to work with
arbitrary repositories.
"

    In Algorithm 2, what is a parent? Are we dealing with the normal DAG
    or the inverted DAG? Apparently neither, because Algorithm 2 does not
    mention any symbol called DAG. But if Algorithm 2 starts from the root
    commit, and according to Section II the root has no ancestors, who are
    the parents being accessed in line 6 of the algorithm (and everywhere
    else)? It would be logical to assume that Algorithm 2, starting from
    the root, would have no parents and thus only run once and exit. I
    assume that by “root” the authors mean any master commit (so it
    would be the root of one merge-tree, and not of the whole
    repository).

This is addressed by specifying the terminology surrounding parents and
children in the first paragraph of the Merge-Tree model section.

"In this model, not only has the DAG been inverted (and simplified), but
the entire notion of parent-child has been reversed. Due to this
property, the terminology will be different depending on the model we
are referring to; when we are referring to the Merge-Tree, the parent is
the next node toward the merge into the master branch, or the root of
the tree. When referring to the DAG, the parent relationship is in the
opposite direction, from the root toward the branch-point."

    The explanation of Algorithm 2 is, in its current state, generally
    very hard to follow. An example should be included, in a similar way
    as Figures 3, 4, and 5 helped to comprehend the first algorithm. The
    authors must seriously consider simplifying this explanation
    significantly, or the reader is simply lost. A step-by-step graphical
    explanation, using e.g. a flowchart or a diagram, would be much
    appreciated.

An example is included using the DAG in Figure 4. The steps are shown in
Table VIII. A high level description of the algorithm is provided in
paragraphs 3, with the subsequent paragraphs going into more detail.

    In Section C, the authors mention that “The left-most child of a merge
    is on the same branch as the merge.” But what does left-most mean in
    this context? What is left and right when considering the children of
    a node? This was not introduced.

This is no longer applicable. The 'left-most' was in reference to the
figure showing the visualization of the tree in the plugin. The section
has since been removed.

    Evaluation
    Figure 6 would benefit from a frame, or any other clearer visual cue
    that its text is not part of the main text, such as a different
    background color. Right now it merges (no pun intended) with the rest
    of the text.

A bounding box has been placed around the figure.

    In Section IV.A, the authors mention: “…compared the number of commits
    according to the log with the number of commits in the merge tree
    rooted in this commit.” However, it is not clear what “this commit”
    means, as there is no mention to a specific commit being analyzed.
    This is a little confusing and must be clarified. The way I
    understand, what the authors mean is that for each merge commit (mc),
    they (1) extract mc’s log message, (ii) extract from the log message
    the number of commits that were merged in mc, and (3) compare this
    number to the number of commits on mc’s merge-tree.

We replaced the section with a slight re-wording:

"
We extracted merges by Linus Torvalds using the command
git log --merges --author=’Torvalds’ and
compared the number of commits stated in the log message
with the number of commits in the Merge-Tree. We also
used the summaries of the commits found in the merge
(not necessarily all—see above) to make sure those commits
were in their corresponding Merge-Tree.
"

Basically, we look at the commit log. If it says there are 24 commits
and there are 24 commits in the merge tree, it is correct. If there are
more or fewer, then it is incorrect.

    In Section IV.B the authors suddenly mention a bitbucket limitation,
    but so far nothing has been introduced or explained about why the
    proposed technique is dependent on bitbucket. Also, the authors
    mention test cases, but don’t describe them concretey and
    unambiguously, using “…, and other various tests.” For reproducibility
    it would be preferred that the test suite is at least quickly
    described in full (if it is too long, then it should be provided as
    supplemental material to be available to the reader on demand). I
    recommend a rewrite of Section IV.B into a more concise, objective
    description of the tests and results.

Not applicable. This section has been removed, and the algorithm
generalization has been made into a discussion piece. The references to
bitbucket have been removed, as they are not necessary anymore.

    In Section V, in the sentence “These visualizations can be…”, what are
    these visualizations? No visualizations have been introduced so far.
    Also, the reference for the Reingold-Tilford tree should be shown in
    the first occurrence.

This part has been removed, instead opting to go directly into the
original use-cases that we were designing for.

"
The goal of Linvis is to simplify the navigation of the kernel
commit information, specifically focusing on merges. This is
done by leveraging the Merge-Tree to inspect how commits
are merged on the path to the master branch of the repository.

A. Use cases
"

    In Section 6, Reingold-Tilford tree, the authors describe that
    clicking a node will take me to that commit’s page, but I could not
    reproduce that behavior on the web application. Clicking a node simply
    centers that node on the screen.

Clicking a node will select that node (and center it), but not navigate to it.
Navigation requires clicking on the title that gets updated.

    Bubble Tree: there is now intermediate way of zooming in this
    representation, only from the root to the leaves and vice-versa. For
    big trees, this makes it very hard to navigate the tree. From an
    overview, the leaves are simply too small to be chosen (or clicked on)
    effectively and with accuracy. The navigation of the tree in its
    current state is not very useful.

Clicking on other circles will select that circle. Clicking on an
already selected circle will take you to the root. The interface was
updated slightly to increase the size of the border when the cursor is
hovering over a circle to indicate this.

For incredibly large trees, this is true, and I do not have ideas for
fixing this. Most trees are small though, containing less than 100
nodes.

    Although the two different tree visualizations are described, it is
    not clear how they can be actually used in order to extract new
    knowledge from the data. The authors are advised to add at least one
    interesting use case/example where the reader can follow,
    step-by-step, how the visualizations can be used to obtain insight
    from the data. Since the visualizations themselves are not novel, the
    novel use in this application context must be better motivated.

The new information is showing how a commit is merged into the master
branch of a repository, using different visualization metaphors. This
gives a person choices depending on which metaphor they like.

    Discussion
    In Section VIII, I agree that the proposed tool is more concise in
    showing the information than Gitk, however the practical use of the
    tool is still unclear. In order to replace the use that Gitk offers,
    the interface should be improved to include easier navigation and more
    information about each commit that is shown. In its current state, we
    see the nodes but it is very hard to have any kind of extra
    information about which are the commits we are seeing. The authors
    mention that “With Linvis we are able to immediately see what section
    of the kernel these commits and merges pertain to” but just by looking
    at Figure 18 that is not the case. In the paragraph starting on Page
    12, Line 19, the author talks about files that are nowhere to be
    found; maybe there is a missing figure to explain this point?

The discussion has shifted focus from the tool toward the results of the
study. For the specific use-cases described in section IV Linvis is
better than Gitk, but Linvis will never replace gitk.

    General Style Notes and Typos

Thank you, these have been addressed.

Responses to Reviewer 2:

    The focus of this paper is about two things notably the merge tree
    algorithm and visualizations. The merge tree algorithm part is
    impressive, while the visualizations are weak and not particularly
    novel. The paper shows a strong engineering effort. I am an advocate
    of this paper. Even though the algorithm is the stronger part, the
    contribution on visualization could be improved.

Thank you. The goal is to implement a visualization with the merge-tree
model so we can verify that it works, not to design a novel
visualization metaphor for trees.

    It is not clear how visualization will actually help in this project.
    There is a claim that it will, but no evidence supports that claim.
    Even with the visualizations that are presented it is not clear how to
    interpret them, interact with them, and what we can learn from them.
    An example that steps through the visualization techniques would be
    useful here, much like some of the panels described in the tool, to
    illustrate how effective the visualizations really are.

We conducted a user study to show that our tool/visualizations are able
to improve the correctness, accuracy, and timing when trying to
understand and summarize the events in the repository. See section VII.
We found that Linvis does impact comprehension and summarization.

    The LinVis tool is more like a search tool with visualizations as an
    added bonus. The visualizations almost feel like an afterthought. The
    paper does not describe in detail the justification of these
    visualization techniques over other ones. The visualizations
    themselves are not particularly novel. There is claims in the paper
    that the tool is effective, supports better navigation, and better
    explanation; there is simply no evidence. There is no evaluation of
    any kind. It is not clear how these claims are quantified. I
    understand an evaluation is a follow up project, but some details
    should be provided about how an evaluation is planned to be conducted.
    A few additional paragraphs to justify the visualization design and
    evaluation should suffice here.

From section V.C on page 7.

List tree:

"The nesting indicates the parent-child relationship. This visualization
is text-based enabling fast navigation to commits using the built-in
search in most web-browsers."

It allows users to quickly search for an exact commit, but isn't as good
as the others for visualizing the structure.

Reingold-Tilford Tree:

"is the classic tree visualization"

This is the tree diagram that computer scientists will draw if asked to
draw a tree structure. This tree is easy to understand and see how
commits are related and to which merge they belong, but is not great
with very wide trees.

Pack Tree:

"are useful for displaying an overview of large datasets. The initial
goal for the pack tree was for visualizing file systems, which are
similar to git repositories in that they are relatively shallow, but
very wide."

The pack tree visualization is designed to alleviate the remaining
issue.

    Missing related Work

Yes, added this to related work. There are likely many other
visualizations missing, but none that we found are designed with same
goals of showing how a commit is integrated into the master branch of a
repository.

    Regarding the term user

We are unaware of another concise term that is more gender-neutral than
the term "user" for describing people who use something. If there is a
term that you would prefer, we will be happy to looking into the
changes. For this reason, we retain the term "user" for general people
who will use the tool, regardless of intent.

We refer to the members of our user study as "participants".

    Minor Changes

Yes thank you. The structure of the paper has vastly changed since the
original paper was submitted. I believe that all of these have been
taken care of since then.

### 2017-07-22 paper/design.tex
Journal Formatting

@@
 % vim:set et sw=2 ts=4 tw=72:
+
+\section{Design and Implementation}
+\label{sec:design}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/linvis/search_results.png}
+  \includegraphics[width=\figwidth]{figures/linvis/search_results.png}
   \caption{A single \mt tree in the search results of \tool, showing
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/linvis/linvis_files.png}
+  \includegraphics[width=\figwidth]{figures/linvis/linvis_files.png}
   \caption{Table showing the modified files in a merge, with the second
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/linvis/linvis_authors.png}
+  \includegraphics[width=\figwidth]{figures/linvis/linvis_authors.png}
   \caption{Table showing the authors who made changes in a merge. The
@@
         \centering
-        \includegraphics[width=0.9\linewidth]{figures/linvis/linvis_list_tree.png}
+        \includegraphics[width=0.9\figwidth]{figures/linvis/linvis_list_tree.png}
         \caption{The list tree visualization}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
+  \includegraphics[width=0.8\figwidth]{figures/linvis/linvis_bubble.pdf}
   \caption{The pack tree visualization, with the root depicted as the

### 2017-07-22 paper/study.tex
adding reference to Figure 13
Not sure how we missed it, but we missed the reference to the main
results figure from the study.
Journal Formatting

@@
   \label{tab:summarization_tasks}
-  \begin{tabulary}{\linewidth}{LLL}
+  \begin{tabular}{lll}
     \toprule
@@
     \bottomrule
-  \end{tabulary}
+  \end{tabular}
 \end{table}
@@
   \label{tab:null_alt_hypoths}
-  \begin{tabulary}{0.98\linewidth}{CCLL}
+  \begin{tabulary}{0.9\textwidth}{CCLL}
     \toprule
     Task & Metric             & Null Hypothesis                                                      & Alternative Hypothesis\\\midrule
-    T4   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T4   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T4   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T4   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
+    T4-T10   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
+    T4-T10   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
+    T4-T10   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
+    T4-T10   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
 
-    T5   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T5   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T5   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T5   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
-
-    T6   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T6   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T6   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T6   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
-
-    T7   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T7   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T7   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T7   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
-
-    T8   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T8   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T8   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T8   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
-
-    T9   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T9   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T9   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T9   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
-
-    T10   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
-    T10   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
-    T10   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
-    T10   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
     \bottomrule
@@
   \label{tab:opinion_questions}
-  \begin{tabulary}{0.9\linewidth}{LL}
+  \begin{tabular}{ll}
     \toprule
@@
 
-\evan{This looks weird not having text between the subsections... Should
-  I put something here?...}
-
 \subsubsection{Conceptual Tasks}
@@
 
-Table~\ref{tab:cross_commit_results} shows whether the \mt size
-has an effect on the three studied metrics. An interesting observation
-from the figure is that the correctness for T5 and T9 are effected by
-the size of the \mt, but the accuracy is not. The correctness
-results for T5 and T9 are shown in Figure~\ref{fig:correctness_t5_t9},
-and the timing results for T7 are shown in Figure~\ref{fig:timing_t7}.
+Table~\ref{tab:cross_commit_results} shows whether the \mt size has an
+effect on the three studied metrics. An interesting observation from the
+figure is that the correctness for T5 and T9 are effected by the size of
+the \mt, but the accuracy is not. The correctness results for T5 and T9
+are shown in Figure~\ref{fig:correctness_t5_t9}, and the timing results
+for T7 are shown in Figure~\ref{fig:timing_t7}.
+Figure~\ref{fig:summarization_results} shows the combined results from
+both commits for the other tasks.
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/userstudy/time/7.pdf}
+  \includegraphics[width=0.8\figwidth]{figures/userstudy/time/7.pdf}
   \caption{Unaggregated Timing results for T7}
